### 2020-09-05 src/The increasing cost of happiness.Rmd
Add draft

---
title: "The increasing cost of happiness"
author: "Dr Richard Morris"
date: "Updated: `r Sys.Date()`"
bibliography: references.bib
csl: elsevier-harvard.csl
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE,
                      fig.path = "../results/")

#### To do ####
# https://www.psychologytoday.com/au/blog/how-happiness/201409/does-money-really-buy-happiness
```

## Abstract
Goes here.

<br>

### Introduction
***
A fundamental question for the psychology and economics of wellbeing is just how much happiness does a dollar buy? Increasing wealth is commonly associated with increasing happiness, however a point at which wealth no longer increases happiness has also been [widely observed](https://www-sciencedirect-com.ezproxy2.library.usyd.edu.au/science/article/pii/S0167487007000694){target="_blank"} [@dolan2008we; @easterlin1974growth]. Given that a central goal of nations and governments is to improve wealth under the assumption that wealth always increases wellbeing, challenges to this notion have far reaching consequences [@frijters2020happy].

For instance, one [survey of 1,000 Americans](https://www.pnas.org/content/107/38/16489.full){target="_blank"}, conducted in 2010, concluded that money does make us happier – but only up to a certain point. The findings showed that self-reported levels of happiness increased with household wealth up to \$75,000 a year. But after that, increasing amounts of money had no further effect on happiness [@kahneman2010high]. Results such as these have reinforced the view among psychologists and economists that the relationship between income and measures of subjective wellbeing, such as happiness, increases linearly up to a point after which there is little gained by further increments in income [@clark2018four; @dolan2008we].

However a study reported in 2013 challenged the idea that the positive effect of money plateaus [@stevenson2013subjective]. After comparing life satisfaction levels in both rich and poor countries, and rich and poor people within a country – with “rich” being defined as an income greater than $15,000 per person – Stevenson & Wolfers concluded: “The relationship between wellbeing and income … does not diminish as income rises. If there is a satiation point, we are yet to reach it.”

Subjective wellbeing is not a unitary entity [@diener2017findings]; studies typically distinguish between _life satisfaction_, the cognitive appraisal of one's own accomplishments, and _happiness_, one's prevailing affective state or emotional mood. Money can have different effects on each. For instance, we have recently reported that positive life events, such as a major financial windfall, can have a greater impact on a person's satisfaction than their happiness [@kettlewell2020differential]. And while Kahneman observed the association between household wealth and happiness plateaud after $75,000; he also reported the association between wealth and life satisfaction was consistentily positive. This raises the possibility that contradictory associations between wealth and wellbeing previously reported may be due to differences in the construct measured. It is easy to imagine a person who is satisfied (for example because they have a good job and healthy family) but is still unhappy on a day-to-day basis. And researchers studying the relationship between wealth and wellbeing tend to focus on surveys of _life satisfaction_, which reflects only one domain of wellbeing, and ignore other aspects of wellbeing such as happiness [e.g., @stevenson2013subjective; @frijters2020happy].

Since 2013, a distinct effect of wealth and happiness has also been challenged. [A survey of 1.7 million people representing 164 countries over 12 years (2005-2016) in the Gallup World Poll](https://www-nature-com.ezproxy2.library.usyd.edu.au/articles/s41562-017-0277-0){target="_blank"} reported money was associated with no further improvement in happiness after \$60-75,000, _nor any improvement in satisfaction_ after $95,000 [@jebb2018happiness]. Thus while the location differed, money had inflected effects with both satisfaction and happiness. On the other hand, in a nationally representative sample of 44,000 adult Americans over 44 years in the General Social Survey (GSS) happiness continued to increase with money [@twenge2020expanding], implying no inflection point existed between money and happiness.

Both these reports provide contradictory results regarding the nature of the relationship between wealth and wellbeing, and both are based on data spanning long periods of time (12-44 years). And while such research has revealed variations due to world region, gender, and education, there has been no investigation at to whether the relationship between wealth and wellbeing has changed over time. The value of money of course changes over time with inflation, however the distribution of wealth has also changed over the last 30 years in most OECD countries (as the rich got richer faster than the poor). Such changes in the distribution of wealth are also likely to change the relationship with wellbeing when it is governed by relative wealth rather than absolute wealth.

We used household economic panel data from Australia (HILDA) to provide the first investigation of whether changes in wealth and wellbeing have changed over the last 16 years (2002-2018). We distinguished between satisfaction and happiness as different components of subjective wellbeing, and evaluated how each varies with household wealth. Economic security is better represented by household wealth, since members of the same household share expenses as well as risks; i.e., they can provide a direct and immediate support network when financial shocks occur. The other major studies also used household after-tax income as the indicator of wealth and economic security, and so we follow the same standard here as well.

After adjusting for age, relationship status, and education level, we find that satisfaction increases linearly with household wealth, but the same purchasing power has diminishing returns on happiness. We also discover that the location of the inflection point has increased in real dollar value since 2002.

<br>

#### Wealth
We have a variety of income & wealth information in HILDA, most of it collected annually. However the household net worth is collected every four years. In the present report we use:

1. [hwnwip](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=hwnwip){target="_blank"} Household Net Worth ($) positive values [imputed] [weighted topcode]. Calculated as assets minus debts for the household.
2. [hifdip](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=hifdip){target="_blank"} Household financial year disposable regular income ($) positive values [imputed] [weighted topcode]. The sum across all household members of financial year gross regular income minus taxes on financial year gross regular income.
3. [tifdip](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=tifdip){target="_blank"} Personal disposable income ($) positive values [weighted topcode]. Financial year gross regular income minus taxes on financial year gross regular income.

_Note_. We also examined household and personal gross wages, and household and personal total disposable income.

Income and wealth variables were provided by the University of Melbourne as imputed, and with weighted topcodes to preserve anonymity. We removed individuals with values above the threshold indicated for top-coding.

Household wealth variables were adjusted for the number of people in each household in the same manner as Headey, Muffels & Wooden, 2004; i.e., income divided by the square root of household size (Buhmann et al, 1988; Coulter et al, 1992).

<br>

#### Subjective Wellbeing
There are a variety of variables related to subjective well-being collected annually, but the two we used here matched the variables we used in our previous paper:

1. `losat` Life satisfaction - How satisfied are you with your life (0 to 10)
2. `gh9a` to `gh9i` Affective wellbeing - Sum of SF-36 item 9 (mental health and vitality)

_Note_. We also examined `ghmh` (the 5-item subset of mental health questions, sometimes called the **MHi-5**, from the SF-36), and the patterns were almost identical to our sum of 9a to 9i.

<br>

#### Modelling and model-selection

_Modelling_  
To determine whether wellbeing was a decreasing function of wealth or a linear function of wealth, we modelled the relationship between individual wealth and wellbeing using a simple B-spline basis function with 2 degrees of freedom (i.e., 1 y-intercept and 1 inflection point or _knot_). This was equivalent to a "piecewise" or "broken-stick" linear regression, and allowed us to determine the direction and location of the inflection point between wealth and wellbeing in each year.

_Model selection_  
We compared the fit of the "broken-stick" model with a simpler linear model ("unbroken"), adjusted for degrees of freedom by the Bayesian Information Criterion (BIC). BIC was chosen over AIC or LRT because it will penalize our more complex model (the broken stick) more heavily. Following Raftery, differences in BIC greater than 10 were taken as "very strong evidence"; greater than 5 were "strong evidence"; greater than 2 were "positive evidence"; and less than 2 (but greater than zero) were "weak evidence" for the model with the smallest BIC ( [Raftery, 1995](https://www.sciencedirect.com/topics/pharmacology-toxicology-and-pharmaceutical-science/bayesian-information-criterion){target="_blank"}).

For visualization purposes only, due to the large number of individual data points in each year, we grouped individuals into 10 equal-sized subgroups according to their wealth decile ($). Thus in each plot, the mean wealth and mean wellbeing score for each decile is presented, rather than each individual data point. Note that the line-of-best-fit and 95% confidence intervals from the broken stick regression model _of all individuals_ is shown in overlay.

<br>

#### Covariates

Covariates included in subsequent sensitivity analyses were: age, age^2^, years of education, and dummy variables for partnered, single (never married), and unemployed.

Years of education were determined from the [Australian Qualifications Framework (AQF)](https://en.wikipedia.org/wiki/Australian_Qualifications_Framework#Certificates_I-IV){target="_blank"} from the Department of Industry.

Separate analyses for women are included in the Supplementary material.

<br>

### Results
***

#### Demographics and sample characteristics

After stratification by decile (personal disposable income) we see that age, proportion of males, couples, years of education and education levels all increase with decile. Conversely, proportion of females, single people (never married), unemployed, and fulltime students decreases with decile. Of particular interest is that wellbeing, household wealth and SEIFA index follow a U-shaped function with the minimum value falling between the third and sixth decile.

<br>

#### Household disposable income (adjusted for household size)

**Key points**

- Cognitive wellbeing (satisfaction) had a relatively linear, positive relationship wtih household disposable income. The inflection point was equally positive and negative across years, without a uniform pattern. 
- By contrast, the relationship between affective wellbeing (happiness) and disposable income tended to be a positive but decreasing ("*broken*") function with a uniform inflection point.

<br>

### Conclusions
***

The main conclusions of this report so far are:

- The relationship between subjective wellbeing and wealth is positive but happiness and satisfaction have different (positive) relationships. Satisfaction increases linearly with wealth while happiness increases rapidly up to a point, after which further increments in wealth produce less change - this is the **cost point** of happiness. So money buys happiness when economic security is scarce, but not so much once the cost point is reached.
- The cost point after which money provides decreasing returns in happiness is increasing over time. In 2002, the cost point of happiness represented a 13% increase over median wealth, while in 2018 it represented a 67% increase over median wealth.
- According to @frijters2020happy, coming up with a consensus estimate for income to translate into welbeing features high on the wellbeing research agenda. However evidence of a linear relationship between wellbeing and wealth calls into question the need to supplant traditional measures of economic success such as GDP with measures of wellbeing, such as "life satisfaction" surveys.

### References

### 2020-09-05 src/The increasing cost of happiness.Rmd
Updated draft

@@
 csl: elsevier-harvard.csl
-output: github_document
+output: 
+  html_document:
+    keep_md: true
 ---
@@
 ## Abstract
-Goes here.
+A fundamental question for governments and people is how much happiness does a dollar buy?
+The accepted view among economists and psychologists is that money and happiness increase together up to a point, after which there is little further gain from increasing wealth.
+While the location of this _change point_ has been determined, and the cost of happiness reportedly ranges between USD\$60-95K, there has been no investigation as to whether the cost of happiness has increased or decreased over time.
 
@@
 But after that, increasing amounts of money had no further effect on happiness [@kahneman2010high].
-Results such as these have reinforced the view among psychologists and economists that the relationship between income and measures of subjective wellbeing, such as happiness, increases linearly up to a point after which there is little gained by further increments in income [@clark2018four; @dolan2008we].
+Results such as these have reinforced the view among psychologists and economists that the relationship between wealth and measures of subjective wellbeing, such as happiness, increases linearly up to a point after which there is little gained by further increments in wealth [@clark2018four; @dolan2008we].
 

### 2020-09-05 src/The increasing cost of happiness.Rmd
Updated draft

@@
 ```
+***
 
@@
 While the location of this _change point_ has been determined, and the cost of happiness reportedly ranges between USD\$60-95K, there has been no investigation as to whether the cost of happiness has increased or decreased over time.
+We use household economic data from Australia between 2002-2016 to test the relationship with both happiness and life satisfaction.
+We discovered no change point existed for satisfaction, but the cost of happiness has increased over those 16 years faster than inflation (i.e., cost of living).
 

### 2020-09-05 src/The increasing cost of happiness.Rmd
Updated to github_document

@@
 csl: elsevier-harvard.csl
-output: 
-  html_document:
-    keep_md: true
+output: github_document
 ---

### 2020-09-08 src/The increasing cost of happiness.Rmd
Add figure

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(eval = FALSE,
-                      fig.path = "../results/")
+library(tidyverse)
+library(bayesplot)
+
+knitr::opts_chunk$set(eval = TRUE,
+                      echo = FALSE,
+                      warning = FALSE,
+                      message = FALSE,
+                      fig.path = "../figures/")
 
@@
 While the location of this _change point_ has been determined, and the cost of happiness reportedly ranges between USD\$60-95K, there has been no investigation as to whether the cost of happiness has increased or decreased over time.
-We use household economic data from Australia between 2002-2016 to test the relationship with both happiness and life satisfaction.
-We discovered no change point existed for satisfaction, but the cost of happiness has increased over those 16 years faster than inflation (i.e., cost of living).
+We tested the relationship between money and both happiness and life satisfaction using household economic data from Australia between 2002-2016.
+We discovered no change point with wealth existed for satisfaction, but the cost of happiness has increased over those 16 years faster than inflation (i.e., cost of living).
+Results such as these suggest we need to understand why the cost of happiness has increased, and to consider whether policy-goals to improve wellbeing by increasing wealth are feasible.
 
@@
 _Model selection_  
-We compared the fit of the "broken-stick" model with a simpler linear model ("unbroken"), adjusted for degrees of freedom by the Bayesian Information Criterion (BIC).
-BIC was chosen over AIC or LRT because it will penalize our more complex model (the broken stick) more heavily.
-Following Raftery, differences in BIC greater than 10 were taken as "very strong evidence"; greater than 5 were "strong evidence"; greater than 2 were "positive evidence"; and less than 2 (but greater than zero) were "weak evidence" for the model with the smallest BIC ( [Raftery, 1995](https://www.sciencedirect.com/topics/pharmacology-toxicology-and-pharmaceutical-science/bayesian-information-criterion){target="_blank"}).
+The [Widely Applicable Information Criterion (WAIC) was used for model selection](https://bookdown.org/ajkurz/Statistical_Rethinking_recoded/overfitting-regularization-and-information-criteria.html#the-problem-with-parameters){target="_blank"}.
+WAIC was defined as:
+
+WAIC = -2(lppd - *p*~WAIC~)
+
+Where lppd (log pointwise predictive density) is the total across observations of the log of the average likelihood of each observation, and *p*~WAIC~ is the effective number of free parameters determined by the sum of the variance in log-likelihood for each observation (*i*).
 
@@
 
-Covariates included in subsequent sensitivity analyses were: age, age^2^, years of education, and dummy variables for partnered, single (never married), and unemployed.
+Age (and age^2^) and education were included as covariates, and students were removed.
 
@@
 
+##### Bayesian change point results
+
+Shaded areas to the right of the vertical dotted line are credibly (95%) larger than the base year (2002).
+
+```{r happy_change_time, fig.width = 4.9, fig.height = 9}
+hifdip_gh9_fixed_fit <- read_rds("../results/hifdip_gh9_fixed_full.RDS")
+
+as.data.frame(hifdip_gh9_fixed_fit) %>%
+  select(b_alpha_Intercept, b_alpha_year2006, b_alpha_year2010, 
+         b_alpha_year2014, b_alpha_year2018) %>%
+  transmute(
+    fy2002 = b_alpha_Intercept,
+    fy2006 = mean(b_alpha_Intercept) + b_alpha_year2006,
+    fy2010 = mean(b_alpha_Intercept) + b_alpha_year2010,
+    fy2014 = mean(b_alpha_Intercept) + b_alpha_year2014,
+    fy2018 = mean(b_alpha_Intercept) + b_alpha_year2018
+  ) %>%
+  mutate_all(.funs = ~brms::inv_logit_scaled(., 0, 150)*10) -> d1
+
+color_scheme_set("red")
+mcmc_areas(
+  d1,
+  pars = rev(colnames(d1)),
+  prob = 0.95, # 80% intervals
+  prob_outer = 1, # 99%
+  point_est = "mean"
+  ) +
+  theme_test() +
+  geom_vline(aes(xintercept = mean(d1$fy2002)), color = "gray", linetype = 3) +
+  labs(
+    subtitle = "Posterior change points (mean & 95% distribution)",
+    x = "Household wealth ($000s)")
+```
+
+<br>
+
 **Key points**
 
-- Cognitive wellbeing (satisfaction) had a relatively linear, positive relationship wtih household disposable income.
-The inflection point was equally positive and negative across years, without a uniform pattern.
-- By contrast, the relationship between affective wellbeing (happiness) and disposable income tended to be a positive but decreasing ("*broken*") function with a uniform inflection point.
+- The change point for happiness (affective wellbeing) shifts to the right over time and requires greater household wealth.
 

### 2020-09-08 drafts/The increasing cost of happiness.Rmd
added drafts

--- 2020-09-08 src/The increasing cost of happiness.Rmd
@@
 date: "Updated: `r Sys.Date()`"
-bibliography: references.bib
-csl: elsevier-harvard.csl
+bibliography: ../src/references.bib
+csl: ../src/elsevier-harvard.csl
 output: github_document

### 2020-09-08 drafts/draft0.Rmd
updated draft0

--- 2020-09-08 drafts/The increasing cost of happiness.Rmd
@@
 A fundamental question for the psychology and economics of wellbeing is just how much happiness does a dollar buy?
-Increasing wealth is commonly associated with increasing happiness, however a point at which wealth no longer increases happiness has also been [widely observed](https://www-sciencedirect-com.ezproxy2.library.usyd.edu.au/science/article/pii/S0167487007000694){target="_blank"} [@dolan2008we; @easterlin1974growth].
+Increasing wealth is commonly associated with increasing happiness, however a point at which wealth no longer increases happiness has also been [widely observed](https://www-sciencedirect-com.ezproxy2.library.usyd.edu.au/science/article/pii/S0167487007000694){target="_blank"} [@clark2008relative; @dolan2008we; @easterlin1974growth].
 Given that a central goal of nations and governments is to improve wealth under the assumption that wealth always increases wellbeing, challenges to this notion have far reaching consequences [@frijters2020happy].
@@
 
-However a study reported in 2013 challenged the idea that the positive effect of money plateaus [@stevenson2013subjective].
-After comparing life satisfaction levels in both rich and poor countries, and rich and poor people within a country – with “rich” being defined as an income greater than $15,000 per person – Stevenson & Wolfers concluded: “The relationship between wellbeing and income … does not diminish as income rises.
-If there is a satiation point, we are yet to reach it.”
-
 Subjective wellbeing is not a unitary entity [@diener2017findings]; studies typically distinguish between _life satisfaction_, the cognitive appraisal of one's own accomplishments, and _happiness_, one's prevailing affective state or emotional mood.
 Money can have different effects on each.
-For instance, we have recently reported that positive life events, such as a major financial windfall, can have a greater impact on a person's satisfaction than their happiness [@kettlewell2020differential].
-And while Kahneman observed the association between household wealth and happiness plateaud after $75,000; he also reported the association between wealth and life satisfaction was consistentily positive.
-This raises the possibility that contradictory associations between wealth and wellbeing previously reported may be due to differences in the construct measured.
-It is easy to imagine a person who is satisfied (for example because they have a good job and healthy family) but is still unhappy on a day-to-day basis.
-And researchers studying the relationship between wealth and wellbeing tend to focus on surveys of _life satisfaction_, which reflects only one domain of wellbeing, and ignore other aspects of wellbeing such as happiness [e.g., @stevenson2013subjective; @frijters2020happy].
+For instance, we have recently reported that positive life events, such as a major financial windfall, can have a greater impact on an individual's satisfaction than their happiness [@kettlewell2020differential].
+However these effects were observed as a change within individuals, and the change was transitory as both satisfaction and happiness returned to baseline after two years [see also @frijters2011life].
+Evidence for a distinct relationship with happiness and satisfaction across individuals also exists: while Kahneman observed the association between household wealth and happiness plateaud after $75,000, he also reported the association between wealth and life satisfaction was consistently positive.
+Indeed, several researchers have observed a difference between two questions that are often used in surveys of subjective well-being: “How satisfied are you with your life?” and “How happy are you these days?” [@veenhoven2006rising; @howell2008relation].
+The common conclusion is that wealth is more strongly related to satisfaction than to happiness.
 
-Since 2013, a distinct effect of wealth and happiness has also been challenged.
-[A survey of 1.7 million people representing 164 countries over 12 years (2005-2016) in the Gallup World Poll](https://www-nature-com.ezproxy2.library.usyd.edu.au/articles/s41562-017-0277-0){target="_blank"} reported money was associated with no further improvement in happiness after \$60-75,000, _nor any improvement in satisfaction_ after $95,000 [@jebb2018happiness].
-Thus while the location differed, money had inflected effects with both satisfaction and happiness.
-On the other hand, in a nationally representative sample of 44,000 adult Americans over 44 years in the General Social Survey (GSS) happiness continued to increase with money [@twenge2020expanding], implying no inflection point existed between money and happiness.
+The relationship between wealth and happiness and satisfaction are relevant for the growing popularity of using subjective well-being as a policy guide for government and agencies.
+If there are distinct effects of wealth on happiness and satisfaction then the question of which is better suited to assess human welfare and to guide policy seems critical.
+The observation that happiness plateaus with increasing wealth while satisfaction keeps rising acquires some importance in this context.
+Not everyone will agree that enhancing the happiness experienced by those who are already quite happy is a legitimate policy objective.
 
-Both these reports provide contradictory results regarding the nature of the relationship between wealth and wellbeing, and both are based on data spanning long periods of time (12-44 years).
-And while such research has revealed variations due to world region, gender, and education, there has been no investigation at to whether the relationship between wealth and wellbeing has changed over time.
-The value of money of course changes over time with inflation, however the distribution of wealth has also changed over the last 30 years in most OECD countries (as the rich got richer faster than the poor).
-Such changes in the distribution of wealth are also likely to change the relationship with wellbeing when it is governed by relative wealth rather than absolute wealth.
+And while research since 2010 has revealed variations in the plateau or _change point_ due to [world region, gender, and education](https://www-nature-com.ezproxy2.library.usyd.edu.au/articles/s41562-017-0277-0){target="_blank"} [@jebb2018happiness], there has been no investigation as to whether the relationship between wealth and happiness is changing over time [although see Fig.
+4 in @clark2008relative; for a speculative mechanism].
+The value of money of course changes over time with inflation, however the distribution of wealth has also changed over the last 30 years in many OECD countries including Australia (as the rich got richer faster than the poor, ref needed).
+Such changes in the distribution of wealth are also likely to change the relationship with wellbeing when it is governed by relative wealth rather than absolute wealth [@clark2008relative].
 
@@
 
+<br>
+
+A study reported in 2013 challenged the idea that the positive effect of money plateaus [@stevenson2013subjective].
+After comparing life satisfaction levels in both rich and poor countries, and rich and poor people within a country – with “rich” being defined as an income greater than $15,000 per person – Stevenson & Wolfers concluded: “The relationship between wellbeing and income … does not diminish as income rises.
+If there is a satiation point, we are yet to reach it.”
+
+Since 2013, the distinct effect of wealth and happiness has also been challenged.
+[A survey of 1.7 million people representing 164 countries over 12 years (2005-2016) in the Gallup World Poll](https://www-nature-com.ezproxy2.library.usyd.edu.au/articles/s41562-017-0277-0){target="_blank"} reported money was associated with no further improvement in happiness after \$60-75,000, _nor any improvement in satisfaction_ after $95,000 [@jebb2018happiness].
+Thus while the location differed, money had inflected effects with both satisfaction and happiness.
+On the other hand, in a nationally representative sample of 44,000 adult Americans over 44 years in the General Social Survey (GSS) happiness continued to increase with money [@twenge2020expanding], implying no inflection point existed between money and happiness.
+
+Both these reports provide contradictory results regarding the nature of the relationship between wealth and wellbeing, and both are based on data spanning long periods of time (12-44 years).
+Perhaps the different results are due to differences in the constructs being measured, or perhaps it is due to different effects of time between studies.
+
 ### References

### 2020-09-10 drafts/draft0.Rmd
Added results

@@
 
-After adjusting for age, relationship status, and education level, we find that satisfaction increases linearly with household wealth, but the same purchasing power has diminishing returns on happiness.
+After adjusting for age, gender, and education level, we find that satisfaction increases linearly with household wealth, but the same purchasing power has diminishing returns on happiness.
 We also discover that the location of the inflection point has increased in real dollar value since 2002.
@@
 #### Wealth
-We have a variety of income & wealth information in HILDA, most of it collected annually. However the household net worth is collected every four years. In the present report we use:
+We have a variety of income & wealth information in HILDA, most of it collected annually.
+However the household net worth is collected every four years.
+In the present report we use [hifdip](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=hifdip){target="_blank"} Household financial year disposable regular income ($).
+The sum across all household members of financial year gross regular income minus taxes on financial year gross regular income.
+Wealth variables were provided by the University of Melbourne as imputed, and with weighted topcodes to preserve anonymity.
+_Nb_.
+We also examined household and personal gross wages and disposable income (not shown).
 
-1.
-[hwnwip](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=hwnwip){target="_blank"} Household Net Worth ($) positive values [imputed] [weighted topcode].
-Calculated as assets minus debts for the household.
-2.
-[hifdip](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=hifdip){target="_blank"} Household financial year disposable regular income ($) positive values [imputed] [weighted topcode].
-The sum across all household members of financial year gross regular income minus taxes on financial year gross regular income.
-3.
-[tifdip](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=tifdip){target="_blank"} Personal disposable income ($) positive values [weighted topcode].
-Financial year gross regular income minus taxes on financial year gross regular income.
-
-_Note_. We also examined household and personal gross wages, and household and personal total disposable income.
-
-Income and wealth variables were provided by the University of Melbourne as imputed, and with weighted topcodes to preserve anonymity.
 We removed individuals with values above the threshold indicated for top-coding.
-
-Household wealth variables were adjusted for the number of people in each household in the same manner as Headey, Muffels & Wooden, 2004; i.e., income divided by the square root of household size (Buhmann et al, 1988; Coulter et al, 1992).
+Household wealth was adjusted for the number of people in each household in the same manner as Headey, Muffels & Wooden, 2004; i.e., income divided by the square root of household size (Buhmann et al, 1988; Coulter et al, 1992).
 
@@
 
+```{r funtimes}
+source("../src/happyfun.r")
+```
+
 <br>
@@
 
-Age (and age^2^) and education were included as covariates, and students were removed.
+Age (and age^2^), gender, and education were included as covariates, and students were removed.
 
-Years of education were determined from the [Australian Qualifications Framework (AQF)](https://en.wikipedia.org/wiki/Australian_Qualifications_Framework#Certificates_I-IV){target="_blank"} from the Department of Industry.
-
-Separate analyses for women are included in the Supplementary material.
+```{r read_write_covs}
+happywealth_covs <- read_rds("../results/happywealth_covs.rds")
+```
 
@@
 
-#### Demographics and sample characteristics
+#### Household wealth on satisfaction (cognitive wellbeing)
+```{r losat_hwnwip, fig.height=11}
+library(patchwork)
 
-After stratification by decile (personal disposable income) we see that age, proportion of males, couples, years of education and education levels all increase with decile.
-Conversely, proportion of females, single people (never married), unemployed, and fulltime students decreases with decile.
-Of particular interest is that wellbeing, household wealth and SEIFA index follow a U-shaped function with the minimum value falling between the third and sixth decile.
+p1 <- happywealth_covs %>%
+  plot_linear(losat, adj_hifdip)
+
+p2 <- happywealth_covs %>%
+  plot_inflection_cubic(losat, adj_hifdip)
+
+p3 <- happywealth_covs %>%
+  plot_inflection_piecewise(losat, adj_hifdip)
+
+(p1 + p2 + p3) +
+  plot_annotation(title = "Effect of wealth on satisfaction")
+```
 
@@
 
-#### Household disposable income (adjusted for household size)
+**Key points**
+
+- Generally the relationship between satisfaction and household wealth appears linear
+- There is no satiety point (horizontal inflection) in satisfaction before the highest decile
+- Any inflection point in satisfaction is as likely to be convex as concave
+
+<br>
+
+#### Household wealth on happiness (affective wellbeing)
+```{r gh9_hwnwip, fig.height=11}
+library(patchwork)
+
+p1 <- happywealth_covs %>%
+  plot_linear(gh9, adj_hifdip)
+
+p2 <- happywealth_covs %>%
+  plot_inflection_cubic(gh9, adj_hifdip)
+
+p3 <- happywealth_covs %>%
+  plot_inflection_piecewise(gh9, adj_hifdip)
+
+(p1 + p2 + p3) +
+  plot_annotation(title = "Effect of wealth on happiness")
+```
+
+<br>
+
+**Key points**
+
+- There is no happiness satiety point (horizontal inflection) before the final decile
+- A concave happiness inflection point is evident before or near the median decile
+
+<br>
 
@@
   ) %>%
-  mutate_all(.funs = ~brms::inv_logit_scaled(., 0, 150)*10) -> d1
+  mutate_all(.funs = ~brms::inv_logit_scaled(., 0, 150)*10) -> posteriors
 
@@
 mcmc_areas(
-  d1,
-  pars = rev(colnames(d1)),
+  posteriors,
+  pars = rev(colnames(posteriors)),
   prob = 0.95, # 80% intervals
@@
   theme_test() +
-  geom_vline(aes(xintercept = mean(d1$fy2002)), color = "gray", linetype = 3) +
+  geom_vline(aes(xintercept = mean(posteriors$fy2002)), color = "gray", linetype = 3) +
   labs(
@@
 
-- The change point for happiness (affective wellbeing) shifts to the right over time and requires greater household wealth.
+- The change point for happiness (affective wellbeing) shifts to the right over time and occurs at greater household wealth levels.
 
 <br>
+
+#### Change points over time
+
+The mean change point location has moved further right (higher, faster) than the median household wealth over time:
+
+```{r, table1, ft.align="left"}
+library(knitr)
+
+happywealth_covs %>%
+  select(year, xwaveid, adj_hifdip) %>%
+  mutate(
+    year = paste0("fy", year),
+    adj_hifdip = adj_hifdip/1000) %>%
+  left_join(
+    gather(
+      summarise_all(posteriors, list(median)), year, change_point
+    )) %>%
+  group_by(year) %>%
+  summarise(
+    `median ($000s)` = median(adj_hifdip),
+    `change point ($000s)` = mean(change_point),
+    `%` = mean(adj_hifdip > mean(change_point))
+  ) %>%
+  kable(format = "html", caption = "Household wealth median, change point and percentile")
+```
+
+<br><br>
 

### 2020-09-10 src/happyfun.R

@@ -1,0 +1,347 @@
+plot_linear <- function(.data, swb_col, dollars_col) {
+ 
+
+  swb_colname <- rlang::enexpr(swb_col)
+  dollars_colname <- rlang::enexpr(dollars_col)
+  
+  .data %>%
+    select(xwaveid, 
+           year, 
+           # swb = enexpr(swb_col), dollars = enexpr(dollars_col)) %>%
+           swb = !!swb_colname, dollars = !!dollars_colname) %>%
+    filter(swb > 0) %>%
+    filter(dollars > 0) %>%
+    filter(year %in% c(2002, 2006, 2010, 2014, 2018)) -> df
+  
+  df %>%
+    group_by(year) %>%
+    mutate(
+      dollar_bin = ntile(dollars, 10)
+    ) %>%
+    group_by(year, dollar_bin) %>%
+    summarize(
+      wellbeing = mean(swb, na.rm=T),
+      wealth = mean(dollars, na.rm=T)/1000,
+      n = n()
+    ) -> df.summary
+  
+  df %>%
+    mutate(
+      wealth = dollars/1000,
+      wellbeing = swb
+    ) -> df.plot
+  
+  # general additive model
+  ggplot(df.plot, aes(x = wealth, y = wellbeing)) +
+    geom_smooth(
+      method = "lm", 
+      size = 0.5,
+      se = TRUE) +
+    coord_cartesian(xlim = c(0, max(df.summary$wealth)),
+                    ylim = c(min(df.summary$wellbeing), max(df.summary$wellbeing))) +
+    facet_wrap(~year, ncol = 1) +
+    labs(subtitle = paste(rlang::as_name(swb_colname), "(linear)"),
+         x = paste(rlang::as_name(dollars_colname), "($000s)"),
+         y = "") +
+    theme_test() +
+    theme(
+      legend.background = element_blank(),
+      legend.title = element_blank(),
+      legend.position = c(0.075, 0.925)) -> p1
+  
+  p1 +
+    geom_point(data = df.summary) 
+  
+}
+
+plot_inflection_piecewise <- function(.data, swb_col, dollars_col) {
+  
+  swb_colname <- rlang::enexpr(swb_col)
+  dollars_colname <- rlang::enexpr(dollars_col)
+  
+  .data %>%
+    select(xwaveid, year, 
+           # swb = enexpr(swb_col), dollars = enexpr(dollars_col)) %>%
+           swb = !!swb_colname, dollars = !!dollars_colname) %>%
[…]

### 2020-09-10 drafts/draft0.Rmd
added maths

@@
 output: github_document
+always_allow_html: yes
 ---
@@
 # https://www.psychologytoday.com/au/blog/how-happiness/201409/does-money-really-buy-happiness
+# 
+# more people could a achieve a level of wealth at which happiness no longer increased
 ```
@@
 However these effects were observed as a change within individuals, and the change was transitory as both satisfaction and happiness returned to baseline after two years [see also @frijters2011life].
-Evidence for a distinct relationship with happiness and satisfaction across individuals also exists: while Kahneman observed the association between household wealth and happiness plateaud after $75,000, he also reported the association between wealth and life satisfaction was consistently positive.
+Across individuals, wealth also has a distinct relationship with happiness and satisfaction as well.
+While Kahneman et al observed the association between household wealth and happiness plateaud after $75,000, they also reported the association between wealth and life satisfaction was consistently positive.
 Indeed, several researchers have observed a difference between two questions that are often used in surveys of subjective well-being: “How satisfied are you with your life?” and “How happy are you these days?” [@veenhoven2006rising; @howell2008relation].
@@
 _Modelling_  
-To determine whether wellbeing was a decreasing function of wealth or a linear function of wealth, we modelled the relationship between individual wealth and wellbeing using a simple B-spline basis function with 2 degrees of freedom (i.e., 1 y-intercept and 1 inflection point or _knot_).
-This was equivalent to a "piecewise" or "broken-stick" linear regression, and allowed us to determine the direction and location of the inflection point between wealth and wellbeing in each year.
+To initially describe the relationship between wealth and wellbeing, we modelled the data from each year in three different ways: 1) a simple linear model of the form y<sub>*i*</sub> = *&beta;*<sub>0</sub> + *&beta;*<sub>1</sub>*x*<sub>*i*</sub> + *&epsilon;*<sub>*i*</sub>; 2) a cubic b-spline model of the form *B*(*x*) = *&Sigma;&beta;*<sub>*i*</sub>*B*<sub>*i,n*</sub>(*x*); and 3) a piecewise (broken-stick) model of the form y<sub>*i*</sub> = *&beta;*<sub>0</sub> + *&beta;*<sub>1</sub>(*x*<sub>*i*</sub> - &omega;)(*x*<sub>*i*</sub> ≤ &omega;) + *&beta;*<sub>2</sub>(*x*<sub>*i*</sub> - &omega;)(*x*<sub>*i*</sub> ≤ &omega;).
+
+These three models were chosen to represent varying degrees of complexity.
+The linear model was obviously the simplest relationship that can exist between wealth and wellbeing.
+The cubic B-spline was a complex model (more degrees of freedom) and chosen because it has been recently used to describe the maximum point at which wealth no longer increases wellbeing [@jebb2018happiness].
+Following Jebb et al, we restricted the model to 4, 5 and 6 knots and show each instantiation here.
+The piecewise model was chosen as the simplest extension of a linear model to identify the change point at which wellbeing no longer increases linearly with wealth.
+
+_Change point estimation_  
+To determine whether the relationship between wellbeing and wealth contained an inflection point ("change point"), we modelled the relationship between wealth and wellbeing across individuals using the piecewise model described above, and estimated the posterior probability of the changepoint location using a Bayesian approach:
+
+Let y<sub>*i*</sub> ~ *N*(*&mu;*<sub>*i*</sub>, *&sigma;*<sup>2</sup><sub>*y*</sub>)
+
+*&mu;*<sub>i</sub> = *&beta;*<sub>0</sub> + *&beta;*<sub>1</sub>(*x*<sub>*i*</sub> - &omega;<sub>*t*</sub>)(*x*<sub>*i*</sub> ≤ &omega;<sub>*t*</sub>) + *&beta;*<sub>2</sub>(*x*<sub>*i*</sub> - &omega;<sub>*t*</sub>)(*x*<sub>*i*</sub> > &omega;<sub>*t*</sub>)  \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \  (1)
+
+Where *x*<sub>i</sub> and *y*<sub>i</sub> was an individuals' household wealth (\$) and wellbeing, respectively; and &omega;<sub>*t*</sub> was the changepoint in $ for each year (*t*).
+
+The location of the changepoint was sampled randomly from a uniform prior over the resticted range of wealth values, and the posterior probability of the location for each year is presented.
+
+<br>
 
 _Model selection_  
-The [Widely Applicable Information Criterion (WAIC) was used for model selection](https://bookdown.org/ajkurz/Statistical_Rethinking_recoded/overfitting-regularization-and-information-criteria.html#the-problem-with-parameters){target="_blank"}.
+To determine whether wellbeing was a linear or non-linear (e.g., piecewise) function of wealth, we compared the linear and piecewise model fits using the [Widely Applicable Information Criterion (WAIC)](https://bookdown.org/ajkurz/Statistical_Rethinking_recoded/overfitting-regularization-and-information-criteria.html#the-problem-with-parameters){target="_blank"}.
+The WAIC is the log-posterior predictive density plus a penalty proportional to the variance in the posterior distribution.
+Thus it provides an approximation of the out-of-sample deviance that converges to the cross-validation approximation in a large sample, with a penalty for the effective number of parameters (degrees of freedom).
+For this reason is it useful to compare two models of varying complexity.
+
 WAIC was defined as:
@@
 
-- Generally the relationship between satisfaction and household wealth appears linear
-- There is no satiety point (horizontal inflection) in satisfaction before the highest decile
-- Any inflection point in satisfaction is as likely to be convex as concave
+- Generally the relationship between satisfaction and household wealth appeared linear
+- The linear relationship between satisfaction and wealth became weaker (less steep) over years
+- There is no satiety point (horizontal inflection) in satisfaction before the highest decile in any year
+- Any inflection point in satisfaction was as likely to be convex as concave
 
@@
 p3 <- happywealth_covs %>%
-  plot_inflection_piecewise(gh9, adj_hifdip)
+  plot_inflection_custom(gh9, adj_hifdip, knots = c(32, 46, 52.5, 86, 124.5))
 
@@
 
-- There is no happiness satiety point (horizontal inflection) before the final decile
-- A concave happiness inflection point is evident before or near the median decile
+- The linear relationship between happiness and wealth became weaker (less steep) over years
+- There was no happiness satiety point (horizontal inflection) before the final decile in any year
+- A concave inflection point with happiness occurred in each year, and appeared to shift right with time.
 
@@
   summarise(
-    `median ($000s)` = median(adj_hifdip),
-    `change point ($000s)` = mean(change_point),
-    `%` = mean(adj_hifdip > mean(change_point))
+    `median ($000s)` = round(median(adj_hifdip), 2),
+    `change point ($000s)` = round(mean(change_point), 2),
+    `%` = round(mean(adj_hifdip > mean(change_point)), 2)
   ) %>%
-  kable(format = "html", caption = "Household wealth median, change point and percentile")
+  kable(format = "html", 
+        caption = "Household wealth median, change point and percentile")
 ```

### 2020-09-10 src/happyfun.R

@@ -54,7 +54,7 @@ plot_linear <- function(.data, swb_col, dollars_col) {
   
 }
 
-plot_inflection_piecewise <- function(.data, swb_col, dollars_col) {
+plot_inflection_piecewise <- function(.data, swb_col, dollars_col, knots) {
   
   swb_colname <- rlang::enexpr(swb_col)
   dollars_colname <- rlang::enexpr(dollars_col)
@@ -76,14 +76,12 @@ plot_inflection_piecewise <- function(.data, swb_col, dollars_col) {
   df %>%
     group_by(year) %>%
     mutate(
-      dollar_bin = ntile(dollars, 10),
-      knots = attr(splines::bs(dollars, degree = 1, df = 2), which = "knots")
+      dollar_bin = ntile(dollars, 10)
     ) %>%
     group_by(year, dollar_bin) %>%
     summarize(
       wellbeing = mean(swb, na.rm=T),
       wealth = mean(dollars, na.rm=T)/1000,
-      knot = unique(knots)/1000,
       n = n()
     ) -> df.summary
   
@@ -94,6 +92,74 @@ plot_inflection_piecewise <- function(.data, swb_col, dollars_col) {
                 aes(x = wealth, y = wellbeing),
                 method = lm, 
                 formula = y ~ splines::bs(x, df = 2, degree = 1),
+                size = 0.5) +
+    coord_cartesian(xlim = c(0, max(df.summary$wealth)),
+                    ylim = c(min(df.summary$wellbeing), max(df.summary$wellbeing))) +
+    facet_wrap(~year, ncol = 1) +
+    labs(subtitle = paste(rlang::as_name(swb_colname), "(piecewise)"),
+         x = paste(rlang::as_name(dollars_colname), "($000s)"),
+         y = "") +
+    theme_test()
+}
+
+plot_inflection_custom <- function(.data, swb_col, dollars_col, knots) {
+  
+  swb_colname <- rlang::enexpr(swb_col)
+  dollars_colname <- rlang::enexpr(dollars_col)
+  
+  .data %>%
+    select(xwaveid, year, 
+           # swb = enexpr(swb_col), dollars = enexpr(dollars_col)) %>%
+           swb = !!swb_colname, dollars = !!dollars_colname) %>%
+    filter(swb >= 0) %>%
+    filter(dollars >= 0) %>%
+    filter(year %in% c(2002, 2006, 2010, 2014, 2018)) -> df
[…]

### 2020-09-12 drafts/draft0.Rmd
added Clark fig 4, transposed plots

@@
 
-### Introduction
+## Background
 ***
@@
 For instance, we have recently reported that positive life events, such as a major financial windfall, can have a greater impact on an individual's satisfaction than their happiness [@kettlewell2020differential].
-However these effects were observed as a change within individuals, and the change was transitory as both satisfaction and happiness returned to baseline after two years [see also @frijters2011life].
-Across individuals, wealth also has a distinct relationship with happiness and satisfaction as well.
-While Kahneman et al observed the association between household wealth and happiness plateaud after $75,000, they also reported the association between wealth and life satisfaction was consistently positive.
-Indeed, several researchers have observed a difference between two questions that are often used in surveys of subjective well-being: “How satisfied are you with your life?” and “How happy are you these days?” [@veenhoven2006rising; @howell2008relation].
+However these effects were observed as a change within individuals, and the change was temporary as both satisfaction and happiness returned to baseline after two years [see also @frijters2011life].
+Wealth also has a distinct relationship with happiness and satisfaction across individuals was well.
+As described above, Kahneman et al (2010) observed the association between household wealth and happiness plateaud after $75,000, but they also reported that life satisfaction continued to increase with wealth.
+Indeed, several researchers have observed a similar difference between two questions that are often used in surveys of subjective well-being: “How satisfied are you with your life?” and “How happy are you these days?” [@veenhoven2006rising; @howell2008relation].
 The common conclusion is that wealth is more strongly related to satisfaction than to happiness.
+
+Research since 2010 has revealed variations in the plateau or _change point_ due to [world region, gender, and education](https://www-nature-com.ezproxy2.library.usyd.edu.au/articles/s41562-017-0277-0){target="_blank"} [@jebb2018happiness].
+However to date there has been no investigation as to whether the relationship between wealth and happiness changes over time.
+The value of money of course generally _decreases_ over time with inflation and the cost of living and so this will _raise_ the cost of happiness, however it will not change the cost of happiness over time in real dollar terms (i.e., adjusted for inflation).
+The distribution of wealth has also changed over the last 30 years in many OECD countries including Australia as the rich got richer while the poor stayed poor (ref needed).
+Such changes in the distribution of wealth may increase the cost of happiness if it is derived by relative wealth (e.g., Clarke et al 2008).
+That is, as the ceiling of wealth in dollar terms is raised over time, for any fixed income level at *t*<sub>0</sub> (i.e., the base year) there will be a greater disparity relative to the highest wealth percentile.
+This growing disparity may produce corresponding changes in the level of wealth at which people no longer judge their happiness to be increased relative to others.
 
@@
 If there are distinct effects of wealth on happiness and satisfaction then the question of which is better suited to assess human welfare and to guide policy seems critical.
-The observation that happiness plateaus with increasing wealth while satisfaction keeps rising acquires some importance in this context.
-Not everyone will agree that enhancing the happiness experienced by those who are already quite happy is a legitimate policy objective.
-
-And while research since 2010 has revealed variations in the plateau or _change point_ due to [world region, gender, and education](https://www-nature-com.ezproxy2.library.usyd.edu.au/articles/s41562-017-0277-0){target="_blank"} [@jebb2018happiness], there has been no investigation as to whether the relationship between wealth and happiness is changing over time [although see Fig.
-4 in @clark2008relative; for a speculative mechanism].
-The value of money of course changes over time with inflation, however the distribution of wealth has also changed over the last 30 years in many OECD countries including Australia (as the rich got richer faster than the poor, ref needed).
-Such changes in the distribution of wealth are also likely to change the relationship with wellbeing when it is governed by relative wealth rather than absolute wealth [@clark2008relative].
+The observation that happiness plateaus with increasing wealth while satisfaction keeps rising acquires substantial importance in this context.
+Not everyone will agree that enhancing the happiness experienced by those who are already very happy is a legitimate policy objective.
+Likewise, if the point at which wealth no longer increases happiness is increasing (or decreasing) over time, then policy objectives may need to change to reflect this.
+<span style="color: blue"> (This section needs a clearer statement on the importance of determining the increasing cost of happiness) </span>
 
@@
 We distinguished between satisfaction and happiness as different components of subjective wellbeing, and evaluated how each varies with household wealth.
-Economic security is better represented by household wealth, since members of the same household share expenses as well as risks; i.e., they can provide a direct and immediate support network when financial shocks occur.
-The other major studies also used household after-tax income as the indicator of wealth and economic security, and so we follow the same standard here as well.
-
 After adjusting for age, gender, and education level, we find that satisfaction increases linearly with household wealth, but the same purchasing power has diminishing returns on happiness.
@@
 
+## Methods
+
 #### Wealth
-We have a variety of income & wealth information in HILDA, most of it collected annually.
-However the household net worth is collected every four years.
-In the present report we use [hifdip](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=hifdip){target="_blank"} Household financial year disposable regular income ($).
-The sum across all household members of financial year gross regular income minus taxes on financial year gross regular income.
-Wealth variables were provided by the University of Melbourne as imputed, and with weighted topcodes to preserve anonymity.
-_Nb_.
-We also examined household and personal gross wages and disposable income (not shown).
+Our wealth variable was household after-tax income.
+Economic security is better represented by household wealth, since members of the same household share expenses as well as risks; i.e., they can provide a direct and immediate support network when financial shocks occur.
+The other major studies also used household after-tax income as the indicator of wealth and economic security [e..g, @kahneman2010high; @jebb2018happiness], and so we follow the same standard here as well.
+Household financial year disposable regular income [(hifdip)](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=hifdip){target="_blank"} from every four years of HILDA (2002, 2006, 2010, 2014, 2018) was used.
+This variable represents the sum across all household members of financial year gross regular income less taxes on financial year gross regular income.
 
+Household income was provided by the University of Melbourne as imputed, and with weighted topcodes to preserve anonymity.
 We removed individuals with values above the threshold indicated for top-coding.
-Household wealth was adjusted for the number of people in each household in the same manner as Headey, Muffels & Wooden, 2004; i.e., income divided by the square root of household size (Buhmann et al, 1988; Coulter et al, 1992).
+Household wealth was adjusted for the number of people in each household by dividing by the square root of household size [headey2004effects; jebb2018happiness; kahneman2010high].
 
@@
 #### Subjective Wellbeing
-There are a variety of variables related to subjective well-being collected annually, but the two we used here matched the variables we used in our previous paper:
-
-1. `losat` Life satisfaction - How satisfied are you with your life (0 to 10)
-2. `gh9a` to `gh9i` Affective wellbeing - Sum of SF-36 item 9 (mental health and vitality)
-
-_Note_. We also examined `ghmh` (the 5-item subset of mental health questions, sometimes called the **MHi-5**, from the SF-36), and the patterns were almost identical to our sum of 9a to 9i.
+There are a variety of variables related to subjective well-being collected annually in HILDA, but the two we used here matched the variables we used in our previous paper [@kettlewell2020differential], namely 1) life satisfaction [(losat)](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=losat){target="_blank"}, a single item question asked each survey "How satisfied are you with your life (0 to 10)", and 2) happiness determined by the sum of SF-36 item 9 on mental health and vitality [(gh9a to gh9i)](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=gh9a){target="_blank"}.
+The SF-36 is a widely used self-completion measure of various aspects of physical, emotional and mental health [@ware2000sf].
+We calculated a summed score based on nine questions (9a-9i), where five questions dealt with positive and negative aspects of mental health (e.g., "Felt so down in the dumps nothing could cheer me up", "Been happy"), and four questions dealt with positive and negative aspects of vitality (e.g., "feel full of life", "felt worn out").
+We reverse scored relevant items to ensure the summed score represented better wellbeing with higher scores.
 
@@
 Thus in each plot, the mean wealth and mean wellbeing score for each decile is presented, rather than each individual data point.
-Note that the line-of-best-fit and 95% confidence intervals from the broken stick regression model _of all individuals_ is shown in overlay.
+Note that the line-of-best-fit and 95% confidence intervals from each regression model _of all individuals_ is shown in overlay.
 
@@
 
-### Results
+## Results
 ***
@@
 #### Household wealth on satisfaction (cognitive wellbeing)
-```{r losat_hwnwip, fig.height=11}
+```{r losat_hwnwip, fig.height=8, fig.width=9}
 library(patchwork)
@@
 
-(p1 + p2 + p3) +
+(p1 / p2 / p3) +
   plot_annotation(title = "Effect of wealth on satisfaction")
@@
 #### Household wealth on happiness (affective wellbeing)
-```{r gh9_hwnwip, fig.height=11}
+```{r gh9_hwnwip, fig.height=8, fig.width=9}
 library(patchwork)
@@
 
-(p1 + p2 + p3) +
+(p1 / p2 / p3) +
   plot_annotation(title = "Effect of wealth on happiness")
@@
 
-- The change point for happiness (affective wellbeing) shifts to the right over time and occurs at greater household wealth levels.
+- The level of wealth at which happiness no longer increased at the same rate (the _cost of happiness_) credibly increased between 2002 and 2018, and even as early as 2014.
 
@@
 
-The mean change point location has moved further right (higher, faster) than the median household wealth over time:
+The cost of happiness increased faster than median household wealth. As a result, a smaller percentage of people achieved a level of wealth at which happiness no longer increased in 2018 than in 2002:
 
@@
 
-### Conclusions
+## Conclusions
 ***
@@
 
-### References
+## References

### 2020-09-12 src/happyfun.R

@@ -9,8 +9,8 @@ plot_linear <- function(.data, swb_col, dollars_col) {
            year, 
            # swb = enexpr(swb_col), dollars = enexpr(dollars_col)) %>%
            swb = !!swb_colname, dollars = !!dollars_colname) %>%
-    filter(swb > 0) %>%
-    filter(dollars > 0) %>%
+    filter(swb >= 0) %>%
+    filter(dollars >= 0) %>%
     filter(year %in% c(2002, 2006, 2010, 2014, 2018)) -> df
   
   df %>%
@@ -39,7 +39,7 @@ plot_linear <- function(.data, swb_col, dollars_col) {
       se = TRUE) +
     coord_cartesian(xlim = c(0, max(df.summary$wealth)),
                     ylim = c(min(df.summary$wellbeing), max(df.summary$wellbeing))) +
-    facet_wrap(~year, ncol = 1) +
+    facet_wrap(~year, nrow = 1) +
     labs(subtitle = paste(rlang::as_name(swb_colname), "(linear)"),
          x = paste(rlang::as_name(dollars_colname), "($000s)"),
          y = "") +
@@ -52,6 +52,65 @@ plot_linear <- function(.data, swb_col, dollars_col) {
   p1 +
     geom_point(data = df.summary) 
   
+}
+
+plot_inflection_cubic <- function(.data, swb_col, dollars_col) {
+  
+  swb_colname <- rlang::enexpr(swb_col)
+  dollars_colname <- rlang::enexpr(dollars_col)
+  
+  .data %>%
+    select(xwaveid, year, 
+           # swb = enexpr(swb_col), dollars = enexpr(dollars_col)) %>%
+           swb = !!swb_colname, dollars = !!dollars_colname) %>%
+    filter(swb >= 0) %>%
+    filter(dollars >= 0) %>%
+    filter(year %in% c(2002, 2006, 2010, 2014, 2018)) -> df
+  
+  df %>%
+    mutate(
+      wealth = dollars/1000,
+      wellbeing = swb
+    ) -> df.plot
+  
+  df %>%
+    group_by(year) %>%
+    mutate(
+      dollar_bin = ntile(dollars, 10)
+    ) %>%
+    group_by(year, dollar_bin) %>%
+    summarize(
+      wellbeing = mean(swb, na.rm=T),
+      wealth = mean(dollars, na.rm=T)/1000,
+      n = n()
+    ) -> df.summary
+  
[…]

### 2020-09-12 src/preprocessing.r

@@ -1,0 +1,178 @@
+#### Preprocessing ####
+
+library(haven)
+
+#### Load data ####
+path_to_hilda <- list.files(
+  path = '~/Dropbox/HILDA/data',
+  pattern = '^Combined.*.dta$',
+  full.names = TRUE
+)
+
+hilda <- list()
+for (pathtofile in path_to_hilda) {
+  df <- read_dta(pathtofile)
+  hilda <- append(hilda, list(df))
+  cat('.')
+}
+
+# Helper function
+source('../../src/gather_hilda.R')
+
+# demographic data 
+gather_hilda(hilda, c(
+  "esbrd",   # employment status (broad)
+  "hhda10",  # SEIFA 2001 Decile of socio-economic advantage (higher is better)
+  "hgage",   # age
+  "hgsex",   # sex (male = 1)
+  "helth",   # long term health condition (yes = 1)
+  "hhpers",  # no. persons in household
+  "hhrhid",  # household id (within wave)
+  "mrcurr",  # current marital status (married/de facto ≤ 2)
+  "edhigh1", # highest education achieved
+  "edfts"    # current fulltime student (yes = 1)
+)) %>%
+  spread(code, val) %>%
+  mutate_if(is.double, ~ ifelse(. < 0, NA_real_, .)) -> demographic_items
+
+# financial data
+gather_hilda(hilda, c(
+  "wsfei",   # financial year gross
+  "tifdip",  # Disposable regular income
+  "tifeftp", # Gross total income (including benefits, pensions, foreign, etc)
+  "hiwsfei", # household financial year gross
+  "hwnwip",  # household net worth
+  "hifdip",  # household disposable regular income
+  "hifditp"  # household disposable total income
+)) %>%
+  spread(code, val) %>%
+  mutate_if(is.double, ~ ifelse(. < 0, NA_real_, .)) -> wealth_items
+
+# Affective wellbeing:
+gather_hilda(hilda, c(
+  'losat',
+  'gh9a', # Vitality: feel full of life (lower is better)*
+  'gh9b', # Mental H: Been a nervous person (higher is better)
+  'gh9c', # Mental H: Felt so down in the dumps (higher is better)
+  'gh9d', # Mental H: Felt calm and peaceful (lower is better)*
[…]

### 2020-09-12 drafts/draft0.Rmd
Added model comparisons

@@
 ## Background
-***
 A fundamental question for the psychology and economics of wellbeing is just how much happiness does a dollar buy?
@@
 But after that, increasing amounts of money had no further effect on happiness [@kahneman2010high].
-Results such as these have reinforced the view among psychologists and economists that the relationship between wealth and measures of subjective wellbeing, such as happiness, increases linearly up to a point after which there is little gained by further increments in wealth [@clark2018four; @dolan2008we].
+Results such as these have reinforced the view among psychologists and economists that the relationship between wealth and measures of subjective wellbeing, such as happiness, increases linearly up to a point (i.e., a _change point_) after which there is little gained by further increments in wealth [@clark2018four; @dolan2008we].
+We suggest that the point at which wealth produces no further increases in happiness might be described as the _cost_ of happiness.
 
@@
 
-Research since 2010 has revealed variations in the plateau or _change point_ due to [world region, gender, and education](https://www-nature-com.ezproxy2.library.usyd.edu.au/articles/s41562-017-0277-0){target="_blank"} [@jebb2018happiness].
-However to date there has been no investigation as to whether the relationship between wealth and happiness changes over time.
+Research since 2010 has revealed variations in the change point or cost of happiness due to [world region, gender, and education](https://www-nature-com.ezproxy2.library.usyd.edu.au/articles/s41562-017-0277-0){target="_blank"} [@jebb2018happiness].
+However to date there has been no investigation as to whether the _change point_ varies over time.
+That is, has the point at which happiness no longer increases with wealth itself changed over time?
 The value of money of course generally _decreases_ over time with inflation and the cost of living and so this will _raise_ the cost of happiness, however it will not change the cost of happiness over time in real dollar terms (i.e., adjusted for inflation).
@@
 Such changes in the distribution of wealth may increase the cost of happiness if it is derived by relative wealth (e.g., Clarke et al 2008).
-That is, as the ceiling of wealth in dollar terms is raised over time, for any fixed income level at *t*<sub>0</sub> (i.e., the base year) there will be a greater disparity relative to the highest wealth percentile.
-This growing disparity may produce corresponding changes in the level of wealth at which people no longer judge their happiness to be increased relative to others.
+When the ceiling of wealth in real dollar terms is raised over time; for any fixed income level at *t*<sub>0</sub> (i.e., the base year) there will be a greater disparity relative to the highest wealth percentile at *t*<sub>n+1</sub>.
+If happiness is, even partly, determined by this relative disparity then increasing the disparity will reduce happiness amongst the poorest percentiles (while leaving happiness amongst the richest percentiles unchanged).
+This may increase the level of wealth at which happiness no longer increases.
+That is, as the rich get richer, the poor get unhappier due to the increasing relative wealth disparity and so the cost of happiness increases.
+<span style="color: blue"> (This section may need more clarity) </span>
 
-The relationship between wealth and happiness and satisfaction are relevant for the growing popularity of using subjective well-being as a policy guide for government and agencies.
-If there are distinct effects of wealth on happiness and satisfaction then the question of which is better suited to assess human welfare and to guide policy seems critical.
-The observation that happiness plateaus with increasing wealth while satisfaction keeps rising acquires substantial importance in this context.
-Not everyone will agree that enhancing the happiness experienced by those who are already very happy is a legitimate policy objective.
-Likewise, if the point at which wealth no longer increases happiness is increasing (or decreasing) over time, then policy objectives may need to change to reflect this.
+The relationship between wealth and happiness and satisfaction are increasingly relevant to governments and policy-makers due to the growing recognition that increasing wealth does not necessarily lead to improvements in wellbeing [@clark2018four; @frijters2020happy].
+If satisfaction but not happiness increases with wealth, then the pursuit of both may not be a feasible policy.
+Likewise, if happiness increases up to a point with wealth, or the cost of happiness changes over time, then not everyone will agree that enhancing the happiness of those who are already very happy is a legitimate policy objective.
+Conversely, it may emphasize that the relationship between wealth and happiness is malleable and thus amenable to policy objectives.
 <span style="color: blue"> (This section needs a clearer statement on the importance of determining the increasing cost of happiness) </span>
@@
 After adjusting for age, gender, and education level, we find that satisfaction increases linearly with household wealth, but the same purchasing power has diminishing returns on happiness.
-We also discover that the location of the inflection point has increased in real dollar value since 2002.
+We also discover that the location of the change point (the _cost_ of happiness) has increased in real dollar terms since 2002.
 
@@
 #### Wealth
-Our wealth variable was household after-tax income.
-Economic security is better represented by household wealth, since members of the same household share expenses as well as risks; i.e., they can provide a direct and immediate support network when financial shocks occur.
+Our indicator of wealth and economic security was household after-tax income.
+Household income better represents economic security, since members of the same household share expenses as well as risks; i.e., they can provide a direct and immediate support network when financial shocks occur.
 The other major studies also used household after-tax income as the indicator of wealth and economic security [e..g, @kahneman2010high; @jebb2018happiness], and so we follow the same standard here as well.
-Household financial year disposable regular income [(hifdip)](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=hifdip){target="_blank"} from every four years of HILDA (2002, 2006, 2010, 2014, 2018) was used.
-This variable represents the sum across all household members of financial year gross regular income less taxes on financial year gross regular income.
+Household financial year disposable regular income [(hifdip)](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=hifdip){target="_blank"} from the annual HILDA survey in 2002, 2006, 2010, 2014,and 2018 was selected.
+This variable in HILDA represents the sum across all household members of financial year gross regular income less taxes on financial year gross regular income.
 
@@
 We removed individuals with values above the threshold indicated for top-coding.
-Household wealth was adjusted for the number of people in each household by dividing by the square root of household size [headey2004effects; jebb2018happiness; kahneman2010high].
+Household wealth was adjusted for the number of people in each household by dividing by the square root of household size [@headey2004effects; @jebb2018happiness; @kahneman2010high].
+We converted dollar values to the base year (2002) according to CPI from the RBA, in order to adjust for inflation (https://www.rba.gov.au/calculator/annualDecimal.html).
 
@@
 #### Subjective Wellbeing
-There are a variety of variables related to subjective well-being collected annually in HILDA, but the two we used here matched the variables we used in our previous paper [@kettlewell2020differential], namely 1) life satisfaction [(losat)](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=losat){target="_blank"}, a single item question asked each survey "How satisfied are you with your life (0 to 10)", and 2) happiness determined by the sum of SF-36 item 9 on mental health and vitality [(gh9a to gh9i)](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=gh9a){target="_blank"}.
+There are a variety of variables related to subjective well-being collected annually in HILDA, but the two we used here matched the variables we used in our previous paper [@kettlewell2020differential].
+Namely, 1) life satisfaction [(losat)](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=losat){target="_blank"}, a single item question asked each survey "How satisfied are you with your life (0 to 10)", and 2) happiness determined by item 9 in the SF-36 [(gh9a to gh9i)](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=gh9a){target="_blank"}.
 The SF-36 is a widely used self-completion measure of various aspects of physical, emotional and mental health [@ware2000sf].
-We calculated a summed score based on nine questions (9a-9i), where five questions dealt with positive and negative aspects of mental health (e.g., "Felt so down in the dumps nothing could cheer me up", "Been happy"), and four questions dealt with positive and negative aspects of vitality (e.g., "feel full of life", "felt worn out").
-We reverse scored relevant items to ensure the summed score represented better wellbeing with higher scores.
+Item 9 consists of nine questions relating to mental health and vitality, where five questions deal with positive and negative aspects of mental health (e.g., "Felt so down in the dumps nothing could cheer me up", "Been happy"), and four questions deal with positive and negative aspects of vitality (e.g., "feel full of life", "felt worn out").
+We reverse scored the relevant responses and calculated a summed score so that better wellbeing was represented by higher scores.
 
@@
 
-#### Modelling and model-selection
+#### Modelling
 
-_Modelling_  
-To initially describe the relationship between wealth and wellbeing, we modelled the data from each year in three different ways: 1) a simple linear model of the form y<sub>*i*</sub> = *&beta;*<sub>0</sub> + *&beta;*<sub>1</sub>*x*<sub>*i*</sub> + *&epsilon;*<sub>*i*</sub>; 2) a cubic b-spline model of the form *B*(*x*) = *&Sigma;&beta;*<sub>*i*</sub>*B*<sub>*i,n*</sub>(*x*); and 3) a piecewise (broken-stick) model of the form y<sub>*i*</sub> = *&beta;*<sub>0</sub> + *&beta;*<sub>1</sub>(*x*<sub>*i*</sub> - &omega;)(*x*<sub>*i*</sub> ≤ &omega;) + *&beta;*<sub>2</sub>(*x*<sub>*i*</sub> - &omega;)(*x*<sub>*i*</sub> ≤ &omega;).
+We initially describe the relationship between wealth and each wellbeing variable (happiness and satisfaction) using three different models: 1) a simple **linear model**; 2) a cubic **B-spline model**; and 3) a **piecewise model** (broken-stick).
+These three models were chosen to represent varying degrees of complexity in the relationship between wealth and wellbeing.
+The linear model was the simplest relationship while the cubic B-spline was the most complex model (more degrees of freedom), and chosen because it has been recently used to describe the maximum point at which wellbeing no longer increases with wealth [i.e., "satiety", @jebb2018happiness].
+Following Jebb et al, we restricted the cubic B-spline model to *n* = 4, 5 and 6 knots and show each instantiation here.
+The piecewise model was chosen as the simplest extension of a linear model which can identify a change point (inflection) between wellbeing and wealth.
+The location of the change point was a free parameter which revealed where wellbeing no longer increased linearly with wealth.
 
-These three models were chosen to represent varying degrees of complexity.
-The linear model was obviously the simplest relationship that can exist between wealth and wellbeing.
-The cubic B-spline was a complex model (more degrees of freedom) and chosen because it has been recently used to describe the maximum point at which wealth no longer increases wellbeing [@jebb2018happiness].
-Following Jebb et al, we restricted the model to 4, 5 and 6 knots and show each instantiation here.
-The piecewise model was chosen as the simplest extension of a linear model to identify the change point at which wellbeing no longer increases linearly with wealth.
+We then compared the linear model against the piecewise model to determine if a change point existed in each or any year between wealth and each wellbeing variable (model selection).
+Finally, where a change point existed, we determined the location of the change point for that year (change point estimation).
 
-_Change point estimation_  
-To determine whether the relationship between wellbeing and wealth contained an inflection point ("change point"), we modelled the relationship between wealth and wellbeing across individuals using the piecewise model described above, and estimated the posterior probability of the changepoint location using a Bayesian approach:
+_Model estimation_  
+We adopted a Bayesian approach for estimating the linear and piecewise model in the software Stan [@burkner2017; @rstan2019]. In each case,
 
@@
 
-*&mu;*<sub>i</sub> = *&beta;*<sub>0</sub> + *&beta;*<sub>1</sub>(*x*<sub>*i*</sub> - &omega;<sub>*t*</sub>)(*x*<sub>*i*</sub> ≤ &omega;<sub>*t*</sub>) + *&beta;*<sub>2</sub>(*x*<sub>*i*</sub> - &omega;<sub>*t*</sub>)(*x*<sub>*i*</sub> > &omega;<sub>*t*</sub>)  \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \  (1)
+The linear model was estimated as:
 
-Where *x*<sub>i</sub> and *y*<sub>i</sub> was an individuals' household wealth (\$) and wellbeing, respectively; and &omega;<sub>*t*</sub> was the changepoint in $ for each year (*t*).
+*&mu;*<sub>i</sub> = *&beta;*<sub>0</sub> + *&beta;*<sub>1</sub>*X*<sub>*i*</sub>   \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \  (1)
 
-The location of the changepoint was sampled randomly from a uniform prior over the resticted range of wealth values, and the posterior probability of the location for each year is presented.
+Where *X*<sub>i</sub> was an individual's household wealth (\$) as well as other covariates (age, age^2^, gender, education), and *y*<sub>i</sub> was an individual's wellbeing.
 
@@
 
+The piecewise model was a simple extension of this to include a free parameter to represent the changepoint in wealth (&omega;) as well as the slope after the change point (*&beta;*<sub>2</sub>):
+
+*&mu;*<sub>i</sub> = *&beta;*<sub>0</sub> + *&beta;*<sub>1</sub>(*x*<sub>*i*</sub> - &omega;<sub>*t*</sub>)(*x*<sub>*i*</sub> ≤ &omega;<sub>*t*</sub>) + *&beta;*<sub>2</sub>(*x*<sub>*i*</sub> - &omega;<sub>*t*</sub>)(*x*<sub>*i*</sub> > &omega;<sub>*t*</sub>)  \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \  (2)
+
+<br>
+
+The above models estimated population-level (fixed) effects for each year in the data (*t* = 2002, 2006, 2010, 2014 and 2018).
+Because we were interested in the location of the change point between wealth and wellbeing that existed across individuals within each year, we ignored the panel design of HILDA because the dependency between observations of the same person across years was orthogonal to our effects of interest.
+We assumed an unstructured variance-covariance matrix between years, and specified weakly informed priors for each *&beta;*, and a uniform prior over the restricted range of wealth values for *&omega;*.
+
 _Model selection_  
-To determine whether wellbeing was a linear or non-linear (e.g., piecewise) function of wealth, we compared the linear and piecewise model fits using the [Widely Applicable Information Criterion (WAIC)](https://bookdown.org/ajkurz/Statistical_Rethinking_recoded/overfitting-regularization-and-information-criteria.html#the-problem-with-parameters){target="_blank"}.
+To determine whether wellbeing was a linear or non-linear (e.g., piecewise) function of wealth, we compared the linear and piecewise model posterior probabilities using the [Widely Applicable Information Criterion (WAIC)](https://bookdown.org/ajkurz/Statistical_Rethinking_recoded/overfitting-regularization-and-information-criteria.html#the-problem-with-parameters){target="_blank"}.
 The WAIC is the log-posterior predictive density plus a penalty proportional to the variance in the posterior distribution.
 Thus it provides an approximation of the out-of-sample deviance that converges to the cross-validation approximation in a large sample, with a penalty for the effective number of parameters (degrees of freedom).
-For this reason is it useful to compare two models of varying complexity.
+For this reason is it useful to compare two models of varying complexity, such as our linear and piecewise model.
 
@@
 
-For visualization purposes only, due to the large number of individual data points in each year, we grouped individuals into 10 equal-sized subgroups according to their wealth decile ($).
-Thus in each plot, the mean wealth and mean wellbeing score for each decile is presented, rather than each individual data point.
-Note that the line-of-best-fit and 95% confidence intervals from each regression model _of all individuals_ is shown in overlay.
-
-```{r funtimes}
-source("../src/happyfun.r")
-```
+_Change point estimation_  
+To determine the location of the change point between wellbeing and wealth, we modelled the relationship between wealth and wellbeing across individuals using the piecewise model described above, and estimated the posterior probability of the changepoint location.
+The posterior distribution of the change point location for each year is presented.
 
@@
 ## Results
-***
 
-#### Household wealth on satisfaction (cognitive wellbeing)
+```{r funtimes}
+source("../src/happyfun.r")
+```
+
+For visualization purposes only, due to the large number of individual data points in each year, we grouped individuals into 10 equal-sized subgroups according to their wealth decile ($).
+Thus in each plot, the mean wealth and mean wellbeing score for each decile is presented, rather than each individual data point.
+Note that the line-of-best-fit and 95% confidence intervals from each regression model _of all individuals_ is shown in overlay.
+
+#### Household wealth and satisfaction (cognitive wellbeing)
 ```{r losat_hwnwip, fig.height=8, fig.width=9}
@@
 
-- Generally the relationship between satisfaction and household wealth appeared linear
-- The linear relationship between satisfaction and wealth became weaker (less steep) over years
-- There is no satiety point (horizontal inflection) in satisfaction before the highest decile in any year
-- Any inflection point in satisfaction was as likely to be convex as concave
+- The relationship between satisfaction and household wealth appeared linear
+- The linear relationship between satisfaction and wealth became slightly weaker (less steep) over years
+- The cubic B-spline models indicated no satiety point (horizontal inflection) in satisfaction before the highest decile in any year
+- Any change point in satisfaction was as likely to be convex as concave
 
@@
 
-#### Household wealth on happiness (affective wellbeing)
+#### Household wealth and happiness (affective wellbeing)
 ```{r gh9_hwnwip, fig.height=8, fig.width=9}
@@
 
-- The linear relationship between happiness and wealth became weaker (less steep) over years
-- There was no happiness satiety point (horizontal inflection) before the final decile in any year
-- A concave inflection point with happiness occurred in each year, and appeared to shift right with time.
+- A change point is evident in each year where the linear relationship between happiness and wealth became abruptly weaker (less steep)
+- The overall linear relationship between happiness and wealth, as well as the pre-inflection slope, became weak (less steep) over years
+- The cubic B-spline model indicated there was no happiness satiety point (horizontal inflection) before the final decile in any year
+- The change point between happiness and wealth appeared to shift right with time.
+
+<br>
+
+#### Model selection
+
+We compared the posterior evidence for a linear relationship between wellbeing and wealth with the posterior evidence of a nonlinear (piecewise) relationship (WAIC~linear~ — WAIC~piecewise~).
+Thus a WAIC difference greater than zero indicates evidence for a linear relationship.
+A WAIC difference less than zero indicates evidence for a nonlinear (piecewise) relationship.
+
+##### Model evidence for linear satisfaction
+
+```{r plot_waic_losat}
+waic_result <- read_rds("../results/waic_losat.RDS")
+
+d1 <- data.frame(fy2018 = rnorm(n = 1e5, 
+                                     mean = waic_result[2, 1] * -2, 
+                                     sd   = waic_result[2, 2] *  2)/10,
+                 fy2014 = rnorm(n = 1e5, 
+                                     mean = waic_result[4, 1] * -2, 
+                                     sd   = waic_result[4, 2] *  2)/10,
+                 fy2010 = rnorm(n = 1e5, 
+                                     mean = waic_result[6, 1] * -2, 
+                                     sd   = waic_result[6, 2] *  2)/10,
+                 fy2006 = rnorm(n = 1e5, 
+                                     mean = waic_result[8, 1] * -2, 
+                                     sd   = waic_result[8, 2] *  2)/10,
+                 fy2002 = rnorm(n = 1e5, 
+                                     mean = waic_result[10, 1] * -2, 
+                                     sd   = waic_result[10, 2] *  2)/10
+                )
+
+color_scheme_set("blue")
+mcmc_areas(x = d1,
+  pars = colnames(d1),
+  prob = 0.90,
+  prob_outer = 1
+  ) +
+  scale_x_continuous(
+    labels = function(x) {x*10}) +
+  geom_vline(aes(xintercept = 0), color = "grey90") +
+  theme_test() +
+  labs(
+    subtitle = "Nonlinear over linear fit for satisfaction (mean difference and 90% distribution)",
+    x = "WAIC difference (90% shaded)")
+```
+
+<br>
+
+- A linear fit between household wealth and satisfaction was credibly superior to a nonlinear fit in each year  
+<br>
+
+##### Model evidence for nonlinear happiness
+```{r plot_waic_gh9}
+waic_result <- read_rds("../results/waic_comparisons_gh9_percentiles.RDS")
+
+d1 <- data.frame(fy2018 = rnorm(n = 1e5, 
+                                mean = waic_result[10, 1] * 2, 
+                                sd   = waic_result[10, 2] *  2)/10,
+                 fy2014 = rnorm(n = 1e5, 
+                                mean = waic_result[8, 1] * 2, 
+                                sd   = waic_result[8, 2] *  2)/10,
+                 fy2010 = rnorm(n = 1e5, 
+                                mean = waic_result[6, 1] * 2, 
+                                sd   = waic_result[6, 2] *  2)/10,
+                 fy2006 = rnorm(n = 1e5, 
+                                mean = waic_result[4, 1] * 2, 
+                                sd   = waic_result[4, 2] *  2)/10,
+                 fy2002 = rnorm(n = 1e5, 
+                                mean = waic_result[2, 1] * 2, 
+                                sd   = waic_result[2, 2] *  2)/10
+)
+
+color_scheme_set("red")
+mcmc_areas(x = d1,
+           pars = colnames(d1),
+           prob = 0.90,
+           prob_outer = 1
+) +
+  scale_x_continuous(
+    labels = function(x) {x*10}) +
+  geom_vline(aes(xintercept = 0), color = "grey90") +
+  theme_test() +
+  labs(
+    subtitle = "Nonlinear over linear fit for happiness (mean difference and 90% distribution)",
+    x = "WAIC difference (90% shaded)")
+```
+
+<br>
+
+- A nonlinear fit between household wealth and happiness was superior to a linear fit in each year 
+- The nonlinear fit was credibly superior to a linear fit at 90% for each of three years (2018, 2014, 2006)
+
+<br>
+
+**Key points**
+
+- The posterior evidence indicates a linear model with no inflection point explains the relationship between satisfaction and household wealth in each year
+- However a nonlinear fit with at least one inflection point better explained the relationship between happiness and household wealth
 
@@
 ```{r happy_change_time, fig.width = 4.9, fig.height = 9}
-hifdip_gh9_fixed_fit <- read_rds("../results/hifdip_gh9_fixed_full.RDS")
+hifdip_gh9_fixed_fit <- read_rds("../results/hifdip_gh9_fixed_full_covs.RDS")
 
@@
   prob = 0.95, # 80% intervals
-  prob_outer = 1, # 99%
+  prob_outer = .99, # 99%
   point_est = "mean"
@@
 
-- The level of wealth at which happiness no longer increased at the same rate (the _cost of happiness_) credibly increased between 2002 and 2018, and even as early as 2014.
+- The level of wealth at which happiness no longer increased with wealth (the _cost of happiness_) credibly increased between 2002 and 2018.
 
@@
 ## Conclusions
-***
 
-The main conclusions of this report so far are:
+We observed the relationship between subjective wellbeing and wealth was positive, but happiness and satisfaction have different (positive) relationships.
+Satisfaction increased linearly with wealth while happiness increased rapidly up to a point, after which further increments in wealth produced less change - we referred to this as the **cost** of happiness.
+So household wealth, and the economic security it represents, buys happiness up to a point after which happiness and wealth are rarely related.
+Satisfaction on the other hand always increased with household wealth.
+The difference likely reflects the importance of a numerical dollar value (e.g., bank balance) when cognitively appraising one's life achievements, versus the relevance of that number to our everyday experience of joy or our prevailing mood.
 
-- The relationship between subjective wellbeing and wealth is positive but happiness and satisfaction have different (positive) relationships.
-Satisfaction increases linearly with wealth while happiness increases rapidly up to a point, after which further increments in wealth produce less change - this is the **cost point** of happiness.
-So money buys happiness when economic security is scarce, but not so much once the cost point is reached.
-- The cost point after which money provides decreasing returns in happiness is increasing over time.
-In 2002, the cost point of happiness represented a 13% increase over median wealth, while in 2018 it represented a 67% increase over median wealth.
-- According to @frijters2020happy, coming up with a consensus estimate for income to translate into welbeing features high on the wellbeing research agenda.
-However evidence of a linear relationship between wellbeing and wealth calls into question the need to supplant traditional measures of economic success such as GDP with measures of wellbeing, such as "life satisfaction" surveys.
-
-<br>
-
-A study reported in 2013 challenged the idea that the positive effect of money plateaus [@stevenson2013subjective].
+Previous studies have challenged the notion that the positive effect of money plateaus [@stevenson2013subjective].
 After comparing life satisfaction levels in both rich and poor countries, and rich and poor people within a country – with “rich” being defined as an income greater than $15,000 per person – Stevenson & Wolfers concluded: “The relationship between wellbeing and income … does not diminish as income rises.
@@
 On the other hand, in a nationally representative sample of 44,000 adult Americans over 44 years in the General Social Survey (GSS) happiness continued to increase with money [@twenge2020expanding], implying no inflection point existed between money and happiness.
+Both these most recent reports provide contradictory results regarding the nature of the relationship between wealth and wellbeing, and both are based on data spanning long periods of time (12-44 years).
+Perhaps the different results are due to differences in the constructs being measured, or perhaps they are due to the different effects of time between studies.
 
-Both these reports provide contradictory results regarding the nature of the relationship between wealth and wellbeing, and both are based on data spanning long periods of time (12-44 years).
-Perhaps the different results are due to differences in the constructs being measured, or perhaps it is due to different effects of time between studies.
+We report here for the first time that the cost of happiness increased over time, faster than inflation or the median household income.
+In 2002, the cost point of happiness represented a `r round((29.41/26.86 - 1)*100)`% increase over median wealth, while in 2018 it represented a `r round((105.85/74.54 - 1)*100)`% increase over median wealth.
+This also represents a reduction from 44% to 26% in the proportion of people who have achieved a level of household wealth beyond which their happiness no longer depends.
+
+According to @frijters2020happy, coming up with a consensus estimate for income to translate into welbeing features high on the wellbeing research agenda.
+However evidence of a linear relationship between wellbeing and wealth calls into question the need to supplant traditional measures of economic success such as GDP with measures of wellbeing, such as "life satisfaction" surveys.
+
+<br>
 

### 2020-11-14 drafts/appendix.Rmd
Updated word draft

---
title: "Supplementary material: The increasing cost of happiness"
bibliography: ../src/references.bib
csl: ../src/elsevier-harvard.csl
output:  
  word_document:
    reference_docx: ../src/academic_style.docx
always_allow_html: yes
---

```{r setup, include=FALSE}
library(tidyverse)
library(flextable)
library(bayesplot)

knitr::opts_chunk$set(eval = TRUE,
                      echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      fig.path = "../figures/")

#### To do ####
# https://www.psychologytoday.com/au/blog/how-happiness/201409/does-money-really-buy-happiness
# https://www.pc.gov.au/research/completed/rising-inequality/rising-inequality-highlights.pdf
# https://www.pc.gov.au/research/completed/rising-inequality/3_HILDA_variables.R
# more people could a achieve a level of wealth at which happiness no longer increased
```
\  
\

R.W. Morris^1, 2^, N. Kettlewell^2,3,4^  & N.Glozier^1,5^

1. Central Clinical School, Faculty of Medicine and Health, University of Sydney, NSW, Australia
2. ARC Centre of Excellence for Children and Families over the Life Course, University of Queensland, QLD, Australia
3. School of Economics, University of Technology, NSW, Australia
4. Institute of Labor Economics (IZA), Bonn, Germany
5. Brain and Mind Centre, University of Sydney, NSW, Australia

\

**Corresponding author:**  
Professor Nick Glozier  
Faculty of Medicine and Health,   
University of Sydney,  
NSW 2050,  
Australia  
email: nick.glozier@sydney.edu.au

\

Draft: `r format(Sys.time(), '%d %B, %Y')`  
Word count:  3760 (max 4000)  
Tables:      1  
Figures:     5

**keywords:** Subjective wellbeing, household wealth, HILDA

\

***

## Supplementary Material

#### Modelling

We tested the relationship between wealth and each wellbeing variable (happiness and satisfaction) using a cubic **B-spline model**. A cubic B-spline was the most complex model (more degrees of freedom), and chosen because it has been recently used to describe the maximum point at which wellbeing no longer increases with wealth [i.e., "satiety", @jebb2018happiness]. Following Jebb et al, we restricted the cubic B-spline model to *n* = 4, 5 and 6 knots and show each instantiation here.

_Model estimation_

<br>

## Results

```{r funtimes}
source("../src/happyfun.r")
```

For visualization purposes only, due to the large number of individual data points in each year, we grouped individuals into 10 equal-sized subgroups according to their wealth decile ($). Thus in each plot, the mean wealth and mean wellbeing score for each decile is presented, rather than each individual data point. Note that the line-of-best-fit and 95% confidence intervals from each regression model _of all individuals_ is shown in overlay.

#### Household wealth and wellbeing (satisfaction and happiness)
```{r losat_hwnwip, fig.height=9, fig.width=11}
library(patchwork)

p1 <- happywealth_covs %>%
  plot_inflection_cubic(losat, adj_hifdip)

p2 <- happywealth_covs %>%
  plot_inflection_cubic(gh9, adj_hifdip)

(p1 / p2) +
  plot_annotation(title = "Effect of wealth on satisfaction and happiness")
```

<br>

**Key points**

- The relationship between satisfaction and household wealth appeared linear
- The linear relationship between satisfaction and wealth became slightly weaker (less steep) over years
- The cubic B-spline models indicated no satiety point (horizontal inflection) in satisfaction before the highest decile in any year
- Any change point in satisfaction was as likely to be convex as concave

<br>

## References

### 2020-11-14 drafts/wdraft0.Rmd
Updated word draft

---
title: "The increasing cost of happiness"
author: "R.W. Morris^1,2^, N. Kettlewell^2,3,4^  & N.Glozier^1,5^"
bibliography: ../src/references.bib
csl: ../src/elsevier-harvard.csl
output:  
  word_document:
    reference_docx: ../src/academic_style.docx
always_allow_html: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(flextable)
library(bayesplot)

knitr::opts_chunk$set(eval = TRUE,
                      echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      fig.path = "../figures/")

#### To do ####
# https://www.psychologytoday.com/au/blog/how-happiness/201409/does-money-really-buy-happiness
# https://www.pc.gov.au/research/completed/rising-inequality/rising-inequality-highlights.pdf
# https://www.pc.gov.au/research/completed/rising-inequality/3_HILDA_variables.R
# more people could a achieve a level of wealth at which happiness no longer increased
```

\

1. Central Clinical School, Faculty of Medicine and Health, University of Sydney, NSW, Australia
2. ARC Centre of Excellence for Children and Families over the Life Course, University of Queensland, QLD, Australia
3. School of Economics, University of Technology, NSW, Australia
4. Institute of Labor Economics (IZA), Bonn, Germany
5. Brain and Mind Centre, University of Sydney, NSW, Australia

\

**Corresponding author:**  
```
Professor Nick Glozier  
Faculty of Medicine and Health,   
University of Sydney,  
NSW 2050,  
Australia  
email: nick.glozier@sydney.edu.au
```

\

|            |                                     |
|------------|-------------------------------------|
| Draft      | `r format(Sys.time(), '%d %B, %Y')` |
| Word count | 3760 (max 4000)                     |
| Tables     | 1                                   |
| Figures    | 3                                   |

\ 
\

**keywords:** Subjective wellbeing, household wealth, HILDA

\

<br><br>

# Abstract

A fundamental question for governments and people is how much happiness does a dollar buy? The accepted view among economists and psychologists is that money and happiness increase together up to a point, after which there is little further gain from increasing wealth. While the location of this change point has been determined, and the cost of happiness reportedly ranges between USD\$60 to $95K, there has been no investigation as to whether the cost of happiness has increased or decreased over time. We tested the relationship between household wealth and both happiness and life satisfaction using household economic data from Australia between 2002-2016. We discovered the cost of happiness has increased over those 16 years faster than inflation (i.e., cost of living). This result shows that inequalities in wealth may be driving inequities in happiness and wellbeing, with implications for health and recent government policy-goals to monitor and improve wellbeing.

<br><br>

<br><br>

<br><br>

<br><br>

<br><br>

<br><br>

<br><br>

<br><br>

## Background
A fundamental question for psychology and economics is just how much wellbeing does a dollar buy? Increasing wealth is commonly associated with increasing wellbeing, however a point at which wealth no longer increases subjective wellbeing has also been widely observed [@clark2008relative; @dolan2008we; @easterlin1974growth]. Given that a central goal of nations and governments is to improve wealth under the assumption that wealth always increases wellbeing, challenges to this notion have far reaching consequences [@frijters2020happy].

Subjective wellbeing is not a unitary entity [@diener2017findings]; studies typically distinguish between life satisfaction, the cognitive appraisal of one’s own accomplishments, and happiness, one’s prevailing affective state or emotional mood. Money can have different effects on each. For instance, we have recently reported that positive life events, such as a major financial windfall, have a greater impact on an individual’s satisfaction than their happiness [@kettlewell2020differential]. While the distinct effect of wealth on satisfaction and happiness was observed within individuals, the distinct effects of wealth have also been observed across individuals. For instance, Kahneman & Deaton [-@kahneman2010high] showed that self-reported levels of happiness increased with household wealth up to a point ($75,000). But after that, increasing amounts of money had no further effect on happiness. They also reported that life satisfaction continued to increase with wealth beyond $75,000. Indeed, the difference between the two questions: “How satisfied are you with your life?” and “How happy are you these days?” has been identified as a crucial mediating factor in a meta-analysis of 111 studies on wealth and wellbeing [@howell2008relation; also @veenhoven2006rising]. Results such as these have provided a more nuanced view among psychologists and (some!) economists about the relationship between wealth and wellbeing;  namely that wealth is more strongly related to satisfaction than to happiness.

The distinction between satisfaction and happiness are increasingly relevant to governments and policy-makers due to the growing recognition that increasing wealth does not necessarily lead to improvements in wellbeing [@clark2018four; @frijters2020happy]. If wealth no longer improves wellbeing, then the maximization of wealth may no longer be a legitimate goal of government. Fundamentally, the existence of a change point between wealth and happiness reveals an unacknowledged inequality in the distribution of wellbeing in the economy. Of concern is the point at which wealth produces no further increases in happiness – that is, the change point, or cost of happiness. The cost of happiness represents the point at which material wealth stops driving inequalities in the distribution of happiness in the economy, where lower cost-points represent more equitable distributions of happiness. For instance, Kahneman estimated the cost of happiness among US survey respondents in 2008 to be approximately USD75,000 per year, substantially more than the US median income (USD52,000) in that year. Research since then has revealed variations in the cost of happiness due to world region, gender, and education [@jebb2018happiness], consistent with financial inequality driving inequities in wellbeing and happiness.  However, to date there has been no investigation of whether the cost of happiness has changed over time. In particular has the cost, and therefore the distribution, of happiness become more or less equitable in the last few decades?

We used household economic panel data from Australia (HILDA) to provide the first investigation of whether changes in wealth and wellbeing have changed the cost of happiness over the last 16 years (2002-2018). HILDA provides a representative sample of households in Australia with detailed measurements of wealth and subjective wellbeing in the same sample, which makes it an excellent data source to investigate the present question. We distinguished between satisfaction and happiness as different components of subjective wellbeing, and evaluated how each varies with household wealth. After adjusting for age, gender, and education level, we confirmed that happiness and satisfaction have distinct relationships with increasing wealth, but the cost of happiness has increased in real dollar terms since 2002.

<br>

## Methods

<br>

#### Wealth
Our indicator of wealth and economic security was household after-tax income. Household income better represents economic security, since members of the same household share expenses as well as risks; i.e., they can provide a direct and immediate support network when financial shocks occur. The other major studies also used household after-tax income as the indicator of wealth and economic security [e..g, @kahneman2010high; @jebb2018happiness], and so we follow the same standard here as well. Household financial year disposable regular income [(hifdip)](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=hifdip){target="_blank"} from the annual HILDA survey in 2002, 2006, 2010, 2014,and 2018 was selected. This variable in HILDA represents the sum across all household members of financial year gross regular income less taxes on financial year gross regular income.

Household income was provided by the University of Melbourne as imputed, and with weighted topcodes to preserve anonymity. We removed individuals with values above the threshold indicated for top-coding. Household wealth was adjusted for the number of people in each household by dividing by the square root of household size [@headey2004effects; @jebb2018happiness; @kahneman2010high]. We converted dollar values to the base year (2002) according to CPI from the RBA, in order to adjust for inflation (https://www.rba.gov.au/calculator/annualDecimal.html).

<br>

#### Subjective Wellbeing
There are a variety of variables related to subjective well-being collected annually in HILDA, but the two we used here matched the variables we used in our previous paper [@kettlewell2020differential], namely, life satisfaction as a measure of cognitive wellbeing, and the SF-36 as a measure of affective wellbeing or happiness. Life satisfaction [(losat)](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=losat){target="_blank"} was assessed by a single item question asked each survey: "How satisfied are you with your life (0 to 10)". Happiness was determined by item 9 in the SF-36 [(gh9a to gh9i)](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=gh9a){target="_blank"}. The SF-36 is a widely used self-completion measure of various aspects of physical, emotional and mental health [@ware2000sf]. Item 9 consists of nine questions relating to mental health and vitality, where five questions deal with positive and negative aspects of mental health (e.g., "Felt so down in the dumps nothing could cheer me up", "Been happy"), and four questions deal with positive and negative aspects of vitality (e.g., "feel full of life", "felt worn out"). Each question referred to the past four weeks and agreement was indicated on a six-point scale. We reverse scored the relevant responses and calculated the sum of the nine questions so that higher scores represented better wellbeing.

<br>

#### Modelling

We modelled the relationship between wealth and each wellbeing variable (happiness and satisfaction) using a simple **linear model** and a **piecewise model** (broken-stick). The linear model was the simplest relationship while the piecewise model was chosen as the simplest extension of a linear model which can identify a change point (inflection) between wellbeing and wealth. The location of the change point was a free parameter which revealed where wellbeing no longer increased linearly with wealth. We then compared the linear model against the piecewise model to determine if a change point existed in each or any year between wealth and each wellbeing variable (see model selection). Finally, where a change point existed, we determined the location of the change point for that year (see change point estimation).

_Model estimation_  
We adopted a Bayesian approach for estimating the linear and piecewise model in the software Stan [@burkner2017; @rstan2019]. In each case,

Let $y_i \sim N(\mu_i, \sigma^2_y)$

The linear model was estimated as:

$$
\mu_i = \beta_0 + \beta_1 X_i 
$$

Where $X_i$ was an individual's household wealth (\$) as well as other covariates (age, age^2^, gender, education), and $y_i$ was an individual's wellbeing.

<br>

The piecewise model was a simple extension of this to include a free parameter to represent the changepoint in wealth ($\omega$) as well as the slope after the change point ($\beta_2$):

$$
\mu_i = \beta_0 + \beta_1 (x_i - \omega_t) (x_i ≤ \omega_t) + \beta_2 (x_i - \omega_t) (x_i > \omega_t)
$$
<br>

The above models estimated population-level (fixed) effects for each year in the data (*t* = 2002, 2006, 2010, 2014 and 2018). Because we were interested in the location of the change point between wealth and wellbeing that existed across individuals within each year, we ignored the panel design of HILDA because the dependency between observations of the same person across years was orthogonal to our effects of interest. We assumed an unstructured variance-covariance matrix between years, and specified weakly informed priors for each *&beta;*, and a uniform prior over the restricted range of wealth values for *&omega;*.

_Model selection_  
To determine whether wellbeing was a linear or non-linear (e.g., piecewise) function of wealth, we compared the linear and piecewise model posterior probabilities using the [Widely Applicable Information Criterion (WAIC)](https://bookdown.org/ajkurz/Statistical_Rethinking_recoded/overfitting-regularization-and-information-criteria.html#the-problem-with-parameters){target="_blank"}. The WAIC is the log-posterior predictive density plus a penalty proportional to the variance in the posterior distribution. Thus it provides an approximation of the out-of-sample deviance that converges to the cross-validation approximation in a large sample, with a penalty for the effective number of parameters (degrees of freedom). For this reason is it useful to compare two models of varying complexity, such as our linear and piecewise model.

WAIC was defined as: WAIC = -2(lppd - *p*~WAIC~)

Where lppd (log pointwise predictive density) is the total across observations of the log of the average likelihood of each observation, and *p*~WAIC~ is the effective number of free parameters determined by the sum of the variance in log-likelihood for each observation (*i*).

_Parameter estimation_  
To determine the location of the change point (*&omega;*) between wellbeing and wealth, we modelled the relationship between wealth and wellbeing across individuals using the piecewise model described above, and sampled the posterior probability of *&omega;* over 4000 interations. The complete posterior distribution of *&omega;* for each year is presented along with the expected value (mean).

<br>

_Covariates_  
Age (and age^2^), gender, and education were included as covariates, and fulltime students were removed. Gender was included as a binary variable, and education was coded from the highest level of education achieved ("less than highschool", "highschool", "university graduate", "postgraduate").

```{r read_write_covs}
happywealth_covs <- read_rds("../results/happywealth_covs.rds")
```

<br>

## Results

```{r funtimes}
source("../src/happyfun.r")
```

For visualization purposes only, due to the large number of individual data points in each year, we grouped individuals into 10 equal-sized subgroups according to their wealth decile ($). Thus in each plot, the mean wealth and mean wellbeing score for each decile is presented, rather than each individual data point. Note that the line-of-best-fit and 95% confidence intervals from each regression model _of all individuals_ is shown in overlay.

<br>

##### Figure 1. Household wealth and satisfaction (blue) and happiness (red)
```{r losat_gh9_hifdip}
library(patchwork)

p1 <- happywealth_covs %>%
  rename(Satisfaction = losat, `Household wealth` = adj_hifdip) %>% 
  plot_linear(Satisfaction, `Household wealth`) +
  scale_fill_manual(values = "#b3cde0") +
  scale_color_manual(values = "#005b96")

p2 <- happywealth_covs %>%
  rename(Satisfaction = losat, `Household wealth` = adj_hifdip) %>% 
  plot_inflection_piecewise(Satisfaction, `Household wealth`) +
  scale_fill_manual(values = "#b3cde0") +
  scale_color_manual(values = "#005b96")

p3 <- happywealth_covs %>%
  rename(Happiness = gh9, `Household wealth` = adj_hifdip) %>% 
  plot_linear(Happiness, `Household wealth`) +
  scale_fill_manual(values = "#C79999") +
  scale_color_manual(values = "#8F2727")

p4 <- happywealth_covs %>%
  rename(Happiness = gh9, `Household wealth` = adj_hifdip) %>% 
  plot_inflection_custom(Happiness, `Household wealth`, 
                         knots = c(32, 46, 52.5, 86, 124.5)) +
  scale_fill_manual(values = "#C79999") +
  scale_color_manual(values = "#8F2727")
```
```{r figure_1, fig.height=10, fig.width=9}
p1 / p2 / p3 / p4
```

```
Figure 1 legend: Line-of-best-fit and 95% confidence interval (shaded) from regressions of  wellbeing on household wealth, with age, gender and education included as covariates. Wellbeing was measured as life satisfaction (blue) or happiness (red). The total number of individuals in each model are shown (n). Points represent the mean value of each wealth decile.
```

<br>

Figure 1 shows the relationship between household wealth and satisfaction appeared relatively linear, as the results of the piecewise regression indicated the inflection of the change point could be increasing (e.g., 2002, 2006) as well as decreasing (e.g., 2010, 2014, 2018). By contrast, a decreasing change point was evident in each year of happiness on wealth. The results of a formal comparison between linear and piecewise models is provided in Model Selection below. However note that the change point between wealth and happiness appeared to shift rightwards over time (bottom row). (We can also see the linear relationship between household wealth and each wellbeing variable became slightly weaker over time, i.e., less steep)

<br>

#### Model Selection

We compared the posterior evidence for a linear relationship between wellbeing and wealth with the posterior evidence of a nonlinear (piecewise) relationship (i.e., WAIC~linear~ — WAIC~piecewise~). Thus a WAIC difference greater than zero indicated evidence for a linear relationship. A WAIC difference less than zero indicated evidence for a nonlinear (piecewise) relationship.

<br>

##### Figure 2. Model evidence (mean difference of WAIC~linear~ — WAIC~piecewise~)

```{r plot_waic_losat}
waic_result <- read_rds("../results/waic_losat.RDS")

d1 <- data.frame(fy2018 = rnorm(n = 1e5, 
                                     mean = waic_result[2, 1] * -2, 
                                     sd   = waic_result[2, 2] *  2)/10,
                 fy2014 = rnorm(n = 1e5, 
                                     mean = waic_result[4, 1] * -2, 
                                     sd   = waic_result[4, 2] *  2)/10,
                 fy2010 = rnorm(n = 1e5, 
                                     mean = waic_result[6, 1] * -2, 
                                     sd   = waic_result[6, 2] *  2)/10,
                 fy2006 = rnorm(n = 1e5, 
                                     mean = waic_result[8, 1] * -2, 
                                     sd   = waic_result[8, 2] *  2)/10,
                 fy2002 = rnorm(n = 1e5, 
                                     mean = waic_result[10, 1] * -2, 
                                     sd   = waic_result[10, 2] *  2)/10
                )

color_scheme_set("blue")
p5 <- mcmc_intervals(x = d1,
  pars = colnames(d1),
  prob = 0.0,
  prob_outer = .90
  ) +
  scale_x_continuous(
    limits = c(-10, 40),
    labels = function(x) {x*10}) +
  geom_vline(aes(xintercept = 0), color = "grey90") +
  theme_test() +
  labs(
    subtitle = "Satisfaction",
    x = "")
```

```{r plot_waic_gh9}
waic_result <- read_rds("../results/waic_comparisons_gh9_percentiles.RDS")

d1 <- data.frame(fy2018 = rnorm(n = 1e5, 
                                mean = waic_result[10, 1] * 2, 
                                sd   = waic_result[10, 2] *  2)/10,
                 fy2014 = rnorm(n = 1e5, 
                                mean = waic_result[8, 1] * 2, 
                                sd   = waic_result[8, 2] *  2)/10,
                 fy2010 = rnorm(n = 1e5, 
                                mean = waic_result[6, 1] * 2, 
                                sd   = waic_result[6, 2] *  2)/10,
                 fy2006 = rnorm(n = 1e5, 
                                mean = waic_result[4, 1] * 2, 
                                sd   = waic_result[4, 2] *  2)/10,
                 fy2002 = rnorm(n = 1e5, 
                                mean = waic_result[2, 1] * 2, 
                                sd   = waic_result[2, 2] *  2)/10
                 )

color_scheme_set("red")
p6 <- mcmc_intervals(x = d1,
           pars = colnames(d1),
           prob = 0.00,
           prob_outer = 0.90
           ) +
  scale_x_continuous(
    limits = c(-10, 40),
    labels = function(x) {x*10}) +
  geom_vline(aes(xintercept = 0), color = "grey90") +
  theme_test() +
  labs(
    subtitle = "Happiness",
    x = "WAIC difference (90% interval)")
```

```{r figure_2, fig.width=6.5, fig.height=4, dpi=300}
p5 / p6
```

```
Figure 2 legend: Differences in posterior evidence for a linear fit over a piecewise fit (WAIC) for satisfaction (blue) and happiness (red). The filled circle indicates the mean of the distribution and the horizontal bars represents the 90% credible interval. A difference greater than zero is support for the linear model and below zero is support for the piecewise model.
```

<br><br>

Model selection revealed the posterior evidence for the linear fit of satisfaction on wealth was credibly superior to the nonlinear (piecewise) fit - Figure 2 shows the 90% credible interval of the fit for each year of satisfaction was above zero with no overlap. By contrast, a nonlinear fit of happiness on wealth was generally superior, and the nonlinear fit was credibly superior at 90% for each of three years (2006, 2014, 2018). Thus, the posterior evidence indicates happiness and satisfaction have distinct relationships with household wealth; satisfaction tends to increase linearly with wealth, while a change point exists in the relationship between happiness wealth.

<br>

#### Parameter Estimation

We present the complete posterior distribution of &omega; over 4000 samples for each year, representing the location of the change point between happiness and wealth. Shaded areas to the right of the vertical grey dotted line are credibly (90%) larger than the base year (2002).

<br>

##### Figure 3. Posterior distributions of &omega; (mean & 90% distribution)  
```{r figure_3, fig.width = 6, fig.height = 4.5, dpi=300}
hifdip_gh9_fixed_fit <- read_rds("../results/hifdip_gh9_fixed_full_covs.RDS")

as.data.frame(hifdip_gh9_fixed_fit) %>%
  select(b_alpha_Intercept, b_alpha_year2006, b_alpha_year2010, 
         b_alpha_year2014, b_alpha_year2018) %>%
  transmute(
    fy2002 = b_alpha_Intercept,
    fy2006 = mean(b_alpha_Intercept) + b_alpha_year2006,
    fy2010 = mean(b_alpha_Intercept) + b_alpha_year2010,
    fy2014 = mean(b_alpha_Intercept) + b_alpha_year2014,
    fy2018 = mean(b_alpha_Intercept) + b_alpha_year2018
  ) %>%
  mutate_all(.funs = ~brms::inv_logit_scaled(., 0, 150)*10) -> posteriors

color_scheme_set("red")

mcmc_areas_ridges(
  posteriors,
  pars = rev(colnames(posteriors)),
  prob = 0.90, # 80% intervals
  prob_outer = .99, # 99%
  point_est = "mean"
  ) +
  theme_test() +
  geom_vline(aes(xintercept = mean(posteriors$fy2002)), color = "gray", linetype = 3) +
  labs(
    x = "Household wealth ($000s)")
```

```
Figure 3 legend: Posterior distribution of the change point parameter representing the location in household wealth ($) for each year. Shaded area represents the 90% credible region.
```

<br>

Figure 3 shows the change point between happiness and household wealth (the _cost of happiness_) credibly increased between 2002 and 2018. Table 2 below compares the change in median and mean wealth for each year with the cost of happiness. It also present the percentage of people in each year whose household wealth exceeded the cost point.

<br>

```{r, table_2, ft.align="left"}
happywealth_covs %>%
  select(year, xwaveid, adj_hifdip) %>%
  mutate(
    year = paste0("fy", year),
    adj_hifdip = adj_hifdip/1000) %>%
  left_join(
    gather(
      summarise_all(posteriors, list(median)), year, change_point
    )) %>%
  group_by(year) %>%
  summarise(
    `median ($000s)` = round(median(adj_hifdip), 2),
    `mean ($000s)` = round(mean(adj_hifdip), 2),
    `change point ($000s)` = round(mean(change_point), 2),
    `%` = round(mean(adj_hifdip > mean(change_point))*100, 1)
  ) %>%
  flextable() %>%
  set_caption("Table 1. Household wealth, change points and percentile") %>%
  autofit()
```

<br>

Table 1 shows the cost of happiness increased faster than median household wealth. As a result, a smaller percentage of people achieved a level of wealth at which happiness no longer increased in 2018 than in 2002.

<br><br>

## Conclusions

We found the relationship between subjective wellbeing and wealth was positive, but happiness and satisfaction had different (positive) relationships: Satisfaction increased linearly with wealth, while happiness increased rapidly up to a point after which further increments in wealth produced less change – this confirms the distinct effects of wealth in previous research [e.g., @howell2008relation; @kahneman2010high], and contributes rare evidence from the same sample. Furthermore, we report here for the first time that the change point between wealth and happiness increased over time between 2002 and 2018, faster than inflation or the median household income.

We refer to the change point after which increases in wealth no longer produce similar increases in happiness as the cost of happiness. So household wealth, and the economic security it represents, buys happiness up to a point after which happiness and wealth are rarely related. Satisfaction on the other hand always increased with household wealth. The difference likely reflects the importance of a numerical dollar value (e.g., bank balance) when cognitively appraising one’s life achievements, versus the relevance of that number to our everyday experience of joy and prevailing mood. However an implication of the increasing cost of happiness over the last sixteen years is that wealth inequality is driving increasing _inequities_ in wellbeing. In 2002, the cost point of happiness represented a 9% increase over median wealth, while in 2018 it represented a 42% increase over median wealth. This also represents a reduction from 44% to 26% in the proportion of people who have achieved a level of household wealth beyond which their happiness no longer depends. Thus over time the happiness of more people, i.e., their everyday experience of joy and their prevailing mood, has depended on their material wealth.

Australia has low levels of wealth disparity relative to other OECD countries, and the Gini coefficient has not changed a great deal between 2002 and 2018 in the HILDA dataset [@australian2018rising]. A stable Gini coefficient shows income inequality has remained steady over the time period, and our results do not conflict with this conclusion. Rather what we are revealing is the effect of income inequality on happiness has increased over the same time period. So while income inequality has remained steady, it’s impact on wellbeing inequity has increased. We think this highlights the issue that while traditional measures of wealth and income inequality may not change, their impact on wellbeing and health can still vary. As focus shifts from traditional wealth indicators towards wellbeing measures, findings such as this may become more prevalent.

Some recent studies have challenged either the notion that the positive effect of money plateaus, or that the effect on happiness and satisfaction are distinct. In a nationally representative sample of 44,000 adult Americans in the General Social Survey (GSS), happiness continued to increase with household wealth, implying no change point existed between money and happiness [@twenge2020expanding]. The GSS asks a single item on happiness: “Taken all together, how would you say things are these days – would you say that you are very happy, pretty happy, or not too happy?” The form of this question is quite similar to the single item “life satisfaction” question in HILDA, as both request the respondent to cognitively evaluate their circumstances. By contrast, the nine items we selected to measure happiness covered a range of affect and focused on current feelings: e.g., “How much of the time during the past 4 weeks have you been a happy person?” Critical differences in the operational definition of happiness seem likely to explain the results observed here.

A second recent report from a survey of 1.7 million people representing 164 countries in the Gallup World Poll reported increases in household wealth were associated with change points in happiness and satisfaction [@jebb2018happiness], rather than distinct effects as we found. It may be that differences in the sparsity of the high income data in HILDA (for which extremely high income households are masked) explain the differences here. However including these individuals in our analysis did not change the linear effect between wealth and satisfaction observed here, and the weighted topcodes should have biased the result towards a change point. We also note a linear effect between wealth and satisfaction is consistent with the majority of earlier literature [@howell2008relation; @stevenson2013subjective].

As governments and policy-makers begin to focus on wellbeing, it will be critical to understand how traditional economic indicators such as household wealth, income inequality, and consumption interact with wellbeing and health. According to @frijters2020happy, coming up with a consensus to translate income into wellbeing features high on the wider wellbeing research agenda. Establishing the links between income, household wealth, wellbeing and health, and how inequalities in one drives inequities in the other, will be a critical step in that agenda.

<br>

## References

### 2020-11-14 src/happyfun.R

@@ -15,15 +15,19 @@ plot_linear <- function(.data, swb_col, dollars_col) {
   
   df %>%
     group_by(year) %>%
+    summarise(n = n()) %>%
+    mutate(note = paste("n =", format(n, big.mark = ","))) -> df.sum
+  
+  df %>%
+    group_by(year) %>%
     mutate(
       dollar_bin = ntile(dollars, 10)
     ) %>%
     group_by(year, dollar_bin) %>%
     summarize(
       wellbeing = mean(swb, na.rm=T),
-      wealth = mean(dollars, na.rm=T)/1000,
-      n = n()
-    ) -> df.summary
+      wealth = mean(dollars, na.rm=T)/1000
+      ) -> df.deciles
   
   df %>%
     mutate(
@@ -31,14 +35,14 @@ plot_linear <- function(.data, swb_col, dollars_col) {
       wellbeing = swb
     ) -> df.plot
   
-  # general additive model
+  # Linear model
   ggplot(df.plot, aes(x = wealth, y = wellbeing)) +
-    geom_smooth(
+    geom_smooth(aes(fill = "manual1", color = "manual2"),
       method = "lm", 
-      size = 0.5,
-      se = TRUE) +
-    coord_cartesian(xlim = c(0, max(df.summary$wealth)),
-                    ylim = c(min(df.summary$wellbeing), max(df.summary$wellbeing))) +
+      size = 0.5
+      ) +
+    coord_cartesian(xlim = c(0, max(df.deciles$wealth)),
+                    ylim = c(min(df.deciles$wellbeing), max(df.deciles$wellbeing))) +
     facet_wrap(~year, nrow = 1) +
     labs(subtitle = paste(rlang::as_name(swb_colname), "(linear)"),
          x = paste(rlang::as_name(dollars_colname), "($000s)"),
@@ -47,11 +51,151 @@ plot_linear <- function(.data, swb_col, dollars_col) {
     theme(
       legend.background = element_blank(),
       legend.title = element_blank(),
-      legend.position = c(0.075, 0.925)) -> p1
+      legend.position = "none") -> p1
   
   p1 +
-    geom_point(data = df.summary) 
+    geom_point(data = df.deciles) + # size = 3, alpha = 0.25
+    geom_text(data = df.sum, aes(x = Inf, y = -Inf, label = note),
+              hjust = 1.1, vjust = -.75, colour = 1, size = 3)
   
+}
+
+
+plot_inflection_piecewise <- function(.data, swb_col, dollars_col, knots) {
[…]

### 2020-11-14 drafts/draft0.Rmd
Update to wdraft0 and README

@@
 title: "The increasing cost of happiness"
-author: "Dr Richard Morris"
-date: "Updated: `r Sys.Date()`"
+author: "R.W. Morris^1,2^, N. Kettlewell^2,3,4^  & N.Glozier^1,5^"
 bibliography: ../src/references.bib
 csl: ../src/elsevier-harvard.csl
-output: github_document
+output:  
+  html_document:
+    toc: true
+    toc_depth: 4
+    toc_float:
+      collapsed: false
 always_allow_html: yes
+editor_options: 
+  chunk_output_type: console
 ---
@@
 library(tidyverse)
+library(flextable)
 library(bayesplot)
@@
 # https://www.psychologytoday.com/au/blog/how-happiness/201409/does-money-really-buy-happiness
-# 
+# https://www.pc.gov.au/research/completed/rising-inequality/rising-inequality-highlights.pdf
+# https://www.pc.gov.au/research/completed/rising-inequality/3_HILDA_variables.R
 # more people could a achieve a level of wealth at which happiness no longer increased
 ```
+
+\
+
+1. Central Clinical School, Faculty of Medicine and Health, University of Sydney, NSW, Australia
+2. ARC Centre of Excellence for Children and Families over the Life Course, University of Queensland, QLD, Australia
+3. School of Economics, University of Technology, NSW, Australia
+4. Institute of Labor Economics (IZA), Bonn, Germany
+5. Brain and Mind Centre, University of Sydney, NSW, Australia
+
+\
+
+**Corresponding author:**  
+```
+Professor Nick Glozier  
+Faculty of Medicine and Health,   
+University of Sydney,  
+NSW 2050,  
+Australia  
+email: nick.glozier@sydney.edu.au
+```
+
+\
+
+|            |                                     |
+|------------|-------------------------------------|
+| Draft      | `r format(Sys.time(), '%d %B, %Y')` |
+| Word count | 3760 (max 4000)                     |
+| Tables     | 1                                   |
+| Figures    | 3                                   |
+
+\ 
+\
+
+**keywords:** Subjective wellbeing, household wealth, HILDA
+
+\
+
+<br><br>
+
+# Abstract
+
+A fundamental question for governments and people is how much happiness does a dollar buy?
+The accepted view among economists and psychologists is that money and happiness increase together up to a point, after which there is little further gain from increasing wealth.
+While the location of this change point has been determined, and the cost of happiness reportedly ranges between USD\$60 to $95K, there has been no investigation as to whether the cost of happiness has increased or decreased over time.
+We tested the relationship between household wealth and both happiness and life satisfaction using household economic data from Australia between 2002-2016.
+We discovered the cost of happiness has increased over those 16 years faster than inflation (i.e., cost of living).
+This result shows that inequalities in wealth may be driving inequities in happiness and wellbeing, with implications for health and recent government policy-goals to monitor and improve wellbeing.
+
+<br><br>
+
+<br><br>
+
+<br><br>
+
+<br><br>
+
+<br><br>
+
+<br><br>
+
 ***
 
-## Abstract
-A fundamental question for governments and people is how much happiness does a dollar buy?
-The accepted view among economists and psychologists is that money and happiness increase together up to a point, after which there is little further gain from increasing wealth.
-While the location of this _change point_ has been determined, and the cost of happiness reportedly ranges between USD\$60-95K, there has been no investigation as to whether the cost of happiness has increased or decreased over time.
-We tested the relationship between money and both happiness and life satisfaction using household economic data from Australia between 2002-2016.
-We discovered no change point with wealth existed for satisfaction, but the cost of happiness has increased over those 16 years faster than inflation (i.e., cost of living).
-Results such as these suggest we need to understand why the cost of happiness has increased, and to consider whether policy-goals to improve wellbeing by increasing wealth are feasible.
-
-<br>
-
 ## Background
-A fundamental question for the psychology and economics of wellbeing is just how much happiness does a dollar buy?
-Increasing wealth is commonly associated with increasing happiness, however a point at which wealth no longer increases happiness has also been [widely observed](https://www-sciencedirect-com.ezproxy2.library.usyd.edu.au/science/article/pii/S0167487007000694){target="_blank"} [@clark2008relative; @dolan2008we; @easterlin1974growth].
+A fundamental question for psychology and economics is just how much wellbeing does a dollar buy?
+Increasing wealth is commonly associated with increasing wellbeing, however a point at which wealth no longer increases subjective wellbeing has also been widely observed [@clark2008relative; @dolan2008we; @easterlin1974growth].
 Given that a central goal of nations and governments is to improve wealth under the assumption that wealth always increases wellbeing, challenges to this notion have far reaching consequences [@frijters2020happy].
 
-For instance, one [survey of 1,000 Americans](https://www.pnas.org/content/107/38/16489.full){target="_blank"}, conducted in 2010, concluded that money does make us happier – but only up to a certain point.
-The findings showed that self-reported levels of happiness increased with household wealth up to \$75,000 a year.
-But after that, increasing amounts of money had no further effect on happiness [@kahneman2010high].
-Results such as these have reinforced the view among psychologists and economists that the relationship between wealth and measures of subjective wellbeing, such as happiness, increases linearly up to a point (i.e., a _change point_) after which there is little gained by further increments in wealth [@clark2018four; @dolan2008we].
-We suggest that the point at which wealth produces no further increases in happiness might be described as the _cost_ of happiness.
+Subjective wellbeing is not a unitary entity [@diener2017findings]; studies typically distinguish between life satisfaction, the cognitive appraisal of one’s own accomplishments, and happiness, one’s prevailing affective state or emotional mood.
+Money can have different effects on each.
+For instance, we have recently reported that positive life events, such as a major financial windfall, have a greater impact on an individual’s satisfaction than their happiness [@kettlewell2020differential].
+While the distinct effect of wealth on satisfaction and happiness was observed within individuals, the distinct effects of wealth have also been observed across individuals.
+For instance, Kahneman & Deaton [-@kahneman2010high] showed that self-reported levels of happiness increased with household wealth up to a point ($75,000).
+But after that, increasing amounts of money had no further effect on happiness.
+They also reported that life satisfaction continued to increase with wealth beyond $75,000.
+Indeed, the difference between the two questions: “How satisfied are you with your life?” and “How happy are you these days?” has been identified as a crucial mediating factor in a meta-analysis of 111 studies on wealth and wellbeing [@howell2008relation; also @veenhoven2006rising].
+Results such as these have provided a more nuanced view among psychologists and (some!) economists about the relationship between wealth and wellbeing;  namely that wealth is more strongly related to satisfaction than to happiness.
 
-Subjective wellbeing is not a unitary entity [@diener2017findings]; studies typically distinguish between _life satisfaction_, the cognitive appraisal of one's own accomplishments, and _happiness_, one's prevailing affective state or emotional mood.
-Money can have different effects on each.
-For instance, we have recently reported that positive life events, such as a major financial windfall, can have a greater impact on an individual's satisfaction than their happiness [@kettlewell2020differential].
-However these effects were observed as a change within individuals, and the change was temporary as both satisfaction and happiness returned to baseline after two years [see also @frijters2011life].
-Wealth also has a distinct relationship with happiness and satisfaction across individuals was well.
-As described above, Kahneman et al (2010) observed the association between household wealth and happiness plateaud after $75,000, but they also reported that life satisfaction continued to increase with wealth.
-Indeed, several researchers have observed a similar difference between two questions that are often used in surveys of subjective well-being: “How satisfied are you with your life?” and “How happy are you these days?” [@veenhoven2006rising; @howell2008relation].
-The common conclusion is that wealth is more strongly related to satisfaction than to happiness.
+The distinction between satisfaction and happiness are increasingly relevant to governments and policy-makers due to the growing recognition that increasing wealth does not necessarily lead to improvements in wellbeing [@clark2018four; @frijters2020happy].
+If wealth no longer improves wellbeing, then the maximization of wealth may no longer be a legitimate goal of government.
+Fundamentally, the existence of a change point between wealth and happiness reveals an unacknowledged inequality in the distribution of wellbeing in the economy.
+Of concern is the point at which wealth produces no further increases in happiness – that is, the change point, or cost of happiness.
+The cost of happiness represents the point at which material wealth stops driving inequalities in the distribution of happiness in the economy, where lower cost-points represent more equitable distributions of happiness.
+For instance, Kahneman estimated the cost of happiness among US survey respondents in 2008 to be approximately USD75,000 per year, substantially more than the US median income (USD52,000) in that year.
+Research since then has revealed variations in the cost of happiness due to world region, gender, and education [@jebb2018happiness], consistent with financial inequality driving inequities in wellbeing and happiness.
+However, to date there has been no investigation of whether the cost of happiness has changed over time.
+In particular has the cost, and therefore the distribution, of happiness become more or less equitable in the last few decades?
 
-Research since 2010 has revealed variations in the change point or cost of happiness due to [world region, gender, and education](https://www-nature-com.ezproxy2.library.usyd.edu.au/articles/s41562-017-0277-0){target="_blank"} [@jebb2018happiness].
-However to date there has been no investigation as to whether the _change point_ varies over time.
-That is, has the point at which happiness no longer increases with wealth itself changed over time?
-The value of money of course generally _decreases_ over time with inflation and the cost of living and so this will _raise_ the cost of happiness, however it will not change the cost of happiness over time in real dollar terms (i.e., adjusted for inflation).
-The distribution of wealth has also changed over the last 30 years in many OECD countries including Australia as the rich got richer while the poor stayed poor (ref needed).
-Such changes in the distribution of wealth may increase the cost of happiness if it is derived by relative wealth (e.g., Clarke et al 2008).
-When the ceiling of wealth in real dollar terms is raised over time; for any fixed income level at *t*<sub>0</sub> (i.e., the base year) there will be a greater disparity relative to the highest wealth percentile at *t*<sub>n+1</sub>.
-If happiness is, even partly, determined by this relative disparity then increasing the disparity will reduce happiness amongst the poorest percentiles (while leaving happiness amongst the richest percentiles unchanged).
-This may increase the level of wealth at which happiness no longer increases.
-That is, as the rich get richer, the poor get unhappier due to the increasing relative wealth disparity and so the cost of happiness increases.
-<span style="color: blue"> (This section may need more clarity) </span>
-
-The relationship between wealth and happiness and satisfaction are increasingly relevant to governments and policy-makers due to the growing recognition that increasing wealth does not necessarily lead to improvements in wellbeing [@clark2018four; @frijters2020happy].
-If satisfaction but not happiness increases with wealth, then the pursuit of both may not be a feasible policy.
-Likewise, if happiness increases up to a point with wealth, or the cost of happiness changes over time, then not everyone will agree that enhancing the happiness of those who are already very happy is a legitimate policy objective.
-Conversely, it may emphasize that the relationship between wealth and happiness is malleable and thus amenable to policy objectives.
-<span style="color: blue"> (This section needs a clearer statement on the importance of determining the increasing cost of happiness) </span>
-
-We used household economic panel data from Australia (HILDA) to provide the first investigation of whether changes in wealth and wellbeing have changed over the last 16 years (2002-2018).
+We used household economic panel data from Australia (HILDA) to provide the first investigation of whether changes in wealth and wellbeing have changed the cost of happiness over the last 16 years (2002-2018).
+HILDA provides a representative sample of households in Australia with detailed measurements of wealth and subjective wellbeing in the same sample, which makes it an excellent data source to investigate the present question.
 We distinguished between satisfaction and happiness as different components of subjective wellbeing, and evaluated how each varies with household wealth.
-After adjusting for age, gender, and education level, we find that satisfaction increases linearly with household wealth, but the same purchasing power has diminishing returns on happiness.
-We also discover that the location of the change point (the _cost_ of happiness) has increased in real dollar terms since 2002.
+After adjusting for age, gender, and education level, we confirmed that happiness and satisfaction have distinct relationships with increasing wealth, but the cost of happiness has increased in real dollar terms since 2002.
 
@@
 ## Methods
+
+<br>
 
@@
 #### Subjective Wellbeing
-There are a variety of variables related to subjective well-being collected annually in HILDA, but the two we used here matched the variables we used in our previous paper [@kettlewell2020differential].
-Namely, 1) life satisfaction [(losat)](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=losat){target="_blank"}, a single item question asked each survey "How satisfied are you with your life (0 to 10)", and 2) happiness determined by item 9 in the SF-36 [(gh9a to gh9i)](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=gh9a){target="_blank"}.
+There are a variety of variables related to subjective well-being collected annually in HILDA, but the two we used here matched the variables we used in our previous paper [@kettlewell2020differential], namely, life satisfaction as a measure of cognitive wellbeing, and the SF-36 as a measure of affective wellbeing or happiness.
+Life satisfaction [(losat)](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=losat){target="_blank"} was assessed by a single item question asked each survey: "How satisfied are you with your life (0 to 10)".
+Happiness was determined by item 9 in the SF-36 [(gh9a to gh9i)](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=gh9a){target="_blank"}.
 The SF-36 is a widely used self-completion measure of various aspects of physical, emotional and mental health [@ware2000sf].
 Item 9 consists of nine questions relating to mental health and vitality, where five questions deal with positive and negative aspects of mental health (e.g., "Felt so down in the dumps nothing could cheer me up", "Been happy"), and four questions deal with positive and negative aspects of vitality (e.g., "feel full of life", "felt worn out").
-We reverse scored the relevant responses and calculated a summed score so that better wellbeing was represented by higher scores.
+Each question referred to the past four weeks and agreement was indicated on a six-point scale.
+We reverse scored the relevant responses and calculated the sum of the nine questions so that higher scores represented better wellbeing.
 
@@
 
-We initially describe the relationship between wealth and each wellbeing variable (happiness and satisfaction) using three different models: 1) a simple **linear model**; 2) a cubic **B-spline model**; and 3) a **piecewise model** (broken-stick).
-These three models were chosen to represent varying degrees of complexity in the relationship between wealth and wellbeing.
-The linear model was the simplest relationship while the cubic B-spline was the most complex model (more degrees of freedom), and chosen because it has been recently used to describe the maximum point at which wellbeing no longer increases with wealth [i.e., "satiety", @jebb2018happiness].
-Following Jebb et al, we restricted the cubic B-spline model to *n* = 4, 5 and 6 knots and show each instantiation here.
-The piecewise model was chosen as the simplest extension of a linear model which can identify a change point (inflection) between wellbeing and wealth.
+We modelled the relationship between wealth and each wellbeing variable (happiness and satisfaction) using a simple **linear model** and a **piecewise model** (broken-stick).
+The linear model was the simplest relationship while the piecewise model was chosen as the simplest extension of a linear model which can identify a change point (inflection) between wellbeing and wealth.
 The location of the change point was a free parameter which revealed where wellbeing no longer increased linearly with wealth.
-
-We then compared the linear model against the piecewise model to determine if a change point existed in each or any year between wealth and each wellbeing variable (model selection).
-Finally, where a change point existed, we determined the location of the change point for that year (change point estimation).
+We then compared the linear model against the piecewise model to determine if a change point existed in each or any year between wealth and each wellbeing variable (see model selection).
+Finally, where a change point existed, we determined the location of the change point for that year (see change point estimation).
 
@@
 
-Let y<sub>*i*</sub> ~ *N*(*&mu;*<sub>*i*</sub>, *&sigma;*<sup>2</sup><sub>*y*</sub>)
+Let $y_i \sim N(\mu_i, \sigma^2_y)$
 
@@
 
-*&mu;*<sub>i</sub> = *&beta;*<sub>0</sub> + *&beta;*<sub>1</sub>*X*<sub>*i*</sub>   \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \  (1)
+$$
+\mu_i = \beta_0 + \beta_1 X_i 
+$$
 
-Where *X*<sub>i</sub> was an individual's household wealth (\$) as well as other covariates (age, age^2^, gender, education), and *y*<sub>i</sub> was an individual's wellbeing.
+Where $X_i$ was an individual's household wealth (\$) as well as other covariates (age, age^2^, gender, education), and $y_i$ was an individual's wellbeing.
 
@@
 
-The piecewise model was a simple extension of this to include a free parameter to represent the changepoint in wealth (&omega;) as well as the slope after the change point (*&beta;*<sub>2</sub>):
+The piecewise model was a simple extension of this to include a free parameter to represent the changepoint in wealth ($\omega$) as well as the slope after the change point ($\beta_2$):
 
-*&mu;*<sub>i</sub> = *&beta;*<sub>0</sub> + *&beta;*<sub>1</sub>(*x*<sub>*i*</sub> - &omega;<sub>*t*</sub>)(*x*<sub>*i*</sub> ≤ &omega;<sub>*t*</sub>) + *&beta;*<sub>2</sub>(*x*<sub>*i*</sub> - &omega;<sub>*t*</sub>)(*x*<sub>*i*</sub> > &omega;<sub>*t*</sub>)  \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \  (2)
-
+$$
+\mu_i = \beta_0 + \beta_1 (x_i - \omega_t) (x_i ≤ \omega_t) + \beta_2 (x_i - \omega_t) (x_i > \omega_t)
+$$
 <br>
@@
 
-WAIC was defined as:
-
-WAIC = -2(lppd - *p*~WAIC~)
+WAIC was defined as: WAIC = -2(lppd - *p*~WAIC~)
 
@@
 
-_Change point estimation_  
-To determine the location of the change point between wellbeing and wealth, we modelled the relationship between wealth and wellbeing across individuals using the piecewise model described above, and estimated the posterior probability of the changepoint location.
-The posterior distribution of the change point location for each year is presented.
+_Parameter estimation_  
+To determine the location of the change point (*&omega;*) between wellbeing and wealth, we modelled the relationship between wealth and wellbeing across individuals using the piecewise model described above, and sampled the posterior probability of *&omega;* over 4000 interations.
+The complete posterior distribution of *&omega;* for each year is presented along with the expected value (mean).
 
@@
 
-#### Covariates
-
-Age (and age^2^), gender, and education were included as covariates, and students were removed.
+_Covariates_  
+Age (and age^2^), gender, and education were included as covariates, and fulltime students were removed.
+Gender was included as a binary variable, and education was coded from the highest level of education achieved ("less than highschool", "highschool", "university graduate", "postgraduate").
 
@@
 
-#### Household wealth and satisfaction (cognitive wellbeing)
-```{r losat_hwnwip, fig.height=8, fig.width=9}
+<br>
+
+##### Figure 1. Household wealth and satisfaction (blue) and happiness (red)
+```{r losat_gh9_hifdip}
 library(patchwork)
@@
 p1 <- happywealth_covs %>%
-  plot_linear(losat, adj_hifdip)
+  rename(Satisfaction = losat, `Household wealth` = adj_hifdip) %>% 
+  plot_linear(Satisfaction, `Household wealth`) +
+  scale_fill_manual(values = "#b3cde0") +
+  scale_color_manual(values = "#005b96")
 
 p2 <- happywealth_covs %>%
-  plot_inflection_cubic(losat, adj_hifdip)
+  rename(Satisfaction = losat, `Household wealth` = adj_hifdip) %>% 
+  plot_inflection_piecewise(Satisfaction, `Household wealth`) +
+  scale_fill_manual(values = "#b3cde0") +
+  scale_color_manual(values = "#005b96")
 
 p3 <- happywealth_covs %>%
-  plot_inflection_piecewise(losat, adj_hifdip)
+  rename(Happiness = gh9, `Household wealth` = adj_hifdip) %>% 
+  plot_linear(Happiness, `Household wealth`) +
+  scale_fill_manual(values = "#C79999") +
+  scale_color_manual(values = "#8F2727")
 
-(p1 / p2 / p3) +
-  plot_annotation(title = "Effect of wealth on satisfaction")
+p4 <- happywealth_covs %>%
+  rename(Happiness = gh9, `Household wealth` = adj_hifdip) %>% 
+  plot_inflection_custom(Happiness, `Household wealth`, 
+                         knots = c(32, 46, 52.5, 86, 124.5)) +
+  scale_fill_manual(values = "#C79999") +
+  scale_color_manual(values = "#8F2727")
+```
+```{r figure_1, fig.height=10, fig.width=9}
+p1 / p2 / p3 / p4
+```
+
+```
+Figure 1 legend: Line-of-best-fit and 95% confidence interval (shaded) from regressions of  wellbeing on household wealth, with age, gender and education included as covariates.
+Wellbeing was measured as life satisfaction (blue) or happiness (red).
+The total number of individuals in each model are shown (n).
+Points represent the mean value of each wealth decile.
 ```
@@
 
-**Key points**
-
-- The relationship between satisfaction and household wealth appeared linear
-- The linear relationship between satisfaction and wealth became slightly weaker (less steep) over years
-- The cubic B-spline models indicated no satiety point (horizontal inflection) in satisfaction before the highest decile in any year
-- Any change point in satisfaction was as likely to be convex as concave
+Figure 1 shows the relationship between household wealth and satisfaction appeared relatively linear, as the results of the piecewise regression indicated the inflection of the change point could be increasing (e.g., 2002, 2006) as well as decreasing (e.g., 2010, 2014, 2018).
+By contrast, a decreasing change point was evident in each year of happiness on wealth.
+The results of a formal comparison between linear and piecewise models is provided in Model Selection below.
+However note that the change point between wealth and happiness appeared to shift rightwards over time (bottom row).
+(We can also see the linear relationship between household wealth and each wellbeing variable became slightly weaker over time, i.e., less steep)
 
@@
 
-#### Household wealth and happiness (affective wellbeing)
-```{r gh9_hwnwip, fig.height=8, fig.width=9}
-library(patchwork)
+#### Model Selection
 
-p1 <- happywealth_covs %>%
-  plot_linear(gh9, adj_hifdip)
-
-p2 <- happywealth_covs %>%
-  plot_inflection_cubic(gh9, adj_hifdip)
-
-p3 <- happywealth_covs %>%
-  plot_inflection_custom(gh9, adj_hifdip, knots = c(32, 46, 52.5, 86, 124.5))
-
-(p1 / p2 / p3) +
-  plot_annotation(title = "Effect of wealth on happiness")
-```
+We compared the posterior evidence for a linear relationship between wellbeing and wealth with the posterior evidence of a nonlinear (piecewise) relationship (i.e., WAIC~linear~ — WAIC~piecewise~).
+Thus a WAIC difference greater than zero indicated evidence for a linear relationship.
+A WAIC difference less than zero indicated evidence for a nonlinear (piecewise) relationship.
 
@@
 
-**Key points**
-
-- A change point is evident in each year where the linear relationship between happiness and wealth became abruptly weaker (less steep)
-- The overall linear relationship between happiness and wealth, as well as the pre-inflection slope, became weak (less steep) over years
-- The cubic B-spline model indicated there was no happiness satiety point (horizontal inflection) before the final decile in any year
-- The change point between happiness and wealth appeared to shift right with time.
-
-<br>
-
-#### Model selection
-
-We compared the posterior evidence for a linear relationship between wellbeing and wealth with the posterior evidence of a nonlinear (piecewise) relationship (WAIC~linear~ — WAIC~piecewise~).
-Thus a WAIC difference greater than zero indicates evidence for a linear relationship.
-A WAIC difference less than zero indicates evidence for a nonlinear (piecewise) relationship.
-
-##### Model evidence for linear satisfaction
+##### Figure 2. Model evidence (mean difference of WAIC~linear~ — WAIC~piecewise~)
 
@@
 color_scheme_set("blue")
-mcmc_areas(x = d1,
+p5 <- mcmc_intervals(x = d1,
   pars = colnames(d1),
-  prob = 0.90,
-  prob_outer = 1
+  prob = 0.0,
+  prob_outer = .90
   ) +
   scale_x_continuous(
+    limits = c(-10, 40),
     labels = function(x) {x*10}) +
@@
   labs(
-    subtitle = "Nonlinear over linear fit for satisfaction (mean difference and 90% distribution)",
-    x = "WAIC difference (90% shaded)")
+    subtitle = "Satisfaction",
+    x = "")
 ```
 
-<br>
-
-- A linear fit between household wealth and satisfaction was credibly superior to a nonlinear fit in each year  
-<br>
-
-##### Model evidence for nonlinear happiness
 ```{r plot_waic_gh9}
@@
 color_scheme_set("red")
-mcmc_areas(x = d1,
+p6 <- mcmc_intervals(x = d1,
            pars = colnames(d1),
-           prob = 0.90,
-           prob_outer = 1
+           prob = 0.00,
+           prob_outer = 0.90
            ) +
   scale_x_continuous(
+    limits = c(-10, 40),
     labels = function(x) {x*10}) +
@@
   labs(
-    subtitle = "Nonlinear over linear fit for happiness (mean difference and 90% distribution)",
-    x = "WAIC difference (90% shaded)")
+    subtitle = "Happiness",
+    x = "WAIC difference (90% interval)")
 ```
+
+```{r figure_2, fig.width=6.5, fig.height=4, dpi=300}
+p5 / p6
+```
+
+```
+Figure 2 legend: Differences in posterior evidence for a linear fit over a piecewise fit (WAIC) for satisfaction (blue) and happiness (red).
+The filled circle indicates the mean of the distribution and the horizontal bars represents the 90% credible interval.
+A difference greater than zero is support for the linear model and below zero is support for the piecewise model.
+```
+
+<br><br>
+
+Model selection revealed the posterior evidence for the linear fit of satisfaction on wealth was credibly superior to the nonlinear (piecewise) fit - Figure 2 shows the 90% credible interval of the fit for each year of satisfaction was above zero with no overlap.
+By contrast, a nonlinear fit of happiness on wealth was generally superior, and the nonlinear fit was credibly superior at 90% for each of three years (2006, 2014, 2018).
+Thus, the posterior evidence indicates happiness and satisfaction have distinct relationships with household wealth; satisfaction tends to increase linearly with wealth, while a change point exists in the relationship between happiness wealth.
 
@@
 
-- A nonlinear fit between household wealth and happiness was superior to a linear fit in each year 
-- The nonlinear fit was credibly superior to a linear fit at 90% for each of three years (2018, 2014, 2006)
+#### Parameter Estimation
+
+We present the complete posterior distribution of &omega; over 4000 samples for each year, representing the location of the change point between happiness and wealth.
+Shaded areas to the right of the vertical grey dotted line are credibly (90%) larger than the base year (2002).
 
@@
 
-**Key points**
-
-- The posterior evidence indicates a linear model with no inflection point explains the relationship between satisfaction and household wealth in each year
-- However a nonlinear fit with at least one inflection point better explained the relationship between happiness and household wealth
-
-<br>
-
-##### Bayesian change point results
-
-Shaded areas to the right of the vertical dotted line are credibly (95%) larger than the base year (2002).
-
-```{r happy_change_time, fig.width = 4.9, fig.height = 9}
+##### Figure 3. Posterior distributions of &omega; (mean & 90% distribution)  
+```{r figure_3, fig.width = 6, fig.height = 4.5, dpi=300}
 hifdip_gh9_fixed_fit <- read_rds("../results/hifdip_gh9_fixed_full_covs.RDS")
@@
 color_scheme_set("red")
-mcmc_areas(
+
+mcmc_areas_ridges(
   posteriors,
   pars = rev(colnames(posteriors)),
-  prob = 0.95, # 80% intervals
+  prob = 0.90, # 80% intervals
   prob_outer = .99, # 99%
@@
   labs(
-    subtitle = "Posterior change points (mean & 95% distribution)",
     x = "Household wealth ($000s)")
+```
+
+```
+Figure 3 legend: Posterior distribution of the change point parameter representing the location in household wealth ($) for each year. Shaded area represents the 90% credible region.
 ```
@@
 
-**Key points**
-
-- The level of wealth at which happiness no longer increased with wealth (the _cost of happiness_) credibly increased between 2002 and 2018.
+Figure 3 shows the change point between happiness and household wealth (the _cost of happiness_) credibly increased between 2002 and 2018.
+Table 2 below compares the change in median and mean wealth for each year with the cost of happiness.
+It also present the percentage of people in each year whose household wealth exceeded the cost point.
 
@@
 
-#### Change points over time
-
-The cost of happiness increased faster than median household wealth. As a result, a smaller percentage of people achieved a level of wealth at which happiness no longer increased in 2018 than in 2002:
-
-```{r, table1, ft.align="left"}
-library(knitr)
-
+```{r, table_2, ft.align="left"}
 happywealth_covs %>%
@@
     `median ($000s)` = round(median(adj_hifdip), 2),
+    `mean ($000s)` = round(mean(adj_hifdip), 2),
     `change point ($000s)` = round(mean(change_point), 2),
-    `%` = round(mean(adj_hifdip > mean(change_point)), 2)
+    `%` = round(mean(adj_hifdip > mean(change_point))*100, 1)
   ) %>%
-  kable(format = "html", 
-        caption = "Household wealth median, change point and percentile")
+  flextable() %>%
+  set_caption("Table 1. Household wealth, change points and percentile") %>%
+  autofit()
 ```
+
+<br>
+
+Table 1 shows the cost of happiness increased faster than median household wealth.
+As a result, a smaller percentage of people achieved a level of wealth at which happiness no longer increased in 2018 than in 2002.
 
@@
 
-We observed the relationship between subjective wellbeing and wealth was positive, but happiness and satisfaction have different (positive) relationships.
-Satisfaction increased linearly with wealth while happiness increased rapidly up to a point, after which further increments in wealth produced less change - we referred to this as the **cost** of happiness.
+We found the relationship between subjective wellbeing and wealth was positive, but happiness and satisfaction had different (positive) relationships: Satisfaction increased linearly with wealth, while happiness increased rapidly up to a point after which further increments in wealth produced less change – this confirms the distinct effects of wealth in previous research [e.g., @howell2008relation; @kahneman2010high], and contributes rare evidence from the same sample.
+Furthermore, we report here for the first time that the change point between wealth and happiness increased over time between 2002 and 2018, faster than inflation or the median household income.
+
+We refer to the change point after which increases in wealth no longer produce similar increases in happiness as the cost of happiness.
 So household wealth, and the economic security it represents, buys happiness up to a point after which happiness and wealth are rarely related.
 Satisfaction on the other hand always increased with household wealth.
-The difference likely reflects the importance of a numerical dollar value (e.g., bank balance) when cognitively appraising one's life achievements, versus the relevance of that number to our everyday experience of joy or our prevailing mood.
+The difference likely reflects the importance of a numerical dollar value (e.g., bank balance) when cognitively appraising one’s life achievements, versus the relevance of that number to our everyday experience of joy and prevailing mood.
+However an implication of the increasing cost of happiness over the last sixteen years is that wealth inequality is driving increasing _inequities_ in wellbeing.
+In 2002, the cost point of happiness represented a 9% increase over median wealth, while in 2018 it represented a 42% increase over median wealth.
+This also represents a reduction from 44% to 26% in the proportion of people who have achieved a level of household wealth beyond which their happiness no longer depends.
+Thus over time the happiness of more people, i.e., their everyday experience of joy and their prevailing mood, has depended on their material wealth.
 
-Previous studies have challenged the notion that the positive effect of money plateaus [@stevenson2013subjective].
-After comparing life satisfaction levels in both rich and poor countries, and rich and poor people within a country – with “rich” being defined as an income greater than $15,000 per person – Stevenson & Wolfers concluded: “The relationship between wellbeing and income … does not diminish as income rises.
-If there is a satiation point, we are yet to reach it.”
+Australia has low levels of wealth disparity relative to other OECD countries, and the Gini coefficient has not changed a great deal between 2002 and 2018 in the HILDA dataset [@australian2018rising].
+A stable Gini coefficient shows income inequality has remained steady over the time period, and our results do not conflict with this conclusion.
+Rather what we are revealing is the effect of income inequality on happiness has increased over the same time period.
+So while income inequality has remained steady, it’s impact on wellbeing inequity has increased.
+We think this highlights the issue that while traditional measures of wealth and income inequality may not change, their impact on wellbeing and health can still vary.
+As focus shifts from traditional wealth indicators towards wellbeing measures, findings such as this may become more prevalent.
 
-Since 2013, the distinct effect of wealth and happiness has also been challenged.
-[A survey of 1.7 million people representing 164 countries over 12 years (2005-2016) in the Gallup World Poll](https://www-nature-com.ezproxy2.library.usyd.edu.au/articles/s41562-017-0277-0){target="_blank"} reported money was associated with no further improvement in happiness after \$60-75,000, _nor any improvement in satisfaction_ after $95,000 [@jebb2018happiness].
-Thus while the location differed, money had inflected effects with both satisfaction and happiness.
-On the other hand, in a nationally representative sample of 44,000 adult Americans over 44 years in the General Social Survey (GSS) happiness continued to increase with money [@twenge2020expanding], implying no inflection point existed between money and happiness.
-Both these most recent reports provide contradictory results regarding the nature of the relationship between wealth and wellbeing, and both are based on data spanning long periods of time (12-44 years).
-Perhaps the different results are due to differences in the constructs being measured, or perhaps they are due to the different effects of time between studies.
+Some recent studies have challenged either the notion that the positive effect of money plateaus, or that the effect on happiness and satisfaction are distinct.
+In a nationally representative sample of 44,000 adult Americans in the General Social Survey (GSS), happiness continued to increase with household wealth, implying no change point existed between money and happiness [@twenge2020expanding].
+The GSS asks a single item on happiness: “Taken all together, how would you say things are these days – would you say that you are very happy, pretty happy, or not too happy?” The form of this question is quite similar to the single item “life satisfaction” question in HILDA, as both request the respondent to cognitively evaluate their circumstances.
+By contrast, the nine items we selected to measure happiness covered a range of affect and focused on current feelings: e.g., “How much of the time during the past 4 weeks have you been a happy person?” Critical differences in the operational definition of happiness seem likely to explain the results observed here.
 
-We report here for the first time that the cost of happiness increased over time, faster than inflation or the median household income.
-In 2002, the cost point of happiness represented a `r round((29.41/26.86 - 1)*100)`% increase over median wealth, while in 2018 it represented a `r round((105.85/74.54 - 1)*100)`% increase over median wealth.
-This also represents a reduction from 44% to 26% in the proportion of people who have achieved a level of household wealth beyond which their happiness no longer depends.
+A second recent report from a survey of 1.7 million people representing 164 countries in the Gallup World Poll reported increases in household wealth were associated with change points in happiness and satisfaction [@jebb2018happiness], rather than distinct effects as we found.
+It may be that differences in the sparsity of the high income data in HILDA (for which extremely high income households are masked) explain the differences here.
+However including these individuals in our analysis did not change the linear effect between wealth and satisfaction observed here, and the weighted topcodes should have biased the result towards a change point.
+We also note a linear effect between wealth and satisfaction is consistent with the majority of earlier literature [@howell2008relation; @stevenson2013subjective].
 
-According to @frijters2020happy, coming up with a consensus estimate for income to translate into welbeing features high on the wellbeing research agenda.
-However evidence of a linear relationship between wellbeing and wealth calls into question the need to supplant traditional measures of economic success such as GDP with measures of wellbeing, such as "life satisfaction" surveys.
+As governments and policy-makers begin to focus on wellbeing, it will be critical to understand how traditional economic indicators such as household wealth, income inequality, and consumption interact with wellbeing and health.
+According to @frijters2020happy, coming up with a consensus to translate income into wellbeing features high on the wider wellbeing research agenda.
+Establishing the links between income, household wealth, wellbeing and health, and how inequalities in one drives inequities in the other, will be a critical step in that agenda.
 

### 2020-11-14 drafts/wdraft0.Rmd
Update to wdraft0 and README

@@
   word_document:
-    reference_docx: ../src/academic_style.docx
+    reference_docx: ../src/Rmarkdown-for-Word-styles.docx
 always_allow_html: yes

### 2020-11-16 src/Clark et al 2008 Figure 4.Rmd
new docs folder

---
title: "Clark et al 2008 Figure 4"
author: "Dr Richard Morris & Prof. Nicholas Glozier"
date: "Updated: `r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.path = "../figures/")
library(tidyverse)
```
***
<br>

Clark, A. E., Frijters, P., & Shields, M. A. (2008). Relative income, happiness, and utility: An explanation for the Easterlin paradox and other puzzles. *Journal of Economic literature, 46*(1), 95-144.

<br>

*Figure 4.* The Relationship between Income and Happiness at the Individual and the Aggregate Level

![](../figures/Figure 4.png)

<br><br>

##### Household wealth and happiness (HILDA 2002-2018)

```{r plot_gh9}
happywealth <- read_rds("../results/happywealth.rds")

happywealth %>%
  select(xwaveid, year, 
         # swb = enexpr(swb_col), dollars = enexpr(dollars_col)) %>%
         swb = gh9, dollars = hifdip) %>%
  filter(swb >= 0) %>%
  filter(dollars >= 0) %>%
  filter(year %in% c(2002, 2010, 2018)) %>%
  mutate(year = as.factor(year)) -> df

df %>%
  group_by(year) %>%
  mutate(
    dollar_bin = ntile(dollars, 10)
  ) %>%
  group_by(year, dollar_bin) %>%
  summarize(
    wellbeing = mean(swb, na.rm=T),
    wealth = mean(dollars, na.rm=T)/1000,
    n = n()
  ) -> df.summary

df.summary %>%
  transmute(
    wealth = mean(wealth),
    wellbeing = mean(wellbeing)
  ) %>%
  distinct() -> df.summary.year

# piecewise model
ggplot(df.summary, aes(x = wealth, y = wellbeing)) +
  geom_point(aes(color = year)) +
  geom_smooth(aes(color = year), 
              method = "lm", formula = y ~ splines::bs(x, df = 2, degree = 1),
              # method = "lm", formula = y ~ log(1/(x+1)),
              linetype = 2,
              size = .5,
              se=F) +
  geom_smooth(data = df.summary.year, 
              method = "lm", formula = y ~ log(1/(x+1)), 
              color = "black",
              se=F) + 
  # scale_x_continuous(breaks = c(10, 30, 100, 300), trans = "log") +
  theme_test() +
  labs(subtitle = "SF-36 item 9 (summed)", 
       y = "", x = "household wealth ($000s)")
```

<br><br>

##### Household wealth and life satisfaction (HILDA 2002-2018)

```{r plot_losat}
happywealth %>%
  select(xwaveid, year, 
         # swb = enexpr(swb_col), dollars = enexpr(dollars_col)) %>%
         swb = losat, dollars = hifdip) %>%
  filter(swb >= 0) %>%
  filter(dollars >= 0) %>%
  filter(year %in% c(2002, 2010, 2018)) %>%
  mutate(year = as.factor(year)) -> df

df %>%
  group_by(year) %>%
  mutate(
    dollar_bin = ntile(dollars, 10)
  ) %>%
  group_by(year, dollar_bin) %>%
  summarize(
    wellbeing = mean(swb, na.rm=T),
    wealth = mean(dollars, na.rm=T)/1000,
    n = n()
  ) -> df.summary

df.summary %>%
  transmute(
    wealth = mean(wealth),
    wellbeing = mean(wellbeing)
  ) %>%
  distinct() -> df.summary.year

# piecewise model
ggplot(df.summary, aes(x = wealth, y = wellbeing)) +
  geom_point(aes(color = year)) +
  geom_smooth(aes(color = year), 
              method = "lm", formula = y ~ splines::bs(x, df = 2, degree = 1),
              # method = "lm", formula = y ~ log(1/(x+1)),
              linetype = 2,
              size = .5,
              se=F) +
  geom_smooth(data = df.summary.year, 
              method = "lm", formula = y ~ log(1/(x+1)), 
              color = "black",
              se=F) + 
  # scale_x_continuous(breaks = c(10, 30, 100, 300), trans = "log") +
  theme_test() +
  labs(subtitle = "'All things considered, how satisfied are you with your life?' (0-10)", 
       y = "", x = "household wealth ($000s)")
```

<br><br>

#### Methods

The wealth and wellbeing variables were used from HILDA v18 without further adjustment or indexing.

##### Appendix

```{r plot_gini, eval=F}
# From http://freerangestats.info/blog/2017/08/05/weighted-gini
# Duoba and MacGibbon's function, reproduced as in the original paper:
StatsGini <- function(x, w = rep(1, length(x))){
  # x and w are vectors
  # w can be left blank when calling the fn (i.e. no weighting)
  # Examples:
  # x <- c(3, 1, 7, 2, 5)
  # w <- c(1, 2, 3, 4, 5)
  # StatsGini(x, w) should yield 0.2983050847
  # StatsGini(c(0.25, 0.75), c(1, 1)) should yield 0.25

  n <- length(x)                               # number of wealth observations (N)
  wxsum <- sum(w * x)                          # sum of (weighted) wealth observations
  wsum <- sum(w)                               # sum of weights (e.g., N when unweighted)
  sxw <- order(x, w)                           # Order of (weighted) wealth observations
  sx <- w[sxw] * x[sxw]                        # Reorders the (weighted) wealth observations
  sw <- w[sxw]
  pxi <- vector(mode = "numeric", length = n)  # Vector of zeros
  pci <- vector(mode = "numeric", length = n)  # Vector of zeros
  pxi <- cumsum(sx) / wxsum
  pci <- cumsum(sw) / wsum
  G <- 0.0
  for (i in 2:n){
    G <- G - (pci[i] * pxi[i - 1] - pxi[i] * pci[i - 1] )
  }
  return(G)
}

happywealth %>%
  select(year, xwaveid, hifdip) %>%
  group_by(year) %>%
  summarise(
    gini = StatsGini(hifdip)
  ) %>%
  ggplot(aes(x = year, y = gini)) +
    geom_line() +
    ylim(x = c(0.25,0.35)) +
    labs(subtitle = "Gini coefficient", y = "") +
    theme_test()
```

### 2020-11-19 src/appendix.Rmd
New version ready for Nick

---
title: "Supplementary material: The increasing cost of happiness"
author: "R.W. Morris^1,2^, N. Kettlewell^2,3,4^  & N.Glozier^1,5^"
bibliography: ../src/references.bib
csl: ../src/elsevier-harvard.csl
output:  
  html_document:
    reference_docx: ../src/academic_style.docx
always_allow_html: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(flextable)
library(bayesplot)

knitr::opts_chunk$set(eval = TRUE,
                      echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      fig.path = "../figures/")
```
\  
\

1. Central Clinical School, Faculty of Medicine and Health, University of Sydney, NSW, Australia
2. ARC Centre of Excellence for Children and Families over the Life Course, University of Queensland, QLD, Australia
3. School of Economics, University of Technology, NSW, Australia
4. Institute of Labor Economics (IZA), Bonn, Germany
5. Brain and Mind Centre, University of Sydney, NSW, Australia

\

**Corresponding author:**  
Professor Nick Glozier  
Faculty of Medicine and Health,   
University of Sydney,  
NSW 2050,  
Australia  
email: nick.glozier@sydney.edu.au

\

Draft: `r format(Sys.time(), '%d %B, %Y')`  
Supplementary tables:      1  
Supplementary figures:     2

**keywords:** Subjective wellbeing, household wealth, HILDA

\

***

## Supplementary Material

```{r read_write_covs}
happywealth_covs <- read_rds("../results/happywealth_covs.rds")
```

```{r funtimes}
source("../src/happyfun.r")
```

We tested the relationship between wealth and each wellbeing variable (happiness and satisfaction) using cubic **B-spline models**. A cubic B-spline was the most complex model (more degrees of freedom), and chosen because it has been recently used to describe the maximum point at which wellbeing no longer increases with wealth [i.e., "satiety", @jebb2018happiness]. Following Jebb et al [-@jebb2018happiness], we restricted the cubic B-spline model to *n* = 4, 5 and 6 knots and display wealth on a log scale.

<br>

##### Figure S1. Cubic spline of household wealth on satisfaction and happiness
```{r losat_gh9_hwnwip}
library(patchwork)

p1 <- happywealth_covs %>%
  filter(adj_hifdip > 0) %>%
  rename(Satisfaction = losat, `Household wealth` = adj_hifdip) %>% 
  plot_inflection_cubic(Satisfaction, `Household wealth`) +
  scale_x_continuous(trans = "log", 
                     breaks = c(5, 10, 20, 40, 80, 160),
                     limits = c(2.5, 170))

p2 <- happywealth_covs %>%
  filter(adj_hifdip > 0) %>%
  rename(Happiness = gh9, `Household wealth` = adj_hifdip) %>% 
  plot_inflection_cubic(Happiness, `Household wealth`) +
  scale_x_continuous(trans = "log", 
                     breaks = c(5, 10, 20, 40, 80, 160),
                     limits = c(2.5, 170))
```
```{r cubic_plot, fig.height=9, fig.width=11}
p1 / p2
```

<br>

The cubic B-spline models failed to reveal a complete satiety point (horizontal inflection) in satisfaction before the highest decile in any year. Also note that despite using a log scale, we do not observe a clear, regular logarithmic function with either wellbeing measure. Other researchers have argued that wealth represents a quantitative dimension subject to Weber's Law of perception [@kahneman2010high]. Under this view, similar proportional changes in income or wealth produce similar evaluations of satisfaction or happiness, rather than the absolute change. That is, a \$100 increase in wealth will produce a greater impact on someone earning a minimum wage than someone earning many orders of magnitude of that amount. However while it is true that wealth clearly represents a quantitative dimension, the effects of wealth observed here represent cross-sectional differences in the population rather than changes experienced by an individual over time. It is not clear the extent to which Weber's Law applies to cross-sectional or between-subject differences in quantitative dimensions such as wealth.

<br>

#### Change point model parameter estimates

There were _df_ = 59,876 total observations. 4000 samples (post-warmup) were drawn using sampling(NUTS). For each parameter, Bulk_ESS and Tail_ESS are effective sample size measures, and Rhat is the potential scale reduction factor on split chains (at convergence, Rhat = 1).

<br>

_Coefficient definitions_:

_b0_ is the intercept denoting the expected value of wellbeing (y) at the changepoint  
_b1_ is the linear slope before the changepoint  
_b2_ is the linear slope after the changepoint  
_b3_ is the coefficient of education (graduate)  
_b4_ is the coefficient of sex (male)  
_b5_ is the coefficient of age  
_b6_ is the coefficient of age^2^  
_alpha_ is the expected value of wealth (x) at the changepoint

<br>

```{r fit_summary, ft.align="left"}
hifdip_gh9_fixed_fit <- read_rds("../results/hifdip_gh9_fixed_full_covs.RDS")

fit_summary <- summary(hifdip_gh9_fixed_fit)

sigma <- fit_summary$spec_pars %>% 
  data.frame() %>%
  rownames_to_column("Coef") %>%
  mutate(Parameter = "Family Specific")

param_key = c(Intercept = "2002 (Intercept)",
              year2006 = "2006",
              year2010 = "2010",
              year2014 = "2014",
              year2018 = "2018")

fit_summary$fixed %>%
  data.frame() %>%
  rownames_to_column("Parameter") %>%
  separate(col = `Parameter`, into = c("Coef", "Parameter")) %>%
  mutate(Parameter = recode(Parameter, !!!param_key)) %>%
  bind_rows(sigma) %>%
  flextable() %>%
  merge_v(j = 1) %>%
  valign(j = 1, valign = "top") %>%
  bg(i = ~Coef %in% c("b0", "b2", "b4", "b6"), bg = "grey90", part = "body") %>%
  bg(i = ~Coef %in% c("sigma"), bg = "#b3cde0", part = "body") %>%
  colformat_num(j = 3:6, digits= 3) %>%
  colformat_num(j = 7, digits= 1) %>%
  set_caption("Table S1. Complete list of model parameter estimates") %>%
  autofit()
```

<br><br>

##### Figure S2. Posterior predictive distribution of marginal effect of wealth by year (95% CI)

```{r figure_s2, fig.height = 3}
library(brms)

marginals <- conditional_effects(
  hifdip_gh9_fixed_fit,
  effects = c("dollars:year"), # year, dollars, 
  method = "fitted"
  )

marginals$`dollars:year`$year <- factor(marginals$`dollars:year`$year, 
                                        levels = c("2002", "2006", "2010", "2014", "2018"))

plot(marginals, plot=F)[[1]] +
  xlim(c(0, 30)) +
  ylim(c(-0.5, 1)) +
  facet_wrap(~year, nrow=1) +
  theme_test() +
  theme(legend.position = "none") +
  labs(x = "", y = "")
```

<br><br>

## References

### 2020-11-19 src/draft0.Rmd
New version ready for Nick

--- 2020-11-14 drafts/draft0.Rmd
@@
 | Draft      | `r format(Sys.time(), '%d %B, %Y')` |
-| Word count | 3760 (max 4000)                     |
-| Tables     | 1                                   |
+| Word count | 3061                                |
+| Tables     | 2                                   |
 | Figures    | 3                                   |
@@
 This result shows that inequalities in wealth may be driving inequities in happiness and wellbeing, with implications for health and recent government policy-goals to monitor and improve wellbeing.
+(150 words)
 
@@
 
-***
+<br><br>
+
+<br><br>
 
@@
 We modelled the relationship between wealth and each wellbeing variable (happiness and satisfaction) using a simple **linear model** and a **piecewise model** (broken-stick).
-The linear model was the simplest relationship while the piecewise model was chosen as the simplest extension of a linear model which can identify a change point (inflection) between wellbeing and wealth.
-The location of the change point was a free parameter which revealed where wellbeing no longer increased linearly with wealth.
-We then compared the linear model against the piecewise model to determine if a change point existed in each or any year between wealth and each wellbeing variable (see model selection).
-Finally, where a change point existed, we determined the location of the change point for that year (see change point estimation).
+The piecewise model was chosen as the simplest extension of a linear model which can identify a change point (inflection) between wellbeing and wealth.
+The location of the change point was a free parameter which revealed where wellbeing no longer increased at a uniform rate with wealth.
+We then compared the linear model against the piecewise model to determine if a change point existed in any year between household wealth and each wellbeing variable (see Model Selection).
+Finally, where a change point existed, we determined the location of the change point for that year (see Parameter Estimation).
 
-_Model estimation_  
+_Model Estimation_  
 We adopted a Bayesian approach for estimating the linear and piecewise model in the software Stan [@burkner2017; @rstan2019]. In each case,
@@
 
-_Model selection_  
+_Model Selection_  
 To determine whether wellbeing was a linear or non-linear (e.g., piecewise) function of wealth, we compared the linear and piecewise model posterior probabilities using the [Widely Applicable Information Criterion (WAIC)](https://bookdown.org/ajkurz/Statistical_Rethinking_recoded/overfitting-regularization-and-information-criteria.html#the-problem-with-parameters){target="_blank"}.
@@
 
-_Parameter estimation_  
+_Parameter Estimation_  
 To determine the location of the change point (*&omega;*) between wellbeing and wealth, we modelled the relationship between wealth and wellbeing across individuals using the piecewise model described above, and sampled the posterior probability of *&omega;* over 4000 interations.
@@
 Age (and age^2^), gender, and education were included as covariates, and fulltime students were removed.
-Gender was included as a binary variable, and education was coded from the highest level of education achieved ("less than highschool", "highschool", "university graduate", "postgraduate").
+Gender was included as a binary variable (Male = 1), and education was a binary variable coded from the highest level of education achieved (university/college graduate = 1).
 
@@
 
+The broad demographic characteristics of the sample from each year are presented below in Table 1.
+
+```{r table_1, ft.align="left"}
+library(gtsummary)
+
+happywealth_covs %>%
+  filter(!student, (!is.na(losat)|!is.na(gh9))) %>%
+  mutate(
+    Sex = if_else(male, "Male", "Female"),
+    SEIFA = as.integer(extract_numeric(SEIFA)),
+    SEIFA = if_else(SEIFA > 0, SEIFA, NA_integer_),
+    Education = fct_collapse(
+      edu,
+      `High school or less` = c("None", "In school", "Highschool", "Trade"),
+      Graduate = c("Grad", "Postgrad")
+      ),
+    Workforce = case_when(
+      workforce & !unemployed ~ "Employed",
+      unemployed ~ "Unemployed",
+      TRUE ~ "Not in labour force"
+      ),
+    `Relationship status` = case_when(
+      single ~ "Single",
+      coupled ~ "Married",
+      TRUE ~ "Divorced/Widow"
+      ),
+    Workforce = fct_relevel(Workforce, "Employed", "Unemployed"),
+    `Relationship status` = fct_relevel(`Relationship status`, "Single", "Married")
+    ) %>%
+  select(
+    year, 
+    # `Life satisfaction` = losat, `Happiness` = gh9, 
+    # `Household wealth ($)` = adj_hifdip, 
+    Sex, `Age (years)` = age, Education, Workforce, `Relationship status`, 
+    `Chronic illness` = chronic, SEIFA, `Household size` = hhsize,
+  ) %>% 
+  tbl_summary(
+    by = year,
+    missing = "no",
+    statistic = 
+      list(all_continuous() ~ "{median} ({p25}-{p75})",
+           all_categorical() ~ "{n} ({p}%)"),
+    digits = all_continuous() ~ 0,
+    missing_text = "(Missing)" 
+    ) %>%
+  modify_header(stat_by = "{level}\nN = {style_number(n)}") %>%
+  as_flex_table() %>%
+  set_caption(caption = "Table 1. Demographic characteristics") %>%
+  align(part = "body") %>%
+  valign(valign = "top", part = "header") %>%
+  fontsize(size = 9.5, part = "all") %>%
+  width(j = 1, width = 1.5) %>%
+  width(j = 2:6, width = 1)
+```
+
+<br>
+
+The proportions of each sex and relationship status were stable over time, as were the average household size and SEIFA index.
+However age and chronic health conditions tended to increase over time, as expected in an aging longitudinal sample.
+Education levels (i.e., graduate percentage) also tended to increase over time, while changes in the workforce varied both up and down.
+
+<br>
+
 ```{r funtimes}
@@
 
-For visualization purposes only, due to the large number of individual data points in each year, we grouped individuals into 10 equal-sized subgroups according to their wealth decile ($).
-Thus in each plot, the mean wealth and mean wellbeing score for each decile is presented, rather than each individual data point.
-Note that the line-of-best-fit and 95% confidence intervals from each regression model _of all individuals_ is shown in overlay.
+Below we show the relationship between household wealth and wellbeing (Figure 1).
+For visualization purposes only, due to the large numbers of individual data in each year, we display the mean levels of wealth and wellbeing for each (equal-sized) wealth decile rather than every individual data point.
+Note that the line-of-best-fit and 95% credible intervals (shaded) from each regression model _of all individuals_ is shown in overlay.
 
@@
   rename(Satisfaction = losat, `Household wealth` = adj_hifdip) %>% 
+  filter(!student) %>%
   plot_linear(Satisfaction, `Household wealth`) +
@@
   rename(Satisfaction = losat, `Household wealth` = adj_hifdip) %>% 
+  filter(!student) %>%
   plot_inflection_piecewise(Satisfaction, `Household wealth`) +
@@
   rename(Happiness = gh9, `Household wealth` = adj_hifdip) %>% 
+  filter(!student) %>%
   plot_linear(Happiness, `Household wealth`) +
@@
   rename(Happiness = gh9, `Household wealth` = adj_hifdip) %>% 
+  filter(!student) %>%
   plot_inflection_custom(Happiness, `Household wealth`, 
@@
 ```
-Figure 1 legend: Line-of-best-fit and 95% confidence interval (shaded) from regressions of  wellbeing on household wealth, with age, gender and education included as covariates.
+Figure 1 legend: Posterior predictive distributions (95% credible intervals) showing the model fit from regressions of wellbeing on household wealth, overlaid on summary data points (wealth deciles).
 Wellbeing was measured as life satisfaction (blue) or happiness (red).
-The total number of individuals in each model are shown (n).
-Points represent the mean value of each wealth decile.
+The total number of individuals contributing to each year are shown (n).
 ```
@@
 
-We present the complete posterior distribution of &omega; over 4000 samples for each year, representing the location of the change point between happiness and wealth.
+We present the complete posterior distribution of &omega; over 4000 samples drawn from the piecewise model, representing the location of the change point between happiness and household wealth for each year.
 Shaded areas to the right of the vertical grey dotted line are credibly (90%) larger than the base year (2002).
@@
 
-##### Figure 3. Posterior distributions of &omega; (mean & 90% distribution)  
+##### Figure 3. Posterior distributions of &omega; (90% distribution)  
 ```{r figure_3, fig.width = 6, fig.height = 4.5, dpi=300}
@@
 Table 2 below compares the change in median and mean wealth for each year with the cost of happiness.
-It also present the percentage of people in each year whose household wealth exceeded the cost point.
+It presents the percentage of people in each year whose household wealth exceeded the cost point of happiness.
 
@@
   flextable() %>%
-  set_caption("Table 1. Household wealth, change points and percentile") %>%
+  set_caption("Table 2. Household wealth, change points and percentile") %>%
   autofit()
@@
 
-Table 1 shows the cost of happiness increased faster than median household wealth.
-As a result, a smaller percentage of people achieved a level of wealth at which happiness no longer increased in 2018 than in 2002.
+Table 2 shows the cost of happiness increased faster than the median or mean household wealth between 2002 and 2018.
+As a result, a smaller percentage of people in 2018 achieved a level of financial security on which their happiness was no longer dependent on their wealth, relative to 2002.
 
@@
 
-We found the relationship between subjective wellbeing and wealth was positive, but happiness and satisfaction had different (positive) relationships: Satisfaction increased linearly with wealth, while happiness increased rapidly up to a point after which further increments in wealth produced less change – this confirms the distinct effects of wealth in previous research [e.g., @howell2008relation; @kahneman2010high], and contributes rare evidence from the same sample.
-Furthermore, we report here for the first time that the change point between wealth and happiness increased over time between 2002 and 2018, faster than inflation or the median household income.
+We found the relationship between subjective wellbeing and household wealth was positive, but happiness and satisfaction had different (positive) relationships: Satisfaction increased linearly with wealth, while happiness increased rapidly up to a point after which further increments in wealth produced less change – this confirms the distinct effects of wealth in previous research [e.g., @howell2008relation; @kahneman2010high], and contributes rare evidence from the same sample.
+Furthermore, we report here for the first time that the change point between household wealth and happiness increased over time between 2002 and 2018, faster than inflation or the median household income.
 
 We refer to the change point after which increases in wealth no longer produce similar increases in happiness as the cost of happiness.
-So household wealth, and the economic security it represents, buys happiness up to a point after which happiness and wealth are rarely related.
-Satisfaction on the other hand always increased with household wealth.
-The difference likely reflects the importance of a numerical dollar value (e.g., bank balance) when cognitively appraising one’s life achievements, versus the relevance of that number to our everyday experience of joy and prevailing mood.
-However an implication of the increasing cost of happiness over the last sixteen years is that wealth inequality is driving increasing _inequities_ in wellbeing.
+After this point, happiness is no longer as dependent on household wealth, and the economic security it represents.
+Presumably after this point further increases in happiness depend on other life factors (e.g., social connections).
+Life satisfaction on the other hand always increased with household wealth.
+The difference likely reflects the importance of a numerical dollar value (e.g., bank balance, house value) when cognitively appraising one’s life achievements, versus the relevance of that number to our everyday experience of joy and prevailing mood.
+
+An implication of the increasing cost of happiness over the last sixteen years is that wealth inequality is driving increasing _inequities_ in wellbeing.
 In 2002, the cost point of happiness represented a 9% increase over median wealth, while in 2018 it represented a 42% increase over median wealth.
-This also represents a reduction from 44% to 26% in the proportion of people who have achieved a level of household wealth beyond which their happiness no longer depends.
-Thus over time the happiness of more people, i.e., their everyday experience of joy and their prevailing mood, has depended on their material wealth.
+This also represents a reduction from 44% to 26% in the proportion of people who have achieved a level of financial security beyond which their happiness no longer depends.
+Thus we can see that over time in the last sixteen years the happiness of more people, i.e., their everyday experience of joy and their prevailing mood, has depended on their material wealth.
 
@@
 So while income inequality has remained steady, it’s impact on wellbeing inequity has increased.
-We think this highlights the issue that while traditional measures of wealth and income inequality may not change, their impact on wellbeing and health can still vary.
+We think this highlights the issue that while traditional measures of wealth and income inequality may be relatively stable and exhibit little change, their impact on wellbeing and health can still vary.
 As focus shifts from traditional wealth indicators towards wellbeing measures, findings such as this may become more prevalent.
 
-Some recent studies have challenged either the notion that the positive effect of money plateaus, or that the effect on happiness and satisfaction are distinct.
+Some recent studies have challenged either the notion that the positive effect of household wealth plateaus, or that the effect on happiness and satisfaction are distinct.
 In a nationally representative sample of 44,000 adult Americans in the General Social Survey (GSS), happiness continued to increase with household wealth, implying no change point existed between money and happiness [@twenge2020expanding].
 The GSS asks a single item on happiness: “Taken all together, how would you say things are these days – would you say that you are very happy, pretty happy, or not too happy?” The form of this question is quite similar to the single item “life satisfaction” question in HILDA, as both request the respondent to cognitively evaluate their circumstances.
-By contrast, the nine items we selected to measure happiness covered a range of affect and focused on current feelings: e.g., “How much of the time during the past 4 weeks have you been a happy person?” Critical differences in the operational definition of happiness seem likely to explain the results observed here.
+By contrast, the nine items we selected to measure happiness covered a range of affect and focused on current feelings: e.g., “How much of the time during the past 4 weeks have you been a happy person?”.
+Critical differences in the operational definition of happiness seem likely to explain the distinct results observed here.
 

### 2020-11-19 src/wdraft0.Rmd
New version ready for Nick

--- 2020-11-19 src/draft0.Rmd
@@
 output:  
-  html_document:
-    toc: true
-    toc_depth: 4
-    toc_float:
-      collapsed: false
+  word_document:
+    reference_docx: ../src/Rmarkdown-for-Word-styles.docx
 always_allow_html: yes

### 2020-11-19 src/happyfun.R

@@ -248,10 +248,10 @@ plot_inflection_cubic <- function(.data, swb_col, dollars_col) {
                 colour = "navy",
                 size = 0.5,
                 se = FALSE) +
-    coord_cartesian(xlim = c(0, max(df.summary$wealth)),
-                    ylim = c(min(df.summary$wellbeing), max(df.summary$wellbeing))) +
+    # coord_cartesian(xlim = c(0, max(df.summary$wealth)),
+    #                 ylim = c(min(df.summary$wellbeing), max(df.summary$wellbeing))) +
     facet_wrap(~year, nrow = 1) +
-    labs(subtitle = paste(rlang::as_name(swb_colname), "(b-spline)"),
+    labs(subtitle = paste(rlang::as_name(swb_colname), "(b-spline with knots = 4, 5, 6)"),
          x = paste(rlang::as_name(dollars_colname), "($000s)"),
          y = "") +
     theme_test()

### 2020-11-23 src/Bin size plots.Rmd
Fixed deflation multiplier and calculated results for all years 2002-2018

---
title: "Bin size experiments"
author: "Dr R. Morris, School of Medicine, University of Sydney"
date: "23/11/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE,
                      echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      fig.path = "../figures/")

library(tidyverse)
```

## Depicting large data

This is a brief examination of the effect of the bin design on data visualization (e.g., scatter plots).

HILDA provides income and household wealth data on approximately 10,000 individuals each year. When visualizing large datasets of n > 10,000 there are often too many points to distinguish patterns among the individuals or even the overall trend. Models, or some kind of summary function, is needed to provide an accurate or fair depiction of trends and patterns. Measures of central tendency, or the expected value can also be used to reduce the number of indiivdual points, however then the data must be grouped or binned. Binning can occur by equal length or equal size.

```{r read_covs}
happywealth_covs <- read_rds("../data/happywealth_covs.rds")
```

```{r plot_all}
plot_all <- function(.data, swb_col, dollars_col) {

  swb_colname <- rlang::enexpr(swb_col)
  dollars_colname <- rlang::enexpr(dollars_col)

  .data %>%
    select(xwaveid, 
           year, 
           swb = !!swb_colname, dollars = !!dollars_colname) %>%
    filter(swb >= 0) %>%
    filter(dollars >= 0) %>%
    filter(year %in% c(2002, 2006, 2010, 2014, 2018)) -> df

  df %>%
    group_by(year) %>%
    summarise(n = n()) %>%
    mutate(note = paste("n =", format(n, big.mark = ","))) -> df.sum

  df %>%
    mutate(
      wealth = dollars/1000,
      wellbeing = swb
    ) -> df.plot

  # smooth 
  ggplot(df.plot, aes(x = wealth, y = wellbeing)) +
    geom_point(
      alpha = 0.05
      ) +
    geom_smooth(
      aes(fill = "manual1", color = "manual2"),
      method = "gam",
      size = 0.5
      ) +
    coord_cartesian(
      ylim = c(min(df.plot$wellbeing), max(df.plot$wellbeing))
      ) +
    facet_wrap(
      ~year, 
      nrow = 1,
      scales = "free_x"
      ) +
    labs(
      subtitle = paste(rlang::as_name(swb_colname), "(all)"),
      x = paste(rlang::as_name(dollars_colname), "($000s)"),
      y = ""
      ) +
    theme_test() +
    theme(
      legend.background = element_blank(),
      legend.title = element_blank(),
      legend.position = "none"
      ) -> p1

  p1 +
    geom_text(
      data = df.sum, 
      aes(x = Inf, y = -Inf, label = note),
      hjust = 1.1, vjust = -.75, colour = 1, size = 3)

}
```

```{r plot_cut_interval}
plot_cut_interval <- function(.data, swb_col, dollars_col) {

  swb_colname <- rlang::enexpr(swb_col)
  dollars_colname <- rlang::enexpr(dollars_col)

  .data %>%
    select(xwaveid, year, swb = enexpr(swb_col), dollars = enexpr(dollars_col)) %>%
    filter(swb >= 0) %>%
    filter(dollars >= 0) %>%
    filter(year %in% c(2002, 2006, 2010, 2014, 2018)) -> df

  df %>%
    mutate(
      wealth = dollars/1000,
      wellbeing = swb
    ) -> df.plot

  df %>%
    group_by(year) %>%
    mutate(
      dollar_bin = cut_interval(dollars, 10)
      ) %>%
    group_by(year, dollar_bin) %>%
    summarize(
      wellbeing = mean(swb, na.rm=T),
      wealth = mean(dollars, na.rm=T)/1000,
      n = n()
      ) -> df.summary

  # smooth
  ggplot(df.plot, aes(x = wealth, y = wellbeing)) +
    geom_smooth(
      aes(fill = "manual1", color = "manual2"),
      method = "gam", 
      size = 0.5
      ) +
    facet_wrap(~year, nrow = 1, scale = "free_x") +
    labs(
      subtitle = paste(rlang::as_name(swb_colname), "(equal intervals)"),
      x = paste(rlang::as_name(dollars_colname), "($000s)"),
      y = "") +
    theme_test() +
    theme(
      legend.position = "none") -> p1

  p1 +
    geom_point(data = df.summary, aes(size = sqrt(n)), alpha = 0.5)
}
```

```{r plot_deciles}
plot_deciles <- function(.data, swb_col, dollars_col) {

  swb_colname <- rlang::enexpr(swb_col)
  dollars_colname <- rlang::enexpr(dollars_col)

  .data %>%
    select(xwaveid, 
           year, 
           swb = !!swb_colname, dollars = !!dollars_colname) %>%
    filter(swb >= 0) %>%
    filter(dollars >= 0) %>%
    filter(year %in% c(2002, 2006, 2010, 2014, 2018)) -> df

  df %>%
    group_by(year) %>%
    summarise(n = n()) %>%
    mutate(note = paste("n =", format(n, big.mark = ","))) -> df.sum

  df %>%
    group_by(year) %>%
    mutate(
      dollar_bin = ntile(dollars, 10)
    ) %>%
    group_by(year, dollar_bin) %>%
    summarize(
      wellbeing = mean(swb, na.rm=T),
      wealth = mean(dollars, na.rm=T)/1000
      ) -> df.deciles

  df %>%
    mutate(
      wealth = dollars/1000,
      wellbeing = swb
    ) -> df.plot

  # smooth
  ggplot(df.plot, aes(x = wealth, y = wellbeing)) +
    geom_smooth(aes(fill = "manual1", color = "manual2"),
      method = "gam", 
      size = 0.5
      ) +
    facet_wrap(~year, nrow = 1, scales = "free_x") +
    labs(subtitle = paste(rlang::as_name(swb_colname), "(deciles)"),
         x = paste(rlang::as_name(dollars_colname), "($000s)"),
         y = "") +
    theme_test() +
    theme(
      legend.position = "none") -> p1

  p1 +
    geom_point(data = df.deciles)  # size = 3, alpha = 0.25
    # geom_text(data = df.sum, aes(x = Inf, y = -Inf, label = note),
    #           hjust = 1.1, vjust = -.75, colour = 1, size = 3)

}
```

#### Unsummarized plots
```{r unsum_plot, fig.width=9}
happywealth_covs %>%
  rename(Happiness = gh9, `Household wealth` = adj_hifdip) %>% 
  filter(!student) %>%
  plot_all(Happiness, `Household wealth`) +
  scale_fill_manual(values = "#b3cde0") +
  scale_color_manual(values = "#005b96")
```

#### Equal interval bins
```{r cut_interval_plot, fig.width=9}
happywealth_covs %>%
  rename(Happiness = gh9, `Household wealth` = adj_hifdip) %>% 
  filter(!student) %>%
  plot_cut_interval(Happiness, `Household wealth`) +
  scale_fill_manual(values = "#b3cde0") +
  scale_color_manual(values = "#005b96")
```

#### Decile bins
```{r decile_plot, fig.width=9}
happywealth_covs %>%
  rename(Happiness = gh9, `Household wealth` = adj_hifdip) %>% 
  filter(!student) %>%
  plot_deciles(Happiness, `Household wealth`) +
  scale_fill_manual(values = "#b3cde0") +
  scale_color_manual(values = "#005b96")
```

### 2020-11-23 src/draft0.Rmd
Fixed deflation multiplier and calculated results for all years 2002-2018

@@
 
-Where $X_i$ was an individual's household wealth (\$) as well as other covariates (age, age^2^, gender, education), and $y_i$ was an individual's wellbeing.
+Where $X_i$ was an individual's household wealth (\$) as well as other covariates (age, age^2^, sex, education), and $y_i$ was an individual's wellbeing.
 
@@
 
-The piecewise model was a simple extension of this to include a free parameter to represent the changepoint in wealth ($\omega$) as well as the slope after the change point ($\beta_2$):
+The piecewise model was a simple extension of this to include a free parameter to represent the changepoint in wealth ($\omega$) as well as the slope before the change point ($\beta_1$) and after the change point ($\beta_2$):
 
 $$
-\mu_i = \beta_0 + \beta_1 (x_i - \omega_t) (x_i ≤ \omega_t) + \beta_2 (x_i - \omega_t) (x_i > \omega_t)
+\mu_i = \beta_0 + \beta_1 (x_i - \omega) (x_i ≤ \omega) + \beta_2 (x_i - \omega) (x_i > \omega) + \beta_3 X_i
 $$
+
+Where $X_i$ were covariates for age, age^2^, sex and education.
+
 <br>
@@
 ```{r read_write_covs}
-happywealth_covs <- read_rds("../results/happywealth_covs.rds")
+happywealth_covs <- read_rds("../data/happywealth_covs.rds")
 ```
@@
 happywealth_covs %>%
+  filter(year %in% c(2002, 2006, 2010, 2014, 2018)) %>%
   filter(!student, (!is.na(losat)|!is.na(gh9))) %>%
@@
     year, 
-    # `Life satisfaction` = losat, `Happiness` = gh9, 
+    `Life satisfaction` = losat, `Happiness` = gh9,
     # `Household wealth ($)` = adj_hifdip, 
@@
   plot_inflection_custom(Happiness, `Household wealth`, 
-                         knots = c(32, 46, 52.5, 86, 124.5)) +
+                         knots = c(43.7, 49.2, 47.4, 53.2, 72.3)) +
   scale_fill_manual(values = "#C79999") +
@@
 
-##### Figure 2. Model evidence (mean difference of WAIC~linear~ — WAIC~piecewise~)
+##### Figure 2. Linear model evidence (WAIC~linear~ — WAIC~piecewise~)
 
 ```{r plot_waic_losat}
-waic_result <- read_rds("../results/waic_losat.RDS")
+waic_result <- read_rds("../results/model_comparisons_waic_R2_losat.RDS")
 
-d1 <- data.frame(fy2018 = rnorm(n = 1e5, 
-                                     mean = waic_result[2, 1] * -2, 
-                                     sd   = waic_result[2, 2] *  2)/10,
-                 fy2014 = rnorm(n = 1e5, 
-                                     mean = waic_result[4, 1] * -2, 
-                                     sd   = waic_result[4, 2] *  2)/10,
-                 fy2010 = rnorm(n = 1e5, 
-                                     mean = waic_result[6, 1] * -2, 
-                                     sd   = waic_result[6, 2] *  2)/10,
-                 fy2006 = rnorm(n = 1e5, 
-                                     mean = waic_result[8, 1] * -2, 
-                                     sd   = waic_result[8, 2] *  2)/10,
-                 fy2002 = rnorm(n = 1e5, 
-                                     mean = waic_result[10, 1] * -2, 
-                                     sd   = waic_result[10, 2] *  2)/10
-                )
+waic_result %>%
+  select(model, year, elpd_diff, se_diff) %>%
+  filter(elpd_diff != 0) %>%
+  filter(year %in% c(2002, 2006, 2010, 2014, 2018)) %>%
+  rowwise() %>%
+  mutate(samples = list(rnorm(n = 1e5, 
+                              mean = elpd_diff * -2, 
+                              sd   = se_diff *  2)),
+         year = paste0("AD", year)
+         ) %>%
+  select(year, samples) %>%
+  group_by(year) %>%
+  unnest(samples) %>% 
+  mutate(sample_id = c(1:1e5)) %>%
+  pivot_wider(id_col = "sample_id", names_from = "year", values_from = "samples") %>%
+  select(-sample_id) -> d1
+
+# "fix" d1
+colnames(d1) <- rev(colnames(d1))
 
 color_scheme_set("blue")
+
 p5 <- mcmc_intervals(x = d1,
@@
   prob = 0.0,
-  prob_outer = .90
+  prob_outer = .95
   ) +
   scale_x_continuous(
-    limits = c(-10, 40),
-    labels = function(x) {x*10}) +
+    limits = c(-200, 150)
+    ) +
   geom_vline(aes(xintercept = 0), color = "grey90") +
@@
 ```{r plot_waic_gh9}
-waic_result <- read_rds("../results/waic_comparisons_gh9_percentiles.RDS")
+waic_result <- read_rds("../results/model_comparisons_waic_R2_gh9.RDS")
 
-d1 <- data.frame(fy2018 = rnorm(n = 1e5, 
-                                mean = waic_result[10, 1] * 2, 
-                                sd   = waic_result[10, 2] *  2)/10,
-                 fy2014 = rnorm(n = 1e5, 
-                                mean = waic_result[8, 1] * 2, 
-                                sd   = waic_result[8, 2] *  2)/10,
-                 fy2010 = rnorm(n = 1e5, 
-                                mean = waic_result[6, 1] * 2, 
-                                sd   = waic_result[6, 2] *  2)/10,
-                 fy2006 = rnorm(n = 1e5, 
-                                mean = waic_result[4, 1] * 2, 
-                                sd   = waic_result[4, 2] *  2)/10,
-                 fy2002 = rnorm(n = 1e5, 
-                                mean = waic_result[2, 1] * 2, 
-                                sd   = waic_result[2, 2] *  2)/10
-                 )
+waic_result %>%
+  select(model, year, elpd_diff, se_diff) %>%
+  filter(elpd_diff != 0) %>%
+  filter(year %in% c(2002, 2006, 2010, 2014, 2018)) %>%
+  rowwise() %>%
+  mutate(samples = list(rnorm(n = 1e5, 
+                              mean = elpd_diff * 2, 
+                              sd   = se_diff *  2)),
+         year = paste0("AD", year)
+         ) %>%
+  select(year, samples) %>%
+  group_by(year) %>%
+  unnest(samples) %>% 
+  mutate(sample_id = c(1:1e5)) %>%
+  pivot_wider(id_col = "sample_id", names_from = "year", values_from = "samples") %>%
+  select(-sample_id) -> d2
 
 color_scheme_set("red")
-p6 <- mcmc_intervals(x = d1,
-           pars = colnames(d1),
+
+p6 <- mcmc_intervals(x = d2,
+           pars = rev(colnames(d2)),
            prob = 0.00,
-           prob_outer = 0.90
+           prob_outer = 0.95
            ) +
   scale_x_continuous(
-    limits = c(-10, 40),
-    labels = function(x) {x*10}) +
+    limits = c(-200, 150)
+    ) +
   geom_vline(aes(xintercept = 0), color = "grey90") +
@@
     subtitle = "Happiness",
-    x = "WAIC difference (90% interval)")
+    x = "WAIC difference (95% interval)")
 ```
@@
 Figure 2 legend: Differences in posterior evidence for a linear fit over a piecewise fit (WAIC) for satisfaction (blue) and happiness (red).
-The filled circle indicates the mean of the distribution and the horizontal bars represents the 90% credible interval.
-A difference greater than zero is support for the linear model and below zero is support for the piecewise model.
+The filled circle indicates the mean of the distribution and the horizontal bars represents the 95% credible interval.
+A difference greater than zero is support for the linear model and a difference below zero is support for the piecewise model.
 ```
@@
 
-##### Figure 3. Posterior distributions of &omega; (90% distribution)  
+##### Figure 3. Posterior distributions of &omega; (95% distribution)  
 ```{r figure_3, fig.width = 6, fig.height = 4.5, dpi=300}
-hifdip_gh9_fixed_fit <- read_rds("../results/hifdip_gh9_fixed_full_covs.RDS")
+path_to_fits <- list.files(
+  path = "../results",#'~/Dropbox/HILDA/data',
+  pattern = '^fit_cp_hifdip_gh9_cov_.*.RDS$',
+  full.names = TRUE
+)
 
-as.data.frame(hifdip_gh9_fixed_fit) %>%
-  select(b_alpha_Intercept, b_alpha_year2006, b_alpha_year2010, 
-         b_alpha_year2014, b_alpha_year2018) %>%
-  transmute(
-    fy2002 = b_alpha_Intercept,
-    fy2006 = mean(b_alpha_Intercept) + b_alpha_year2006,
-    fy2010 = mean(b_alpha_Intercept) + b_alpha_year2010,
-    fy2014 = mean(b_alpha_Intercept) + b_alpha_year2014,
-    fy2018 = mean(b_alpha_Intercept) + b_alpha_year2018
-  ) %>%
-  mutate_all(.funs = ~brms::inv_logit_scaled(., 0, 150)*10) -> posteriors
+posterior_alphas <- list()
+for (pathtofile in path_to_fits) {
+  df <- read_rds(pathtofile) %>%
+    as.data.frame() %>%
+    select(b_alpha_Intercept)
+
+  colnames(df) <- paste0("b_alpha_", substring(pathtofile, 34, 37))
+
+  posterior_alphas <- bind_cols(posterior_alphas, df)
+}
+
+posterior_alphas %>%
+  mutate_all(.funs = ~brms::inv_logit_scaled(., 0, 100)*10) -> posterior_omegas
+
+# select posteriors for plotting
+posteriors_selected <- posterior_omegas %>%
+  mutate(b_alpha_2010 = (b_alpha_2010 + b_alpha_2011)/2) %>%
+  select(b_alpha_2002, b_alpha_2006, b_alpha_2010, b_alpha_2014, b_alpha_2018)
 
@@
 mcmc_areas_ridges(
-  posteriors,
-  pars = rev(colnames(posteriors)),
-  prob = 0.90, # 80% intervals
-  prob_outer = .99, # 99%
-  point_est = "mean"
+  posteriors_selected,
+  pars = rev(colnames(posteriors_selected)),
+  prob = 0.95, # 80% intervals
+  prob_outer = .99 # 99%
   ) +
   theme_test() +
-  geom_vline(aes(xintercept = mean(posteriors$fy2002)), color = "gray", linetype = 3) +
-  labs(
-    x = "Household wealth ($000s)")
+  geom_vline(aes(xintercept = median(posteriors_selected$b_alpha_2002)), 
+             color = "gray", 
+             linetype = 3) +
+  labs(x = "Household wealth ($000s)")
 ```
@@
 ```{r, table_2, ft.align="left"}
+# Trim posteriors for 2003 outlier
+posterior_omegas %>%
+  mutate(b_alpha_2003 = (b_alpha_2002 + b_alpha_2004)/2) -> posteriors_trimmed
+
+# Tabulate against median wealth
 happywealth_covs %>%
@@
   mutate(
-    year = paste0("fy", year),
+    year = paste0("b_alpha_", year),
     adj_hifdip = adj_hifdip/1000) %>%
@@
     gather(
-      summarise_all(posteriors, list(median)), year, change_point
+      summarise_all(posteriors_trimmed, list(median)), year, change_point
     )) %>%
+  mutate(
+    year = extract_numeric(year)
+    ) %>%
   group_by(year) %>%
   summarise(
-    `median ($000s)` = round(median(adj_hifdip), 2),
-    `mean ($000s)` = round(mean(adj_hifdip), 2),
-    `change point ($000s)` = round(mean(change_point), 2),
-    `%` = round(mean(adj_hifdip > mean(change_point))*100, 1)
-  ) %>%
+    `Median household wealth ($000s)` = round(median(adj_hifdip), 2),
+    # `mean ($000s)` = round(mean(adj_hifdip), 2),
+    `Cost of happiness ($000s)` = round(mean(change_point), 2),
+    `Percent pop. afford happiness` = round(mean(adj_hifdip > mean(change_point))*100, 1)
+    ) -> results
+
+results %>%
+  filter(year %in% c(2002, 2006, 2010, 2014, 2018)) %>%
   flextable() %>%
@@
 As a result, a smaller percentage of people in 2018 achieved a level of financial security on which their happiness was no longer dependent on their wealth, relative to 2002.
+
+<br>
+
+##### Figure 4.  
+```{r figure_4}
+results %>%
+  gather("key", "value", -year) %>%
+  mutate(
+    key = paste0(key, "\n"),
+    key = fct_relevel(key, "Median household wealth ($000s)\n")) %>%
+  select(key) %>% distinct() %>%
+  mutate(string = c("GFC occurred\nin 2009", "", "")) -> df.note
+
+p <- results %>%
+  gather("key", "value", -year) %>%
+  mutate(
+    key = paste0(key, "\n"),
+    key = fct_relevel(key, "Median household wealth ($000s)\n")) %>%
+  ggplot(aes(x = year, y = value, color = key)) +
+    geom_line(aes(group = key), size = 2, alpha = 0.5) +
+    # geom_smooth(aes(group = key), se = F, span = 2) +
+    labs(title = "Wealth & happiness in Australia between 2002-2018",
+         caption = "Source: HILDA, Melbourne University",
+         x = "  ", y = "") +
+    geom_vline(aes(xintercept = 2009), color = "red", linetype = 3) +
+    geom_text(data = df.note, aes(x = 2009.5, y = 55, label = string),
+              colour = "red", size = 3, hjust = "left") +
+    facet_wrap(~key)
+
+# p.wsj <- p +
+#   ggthemes::scale_colour_wsj("colors6", "") +
+#   ggthemes::theme_wsj() +
+#   theme(legend.position = "none",
+#         plot.title = element_text(size = 19),
+#         plot.caption = element_text(size = 12),
+#         strip.text = element_text(size = 12))
+
+p.econ <- p + 
+  labs(subtitle = "  ") +
+  ggthemes::scale_color_economist() +
+  ggthemes::theme_economist(base_family="Verdana", base_size = 8) +
+  theme(legend.position = "none",
+        strip.text = element_text(size = 10)
+        )
+p.econ
+```
 

### 2020-11-23 src/happyfun.R

@@ -49,8 +49,6 @@ plot_linear <- function(.data, swb_col, dollars_col) {
          y = "") +
     theme_test() +
     theme(
-      legend.background = element_blank(),
-      legend.title = element_blank(),
       legend.position = "none") -> p1
   
   p1 +
@@ -256,3 +254,6 @@ plot_inflection_cubic <- function(.data, swb_col, dollars_col) {
          y = "") +
     theme_test()
 }
+
+
+

### 2020-11-23 src/model_selection.r

@@ -1,0 +1,91 @@
+#### Comparing models ####
+# https://bookdown.org/ajkurz/Statistical_Rethinking_recoded/overfitting-
+# regularization-and-information-criteria.html#the-problem-with-parameters
+# 
+# WAIC is a (penalized) estimate of out-of-sample deviance so smaller values (closer to 
+# zero) are better. Note also that p_waic is the penalty term of WAIC, and so it
+# will be close to the number of dimensions in the posterior of each model  
+
+lm_fit <- readRDS("../results/fit_lm_hifdip_losat_cov_2018.RDS")
+pw_fit <- readRDS("../results/fit_cp_hifdip_losat_cov_2018.RDS")
+
+lm_fit <- add_criterion(lm_fit, "waic")
+pw_fit <- add_criterion(pw_fit, "waic")
+
+# Comparing WAIC using loo_compare()
+# The best model will be reported at the top of the list with elpd_diff = 0
+ans <- loo_compare(lm_fit, pw_fit, criterion = "waic")
+
+print(ans, simplify = F)
+#         elpd_diff se_diff
+# pw_fit   0.0       0.0  
+# lm_fit -69.6      12.2  
+
+waic_results <- tibble()
+
+for (i in c(2002:2018)) {
+  
+  lm_fit <- readRDS(paste0("../results/fit_lm_hifdip_losat_cov_", i, ".RDS"))
+  pw_fit <- readRDS(paste0("../results/fit_cp_hifdip_losat_cov_", i, ".RDS"))
+  
+  lm_fit <- add_criterion(lm_fit, "waic")
+  pw_fit <- add_criterion(pw_fit, "waic")
+
+  saveRDS(lm_fit, paste0("../results/fit_lm_hifdip_losat_cov_", i, ".RDS"))
+  saveRDS(pw_fit, paste0("../results/fit_cp_hifdip_losat_cov_", i, ".RDS"))
+  
+  ans <- loo_compare(lm_fit, pw_fit, criterion = "waic")
+  
+  as_tibble(ans, rownames = "model") %>%
+    mutate(year = i) %>%
+    mutate(across(elpd_diff:year, as.numeric)) -> tb
+  
+  # Add R-squared
+  lm_r2 <- bayes_R2(lm_fit) %>%
+    as_tibble() %>%
+    rename(R2 = Estimate, R2.Error = Est.Error, R2_Q2.5 = Q2.5, R2_Q97.5 = Q97.5) %>%
[…]

### 2020-11-23 src/modeling.r

@@ -1,0 +1,243 @@
+#### The Increasing Cost of Happiness ####
+# RW Morris, N Kettlewell, N Glozier
+# 
+# Model specification and fitting
+# Nov 21st 2020
+# 
+#### libraries ####
+library(dplyr)
+library(ggplot2)
+library(brms)
+options(mc.cores = 4)
+
+#### Import ####
+happywealth_covs <- readRDS("../data/happywealth_covs.rds")
+
+#### Preprocessing ####
+happywealth_covs %>%
+  filter(!student, year == 2002) %>%
+  transmute(
+    y = as.vector(scale(losat)), # gh9
+    dollars = adj_hifdip/10000, 
+    male = as.numeric(male),
+    age = as.vector(scale(age)),
+    age_sq = age^2,
+    grad = as.numeric(edu %in% c("Grad", "Postgrad"))
+  ) %>%
+  na.omit() -> df
+
+# Checking preprocessing...
+df %>%
+  mutate(decile = ntile(dollars, 10)) %>% 
+  group_by(decile) %>%
+  summarise(dollars = mean(dollars), 
+            y = mean(y)) %>%
+  ggplot(aes(x = dollars, y = y)) + geom_point()
+
+summary(df$dollars)
+#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
+#  0.000   2.697   4.176   4.758   6.065  40.311 2002 | gh9
+#  0.000   3.588   5.362   6.173   7.706  85.139 2018 | gh9
+
+summary(df$y)
+#  Min.     1st Qu.  Median  Mean    3rd Qu. Max. 
+#  -4.0191  -0.6054  0.1824  0.0000  0.7076  1.8893  2002 | gh9
+#  -4.90231 -0.54261 0.08021 0.00000 0.70302 1.32584 2002 | losat
+
+hist(df$dollars, breaks = 100)
+hist(df$y, breaks = 10)
+
+#### Fit the model ####
+#
+# This model was suggested at: 
+# https://discourse.mc-stan.org/t/piecewise-linear-mixed-models-with-a-random-change-point/5306/5
+# 
+# Model definition
+# b0 is the intercept denoting the expected value of wellbeing (y) at the changepoint
+# b1 is the linear slope before the changepoint
+# b2 is the linear slope after the changepoint
+# alpha is the coefficient of wealth (x) at the changepoint
+
+bfixed <- bf(
+  y ~ b0 + b1 * (dollars - omega) * step(omega - dollars) + 
[…]

### 2020-11-23 src/piecewise_model_simulation.r

@@ -1,0 +1,69 @@
+tibble(
+  year = rep(1:16, each = 20),
+  dollars = rep(1:10, 32)
+  ) %>% 
+  rowwise() %>%
+  mutate(
+    dollars = dollars + rnorm(1),
+    b0 = 8,
+    b1 = 2,
+    b2 = 0.05,
+    omega = 5, #ceiling((year + 2)/2.5),
+    y = b0 + b1*(dollars - omega)*(omega > dollars) + b2*(dollars - omega)*(dollars >= omega) + rnorm(1),
+    age = sample(18:100, 1),
+    male = sample(0:1, 1),
+    grad = rbinom(1, 1, 0.25)) %>%
+  ungroup() %>%
+  mutate(
+    y = as.vector(scale(y)),
+    age = as.vector(scale(age)),
+    agesq = age^2
+    ) -> df
+
+ggplot(df, aes(x = dollars, y = y)) +
+  geom_point() +
+  facet_wrap(~year)
+
+hist(df$wealth)
+hist(df$y)
+summary(df$dollars)
+#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
+# -2.094   3.038   5.559   5.478   7.930  12.232 
+
+bfixed <- bf(
+  y ~ b0 + b1 * (dollars - omega) * step(omega - dollars) + 
+    b2 * (dollars - omega) * step(dollars - omega) + b3*age + b4*agesq + b5*male + b6*grad,
+  b0 + b1 + b2 + b3 + b4 + b5 + b6 + alpha ~ 1, #+ (1|male + grad),  # fixed intercept + (random intercept per sex and grad)
+  # Sigmoid transform to keep omega within the range of 0 to x
+  nlf(omega ~ inv_logit(alpha) * 15),
+  nl = TRUE
+)
+
+# priors                                      I tried tighter priors:
+bprior <- prior(normal(0, 3), nlpar = "b0") + # normal(0, 1) did not converge Sep 6th
+  prior(normal(0, 3), nlpar = "b1") +         # normal(0, 1) did not converge Sep 6th
+  prior(normal(0, 3), nlpar = "b2") +         # normal(0, 1) did not converge Sep 6th
+  prior(normal(0, 3), nlpar = "b3") +
+  prior(normal(0, 3), nlpar = "b4") +
+  prior(normal(0, 3), nlpar = "b5") +
+  prior(normal(0, 3), nlpar = "b6") +
[…]

### 2020-11-23 src/preprocessing.r

@@ -1,5 +1,11 @@
-#### Preprocessing ####
-
+#### The Increasing Cost of Happiness ####
+# RW Morris, N Kettlewell, N Glozier
+# 
+# Preprocessing 
+# Nov 21st 2020
+# 
+# Libraries
+library(tidyverse)
 library(haven)
 
 #### Load data ####
@@ -19,7 +25,7 @@ for (pathtofile in path_to_hilda) {
 # Helper function
 source('../../src/gather_hilda.R')
 
-# demographic data 
+# DEMOGRAPHIC VARIABLES 
 gather_hilda(hilda, c(
   "esbrd",   # employment status (broad)
   "hhda10",  # SEIFA 2001 Decile of socio-economic advantage (higher is better)
@@ -35,22 +41,15 @@ gather_hilda(hilda, c(
   spread(code, val) %>%
   mutate_if(is.double, ~ ifelse(. < 0, NA_real_, .)) -> demographic_items
 
-# financial data
+# MODEL VARIABLES (household wealth, satisfaction, happiness):
+# Note that regular HILDA income variables were used (not total income, which 
+# include irregular income) to match with the ABS, and because more consistent
+# prior to 2012. See Appendix in Wilkins (2014) 'Derived income variables in the
+# HILDA survey' for explanation
+# 
 gather_hilda(hilda, c(
-  "wsfei",   # financial year gross
-  "tifdip",  # Disposable regular income
-  "tifeftp", # Gross total income (including benefits, pensions, foreign, etc)
-  "hiwsfei", # household financial year gross
-  "hwnwip",  # household net worth
-  "hifdip",  # household disposable regular income
-  "hifditp"  # household disposable total income
-)) %>%
-  spread(code, val) %>%
-  mutate_if(is.double, ~ ifelse(. < 0, NA_real_, .)) -> wealth_items
-
-# Affective wellbeing:
-gather_hilda(hilda, c(
-  'losat',
+  "hifdip", # household disposable regular income
+  'losat', # Life satisfaction
   'gh9a', # Vitality: feel full of life (lower is better)*
   'gh9b', # Mental H: Been a nervous person (higher is better)
   'gh9c', # Mental H: Felt so down in the dumps (higher is better)
@@ -63,11 +62,11 @@ gather_hilda(hilda, c(
 )) %>%
   # Recode missing to NA
   spread(code, val) %>%
[…]

### 2020-12-02 src/draft0.Rmd
fixed some shit

@@
 While the location of this change point has been determined, and the cost of happiness reportedly ranges between USD\$60 to $95K, there has been no investigation as to whether the cost of happiness has increased or decreased over time.
-We tested the relationship between household wealth and both happiness and life satisfaction using household economic data from Australia between 2002-2016.
-We discovered the cost of happiness has increased over those 16 years faster than inflation (i.e., cost of living).
+We tested the relationship between wealth and both happiness and life satisfaction using household economic data from Australia between 2002-2018.
+We discovered the cost of happiness has increased over those 18 years faster than inflation (i.e., cost of living).
 This result shows that inequalities in wealth may be driving inequities in happiness and wellbeing, with implications for health and recent government policy-goals to monitor and improve wellbeing.
@@
 
-We used household economic panel data from Australia (HILDA) to provide the first investigation of whether changes in wealth and wellbeing have changed the cost of happiness over the last 16 years (2002-2018).
+We used household economic panel data from Australia (HILDA) to provide the first investigation of whether changes in wealth and wellbeing have changed the cost of happiness over the last 18 years (2002-2018).
 HILDA provides a representative sample of households in Australia with detailed measurements of wealth and subjective wellbeing in the same sample, which makes it an excellent data source to investigate the present question.
@@
 #### Wealth
-Our indicator of wealth and economic security was household after-tax income.
+Our indicator of wealth and economic security was household income.
 Household income better represents economic security, since members of the same household share expenses as well as risks; i.e., they can provide a direct and immediate support network when financial shocks occur.
 The other major studies also used household after-tax income as the indicator of wealth and economic security [e..g, @kahneman2010high; @jebb2018happiness], and so we follow the same standard here as well.
-Household financial year disposable regular income [(hifdip)](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=hifdip){target="_blank"} from the annual HILDA survey in 2002, 2006, 2010, 2014,and 2018 was selected.
-This variable in HILDA represents the sum across all household members of financial year gross regular income less taxes on financial year gross regular income.
+The ‘real household annual disposable income’ was calculated from the combined income of all household members after receipt of government pensions and benefits and deduction of income taxes in the financial year ended 30th June of the year of the wave (e.g., 2002 in wave 2).
+This was then adjusted for inflation - the rise in the general price level of the economy - using the Australian Bureau of Statistics (ABS) Consumer Price Index, so that income in all waves is expressed in FY 2017/18 prices, to give real income.
 
-Household income was provided by the University of Melbourne as imputed, and with weighted topcodes to preserve anonymity.
-We removed individuals with values above the threshold indicated for top-coding.
-Household wealth was adjusted for the number of people in each household by dividing by the square root of household size [@headey2004effects; @jebb2018happiness; @kahneman2010high].
-We converted dollar values to the base year (2002) according to CPI from the RBA, in order to adjust for inflation (https://www.rba.gov.au/calculator/annualDecimal.html).
+The _equivalised_ household income was obtained by adjusting for household size (the number of adult and child household members).
+In this instance, we have used the 'modified OECD' scale (Hagenaars et al., 1994), which divides household income by 1 for the first household member plus 0.5 for each other household member aged 15 or over, plus 0.3 for each child under 15.
+A family comprising two adults and two children under 15 years of age would therefore have an equivalence scale of 2.1 (1 + 0.5 + 0.3 + 0.3), meaning that the family would need to have an income 2.1 times that of a single-person household in order to achieve the same standard of living.
+This scale recognises that larger households require more income, but it also recognises that there are economies of scale in consumption and that children require less than adults.
+The equivalised income calculated for a household is then assigned to each member of the household.
 
@@
 There are a variety of variables related to subjective well-being collected annually in HILDA, but the two we used here matched the variables we used in our previous paper [@kettlewell2020differential], namely, life satisfaction as a measure of cognitive wellbeing, and the SF-36 as a measure of affective wellbeing or happiness.
-Life satisfaction [(losat)](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=losat){target="_blank"} was assessed by a single item question asked each survey: "How satisfied are you with your life (0 to 10)".
+
+Life satisfaction [(losat)](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=losat){target="_blank"} was assessed by a single item question asked each survey: "All things considered, how satisfied are you with your life (0 to 10)".
+
 Happiness was determined by item 9 in the SF-36 [(gh9a to gh9i)](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=gh9a){target="_blank"}.
@@
 We reverse scored the relevant responses and calculated the sum of the nine questions so that higher scores represented better wellbeing.
+To aid interpretability, we rescaled the final sum to a score between 1-100, where 100 represents the maximum happiness achievable.
+
+For modelling, both dependent variables were rescaled with a mean of zero and a SD of 1 (z-scores) for each year.
 
@@
 
-Where $X_i$ were covariates for age, age^2^, sex and education.
+Where $x_i$ was an individual's household wealth, and $X_i$ were covariates for age, age^2^, sex and education.
 
@@
 
-The above models estimated population-level (fixed) effects for each year in the data (*t* = 2002, 2006, 2010, 2014 and 2018).
+The above models estimated population-level effects separately for each year (*t* = 2002, 2006, 2010, 2014 and 2018).
 Because we were interested in the location of the change point between wealth and wellbeing that existed across individuals within each year, we ignored the panel design of HILDA because the dependency between observations of the same person across years was orthogonal to our effects of interest.
-We assumed an unstructured variance-covariance matrix between years, and specified weakly informed priors for each *&beta;*, and a uniform prior over the restricted range of wealth values for *&omega;*.
+We specified weakly informed priors for each *&beta;*, and a uniform prior over the restricted range of wealth values for *&omega;*.
 
@@
 _Covariates_  
-Age (and age^2^), gender, and education were included as covariates, and fulltime students were removed.
+Age (and age^2^), gender, and education were included as covariates.
+Full-time students were removed, as well as individuals with an annual household disposable income that was indicated as _topcoded_ by the University of Melbourne (topcoding occurs to ensure privacy of high wealth individuals).
 Gender was included as a binary variable (Male = 1), and education was a binary variable coded from the highest level of education achieved (university/college graduate = 1).
@@
 ```{r read_write_covs}
-happywealth_covs <- read_rds("../data/happywealth_covs.rds")
+happywealth <- read_rds("../data/happywealth.rds")
 ```
@@
 
-The broad demographic characteristics of the sample from each year are presented below in Table 1.
+The broad demographic characteristics of the sample over the time period are presented below in Table 1 (at four year intervals).
 
@@
 
-happywealth_covs %>%
+happywealth %>%
   filter(year %in% c(2002, 2006, 2010, 2014, 2018)) %>%
-  filter(!student, (!is.na(losat)|!is.na(gh9))) %>%
+  filter(!student, !top_hifdip, (!is.na(losat)|!is.na(gh9))) %>%
   mutate(
     Sex = if_else(male, "Male", "Female"),
-    SEIFA = as.integer(extract_numeric(SEIFA)),
-    SEIFA = if_else(SEIFA > 0, SEIFA, NA_integer_),
     Education = fct_collapse(
       edu,
-      `High school or less` = c("None", "In school", "Highschool", "Trade"),
-      Graduate = c("Grad", "Postgrad")
+      `High school or less` = c("None", "In school", "Highschool"),
+      Graduate = c("Grad")
       ),
@@
     `Life satisfaction` = losat, `Happiness` = gh9,
-    # `Household wealth ($)` = adj_hifdip, 
     Sex, `Age (years)` = age, Education, Workforce, `Relationship status`, 
@@
     statistic = 
-      list(all_continuous() ~ "{median} ({p25}-{p75})",
+      list(all_continuous() ~ "{mean} (±{sd})",
            all_categorical() ~ "{n} ({p}%)"),
-    digits = all_continuous() ~ 0,
+    digits = all_continuous() ~ 1,
     missing_text = "(Missing)" 
@@
 
+Average life satisfaction levels were very steady between 2002-2018, while average happiness score decreased slightly over the 18 years.
 The proportions of each sex and relationship status were stable over time, as were the average household size and SEIFA index.
-However age and chronic health conditions tended to increase over time, as expected in an aging longitudinal sample.
+However age and chronic health conditions tended to increase over time, as expected in an aging sample.
 Education levels (i.e., graduate percentage) also tended to increase over time, while changes in the workforce varied both up and down.
@@
 
-Below we show the relationship between household wealth and wellbeing (Figure 1).
-For visualization purposes only, due to the large numbers of individual data in each year, we display the mean levels of wealth and wellbeing for each (equal-sized) wealth decile rather than every individual data point.
+Below we show the relationship between household income and wellbeing (Figure 1).
+For visualization purposes only, due to the large numbers of individual data in each year, we display the mean levels of income and wellbeing for each (equal-sized) income decile rather than every individual data point.
 Note that the line-of-best-fit and 95% credible intervals (shaded) from each regression model _of all individuals_ is shown in overlay.
@@
 
-##### Figure 1. Household wealth and satisfaction (blue) and happiness (red)
+##### Figure 1. Household income and satisfaction (blue) and happiness (red)
 ```{r losat_gh9_hifdip}
@@
 
-p1 <- happywealth_covs %>%
-  rename(Satisfaction = losat, `Household wealth` = adj_hifdip) %>% 
-  filter(!student) %>%
-  plot_linear(Satisfaction, `Household wealth`) +
+p1 <- happywealth %>%
+  rename(Satisfaction = losat, `Household income` = hh_disp_inc_eq) %>% 
+  filter(!student, !top_hifdip) %>%
+  plot_linear(Satisfaction, `Household income`) +
   scale_fill_manual(values = "#b3cde0") +
@@
 
-p2 <- happywealth_covs %>%
-  rename(Satisfaction = losat, `Household wealth` = adj_hifdip) %>% 
-  filter(!student) %>%
-  plot_inflection_piecewise(Satisfaction, `Household wealth`) +
+p2 <- happywealth %>%
+  rename(Satisfaction = losat, `Household income` = hh_disp_inc_eq) %>% 
+  filter(!student, !top_hifdip) %>%
+  plot_inflection_piecewise(Satisfaction, `Household income`) +
   scale_fill_manual(values = "#b3cde0") +
@@
 
-p3 <- happywealth_covs %>%
-  rename(Happiness = gh9, `Household wealth` = adj_hifdip) %>% 
-  filter(!student) %>%
-  plot_linear(Happiness, `Household wealth`) +
+p3 <- happywealth %>%
+  rename(Happiness = gh9, `Household income` = hh_disp_inc_eq) %>% 
+  filter(!student, !top_hifdip) %>%
+  plot_linear(Happiness, `Household income`) +
   scale_fill_manual(values = "#C79999") +
@@
 
-p4 <- happywealth_covs %>%
-  rename(Happiness = gh9, `Household wealth` = adj_hifdip) %>% 
-  filter(!student) %>%
-  plot_inflection_custom(Happiness, `Household wealth`, 
+p4 <- happywealth %>%
+  rename(Happiness = gh9, `Household income` = hh_disp_inc_eq) %>% 
+  filter(!student, !top_hifdip) %>%
+  plot_inflection_custom(Happiness, `Household income`, 
                          knots = c(43.7, 49.2, 47.4, 53.2, 72.3)) +
@@
 ```
-Figure 1 legend: Posterior predictive distributions (95% credible intervals) showing the model fit from regressions of wellbeing on household wealth, overlaid on summary data points (wealth deciles).
+Figure 1 legend: Posterior predictive distributions (95% credible intervals) showing the model fit from regressions of wellbeing on household income, overlaid on summary data points (income deciles).
 Wellbeing was measured as life satisfaction (blue) or happiness (red).
@@
 
-Figure 1 shows the relationship between household wealth and satisfaction appeared relatively linear, as the results of the piecewise regression indicated the inflection of the change point could be increasing (e.g., 2002, 2006) as well as decreasing (e.g., 2010, 2014, 2018).
-By contrast, a decreasing change point was evident in each year of happiness on wealth.
+Figure 1 shows the relationship between household income and satisfaction appeared relatively linear, as the results of the piecewise regression indicated the inflection of the change point could be increasing (e.g., 2002, 2006) as well as decreasing (e.g., 2010, 2014, 2018).
+By contrast, a decreasing change point was evident in each year of happiness on income.
 The results of a formal comparison between linear and piecewise models is provided in Model Selection below.
-However note that the change point between wealth and happiness appeared to shift rightwards over time (bottom row).
-(We can also see the linear relationship between household wealth and each wellbeing variable became slightly weaker over time, i.e., less steep)
+However note that the change point between income and happiness appeared to shift rightwards over time (bottom row).
+(We can also see the linear relationship between household income and each wellbeing variable became slightly weaker over time, i.e., less steep)
 
@@
 
-We compared the posterior evidence for a linear relationship between wellbeing and wealth with the posterior evidence of a nonlinear (piecewise) relationship (i.e., WAIC~linear~ — WAIC~piecewise~).
+We compared the posterior evidence for a linear relationship between wellbeing and income with the posterior evidence of a nonlinear (piecewise) relationship (i.e., WAIC~linear~ — WAIC~piecewise~).
 Thus a WAIC difference greater than zero indicated evidence for a linear relationship.
@@
 ##### Figure 2. Linear model evidence (WAIC~linear~ — WAIC~piecewise~)
+
+```{r plot_waic_gh9}
+waic_result <- read_rds("../results/model_comparisons_waic_R2_gh9.RDS")
+
+waic_result %>%
+  select(model, year, elpd_diff, se_diff) %>%
+  filter(elpd_diff != 0) %>%
+  filter(year %in% c(2002, 2006, 2010, 2014, 2018)) %>%
+  rowwise() %>%
+  mutate(samples = list(rnorm(n = 1e5, 
+                              mean = elpd_diff * 2, # for pw advantage
+                              sd   = se_diff *  2)),
+         year = paste0("AD", year)
+         ) %>%
+  select(year, samples) %>%
+  group_by(year) %>%
+  unnest(samples) %>% 
+  mutate(sample_id = c(1:1e5)) %>%
+  pivot_wider(id_col = "sample_id", names_from = "year", values_from = "samples") %>%
+  select(-sample_id) -> d2
+
+color_scheme_set("red")
+
+p5 <- mcmc_intervals(x = d2,
+           pars = rev(colnames(d2)),
+           prob = 0.00,
+           prob_outer = 0.95
+           ) +
+  scale_x_continuous(
+    limits = c(-200, 150)
+    ) +
+  geom_vline(aes(xintercept = 0), color = "grey90") +
+  theme_test() +
+  labs(
+    subtitle = "Happiness")
+```
 
@@
   mutate(samples = list(rnorm(n = 1e5, 
-                              mean = elpd_diff * -2, 
+                              mean = elpd_diff * -2, # for linear advantage
                               sd   = se_diff *  2)),
@@
 
-p5 <- mcmc_intervals(x = d1,
+p6 <- mcmc_intervals(x = d1,
   pars = colnames(d1),
@@
     subtitle = "Satisfaction",
-    x = "")
-```
-
-```{r plot_waic_gh9}
-waic_result <- read_rds("../results/model_comparisons_waic_R2_gh9.RDS")
-
-waic_result %>%
-  select(model, year, elpd_diff, se_diff) %>%
-  filter(elpd_diff != 0) %>%
-  filter(year %in% c(2002, 2006, 2010, 2014, 2018)) %>%
-  rowwise() %>%
-  mutate(samples = list(rnorm(n = 1e5, 
-                              mean = elpd_diff * 2, 
-                              sd   = se_diff *  2)),
-         year = paste0("AD", year)
-         ) %>%
-  select(year, samples) %>%
-  group_by(year) %>%
-  unnest(samples) %>% 
-  mutate(sample_id = c(1:1e5)) %>%
-  pivot_wider(id_col = "sample_id", names_from = "year", values_from = "samples") %>%
-  select(-sample_id) -> d2
-
-color_scheme_set("red")
-
-p6 <- mcmc_intervals(x = d2,
-           pars = rev(colnames(d2)),
-           prob = 0.00,
-           prob_outer = 0.95
-           ) +
-  scale_x_continuous(
-    limits = c(-200, 150)
-    ) +
-  geom_vline(aes(xintercept = 0), color = "grey90") +
-  theme_test() +
-  labs(
-    subtitle = "Happiness",
     x = "WAIC difference (95% interval)")
@@
 
-Model selection revealed the posterior evidence for the linear fit of satisfaction on wealth was credibly superior to the nonlinear (piecewise) fit - Figure 2 shows the 90% credible interval of the fit for each year of satisfaction was above zero with no overlap.
-By contrast, a nonlinear fit of happiness on wealth was generally superior, and the nonlinear fit was credibly superior at 90% for each of three years (2006, 2014, 2018).
-Thus, the posterior evidence indicates happiness and satisfaction have distinct relationships with household wealth; satisfaction tends to increase linearly with wealth, while a change point exists in the relationship between happiness wealth.
+Model selection revealed the posterior evidence for the linear fit of satisfaction on income was credibly superior to the nonlinear (piecewise) fit - Figure 2 shows the 90% credible interval of the fit for each year of satisfaction was above zero with no overlap.
+By contrast, a nonlinear fit of happiness on income was generally superior, and the nonlinear fit was credibly superior at 90% for each of three years (2006, 2014, 2018).
+Thus, the posterior evidence indicates happiness and satisfaction have distinct relationships with household income; satisfaction tends to increase linearly with income, while a change point exists in the relationship between happiness and income.
 
@@
 
-We present the complete posterior distribution of &omega; over 4000 samples drawn from the piecewise model, representing the location of the change point between happiness and household wealth for each year.
-Shaded areas to the right of the vertical grey dotted line are credibly (90%) larger than the base year (2002).
+Figure 3 below presents the complete posterior distribution of &omega; over 4000 samples drawn from the piecewise model, representing the location of the change point between happiness and household income in four separate years.
+Shaded areas to the right of the vertical grey dotted line are credibly (95%) larger than 2002.
+The figure shows the change point between happiness and household income (the _cost of happiness_) credibly increased between 2002 and 2018.
 
@@
 path_to_fits <- list.files(
-  path = "../results",#'~/Dropbox/HILDA/data',
+  path = "../results",
   pattern = '^fit_cp_hifdip_gh9_cov_.*.RDS$',
@@
 
-  colnames(df) <- paste0("b_alpha_", substring(pathtofile, 34, 37))
+  colnames(df) <- paste0("param_", substring(pathtofile, 34, 37))
 
@@
 
+posteriors_trim <- posterior_omegas %>%
+  mutate(
+    param_2010 = (param_2010 + param_2011)/2,
+    param_2003 = (param_2002 + param_2004)/2)
+
 # select posteriors for plotting
-posteriors_selected <- posterior_omegas %>%
-  mutate(b_alpha_2010 = (b_alpha_2010 + b_alpha_2011)/2) %>%
-  select(b_alpha_2002, b_alpha_2006, b_alpha_2010, b_alpha_2014, b_alpha_2018)
+posteriors_select <- posterior_omegas %>%
+  select(param_2002, param_2006, param_2010, param_2014, param_2018)
 
@@
 mcmc_areas_ridges(
-  posteriors_selected,
-  pars = rev(colnames(posteriors_selected)),
-  prob = 0.95, # 80% intervals
-  prob_outer = .99 # 99%
+  posteriors_select,
+  pars = rev(colnames(posteriors_select)),
+  prob = 0.95, 
+  prob_outer = .99 
   ) +
   theme_test() +
-  geom_vline(aes(xintercept = median(posteriors_selected$b_alpha_2002)), 
+  geom_vline(aes(xintercept = median(posteriors_select$param_2002)), 
              color = "gray", 
              linetype = 3) +
-  labs(x = "Household wealth ($000s)")
+  labs(x = "Household income ($000s)")
 ```
@@
 ```
-Figure 3 legend: Posterior distribution of the change point parameter representing the location in household wealth ($) for each year. Shaded area represents the 90% credible region.
+Figure 3 legend: Posterior distribution of the change point parameter representing the location in household income ($) for each year. Shaded area represents the 90% credible region.
 ```
@@
 
-Figure 3 shows the change point between happiness and household wealth (the _cost of happiness_) credibly increased between 2002 and 2018.
-Table 2 below compares the change in median and mean wealth for each year with the cost of happiness.
-It presents the percentage of people in each year whose household wealth exceeded the cost point of happiness.
+#### Wealth and happiness
+
+##### Figure 4. Estimated happiness (SD units) at $50K and $80K from 2002-2018 
+```{r figure_4}
+predicted_fits <- list()
+for (pathtofile in path_to_fits) {
+  fit <- read_rds(pathtofile)
+
+  ans <- model.frame(fit) %>%
+    summarise_all(median)
+
+  newdat = data.frame(
+    dollars = c(5, ans[, "dollars"], 8),
+    age = rep(ans[, "age"], 3),
+    age_sq = rep(ans[, "age_sq"], 3),
+    male = rep(ans[, "male"], 3), 
+    grad = rep(ans[, "grad"], 3)
+  )
+
+  df <- fitted(fit, newdata = newdat) %>% 
+    as_tibble() %>%
+    mutate(
+      Dollars = c("$50K", "\u03bc", "$80K"),
+      year = as.numeric(substring(pathtofile, 34, 37)))
+
+  predicted_fits <- bind_rows(predicted_fits, df)
+}
+
+predicted_fits %>%
+  select(Estimate, Dollars, year) %>%
+  filter(Dollars %in% c("$80K", "$50K")) %>%
+  spread(Dollars, Estimate) %>%
+  mutate(`$80K - $50K (∆)` = `$80K` - `$50K`) %>%
+  gather(income, happiness, -year) %>%
+  mutate(income = fct_relevel(income, "$50K", "$80K")) %>%
+  ggplot(aes(x = year, y = happiness, color = income)) +
+  geom_point(size = 4, alpha = 0.25) +
+  geom_smooth(method = "loess", se = F, span = 3) +
+  facet_wrap(~income) +
+  labs(caption = "Source: Household disposable income, HILDA, Melbourne University", 
+       x = "", y = "") -> p
+
+source("ft_themes.r")
+
+p +
+  ft_theme() +
+  ggthemes::scale_colour_wsj() +
+  theme(legend.position = "none",
+        plot.title = element_text(size = 19),
+        strip.text = element_text(size = 12))
+```
+
+```
+Figure 4 legend: The difference (∆) in happiness between a household earning $50K per year and a household earning $80K per year has increased since 2009.
+```
 
@@
 
-```{r, table_2, ft.align="left"}
-# Trim posteriors for 2003 outlier
-posterior_omegas %>%
-  mutate(b_alpha_2003 = (b_alpha_2002 + b_alpha_2004)/2) -> posteriors_trimmed
+Figure 5 presents house income levels adjusted for the Australian population (by age, sex, marital status, labour force participation and region).
+It shows the cost of happiness increased faster than household income between 2002 and 2018.
+As a result, a smaller percentage of the Australian population in 2018 achieved a level of financial security on which their happiness was no longer dependent on their income, relative to 2002.
 
-# Tabulate against median wealth
-happywealth_covs %>%
-  select(year, xwaveid, adj_hifdip) %>%
+<br>
+
+##### Figure 5.  
+```{r figure_5, fig.width=7, fig.height=6}
+# Tabulate against median income
+happywealth %>%
+  filter(!student) %>%
+  filter(!top_hifdip) %>%
+  select(year, xwaveid, income = hh_disp_inc_eq) %>%
+  filter(year %in% 2002:2018) %>%
   mutate(
-    year = paste0("b_alpha_", year),
-    adj_hifdip = adj_hifdip/1000) %>%
+    year = paste0("param_", year),
+    income = income/1000) %>%
   left_join(
     gather(
-      summarise_all(posteriors_trimmed, list(median)), year, change_point
+      summarise_all(posteriors_trim, 
+                    list(median)), 
+      year, 
+      change_point
     )) %>%
@@
   summarise(
-    `Median household wealth ($000s)` = round(median(adj_hifdip), 2),
-    # `mean ($000s)` = round(mean(adj_hifdip), 2),
-    `Cost of happiness ($000s)` = round(mean(change_point), 2),
-    `Percent pop. afford happiness` = round(mean(adj_hifdip > mean(change_point))*100, 1)
-    ) -> results
-
-results %>%
-  filter(year %in% c(2002, 2006, 2010, 2014, 2018)) %>%
-  flextable() %>%
-  set_caption("Table 2. Household wealth, change points and percentile") %>%
-  autofit()
-```
-
-<br>
-
-Table 2 shows the cost of happiness increased faster than the median or mean household wealth between 2002 and 2018.
-As a result, a smaller percentage of people in 2018 achieved a level of financial security on which their happiness was no longer dependent on their wealth, relative to 2002.
-
-<br>
-
-##### Figure 4.  
-```{r figure_4}
-results %>%
+    `Median household income\n ($000s)` = round(median(income), 
+                                                2),
+    `Cost of happiness\n ($000s)` = round(mean(change_point), 
+                                          2),
+    `Individuals affording happiness\n (Population %)` = round(mean(income > mean(change_point))
+                                                           *100,
+                                                           1)
+    ) %>%
   gather("key", "value", -year) %>%
@@
     key = paste0(key, "\n"),
-    key = fct_relevel(key, "Median household wealth ($000s)\n")) %>%
+    key = fct_relevel(key, "Median household income\n ($000s)\n")) -> df.plot
+
+df.plot %>%
   select(key) %>% distinct() %>%
-  mutate(string = c("GFC occurred\nin 2009", "", "")) -> df.note
+  mutate(string = c("post-GFC\n(2009 onwards)", "", "")) -> gfc.note
 
-p <- results %>%
-  gather("key", "value", -year) %>%
+df.plot %>%
+  filter(year %in% c(min(year), 2018)) %>%
   mutate(
-    key = paste0(key, "\n"),
-    key = fct_relevel(key, "Median household wealth ($000s)\n")) %>%
-  ggplot(aes(x = year, y = value, color = key)) +
+    val = as.character(
+      if_else(str_detect(key, "afford"),
+              paste0(round(value), "%"),
+              paste0("$", round(value), "K"))),
+    vpos = if_else(val %in% c("$49K", "$72K", "41%"), -0.75, 1.65),
+    hpos = if_else(year == min(year), 0.3, 0.8)
+    ) -> df.label
+
+ggplot(df.plot, aes(x = year, y = value, color = key)) +
+  # geom_point() +
   geom_line(aes(group = key), size = 2, alpha = 0.5) +
-    # geom_smooth(aes(group = key), se = F, span = 2) +
+  # geom_smooth(aes(group = key), se = F, span = .5) +
   labs(title = "Wealth & happiness in Australia between 2002-2018",
-         caption = "Source: HILDA, Melbourne University",
+       caption = "Source: HILDA sample weighted for Australian population
+                  Household equivalised income per person (modified OECD scale)
+                  Excluding full-time students and topcoded individuals",
        x = "  ", y = "") +
   geom_vline(aes(xintercept = 2009), color = "red", linetype = 3) +
-    geom_text(data = df.note, aes(x = 2009.5, y = 55, label = string),
+  geom_text(data = gfc.note, aes(x = 2009.5, y = 57, label = string),
             colour = "red", size = 3, hjust = "left") +
-    facet_wrap(~key)
-
-# p.wsj <- p +
-#   ggthemes::scale_colour_wsj("colors6", "") +
-#   ggthemes::theme_wsj() +
-#   theme(legend.position = "none",
-#         plot.title = element_text(size = 19),
-#         plot.caption = element_text(size = 12),
-#         strip.text = element_text(size = 12))
+  geom_text(data = df.label,
+            aes(label = val, vjust = vpos, hjust = hpos)) +
+  expand_limits(y = c(22, 73)) +
+  facet_wrap(~key) -> p
 
@@
 
+```
+Figure 5 legend: Household income has stagnated in Australia since the GFC while the cost of happiness has increased.
+Consequently fewer Australians have a household income exceeding the cost of happiness in 2018.
+```
+
 <br><br>
@@
 
-We found the relationship between subjective wellbeing and household wealth was positive, but happiness and satisfaction had different (positive) relationships: Satisfaction increased linearly with wealth, while happiness increased rapidly up to a point after which further increments in wealth produced less change – this confirms the distinct effects of wealth in previous research [e.g., @howell2008relation; @kahneman2010high], and contributes rare evidence from the same sample.
-Furthermore, we report here for the first time that the change point between household wealth and happiness increased over time between 2002 and 2018, faster than inflation or the median household income.
+We found the relationship between subjective wellbeing and household income was positive, but happiness and satisfaction had different (positive) relationships: Satisfaction increased linearly with income, while happiness increased rapidly up to a point after which further increments in income produced less change – this confirms the distinct effects of wealth in previous research [e.g., @howell2008relation; @kahneman2010high], and contributes rare evidence from the same sample.
+Furthermore, we report here for the first time that the change point between household income and happiness increased over time between 2002 and 2018, faster than inflation or the median household income.
 
 We refer to the change point after which increases in wealth no longer produce similar increases in happiness as the cost of happiness.
-After this point, happiness is no longer as dependent on household wealth, and the economic security it represents.
+After this point, happiness is no longer as dependent on household income, and the economic security it represents.
 Presumably after this point further increases in happiness depend on other life factors (e.g., social connections).
-Life satisfaction on the other hand always increased with household wealth.
+Life satisfaction on the other hand always increased with household income.
 The difference likely reflects the importance of a numerical dollar value (e.g., bank balance, house value) when cognitively appraising one’s life achievements, versus the relevance of that number to our everyday experience of joy and prevailing mood.
@@
 
-Some recent studies have challenged either the notion that the positive effect of household wealth plateaus, or that the effect on happiness and satisfaction are distinct.
-In a nationally representative sample of 44,000 adult Americans in the General Social Survey (GSS), happiness continued to increase with household wealth, implying no change point existed between money and happiness [@twenge2020expanding].
+Some recent studies have challenged either the notion that the positive effect of household income plateaus, or that the effect on happiness and satisfaction are distinct.
+In a nationally representative sample of 44,000 adult Americans in the General Social Survey (GSS), happiness continued to increase with household income, implying no change point existed between money and happiness [@twenge2020expanding].
 The GSS asks a single item on happiness: “Taken all together, how would you say things are these days – would you say that you are very happy, pretty happy, or not too happy?” The form of this question is quite similar to the single item “life satisfaction” question in HILDA, as both request the respondent to cognitively evaluate their circumstances.
@@
 
-A second recent report from a survey of 1.7 million people representing 164 countries in the Gallup World Poll reported increases in household wealth were associated with change points in happiness and satisfaction [@jebb2018happiness], rather than distinct effects as we found.
+A second recent report from a survey of 1.7 million people representing 164 countries in the Gallup World Poll reported increases in household income were associated with change points in happiness and satisfaction [@jebb2018happiness], rather than distinct effects as we found.
 It may be that differences in the sparsity of the high income data in HILDA (for which extremely high income households are masked) explain the differences here.
@@
 
-As governments and policy-makers begin to focus on wellbeing, it will be critical to understand how traditional economic indicators such as household wealth, income inequality, and consumption interact with wellbeing and health.
+As governments and policy-makers begin to focus on wellbeing, it will be critical to understand how traditional economic indicators such as household income, wealth inequality, and consumption interact with wellbeing and health.
 According to @frijters2020happy, coming up with a consensus to translate income into wellbeing features high on the wider wellbeing research agenda.
-Establishing the links between income, household wealth, wellbeing and health, and how inequalities in one drives inequities in the other, will be a critical step in that agenda.
+Establishing the links between wealth, household income, wellbeing and health, and how inequalities in one drives inequities in the other, will be a critical step in that agenda.
 

### 2020-12-02 src/supplementary.Rmd
fixed some shit

---
title: "Supplementary material: The increasing cost of happiness"
author: "R.W. Morris^1,2^, N. Kettlewell^2,3,4^  & N.Glozier^1,5^"
bibliography: ../src/references.bib
csl: ../src/elsevier-harvard.csl
output:  
  html_document:
    reference_docx: ../src/academic_style.docx
always_allow_html: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(flextable)
library(bayesplot)
library(patchwork)

knitr::opts_chunk$set(eval = TRUE,
                      echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      fig.path = "../figures/")
```
\  
\

1. Central Clinical School, Faculty of Medicine and Health, University of Sydney, NSW, Australia
2. ARC Centre of Excellence for Children and Families over the Life Course, University of Queensland, QLD, Australia
3. School of Economics, University of Technology, NSW, Australia
4. Institute of Labor Economics (IZA), Bonn, Germany
5. Brain and Mind Centre, University of Sydney, NSW, Australia

\

**Corresponding author:**  
Professor Nick Glozier  
Faculty of Medicine and Health,   
University of Sydney,  
NSW 2050,  
Australia  
email: nick.glozier@sydney.edu.au

\

Draft: `r format(Sys.time(), '%d %B, %Y')`  
Supplementary tables:      4  
Supplementary figures:     3

**keywords:** Subjective wellbeing, household wealth, HILDA

\

***

## Supplementary Material

#### Model Selection (2002-2018)

##### Figure S1. Linear model evidence (WAIC~linear~ — WAIC~piecewise~)
```{r complete_waic}
waic_result <- read_rds("../results/model_comparisons_waic_R2_losat.RDS")

waic_result %>%
  select(model, year, elpd_diff, se_diff) %>%
  filter(elpd_diff != 0) %>%
  rowwise() %>%
  mutate(samples = list(rnorm(n = 1e5, 
                              mean = elpd_diff * -2, # for linear advantage
                              sd   = se_diff *  2)),
         year = paste0("AD", year)
         ) %>%
  select(year, samples) %>%
  group_by(year) %>%
  unnest(samples) %>% 
  mutate(sample_id = c(1:1e5)) %>%
  pivot_wider(id_col = "sample_id", names_from = "year", values_from = "samples") %>%
  select(-sample_id) -> d1

# "fix" d1
colnames(d1) <- rev(colnames(d1))

color_scheme_set("blue")

p2 <- mcmc_intervals(x = d1,
  pars = colnames(d1),
  prob = 0.0,
  prob_outer = .95
  ) +
  scale_x_continuous(
    limits = c(-200, 150)
    ) +
  geom_vline(aes(xintercept = 0), color = "grey90") +
  theme_test() +
  labs(
    subtitle = "Satisfaction",
    x = "WAIC difference (95% interval)")

waic_result <- read_rds("../results/model_comparisons_waic_R2_gh9.RDS")

waic_result %>%
  select(model, year, elpd_diff, se_diff) %>%
  filter(elpd_diff != 0) %>%
  rowwise() %>%
  mutate(samples = list(rnorm(n = 1e5, 
                              mean = elpd_diff * 2, # for pw advantage
                              sd   = se_diff *  2)),
         year = paste0("AD", year)
         ) %>%
  select(year, samples) %>%
  group_by(year) %>%
  unnest(samples) %>% 
  mutate(sample_id = c(1:1e5)) %>%
  pivot_wider(id_col = "sample_id", names_from = "year", values_from = "samples") %>%
  select(-sample_id) -> d2

color_scheme_set("red")

p1 <- mcmc_intervals(x = d2,
           pars = rev(colnames(d2)),
           prob = 0.00,
           prob_outer = 0.95
           ) +
  scale_x_continuous(
    limits = c(-200, 150)
    ) +
  geom_vline(aes(xintercept = 0), color = "grey90") +
  theme_test() +
  labs(
    subtitle = "Happiness")
```

```{r Figure_S1, fig.width=6.5, fig.height=7, dpi=300}
p1 / p2
```

```
Figure S1 legend: Differences in posterior evidence (WAIC) for a linear fit over a piecewise fit for satisfaction (blue) and happiness (red). The filled circle indicates the mean of the distribution and the horizontal bars represents the 95% credible interval. A difference greater than zero is support for the linear model and a difference below zero is support for the piecewise model.
```

<br><br>

#### Parameter Estimation (2002-2018)

##### Figure S2. Posterior distributions of &omega; (95% distribution)  
```{r gh9_omega}
path_to_gh9_fits <- list.files(
  path = "../results",
  pattern = '^fit_cp_hifdip_gh9_cov_.*.RDS$',
  full.names = TRUE
)

posterior_alphas <- list()
for (pathtofile in path_to_gh9_fits) {
  df <- read_rds(pathtofile) %>%
    as.data.frame() %>%
    select(b_alpha_Intercept)

  colnames(df) <- paste0("param_", substring(pathtofile, 34, 37))

  posterior_alphas <- bind_cols(posterior_alphas, df)
}

posterior_alphas %>%
  mutate_all(.funs = ~brms::inv_logit_scaled(., 0, 100)*10) -> posterior_omegas

posteriors_trim <- posterior_omegas %>%
  mutate(
    # param_2010 = (param_2010 + param_2011)/2,
    param_2003 = (param_2002 + param_2004)/2)

color_scheme_set("red")

p3 <- mcmc_intervals(
  posteriors_trim,
  pars = rev(colnames(posteriors_trim)),
  prob = 0.0, 
  prob_outer = .95 
  ) +
  geom_vline(aes(xintercept = median(posteriors_trim$param_2002)), 
             color = "gray", 
             linetype = 3) +
  theme_test() +
  coord_cartesian(xlim = c(20, 200)) +
  labs(
    subtitle = "Happiness",
    x = "")
```

```{r losat_omega}
path_to_losat_fits <- list.files(
  path = "../results",
  pattern = '^fit_cp_hifdip_losat_cov_.*.RDS$',
  full.names = TRUE
)

posterior_alphas <- list()
for (pathtofile in path_to_losat_fits) {
  df <- read_rds(pathtofile) %>%
    as.data.frame() %>%
    select(b_alpha_Intercept)

  colnames(df) <- paste0("param_", substring(pathtofile, 36, 39))

  posterior_alphas <- bind_cols(posterior_alphas, df)
}

posterior_alphas %>%
  mutate_all(.funs = ~brms::inv_logit_scaled(., 0, 100)*10) -> posterior_omegas

posteriors_trim <- posterior_omegas %>%
  mutate(
    param_2006 = (param_2007 + param_2005)/2)

color_scheme_set("blue")

p4 <- mcmc_intervals(
  posteriors_trim,
  pars = rev(colnames(posteriors_trim)),
  prob = 0.0, 
  prob_outer = .95 
  ) +
  geom_vline(aes(xintercept = median(posteriors_trim$param_2002)), 
             color = "gray", 
             linetype = 3) +
  theme_test() +
  coord_cartesian(xlim = c(20, 200)) +
  labs(
    subtitle = "Satisfaction",
    x = "Household income ($000s)")
```

```{r Figure_S2, fig.width = 6, fig.height = 7, dpi=300}
p3 / p4
```

```
Figure S2 legend: Posterior distribution of the change point parameter representing the location in household income ($) for each year. Horizontal bar represents the 95% credible interval.
```

<br>

#### Non-linear fits

We tested the relationship between wealth and each wellbeing variable (happiness and satisfaction) using cubic **B-spline models**. A cubic B-spline was the most complex model (more degrees of freedom), and included here because it has been recently used to describe the maximum point at which wellbeing no longer increases with wealth [i.e., "satiety", @jebb2018happiness]. Following Jebb et al [-@jebb2018happiness], we restricted the cubic B-spline model to *n* = 4, 5 and 6 knots and display wealth on a log scale.

<br>

##### Figure S3. Cubic spline of household wealth on satisfaction and happiness
```{r losat_gh9_hwnwip}
source("../src/happyfun.r")
happywealth <- read_rds("../data/happywealth.rds")

p1 <- happywealth %>%
  filter(hifdip_adj > 0) %>%
  rename(Satisfaction = losat, `Household wealth` = hifdip_adj) %>% 
  plot_inflection_cubic(Satisfaction, `Household wealth`) +
  scale_x_continuous(trans = "log", 
                     breaks = c(5, 10, 20, 40, 80, 160),
                     limits = c(2.5, 170))

p2 <- happywealth %>%
  filter(hifdip_adj > 0) %>%
  rename(Happiness = gh9, `Household wealth` = hifdip_adj) %>% 
  plot_inflection_cubic(Happiness, `Household wealth`) +
  scale_x_continuous(trans = "log", 
                     breaks = c(5, 10, 20, 40, 80, 160),
                     limits = c(2.5, 170))
```
```{r cubic_plot, fig.height=9, fig.width=11}
p1 / p2
```

<br>

The cubic B-spline models failed to reveal a complete satiety point (horizontal inflection) in satisfaction before the highest decile in any year. Also note that despite using a log scale, we do not observe a clear, regular logarithmic function with either wellbeing measure. Other researchers have argued that wealth represents a quantitative dimension subject to Weber's Law of perception [@kahneman2010high]. Under this view, similar proportional changes in income or wealth produce similar evaluations of satisfaction or happiness, rather than the absolute change. That is, a \$100 increase in wealth will produce a greater impact on someone earning a minimum wage than someone earning many orders of magnitude of that amount. However while it is true that wealth clearly represents a quantitative dimension, the effects of wealth observed here represent cross-sectional differences in the population rather than changes experienced by an individual over time. It is not clear the extent to which Weber's Law applies to cross-sectional or between-subject differences in quantitative dimensions such as wealth.

<br>

#### Change point model parameter estimates

There were _df_ = 59,876 total observations. 4000 samples (post-warmup) were drawn using sampling(NUTS). For each parameter, Bulk_ESS and Tail_ESS are effective sample size measures, and Rhat is the potential scale reduction factor on split chains (at convergence, Rhat = 1).

The Bayes R^2^ = 0.03964581 ±0.001497172 (0.03672371-0.04262236)

<br>

_Coefficient definitions_:

_alpha_ is the log value of wealth (x) at the changepoint  
_b0_ is the intercept denoting the expected value of wellbeing (y) at the changepoint  
_b1_ is the linear slope before the changepoint  
_b2_ is the linear slope after the changepoint  
_b3_ is the coefficient of education (graduate)  
_b4_ is the coefficient of age  
_b5_ is the coefficient of age^2^  
_b6_ is the coefficient of sex (male)

<br>

```{r gh9_pw_summary, ft.align="left"}
population_summary <- tibble()
for (pathtofile in path_to_gh9_fits) {
  fit_summary <- read_rds(pathtofile) %>%
    summary()

  sigma_summary <- fit_summary$spec_pars %>%
    as_tibble(rownames = "Parameter")

  fixed_summary <- fit_summary$fixed %>%
    as_tibble(rownames = "Parameter") %>%
    bind_rows(sigma_summary) %>%
    mutate(Year = substring(pathtofile, 34, 37)) %>%
    select(Parameter, Year, everything())

  population_summary <- bind_rows(population_summary, fixed_summary)
}

population_summary %>%
  mutate(Parameter = str_remove(Parameter, "_Intercept")) %>%
  arrange(Parameter, desc(Year)) %>%
  flextable() %>%
  merge_v(j = 1) %>%
  valign(j = 1, valign = "top") %>%
  bg(i = ~Parameter %in% c("b0", 
                           "b2", 
                           "b4", 
                           "b6"), 
     bg = "grey90", 
     part = "body") %>%
  bg(i = ~Parameter %in% c("sigma"), 
     bg = "#b3cde0", 
     part = "body") %>%
  colformat_num(j = 3:6, digits= 3) %>%
  colformat_num(j = 7, digits= 1) %>%
  set_caption("Table S1. Population-level piecewise parameter estimates for happiness") %>%
  autofit()
```

<br>

```{r gh9_lm_summary, ft.align="left"}
path_to_gh9_lm_fits <- list.files(
  path = "../results",
  pattern = '^fit_lm_hifdip_gh9_cov_.*.RDS$',
  full.names = TRUE
)

population_summary <- tibble()
for (pathtofile in path_to_gh9_lm_fits) {
  fit_summary <- read_rds(pathtofile) %>%
    summary()

  sigma_summary <- fit_summary$spec_pars %>%
    as_tibble(rownames = "Parameter")

  fixed_summary <- fit_summary$fixed %>%
    as_tibble(rownames = "Parameter") %>%
    bind_rows(sigma_summary) %>%
    mutate(Year = substring(pathtofile, 34, 37)) %>%
    select(Parameter, Year, everything())

  population_summary <- bind_rows(population_summary, fixed_summary)
}

population_summary %>%
  mutate(Parameter = fct_relevel(Parameter, "Intercept", "dollars", "grad")) %>%
  arrange(Parameter, desc(Year)) %>%
  flextable() %>%
  merge_v(j = 1) %>%
  valign(j = 1, valign = "top") %>%
  bg(i = ~Parameter %in% c("dollars",
                           "age",
                           "male"),
     bg = "grey90",
     part = "body") %>%
  bg(i = ~Parameter %in% c("sigma"), 
     bg = "#b3cde0", 
     part = "body") %>%
  colformat_num(j = 3:6, digits= 3) %>%
  colformat_num(j = 7, digits= 1) %>%
  set_caption("Table S2. Population-level linear parameter estimates for happiness") %>%
  autofit()
```

<br>

```{r losat_pw_summary, ft.align="left"}
population_summary <- tibble()
for (pathtofile in path_to_losat_fits) {
  fit_summary <- read_rds(pathtofile) %>%
    summary()

  sigma_summary <- fit_summary$spec_pars %>%
    as_tibble(rownames = "Parameter")

  fixed_summary <- fit_summary$fixed %>%
    as_tibble(rownames = "Parameter") %>%
    bind_rows(sigma_summary) %>%
    mutate(Year = substring(pathtofile, 36, 39)) %>%
    select(Parameter, Year, everything())

  population_summary <- bind_rows(population_summary, fixed_summary)
}

population_summary %>%
  mutate(Parameter = str_remove(Parameter, "_Intercept")) %>%
  arrange(Parameter, desc(Year)) %>%
  flextable() %>%
  merge_v(j = 1) %>%
  valign(j = 1, valign = "top") %>%
  bg(i = ~Parameter %in% c("b0", 
                           "b2", 
                           "b4", 
                           "b6"), 
     bg = "grey90", 
     part = "body") %>%
  bg(i = ~Parameter %in% c("sigma"), 
     bg = "#b3cde0", 
     part = "body") %>%
  colformat_num(j = 3:6, digits= 3) %>%
  colformat_num(j = 7, digits= 1) %>%
  set_caption("Table S3. Population-level piecewise parameter estimates for satisfaction") %>%
  autofit()
```

<br>

```{r losat_lm_summary, ft.align="left"}
path_to_losat_lm_fits <- list.files(
  path = "../results",
  pattern = '^fit_lm_hifdip_losat_cov_.*.RDS$',
  full.names = TRUE
)

population_summary <- tibble()
for (pathtofile in path_to_gh9_lm_fits) {
  fit_summary <- read_rds(pathtofile) %>%
    summary()

  sigma_summary <- fit_summary$spec_pars %>%
    as_tibble(rownames = "Parameter")

  fixed_summary <- fit_summary$fixed %>%
    as_tibble(rownames = "Parameter") %>%
    bind_rows(sigma_summary) %>%
    mutate(Year = substring(pathtofile, 34, 37)) %>%
    select(Parameter, Year, everything())

  population_summary <- bind_rows(population_summary, fixed_summary)
}

population_summary %>%
  mutate(Parameter = fct_relevel(Parameter, "Intercept", "dollars", "grad")) %>%
  arrange(Parameter, desc(Year)) %>%
  flextable() %>%
  merge_v(j = 1) %>%
  valign(j = 1, valign = "top") %>%
  bg(i = ~Parameter %in% c("dollars",
                           "age",
                           "male"),
     bg = "grey90",
     part = "body") %>%
  bg(i = ~Parameter %in% c("sigma"), 
     bg = "#b3cde0", 
     part = "body") %>%
  colformat_num(j = 3:6, digits= 3) %>%
  colformat_num(j = 7, digits= 1) %>%
  set_caption("Table S4. Population-level linear parameter estimates for happiness") %>%
  autofit()
```

<br><br>

##### Figure S5. Posterior predictive distribution of marginal effect of wealth by year (95% CI)  
(not run)

```{r figure_s2, fig.height = 3, eval=F}
library(brms)

marginals <- conditional_effects(
  hifdip_gh9_fixed_fit,
  effects = c("dollars:year"), # year, dollars, 
  method = "fitted"
  )

marginals$`dollars:year`$year <- factor(marginals$`dollars:year`$year, 
                                        levels = c("2002", "2006", "2010", "2014", "2018"))

plot(marginals, plot=F)[[1]] +
  xlim(c(0, 30)) +
  ylim(c(-0.5, 1)) +
  facet_wrap(~year, nrow=1) +
  theme_test() +
  theme(legend.position = "none") +
  labs(x = "", y = "")
```

<br><br>

## References

### 2020-12-02 src/figure 1.r

@@ -1,0 +1,182 @@
+library(tidyverse)
+library(haven)
+
+#### Load data ####
+path_to_hilda <- list.files(
+  path = '~/Dropbox/HILDA/data',
+  pattern = '^Combined.*.dta$',
+  full.names = TRUE
+)
+
+hilda <- list()
+for (pathtofile in path_to_hilda) {
+  df <- read_dta(pathtofile)
+  hilda <- append(hilda, list(df))
+  cat('.')
+}
+
+# Helper function
+source('~/Dropbox/HILDA/src/gather_hilda.R')
+
+
+#### Satisfaction ####
+satisfaction <- gather_hilda(hilda, c(
+  "losat",   # Life satisfaction
+  "hhwtrps") # sample weights for life satisfaction (RP)
+  ) %>%
+  spread(code, val) %>%
+  mutate_if(is.double, ~ ifelse(. < 0, NA_real_, .)) %>%
+  transmute(xwaveid, wave, losat, losat_wt = losat * hhwtrps)
+
+#### Income ####
+# CPI values from https://www.pc.gov.au/research/completed/rising-inequality
+CPI = data.frame(
+  wave = letters[1:18],
+  deflator = c(1.52547554, # 2000/01
+               1.48315720, # 2001/02
+               1.43988458, # 2002/03
+               1.40607389, # 2003/04
+               1.37297463, # 2004/05
+               1.33027251, # 2005/06
+               1.29200230, # 2006/07
+               1.24993042, # 2007/08
+               1.21214575, # 2008/09
+               1.18464785, # 2009/10
+               1.14888718, # 2010/11
+               1.12303076, # 2011/12
+               1.09804401, # 2012/13
+               1.06903118, # 2013/14
+               1.05101802, # 2014/15
+               1.03670360, # 2015/16
+               1.01929187, # 2016/17
+               1)) %>% mutate(inflator = rev(deflator))
+
+thresholds <- read_csv("../../docs/Frequencies 180c/HILDA-thresholds-by-wave-180.csv") %>%
+  filter(variable %in% c("hifdip")) %>%
+  gather("wave", "threshold", -variable, -label) %>%
+  separate(wave, into = c("temp", "wave"), sep = "e") %>%
+  mutate(
+    variable = paste0(variable, "_thld"),
+    wave = letters[as.numeric(wave)]) %>%
[…]

### 2020-12-02 src/figure 5.r

@@ -1,0 +1,62 @@
+
+predicted_fits <- list()
+for (pathtofile in path_to_fits) {
+  fit <- read_rds(pathtofile) 
+  
+  ans <- model.frame(fit) %>%
+    summarise_all(median)
+  
+  newdat = data.frame(
+    dollars = c(5, ans[, "dollars"], 8),
+    age = rep(ans[, "age"], 3),
+    age_sq = rep(ans[, "age_sq"], 3),
+    male = rep(ans[, "male"], 3), 
+    grad = rep(ans[, "grad"], 3)
+  )
+  
+  df <- fitted(fit, newdata = newdat) %>% 
+    as_tibble() %>%
+    mutate(
+      Dollars = c("$50K", "\u03bc", "$80K"),
+      year = as.numeric(substring(pathtofile, 34, 37)))
+  
+  predicted_fits <- bind_rows(predicted_fits, df)
+}
+
+#### plot those fits
+predicted_fits %>%
+  mutate(Dollars = fct_relevel(Dollars, "$80K", "\u03bc")) %>%
+  ggplot(aes(x = year, y = Estimate, color = Dollars)) +
+    geom_point(size = 4, alpha = 0.25) +
+    geom_smooth(method = "loess", se = F, span = 3) +
+    facet_wrap(~Dollars) +
+    labs(title = "Predicted happiness at varying income levels") +
+    ggthemes::scale_colour_wsj() +
+    ggthemes::theme_wsj() +
+    theme(legend.position = "none",
+          plot.title = element_text(size = 19),
+          plot.caption = element_text(size = 12),
+          strip.text = element_text(size = 12))
+
+#### plot those fits
+predicted_fits %>%
+  select(Estimate, Dollars, year) %>%
+  filter(Dollars %in% c("$80K", "$50K")) %>%
+  spread(Dollars, Estimate) %>%
+  # mutate(`∆` = `$80K` - `$50K`) %>%
+  gather(income, happiness, -year) %>%
+  mutate(income = fct_relevel(income, "$50K", "$80K")) %>%
+  ggplot(aes(x = year, y = happiness, color = income)) +
+  geom_point(size = 4, alpha = 0.25) +
+  geom_smooth(method = "loess", se = F, span = 3) +
+  facet_wrap(~income) +
+  labs(title = "Estimated happiness at varying income\nin Australia from 2002-2018",
+       caption = "Source: HILDA, Melbourne University") -> p
+
+p +
+  ggthemes::scale_colour_wsj() +
[…]

### 2020-12-02 src/ft_themes.r

@@ -1,0 +1,247 @@
+#' ggplot2 theme for FT graphs.
+#' @param legend_right Logical indicating whether legend should be placed to
+#' the right of the plot. If FALSE, the default, legend is positioned above the
+#' plot.
+#' @param base_size The base font size
+#' @param base_family Font family
+#' @param base_line_size Default
+#' @param base_rect_size Default
+#' @importFrom ggplot2 %+replace%
+#' @export
+#' @examples
+#' library(ggplot2)
+#' ggplot(mpg, aes(displ, hwy, color = class)) +
+#'   geom_point() +
+#'   ft_theme()
+#'
+#' ggplot(mpg, aes(displ, hwy)) +
+#'   geom_point() +
+#'   facet_wrap(vars(class)) +
+#'   ft_theme()
+#'
+ft_theme <- function(legend_right = FALSE,
+                     base_size = 12,
+                     base_family = "",
+                     base_line_size = base_size / 170,
+                     base_rect_size = base_size / 170) {
+  
+  half_line <- base_size/2
+  grid_line_color <- ft_colors("black-20")
+  grid_line_size <- 0.2
+  title_text_color <- ft_colors("black")
+  other_text_color <- ft_colors("black-70")
+  
+  if(legend_right == TRUE){
+    spec_legend_position <- "right"
+    spec_legend_direction <- "vertical"
+    legend_justification_spec <- "center"
+    legend_box_spacing_spec = ggplot2::unit(2 * half_line, "pt")
+  } else {
+    spec_legend_position <- "top"
+    spec_legend_direction <- "horizontal"
+    legend_justification_spec <- c(0,0)
+    legend_box_spacing_spec <- ggplot2::unit(0, "char")
+  }
+  
+  ggplot2::theme_minimal(base_size = base_size,
+                         base_family = base_family,
+                         base_line_size = base_line_size) %+replace%
+    ggplot2::theme(
+      plot.title = ggplot2::element_text(
+        color = title_text_color,
+        size = ggplot2::rel(1.2),
+        face = "bold",
+        hjust = 0,
+        margin = ggplot2::margin(b = half_line)
+      ),
+      plot.subtitle = ggplot2::element_text(
+        color = other_text_color,
+        face = "bold",
+        hjust = 0,
[…]

### 2020-12-02 src/model_selection.r

@@ -59,16 +59,17 @@ for (i in c(2002:2018)) {
 }
 
 saveRDS(waic_results, "../results/model_comparisons_waic_R2_losat.RDS")
-waic_results <- readRDS("../results/model_comparisons_waic_R2_losat.RDS")
+waic_results_losat <- readRDS("../results/model_comparisons_waic_R2_losat.RDS")
+waic_results_gh9 <- readRDS("../results/model_comparisons_waic_R2_gh9.RDS")
 
 #### Plot results ####
-waic_results %>%
+waic_results_gh9 %>%
   select(model, year, elpd_diff, se_diff) %>%
   filter(elpd_diff != 0) %>%
-  filter(year %in% c(2002, 2006, 2010, 2014, 2018)) %>%
+  # filter(year %in% c(2002, 2006, 2010, 2014, 2018)) %>%
   rowwise() %>%
   mutate(samples = list(rnorm(n = 1e5, 
-                              mean = elpd_diff * -2, 
+                              mean = elpd_diff * 2, 
                               sd   = se_diff *  2)),
          year = paste0("AD", year)
          ) %>%
@@ -87,5 +88,5 @@ mcmc_intervals(x = d1,
   geom_vline(aes(xintercept = 0), color = "grey90") +
   theme_test() +
   labs(
-    subtitle = "Happiness",
+    subtitle = "Happiness (linear deviance over piecewise deviance)",
     x = "WAIC difference (95% interval)")

### 2020-12-02 src/modeling.r

@@ -11,14 +11,14 @@ library(brms)
 options(mc.cores = 4)
 
 #### Import ####
-happywealth_covs <- readRDS("../data/happywealth_covs.rds")
+happywealth <- readRDS("../data/happywealth.rds")
 
 #### Preprocessing ####
-happywealth_covs %>%
-  filter(!student, year == 2002) %>%
+happywealth %>%
+  filter(!student, year == 2001) %>%
   transmute(
-    y = as.vector(scale(losat)), # gh9
-    dollars = adj_hifdip/10000, 
+    y = as.vector(scale(gh9)), # gh9
+    dollars = hifdip_adj/10000, 
     male = as.numeric(male),
     age = as.vector(scale(age)),
     age_sq = age^2,
@@ -76,7 +76,7 @@ bprior <- prior(normal(0, 3), nlpar = "b0") + # normal(0, 1) did not converge Se
   prior(normal(0, 3), nlpar = "b5") +  
   prior(normal(0, 3), nlpar = "b6") +
   prior(normal(0, 3), nlpar = "alpha")      # normal(0, 3) with init = "0" 12/09/2020
-                                            # normal(0, sqrt(3)) with default init did not converge
+                                            # normal(0, sqrt(3)) with default init didn't converge
 
 # e.g., plot prior for alpha
 hist(inv_logit_scaled(rnorm(1000, mean = 0, sd = 3), 0, 100))
@@ -174,7 +174,7 @@ posteriors %>%
 # omega             $54.5K 6.22  45.9   70.5   2014
 # omega             $72.6K 4.83  62.9   82.1   2018
 
-# saveRDS(fit, "../results/hifdip_gh9_fixed_cov_2002.RDS")
+saveRDS(fit, "../results/fit_cp_hifdip_gh9_cov_2001.RDS")
 
 #### Piece-wise fits ####
 for (i in c(2018:2010)) {

### 2020-12-02 src/piecewise_model_simulation.r

@@ -1,3 +1,6 @@
+library(tidyverse)
+library(brms)
+
 tibble(
   year = rep(1:16, each = 20),
   dollars = rep(1:10, 32)
@@ -11,8 +14,8 @@ tibble(
     omega = 5, #ceiling((year + 2)/2.5),
     y = b0 + b1*(dollars - omega)*(omega > dollars) + b2*(dollars - omega)*(dollars >= omega) + rnorm(1),
     age = sample(18:100, 1),
-    male = sample(0:1, 1),
-    grad = rbinom(1, 1, 0.25)) %>%
+    male = as.logical(sample(0:1, 1)),
+    grad = as.logical(rbinom(1, 1, 0.25))) %>%
   ungroup() %>%
   mutate(
     y = as.vector(scale(y)),
@@ -39,17 +42,17 @@ bfixed <- bf(
   nl = TRUE
 )
 
-# priors                                      I tried tighter priors:
-bprior <- prior(normal(0, 3), nlpar = "b0") + # normal(0, 1) did not converge Sep 6th
-  prior(normal(0, 3), nlpar = "b1") +         # normal(0, 1) did not converge Sep 6th
-  prior(normal(0, 3), nlpar = "b2") +         # normal(0, 1) did not converge Sep 6th
+# priors                                      
+bprior <- prior(normal(0, 3), nlpar = "b0") + 
+  prior(normal(0, 3), nlpar = "b1") +         
+  prior(normal(0, 3), nlpar = "b2") +         
   prior(normal(0, 3), nlpar = "b3") +
   prior(normal(0, 3), nlpar = "b4") +
   prior(normal(0, 3), nlpar = "b5") +
   prior(normal(0, 3), nlpar = "b6") +
   prior(normal(0, sqrt(3)), nlpar = "alpha")      # normal(0, 3) with init = "0" 12/09/2020
 
-# fitting (try more iterations)
+# fitting 
 fit <- brm(bfixed, data = select(df, y, dollars, age, agesq, male, grad), 
            prior = bprior, 
            # iter = 4000,
@@ -60,10 +63,53 @@ fit <- brm(bfixed, data = select(df, y, dollars, age, agesq, male, grad),
 
 summary(fit)
 
+# Plot the marginal effect of the breakpoint
 breakpoint <- inv_logit_scaled(fixef(fit)['alpha_Intercept', 'Estimate'], 0, 15)
[…]

### 2020-12-02 src/plotting cost of happiness 2012-2018.r

@@ -1,0 +1,113 @@
+library(bayesplot)
+
+
+fit <- read_rds("../results/fit_cp_hifdip_losat_cov_2006.RDS")
+
+plot(fit)
+
+path_to_fits <- list.files(
+  path = "../results",#'~/Dropbox/HILDA/data',
+  pattern = '^fit_cp_hifdip_losat_cov_.*.RDS$',
+  full.names = TRUE
+)
+
+alphas <- list()
+for (pathtofile in path_to_fits) {
+  df <- read_rds(pathtofile) %>%
+    as.data.frame() %>%
+    select(b_alpha_Intercept) 
+  
+  # colnames(df) <- paste0("b_alpha_", substring(pathtofile, 34, 37)) # gh9
+  colnames(df) <- paste0("b_alpha_", substring(pathtofile, 36, 39)) # losat
+  
+  alphas <- bind_cols(alphas, df)
+}
+
+alphas %>%
+  mutate_all(.funs = ~brms::inv_logit_scaled(., 0, 100)*10) -> posteriors
+
+# OPTIONAL: Trim posteriors for 2003 outlier
+posteriors %>%
+  mutate(b_alpha_2003 = (b_alpha_2002 + b_alpha_2004)/2) -> posteriors
+
+# OPTIONAL: Trim posteriors for 2006 outlier (did not converge)
+posteriors %>%
+  mutate(b_alpha_2006 = (b_alpha_2005 + b_alpha_2007)/2) -> posteriors
+
+color_scheme_set("blue")
+
+mcmc_areas_ridges(
+  posteriors,
+  pars = rev(colnames(posteriors)),
+  prob = 0.90, # 80% intervals
+  prob_outer = .99 # 99%
+  ) +
+  theme_test() +
+  geom_vline(aes(xintercept = median(posteriors$b_alpha_2002)), 
+             color = "gray", 
+             linetype = 3) +
+  labs(subtitle = "Cost of happiness by year and household wealth",
+       x = "Household wealth ($000s)")
+
+happywealth_covs %>%
+  select(year, xwaveid, adj_hifdip) %>%
+  mutate(
+    year = paste0("b_alpha_", year),
+    adj_hifdip = adj_hifdip/1000) %>%
+  left_join(
+    gather(
+      summarise_all(posteriors, list(median)), year, change_point
+    )) %>%
+  mutate(
+    year = extract_numeric(year)
+    ) %>%
+  group_by(year) %>%
+  summarise(
[…]

### 2020-12-02 src/preprocessing.r

@@ -2,7 +2,7 @@
 # RW Morris, N Kettlewell, N Glozier
 # 
 # Preprocessing 
-# Nov 21st 2020
+# Nov 27th 2020
 # 
 # Libraries
 library(tidyverse)
@@ -23,9 +23,9 @@ for (pathtofile in path_to_hilda) {
 }
 
 # Helper function
-source('../../src/gather_hilda.R')
+source('~/Dropbox/HILDA/src/gather_hilda.R')
 
-# DEMOGRAPHIC VARIABLES 
+#### DEMOGRAPHICS ####
 gather_hilda(hilda, c(
   "esbrd",   # employment status (broad)
   "hhda10",  # SEIFA 2001 Decile of socio-economic advantage (higher is better)
@@ -33,40 +33,80 @@ gather_hilda(hilda, c(
   "hgsex",   # sex (male = 1)
   "helth",   # long term health condition (yes = 1)
   "hhpers",  # no. persons in household
-  "hhrhid",  # household id (within wave)
+  # "hhrhid",  # household id (within wave)
   "mrcurr",  # current marital status (married/de facto ≤ 2)
   "edhigh1", # highest education achieved
   "edfts"    # current fulltime student (yes = 1)
-)) %>%
+  )) %>%
   spread(code, val) %>%
   mutate_if(is.double, ~ ifelse(. < 0, NA_real_, .)) -> demographic_items
 
-# MODEL VARIABLES (household wealth, satisfaction, happiness):
-# Note that regular HILDA income variables were used (not total income, which 
-# include irregular income) to match with the ABS, and because more consistent
-# prior to 2012. See Appendix in Wilkins (2014) 'Derived income variables in the
-# HILDA survey' for explanation
-# 
-gather_hilda(hilda, c(
-  "hifdip", # household disposable regular income
-  'losat', # Life satisfaction
-  'gh9a', # Vitality: feel full of life (lower is better)*
-  'gh9b', # Mental H: Been a nervous person (higher is better)
-  'gh9c', # Mental H: Felt so down in the dumps (higher is better)
-  'gh9d', # Mental H: Felt calm and peaceful (lower is better)*
-  'gh9e', # Vitality: Have a lot of energy (lower is better)*
-  'gh9f', # Mental H: Felt down (higher is better)
-  'gh9g', # Vitality: Felt worn out (higher is better)
-  'gh9h', # Mental H: Been happy (lower is better)*
-  'gh9i'  # Vitality: Felt tired (higher is better)
[…]

### 2020-12-02 src/draft0.Rmd
November draft ready

@@
 We tested the relationship between wealth and both happiness and life satisfaction using household economic data from Australia between 2002-2018.
-We discovered the cost of happiness has increased over those 18 years faster than inflation (i.e., cost of living).
+We discovered the cost of happiness has increased over those 17 years faster than inflation (i.e., cost of living).
 This result shows that inequalities in wealth may be driving inequities in happiness and wellbeing, with implications for health and recent government policy-goals to monitor and improve wellbeing.
@@
 
-We used household economic panel data from Australia (HILDA) to provide the first investigation of whether changes in wealth and wellbeing have changed the cost of happiness over the last 18 years (2002-2018).
+We used household economic panel data from Australia (HILDA) to provide the first investigation of whether changes in wealth and wellbeing have changed the cost of happiness over the last 17 years (2002-2018).
 HILDA provides a representative sample of households in Australia with detailed measurements of wealth and subjective wellbeing in the same sample, which makes it an excellent data source to investigate the present question.
@@
 
-Average life satisfaction levels were very steady between 2002-2018, while average happiness score decreased slightly over the 18 years.
+Average life satisfaction levels were very steady between 2002-2018, while average happiness score decreased slightly over the 17 years.
 The proportions of each sex and relationship status were stable over time, as were the average household size and SEIFA index.
@@
 ```
-```{r figure_1, fig.height=10, fig.width=9}
+```{r figure_1, fig.height=10, fig.width=9, dpi=300}
 p1 / p2 / p3 / p4
@@
 
-##### Figure 4. Estimated happiness (SD units) at $50K and $80K from 2002-2018 
-```{r figure_4}
+##### Figure 4. Estimated happiness (SD units) at $50K/yr and $80K/yr from 2002-2018 
+```{r figure_4, fig.width=7, fig.height=5, dpi=300}
 predicted_fits <- list()
@@
   facet_wrap(~income) +
-  labs(caption = "Source: Household disposable income, HILDA, Melbourne University", 
+  labs(caption = "Source:     HILDA, University of Melbourne
+                 Predicted happiness scores (SD unit) from piecewise regression on real household income", 
        x = "", y = "") -> p
@@
 ```
-Figure 4 legend: The difference (∆) in happiness between a household earning $50K per year and a household earning $80K per year has increased since 2009.
+Figure 4 legend: The difference (∆) in happiness between individuals with a household income of $50K per year and a household income of $80K per year has increased since 2009.
 ```
@@
 
-Figure 5 presents house income levels adjusted for the Australian population (by age, sex, marital status, labour force participation and region).
+Figure 5 presents house income levels weighted for the Australian population (by age, sex, marital status, labour force participation and region).
 It shows the cost of happiness increased faster than household income between 2002 and 2018.
@@
 ##### Figure 5.  
-```{r figure_5, fig.width=7, fig.height=6}
+```{r figure_5, fig.width=7, fig.height=6, dpi=300}
 # Tabulate against median income
@@
   labs(title = "Wealth & happiness in Australia between 2002-2018",
-       caption = "Source: HILDA sample weighted for Australian population
-                  Household equivalised income per person (modified OECD scale)
-                  Excluding full-time students and topcoded individuals",
+       caption = "Source: HILDA sample, University of Melbourne
+                  Real household income (modified OECD scale) weighted for the Australian population",
        x = "  ", y = "") +
-  geom_vline(aes(xintercept = 2009), color = "red", linetype = 3) +
-  geom_text(data = gfc.note, aes(x = 2009.5, y = 57, label = string),
-            colour = "red", size = 3, hjust = "left") +
+  # geom_vline(aes(xintercept = 2009), color = "red", linetype = 3) +
+  # geom_text(data = gfc.note, aes(x = 2009.5, y = 57, label = string),
+  #           colour = "red", size = 3, hjust = "left") +
   geom_text(data = df.label,
@@
 ```
-Figure 5 legend: Household income has stagnated in Australia since the GFC while the cost of happiness has increased.
+Figure 5 legend: Real household income has stagnated in Australia since 2009 (post GFC) while the cost of happiness has increased.
 Consequently fewer Australians have a household income exceeding the cost of happiness in 2018.

### 2020-12-02 src/wdraft0.Rmd
November draft ready

@@
 | Draft      | `r format(Sys.time(), '%d %B, %Y')` |
-| Word count | 3061                                |
-| Tables     | 2                                   |
-| Figures    | 3                                   |
+| Word count | 3767                                |
+| Tables     | 1                                   |
+| Figures    | 5                                   |
 
@@
 
-**keywords:** Subjective wellbeing, household wealth, HILDA
+**keywords:** Subjective wellbeing, wealth, household income, HILDA
 
@@
 While the location of this change point has been determined, and the cost of happiness reportedly ranges between USD\$60 to $95K, there has been no investigation as to whether the cost of happiness has increased or decreased over time.
-We tested the relationship between household wealth and both happiness and life satisfaction using household economic data from Australia between 2002-2016.
-We discovered the cost of happiness has increased over those 16 years faster than inflation (i.e., cost of living).
+We tested the relationship between wealth and both happiness and life satisfaction using household economic data from Australia between 2002-2018.
+We discovered the cost of happiness has increased over those 17 years faster than inflation (i.e., cost of living).
 This result shows that inequalities in wealth may be driving inequities in happiness and wellbeing, with implications for health and recent government policy-goals to monitor and improve wellbeing.
@@
 
-We used household economic panel data from Australia (HILDA) to provide the first investigation of whether changes in wealth and wellbeing have changed the cost of happiness over the last 16 years (2002-2018).
+We used household economic panel data from Australia (HILDA) to provide the first investigation of whether changes in wealth and wellbeing have changed the cost of happiness over the last 17 years (2002-2018).
 HILDA provides a representative sample of households in Australia with detailed measurements of wealth and subjective wellbeing in the same sample, which makes it an excellent data source to investigate the present question.
@@
 
-The piecewise model was a simple extension of this to include a free parameter to represent the changepoint in wealth ($\omega$) as well as the slope after the change point ($\beta_2$):
+The piecewise model was a simple extension of this to include a free parameter to represent the changepoint in wealth ($\omega$) as well as the slope before the change point ($\beta_1$) and after the change point ($\beta_2$):
 
 $$
-\mu_i = \beta_0 + \beta_1 (x_i - \omega_t) (x_i ≤ \omega_t) + \beta_2 (x_i - \omega_t) (x_i > \omega_t)
+\mu_i = \beta_0 + \beta_1 (x_i - \omega) (x_i ≤ \omega) + \beta_2 (x_i - \omega) (x_i > \omega) + \beta_3 X_i
 $$
+
+Where $x_i$ was an individual's household wealth, and $X_i$ were covariates for age, age^2^, sex and education.
+
 <br>
 
-The above models estimated population-level (fixed) effects for each year in the data (*t* = 2002, 2006, 2010, 2014 and 2018).
+The above models estimated population-level effects separately for each year (*t* = 2002, 2006, 2010, 2014 and 2018).
 Because we were interested in the location of the change point between wealth and wellbeing that existed across individuals within each year, we ignored the panel design of HILDA because the dependency between observations of the same person across years was orthogonal to our effects of interest.
-We assumed an unstructured variance-covariance matrix between years, and specified weakly informed priors for each *&beta;*, and a uniform prior over the restricted range of wealth values for *&omega;*.
+We specified weakly informed priors for each *&beta;*, and a uniform prior over the restricted range of wealth values for *&omega;*.
 
@@
 ```{r read_write_covs}
-happywealth_covs <- read_rds("../results/happywealth_covs.rds")
+happywealth <- read_rds("../data/happywealth.rds")
 ```
@@
 
-happywealth_covs %>%
-  filter(!student, (!is.na(losat)|!is.na(gh9))) %>%
+happywealth %>%
+  filter(year %in% c(2002, 2006, 2010, 2014, 2018)) %>%
+  filter(!student, !top_hifdip, (!is.na(losat)|!is.na(gh9))) %>%
   mutate(
     Sex = if_else(male, "Male", "Female"),
-    SEIFA = as.integer(extract_numeric(SEIFA)),
-    SEIFA = if_else(SEIFA > 0, SEIFA, NA_integer_),
     Education = fct_collapse(
       edu,
-      `High school or less` = c("None", "In school", "Highschool", "Trade"),
-      Graduate = c("Grad", "Postgrad")
+      `High school or less` = c("None", "In school", "Highschool"),
+      Graduate = c("Grad")
       ),
@@
     year, 
-    # `Life satisfaction` = losat, `Happiness` = gh9, 
-    # `Household wealth ($)` = adj_hifdip, 
+    `Life satisfaction` = losat, `Happiness` = gh9,
     Sex, `Age (years)` = age, Education, Workforce, `Relationship status`, 
@@
 
+Average life satisfaction levels were very steady between 2002-2018, while average happiness score decreased slightly over the 17 years.
 The proportions of each sex and relationship status were stable over time, as were the average household size and SEIFA index.
-However age and chronic health conditions tended to increase over time, as expected in an aging longitudinal sample.
+However age and chronic health conditions tended to increase over time, as expected in an aging sample.
 Education levels (i.e., graduate percentage) also tended to increase over time, while changes in the workforce varied both up and down.
@@
 
-p4 <- happywealth_covs %>%
-  rename(Happiness = gh9, `Household wealth` = adj_hifdip) %>% 
-  filter(!student) %>%
-  plot_inflection_custom(Happiness, `Household wealth`, 
-                         knots = c(32, 46, 52.5, 86, 124.5)) +
+p4 <- happywealth %>%
+  rename(Happiness = gh9, `Household income` = hh_disp_inc_eq) %>% 
+  filter(!student, !top_hifdip) %>%
+  plot_inflection_custom(Happiness, `Household income`, 
+                         knots = c(43.7, 49.2, 47.4, 53.2, 72.3)) +
   scale_fill_manual(values = "#C79999") +
@@
 
-##### Figure 2. Model evidence (mean difference of WAIC~linear~ — WAIC~piecewise~)
+##### Figure 2. Linear model evidence (WAIC~linear~ — WAIC~piecewise~)
+
+```{r plot_waic_gh9}
+waic_result <- read_rds("../results/model_comparisons_waic_R2_gh9.RDS")
+
+waic_result %>%
+  select(model, year, elpd_diff, se_diff) %>%
+  filter(elpd_diff != 0) %>%
+  filter(year %in% c(2002, 2006, 2010, 2014, 2018)) %>%
+  rowwise() %>%
+  mutate(samples = list(rnorm(n = 1e5, 
+                              mean = elpd_diff * 2, # for pw advantage
+                              sd   = se_diff *  2)),
+         year = paste0("AD", year)
+         ) %>%
+  select(year, samples) %>%
+  group_by(year) %>%
+  unnest(samples) %>% 
+  mutate(sample_id = c(1:1e5)) %>%
+  pivot_wider(id_col = "sample_id", names_from = "year", values_from = "samples") %>%
+  select(-sample_id) -> d2
+
+color_scheme_set("red")
+
+p5 <- mcmc_intervals(x = d2,
+           pars = rev(colnames(d2)),
+           prob = 0.00,
+           prob_outer = 0.95
+           ) +
+  scale_x_continuous(
+    limits = c(-200, 150)
+    ) +
+  geom_vline(aes(xintercept = 0), color = "grey90") +
+  theme_test() +
+  labs(
+    subtitle = "Happiness")
+```
 
 ```{r plot_waic_losat}
-waic_result <- read_rds("../results/waic_losat.RDS")
+waic_result <- read_rds("../results/model_comparisons_waic_R2_losat.RDS")
 
-d1 <- data.frame(fy2018 = rnorm(n = 1e5, 
-                                     mean = waic_result[2, 1] * -2, 
-                                     sd   = waic_result[2, 2] *  2)/10,
-                 fy2014 = rnorm(n = 1e5, 
-                                     mean = waic_result[4, 1] * -2, 
-                                     sd   = waic_result[4, 2] *  2)/10,
-                 fy2010 = rnorm(n = 1e5, 
-                                     mean = waic_result[6, 1] * -2, 
-                                     sd   = waic_result[6, 2] *  2)/10,
-                 fy2006 = rnorm(n = 1e5, 
-                                     mean = waic_result[8, 1] * -2, 
-                                     sd   = waic_result[8, 2] *  2)/10,
-                 fy2002 = rnorm(n = 1e5, 
-                                     mean = waic_result[10, 1] * -2, 
-                                     sd   = waic_result[10, 2] *  2)/10
-                )
+waic_result %>%
+  select(model, year, elpd_diff, se_diff) %>%
+  filter(elpd_diff != 0) %>%
+  filter(year %in% c(2002, 2006, 2010, 2014, 2018)) %>%
+  rowwise() %>%
+  mutate(samples = list(rnorm(n = 1e5, 
+                              mean = elpd_diff * -2, # for linear advantage
+                              sd   = se_diff *  2)),
+         year = paste0("AD", year)
+         ) %>%
+  select(year, samples) %>%
+  group_by(year) %>%
+  unnest(samples) %>% 
+  mutate(sample_id = c(1:1e5)) %>%
+  pivot_wider(id_col = "sample_id", names_from = "year", values_from = "samples") %>%
+  select(-sample_id) -> d1
+
+# "fix" d1
+colnames(d1) <- rev(colnames(d1))
 
 color_scheme_set("blue")
-p5 <- mcmc_intervals(x = d1,
+
+p6 <- mcmc_intervals(x = d1,
   pars = colnames(d1),
   prob = 0.0,
-  prob_outer = .90
+  prob_outer = .95
   ) +
   scale_x_continuous(
-    limits = c(-10, 40),
-    labels = function(x) {x*10}) +
+    limits = c(-200, 150)
+    ) +
   geom_vline(aes(xintercept = 0), color = "grey90") +
@@
     subtitle = "Satisfaction",
-    x = "")
-```
-
-```{r plot_waic_gh9}
-waic_result <- read_rds("../results/waic_comparisons_gh9_percentiles.RDS")
-
-d1 <- data.frame(fy2018 = rnorm(n = 1e5, 
-                                mean = waic_result[10, 1] * 2, 
-                                sd   = waic_result[10, 2] *  2)/10,
-                 fy2014 = rnorm(n = 1e5, 
-                                mean = waic_result[8, 1] * 2, 
-                                sd   = waic_result[8, 2] *  2)/10,
-                 fy2010 = rnorm(n = 1e5, 
-                                mean = waic_result[6, 1] * 2, 
-                                sd   = waic_result[6, 2] *  2)/10,
-                 fy2006 = rnorm(n = 1e5, 
-                                mean = waic_result[4, 1] * 2, 
-                                sd   = waic_result[4, 2] *  2)/10,
-                 fy2002 = rnorm(n = 1e5, 
-                                mean = waic_result[2, 1] * 2, 
-                                sd   = waic_result[2, 2] *  2)/10
-                 )
-
-color_scheme_set("red")
-p6 <- mcmc_intervals(x = d1,
-           pars = colnames(d1),
-           prob = 0.00,
-           prob_outer = 0.90
-           ) +
-  scale_x_continuous(
-    limits = c(-10, 40),
-    labels = function(x) {x*10}) +
-  geom_vline(aes(xintercept = 0), color = "grey90") +
-  theme_test() +
-  labs(
-    subtitle = "Happiness",
-    x = "WAIC difference (90% interval)")
+    x = "WAIC difference (95% interval)")
 ```
@@
 
-##### Figure 3. Posterior distributions of &omega; (90% distribution)  
+##### Figure 3. Posterior distributions of &omega; (95% distribution)  
 ```{r figure_3, fig.width = 6, fig.height = 4.5, dpi=300}
-hifdip_gh9_fixed_fit <- read_rds("../results/hifdip_gh9_fixed_full_covs.RDS")
+path_to_fits <- list.files(
+  path = "../results",
+  pattern = '^fit_cp_hifdip_gh9_cov_.*.RDS$',
+  full.names = TRUE
+)
 
-as.data.frame(hifdip_gh9_fixed_fit) %>%
-  select(b_alpha_Intercept, b_alpha_year2006, b_alpha_year2010, 
-         b_alpha_year2014, b_alpha_year2018) %>%
-  transmute(
-    fy2002 = b_alpha_Intercept,
-    fy2006 = mean(b_alpha_Intercept) + b_alpha_year2006,
-    fy2010 = mean(b_alpha_Intercept) + b_alpha_year2010,
-    fy2014 = mean(b_alpha_Intercept) + b_alpha_year2014,
-    fy2018 = mean(b_alpha_Intercept) + b_alpha_year2018
-  ) %>%
-  mutate_all(.funs = ~brms::inv_logit_scaled(., 0, 150)*10) -> posteriors
+posterior_alphas <- list()
+for (pathtofile in path_to_fits) {
+  df <- read_rds(pathtofile) %>%
+    as.data.frame() %>%
+    select(b_alpha_Intercept)
+
+  colnames(df) <- paste0("param_", substring(pathtofile, 34, 37))
+
+  posterior_alphas <- bind_cols(posterior_alphas, df)
+}
+
+posterior_alphas %>%
+  mutate_all(.funs = ~brms::inv_logit_scaled(., 0, 100)*10) -> posterior_omegas
+
+posteriors_trim <- posterior_omegas %>%
+  mutate(
+    param_2010 = (param_2010 + param_2011)/2,
+    param_2003 = (param_2002 + param_2004)/2)
+
+# select posteriors for plotting
+posteriors_select <- posterior_omegas %>%
+  select(param_2002, param_2006, param_2010, param_2014, param_2018)
 
@@
 mcmc_areas_ridges(
-  posteriors,
-  pars = rev(colnames(posteriors)),
-  prob = 0.90, # 80% intervals
-  prob_outer = .99, # 99%
-  point_est = "mean"
+  posteriors_select,
+  pars = rev(colnames(posteriors_select)),
+  prob = 0.95, 
+  prob_outer = .99 
   ) +
   theme_test() +
-  geom_vline(aes(xintercept = mean(posteriors$fy2002)), color = "gray", linetype = 3) +
-  labs(
-    x = "Household wealth ($000s)")
+  geom_vline(aes(xintercept = median(posteriors_select$param_2002)), 
+             color = "gray", 
+             linetype = 3) +
+  labs(x = "Household income ($000s)")
 ```
@@
 
-Figure 3 shows the change point between happiness and household wealth (the _cost of happiness_) credibly increased between 2002 and 2018.
-Table 2 below compares the change in median and mean wealth for each year with the cost of happiness.
-It presents the percentage of people in each year whose household wealth exceeded the cost point of happiness.
+#### Wealth and happiness
 
-<br>
+##### Figure 4. Estimated happiness (SD units) at $50K/yr and $80K/yr from 2002-2018 
+```{r figure_4, fig.width=7, fig.height=5, dpi=300}
+predicted_fits <- list()
+for (pathtofile in path_to_fits) {
+  fit <- read_rds(pathtofile)
 
-```{r, table_2, ft.align="left"}
-happywealth_covs %>%
-  select(year, xwaveid, adj_hifdip) %>%
+  ans <- model.frame(fit) %>%
+    summarise_all(median)
+
+  newdat = data.frame(
+    dollars = c(5, ans[, "dollars"], 8),
+    age = rep(ans[, "age"], 3),
+    age_sq = rep(ans[, "age_sq"], 3),
+    male = rep(ans[, "male"], 3), 
+    grad = rep(ans[, "grad"], 3)
+  )
+
+  df <- fitted(fit, newdata = newdat) %>% 
+    as_tibble() %>%
     mutate(
-    year = paste0("fy", year),
-    adj_hifdip = adj_hifdip/1000) %>%
-  left_join(
-    gather(
-      summarise_all(posteriors, list(median)), year, change_point
-    )) %>%
-  group_by(year) %>%
-  summarise(
-    `median ($000s)` = round(median(adj_hifdip), 2),
-    `mean ($000s)` = round(mean(adj_hifdip), 2),
-    `change point ($000s)` = round(mean(change_point), 2),
-    `%` = round(mean(adj_hifdip > mean(change_point))*100, 1)
-  ) %>%
-  flextable() %>%
-  set_caption("Table 2. Household wealth, change points and percentile") %>%
-  autofit()
+      Dollars = c("$50K", "\u03bc", "$80K"),
+      year = as.numeric(substring(pathtofile, 34, 37)))
+
+  predicted_fits <- bind_rows(predicted_fits, df)
+}
+
+predicted_fits %>%
+  select(Estimate, Dollars, year) %>%
+  filter(Dollars %in% c("$80K", "$50K")) %>%
+  spread(Dollars, Estimate) %>%
+  mutate(`$80K - $50K (∆)` = `$80K` - `$50K`) %>%
+  gather(income, happiness, -year) %>%
+  mutate(income = fct_relevel(income, "$50K", "$80K")) %>%
+  ggplot(aes(x = year, y = happiness, color = income)) +
+  geom_point(size = 4, alpha = 0.25) +
+  geom_smooth(method = "loess", se = F, span = 3) +
+  facet_wrap(~income) +
+  labs(caption = "Source:     HILDA, University of Melbourne
+                 Predicted happiness scores (SD units) from piecewise regression on real household income", 
+       x = "", y = "") -> p
+
+source("ft_themes.r")
+
+p +
+  ft_theme() +
+  ggthemes::scale_colour_wsj() +
+  theme(legend.position = "none",
+        plot.title = element_text(size = 19),
+        strip.text = element_text(size = 12))
+```
+
+```
+Figure 4 legend: The difference (∆) in happiness between individuals with a household income of $50K per year and a household income of $80K per year has increased since 2009.
 ```
@@
 
-Table 2 shows the cost of happiness increased faster than the median or mean household wealth between 2002 and 2018.
-As a result, a smaller percentage of people in 2018 achieved a level of financial security on which their happiness was no longer dependent on their wealth, relative to 2002.
+Figure 5 presents house income levels weighted for the Australian population (by age, sex, marital status, labour force participation and region).
+It shows the cost of happiness increased faster than household income between 2002 and 2018.
+As a result, a smaller percentage of the Australian population in 2018 achieved a level of financial security on which their happiness was no longer dependent on their income, relative to 2002.
+
+<br>
+
+##### Figure 5.  
+```{r figure_5, fig.width=7, fig.height=6, dpi=300}
+# Tabulate against median income
+happywealth %>%
+  filter(!student) %>%
+  filter(!top_hifdip) %>%
+  select(year, xwaveid, income = hh_disp_inc_eq) %>%
+  filter(year %in% 2002:2018) %>%
+  mutate(
+    year = paste0("param_", year),
+    income = income/1000) %>%
+  left_join(
+    gather(
+      summarise_all(posteriors_trim, 
+                    list(median)), 
+      year, 
+      change_point
+    )) %>%
+  mutate(
+    year = extract_numeric(year)
+    ) %>%
+  group_by(year) %>%
+  summarise(
+    `Median household income\n ($000s)` = round(median(income), 
+                                                2),
+    `Cost of happiness\n ($000s)` = round(mean(change_point), 
+                                          2),
+    `Individuals affording happiness\n (Population %)` = round(mean(income > mean(change_point))
+                                                           *100,
+                                                           1)
+    ) %>%
+  gather("key", "value", -year) %>%
+  mutate(
+    key = paste0(key, "\n"),
+    key = fct_relevel(key, "Median household income\n ($000s)\n")) -> df.plot
+
+df.plot %>%
+  select(key) %>% distinct() %>%
+  mutate(string = c("post-GFC\n(2009 onwards)", "", "")) -> gfc.note
+
+df.plot %>%
+  filter(year %in% c(min(year), 2018)) %>%
+  mutate(
+    val = as.character(
+      if_else(str_detect(key, "afford"),
+              paste0(round(value), "%"),
+              paste0("$", round(value), "K"))),
+    vpos = if_else(val %in% c("$49K", "$72K", "41%"), -0.75, 1.65),
+    hpos = if_else(year == min(year), 0.3, 0.8)
+    ) -> df.label
+
+ggplot(df.plot, aes(x = year, y = value, color = key)) +
+  # geom_point() +
+  geom_line(aes(group = key), size = 2, alpha = 0.5) +
+  # geom_smooth(aes(group = key), se = F, span = .5) +
+  labs(title = "Wealth & happiness in Australia between 2002-2018",
+       caption = "Source: HILDA sample, University of Melbourne
+                  Real household income (modified OECD scale) weighted for the Australian population",
+       x = "  ", y = "") +
+  # geom_vline(aes(xintercept = 2009), color = "red", linetype = 3) +
+  # geom_text(data = gfc.note, aes(x = 2009.5, y = 57, label = string),
+  #           colour = "red", size = 3, hjust = "left") +
+  geom_text(data = df.label,
+            aes(label = val, vjust = vpos, hjust = hpos)) +
+  expand_limits(y = c(22, 73)) +
+  facet_wrap(~key) -> p
+
+p.econ <- p + 
+  labs(subtitle = "  ") +
+  ggthemes::scale_color_economist() +
+  ggthemes::theme_economist(base_family="Verdana", base_size = 8) +
+  theme(legend.position = "none",
+        strip.text = element_text(size = 10)
+        )
+p.econ
+```
+
+```
+Figure 5 legend: Real household income has stagnated in Australia since 2009 (post GFC) while the cost of happiness has increased.
+Consequently fewer Australians have a household income exceeding the cost of happiness in 2018.
+```
 

### 2020-12-05 src/draft0.Rmd
minor update to remove the term "wealth"

@@
 library(flextable)
+library(gtsummary)
 library(bayesplot)
@@
 # https://www.pc.gov.au/research/completed/rising-inequality/3_HILDA_variables.R
-# more people could a achieve a level of wealth at which happiness no longer increased
+# more people could a achieve a level of income at which happiness no longer increased
 ```
@@
 
-**keywords:** Subjective wellbeing, household wealth, HILDA
+**keywords:** Subjective wellbeing, household income, HILDA
 
@@
 While the location of this change point has been determined, and the cost of happiness reportedly ranges between USD\$60 to $95K, there has been no investigation as to whether the cost of happiness has increased or decreased over time.
-We tested the relationship between wealth and both happiness and life satisfaction using household economic data from Australia between 2002-2018.
+We tested the relationship between income and both happiness and life satisfaction using household economic data from Australia between 2002-2018.
 We discovered the cost of happiness has increased over those 17 years faster than inflation (i.e., cost of living).
@@
 A fundamental question for psychology and economics is just how much wellbeing does a dollar buy?
-Increasing wealth is commonly associated with increasing wellbeing, however a point at which wealth no longer increases subjective wellbeing has also been widely observed [@clark2008relative; @dolan2008we; @easterlin1974growth].
+Increasing income is commonly associated with increasing wellbeing, however a point at which income no longer increases subjective wellbeing has also been widely observed [@clark2008relative; @dolan2008we; @easterlin1974growth].
 Given that a central goal of nations and governments is to improve wealth under the assumption that wealth always increases wellbeing, challenges to this notion have far reaching consequences [@frijters2020happy].
@@
 For instance, we have recently reported that positive life events, such as a major financial windfall, have a greater impact on an individual’s satisfaction than their happiness [@kettlewell2020differential].
-While the distinct effect of wealth on satisfaction and happiness was observed within individuals, the distinct effects of wealth have also been observed across individuals.
-For instance, Kahneman & Deaton [-@kahneman2010high] showed that self-reported levels of happiness increased with household wealth up to a point ($75,000).
+While the distinct effect of money on satisfaction and happiness was observed within individuals, the distinct effects of money have also been observed across individuals.
+For instance, Kahneman & Deaton [-@kahneman2010high] showed that self-reported levels of happiness increased with household income up to a point (\$75,000).
 But after that, increasing amounts of money had no further effect on happiness.
-They also reported that life satisfaction continued to increase with wealth beyond $75,000.
-Indeed, the difference between the two questions: “How satisfied are you with your life?” and “How happy are you these days?” has been identified as a crucial mediating factor in a meta-analysis of 111 studies on wealth and wellbeing [@howell2008relation; also @veenhoven2006rising].
-Results such as these have provided a more nuanced view among psychologists and (some!) economists about the relationship between wealth and wellbeing;  namely that wealth is more strongly related to satisfaction than to happiness.
+They also reported that life satisfaction continued to increase with income beyond \$75,000.
+Indeed, the difference between the two questions: “How satisfied are you with your life?” and “How happy are you these days?” has been identified as a crucial mediating factor in a meta-analysis of 111 studies on money and wellbeing [@howell2008relation; also @veenhoven2006rising].
+Results such as these have provided a more nuanced view among psychologists and (some!) economists about the relationship between money and wellbeing; namely that money is more strongly related to satisfaction than to happiness.
 
 The distinction between satisfaction and happiness are increasingly relevant to governments and policy-makers due to the growing recognition that increasing wealth does not necessarily lead to improvements in wellbeing [@clark2018four; @frijters2020happy].
-If wealth no longer improves wellbeing, then the maximization of wealth may no longer be a legitimate goal of government.
-Fundamentally, the existence of a change point between wealth and happiness reveals an unacknowledged inequality in the distribution of wellbeing in the economy.
-Of concern is the point at which wealth produces no further increases in happiness – that is, the change point, or cost of happiness.
+If money no longer improves wellbeing, then the maximization of wealth may no longer be a legitimate goal of government.
+Fundamentally, the existence of a change point between income and happiness reveals an unacknowledged inequality in the distribution of wellbeing in the economy.
+Of concern is the point at which income produces no further increases in happiness – that is, the change point, or cost of happiness.
 The cost of happiness represents the point at which material wealth stops driving inequalities in the distribution of happiness in the economy, where lower cost-points represent more equitable distributions of happiness.
@@
 
-We used household economic panel data from Australia (HILDA) to provide the first investigation of whether changes in wealth and wellbeing have changed the cost of happiness over the last 17 years (2002-2018).
-HILDA provides a representative sample of households in Australia with detailed measurements of wealth and subjective wellbeing in the same sample, which makes it an excellent data source to investigate the present question.
-We distinguished between satisfaction and happiness as different components of subjective wellbeing, and evaluated how each varies with household wealth.
-After adjusting for age, gender, and education level, we confirmed that happiness and satisfaction have distinct relationships with increasing wealth, but the cost of happiness has increased in real dollar terms since 2002.
+We used household economic panel data from Australia (HILDA) to provide the first investigation of whether changes in income and wellbeing have changed the cost of happiness over the last 17 years (2002-2018).
+HILDA provides a representative sample of households in Australia with detailed measurements of income and subjective wellbeing in the same sample, which makes it an excellent data source to investigate the present question.
+We distinguished between satisfaction and happiness as different components of subjective wellbeing, and evaluated how each varies with household income.
+After adjusting for age, gender, and education level, we confirmed that happiness and satisfaction have distinct relationships with increasing income, but the cost of happiness has increased in real dollar terms since 2002.
 
@@
 
-#### Wealth
-Our indicator of wealth and economic security was household income.
-Household income better represents economic security, since members of the same household share expenses as well as risks; i.e., they can provide a direct and immediate support network when financial shocks occur.
+#### Income
+Our indicator of income and economic security was household income.
+Household income better represents economic security than personal income, since members of the same household share expenses as well as risks; i.e., they can provide a direct and immediate support network when financial shocks occur.
 The other major studies also used household after-tax income as the indicator of wealth and economic security [e..g, @kahneman2010high; @jebb2018happiness], and so we follow the same standard here as well.
@@
 
-We modelled the relationship between wealth and each wellbeing variable (happiness and satisfaction) using a simple **linear model** and a **piecewise model** (broken-stick).
-The piecewise model was chosen as the simplest extension of a linear model which can identify a change point (inflection) between wellbeing and wealth.
-The location of the change point was a free parameter which revealed where wellbeing no longer increased at a uniform rate with wealth.
-We then compared the linear model against the piecewise model to determine if a change point existed in any year between household wealth and each wellbeing variable (see Model Selection).
+We modelled the relationship between income and each wellbeing variable (happiness and satisfaction) using a simple **linear model** and a **piecewise model** (broken-stick).
+The piecewise model was chosen as the simplest extension of a linear model which can identify a change point (inflection) between wellbeing and income.
+The location of the change point was a free parameter which revealed where wellbeing no longer increased at a uniform rate with income.
+We then compared the linear model against the piecewise model to determine if a change point existed in any year between household income and each wellbeing variable (see Model Selection).
 Finally, where a change point existed, we determined the location of the change point for that year (see Parameter Estimation).
@@
 
-Where $X_i$ was an individual's household wealth (\$) as well as other covariates (age, age^2^, sex, education), and $y_i$ was an individual's wellbeing.
+Where $X_i$ was an individual's household income (\$) as well as other covariates (age, age^2^, sex, education), and $y_i$ was an individual's wellbeing.
 
@@
 
-The piecewise model was a simple extension of this to include a free parameter to represent the changepoint in wealth ($\omega$) as well as the slope before the change point ($\beta_1$) and after the change point ($\beta_2$):
+The piecewise model was a simple extension of this to include a free parameter to represent the changepoint in income ($\omega$) as well as the slope before the change point ($\beta_1$) and after the change point ($\beta_2$):
 
@@
 
-Where $x_i$ was an individual's household wealth, and $X_i$ were covariates for age, age^2^, sex and education.
+Where $x_i$ was an individual's household income, and $X_i$ were covariates for age, age^2^, sex and education.
 
@@
 The above models estimated population-level effects separately for each year (*t* = 2002, 2006, 2010, 2014 and 2018).
-Because we were interested in the location of the change point between wealth and wellbeing that existed across individuals within each year, we ignored the panel design of HILDA because the dependency between observations of the same person across years was orthogonal to our effects of interest.
-We specified weakly informed priors for each *&beta;*, and a uniform prior over the restricted range of wealth values for *&omega;*.
+Because we were interested in the location of the change point between income and wellbeing that existed across individuals within each year, we ignored the panel design of HILDA because the dependency between observations of the same person across years was orthogonal to our effects of interest.
+We specified weakly informed priors for each *&beta;*, and a uniform prior over the restricted range of income values for *&omega;*.
 
 _Model Selection_  
-To determine whether wellbeing was a linear or non-linear (e.g., piecewise) function of wealth, we compared the linear and piecewise model posterior probabilities using the [Widely Applicable Information Criterion (WAIC)](https://bookdown.org/ajkurz/Statistical_Rethinking_recoded/overfitting-regularization-and-information-criteria.html#the-problem-with-parameters){target="_blank"}.
+To determine whether wellbeing was a linear or non-linear (e.g., piecewise) function of income, we compared the linear and piecewise model posterior probabilities using the [Widely Applicable Information Criterion (WAIC)](https://bookdown.org/ajkurz/Statistical_Rethinking_recoded/overfitting-regularization-and-information-criteria.html#the-problem-with-parameters){target="_blank"}.
 The WAIC is the log-posterior predictive density plus a penalty proportional to the variance in the posterior distribution.
@@
 _Parameter Estimation_  
-To determine the location of the change point (*&omega;*) between wellbeing and wealth, we modelled the relationship between wealth and wellbeing across individuals using the piecewise model described above, and sampled the posterior probability of *&omega;* over 4000 interations.
+To determine the location of the change point (*&omega;*) between wellbeing and income, we modelled the relationship between income and wellbeing across individuals using the piecewise model described above, and sampled the posterior probability of *&omega;* over 4000 interations.
 The complete posterior distribution of *&omega;* for each year is presented along with the expected value (mean).
@@
 
-#### Wealth and happiness
+#### Income and happiness
 
@@
   # geom_smooth(aes(group = key), se = F, span = .5) +
-  labs(title = "Wealth & happiness in Australia between 2002-2018",
+  labs(title = "Income & happiness in Australia between 2002-2018",
        caption = "Source: HILDA sample, University of Melbourne
@@
 
-We found the relationship between subjective wellbeing and household income was positive, but happiness and satisfaction had different (positive) relationships: Satisfaction increased linearly with income, while happiness increased rapidly up to a point after which further increments in income produced less change – this confirms the distinct effects of wealth in previous research [e.g., @howell2008relation; @kahneman2010high], and contributes rare evidence from the same sample.
+We found the relationship between subjective wellbeing and household income was positive, but happiness and satisfaction had different (positive) relationships: Satisfaction increased linearly with income, while happiness increased rapidly up to a point after which further increments in income produced less change – this confirms the distinct effects of money in previous research [e.g., @howell2008relation; @kahneman2010high], and contributes rare evidence from the same sample.
 Furthermore, we report here for the first time that the change point between household income and happiness increased over time between 2002 and 2018, faster than inflation or the median household income.
 
-We refer to the change point after which increases in wealth no longer produce similar increases in happiness as the cost of happiness.
+We refer to the change point after which increases in income no longer produce similar increases in happiness as the cost of happiness.
 After this point, happiness is no longer as dependent on household income, and the economic security it represents.
@@
 An implication of the increasing cost of happiness over the last sixteen years is that wealth inequality is driving increasing _inequities_ in wellbeing.
-In 2002, the cost point of happiness represented a 9% increase over median wealth, while in 2018 it represented a 42% increase over median wealth.
+In 2002, the cost point of happiness represented a 9% increase over median income, while in 2018 it represented a 42% increase over median income.
 This also represents a reduction from 44% to 26% in the proportion of people who have achieved a level of financial security beyond which their happiness no longer depends.
@@
 
-Australia has low levels of wealth disparity relative to other OECD countries, and the Gini coefficient has not changed a great deal between 2002 and 2018 in the HILDA dataset [@australian2018rising].
+Australia has low levels of income disparity relative to other OECD countries, and the Gini coefficient has not changed a great deal between 2002 and 2018 in the HILDA dataset [@australian2018rising].
 A stable Gini coefficient shows income inequality has remained steady over the time period, and our results do not conflict with this conclusion.
@@
 It may be that differences in the sparsity of the high income data in HILDA (for which extremely high income households are masked) explain the differences here.
-However including these individuals in our analysis did not change the linear effect between wealth and satisfaction observed here, and the weighted topcodes should have biased the result towards a change point.
-We also note a linear effect between wealth and satisfaction is consistent with the majority of earlier literature [@howell2008relation; @stevenson2013subjective].
+However including these individuals in our analysis did not change the linear effect between income and satisfaction observed here, and the weighted topcodes should have biased the result towards a change point.
+We also note a linear effect between income and satisfaction is consistent with the majority of earlier literature [@howell2008relation; @stevenson2013subjective].
 

### 2020-12-05 src/wdraft0.Rmd
minor update to remove the term "wealth"

@@
 
-**keywords:** Subjective wellbeing, wealth, household income, HILDA
+**keywords:** Subjective wellbeing, household income, HILDA
 
@@
   labs(caption = "Source:     HILDA, University of Melbourne
-                 Predicted happiness scores (SD units) from piecewise regression on real household income", 
+                 Predicted happiness scores (SD unit) from piecewise regression on real household income", 
        x = "", y = "") -> p

### 2020-12-05 src/figure 1.r

@@ -156,7 +156,7 @@ p <- dat %>%
     key = fct_relevel(key, "Median household income \n($000s)\n")) %>%
   ggplot(aes(x = year, y = value, color = key)) +
   geom_line(aes(group = key), size = 2, alpha = 0.5) +
-  labs(title = "Wealth & wellbeing in Australia between 2001-2018",
+  labs(title = "Income & wellbeing in Australia between 2001-2018",
        caption = "Source: HILDA, Melbourne University
                   Household equivalised income per person (modified OECD scale)",
        x = "  ", y = "") +

### 2020-12-05 src/model_fitting.R

@@ -1,0 +1,93 @@
+#### The Increasing Cost of Happiness ####
+# RW Morris, N Kettlewell, N Glozier
+# 
+# Model specification and fitting
+# Dec 4th 2020
+# 
+#### libraries ####
+library(dplyr)
+library(Rcpp)
+library(brms)
+options(mc.cores = parallel::detectCores())
+ncores = parallel::detectCores()
+
+#### Import ####
+happywealth <- readRDS("../data/happywealth.rds")
+
+#### Fit the model ####
+#
+# This model was suggested at: 
+# https://discourse.mc-stan.org/t/piecewise-linear-mixed-models-with-a-random-change-point/5306/5
+# 
+# Model definition
+# b0 is the intercept denoting the expected value of wellbeing (y) at the changepoint
+# b1 is the linear slope before the changepoint
+# b2 is the linear slope after the changepoint
+# alpha is the coefficient of wealth (x) at the changepoint
+
+bfixed <- bf(
+  y ~ b0 + b1 * (dollars - omega) * step(omega - dollars) + 
+    b2 * (dollars - omega) * step(dollars - omega) + b3*grad + b4*age + b5*age_sq + b6*male,
+  b0 + b1 + b2 + b3 + b4 + b5 + b6 + alpha ~ 1,  # fixed intercepts for each year
+  # Sigmoid transform to keep omega within the range of 0 to x
+  nlf(omega ~ inv_logit(alpha) * 100),
+  nl = TRUE
+)
+
+# priors                                      I tried tighter priors:
+bprior <- prior(normal(0, 3), nlpar = "b0") + # normal(0, 1) did not converge Sep 6th
+  prior(normal(0, 3), nlpar = "b1") +         # normal(0, 1) did not converge Sep 6th
+  prior(normal(0, 3), nlpar = "b2") +         # normal(0, 1) did not converge Sep 6th
+  prior(normal(0, 3), nlpar = "b3") +  
+  prior(normal(0, 3), nlpar = "b4") +  
+  prior(normal(0, 3), nlpar = "b5") +  
+  prior(normal(0, 3), nlpar = "b6") +
+  prior(normal(0, 3), nlpar = "alpha")      # normal(0, 3) with init = "0" 12/09/2020
+                                            # normal(0, sqrt(3)) with default init didn't converge
+
+#### Piece-wise fits ####
+for (i in c(2001:2018)) {
+  
+  #### Preprocessing ####
[…]

### 2021-01-12 src/draft0.Rmd
Added population weights and residual plots

@@
 | Draft      | `r format(Sys.time(), '%d %B, %Y')` |
-| Word count | 3767                                |
+| Words      | 4460                                |
 | Tables     | 1                                   |
@@
 The accepted view among economists and psychologists is that money and happiness increase together up to a point, after which there is little further gain from increasing wealth.
-While the location of this change point has been determined, and the cost of happiness reportedly ranges between USD\$60 to $95K, there has been no investigation as to whether the cost of happiness has increased or decreased over time.
+While the location of this change point has been determined, and the cost of happiness reportedly ranges between USD\$60K to $95K, there has been no investigation as to whether the cost of happiness has increased or decreased over time.
 We tested the relationship between income and both happiness and life satisfaction using household economic data from Australia between 2002-2018.
 We discovered the cost of happiness has increased over those 17 years faster than inflation (i.e., cost of living).
-This result shows that inequalities in wealth may be driving inequities in happiness and wellbeing, with implications for health and recent government policy-goals to monitor and improve wellbeing.
+This result shows that inequalities in income may be driving inequities in happiness and wellbeing, with implications for health and recent government policy-goals to monitor and improve wellbeing.
 (150 words)
@@
 Increasing income is commonly associated with increasing wellbeing, however a point at which income no longer increases subjective wellbeing has also been widely observed [@clark2008relative; @dolan2008we; @easterlin1974growth].
-Given that a central goal of nations and governments is to improve wealth under the assumption that wealth always increases wellbeing, challenges to this notion have far reaching consequences [@frijters2020happy].
+Given that a central goal of nations and governments is to improve income under the assumption that income always increases wellbeing, challenges to this notion have far reaching consequences [@frijters2020happy].
 
@@
 
-The distinction between satisfaction and happiness are increasingly relevant to governments and policy-makers due to the growing recognition that increasing wealth does not necessarily lead to improvements in wellbeing [@clark2018four; @frijters2020happy].
+The distinction between satisfaction and happiness are increasingly relevant to governments and policy-makers due to the growing recognition that increasing income does not necessarily lead to improvements in wellbeing [@clark2018four; @frijters2020happy].
 If money no longer improves wellbeing, then the maximization of wealth may no longer be a legitimate goal of government.
@@
 Household income better represents economic security than personal income, since members of the same household share expenses as well as risks; i.e., they can provide a direct and immediate support network when financial shocks occur.
-The other major studies also used household after-tax income as the indicator of wealth and economic security [e..g, @kahneman2010high; @jebb2018happiness], and so we follow the same standard here as well.
+The other major studies also used household after-tax income as the indicator of income and economic security [e..g, @kahneman2010high; @jebb2018happiness], and so we follow the same standard here as well.
 The ‘real household annual disposable income’ was calculated from the combined income of all household members after receipt of government pensions and benefits and deduction of income taxes in the financial year ended 30th June of the year of the wave (e.g., 2002 in wave 2).
@@
 _Covariates_  
-Age (and age^2^), gender, and education were included as covariates.
+Age (and age^2^), gender, and university graduate education were included as covariates.
+Cross-sectional population weights for Australia provided by the University of Melbourne in each year were also included as a covariate to adjust for differences in the sample representativeness according to sex by broad age, marital status, region, and labour force status.
 Full-time students were removed, as well as individuals with an annual household disposable income that was indicated as _topcoded_ by the University of Melbourne (topcoding occurs to ensure privacy of high wealth individuals).
@@
 
-##### Figure 1. Household income and satisfaction (blue) and happiness (red)
+##### Figure 1. Household income and happiness (red) and satisfaction (blue)  
 ```{r losat_gh9_hifdip}
@@
 ```{r figure_1, fig.height=10, fig.width=9, dpi=300}
-p1 / p2 / p3 / p4
+p3 / p4 / p1 / p2
 ```
@@
 Figure 1 legend: Posterior predictive distributions (95% credible intervals) showing the model fit from regressions of wellbeing on household income, overlaid on summary data points (income deciles).
-Wellbeing was measured as life satisfaction (blue) or happiness (red).
+Wellbeing was measured as happiness (red) or life satisfaction (blue).
 The total number of individuals contributing to each year are shown (n).
@@
 
-Figure 1 shows the relationship between household income and satisfaction appeared relatively linear, as the results of the piecewise regression indicated the inflection of the change point could be increasing (e.g., 2002, 2006) as well as decreasing (e.g., 2010, 2014, 2018).
-By contrast, a decreasing change point was evident in each year of happiness on income.
-The results of a formal comparison between linear and piecewise models is provided in Model Selection below.
-However note that the change point between income and happiness appeared to shift rightwards over time (bottom row).
-(We can also see the linear relationship between household income and each wellbeing variable became slightly weaker over time, i.e., less steep)
+Figure 1 shows the relationship between household income and happiness (red) and satisfaction (blue).
+A linear fit is shown in the 1st and 3rd row for happiness and satisfaction respectively.
+The nonlinear relationship between happiness and income is shown in the 2nd row and it was negatively inflected, with a consistently decreasing change point in each year.
+By contrast the nonlinear relationship between satisfaction and income shown in the 4th row appeared to be positively inflected in 2002 and 2006, and negatively inflected 2014 and 2018.
+The relationship between satisfaction and income was relatively linear (no apparent change point) in 2014.
 
-<br>
+<br><br>
 
@@
 
-We compared the posterior evidence for a linear relationship between wellbeing and income with the posterior evidence of a nonlinear (piecewise) relationship (i.e., WAIC~linear~ — WAIC~piecewise~).
+We compared the posterior evidence for a linear relationship between each wellbeing variable and income with the posterior evidence of a nonlinear (piecewise) relationship (i.e., WAIC~linear~ — WAIC~piecewise~).
 Thus a WAIC difference greater than zero indicated evidence for a linear relationship.
@@
 ```
-Figure 2 legend: Differences in posterior evidence for a linear fit over a piecewise fit (WAIC) for satisfaction (blue) and happiness (red).
+Figure 2 legend: Differences in posterior evidence for a linear fit over a piecewise fit (WAIC) for happiness (red) and satisfaction (blue).
 The filled circle indicates the mean of the distribution and the horizontal bars represents the 95% credible interval.
@@
 
+<br>
+
+The posterior evidence from model selection revealed the nonlinear (piecewise) fit of _happiness_, but not _satisfaction_, was credibly superior to a linear fit.
+Figure 2 shows the 95% credible interval of the difference in WAIC values for the linear over nonlinear fits for happiness was below zero with no overlap for each year, indicating consistent support for the nonlinear model.
+By contrast, the WAIC differences for the satisfaction models were generally positive, indicating the superiority of the linear fit for this variable.
+The linear fit for satisfaction was credibly superior (95%) for 2002, 2006, 2010 and 2018, and above but near zero for 2014.
+Overall the posterior evidence suggests that happiness and satisfaction have distinct relationships with household income; satisfaction tends to increase linearly with income, while a change point exists in the relationship between happiness and household income.
+
 <br><br>
-
-Model selection revealed the posterior evidence for the linear fit of satisfaction on income was credibly superior to the nonlinear (piecewise) fit - Figure 2 shows the 90% credible interval of the fit for each year of satisfaction was above zero with no overlap.
-By contrast, a nonlinear fit of happiness on income was generally superior, and the nonlinear fit was credibly superior at 90% for each of three years (2006, 2014, 2018).
-Thus, the posterior evidence indicates happiness and satisfaction have distinct relationships with household income; satisfaction tends to increase linearly with income, while a change point exists in the relationship between happiness and income.
-
-<br>
 
@@
 
-Figure 3 below presents the complete posterior distribution of &omega; over 4000 samples drawn from the piecewise model, representing the location of the change point between happiness and household income in four separate years.
-Shaded areas to the right of the vertical grey dotted line are credibly (95%) larger than 2002.
-The figure shows the change point between happiness and household income (the _cost of happiness_) credibly increased between 2002 and 2018.
+Figure 3 below presents the complete posterior distribution of &omega; over 12,000 samples were drawn from each piecewise model, representing the location of the change point between happiness and household income in four separate years.
+Shaded areas to the right of the vertical grey dotted line are credibly (95%) higher than 2002.
+The figure shows the change point between happiness and household income was credibly higher in the two most recent years (2014 and 2018) relative to 2002.
 
@@
 path_to_fits <- list.files(
-  path = "../results",
-  pattern = '^fit_cp_hifdip_gh9_cov_.*.RDS$',
+  path = "../results/incl_topcodes/with_pweights/",
+  pattern = "^fit_nl_hifdip_gh9_4cov_pwt_.*.RDS$",
   full.names = TRUE
@@
 
-  colnames(df) <- paste0("param_", substring(pathtofile, 34, 37))
+  colnames(df) <- paste0(
+    "param_", 
+    substring(pathtofile, nchar(pathtofile)-7, nchar(pathtofile)-4)
+    )
 
@@
 # select posteriors for plotting
-posteriors_select <- posterior_omegas %>%
+posteriors_select <- posteriors_trim %>%
   select(param_2002, param_2006, param_2010, param_2014, param_2018)
@@
 
-mcmc_areas_ridges(
+mcmc_areas(
   posteriors_select,
@@
   prob = 0.95, 
-  prob_outer = .99 
+  prob_outer = .99,
+  point_est = "median",
   ) +
@@
 ```
-Figure 3 legend: Posterior distribution of the change point parameter representing the location in household income ($) for each year. Shaded area represents the 90% credible region.
+Figure 3 legend: Posterior distribution of the change point parameter representing the location in real household income (2018 dollars) for each year.
+Shaded area represents the 95% credible region and the solid vertical line indicates the expected value (median) of each distribution.
+Dotted line indicates the 2002 change point parameter (median) for comparison.
 ```
 
-<br>
+<br><br>
 
-#### Income and happiness
+#### Income and the cost of happiness
+
+The posterior estimates of the change point between happiness and household income indicates that the relationship between happiness and income has changed between 2002 and 2018.
+One  implication of such a change is the disparity in happiness between income groups has increased.
+This will occur as more people fall below the change point over time and so are subject to the steep region of the function where income and happiness are strongly related.
+However the relationship between income and happiness depends on all the parameters of the function such as the intercept, pre-change slope and the post-change slope rather than solely the change point location.
+So to directly examine the implications of our model for the distribution of happiness, we used the model to estimate the happiness levels for two different income groups: one group for whom the change point increased above their household income over the time period (\$50K/yr); and another group which always remained safely above the change point over the same period ($80K/yr).
+Figure 4 below presents the standardized happiness levels for each (hypothetical) group as well as the difference in happiness between them.
 
@@
     age_sq = rep(ans[, "age_sq"], 3),
-    male = rep(ans[, "male"], 3), 
-    grad = rep(ans[, "grad"], 3)
+    male = as.logical(rep(ans[, "male"], 3)), 
+    grad = as.logical(rep(ans[, "grad"], 3)),
+    pweights = rep(ans[, "pweights"], 3)
   )
@@
       Dollars = c("$50K", "\u03bc", "$80K"),
-      year = as.numeric(substring(pathtofile, 34, 37)))
+      year = as.numeric(
+        substring(pathtofile, nchar(pathtofile)-7, nchar(pathtofile)-4)
+        )
+      )
 
@@
 
+##### Alt Figure 4. Observed happiness (SD units) at $50K/yr and $80K/yr from 2002-2018 
+```{r alt_figure_4, fig.width=7, fig.height=5, dpi = 300}
+happywealth %>%
+  filter(!student) %>%
+  filter(!top_hifdip) %>%
+  select(year, xwaveid, gh9, income = hh_disp_inc_eq, weights) %>%
+  filter(year %in% 2002:2018) %>%
+  mutate(
+    year = paste0("param_", year),
+    income = income/1000) %>%
+  left_join(
+    gather(
+      summarise_all(posteriors_trim, list(median)), 
+      year, 
+      change_point
+    )) %>%
+  mutate(
+    year = extract_numeric(year),
+    income_group = case_when(
+      income > 35 & income < 45 ~ "$50K",
+      income > 65 & income < 75 ~ "$80K"
+    )) %>%
+  group_by(year) %>%
+  mutate(
+    happiness = as.vector(scale(gh9)),
+    ) %>%
+  filter(!is.na(income_group)) %>%
+  group_by(year, income_group) %>%
+  summarise(
+    Happiness = mean(happiness, na.rm = T)
+    ) %>% 
+  ungroup() %>%
+  spread(income_group, Happiness) %>%
+  mutate(`$80K - $50K (∆)` = `$80K` - `$50K`) %>%
+  gather(key, val, `$50K`:`$80K - $50K (∆)`) %>%
+  mutate(key = fct_relevel(key, c("$50K", "$80K"))) %>%
+  ggplot(aes(x = year, y = val, color = key)) +
+    geom_point(size = 4, alpha = 0.25) +
+    geom_smooth(span = 3, se = F) +
+    facet_wrap(~key) +
+    labs(caption = "Source:     HILDA, University of Melbourne
+                 Observed happiness scores (SD unit)", 
+     x = "", y = "") -> p
+
+p +
+  ft_theme() +
+  ggthemes::scale_colour_wsj() +
+  theme(legend.position = "none",
+        plot.title = element_text(size = 19),
+        strip.text = element_text(size = 12))
+```
+
 ```
@@
 
-Figure 5 presents house income levels weighted for the Australian population (by age, sex, marital status, labour force participation and region).
-It shows the cost of happiness increased faster than household income between 2002 and 2018.
-As a result, a smaller percentage of the Australian population in 2018 achieved a level of financial security on which their happiness was no longer dependent on their income, relative to 2002.
+Figure 4 shows the estimated disparity in happiness between two fixed income levels (∆).
+From 2002 to 2009, a hypothetical person with a household income \$80K per year was approximately 0.1 SD units happier than a person with an income of $50K per year.
+However by 2018 the difference had doubled to 0.2 SD units.
+Interestingly this increase was not due to happiness levels increasing among \$80K/year households year-on-year, but rather both income groups had declining happiness levels between 2002-2009 and this decline was arrested and reversed among \$80K/year households after 2009, but not among \$50K/households.
+The increasing difference in happiness between our two income levels shown in the third panel confirms the model fully implies that happiness levels became more inequitable over the period.
+
+A second implication of the increase in the change point between income and happiness is that over time fewer people will enjoy a level of financial security beyond which their happiness no longer depends on income.
+This occurs because the income of fewer people fall above the change point, where the flat (or gradual) region of the function indicates happiness is less strongly determined by income.
+Asuming median income is also not increasing at a faster rate, any increase in the change point should reduce the number of people who fall above it over time.
+Figure 5 below presents median household income levels weighted for the Australian population (by age, sex, marital status, labour force participation and geographical region).
+It also shows the change point between income and happiness increased faster than household income between 2002 and 2018.
+The third panel shows that as a result, a smaller percentage of the Australian population in 2018 had a household income above the changepoint than in 2002.
+This confirms that fewer Australians can afford a level of financial security beyond which their happiness no longer depends on income in 2018 (relative to 2002).
 
@@
 
-##### Figure 5.  
+##### Figure 5. The increasing cost of happiness in Australia 2002-2018 
 ```{r figure_5, fig.width=7, fig.height=6, dpi=300}
@@
   filter(!top_hifdip) %>%
-  select(year, xwaveid, income = hh_disp_inc_eq) %>%
+  select(year, xwaveid, income = hh_disp_inc_eq, weights) %>%
   filter(year %in% 2002:2018) %>%
@@
     gather(
-      summarise_all(posteriors_trim, 
-                    list(median)), 
+      summarise_all(posteriors_trim, list(median)), 
       year, 
@@
   mutate(
-    year = extract_numeric(year)
+    year = extract_numeric(year),
+    over_change = income > change_point,
+    weighted_over = over_change * weights
   ) %>%
@@
   summarise(
-    `Median household income\n ($000s)` = round(median(income), 
-                                                2),
-    `Cost of happiness\n ($000s)` = round(mean(change_point), 
-                                          2),
-    `Individuals affording happiness\n (Population %)` = round(mean(income > mean(change_point))
-                                                           *100,
-                                                           1)
+    `Median household income\n ($000s)` = round(median(income), 2),
+    # `Median household income\n ($000s)` = round(matrixStats::weightedMedian(income, weights)),
+
+    `Cost of happiness\n ($000s)` = round(mean(change_point), 2),
+
+    `Proportion of Australians\n Afford (%)` = round(
+      sum(weighted_over) / sum(weights) *100)
+
     ) %>%
@@
     key = paste0(key, "\n"),
-    key = fct_relevel(key, "Median household income\n ($000s)\n")) -> df.plot
+    key = fct_relevel(key, "Median household income\n ($000s)\n")
+    ) -> df.plot
 
@@
     val = as.character(
-      if_else(str_detect(key, "afford"),
+      if_else(str_detect(key, "Proportion"),
               paste0(round(value), "%"),
               paste0("$", round(value), "K"))),
-    vpos = if_else(val %in% c("$49K", "$72K", "41%"), -0.75, 1.65),
+    vpos = if_else(val %in% c("$49K", "$72K", "42%"), -0.75, 1.65),
     hpos = if_else(year == min(year), 0.3, 0.8)
@@
 ggplot(df.plot, aes(x = year, y = value, color = key)) +
-  # geom_point() +
   geom_line(aes(group = key), size = 2, alpha = 0.5) +
-  # geom_smooth(aes(group = key), se = F, span = .5) +
-  labs(title = "Income & happiness in Australia between 2002-2018",
-       caption = "Source: HILDA sample, University of Melbourne
+  labs(caption = "Source:     HILDA sample, University of Melbourne
                   Real household income (modified OECD scale) weighted for the Australian population",
        x = "  ", y = "") +
-  # geom_vline(aes(xintercept = 2009), color = "red", linetype = 3) +
-  # geom_text(data = gfc.note, aes(x = 2009.5, y = 57, label = string),
-  #           colour = "red", size = 3, hjust = "left") +
   geom_text(data = df.label,
@@
 p.econ <- p + 
-  labs(subtitle = "  ") +
+  labs(title = "Income & happiness in Australia between 2002-2018",
+       subtitle = " ") +
   ggthemes::scale_color_economist() +
@@
         )
-p.econ
+
+p + 
+  ft_theme() +
+  ggthemes::scale_colour_wsj() +
+  theme(legend.position = "none",
+        plot.title = element_text(size = 19),
+        strip.text = element_text(size = 12))
 ```
@@
 Figure 5 legend: Real household income has stagnated in Australia since 2009 (post GFC) while the cost of happiness has increased.
-Consequently fewer Australians have a household income exceeding the cost of happiness in 2018.
+Consequently fewer Australians have a household income which affords them happiness not determined by financial security in 2018.
 ```
@@
 Life satisfaction on the other hand always increased with household income.
-The difference likely reflects the importance of a numerical dollar value (e.g., bank balance, house value) when cognitively appraising one’s life achievements, versus the relevance of that number to our everyday experience of joy and prevailing mood.
+The difference likely reflects the importance of a numerical dollar value (e.g., bank balance, house value) when cognitively appraising one’s life achievements, versus the relevance of that number to our everyday experience of joy and our prevailing mood.
 
-An implication of the increasing cost of happiness over the last sixteen years is that wealth inequality is driving increasing _inequities_ in wellbeing.
-In 2002, the cost point of happiness represented a 9% increase over median income, while in 2018 it represented a 42% increase over median income.
-This also represents a reduction from 44% to 26% in the proportion of people who have achieved a level of financial security beyond which their happiness no longer depends.
-Thus we can see that over time in the last sixteen years the happiness of more people, i.e., their everyday experience of joy and their prevailing mood, has depended on their material wealth.
+An implication of the changing relationship between happiness and income over the last sixteen years is that income inequality may be driving increasing _inequities_ in wellbeing.
+This is particularly clear in Figure 4, where the difference in happiness between \$50K and $80K households represented 0.1 SD units in happiness, however by 2018 the difference had doubled to 0.2 SD units.
+The inequity was also highlighted in Figure 5 where the change point of happiness represented a 9% increase over median income in 2002, while in 2018 it represented a 42% increase over median income.
+This increase relative to median income also represented a reduction from 42% to 24% in the proportion of people who have achieved a level of income beyond which their happiness no longer depends.
+Thus we can see that over the last sixteen years the difference in happiness between fixed income levels has increased, and the proportion of people whose happiness, i.e., their everyday experience of joy and their prevailing mood, no longer depends on their financial security has decreased.
 
@@
 We think this highlights the issue that while traditional measures of wealth and income inequality may be relatively stable and exhibit little change, their impact on wellbeing and health can still vary.
-As focus shifts from traditional wealth indicators towards wellbeing measures, findings such as this may become more prevalent.
+As focus shifts from traditional financial indicators towards health and wellbeing measures, findings such as this may become more prevalent.
 
@@
 
+It is worth noting that other models will of course provide better fits to this data.
+For instance, log-normal models are typically used to model income effects.
+But such models will not provide a clear or distinct change point.
+One implication of this weakness is that the inflection point we identify may be much smoother than implied by our analysis.
+We concede this point but would still contend the location of the inflection point in such models will display a comparable shift over time, if it can be detected with certainty.
+
 As governments and policy-makers begin to focus on wellbeing, it will be critical to understand how traditional economic indicators such as household income, wealth inequality, and consumption interact with wellbeing and health.
 According to @frijters2020happy, coming up with a consensus to translate income into wellbeing features high on the wider wellbeing research agenda.
-Establishing the links between wealth, household income, wellbeing and health, and how inequalities in one drives inequities in the other, will be a critical step in that agenda.
+Establishing the links between wealth, household income, wellbeing and health, and how inequalities in one drives inequities in the other, will be a critical step in the success of that agenda.
 

### 2021-01-12 src/supplementary.Rmd
Added population weights and residual plots

@@
 
-#### Non-linear fits
-
-We tested the relationship between wealth and each wellbeing variable (happiness and satisfaction) using cubic **B-spline models**.
-A cubic B-spline was the most complex model (more degrees of freedom), and included here because it has been recently used to describe the maximum point at which wellbeing no longer increases with wealth [i.e., "satiety", @jebb2018happiness].
-Following Jebb et al [-@jebb2018happiness], we restricted the cubic B-spline model to *n* = 4, 5 and 6 knots and display wealth on a log scale.
-
-<br>
-
-##### Figure S3. Cubic spline of household wealth on satisfaction and happiness
-```{r losat_gh9_hwnwip}
-source("../src/happyfun.r")
-happywealth <- read_rds("../data/happywealth.rds")
-
-p1 <- happywealth %>%
-  filter(hifdip_adj > 0) %>%
-  rename(Satisfaction = losat, `Household wealth` = hifdip_adj) %>% 
-  plot_inflection_cubic(Satisfaction, `Household wealth`) +
-  scale_x_continuous(trans = "log", 
-                     breaks = c(5, 10, 20, 40, 80, 160),
-                     limits = c(2.5, 170))
-
-p2 <- happywealth %>%
-  filter(hifdip_adj > 0) %>%
-  rename(Happiness = gh9, `Household wealth` = hifdip_adj) %>% 
-  plot_inflection_cubic(Happiness, `Household wealth`) +
-  scale_x_continuous(trans = "log", 
-                     breaks = c(5, 10, 20, 40, 80, 160),
-                     limits = c(2.5, 170))
-```
-```{r cubic_plot, fig.height=9, fig.width=11}
-p1 / p2
-```
-
-<br>
-
-The cubic B-spline models failed to reveal a complete satiety point (horizontal inflection) in satisfaction before the highest decile in any year.
-Also note that despite using a log scale, we do not observe a clear, regular logarithmic function with either wellbeing measure.
-Other researchers have argued that wealth represents a quantitative dimension subject to Weber's Law of perception [@kahneman2010high].
-Under this view, similar proportional changes in income or wealth produce similar evaluations of satisfaction or happiness, rather than the absolute change.
-That is, a \$100 increase in wealth will produce a greater impact on someone earning a minimum wage than someone earning many orders of magnitude of that amount.
-However while it is true that wealth clearly represents a quantitative dimension, the effects of wealth observed here represent cross-sectional differences in the population rather than changes experienced by an individual over time.
-It is not clear the extent to which Weber's Law applies to cross-sectional or between-subject differences in quantitative dimensions such as wealth.
-
-<br>
-
 #### Change point model parameter estimates
@@
 4000 samples (post-warmup) were drawn using sampling(NUTS).
-For each parameter, Bulk_ESS and Tail_ESS are effective sample size measures, and Rhat is the potential scale reduction factor on split chains (at convergence, Rhat = 1).
+For each parameter, Bulk_ESS and Tail_ESS are effective sample size (ESS) measures, and Rhat is the potential scale reduction factor on split chains.
+In each case the Rhat apprached 1, indicating convergance.
+The smallest ESS was _n_ = 472, which is greater than a conservative estimate of the sufficient _n_ for estimation (*n* > 300).
 
@@
 
-##### Figure S5. Posterior predictive distribution of marginal effect of wealth by year (95% CI)  
-(not run)
+##### Figure S3. Residual plots for piecewise happiness models
+```{r figure_s3, fig.height=9, fig.width=9}
+# https://www.flutterbys.com.au/stats/tut/tut7.2b.html
+# resid = resid(data.brms)[, "Estimate"]
+# fit = fitted(data.brms)[, "Estimate"]
 
-```{r figure_s2, fig.height = 3, eval=F}
+resid_df <- readRDS("../results/residual_data.RDS")
+
+resid_df %>%
+  group_by(year) %>%
+  slice_sample(n = 1000) %>% # to reduce overplotting
+  ggplot(aes(x = fitteds, y = resids)) +
+    geom_point(alpha = 0.1, size = 0.5) +
+    facet_wrap(~year, scales = "free") +
+    theme_test()
+```
+
+```
+Figure S3 legend: The joint distribution between residual and fitted values shows no common patterns across years and appears clustered around zero in each case.
+```
+
+<br>
+
+Examination of the residual plots from the piecewise models of happiness confirms the residuals were evenly distributed around zero and there was little indication of heterscedascity.
+There was little evidence that the piecewise models of happiness were not appropriate for the data.
+
+<br><br>
+
+#### Non-linear fits
+
+We tested the relationship between wealth and each wellbeing variable (happiness and satisfaction) using cubic **B-spline models**.
+A cubic B-spline was the most complex model (more degrees of freedom), and included here because it has been recently used to describe the maximum point at which wellbeing no longer increases with wealth [i.e., "satiety", @jebb2018happiness].
+Following Jebb et al [-@jebb2018happiness], we restricted the cubic B-spline model to *n* = 4, 5 and 6 knots and display wealth on a log scale.
+
+<br>
+
+##### Figure S4. Cubic spline of household wealth on satisfaction and happiness
+```{r losat_gh9_hwnwip}
+source("../src/happyfun.r")
+happywealth <- read_rds("../data/happywealth.rds")
+
+p1 <- happywealth %>%
+  filter(hifdip_adj > 0) %>%
+  rename(Satisfaction = losat, `Household wealth` = hifdip_adj) %>% 
+  plot_inflection_cubic(Satisfaction, `Household wealth`) +
+  scale_x_continuous(trans = "log", 
+                     breaks = c(5, 10, 20, 40, 80, 160),
+                     limits = c(2.5, 170))
+
+p2 <- happywealth %>%
+  filter(hifdip_adj > 0) %>%
+  rename(Happiness = gh9, `Household wealth` = hifdip_adj) %>% 
+  plot_inflection_cubic(Happiness, `Household wealth`) +
+  scale_x_continuous(trans = "log", 
+                     breaks = c(5, 10, 20, 40, 80, 160),
+                     limits = c(2.5, 170))
+```
+```{r cubic_plot, fig.height=9, fig.width=11}
+p1 / p2
+```
+
+<br>
+
+The cubic B-spline models failed to reveal a complete satiety point (horizontal inflection) in satisfaction before the highest decile in any year.
+Also note that despite using a log scale, we do not observe a clear, regular logarithmic function with either wellbeing measure.
+Other researchers have argued that wealth represents a quantitative dimension subject to Weber's Law of perception [@kahneman2010high].
+Under this view, similar proportional changes in income or wealth produce similar evaluations of satisfaction or happiness, rather than the absolute change.
+That is, a \$100 increase in wealth will produce a greater impact on someone earning a minimum wage than someone earning many orders of magnitude of that amount.
+However while it is true that wealth clearly represents a quantitative dimension, the effects of wealth observed here represent cross-sectional differences in the population rather than changes experienced by an individual over time.
+It is not clear the extent to which Weber's Law applies to cross-sectional or between-subject differences in quantitative dimensions such as wealth.
+
+<br><br>
+
+##### Figure S5. Prior predictive plot
+(Not included)
+
+```{r figure_s5, eval=F}
+# prior_fit <- readRDS("../results/prior_nl_hifdip_losat_cov_2018.RDS")
+
+# Check priors. The fixed estimates should be near zero, the fixed error should
+# be ~3 respectively
+# 
+# summary(prior_fit)
+
+# Extract prior predictions
+# d <- data.frame(predict(prior_fit)) # takes a long time to run
+# prior_predictions <- bind_cols(prior_fit$data, d)
+# write_rds(prior_predictions, "../results/prior_prediction_data.RDS")
+
+prior_predictions <- read_rds("../results/prior_prediction_data.RDS")
+
+prior_predictions %>%
+  mutate(
+      dollar_bin = ntile(dollars, 10)
+      ) %>%
+    group_by(dollar_bin) %>%
+    summarize(
+      wellbeing = mean(Estimate, na.rm=T),
+      wealth = mean(dollars, na.rm=T),
+      n = n()
+    ) -> df.deciles
+
+prior_predictions %>%
+  rename(wellbeing = Estimate,
+         wealth = dollars) -> df.plot
+
+# piecewise model
+p0 <- ggplot(df.deciles, aes(x = wealth, y = wellbeing)) +
+  geom_point() +
+  geom_smooth(data = df.plot,
+              aes(x = wealth, y = wellbeing,
+                  fill = "#C79999", color = "#8F2727"),
+              size = 0.5) +
+  coord_cartesian(xlim = c(0, max(df.deciles$wealth)),
+                  ylim = c(min(df.deciles$wellbeing), max(df.deciles$wellbeing))
+                  ) +
+  theme_test() +
+  theme(legend.position = "none")
+
+p1 <- ggplot(df.deciles, aes(x = wealth, y = wellbeing)) +
+  geom_point() +
+  geom_smooth(data = df.plot,
+              aes(x = wealth, y = wellbeing, fill = "manual"),
+              method = "lm",
+              size = 0.5) +
+  coord_cartesian(xlim = c(0, max(df.deciles$wealth)),
+                  ylim = c(min(df.deciles$wellbeing), max(df.deciles$wellbeing))
+                  ) +
+  theme_test() +
+  theme(legend.position = "none") +
+  scale_fill_manual(values = "#b3cde0")
+
+source("happyfun.R")
+
+p2 <- prior_predictions %>%
+  rownames_to_column(var = "xwaveid") %>%
+  mutate(year = 2018,
+         dollars = dollars*10000) %>%
+  rename(Wellbeing = Estimate, `Household income` = dollars) %>% 
+  plot_inflection_piecewise(Wellbeing, `Household income`, knots = 1)
+
+p3 <- prior_predictions %>%
+  rownames_to_column(var = "xwaveid") %>%
+  mutate(year = 2018,
+         dollars = dollars*10000) %>%
+  rename(Wellbeing = Estimate, `Household income` = dollars) %>% 
+  plot_inflection_cubic(Wellbeing, `Household income`)
+
+p4 <- prior_predictions %>%
+  rownames_to_column(var = "xwaveid") %>%
+  mutate(year = 2018,
+         dollars = dollars*10000) %>%
+  rename(Wellbeing = Estimate, `Household income` = dollars) %>% 
+  plot_linear(Wellbeing, `Household income`) +
+  scale_fill_manual(values = "#b3cde0") +
+  scale_color_manual(values = "#005b96")
+
+p0 + p1
+
+p2 + p3 + p4
+```
+
+<br><br><br>
+
+##### Figure S6. Marginal effect of wealth on satisfaction by year (95% CI)  
+(Not included)
+```{r figure_s6_df, eval=F}
+# Do not run
+
 library(brms)
 
-marginals <- conditional_effects(
-  hifdip_gh9_fixed_fit,
-  effects = c("dollars:year"), # year, dollars, 
-  method = "fitted"
+# import the fits
+nl_losat_paths <- list.files(
+  path = "../results/no_topcodes/",
+  pattern = "fit_nl_hifdip_losat_cov_*",
+  full.names = TRUE
 )
 
-marginals$`dollars:year`$year <- factor(marginals$`dollars:year`$year, 
-                                        levels = c("2002", "2006", "2010", "2014", "2018"))
+list_fits = list()
+for (f in nl_losat_paths) {
+  fit <- readRDS(f)
+  list_fits[str_extract(f, "[[:digit:]]+")] <- list(fit)
+}
 
-plot(marginals, plot=F)[[1]] +
-  xlim(c(0, 30)) +
-  ylim(c(-0.5, 1)) +
-  facet_wrap(~year, nrow=1) +
+# turn each fit into a plot dataframe (this takes a few minutes)
+plot_data <- map_dfr(.x = list_fits, 
+                     .f = function(x) conditional_effects(x)[[1]],
+                     .id = "Year")
+
+saveRDS(plot_data, "../results/nl_marginal_effects_losat.RDS")
+```
+```{r, figure_s6, eval=F}
+readRDS("../results/nl_marginal_effects_losat.RDS") %>%
+  ggplot(aes(x = dollars, y = estimate__, fill = Year)) +
+    geom_ribbon(aes(ymin = lower__, ymax = upper__), alpha = 0.25) +
+    geom_line(aes(color = Year)) +
+    facet_wrap(~Year, scale = "free") +
+    labs(y = "") +
     theme_test() +
-  theme(legend.position = "none") +
-  labs(x = "", y = "")
+    theme(legend.position = "none")
 ```

### 2021-01-12 src/happyfun.R

@@ -59,7 +59,7 @@ plot_linear <- function(.data, swb_col, dollars_col) {
 }
 
 
-plot_inflection_piecewise <- function(.data, swb_col, dollars_col, knots) {
+plot_inflection_piecewise <- function(.data, swb_col, dollars_col, knots = 1) {
   
   swb_colname <- rlang::enexpr(swb_col)
   dollars_colname <- rlang::enexpr(dollars_col)
@@ -103,7 +103,7 @@ plot_inflection_piecewise <- function(.data, swb_col, dollars_col, knots) {
                 aes(x = wealth, y = wellbeing,
                 fill = "manual1", color = "manual2"),
                 method = lm, 
-                formula = y ~ splines::bs(x, df = 2, degree = 1),
+                formula = y ~ splines::bs(x, df = knots+1, degree = 1),
                 size = 0.5) +
     coord_cartesian(xlim = c(0, max(df.deciles$wealth)),
                     ylim = c(min(df.deciles$wellbeing), max(df.deciles$wellbeing))) +
@@ -203,7 +203,6 @@ plot_inflection_cubic <- function(.data, swb_col, dollars_col) {
   
   .data %>%
     select(xwaveid, year, 
-           # swb = enexpr(swb_col), dollars = enexpr(dollars_col)) %>%
            swb = !!swb_colname, dollars = !!dollars_colname) %>%
     filter(swb >= 0) %>%
     filter(dollars >= 0) %>%

### 2021-01-12 src/model fitting.r

@@ -2,7 +2,7 @@
 # RW Morris, N Kettlewell, N Glozier
 # 
 # Model specification and fitting
-# Nov 21st 2020
+# Jan 8th 2021
 # 
 #### libraries ####
 library(dplyr)

### 2021-01-12 src/model selection.r

@@ -1,0 +1,83 @@
+#### Comparing models ####
+# https://bookdown.org/ajkurz/Statistical_Rethinking_recoded/overfitting-
+# regularization-and-information-criteria.html#the-problem-with-parameters
+# 
+# WAIC is a (penalized) estimate of out-of-sample deviance so smaller values (closer to 
+# zero) are better. Note also that p_waic is the penalty term of WAIC, and so it
+# will be close to the number of dimensions in the posterior of each model  
+
+lm_fit <- readRDS("../results/no_topcodes/fit_lm_hifdip_losat_cov_2002.RDS")
+pw_fit <- readRDS("../results/no_topcodes/fit_nl_hifdip_losat_cov_2002.RDS")
+
+# Comparing WAIC using loo_compare()
+# The best model will be reported at the top of the list with elpd_diff = 0
+ans <- loo_compare(lm_fit, pw_fit, criterion = "waic")
+
+print(ans, simplify = F)
+#         elpd_diff se_diff
+# pw_fit   0.0       0.0  
+# lm_fit -69.6      12.2  
+
+waic_results <- tibble()
+
+for (i in c(2001:2018)) {
+  
+  lm_fit <- readRDS(paste0("../results/no_topcodes/fit_lm_hifdip_losat_cov_", i, ".RDS"))
+  pw_fit <- readRDS(paste0("../results/no_topcodes/fit_nl_hifdip_losat_cov_", i, ".RDS"))
+  
+  ans <- loo_compare(lm_fit, pw_fit, criterion = "waic")
+  
+  as_tibble(ans, rownames = "model") %>%
+    mutate(year = i) %>%
+    mutate(across(elpd_diff:year, as.numeric)) -> tb
+  
+  # Add R-squared
+  lm_r2 <- bayes_R2(lm_fit) %>%
+    as_tibble() %>%
+    rename(R2 = Estimate, R2.Error = Est.Error, R2_Q2.5 = Q2.5, R2_Q97.5 = Q97.5) %>%
+    mutate(year = i, model = "lm_fit")
+  
+  pw_r2 <- bayes_R2(pw_fit) %>%
+    as_tibble() %>%
+    rename(R2 = Estimate, R2.Error = Est.Error, R2_Q2.5 = Q2.5, R2_Q97.5 = Q97.5) %>%
+    mutate(year = i, model = "pw_fit")
+  
+  tb %>%
[…]

### 2021-01-12 src/model_fitting.R

@@ -1,93 +1,0 @@
-#### The Increasing Cost of Happiness ####
-# RW Morris, N Kettlewell, N Glozier
-# 
-# Model specification and fitting
-# Dec 4th 2020
-# 
-#### libraries ####
-library(dplyr)
-library(Rcpp)
-library(brms)
-options(mc.cores = parallel::detectCores())
-ncores = parallel::detectCores()
-
-#### Import ####
-happywealth <- readRDS("../data/happywealth.rds")
-
-#### Fit the model ####
-#
-# This model was suggested at: 
-# https://discourse.mc-stan.org/t/piecewise-linear-mixed-models-with-a-random-change-point/5306/5
-# 
-# Model definition
-# b0 is the intercept denoting the expected value of wellbeing (y) at the changepoint
-# b1 is the linear slope before the changepoint
-# b2 is the linear slope after the changepoint
-# alpha is the coefficient of wealth (x) at the changepoint
-
-bfixed <- bf(
-  y ~ b0 + b1 * (dollars - omega) * step(omega - dollars) + 
-    b2 * (dollars - omega) * step(dollars - omega) + b3*grad + b4*age + b5*age_sq + b6*male,
-  b0 + b1 + b2 + b3 + b4 + b5 + b6 + alpha ~ 1,  # fixed intercepts for each year
-  # Sigmoid transform to keep omega within the range of 0 to x
-  nlf(omega ~ inv_logit(alpha) * 100),
-  nl = TRUE
-)
-
-# priors                                      I tried tighter priors:
-bprior <- prior(normal(0, 3), nlpar = "b0") + # normal(0, 1) did not converge Sep 6th
-  prior(normal(0, 3), nlpar = "b1") +         # normal(0, 1) did not converge Sep 6th
-  prior(normal(0, 3), nlpar = "b2") +         # normal(0, 1) did not converge Sep 6th
-  prior(normal(0, 3), nlpar = "b3") +  
-  prior(normal(0, 3), nlpar = "b4") +  
-  prior(normal(0, 3), nlpar = "b5") +  
-  prior(normal(0, 3), nlpar = "b6") +
-  prior(normal(0, 3), nlpar = "alpha")      # normal(0, 3) with init = "0" 12/09/2020
-                                            # normal(0, sqrt(3)) with default init didn't converge
-
-#### Piece-wise fits ####
-for (i in c(2001:2018)) {
-  
-  #### Preprocessing ####
[…]

### 2021-01-12 src/model_selection.r

@@ -1,92 +1,0 @@
-#### Comparing models ####
-# https://bookdown.org/ajkurz/Statistical_Rethinking_recoded/overfitting-
-# regularization-and-information-criteria.html#the-problem-with-parameters
-# 
-# WAIC is a (penalized) estimate of out-of-sample deviance so smaller values (closer to 
-# zero) are better. Note also that p_waic is the penalty term of WAIC, and so it
-# will be close to the number of dimensions in the posterior of each model  
-
-lm_fit <- readRDS("../results/fit_lm_hifdip_losat_cov_2018.RDS")
-pw_fit <- readRDS("../results/fit_cp_hifdip_losat_cov_2018.RDS")
-
-lm_fit <- add_criterion(lm_fit, "waic")
-pw_fit <- add_criterion(pw_fit, "waic")
-
-# Comparing WAIC using loo_compare()
-# The best model will be reported at the top of the list with elpd_diff = 0
-ans <- loo_compare(lm_fit, pw_fit, criterion = "waic")
-
-print(ans, simplify = F)
-#         elpd_diff se_diff
-# pw_fit   0.0       0.0  
-# lm_fit -69.6      12.2  
-
-waic_results <- tibble()
-
-for (i in c(2002:2018)) {
-  
-  lm_fit <- readRDS(paste0("../results/fit_lm_hifdip_losat_cov_", i, ".RDS"))
-  pw_fit <- readRDS(paste0("../results/fit_cp_hifdip_losat_cov_", i, ".RDS"))
-  
-  lm_fit <- add_criterion(lm_fit, "waic")
-  pw_fit <- add_criterion(pw_fit, "waic")
-
-  saveRDS(lm_fit, paste0("../results/fit_lm_hifdip_losat_cov_", i, ".RDS"))
-  saveRDS(pw_fit, paste0("../results/fit_cp_hifdip_losat_cov_", i, ".RDS"))
-  
-  ans <- loo_compare(lm_fit, pw_fit, criterion = "waic")
-  
-  as_tibble(ans, rownames = "model") %>%
-    mutate(year = i) %>%
-    mutate(across(elpd_diff:year, as.numeric)) -> tb
-  
-  # Add R-squared
-  lm_r2 <- bayes_R2(lm_fit) %>%
-    as_tibble() %>%
-    rename(R2 = Estimate, R2.Error = Est.Error, R2_Q2.5 = Q2.5, R2_Q97.5 = Q97.5) %>%
[…]

### 2021-01-12 src/piecewise model simulation.r

@@ -1,0 +1,477 @@
+library(tidyverse)
+library(brms)
+
+tibble(
+  year = rep(1:16, each = 20),
+  dollars = rep(1:10, 32)
+  ) %>% 
+  rowwise() %>%
+  mutate(
+    year = as.character(year),
+    dollars = dollars + rnorm(1),
+    b0 = 8,
+    b1 = 2,
+    b2 = 0.05,
+    omega = 5, #ceiling((year + 2)/2.5),
+    y = -log(1/(dollars + 1.5)) + rnorm(1),
+    age = sample(18:100, 1),
+    male = as.logical(sample(0:1, 1)),
+    grad = as.logical(rbinom(1, 1, 0.25))) %>%
+  ungroup() %>%
+  mutate(
+    y = as.vector(scale(y)),
+    age = as.vector(scale(age)),
+    agesq = age^2
+    ) -> df
+
+ggplot(df, aes(x = dollars, y = y)) +
+  geom_point() +
+  facet_wrap(~year)
+
+hist(df$dollars)
+hist(df$y)
+summary(df$dollars)
+#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
+# -2.094   3.038   5.559   5.478   7.930  12.232 
+
+#### Fit the model ####
+#
+# This model was suggested at: 
+# https://discourse.mc-stan.org/t/piecewise-linear-mixed-models-with-a-random-change-point/5306/5
+# 
+# Model definition
+# b0 is the intercept denoting the expected value of wellbeing (y) at the changepoint
+# b1 is the linear slope before the changepoint
+# b2 is the linear slope after the changepoint
+# alpha is the coefficient of wealth (x) at the changepoint
+
+
+bfixed <- bf(
+  y ~ b0 + b1 * (dollars - omega) * step(omega - dollars) + 
+    b2 * (dollars - omega) * step(dollars - omega) + b3*age + b4*agesq + b5*male + b6*grad,
+  # b0 + b1 + b2 + b3 + b4 + b5 + b6 + alpha ~ 1 + year, # fixed intercept/year (complete pooling)
+  # b0 + b1 + b2 + b3 + b4 + b5 + b6 + alpha ~ 1 + (1|year), # random intercept/year (no pooling)
+  # Sigmoid transform to keep omega within the range of 0 to x
+  nlf(omega ~ inv_logit(alpha) * 15),
+  nl = TRUE
+)
+
+# priors                                      
+bprior <- prior(normal(0, 3), nlpar = "b0") + 
[…]

### 2021-01-12 src/piecewise_model_simulation.r

@@ -1,115 +1,0 @@
-library(tidyverse)
-library(brms)
-
-tibble(
-  year = rep(1:16, each = 20),
-  dollars = rep(1:10, 32)
-  ) %>% 
-  rowwise() %>%
-  mutate(
-    dollars = dollars + rnorm(1),
-    b0 = 8,
-    b1 = 2,
-    b2 = 0.05,
-    omega = 5, #ceiling((year + 2)/2.5),
-    y = b0 + b1*(dollars - omega)*(omega > dollars) + b2*(dollars - omega)*(dollars >= omega) + rnorm(1),
-    age = sample(18:100, 1),
-    male = as.logical(sample(0:1, 1)),
-    grad = as.logical(rbinom(1, 1, 0.25))) %>%
-  ungroup() %>%
-  mutate(
-    y = as.vector(scale(y)),
-    age = as.vector(scale(age)),
-    agesq = age^2
-    ) -> df
-
-ggplot(df, aes(x = dollars, y = y)) +
-  geom_point() +
-  facet_wrap(~year)
-
-hist(df$wealth)
-hist(df$y)
-summary(df$dollars)
-#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
-# -2.094   3.038   5.559   5.478   7.930  12.232 
-
-bfixed <- bf(
-  y ~ b0 + b1 * (dollars - omega) * step(omega - dollars) + 
-    b2 * (dollars - omega) * step(dollars - omega) + b3*age + b4*agesq + b5*male + b6*grad,
-  b0 + b1 + b2 + b3 + b4 + b5 + b6 + alpha ~ 1, #+ (1|male + grad),  # fixed intercept + (random intercept per sex and grad)
-  # Sigmoid transform to keep omega within the range of 0 to x
-  nlf(omega ~ inv_logit(alpha) * 15),
-  nl = TRUE
-)
-
-# priors                                      
-bprior <- prior(normal(0, 3), nlpar = "b0") + 
-  prior(normal(0, 3), nlpar = "b1") +         
-  prior(normal(0, 3), nlpar = "b2") +         
-  prior(normal(0, 3), nlpar = "b3") +
-  prior(normal(0, 3), nlpar = "b4") +
-  prior(normal(0, 3), nlpar = "b5") +
-  prior(normal(0, 3), nlpar = "b6") +
[…]

### 2021-01-12 src/preprocessing.r

@@ -2,7 +2,7 @@
 # RW Morris, N Kettlewell, N Glozier
 # 
 # Preprocessing 
-# Nov 27th 2020
+# Jan 8th 2021
 # 
 # Libraries
 library(tidyverse)
@@ -26,6 +26,14 @@ for (pathtofile in path_to_hilda) {
 source('~/Dropbox/HILDA/src/gather_hilda.R')
 
 #### DEMOGRAPHICS ####
+# Demographic variables were selected to match the sample design of HILDA, so 
+# age, sex, employment status, marital status, and region (State) were included.
+
+state_key = c(`1` = "NSW", `2` = "VIC", `3` = "QLD", `4` = "SA", `5` = "WA", 
+              `6` = "TAS", `7` = "NT", `8` = "ACT")
+
+region_key = c(`0` = "capital", `1` = "urban", `2` = "regional", `3` = "rural")
+
 gather_hilda(hilda, c(
   "esbrd",   # employment status (broad)
   "hhda10",  # SEIFA 2001 Decile of socio-economic advantage (higher is better)
@@ -36,10 +44,18 @@ gather_hilda(hilda, c(
   # "hhrhid",  # household id (within wave)
   "mrcurr",  # current marital status (married/de facto ≤ 2)
   "edhigh1", # highest education achieved
-  "edfts"    # current fulltime student (yes = 1)
+  "edfts",   # current fulltime student (yes = 1)
+  "hhstate", # household State of residence (NSW, VIC, QLD, SA, WA, TAS, NT, ACT)
+  "hhssos",  # household section of State (major urban, other urban, boundary, rural)
+  "hhwte"    # enumarated persons cross-sectional weight (sex, broad age, 
+             # region, employment, marital status)
   )) %>%
   spread(code, val) %>%
-  mutate_if(is.double, ~ ifelse(. < 0, NA_real_, .)) -> demographic_items
+  mutate_if(is.double, ~ ifelse(. < 0, NA_real_, .)) %>%
+  mutate(hhssos = replace_na(hhssos, 3),
+         hhssos = recode(hhssos, !!!region_key),
+         hhstate = recode(hhstate, !!!state_key)
+         ) -> demographic_items
 
 demographics <- demographic_items %>%
   transmute(
@@ -70,26 +86,22 @@ demographics <- demographic_items %>%
     chronic = if_else(helth == 1, TRUE, FALSE, missing = FALSE),
[…]

### 2021-01-12 src/residual data.r

@@ -1,0 +1,29 @@
+#### create residual plot ####
+
+library(tidyverse)
+
+path_to_fits <- list.files(
+  path = 'results',
+  pattern = '^fit_cp_hifdip_gh9_cov_.*.RDS$',
+  full.names = TRUE
+)
+
+residual_df <- data.frame()
+
+for (path in path_to_fits) {
+  fit <- readRDS(path)
+  resids = resid(fit)[, "Estimate"]
+  fitteds = fitted(fit)[, "Estimate"]
+  df <- data.frame(resids, fitteds)
+  df$year = substring(path, 31, 34)
+  
+  residual_df <- bind_rows(residual_df, df)
+}
+
+write_rds(residual_df, "results/residual_data.RDS")
+
+resids = resid(pw_fit)[, "Estimate"]
+fitteds = fitted(pw_fit)[, "Estimate"]
+cor(fitteds, resids)
+hist(resids)
+ggplot() + geom_point(data = NULL, aes(y = resids, x = fitteds))

### 2021-01-12 src/piecewise model simulation.r
update to extensions for ignoring

@@ -1,3 +1,6 @@
+#### Author: RMorris
+#### Date: Jan 12th 2021
+
 library(tidyverse)
 library(brms)
 

### 2021-01-12 src/figure 1.r
rm figure 1.r

@@ -1,182 +1,0 @@
-library(tidyverse)
-library(haven)
-
-#### Load data ####
-path_to_hilda <- list.files(
-  path = '~/Dropbox/HILDA/data',
-  pattern = '^Combined.*.dta$',
-  full.names = TRUE
-)
-
-hilda <- list()
-for (pathtofile in path_to_hilda) {
-  df <- read_dta(pathtofile)
-  hilda <- append(hilda, list(df))
-  cat('.')
-}
-
-# Helper function
-source('~/Dropbox/HILDA/src/gather_hilda.R')
-
-
-#### Satisfaction ####
-satisfaction <- gather_hilda(hilda, c(
-  "losat",   # Life satisfaction
-  "hhwtrps") # sample weights for life satisfaction (RP)
-  ) %>%
-  spread(code, val) %>%
-  mutate_if(is.double, ~ ifelse(. < 0, NA_real_, .)) %>%
-  transmute(xwaveid, wave, losat, losat_wt = losat * hhwtrps)
-
-#### Income ####
-# CPI values from https://www.pc.gov.au/research/completed/rising-inequality
-CPI = data.frame(
-  wave = letters[1:18],
-  deflator = c(1.52547554, # 2000/01
-               1.48315720, # 2001/02
-               1.43988458, # 2002/03
-               1.40607389, # 2003/04
-               1.37297463, # 2004/05
-               1.33027251, # 2005/06
-               1.29200230, # 2006/07
-               1.24993042, # 2007/08
-               1.21214575, # 2008/09
-               1.18464785, # 2009/10
-               1.14888718, # 2010/11
-               1.12303076, # 2011/12
-               1.09804401, # 2012/13
-               1.06903118, # 2013/14
-               1.05101802, # 2014/15
-               1.03670360, # 2015/16
-               1.01929187, # 2016/17
-               1)) %>% mutate(inflator = rev(deflator))
-
-thresholds <- read_csv("../../docs/Frequencies 180c/HILDA-thresholds-by-wave-180.csv") %>%
-  filter(variable %in% c("hifdip")) %>%
-  gather("wave", "threshold", -variable, -label) %>%
-  separate(wave, into = c("temp", "wave"), sep = "e") %>%
-  mutate(
-    variable = paste0(variable, "_thld"),
-    wave = letters[as.numeric(wave)]) %>%
[…]

### 2021-01-12 src/figure 5.r
rm figure 5.r

@@ -1,62 +1,0 @@
-
-predicted_fits <- list()
-for (pathtofile in path_to_fits) {
-  fit <- read_rds(pathtofile) 
-  
-  ans <- model.frame(fit) %>%
-    summarise_all(median)
-  
-  newdat = data.frame(
-    dollars = c(5, ans[, "dollars"], 8),
-    age = rep(ans[, "age"], 3),
-    age_sq = rep(ans[, "age_sq"], 3),
-    male = rep(ans[, "male"], 3), 
-    grad = rep(ans[, "grad"], 3)
-  )
-  
-  df <- fitted(fit, newdata = newdat) %>% 
-    as_tibble() %>%
-    mutate(
-      Dollars = c("$50K", "\u03bc", "$80K"),
-      year = as.numeric(substring(pathtofile, 34, 37)))
-  
-  predicted_fits <- bind_rows(predicted_fits, df)
-}
-
-#### plot those fits
-predicted_fits %>%
-  mutate(Dollars = fct_relevel(Dollars, "$80K", "\u03bc")) %>%
-  ggplot(aes(x = year, y = Estimate, color = Dollars)) +
-    geom_point(size = 4, alpha = 0.25) +
-    geom_smooth(method = "loess", se = F, span = 3) +
-    facet_wrap(~Dollars) +
-    labs(title = "Predicted happiness at varying income levels") +
-    ggthemes::scale_colour_wsj() +
-    ggthemes::theme_wsj() +
-    theme(legend.position = "none",
-          plot.title = element_text(size = 19),
-          plot.caption = element_text(size = 12),
-          strip.text = element_text(size = 12))
-
-#### plot those fits
-predicted_fits %>%
-  select(Estimate, Dollars, year) %>%
-  filter(Dollars %in% c("$80K", "$50K")) %>%
-  spread(Dollars, Estimate) %>%
-  # mutate(`∆` = `$80K` - `$50K`) %>%
-  gather(income, happiness, -year) %>%
-  mutate(income = fct_relevel(income, "$50K", "$80K")) %>%
-  ggplot(aes(x = year, y = happiness, color = income)) +
-  geom_point(size = 4, alpha = 0.25) +
-  geom_smooth(method = "loess", se = F, span = 3) +
-  facet_wrap(~income) +
-  labs(title = "Estimated happiness at varying income\nin Australia from 2002-2018",
-       caption = "Source: HILDA, Melbourne University") -> p
-
-p +
-  ggthemes::scale_colour_wsj() +
[…]

### 2021-01-12 src/plotting cost of happiness 2012-2018.r
rm plotting cost of happiness 2012-2018.r

@@ -1,113 +1,0 @@
-library(bayesplot)
-
-
-fit <- read_rds("../results/fit_cp_hifdip_losat_cov_2006.RDS")
-
-plot(fit)
-
-path_to_fits <- list.files(
-  path = "../results",#'~/Dropbox/HILDA/data',
-  pattern = '^fit_cp_hifdip_losat_cov_.*.RDS$',
-  full.names = TRUE
-)
-
-alphas <- list()
-for (pathtofile in path_to_fits) {
-  df <- read_rds(pathtofile) %>%
-    as.data.frame() %>%
-    select(b_alpha_Intercept) 
-  
-  # colnames(df) <- paste0("b_alpha_", substring(pathtofile, 34, 37)) # gh9
-  colnames(df) <- paste0("b_alpha_", substring(pathtofile, 36, 39)) # losat
-  
-  alphas <- bind_cols(alphas, df)
-}
-
-alphas %>%
-  mutate_all(.funs = ~brms::inv_logit_scaled(., 0, 100)*10) -> posteriors
-
-# OPTIONAL: Trim posteriors for 2003 outlier
-posteriors %>%
-  mutate(b_alpha_2003 = (b_alpha_2002 + b_alpha_2004)/2) -> posteriors
-
-# OPTIONAL: Trim posteriors for 2006 outlier (did not converge)
-posteriors %>%
-  mutate(b_alpha_2006 = (b_alpha_2005 + b_alpha_2007)/2) -> posteriors
-
-color_scheme_set("blue")
-
-mcmc_areas_ridges(
-  posteriors,
-  pars = rev(colnames(posteriors)),
-  prob = 0.90, # 80% intervals
-  prob_outer = .99 # 99%
-  ) +
-  theme_test() +
-  geom_vline(aes(xintercept = median(posteriors$b_alpha_2002)), 
-             color = "gray", 
-             linetype = 3) +
-  labs(subtitle = "Cost of happiness by year and household wealth",
-       x = "Household wealth ($000s)")
-
-happywealth_covs %>%
-  select(year, xwaveid, adj_hifdip) %>%
-  mutate(
-    year = paste0("b_alpha_", year),
-    adj_hifdip = adj_hifdip/1000) %>%
-  left_join(
-    gather(
-      summarise_all(posteriors, list(median)), year, change_point
-    )) %>%
-  mutate(
-    year = extract_numeric(year)
-    ) %>%
-  group_by(year) %>%
-  summarise(
[…]

### 2021-01-12 src/preprocessing.R
Update and rename preprocessing.r to preprocessing.R

@@ -2,7 +2,7 @@
 # RW Morris, N Kettlewell, N Glozier
 # 
 # Preprocessing 
-# Jan 8th 2021
+# Jan 12th 2021
 # 
 # Libraries
 library(tidyverse)

### 2021-02-08 src/draft0.Rmd
Major commit: Added counterfactual, added years 2002:2018 to main, updated format for PNAS

--- 2020-12-05 src/draft0.Rmd
@@
 library(flextable)
-library(gtsummary)
 library(bayesplot)
+library(patchwork)
 
@@
 
+source("ft_themes.r")
+
 #### To do ####
@@
 # https://www.pc.gov.au/research/completed/rising-inequality/3_HILDA_variables.R
-# more people could a achieve a level of income at which happiness no longer increased
 ```
@@
 1. Central Clinical School, Faculty of Medicine and Health, University of Sydney, NSW, Australia
-2. ARC Centre of Excellence for Children and Families over the Life Course, University of Queensland, QLD, Australia
+2. ARC Centre of Excellence for Children and Families over the Life Course, University of Sydney, NSW, Australia
 3. School of Economics, University of Technology, NSW, Australia
@@
 | Draft      | `r format(Sys.time(), '%d %B, %Y')` |
-| Word count | 3767                                |
-| Tables     | 1                                   |
-| Figures    | 5                                   |
+| Words      | 4789                                |
+| Tables     | 0                                   |
+| Figures    | 4                                   |
 
@@
 
-A fundamental question for governments and people is how much happiness does a dollar buy?
+A fundamental question for society is how much happiness does a dollar buy?
 The accepted view among economists and psychologists is that money and happiness increase together up to a point, after which there is little further gain from increasing wealth.
-While the location of this change point has been determined, and the cost of happiness reportedly ranges between USD\$60 to $95K, there has been no investigation as to whether the cost of happiness has increased or decreased over time.
-We tested the relationship between income and both happiness and life satisfaction using household economic data from Australia between 2002-2018.
-We discovered the cost of happiness has increased over those 17 years faster than inflation (i.e., cost of living).
-This result shows that inequalities in wealth may be driving inequities in happiness and wellbeing, with implications for health and recent government policy-goals to monitor and improve wellbeing.
-(150 words)
+While the location of this change point reportedly ranges between USD\$60K to $95K, there has been no investigation as to whether this has increased or decreased over time.
+We tested the temporal relationship between income and affective wellbeing (happiness), and income and cognitive wellbeing (life satisfaction), using household economic data from Australia between 2002-2018.
+We discovered the change point between happiness and income has increased over those 17 years faster than inflation (i.e., cost of living).
+This suggests that inequalities in income may be driving increasing inequities in happiness between the rich and the poor, with implications for health and recent government policy-goals to monitor and improve wellbeing.
+(250 words max)
 
@@
 
+## Significance Statement
+Substantial research has shown that higher incomes have decreasing gains on wellbeing and happiness, across countries, cultures and ethnicities, with an inflection or change point at or near USD75,000.
+At incomes below this point happiness is dependent on financial security while at income levels above, happiness is relatively independent; revealing an inequality in the distribution of happiness between rich and poor.
+However we do not know whether or how this point has itself changed over time.
+Has it decreased and reduced inequality, or increased the difference between the rich and poor?
+We examine 17 years of data from a single developed western nation (Australia) and determine that the happiness of more people are dependent on financial security than before.
+
+<br><br>
+
 ## Background
-A fundamental question for psychology and economics is just how much wellbeing does a dollar buy?
-Increasing income is commonly associated with increasing wellbeing, however a point at which income no longer increases subjective wellbeing has also been widely observed [@clark2008relative; @dolan2008we; @easterlin1974growth].
-Given that a central goal of nations and governments is to improve wealth under the assumption that wealth always increases wellbeing, challenges to this notion have far reaching consequences [@frijters2020happy].
+A fundamental question for governments and people is just how much wellbeing does a dollar buy?
+Increasing income is commonly associated with increasing happiness and wellbeing, however a point at which subjective wellbeing no longer increases with income has also been widely observed [@clark2008relative; @dolan2008we; @easterlin1974growth].
+Given that a central goal of nations and governments is to improve income under the assumption that higher income always increases wellbeing, challenges to this notion have far reaching consequences [@frijters2020happy; @laws2014world].
 
-Subjective wellbeing is not a unitary entity [@diener2017findings]; studies typically distinguish between life satisfaction, the cognitive appraisal of one’s own accomplishments, and happiness, one’s prevailing affective state or emotional mood.
+Subjective wellbeing is not a unitary entity [@diener2017findings]; studies typically distinguish between life satisfaction, the cognitive appraisal of one’s own accomplishments, and affective wellbeing, one's prevailing affective state, emotional mood, or everyday experience of happiness.
 Money can have different effects on each.
-For instance, we have recently reported that positive life events, such as a major financial windfall, have a greater impact on an individual’s satisfaction than their happiness [@kettlewell2020differential].
-While the distinct effect of money on satisfaction and happiness was observed within individuals, the distinct effects of money have also been observed across individuals.
-For instance, Kahneman & Deaton [-@kahneman2010high] showed that self-reported levels of happiness increased with household income up to a point (\$75,000).
-But after that, increasing amounts of money had no further effect on happiness.
-They also reported that life satisfaction continued to increase with income beyond \$75,000.
-Indeed, the difference between the two questions: “How satisfied are you with your life?” and “How happy are you these days?” has been identified as a crucial mediating factor in a meta-analysis of 111 studies on money and wellbeing [@howell2008relation; also @veenhoven2006rising].
-Results such as these have provided a more nuanced view among psychologists and (some!) economists about the relationship between money and wellbeing; namely that money is more strongly related to satisfaction than to happiness.
+For instance, we have recently reported that large increases in wealth, such as a major financial windfall, have a greater impact on an individual’s life satisfaction than their happiness [@kettlewell2020differential].
+The distinct effect of money on satisfaction and happiness was observed within an individual over time, however a distinct effect of income have also been observed between people with different income levels.
+For instance, Kahneman & Deaton [-@kahneman2010high] showed that self-reported levels of happiness increased with household income up to a point (USD75,000), but after that, increasing income had little further effect on happiness.
+Conversely life satisfaction continued to increase with income beyond USD75,000.
+Indeed, the difference between the two questions: “How satisfied are you with your life?” and “How happy are you these days?” has been identified as a crucial mediating factor in a meta-analysis of 111 studies on income and wellbeing [@howell2008relation; also @veenhoven2006rising].
+Results such as these have provided a more nuanced view among psychologists and some economists regarding the relationship between income and wellbeing; namely that income is more strongly related to satisfaction than to happiness.
 
-The distinction between satisfaction and happiness are increasingly relevant to governments and policy-makers due to the growing recognition that increasing wealth does not necessarily lead to improvements in wellbeing [@clark2018four; @frijters2020happy].
-If money no longer improves wellbeing, then the maximization of wealth may no longer be a legitimate goal of government.
-Fundamentally, the existence of a change point between income and happiness reveals an unacknowledged inequality in the distribution of wellbeing in the economy.
-Of concern is the point at which income produces no further increases in happiness – that is, the change point, or cost of happiness.
-The cost of happiness represents the point at which material wealth stops driving inequalities in the distribution of happiness in the economy, where lower cost-points represent more equitable distributions of happiness.
-For instance, Kahneman estimated the cost of happiness among US survey respondents in 2008 to be approximately USD75,000 per year, substantially more than the US median income (USD52,000) in that year.
-Research since then has revealed variations in the cost of happiness due to world region, gender, and education [@jebb2018happiness], consistent with financial inequality driving inequities in wellbeing and happiness.
-However, to date there has been no investigation of whether the cost of happiness has changed over time.
-In particular has the cost, and therefore the distribution, of happiness become more or less equitable in the last few decades?
-
-We used household economic panel data from Australia (HILDA) to provide the first investigation of whether changes in income and wellbeing have changed the cost of happiness over the last 17 years (2002-2018).
-HILDA provides a representative sample of households in Australia with detailed measurements of income and subjective wellbeing in the same sample, which makes it an excellent data source to investigate the present question.
-We distinguished between satisfaction and happiness as different components of subjective wellbeing, and evaluated how each varies with household income.
-After adjusting for age, gender, and education level, we confirmed that happiness and satisfaction have distinct relationships with increasing income, but the cost of happiness has increased in real dollar terms since 2002.
+Fundamentally, the existence of a change point in the relationship between income and happiness reveals an unacknowledged source of inequality in the distribution of wellbeing (i.e., happiness) in the economy.
+For instance, the change point of USD75,000 reported by Kahneman in 2008 was substantially more than the US median income of USD52,000 in the same year, indicating that the happiness of the poorest majority of the US population was tied to marginal changes in income while the happiness of a richer minority was not.
+Thus the change point represents the dollar value up to which income drives inequities in the distribution of happiness, such that a lower value represents a more equitable distribution of happiness in the economy.
+Inequities in the distribution of wellbeing are increasingly relevant to governments and policy-makers due to the growing recognition that increasing income does not necessarily lead to equal changes in wellbeing [@clark2018four; @frijters2020happy].
+Even prior to COVID-19, the World Gallup Poll has observed that happiness has been decreasing over the past decade in western Europe, north America, Australia and New Zealand, despite increases in income in the same countries [@sachs2019world].
+However to date there has been no investigation of whether the relationship between income and happiness has changed over time which may contribute to these social trends.
+In particular has the change point between income and happiness, and therefore the distribution of happiness between rich and poor, become more or less equitable in the last few decades?
 
@@
 
+We used household economic panel data from Australia (HILDA) to provide the first investigation of whether changes in income and wellbeing have shifted the change point over the last 17 years (2002-2018).
+HILDA provides a representative sample of households in Australia with detailed measurements of income and subjective wellbeing in the same sample, which makes it an excellent data source to investigate the present question.
+We distinguished between satisfaction and happiness as different components of subjective wellbeing, and evaluated how each varies with household income.
+Full-time students were removed, as well as individuals with an annual household disposable income designated as _topcoded_ (which occurs to ensure confidentiality of very high income individuals who might otherwise be identifiable).
+
 <br>
@@
 #### Income
-Our indicator of income and economic security was household income.
-Household income better represents economic security than personal income, since members of the same household share expenses as well as risks; i.e., they can provide a direct and immediate support network when financial shocks occur.
-The other major studies also used household after-tax income as the indicator of wealth and economic security [e..g, @kahneman2010high; @jebb2018happiness], and so we follow the same standard here as well.
-The ‘real household annual disposable income’ was calculated from the combined income of all household members after receipt of government pensions and benefits and deduction of income taxes in the financial year ended 30th June of the year of the wave (e.g., 2002 in wave 2).
+To allow comparison with other major studies of income and wellbeing we used household after-tax income as the indicator of income and economic security [e..g, @kahneman2010high; @jebb2018happiness].
+Household income better represents economic security than personal income, since members of the same household share expenses as well as risks; i.e., they can provide a direct and immediate support network when financial shocks occur, which a priori might affect wellbeing.
+The ‘real household annual disposable income’ was calculated from the self-reported combined income of all household members after receipt of government pensions and benefits and deduction of income taxes in the financial year ended 30th June of the year of the wave (e.g., 2002 in wave 2).
 This was then adjusted for inflation - the rise in the general price level of the economy - using the Australian Bureau of Statistics (ABS) Consumer Price Index, so that income in all waves is expressed in FY 2017/18 prices, to give real income.
@@
 #### Subjective Wellbeing
-There are a variety of variables related to subjective well-being collected annually in HILDA, but the two we used here matched the variables we used in our previous paper [@kettlewell2020differential], namely, life satisfaction as a measure of cognitive wellbeing, and the SF-36 as a measure of affective wellbeing or happiness.
+There are a variety of variables related to subjective well-being collected annually in HILDA, but the two we used here matched the variables we used in our previous paper [@kettlewell2020differential], namely, life satisfaction as a measure of cognitive wellbeing, and item 9 from the SF-36 (9a-9i) as a measure of affective wellbeing or happiness.
 
@@
 
-Happiness was determined by item 9 in the SF-36 [(gh9a to gh9i)](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=gh9a){target="_blank"}.
+Happiness was determined by 9 questions in the SF-36 [(9a to 9i)](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=gh9a){target="_blank"}.
 The SF-36 is a widely used self-completion measure of various aspects of physical, emotional and mental health [@ware2000sf].
-Item 9 consists of nine questions relating to mental health and vitality, where five questions deal with positive and negative aspects of mental health (e.g., "Felt so down in the dumps nothing could cheer me up", "Been happy"), and four questions deal with positive and negative aspects of vitality (e.g., "feel full of life", "felt worn out").
-Each question referred to the past four weeks and agreement was indicated on a six-point scale.
-We reverse scored the relevant responses and calculated the sum of the nine questions so that higher scores represented better wellbeing.
+A subset of 9 questions assess mental health and vitality, with five questions measuring positive and negative aspects of mental health (e.g., "Felt so down in the dumps nothing could cheer me up", "Been happy"), and four questions on positive and negative aspects of vitality (e.g., "feel full of life", "felt worn out").
+The response scale timeframe is the past four weeks and agreement was indicated on a six-point Likert scale.
+We reverse scored negatively phrased questions and calculated the sum of the nine questions so that higher scores represented better wellbeing.
 To aid interpretability, we rescaled the final sum to a score between 1-100, where 100 represents the maximum happiness achievable.
@@
 We modelled the relationship between income and each wellbeing variable (happiness and satisfaction) using a simple **linear model** and a **piecewise model** (broken-stick).
-The piecewise model was chosen as the simplest extension of a linear model which can identify a change point (inflection) between wellbeing and income.
+The piecewise model was chosen as the simplest extension of a linear model which can identify a change point (inflection) in the relationship between wellbeing and income.
 The location of the change point was a free parameter which revealed where wellbeing no longer increased at a uniform rate with income.
@@
 
+For modelling, both measures of wellbeing were rescaled with a mean of zero and a SD of 1 (z-scores) for each year.
+
 _Model Estimation_  
-We adopted a Bayesian approach for estimating the linear and piecewise model in the software Stan [@burkner2017; @rstan2019]. In each case,
+We adopted a Bayesian approach for estimating the linear and piecewise model in the software Stan [@burkner2017; @rstan2019]. In each case the linear model was estimated as:
 
-Let $y_i \sim N(\mu_i, \sigma^2_y)$
-
-The linear model was estimated as:
+$$
+y_i \sim N(\mu_i, \sigma^2_y)
+$$
 
@@
 
-The piecewise model was a simple extension of this to include a free parameter to represent the changepoint in income ($\omega$) as well as the slope before the change point ($\beta_1$) and after the change point ($\beta_2$):
+The piecewise model was a simple extension of this to include a free parameter to represent the changepoint in income ($\omega$) as well as the slope before the change point ($\beta_1$) and the slope after the change point ($\beta_2$):
 
@@
 
-The above models estimated population-level effects separately for each year (*t* = 2002, 2006, 2010, 2014 and 2018).
+The above models estimated population-level effects separately for each year (*t* = 2002...2018).
 Because we were interested in the location of the change point between income and wellbeing that existed across individuals within each year, we ignored the panel design of HILDA because the dependency between observations of the same person across years was orthogonal to our effects of interest.
@@
 _Covariates_  
-Age (and age^2^), gender, and education were included as covariates.
+Age (and age^2^), gender, and university graduate education were included as covariates.
+Cross-sectional population weights for Australia provided by the University of Melbourne for each year were also included as a covariate to adjust for differences in the sample representativeness according to sex by broad age, marital status, region, and labour force status.
 Full-time students were removed, as well as individuals with an annual household disposable income that was indicated as _topcoded_ by the University of Melbourne (topcoding occurs to ensure privacy of high wealth individuals).
@@
 
-The broad demographic characteristics of the sample over the time period are presented below in Table 1 (at four year intervals).
+The broad demographic characteristics of the sample are presented in Supplementary Materials Table S1.
+Average life satisfaction levels were very steady between 2002-2018, while average happiness score decreased slightly over the 17 years.
+The proportions of each sex and relationship status were stable over time, as were the average household size and SEIFA index.
+However age, education, and chronic health conditions tended to slightly increase over time.
+For instance, average age increased by 1.8 years over the 17 years of the survey, which is obviously less than would occur in a cohort study [@watson2012hilda].
+Changes in the workforce varied with economic circumstances.
 
-```{r table_1, ft.align="left"}
-library(gtsummary)
+#### Happiness and satisfaction have distinct relationships with income
 
-happywealth %>%
-  filter(year %in% c(2002, 2006, 2010, 2014, 2018)) %>%
-  filter(!student, !top_hifdip, (!is.na(losat)|!is.na(gh9))) %>%
-  mutate(
-    Sex = if_else(male, "Male", "Female"),
-    Education = fct_collapse(
-      edu,
-      `High school or less` = c("None", "In school", "Highschool"),
-      Graduate = c("Grad")
-      ),
-    Workforce = case_when(
-      workforce & !unemployed ~ "Employed",
-      unemployed ~ "Unemployed",
-      TRUE ~ "Not in labour force"
-      ),
-    `Relationship status` = case_when(
-      single ~ "Single",
-      coupled ~ "Married",
-      TRUE ~ "Divorced/Widow"
-      ),
-    Workforce = fct_relevel(Workforce, "Employed", "Unemployed"),
-    `Relationship status` = fct_relevel(`Relationship status`, "Single", "Married")
-    ) %>%
-  select(
-    year, 
-    `Life satisfaction` = losat, `Happiness` = gh9,
-    Sex, `Age (years)` = age, Education, Workforce, `Relationship status`, 
-    `Chronic illness` = chronic, SEIFA, `Household size` = hhsize,
-  ) %>% 
-  tbl_summary(
-    by = year,
-    missing = "no",
-    statistic = 
-      list(all_continuous() ~ "{mean} (±{sd})",
-           all_categorical() ~ "{n} ({p}%)"),
-    digits = all_continuous() ~ 1,
-    missing_text = "(Missing)" 
-    ) %>%
-  modify_header(stat_by = "{level}\nN = {style_number(n)}") %>%
-  as_flex_table() %>%
-  set_caption(caption = "Table 1. Demographic characteristics") %>%
-  align(part = "body") %>%
-  valign(valign = "top", part = "header") %>%
-  fontsize(size = 9.5, part = "all") %>%
-  width(j = 1, width = 1.5) %>%
-  width(j = 2:6, width = 1)
-```
+The relationship between household income and happiness (red) and satisfaction (blue) every four years is shown in Figure 1.
+For each wellbeing variable we show the results of a linear fit (rows 1 and 3) and a piecewise fit (rows 2 and 4).
+For visualization purposes only, due to the large numbers of individual data in each year, we display the mean levels of income and wellbeing for each (equal-sized) income decile rather than every individual data point, whereas the line-of-best-fit and 95% credible intervals (shaded) in each regression model are derived from _all individuals_.
 
@@
 
-Average life satisfaction levels were very steady between 2002-2018, while average happiness score decreased slightly over the 17 years.
-The proportions of each sex and relationship status were stable over time, as were the average household size and SEIFA index.
-However age and chronic health conditions tended to increase over time, as expected in an aging sample.
-Education levels (i.e., graduate percentage) also tended to increase over time, while changes in the workforce varied both up and down.
-
-<br>
-
-```{r funtimes}
+##### Figure 1. Household income and happiness (red) and satisfaction (blue)  
+```{r losat_gh9_hifdip}
 source("../src/happyfun.r")
-```
-
-Below we show the relationship between household income and wellbeing (Figure 1).
-For visualization purposes only, due to the large numbers of individual data in each year, we display the mean levels of income and wellbeing for each (equal-sized) income decile rather than every individual data point.
-Note that the line-of-best-fit and 95% credible intervals (shaded) from each regression model _of all individuals_ is shown in overlay.
-
-<br>
-
-##### Figure 1. Household income and satisfaction (blue) and happiness (red)
-```{r losat_gh9_hifdip}
-library(patchwork)
 
 p1 <- happywealth %>%
-  rename(Satisfaction = losat, `Household income` = hh_disp_inc_eq) %>% 
+  rename(`Satisfaction 0-10` = losat, 
+         `Household income` = hh_disp_inc_eq) %>% 
   filter(!student, !top_hifdip) %>%
-  plot_linear(Satisfaction, `Household income`) +
+  plot_linear(`Satisfaction 0-10`, `Household income`) +
   scale_fill_manual(values = "#b3cde0") +
@@
 p2 <- happywealth %>%
-  rename(Satisfaction = losat, `Household income` = hh_disp_inc_eq) %>% 
+  rename(`Satisfaction 0-10` = losat, 
+         `Household income` = hh_disp_inc_eq) %>% 
   filter(!student, !top_hifdip) %>%
-  plot_inflection_piecewise(Satisfaction, `Household income`) +
+  plot_inflection_piecewise(`Satisfaction 0-10`, `Household income`) +
   scale_fill_manual(values = "#b3cde0") +
@@
 p3 <- happywealth %>%
-  rename(Happiness = gh9, `Household income` = hh_disp_inc_eq) %>% 
+  rename(`Happiness 1-100` = gh9, 
+         `Household income` = hh_disp_inc_eq) %>% 
   filter(!student, !top_hifdip) %>%
-  plot_linear(Happiness, `Household income`) +
+  plot_linear(`Happiness 1-100`, `Household income`) +
   scale_fill_manual(values = "#C79999") +
@@
 p4 <- happywealth %>%
-  rename(Happiness = gh9, `Household income` = hh_disp_inc_eq) %>% 
+  rename(`Happiness 1-100` = gh9, 
+         `Household income` = hh_disp_inc_eq) %>% 
   filter(!student, !top_hifdip) %>%
-  plot_inflection_custom(Happiness, `Household income`, 
+  plot_inflection_custom(`Happiness 1-100`, `Household income`, 
                          knots = c(43.7, 49.2, 47.4, 53.2, 72.3)) +
@@
 ```
-Figure 1 legend: Posterior predictive distributions (95% credible intervals) showing the model fit from regressions of wellbeing on household income, overlaid on summary data points (income deciles).
-Wellbeing was measured as life satisfaction (blue) or happiness (red).
-The total number of individuals contributing to each year are shown (n).
+Figure 1 legend: The relationship between income and wellbeing across equal-sized income deciles, overlaid by regression lines from linear and piecewise models (±95%CI).
+Wellbeing was measured as happiness (red) or life satisfaction (blue).
+The total number of individuals contributing to each regression in each year are noted (n).
 ```
@@
 
-Figure 1 shows the relationship between household income and satisfaction appeared relatively linear, as the results of the piecewise regression indicated the inflection of the change point could be increasing (e.g., 2002, 2006) as well as decreasing (e.g., 2010, 2014, 2018).
-By contrast, a decreasing change point was evident in each year of happiness on income.
-The results of a formal comparison between linear and piecewise models is provided in Model Selection below.
-However note that the change point between income and happiness appeared to shift rightwards over time (bottom row).
-(We can also see the linear relationship between household income and each wellbeing variable became slightly weaker over time, i.e., less steep)
+The nonlinear relationship between happiness and income (Figure 1, 2nd row) was consistently and negatively inflected (happiness increased less with income after the change point).
+By contrast the nonlinear relationship between satisfaction and income shown in the 4th row was as likely to be negatively inflected as positively inflected: positively inflected in 2002 and 2006; negatively inflected 2014 and 2018; and no apparent inflection in 2010.
+
+The posterior evidence from model selection revealed the nonlinear (piecewise) fit of _happiness_, but not _satisfaction_, was credibly superior to a linear fit in each year (Supplementary Materials, Figure S2).
+Overall the posterior evidence suggests that happiness and satisfaction have distinct relationships with household income; satisfaction tends to increase linearly with income, while a change point exists in the relationship between happiness and household income.
+
+<br><br>
+
+#### The evolution of income on happiness 2002-2018
+
+Figure 2 below presents the posterior distribution of each parameter from the piecewise model regressing happiness on income: the change point (&omega;), the intercept (&beta;~0~), the pre-change slope (&beta;~1~), and the post-change slope (&beta;~2~).
+Over 12,000 samples were drawn from each piecewise model to determine the posterior distribution of each parameter.
+Horizontal bars represent the 95% credible interval and so intervals which fall completely to the right of the vertical grey dotted line are credibly higher than our base year 2002.
 
@@
 
-#### Model Selection
+##### Figure 2. Posterior parameters of income on happiness  
+```{r get_posteriors}
+path_to_gh9_fits <- list.files(
+  path = "../results/incl_topcodes/with_pweights/",
+  pattern = "^fit_nl_hifdip_gh9_4cov_pwt_.*.RDS$",
+  full.names = TRUE
+)
 
-We compared the posterior evidence for a linear relationship between wellbeing and income with the posterior evidence of a nonlinear (piecewise) relationship (i.e., WAIC~linear~ — WAIC~piecewise~).
-Thus a WAIC difference greater than zero indicated evidence for a linear relationship.
-A WAIC difference less than zero indicated evidence for a nonlinear (piecewise) relationship.
+posteriors_b0 <- list()
+posteriors_b1 <- list()
+posteriors_b2 <- list()
+posteriors_alpha <- list()
 
-<br>
+for (pathtofile in path_to_gh9_fits) {
 
-##### Figure 2. Linear model evidence (WAIC~linear~ — WAIC~piecewise~)
+  df <- read_rds(pathtofile) %>%
+    as.data.frame()
 
-```{r plot_waic_gh9}
-waic_result <- read_rds("../results/model_comparisons_waic_R2_gh9.RDS")
+  b0 <- select(df, b_b0_Intercept)
+  b1 <- select(df, b_b1_Intercept)
+  b2 <- select(df, b_b2_Intercept)
+  alpha <- select(df, b_alpha_Intercept)
 
-waic_result %>%
-  select(model, year, elpd_diff, se_diff) %>%
-  filter(elpd_diff != 0) %>%
-  filter(year %in% c(2002, 2006, 2010, 2014, 2018)) %>%
-  rowwise() %>%
-  mutate(samples = list(rnorm(n = 1e5, 
-                              mean = elpd_diff * 2, # for pw advantage
-                              sd   = se_diff *  2)),
-         year = paste0("AD", year)
-         ) %>%
-  select(year, samples) %>%
-  group_by(year) %>%
-  unnest(samples) %>% 
-  mutate(sample_id = c(1:1e5)) %>%
-  pivot_wider(id_col = "sample_id", names_from = "year", values_from = "samples") %>%
-  select(-sample_id) -> d2
+  column_name <- paste0("param_",
+                        substring(pathtofile, nchar(pathtofile)-7, nchar(pathtofile)-4))
 
-color_scheme_set("red")
+  colnames(b0) <- column_name
+  colnames(b1) <- column_name
+  colnames(b2) <- column_name
+  colnames(alpha) <- column_name
 
-p5 <- mcmc_intervals(x = d2,
-           pars = rev(colnames(d2)),
-           prob = 0.00,
-           prob_outer = 0.95
-           ) +
-  scale_x_continuous(
-    limits = c(-200, 150)
-    ) +
-  geom_vline(aes(xintercept = 0), color = "grey90") +
-  theme_test() +
-  labs(
-    subtitle = "Happiness")
+  posteriors_b0 <- bind_cols(posteriors_b0, b0)
+  posteriors_b1 <- bind_cols(posteriors_b1, b1)
+  posteriors_b2 <- bind_cols(posteriors_b2, b2)
+  posteriors_alpha <- bind_cols(posteriors_alpha, alpha)
+
+}
+
+smooth_by_year <- function(posteriors) {
+
+  posteriors %>%
+    transmute(
+      param_2002 = (param_2001 + param_2002)/2,
+      param_2003 = (param_2002 + param_2004)/2,
+      param_2004 = (param_2003 + param_2004)/2,
+      param_2005 = (param_2004 + param_2005)/2,
+      param_2006 = (param_2005 + param_2006)/2,
+      param_2007 = (param_2006 + param_2007)/2,
+      param_2008 = (param_2007 + param_2008)/2,
+      param_2009 = (param_2008 + param_2009)/2,
+      param_2010 = (param_2009 + param_2010)/2,
+      param_2011 = (param_2010 + param_2011)/2,
+      param_2012 = (param_2011 + param_2012)/2,
+      param_2013 = (param_2012 + param_2013)/2,
+      param_2014 = (param_2013 + param_2014)/2,
+      param_2015 = (param_2014 + param_2015)/2,
+      param_2016 = (param_2016 + param_2016)/2,
+      param_2017 = (param_2017 + param_2017)/2,
+      param_2018 = (param_2018 + param_2018)/2
+    ) -> smoothed
+
+  return(smoothed)
+
+}
 ```
 
-```{r plot_waic_losat}
-waic_result <- read_rds("../results/model_comparisons_waic_R2_losat.RDS")
+```{r figure_2, fig.width = 6, fig.height = 7, dpi=300}
+posteriors_omega <- posteriors_alpha %>%
+  mutate_all(.funs = ~brms::inv_logit_scaled(., 0, 100)*10)
 
-waic_result %>%
-  select(model, year, elpd_diff, se_diff) %>%
-  filter(elpd_diff != 0) %>%
-  filter(year %in% c(2002, 2006, 2010, 2014, 2018)) %>%
-  rowwise() %>%
-  mutate(samples = list(rnorm(n = 1e5, 
-                              mean = elpd_diff * -2, # for linear advantage
-                              sd   = se_diff *  2)),
-         year = paste0("AD", year)
-         ) %>%
-  select(year, samples) %>%
-  group_by(year) %>%
-  unnest(samples) %>% 
-  mutate(sample_id = c(1:1e5)) %>%
-  pivot_wider(id_col = "sample_id", names_from = "year", values_from = "samples") %>%
-  select(-sample_id) -> d1
+smooth_by_year(posteriors_omega) %>% 
+  mutate(param = "omega ($000s)") %>% # (\u03c9)
+  bind_rows(., (mutate(smooth_by_year(posteriors_b0), 
+                       param = paste0("intercept (\u03b2", "0)")))) %>%
+  bind_rows(., (mutate(smooth_by_year(posteriors_b1), 
+                       param = paste0("preslope (\u03b2", "1)")))) %>%
+  bind_rows(., (mutate(smooth_by_year(posteriors_b2), 
+                       param = paste0("postslope (\u03b2", "2)")))) %>%
+  gather(key = "year", val = "sample", -param) %>% 
+  mutate(
+    # year = str_remove(year, "param_"),
+    year = extract_numeric(year),
+    param = fct_relevel(param, "omega ($000s)")) %>% 
+  filter(year > 2001) %>%
+  group_by(param, year) %>%
+  summarise(
+    low = quantile(sample, probs = 0.05, na.rm=T),
+    expected = quantile(sample, probs = 0.5, na.rm=T),
+    upp = quantile(sample, probs = 0.95, na.rm=T)
+  ) -> d0
 
-# "fix" d1
-colnames(d1) <- rev(colnames(d1))
+d1 <- filter(d0, year == 2002)
 
-color_scheme_set("blue")
-
-p6 <- mcmc_intervals(x = d1,
-  pars = colnames(d1),
-  prob = 0.0,
-  prob_outer = .95
-  ) +
-  scale_x_continuous(
-    limits = c(-200, 150)
-    ) +
-  geom_vline(aes(xintercept = 0), color = "grey90") +
-  theme_test() +
-  labs(
-    subtitle = "Satisfaction",
-    x = "WAIC difference (95% interval)")
-```
-
-```{r figure_2, fig.width=6.5, fig.height=4, dpi=300}
-p5 / p6
+ggplot(d0, aes(y = year, x = expected, color = param)) +
+  geom_pointrange(aes(xmin = low, xmax = upp)) +
+  geom_vline(aes(xintercept = expected), d1, color = "gray", 
+             linetype = 3) +
+  scale_y_continuous(breaks = c(2002, 2006, 2010, 2014, 2018)) +
+  facet_wrap(~param, scales = "free_x") +
+  ft_theme(base_size = 12) +
+  scale_color_manual(values = c("red", "grey65", "grey65", "grey65")) +
+  theme(legend.position = "none",
+        axis.title = element_blank())
 ```
@@
 ```
-Figure 2 legend: Differences in posterior evidence for a linear fit over a piecewise fit (WAIC) for satisfaction (blue) and happiness (red).
-The filled circle indicates the mean of the distribution and the horizontal bars represents the 95% credible interval.
-A difference greater than zero is support for the linear model and a difference below zero is support for the piecewise model.
-```
-
-<br><br>
-
-Model selection revealed the posterior evidence for the linear fit of satisfaction on income was credibly superior to the nonlinear (piecewise) fit - Figure 2 shows the 90% credible interval of the fit for each year of satisfaction was above zero with no overlap.
-By contrast, a nonlinear fit of happiness on income was generally superior, and the nonlinear fit was credibly superior at 90% for each of three years (2006, 2014, 2018).
-Thus, the posterior evidence indicates happiness and satisfaction have distinct relationships with household income; satisfaction tends to increase linearly with income, while a change point exists in the relationship between happiness and income.
-
-<br>
-
-#### Parameter Estimation
-
-Figure 3 below presents the complete posterior distribution of &omega; over 4000 samples drawn from the piecewise model, representing the location of the change point between happiness and household income in four separate years.
-Shaded areas to the right of the vertical grey dotted line are credibly (95%) larger than 2002.
-The figure shows the change point between happiness and household income (the _cost of happiness_) credibly increased between 2002 and 2018.
-
-<br>
-
-##### Figure 3. Posterior distributions of &omega; (95% distribution)  
-```{r figure_3, fig.width = 6, fig.height = 4.5, dpi=300}
-path_to_fits <- list.files(
-  path = "../results",
-  pattern = '^fit_cp_hifdip_gh9_cov_.*.RDS$',
-  full.names = TRUE
-)
-
-posterior_alphas <- list()
-for (pathtofile in path_to_fits) {
-  df <- read_rds(pathtofile) %>%
-    as.data.frame() %>%
-    select(b_alpha_Intercept)
-
-  colnames(df) <- paste0("param_", substring(pathtofile, 34, 37))
-
-  posterior_alphas <- bind_cols(posterior_alphas, df)
-}
-
-posterior_alphas %>%
-  mutate_all(.funs = ~brms::inv_logit_scaled(., 0, 100)*10) -> posterior_omegas
-
-posteriors_trim <- posterior_omegas %>%
-  mutate(
-    param_2010 = (param_2010 + param_2011)/2,
-    param_2003 = (param_2002 + param_2004)/2)
-
-# select posteriors for plotting
-posteriors_select <- posterior_omegas %>%
-  select(param_2002, param_2006, param_2010, param_2014, param_2018)
-
-color_scheme_set("red")
-
-mcmc_areas_ridges(
-  posteriors_select,
-  pars = rev(colnames(posteriors_select)),
-  prob = 0.95, 
-  prob_outer = .99 
-  ) +
-  theme_test() +
-  geom_vline(aes(xintercept = median(posteriors_select$param_2002)), 
-             color = "gray", 
-             linetype = 3) +
-  labs(x = "Household income ($000s)")
-```
-
-```
-Figure 3 legend: Posterior distribution of the change point parameter representing the location in household income ($) for each year. Shaded area represents the 90% credible region.
+Figure 2 legend: Posterior distributions of the change point parameter omega representing the location in real household income (2018 dollars), as well as the intercept, preslope and postslope parameters in arbitrary units.
+Horizontal bar represents the 95% credible region and the solid point indicates the expected value (median) of each distribution.
+Vertical dotted line indicates the 2002 parameter value (median) as a base year comparison.
 ```
@@
 
-#### Income and happiness
+The posterior estimates of the change point between happiness and household income indicates that the location (i.e., the real income value) of the change point shows a systematic increasing trend since 2011.
+Changes to the other parameters of the function between income and happiness also occured between 2002 and 2018 (i.e., the pre-slope, post-slope and intercept), however these did not show any sustained trend over the period nor result in a reliable change from the baseline year 2002.
 
-##### Figure 4. Estimated happiness (SD units) at $50K/yr and $80K/yr from 2002-2018 
-```{r figure_4, fig.width=7, fig.height=5, dpi=300}
-predicted_fits <- list()
-for (pathtofile in path_to_fits) {
-  fit <- read_rds(pathtofile)
+It is also worth noting that both the pre-slope and post-slope parameters (&beta;~1~ and &beta;~2~) were credibly larger than zero in most years, indicating there was a reliable dependency between happiness and income at income levels below and above the change point.
+However comparing the pre-slope with the post-slope values makes clear that the relationship between happiness and income among the majority of people in lower income households was an order of magnitude steeper than those in high income households.
 
-  ans <- model.frame(fit) %>%
-    summarise_all(median)
+The change in parameter values between 2002 to 2018 indicates the relationship between happiness and income evolved over time.
+We determined the impact of this evolution on the distribution of happiness over the range of household income in 2018 in a counterfactual analysis.
+The counterfactual analysis is a hypothetical demonstration of how happiness would change if people in 2018 were subject to the function that existed in 2002.
+That is, would happiness increase or decrease if the 2002 function was in place in 2018?
+It controls for changes in the sample which occur over time that are not related to happiness but could nevertheless contribute to changes in the distribution of happiness.
+For example, the increase in the range of (real) income levels in the economy between 2002 and 2018 could produce a greater divergence in happiness due to the increasing gap between the rich and poor - even with a stable relationship between income and happiness.
+Such changes in the sample characteristics may mask or confound the impact of the change point on the distribution of happiness without careful control.
+Because we were interested in the implications of the evolution of the function rather than changes in our sample characteristics _per se_, we estimated happiness levels for each _n_ = 14,280 person in 2018 using the 2002 function.
+These 2002 model-estimates were compared to (subtracted from) the 2018 model-estimates generated from the same sample (*n* = 14,280), to obtain the change (delta) in happiness for each person under the counterfactual.
+This delta must be entirely due to the evolution of the function between 2002 and 2018, since the analysis has held everything else constant (i.e., exactly the same people with the same characteristics were used to estimate happiness from two different functions).
+Figure 3 below presents the 14,280 deltas from such a comparison, along with a smoothed mean (blue overlay) to summarize how happiness evolved with the function across the entire income distribution.
 
-  newdat = data.frame(
-    dollars = c(5, ans[, "dollars"], 8),
-    age = rep(ans[, "age"], 3),
-    age_sq = rep(ans[, "age_sq"], 3),
-    male = rep(ans[, "male"], 3), 
-    grad = rep(ans[, "grad"], 3)
+##### Figure 3. Counterfactual difference in happiness between 2002 and 2018
+```{r counterfactual_estimates, eval=F}
+# Data from 2018
+data2018 <- happywealth %>%
+  filter(!student, year == 2018) %>%
+  transmute(
+    y = as.vector(scale(gh9)), # losat or gh9
+    dollars = hifdip_adj/10000, 
+    pweights = as.vector(scale(weights)),
+    male = as.logical(male),
+    age = as.vector(scale(age)),
+    age_sq = age^2,
+    grad = edu %in% c("Grad", "Postgrad")
+  ) %>%
+  na.omit()
+
+# Model fits from 2002 and 2018
+fit2002 <- read_rds(path_to_fits[2])
+fit2018 <- read_rds(path_to_fits[18])
+
+# Model estimates from fits and data 
+fits_2018 <- fitted(fit2018, newdata = dat2018)
+fits_2002 <- fitted(fit2002, newdata = dat2018)
+
+write_rds(lst(fits_2002, fits_2018), path = "../results/counterfactual_estimates.rds")
+```
+
+```{r figure_3, fig.width=7, fig.height=7, dpi=300}
+estimates2002 <- read_rds("../results/counterfactual_estimates.rds")$fits_2002
+estimates2018 <- read_rds("../results/counterfactual_estimates.rds")$fits_2018
+
+# Data from 2018
+data2018 <- happywealth %>%
+  filter(!student, year == 2018) %>%
+  transmute(
+    y = as.vector(scale(gh9)), # losat or gh9
+    dollars = hifdip_adj/10000, 
+    pweights = as.vector(scale(weights)),
+    male = as.logical(male),
+    age = as.vector(scale(age)),
+    age_sq = age^2,
+    grad = edu %in% c("Grad", "Postgrad")
+  ) %>%
+  na.omit()
+
+# Difference between estimates (∆ = 2018 - 2002)
+data2018$delta = estimates2018[, 'Estimate'] - estimates2002[, 'Estimate']
+
+# Plot deltas
+ggplot(data = data2018, aes(x = dollars*10, y = delta)) +
+  geom_point(size = 0.1, alpha = 0.1) + 
+  geom_smooth() +
+  geom_vline(xintercept = 72, color = "grey50", size = 0.4) +
+  geom_hline(yintercept = 0, color = "grey50", size = 0.4) +
+  annotate("text", x = 96, y = -.25, label = "$72K (2018 change point)") +
+  xlim(0, 150) +
+  ylim(-.3, .3) +
+  labs(subtitle = "Happiness delta(s) from 2018 predictions less 2002 predictions (SD units)",
+       x = "Household income ($000s)", y = "",
+       caption = "Source: Predicted happiness scores from 2018 HILDA sample (n = 14,280)") +
+  theme_test() +
+  theme(
+    plot.caption = ggplot2::element_text(
+      hjust = 0,
+      size = ggplot2::rel(0.8),
+      margin = margin(t = 6)
+    )
+  )
+```
+
+```
+Figure 3 legend: The difference (∆) in model-estimated happiness between 2018 and 2002 for the same n = 14,280 individuals from the 2018 survey.
+Values below zero on the y-axis indicate lower happiness estimates in 2018 relative to 2002.
+Points to the left of the vertial ($72K) are individuals below the 2018 change point.
+The smoothed overlay (blue) indicates how the average delta changes across the income distribution.
+```
+
+<br>
+
+Figure 3 shows that, on average, happiness estimates improved for people with household incomes above \$72K when compared to 2002, as indicated by the average delta (smooth blue line) falling above zero on the right side of the plot.
+Conversely, people with household incomes below \$72K, on average, suffered a decrease in estimated happiness compared to 2002.
+And the greatest decrease in happiness occurred for people whose income levels were transliminal: that is, their income was above the changepoint in 2002 but the shift in the change point resulted in their income falling below the change point in 2018.
+Of course the obtained deltas are due to changes in all the parameters of the function, including the slope before and after the change point.
+However because this comparison was performed on the same individuals (i.e., from the 2018 survey), it held characteristics such as age, income, etc, constant that would otherwise be expected to change over time and possibly contribute to any difference in happiness distribution.
+In this way these results isolate the amount of change entirely due to the _evolution_ of the function between 2002 and 2018, and demonstrates how this has contributed to the unequal distribution of happiness between the rich and the poor over time.
+
+<br><br>
+
+#### The increasing cost of happiness in Australia
+
+A less hypothetical implication of the shifting change point in the relationship between income and happiness is that over time fewer people have an income that affords them happiness, which is (relatively) independent of their financial security.
+Assuming median income is also not increasing at a faster rate, any increase in the change point will reduce the number of people who fall above it over time.
+Figure 5 presents median household income levels weighted for the Australian population (by age, sex, marital status, labour force participation and geographical region).
+It also shows the change point between income and happiness increased faster than rises in median household income between 2002 and 2018.
+The third panel shows that as a result, a smaller proportion of the Australian population in 2018 had a household income above the changepoint than in 2002.
+
+##### Figure 4. Rise in median income, change point, and population in Australia 2002-2018 
+```{r figure_4, fig.width=7, fig.height=6, dpi=300}
+trimmed_omega <- posteriors_omega %>%
+      mutate(
+        # param_2002 = (param_2001 + param_2002)/2,
+        param_2003 = (param_2002 + param_2004)/2,
+        # param_2004 = (param_2003 + param_2004)/2,
+        # param_2005 = (param_2004 + param_2005)/2,
+        # param_2006 = (param_2005 + param_2006)/2,
+        # param_2007 = (param_2006 + param_2007)/2,
+        # param_2008 = (param_2007 + param_2008)/2,
+        param_2009 = (param_2008 + param_2009)/2,
+        param_2010 = (param_2009 + param_2011)/2,
+        # param_2011 = (param_2010 + param_2011)/2,
+        # param_2012 = (param_2011 + param_2012)/2,
+        # param_2013 = (param_2012 + param_2013)/2,
+        # param_2014 = (param_2013 + param_2014)/2,
+        # param_2015 = (param_2014 + param_2015)/2,
+        # param_2016 = (param_2015 + param_2016)/2,
+        # param_2017 = (param_2016 + param_2017)/2,
+        # param_2018 = (param_2017 + param_2018)/2
         )
 
-  df <- fitted(fit, newdata = newdat) %>% 
-    as_tibble() %>%
+# Tabulate against median income
+happywealth %>%
+  filter(!student) %>%
+  filter(!top_hifdip) %>%
+  select(year, xwaveid, income = hh_disp_inc_eq, weights) %>%
+  filter(year %in% 2002:2018) %>%
   mutate(
-      Dollars = c("$50K", "\u03bc", "$80K"),
-      year = as.numeric(substring(pathtofile, 34, 37)))
+    year = paste0("param_", year),
+    income = income/1000) %>%
+  left_join(
+    gather(
+      summarise_all(trimmed_omega, list(median)), 
+      year, 
+      change_point
+    )) %>%
+  mutate(
+    year = extract_numeric(year),
+    over_change = income > change_point,
+    weighted_over = over_change * weights
+  ) %>%
+  group_by(year) %>%
+  summarise(
+    `Median household income\n ($000s)` = round(median(income), 2),
+    # `Median household income\n ($000s)` = round(matrixStats::weightedMedian(income, weights)),
 
-  predicted_fits <- bind_rows(predicted_fits, df)
-}
+    `Change point\n ($000s)` = round(mean(change_point), 2),
 
-predicted_fits %>%
-  select(Estimate, Dollars, year) %>%
-  filter(Dollars %in% c("$80K", "$50K")) %>%
-  spread(Dollars, Estimate) %>%
-  mutate(`$80K - $50K (∆)` = `$80K` - `$50K`) %>%
-  gather(income, happiness, -year) %>%
-  mutate(income = fct_relevel(income, "$50K", "$80K")) %>%
-  ggplot(aes(x = year, y = happiness, color = income)) +
-  geom_point(size = 4, alpha = 0.25) +
-  geom_smooth(method = "loess", se = F, span = 3) +
-  facet_wrap(~income) +
-  labs(caption = "Source:     HILDA, University of Melbourne
-                 Predicted happiness scores (SD unit) from piecewise regression on real household income", 
-       x = "", y = "") -> p
+    `Population above change\n(%)` = round(
+      sum(weighted_over) / sum(weights) *100)
 
-source("ft_themes.r")
+    ) %>%
+  gather("key", "value", -year) %>%
+  mutate(
+    key = paste0(key, "\n"),
+    key = fct_relevel(key, "Median household income\n ($000s)\n")
+    ) -> df.plot
+
+df.plot %>%
+  select(key) %>% distinct() %>%
+  mutate(string = c("post-GFC\n(2009 onwards)", "", "")) -> gfc.note
+
+df.plot %>%
+  filter(year %in% c(min(year), 2018)) %>%
+  mutate(
+    val = as.character(
+      if_else(str_detect(key, "Population"),
+              paste0(round(value), "%"),
+              paste0("$", round(value), "K"))),
+    vpos = if_else(val %in% c("$49K", "$72K", "42%"), -0.75, 1.65),
+    hpos = if_else(year == min(year), 0.3, 0.8)
+    ) -> df.label
+
+ggplot(df.plot, aes(x = year, y = value, color = key)) +
+  geom_line(aes(group = key), size = 2, alpha = 0.5) +
+  labs(caption = "Source: Real household income (modified OECD scale) weighted for the Australian population",
+       x = "  ", y = "") +
+  geom_text(data = df.label,
+            aes(label = val, vjust = vpos, hjust = hpos)) +
+  expand_limits(y = c(22, 73)) +
+  facet_wrap(~key) -> p
+
+p.econ <- p + 
+  labs(title = "Income & happiness in Australia between 2002-2018",
+       subtitle = " ") +
+  ggthemes::scale_color_economist() +
+  ggthemes::theme_economist(base_family="Verdana", base_size = 8) +
+  theme(legend.position = "none",
+        strip.text = element_text(size = 10)
+        )
 
@@
 ```
-Figure 4 legend: The difference (∆) in happiness between individuals with a household income of $50K per year and a household income of $80K per year has increased since 2009.
-```
-
-<br>
-
-Figure 5 presents house income levels weighted for the Australian population (by age, sex, marital status, labour force participation and region).
-It shows the cost of happiness increased faster than household income between 2002 and 2018.
-As a result, a smaller percentage of the Australian population in 2018 achieved a level of financial security on which their happiness was no longer dependent on their income, relative to 2002.
-
-<br>
-
-##### Figure 5.  
-```{r figure_5, fig.width=7, fig.height=6, dpi=300}
-# Tabulate against median income
-happywealth %>%
-  filter(!student) %>%
-  filter(!top_hifdip) %>%
-  select(year, xwaveid, income = hh_disp_inc_eq) %>%
-  filter(year %in% 2002:2018) %>%
-  mutate(
-    year = paste0("param_", year),
-    income = income/1000) %>%
-  left_join(
-    gather(
-      summarise_all(posteriors_trim, 
-                    list(median)), 
-      year, 
-      change_point
-    )) %>%
-  mutate(
-    year = extract_numeric(year)
-    ) %>%
-  group_by(year) %>%
-  summarise(
-    `Median household income\n ($000s)` = round(median(income), 
-                                                2),
-    `Cost of happiness\n ($000s)` = round(mean(change_point), 
-                                          2),
-    `Individuals affording happiness\n (Population %)` = round(mean(income > mean(change_point))
-                                                           *100,
-                                                           1)
-    ) %>%
-  gather("key", "value", -year) %>%
-  mutate(
-    key = paste0(key, "\n"),
-    key = fct_relevel(key, "Median household income\n ($000s)\n")) -> df.plot
-
-df.plot %>%
-  select(key) %>% distinct() %>%
-  mutate(string = c("post-GFC\n(2009 onwards)", "", "")) -> gfc.note
-
-df.plot %>%
-  filter(year %in% c(min(year), 2018)) %>%
-  mutate(
-    val = as.character(
-      if_else(str_detect(key, "afford"),
-              paste0(round(value), "%"),
-              paste0("$", round(value), "K"))),
-    vpos = if_else(val %in% c("$49K", "$72K", "41%"), -0.75, 1.65),
-    hpos = if_else(year == min(year), 0.3, 0.8)
-    ) -> df.label
-
-ggplot(df.plot, aes(x = year, y = value, color = key)) +
-  # geom_point() +
-  geom_line(aes(group = key), size = 2, alpha = 0.5) +
-  # geom_smooth(aes(group = key), se = F, span = .5) +
-  labs(title = "Income & happiness in Australia between 2002-2018",
-       caption = "Source: HILDA sample, University of Melbourne
-                  Real household income (modified OECD scale) weighted for the Australian population",
-       x = "  ", y = "") +
-  # geom_vline(aes(xintercept = 2009), color = "red", linetype = 3) +
-  # geom_text(data = gfc.note, aes(x = 2009.5, y = 57, label = string),
-  #           colour = "red", size = 3, hjust = "left") +
-  geom_text(data = df.label,
-            aes(label = val, vjust = vpos, hjust = hpos)) +
-  expand_limits(y = c(22, 73)) +
-  facet_wrap(~key) -> p
-
-p.econ <- p + 
-  labs(subtitle = "  ") +
-  ggthemes::scale_color_economist() +
-  ggthemes::theme_economist(base_family="Verdana", base_size = 8) +
-  theme(legend.position = "none",
-        strip.text = element_text(size = 10)
-        )
-p.econ
-```
-
-```
-Figure 5 legend: Real household income has stagnated in Australia since 2009 (post GFC) while the cost of happiness has increased.
-Consequently fewer Australians have a household income exceeding the cost of happiness in 2018.
+Figure 4 legend: Real household income has stagnated in Australia since 2009 (post GFC) while the change point between happiness and income has increased.
+Consequently fewer Australians had an income that affords them happiness, which is independent of their financial security in 2018.
 ```
@@
 
-We found the relationship between subjective wellbeing and household income was positive, but happiness and satisfaction had different (positive) relationships: Satisfaction increased linearly with income, while happiness increased rapidly up to a point after which further increments in income produced less change – this confirms the distinct effects of money in previous research [e.g., @howell2008relation; @kahneman2010high], and contributes rare evidence from the same sample.
-Furthermore, we report here for the first time that the change point between household income and happiness increased over time between 2002 and 2018, faster than inflation or the median household income.
+We confirmed previous findings [e.g., @howell2008relation; @kahneman2010high], that the relationship between both types of subjective wellbeing with household income was positive but quite different over a 17 year period: Satisfaction increased linearly with income, while happiness increased rapidly up to a point after which higher levels of income were associated with less improvement.
+For the first time we have shown that this change point in the relationship between household income and happiness increased faster than inflation or the median household income over time between 2002 and 2018.
 
-We refer to the change point after which increases in income no longer produce similar increases in happiness as the cost of happiness.
+We refer to the change point after which increases in income no longer produce similar increases in happiness as the _cost of happiness_.
 After this point, happiness is no longer as dependent on household income, and the economic security it represents.
-Presumably after this point further increases in happiness depend on other life factors (e.g., social connections).
-Life satisfaction on the other hand always increased with household income.
-The difference likely reflects the importance of a numerical dollar value (e.g., bank balance, house value) when cognitively appraising one’s life achievements, versus the relevance of that number to our everyday experience of joy and prevailing mood.
+Presumably after this point further increases in happiness depend more on other life factors (e.g., leisure time, social connections).
+Life satisfaction on the other hand appeared to show consistent increases with household income and we found no evidence of any change point.
+The difference may reflect the importance of a numerical dollar value (e.g., bank balance, house value) when cognitively appraising one’s life achievements, versus the relevance of that number to our everyday experience of joy and our prevailing mood.
 
-An implication of the increasing cost of happiness over the last sixteen years is that wealth inequality is driving increasing _inequities_ in wellbeing.
-In 2002, the cost point of happiness represented a 9% increase over median income, while in 2018 it represented a 42% increase over median income.
-This also represents a reduction from 44% to 26% in the proportion of people who have achieved a level of financial security beyond which their happiness no longer depends.
-Thus we can see that over time in the last sixteen years the happiness of more people, i.e., their everyday experience of joy and their prevailing mood, has depended on their material wealth.
+An implication of the changing relationship between happiness and income since 2002 is that income inequality may be driving increasing inequities in wellbeing.
+This is demonstrated in Figure 3, where the difference in happiness between 2002 and 2018 increased for incomes above $72K/year and decreased for incomes below that level.
+The inequity was also highlighted in Figure 4 where the change point of happiness represented a 9% increase over median income in 2002, while in 2018 it represented a 42% increase over median income.
+This increase relative to median income also represented a reduction from 42% to 24% in the proportion of people whose income fell above the changepoint.
+Thus we can see that over the last sixteen years the difference in happiness between the rich and the poor has increased; while the proportion of people whose happiness no longer depends on their financial security has decreased.
 
-Australia has low levels of income disparity relative to other OECD countries, and the Gini coefficient has not changed a great deal between 2002 and 2018 in the HILDA dataset [@australian2018rising].
+Australia has low levels of income disparity relative to many other OECD countries, and the Gini coefficient has not changed a great deal between 2002 and 2018 [@australian2018rising].
 A stable Gini coefficient shows income inequality has remained steady over the time period, and our results do not conflict with this conclusion.
-Rather what we are revealing is the effect of income inequality on happiness has increased over the same time period.
-So while income inequality has remained steady, it’s impact on wellbeing inequity has increased.
+Rather what we are revealing is that even a static income distribution may have dynamic effects on happiness over time.
 We think this highlights the issue that while traditional measures of wealth and income inequality may be relatively stable and exhibit little change, their impact on wellbeing and health can still vary.
-As focus shifts from traditional wealth indicators towards wellbeing measures, findings such as this may become more prevalent.
+As such, we believe these results may well have relevance to other developed nations in North America and Europe which have enjoyed stable economic growth overall, but have stagnating incomes, and declining happiness levels [@sachs2019world].
+As focus shifts from traditional financial indicators towards health and wellbeing measures, findings such as this may become more prevalent.
 
-Some recent studies have challenged either the notion that the positive effect of household income plateaus, or that the effect on happiness and satisfaction are distinct.
-In a nationally representative sample of 44,000 adult Americans in the General Social Survey (GSS), happiness continued to increase with household income, implying no change point existed between money and happiness [@twenge2020expanding].
-The GSS asks a single item on happiness: “Taken all together, how would you say things are these days – would you say that you are very happy, pretty happy, or not too happy?” The form of this question is quite similar to the single item “life satisfaction” question in HILDA, as both request the respondent to cognitively evaluate their circumstances.
-By contrast, the nine items we selected to measure happiness covered a range of affect and focused on current feelings: e.g., “How much of the time during the past 4 weeks have you been a happy person?”.
-Critical differences in the operational definition of happiness seem likely to explain the distinct results observed here.
+Some recent studies have challenged either the notion that the positive effect of  income plateaus after some level, or that the effect on happiness and satisfaction are distinct [@killingsworth2021experienced; @twenge2020expanding; @jebb2018happiness].
+Most recently the point has been made that experienced happiness continues to increase on a log scale with income and does not plateau after USD75,000 [@killingsworth2021experienced].
+It is worth noting that log-linear models are typically used to investigate income effects, and we would agree that our data do not support a strict view of a "satiety" point or flat gradient between income and happiness which can persist on a log scale [@jebb2018happiness; kahneman2010high].
+But log-linear models mask the fact that marginal dollars matter less the more one earns, which is what we observed for happiness here.
+Furthermore, log models [and more flexible additive models @jebb2018happiness] may produce a better fit, but will not provide a clear or a distinct change point allowing us to test a change in the marginal effect of income.
+And so one implication of our model choice is that the change in slope may be much smoother than implied by our model.
+We concede this point but would still contend the location of the change in slope will display a comparable shift over time, if it can be detected with certainty.
 
-A second recent report from a survey of 1.7 million people representing 164 countries in the Gallup World Poll reported increases in household income were associated with change points in happiness and satisfaction [@jebb2018happiness], rather than distinct effects as we found.
-It may be that differences in the sparsity of the high income data in HILDA (for which extremely high income households are masked) explain the differences here.
-However including these individuals in our analysis did not change the linear effect between income and satisfaction observed here, and the weighted topcodes should have biased the result towards a change point.
-We also note a linear effect between income and satisfaction is consistent with the majority of earlier literature [@howell2008relation; @stevenson2013subjective].
+Past researchers have also argued there is no distinct relationship between income and the different forms of subjective wellbeing (e.g., happiness, satisfaction), and that income has the same impact on both constructs.
+It is rare for such research to directly test differences in the form of the function, usually instead inferring no difference after observing null results.
+One benefit of our model selection was the quantification of the posterior evidence for a linear over a nonlinear function (e.g., Figure S1, Supplementary Materials).
+That showed the evidence for a linear effect was clearly different between our two wellbeing variables, and the evidence for a linear function was also decreasing over time.
+It may be that with more recent data, or in the near future, there is no distinct function between income and the different wellbeing constructs.
+However it is also worth noting that the only meta-analysis conducted in this area, over ten years ago, also suggested the linear relationship between income and life satisfaction is stronger than that with happiness [@howell2008relation].
+We would also note that a linear function between income and life satisfaction as we report here, is consistent with the majority of earlier research [@stevenson2013subjective].
+In general, the importance of differences between the distinct constructs of subjective wellbeing will need to be adjudicated when deciding how to measure income effects.
 
@@
 According to @frijters2020happy, coming up with a consensus to translate income into wellbeing features high on the wider wellbeing research agenda.
-Establishing the links between wealth, household income, wellbeing and health, and how inequalities in one drives inequities in the other, will be a critical step in that agenda.
+Establishing the links between wealth, household income, wellbeing and health, and how inequalities in one drives inequities in the other, will be a critical step in the success of that agenda.
 
-<br>
+<br><br>
+
+## Acknowledgements
+
+This research was not supported by any funding.
+
+<br><br>
+
+## Author contributions
+
+RWM analysed the data and wrote the paper. NK designed the analysis and wrote the paper. NG proposed the idea, planned the analysis, and wrote the paper.
+
+<br><br>
+
+## Competing interests
+
+The authors declare no competing interests.
+
+<br><br>
+
+## Additional information
+
+This paper uses unit record data from Household, Income and Labour Dynamics in Australia Survey [HILDA](https://melbourneinstitute.unimelb.edu.au/hilda) conducted by the Australian Government Department of Social Services (DSS).
+The findings and views reported in this paper, however, are those of the author[s] and should not be attributed to the Australian Government, DSS, or any of DSS’ contractors or partners.
+All code and scripts used in the analysis are available at https://github.com/datarichard/The-increasing-cost-of-happiness
+
+Supplementary information is available [here](supplementary.html).
+
+<br><br>
 

### 2021-02-08 src/supplementary.Rmd
Major commit: Added counterfactual, added years 2002:2018 to main, updated format for PNAS

@@
 
-#### Model Selection (2002-2018)
+<br>
+
+#### Sample characteristics
+
+The broad demographic characteristics of the sample at each fourth year wave are presented below in Table S1.
+
+```{r table_1, ft.align="left"}
+happywealth <- read_rds("../data/happywealth.rds")
+
+happywealth %>%
+  filter(year %in% c(2002, 2006, 2010, 2014, 2018)) %>%
+  filter(!student, !top_hifdip, (!is.na(losat)|!is.na(gh9))) %>%
+  mutate(
+    Sex = if_else(male, "Male", "Female"),
+    Education = fct_collapse(
+      edu,
+      `High school or less` = c("None", "In school", "Highschool"),
+      Graduate = c("Grad")
+      ),
+    Workforce = case_when(
+      workforce & !unemployed ~ "Employed",
+      unemployed ~ "Unemployed",
+      TRUE ~ "Not in labour force"
+      ),
+    `Relationship status` = case_when(
+      single ~ "Single",
+      coupled ~ "Married",
+      TRUE ~ "Divorced/Widow"
+      ),
+    Workforce = fct_relevel(Workforce, "Employed", "Unemployed"),
+    `Relationship status` = fct_relevel(`Relationship status`, "Single", "Married")
+    ) %>%
+  select(
+    year, 
+    `Life satisfaction` = losat, `Happiness` = gh9,
+    Sex, `Age (years)` = age, Education, Workforce, `Relationship status`, 
+    `Chronic illness` = chronic, SEIFA, `Household size` = hhsize,
+  ) %>% 
+  tbl_summary(
+    by = year,
+    missing = "no",
+    statistic = 
+      list(all_continuous() ~ "{mean} (±{sd})",
+           all_categorical() ~ "{n} ({p}%)"),
+    digits = all_continuous() ~ 1,
+    missing_text = "(Missing)" 
+    ) %>%
+  modify_header(stat_by = "{level}\nN = {style_number(n)}") %>%
+  as_flex_table() %>%
+  set_caption(caption = "Table S1. Demographic characteristics of the sample") %>%
+  align(part = "body") %>%
+  valign(valign = "top", part = "header") %>%
+  fontsize(size = 9.5, part = "all") %>%
+  width(j = 1, width = 1.5) %>%
+  width(j = 2:6, width = 1)
+```
+
+<br>
+
+Average life satisfaction levels were very steady between 2002-2018, while average happiness score decreased slightly over the 17 years.
+The proportions of each sex and relationship status were stable over time, as were the average household size and SEIFA index.
+However age and chronic health conditions tended to increase over time, indicating a slight increase in the age of the panel over time (1.8 years) but not that of a cohort study which would be 16 years.
+Education levels (i.e., graduate percentage) also tended to increase over time, while changes in the workforce varied with economic circumstances.
+
+<br><br>
+
+#### Model Selection
+
+We compared the posterior evidence of a linear- over a nonlinear- relationship between each wellbeing variable and income (i.e., WAIC~linear~ — WAIC~piecewise~).
+Thus a WAIC difference greater than zero indicated evidence for a linear relationship.
+A WAIC difference less than zero indicated evidence for a nonlinear (piecewise) relationship.
+
+<br>
 
 ##### Figure S1. Linear model evidence (WAIC~linear~ — WAIC~piecewise~)
+
 ```{r complete_waic}
-waic_result <- read_rds("../results/model_comparisons_waic_R2_losat.RDS")
+waic_gh9 <- read_rds("../results/model_comparisons_waic_R2_gh9.RDS")
 
-waic_result %>%
-  select(model, year, elpd_diff, se_diff) %>%
-  filter(elpd_diff != 0) %>%
-  rowwise() %>%
-  mutate(samples = list(rnorm(n = 1e5, 
-                              mean = elpd_diff * -2, # for linear advantage
-                              sd   = se_diff *  2)),
-         year = paste0("AD", year)
-         ) %>%
-  select(year, samples) %>%
-  group_by(year) %>%
-  unnest(samples) %>% 
-  mutate(sample_id = c(1:1e5)) %>%
-  pivot_wider(id_col = "sample_id", names_from = "year", values_from = "samples") %>%
-  select(-sample_id) -> d1
-
-# "fix" d1
-colnames(d1) <- rev(colnames(d1))
-
-color_scheme_set("blue")
-
-p2 <- mcmc_intervals(x = d1,
-  pars = colnames(d1),
-  prob = 0.0,
-  prob_outer = .95
-  ) +
-  scale_x_continuous(
-    limits = c(-200, 150)
-    ) +
-  geom_vline(aes(xintercept = 0), color = "grey90") +
-  theme_test() +
-  labs(
-    subtitle = "Satisfaction",
-    x = "WAIC difference (95% interval)")
-
-waic_result <- read_rds("../results/model_comparisons_waic_R2_gh9.RDS")
-
-waic_result %>%
+waic_gh9 %>%
   select(model, year, elpd_diff, se_diff) %>%
@@
 p1 <- mcmc_intervals(x = d2,
-           pars = rev(colnames(d2)),
+           pars = colnames(d2),
            prob = 0.00,
@@
     ) +
+  scale_y_discrete(labels = as.character(2002:2018)) +
   geom_vline(aes(xintercept = 0), color = "grey90") +
@@
     subtitle = "Happiness")
+
+waic_losat <- read_rds("../results/model_comparisons_waic_R2_losat.RDS")
+
+waic_losat %>%
+  select(model, year, elpd_diff, se_diff) %>%
+  filter(elpd_diff != 0, year > 2001) %>%
+  rowwise() %>%
+  mutate(samples = list(rnorm(n = 1e5, 
+                              mean = elpd_diff * -2, # for linear advantage
+                              sd   = se_diff *  2)),
+         year = paste0("AD", year)
+         ) %>%
+  select(year, samples) %>%
+  group_by(year) %>%
+  unnest(samples) %>% 
+  mutate(sample_id = c(1:1e5)) %>%
+  pivot_wider(id_col = "sample_id", names_from = "year", values_from = "samples") %>%
+  select(-sample_id) -> d1
+
+color_scheme_set("blue")
+
+p2 <- mcmc_intervals(x = d1,
+  pars = rev(colnames(d1)),
+  prob = 0.0,
+  prob_outer = .95
+  ) +
+  scale_x_continuous(
+    limits = c(-200, 150)
+    ) +
+  scale_y_discrete(labels = as.character(2002:2018)) +
+  geom_vline(aes(xintercept = 0), color = "grey90") +
+  theme_test() +
+  labs(
+    subtitle = "Satisfaction",
+    x = "WAIC difference (95% CI)")
 ```
 
-```{r Figure_S1, fig.width=6.5, fig.height=7, dpi=300}
+```{r figure_s1, fig.width=6.5, fig.height=7, dpi=300}
 p1 / p2
@@
 ```
-Figure S1 legend: Differences in posterior evidence (WAIC) for a linear fit over a piecewise fit for satisfaction (blue) and happiness (red).
+Figure S1 legend: Differences in posterior evidence for a linear fit over a piecewise fit (WAIC) for happiness (red) and satisfaction (blue).
 The filled circle indicates the mean of the distribution and the horizontal bars represents the 95% credible interval.
@@
 
+<br>
+
+The posterior evidence from model selection revealed the nonlinear (piecewise) fit of _happiness_, but not _satisfaction_, was credibly superior to a linear fit in each year.
+Figure S1 shows the 95% credible interval of the difference in WAIC values for the linear over nonlinear fits for happiness was below zero with no overlap for each year, indicating consistent support for the nonlinear model.
+By contrast, the WAIC differences for the satisfaction models were generally positive, indicating the superiority of the linear fit for this dependent variable.
+Overall the posterior evidence suggests that happiness and satisfaction have distinct relationships with household income; satisfaction tends to increase linearly with income, while a change point exists in the relationship between happiness and household income.
+
 <br><br>
 
-#### Parameter Estimation (2002-2018)
+#### Model predictions (2002-2018)
 
-##### Figure S2. Posterior distributions of &omega; (95% distribution)  
-```{r gh9_omega}
+The posterior estimates of the piecewise model indicated that the relationship between happiness and income changed over time between 2002 and 2018.
+One implication of such a change is the disparity in happiness between income groups has increased.
+This will occur as more people fall below the change point over time and so are subject to the steep region of the function where income and happiness are strongly related.
+To directly examine the implications of our model for such inequities in happiness, we used the model to estimate or _predict_ the happiness of two different income levels: one income level which fell below the change points between 2002-2018 (\$40K/yr); and another level which always remained above the change points over the same period (\$75K/yr).
+Figure S2 below presents the happiness levels (in SD units/year) for each income level as well as the difference in happiness between them (∆).
+
+<br>
+
+##### Figure S2. Happiness (SD units) at $40K/yr and $75K/yr from 2002-2018 
+```{r figure_s2, fig.width=7, fig.height=5, dpi=300, eval=T}
 path_to_gh9_fits <- list.files(
-  path = "../results",
-  pattern = '^fit_cp_hifdip_gh9_cov_.*.RDS$',
+  path = "../results/incl_topcodes/with_pweights/",
+  pattern = "^fit_nl_hifdip_gh9_4cov_pwt_.*.RDS$",
   full.names = TRUE
@@
 
-posterior_alphas <- list()
+predicted_fits <- list()
 for (pathtofile in path_to_gh9_fits) {
-  df <- read_rds(pathtofile) %>%
-    as.data.frame() %>%
-    select(b_alpha_Intercept)
+  fit <- read_rds(pathtofile)
 
-  colnames(df) <- paste0("param_", substring(pathtofile, 34, 37))
+  ans <- model.frame(fit) %>%
+    summarise_all(median)
 
-  posterior_alphas <- bind_cols(posterior_alphas, df)
+  newdat = data.frame(
+    dollars = c(4, ans[, "dollars"], 7.5),
+    age = rep(ans[, "age"], 3),
+    age_sq = rep(ans[, "age_sq"], 3),
+    male = as.logical(rep(ans[, "male"], 3)), 
+    grad = as.logical(rep(ans[, "grad"], 3)),
+    pweights = rep(ans[, "pweights"], 3)
+  )
+
+  df <- fitted(fit, newdata = newdat) %>% 
+    as_tibble() %>%
+    mutate(
+      Dollars = c("$40K", "\u03bc", "$75K"),
+      year = as.numeric(
+        substring(pathtofile, nchar(pathtofile)-7, nchar(pathtofile)-4)
+        )
+      )
+
+  predicted_fits <- bind_rows(predicted_fits, df)
 }
 
-posterior_alphas %>%
-  mutate_all(.funs = ~brms::inv_logit_scaled(., 0, 100)*10) -> posterior_omegas
+predicted_fits %>%
+  select(Estimate, Dollars, year) %>%
+  filter(Dollars %in% c("$75K", "$40K")) %>%
+  spread(Dollars, Estimate) %>%
+  mutate(`$75K - $40K (∆)` = `$75K` - `$40K`) %>%
+  gather(income, happiness, -year) %>%
+  mutate(income = fct_relevel(income, "$40K", "$75K")) %>%
+  ggplot(aes(x = year, y = happiness, color = income)) +
+  geom_point(size = 4, alpha = 0.25) +
+  geom_smooth(method = "loess", se = F, span = 3) +
+  facet_wrap(~income) +
+  labs(caption = "Source:     HILDA, University of Melbourne
+                 Predicted happiness scores (SD unit) from piecewise regression on real household income", 
+       x = "", y = "") -> p
 
-posteriors_trim <- posterior_omegas %>%
-  mutate(
-    # param_2010 = (param_2010 + param_2011)/2,
-    param_2003 = (param_2002 + param_2004)/2)
+source("ft_themes.r")
 
-color_scheme_set("red")
-
-p3 <- mcmc_intervals(
-  posteriors_trim,
-  pars = rev(colnames(posteriors_trim)),
-  prob = 0.0, 
-  prob_outer = .95 
-  ) +
-  geom_vline(aes(xintercept = median(posteriors_trim$param_2002)), 
-             color = "gray", 
-             linetype = 3) +
-  theme_test() +
-  coord_cartesian(xlim = c(20, 200)) +
-  labs(
-    subtitle = "Happiness",
-    x = "")
+p +
+  ft_theme() +
+  ggthemes::scale_colour_wsj() +
+  theme(legend.position = "none",
+        plot.title = element_text(size = 19),
+        strip.text = element_text(size = 12))
 ```
 
-```{r losat_omega}
-path_to_losat_fits <- list.files(
-  path = "../results",
-  pattern = '^fit_cp_hifdip_losat_cov_.*.RDS$',
-  full.names = TRUE
-)
+```{r alt_figure_s2, fig.width=7, fig.height=5, dpi = 300, eval=F}
+happywealth %>%
+  filter(!student) %>%
+  filter(!top_hifdip) %>%
+  select(year, xwaveid, gh9, income = hh_disp_inc_eq, weights) %>%
+  filter(year %in% 2002:2018) %>%
+  mutate(
+    year = paste0("param_", year),
+    income = income/1000) %>%
+  mutate(
+    year = extract_numeric(year),
+    income_group = case_when(
+      income > 35 & income < 45 ~ "$40K",
+      income > 70 & income < 80 ~ "$75K"
+    )) %>%
+  group_by(year) %>%
+  mutate(
+    happiness = as.vector(scale(gh9)),
+    ) %>%
+  filter(!is.na(income_group)) %>%
+  group_by(year, income_group) %>%
+  summarise(
+    Happiness = mean(happiness, na.rm = T)
+    ) %>% 
+  ungroup() %>%
+  spread(income_group, Happiness) %>%
+  mutate(`$75K - $40K (∆)` = `$75K` - `$40K`) %>%
+  gather(key, val, `$40K`:`$75K - $40K (∆)`) %>%
+  mutate(key = fct_relevel(key, c("$40K", "$75K"))) %>%
+  ggplot(aes(x = year, y = val, color = key)) +
+    geom_point(size = 4, alpha = 0.25) +
+    geom_smooth(span = 3, se = F) +
+    facet_wrap(~key) +
+    labs(caption = "Source:     HILDA, University of Melbourne
+                 Observed happiness scores (SD unit)", 
+     x = "", y = "") -> p
 
-posterior_alphas <- list()
-for (pathtofile in path_to_losat_fits) {
-  df <- read_rds(pathtofile) %>%
-    as.data.frame() %>%
-    select(b_alpha_Intercept)
-
-  colnames(df) <- paste0("param_", substring(pathtofile, 36, 39))
-
-  posterior_alphas <- bind_cols(posterior_alphas, df)
-}
-
-posterior_alphas %>%
-  mutate_all(.funs = ~brms::inv_logit_scaled(., 0, 100)*10) -> posterior_omegas
-
-posteriors_trim <- posterior_omegas %>%
-  mutate(
-    param_2006 = (param_2007 + param_2005)/2)
-
-color_scheme_set("blue")
-
-p4 <- mcmc_intervals(
-  posteriors_trim,
-  pars = rev(colnames(posteriors_trim)),
-  prob = 0.0, 
-  prob_outer = .95 
-  ) +
-  geom_vline(aes(xintercept = median(posteriors_trim$param_2002)), 
-             color = "gray", 
-             linetype = 3) +
-  theme_test() +
-  coord_cartesian(xlim = c(20, 200)) +
-  labs(
-    subtitle = "Satisfaction",
-    x = "Household income ($000s)")
+p +
+  ft_theme() +
+  ggthemes::scale_colour_wsj() +
+  theme(legend.position = "none",
+        plot.title = element_text(size = 19),
+        strip.text = element_text(size = 12))
 ```
-
-```{r Figure_S2, fig.width = 6, fig.height = 7, dpi=300}
-p3 / p4
 ```
-
-```
-Figure S2 legend: Posterior distribution of the change point parameter representing the location in household income ($) for each year. Horizontal bar represents the 95% credible interval.
+Figure S2 legend: The difference (∆) in happiness between a household income of $40K per year and a household income of $75K per year has increased between 2002 and 2018.
 ```
@@
 
-#### Change point model parameter estimates
+Figure S2 shows the disparity in happiness between two fixed income levels, \$40K/year (below the change points) and \$75K/year (above the change points).
+In 2002, a household income of \$40K/year achieved a near average level of happiness relative to everyone else that year, however by 2018 happiness had declined to lower than average levels relative to the population.
+By contrast, a household income of \$75K/year enjoyed a higher than average level of happiness for the entire period.
+The difference (∆) in happiness between these two income levels doubled over the period.
+The increasing disparity in happiness between the rich and the poor implies that happiness has became more inequitable over the period.
+
+<br><br>
+
+***
+
+##### Figure S3. Residual plots for piecewise happiness models
+```{r figure_s3, fig.height=9, fig.width=9}
+# https://www.flutterbys.com.au/stats/tut/tut7.2b.html
+# resid = resid(data.brms)[, "Estimate"]
+# fit = fitted(data.brms)[, "Estimate"]
+
+resid_df <- readRDS("../results/residual_data.RDS")
+
+resid_df %>%
+  group_by(year) %>%
+  slice_sample(n = 1000) %>% # to reduce overplotting
+  ggplot(aes(x = fitteds, y = resids)) +
+    geom_point(alpha = 0.1, size = 0.5) +
+    facet_wrap(~year, scales = "free") +
+    theme_test()
+```
+
+```
+Figure S3 legend: The joint distribution between residual and fitted values shows no common patterns across years and appears clustered around zero in each case.
+```
+
+<br>
+
+Examination of the residual plots from the piecewise models of happiness confirms the residuals were evenly distributed around zero and there was little indication of heterscedascity.
+Overall, there was little evidence that the piecewise models of happiness were not appropriate for the data.
+
+<br><br>
+
+#### Complete parameter estimates and fit statistics
 
@@
 
-The Bayes R^2^ = 0.03964581 ±0.001497172 (0.03672371-0.04262236)
+The Bayes R^2^ = `r round(0.0403964581, 3)` ±`r round(0.001497172, 3)` (95%CI `r round(0.03672371,3)` — `r round(0.04262236, 3)`)
 
@@
     bind_rows(sigma_summary) %>%
-    mutate(Year = substring(pathtofile, 34, 37)) %>%
+    mutate(
+      Year = substring(pathtofile, nchar(pathtofile)-7, nchar(pathtofile)-4)
+      ) %>%
     select(Parameter, Year, everything())
@@
   arrange(Parameter, desc(Year)) %>%
+  filter(Year > 2001) %>%
   flextable() %>%
@@
   colformat_num(j = 7, digits= 1) %>%
-  set_caption("Table S1. Population-level piecewise parameter estimates for happiness") %>%
+  set_caption("Table S1. Piecewise parameter estimates for happiness") %>%
   autofit()
@@
   colformat_num(j = 7, digits= 1) %>%
-  set_caption("Table S2. Population-level linear parameter estimates for happiness") %>%
+  set_caption("Table S2. Linear parameter estimates for happiness") %>%
   autofit()
@@
 ```{r losat_pw_summary, ft.align="left"}
+path_to_losat_fits <- list.files(
+  path = "../results",
+  pattern = '^fit_cp_hifdip_losat_cov_.*.RDS$',
+  full.names = TRUE
+)
+
 population_summary <- tibble()
@@
   colformat_num(j = 7, digits= 1) %>%
-  set_caption("Table S3. Population-level piecewise parameter estimates for satisfaction") %>%
+  set_caption("Table S3. Piecewise parameter estimates for satisfaction") %>%
   autofit()
@@
   colformat_num(j = 7, digits= 1) %>%
-  set_caption("Table S4. Population-level linear parameter estimates for happiness") %>%
+  set_caption("Table S4. Linear parameter estimates for happiness") %>%
   autofit()
@@
 
-##### Figure S3. Residual plots for piecewise happiness models
-```{r figure_s3, fig.height=9, fig.width=9}
-# https://www.flutterbys.com.au/stats/tut/tut7.2b.html
-# resid = resid(data.brms)[, "Estimate"]
-# fit = fitted(data.brms)[, "Estimate"]
-
-resid_df <- readRDS("../results/residual_data.RDS")
-
-resid_df %>%
-  group_by(year) %>%
-  slice_sample(n = 1000) %>% # to reduce overplotting
-  ggplot(aes(x = fitteds, y = resids)) +
-    geom_point(alpha = 0.1, size = 0.5) +
-    facet_wrap(~year, scales = "free") +
-    theme_test()
-```
-
-```
-Figure S3 legend: The joint distribution between residual and fitted values shows no common patterns across years and appears clustered around zero in each case.
-```
-
-<br>
-
-Examination of the residual plots from the piecewise models of happiness confirms the residuals were evenly distributed around zero and there was little indication of heterscedascity.
-There was little evidence that the piecewise models of happiness were not appropriate for the data.
-
-<br><br>
-
-#### Non-linear fits
+#### Additional non-linear fits
 
@@
 
-##### Figure S4. Cubic spline of household wealth on satisfaction and happiness
-```{r losat_gh9_hwnwip}
+##### Figure S5. Cubic spline of household wealth on satisfaction and happiness
+```{r cubic_plot, fig.height=9, fig.width=11}
 source("../src/happyfun.r")
@@
 p1 <- happywealth %>%
-  filter(hifdip_adj > 0) %>%
-  rename(Satisfaction = losat, `Household wealth` = hifdip_adj) %>% 
-  plot_inflection_cubic(Satisfaction, `Household wealth`) +
-  scale_x_continuous(trans = "log", 
-                     breaks = c(5, 10, 20, 40, 80, 160),
-                     limits = c(2.5, 170))
+  filter(!student, !top_hifdip) %>%
+  rename(Satisfaction = losat, `Household income` = hh_disp_inc_eq) %>% 
+  plot_inflection_cubic(Satisfaction, `Household income`) +
+  scale_x_continuous(
+    # trans = "log", 
+    # breaks = c(5, 10, 20, 40, 80, 160),
+    limits = c(5, 170))
 
 p2 <- happywealth %>%
-  filter(hifdip_adj > 0) %>%
-  rename(Happiness = gh9, `Household wealth` = hifdip_adj) %>% 
-  plot_inflection_cubic(Happiness, `Household wealth`) +
-  scale_x_continuous(trans = "log", 
-                     breaks = c(5, 10, 20, 40, 80, 160),
-                     limits = c(2.5, 170))
-```
-```{r cubic_plot, fig.height=9, fig.width=11}
+  filter(!student, !top_hifdip) %>%
+  rename(Happiness = gh9, `Household income` = hh_disp_inc_eq) %>% 
+  plot_inflection_cubic(Happiness, `Household income`) +
+  scale_x_continuous(
+    # trans = "log", 
+    # breaks = c(5, 10, 20, 40, 80, 160),
+    limits = c(5, 170))
+
 p1 / p2
@@
 
-##### Figure S5. Prior predictive plot
+##### Figure S6. Prior predictive plot
 (Not included)
@@
 
-##### Figure S6. Marginal effect of wealth on satisfaction by year (95% CI)  
+##### Figure S7. Marginal effect of wealth on satisfaction by year (95% CI)  
 (Not included)

### 2021-02-08 src/wdraft0.Rmd
Update to README

--- 2020-12-05 src/wdraft0.Rmd
@@
 According to @frijters2020happy, coming up with a consensus to translate income into wellbeing features high on the wider wellbeing research agenda.
-Establishing the links between wealth, household income, wellbeing and health, and how inequalities in one drives inequities in the other, will be a critical step in that agenda.
+Establishing the links between wealth, household income, wellbeing and health, and how inequalities in one drives inequities in the other, will be a critical step in the success of that agenda.
 
-<br>
+<br><br>
+
+## Acknowledgements
+
+This research was not supported by any funding.
+
+<br><br>
+
+## Author contributions
+
+RWM analysed the data and wrote the paper. NK designed the analysis and wrote the paper. NG proposed the idea, planned the analysis, and wrote the paper.
+
+<br><br>
+
+## Competing interests
+
+The authors declare no competing interests.
+
+<br><br>
+
+## Additional information
+
+This paper uses unit record data from Household, Income and Labour Dynamics in Australia Survey [HILDA](https://melbourneinstitute.unimelb.edu.au/hilda) conducted by the Australian Government Department of Social Services (DSS).
+The findings and views reported in this paper, however, are those of the author[s] and should not be attributed to the Australian Government, DSS, or any of DSS’ contractors or partners.
+All code and scripts used in the analysis are available at https://github.com/datarichard/The-increasing-cost-of-happiness
+
+[Supplementary information](supplementary.html){:target="_blank"} is available online.
+
+<br><br>
 

### 2021-03-12 src/draft0.Rmd
Update to README
fixed link to supplementary
Major update v0.4

@@
 source("ft_themes.r")
-
-#### To do ####
-# https://www.psychologytoday.com/au/blog/how-happiness/201409/does-money-really-buy-happiness
-# https://www.pc.gov.au/research/completed/rising-inequality/rising-inequality-highlights.pdf
-# https://www.pc.gov.au/research/completed/rising-inequality/3_HILDA_variables.R
 ```
@@
 | Draft      | `r format(Sys.time(), '%d %B, %Y')` |
-| Words      | 4789                                |
+| Words      | 4399                                |
 | Tables     | 0                                   |
@@
 While the location of this change point reportedly ranges between USD\$60K to $95K, there has been no investigation as to whether this has increased or decreased over time.
-We tested the temporal relationship between income and affective wellbeing (happiness), and income and cognitive wellbeing (life satisfaction), using household economic data from Australia between 2002-2018.
-We discovered the change point between happiness and income has increased over those 17 years faster than inflation (i.e., cost of living).
+We tested the temporal relationship between income and affective wellbeing (happiness), and income and cognitive wellbeing (life satisfaction), using household economic data from Australia between 2001-2019.
+We discovered the change point between happiness and income has increased over those 19 years faster than inflation (i.e., cost of living).
 This suggests that inequalities in income may be driving increasing inequities in happiness between the rich and the poor, with implications for health and recent government policy-goals to monitor and improve wellbeing.
-(250 words max)
+(149 words)
 
@@
 Substantial research has shown that higher incomes have decreasing gains on wellbeing and happiness, across countries, cultures and ethnicities, with an inflection or change point at or near USD75,000.
-At incomes below this point happiness is dependent on financial security while at income levels above, happiness is relatively independent; revealing an inequality in the distribution of happiness between rich and poor.
+At incomes below this point happiness is dependent on financial security while at income levels above, happiness is relatively independent of income; revealing an inequality in the distribution of happiness between rich and poor.
 However we do not know whether or how this point has itself changed over time.
 Has it decreased and reduced inequality, or increased the difference between the rich and poor?
-We examine 17 years of data from a single developed western nation (Australia) and determine that the happiness of more people are dependent on financial security than before.
+We examine 19 years of data from a single developed western nation (Australia) and determine that the income level at which happiness becomes less influenced by income has increased relative to rises in income and inflation.
+The happiness of an increasing proportion of the population is more dependent on financial security than before, whereas the association between income and life satisfaction has remained linear this millennium.
 
@@
 A fundamental question for governments and people is just how much wellbeing does a dollar buy?
-Increasing income is commonly associated with increasing happiness and wellbeing, however a point at which subjective wellbeing no longer increases with income has also been widely observed [@clark2008relative; @dolan2008we; @easterlin1974growth].
+Increasing income is commonly associated with increasing happiness and subjective wellbeing, however a point at which subjective wellbeing no longer increases with income has also been widely observed [@clark2008relative; @dolan2008we; @easterlin1974growth].
 Given that a central goal of nations and governments is to improve income under the assumption that higher income always increases wellbeing, challenges to this notion have far reaching consequences [@frijters2020happy; @laws2014world].
@@
 For instance, we have recently reported that large increases in wealth, such as a major financial windfall, have a greater impact on an individual’s life satisfaction than their happiness [@kettlewell2020differential].
-The distinct effect of money on satisfaction and happiness was observed within an individual over time, however a distinct effect of income have also been observed between people with different income levels.
+The distinct effect of money on satisfaction and happiness was observed within individuals over time, however a distinct effect of income has also been observed between people with different income levels.
 For instance, Kahneman & Deaton [-@kahneman2010high] showed that self-reported levels of happiness increased with household income up to a point (USD75,000), but after that, increasing income had little further effect on happiness.
@@
 Indeed, the difference between the two questions: “How satisfied are you with your life?” and “How happy are you these days?” has been identified as a crucial mediating factor in a meta-analysis of 111 studies on income and wellbeing [@howell2008relation; also @veenhoven2006rising].
-Results such as these have provided a more nuanced view among psychologists and some economists regarding the relationship between income and wellbeing; namely that income is more strongly related to satisfaction than to happiness.
+Larger and newer data sets now also suggest that higher incomes are associated with more happiness and life satisfaction both within and across nations [@sacks2012new; @jebb2018happiness; @killingsworth2021experienced].
+Results such as these have provided a more nuanced view of the relationship between income and wellbeing; namely that while both happiness and satisfaction continue to increase with income, the happiness returns are decreasing [@diener2010wealth; @haushofer2014psychology].
 
@@
 For instance, the change point of USD75,000 reported by Kahneman in 2008 was substantially more than the US median income of USD52,000 in the same year, indicating that the happiness of the poorest majority of the US population was tied to marginal changes in income while the happiness of a richer minority was not.
-Thus the change point represents the dollar value up to which income drives inequities in the distribution of happiness, such that a lower value represents a more equitable distribution of happiness in the economy.
+Thus the change point represents the dollar value up to which income drives inequities in the distribution of happiness, such that a lower valued change point represents a more equitable distribution of happiness in the economy.
 Inequities in the distribution of wellbeing are increasingly relevant to governments and policy-makers due to the growing recognition that increasing income does not necessarily lead to equal changes in wellbeing [@clark2018four; @frijters2020happy].
@@
 
-We used household economic panel data from Australia (HILDA) to provide the first investigation of whether changes in income and wellbeing have shifted the change point over the last 17 years (2002-2018).
-HILDA provides a representative sample of households in Australia with detailed measurements of income and subjective wellbeing in the same sample, which makes it an excellent data source to investigate the present question.
-We distinguished between satisfaction and happiness as different components of subjective wellbeing, and evaluated how each varies with household income.
-Full-time students were removed, as well as individuals with an annual household disposable income designated as _topcoded_ (which occurs to ensure confidentiality of very high income individuals who might otherwise be identifiable).
+We used household economic panel data from Australia to provide the first investigation of whether changes in income and wellbeing have shifted the change point between 2001 to 2019.
+Australia is a developed western nation that has experienced a relatively stable period of economic growth in the last 19 years, which makes it a good candidate to observe how the function between income and subjective wellbeing has evolved over time without major economic shocks.
+HILDA (the Household, Income, and Labour Dynamics in Australia survey) provides a representative sample of households in Australia with detailed measurements of income and subjective wellbeing each year, which makes it an excellent data source to investigate the present question.
+We distinguished between satisfaction and happiness as different components of subjective wellbeing, and evaluated how each varied with household income over time using piecewise linear regression to determine the change points in each year.
 
@@
 #### Income
-To allow comparison with other major studies of income and wellbeing we used household after-tax income as the indicator of income and economic security [e..g, @kahneman2010high; @jebb2018happiness].
+To allow comparison with other major studies of income and wellbeing we used household after-tax income as the indicator of income and economic security [e..g, @kahneman2010high].
 Household income better represents economic security than personal income, since members of the same household share expenses as well as risks; i.e., they can provide a direct and immediate support network when financial shocks occur, which a priori might affect wellbeing.
-The ‘real household annual disposable income’ was calculated from the self-reported combined income of all household members after receipt of government pensions and benefits and deduction of income taxes in the financial year ended 30th June of the year of the wave (e.g., 2002 in wave 2).
-This was then adjusted for inflation - the rise in the general price level of the economy - using the Australian Bureau of Statistics (ABS) Consumer Price Index, so that income in all waves is expressed in FY 2017/18 prices, to give real income.
+The ‘real household annual disposable income’ was calculated from the self-reported combined income of all household members after receipt of government pensions and benefits and deduction of income taxes in the financial year ended 30th June of the year of the wave (e.g., 2001 in wave 1).
+This was then adjusted for inflation - the rise in the general price level of the economy - using the Australian Bureau of Statistics (ABS) Consumer Price Index, so that income in all waves is expressed in 2019 prices, to give real income.
 
 The _equivalised_ household income was obtained by adjusting for household size (the number of adult and child household members).
-In this instance, we have used the 'modified OECD' scale (Hagenaars et al., 1994), which divides household income by 1 for the first household member plus 0.5 for each other household member aged 15 or over, plus 0.3 for each child under 15.
+We used the 'modified OECD' scale [@hagenaars1994poverty], which divides household income by 1 for the first household member plus 0.5 for each other household member aged 15 or over, plus 0.3 for each child under 15.
 A family comprising two adults and two children under 15 years of age would therefore have an equivalence scale of 2.1 (1 + 0.5 + 0.3 + 0.3), meaning that the family would need to have an income 2.1 times that of a single-person household in order to achieve the same standard of living.
@@
 
-For modelling, both dependent variables were rescaled with a mean of zero and a SD of 1 (z-scores) for each year.
-
 <br>
@@
 
-We modelled the relationship between income and each wellbeing variable (happiness and satisfaction) using a simple **linear model** and a **piecewise model** (broken-stick).
+We modelled the relationship between income and each wellbeing variable (happiness and satisfaction) using a simple linear model and a piecewise model with a single change-point.
 The piecewise model was chosen as the simplest extension of a linear model which can identify a change point (inflection) in the relationship between wellbeing and income.
@@
 
-For modelling, both measures of wellbeing were rescaled with a mean of zero and a SD of 1 (z-scores) for each year.
+For modelling, both measures of wellbeing and income were rescaled with a mean of zero and a SD of 1 (z-scores) for each year.
 
-_Model Estimation_  
+_Model Design_  
 We adopted a Bayesian approach for estimating the linear and piecewise model in the software Stan [@burkner2017; @rstan2019]. In each case the linear model was estimated as:
@@
 
-Where $X_i$ was an individual's household income (\$) as well as other covariates (age, age^2^, sex, education), and $y_i$ was an individual's wellbeing.
+Where $X_i$ was an individual's household income (\$) as well as other covariates (age, age^2^, sex, education, chronic illness), and $y_i$ was an individual's wellbeing.
 
@@
 
-Where $x_i$ was an individual's household income, and $X_i$ were covariates for age, age^2^, sex and education.
+Where $x_i$ was an individual's household income, and $X_i$ were the cross-sectional HILDA population weights to adjust for differences in the sample representativeness according to sex by broad age, marital status, region, and labour force status.
 
@@
 
-The above models estimated population-level effects separately for each year (*t* = 2002...2018).
+The above models estimated population-level effects separately for each year (*t* = 2001...2019).
 Because we were interested in the location of the change point between income and wellbeing that existed across individuals within each year, we ignored the panel design of HILDA because the dependency between observations of the same person across years was orthogonal to our effects of interest.
-We specified weakly informed priors for each *&beta;*, and a uniform prior over the restricted range of income values for *&omega;*.
+
+The parameters of the piecewise model used Gaussian priors centred at zero, with a regularized parameterization for the slopes and intercept of Normal(0, 0.1) for each *&beta;*, and a slightly less skeptical regularization for the change point Normal(0, 0.5).
+These regularizing priors were selected to reduce overfitting while still allowing the model to learn the regular features of the sample, and so provide a more robust population estimate.
+They are skeptical priors and a prior predictive check confirmed they assume no relationship between income and wellbeing, with no difference in gradient before or after the change point (Supplementary materials, Figure S4).
 
 _Model Selection_  
-To determine whether wellbeing was a linear or non-linear (e.g., piecewise) function of income, we compared the linear and piecewise model posterior probabilities using the [Widely Applicable Information Criterion (WAIC)](https://bookdown.org/ajkurz/Statistical_Rethinking_recoded/overfitting-regularization-and-information-criteria.html#the-problem-with-parameters){target="_blank"}.
-The WAIC is the log-posterior predictive density plus a penalty proportional to the variance in the posterior distribution.
-Thus it provides an approximation of the out-of-sample deviance that converges to the cross-validation approximation in a large sample, with a penalty for the effective number of parameters (degrees of freedom).
-For this reason is it useful to compare two models of varying complexity, such as our linear and piecewise model.
+To determine whether wellbeing was a linear or non-linear (e.g., piecewise) function of income, we compared the out-of-sample deviance of the linear and piecewise model fits using the [Widely Applicable Information Criterion (WAIC)](https://bookdown.org/ajkurz/Statistical_Rethinking_recoded/overfitting-regularization-and-information-criteria.html#the-problem-with-parameters){target="_blank"}.
+The WAIC is the log-posterior predictive density plus a penalty proportional to the variance in the posterior distribution, so it represents an approximation of the out-of-sample deviance that converges to the cross-validation accuracy in a large sample, with a penalty for the effective number of parameters (degrees of freedom).
+Thus using the out-of-sample deviance for model selection in combination with regularizing priors in our model design is a dual strategy to reduce overfitting and penalize overfitting.
+This makes it a useful approach for comparing models of varying complexity, such as our linear and piecewise (nonlinear) models.
+As with other deviance metrics, smaller negative WAIC values are better (i.e., indicate more accuracy).
 
@@
 _Covariates_  
-Age (and age^2^), gender, and university graduate education were included as covariates.
-Cross-sectional population weights for Australia provided by the University of Melbourne for each year were also included as a covariate to adjust for differences in the sample representativeness according to sex by broad age, marital status, region, and labour force status.
-Full-time students were removed, as well as individuals with an annual household disposable income that was indicated as _topcoded_ by the University of Melbourne (topcoding occurs to ensure privacy of high wealth individuals).
-Gender was included as a binary variable (Male = 1), and education was a binary variable coded from the highest level of education achieved (university/college graduate = 1).
+Cross-sectional population weights for Australia provided by the University of Melbourne for each year were included as a covariate to adjust for differences in the sample representativeness.
+Full-time students were removed, as well as individuals with an annual household disposable income that was indicated as _topcoded_ by the University of Melbourne.
 
 ```{r read_write_covs}
-happywealth <- read_rds("../data/happywealth.rds")
+hilda_data <- read_rds("../data/hilda_data.rds")
 ```
@@
 The broad demographic characteristics of the sample are presented in Supplementary Materials Table S1.
-Average life satisfaction levels were very steady between 2002-2018, while average happiness score decreased slightly over the 17 years.
-The proportions of each sex and relationship status were stable over time, as were the average household size and SEIFA index.
+Household income and average life satisfaction levels increased between 2001-2019, while average happiness scores decreased slightly over the 19 years.
+The proportions of each sex and couples were stable over time, as were the average household size and SEIFA index.
 However age, education, and chronic health conditions tended to slightly increase over time.
-For instance, average age increased by 1.8 years over the 17 years of the survey, which is obviously less than would occur in a cohort study [@watson2012hilda].
+For instance, average age increased by 2.3 years over the 19 years of the survey, which is obviously less than would occur in a cohort study [@watson2012hilda].
 Changes in the workforce varied with economic circumstances.
@@
 ##### Figure 1. Household income and happiness (red) and satisfaction (blue)  
-```{r losat_gh9_hifdip}
-source("../src/happyfun.r")
+```{r hh_disp_inc_eq}
+source("../src/plotfun.r")
 
-p1 <- happywealth %>%
-  rename(`Satisfaction 0-10` = losat, 
-         `Household income` = hh_disp_inc_eq) %>% 
-  filter(!student, !top_hifdip) %>%
+hilda_data %>%
+  rename(`Household income` = hh_disp_inc_eq) %>%
+  filter(
+    !student,
+    !top_hifdip,
+    year %in% c(2001, 2005, 2010, 2015, 2019), # 2001, 2007, 2013, 2019
+    `Household income` >= 0
+  ) -> figure1_data
+
+p1 <- figure1_data %>%
+  rename(`Happiness 1-100` = gh9) %>% 
+  plot_linear(`Happiness 1-100`, `Household income`) +
+  scale_fill_manual(values = "#C79999") +
+  scale_color_manual(values = "#8F2727")
+
+p2 <- figure1_data %>%
+  rename(`Happiness 1-100` = gh9) %>% 
+  plot_inflection_custom(`Happiness 1-100`, `Household income`) +
+  scale_fill_manual(values = "#C79999") +
+  scale_color_manual(values = "#8F2727")
+
+p3 <- figure1_data %>%
+  rename(`Satisfaction 0-10` = losat) %>% 
   plot_linear(`Satisfaction 0-10`, `Household income`) +
@@
 
-p2 <- happywealth %>%
-  rename(`Satisfaction 0-10` = losat, 
-         `Household income` = hh_disp_inc_eq) %>% 
-  filter(!student, !top_hifdip) %>%
-  plot_inflection_piecewise(`Satisfaction 0-10`, `Household income`) +
+p4 <- figure1_data %>%
+  rename(`Satisfaction 0-10` = losat) %>% 
+  plot_inflection_custom(`Satisfaction 0-10`, `Household income`) +
   scale_fill_manual(values = "#b3cde0") +
   scale_color_manual(values = "#005b96")
-
-p3 <- happywealth %>%
-  rename(`Happiness 1-100` = gh9, 
-         `Household income` = hh_disp_inc_eq) %>% 
-  filter(!student, !top_hifdip) %>%
-  plot_linear(`Happiness 1-100`, `Household income`) +
-  scale_fill_manual(values = "#C79999") +
-  scale_color_manual(values = "#8F2727")
-
-p4 <- happywealth %>%
-  rename(`Happiness 1-100` = gh9, 
-         `Household income` = hh_disp_inc_eq) %>% 
-  filter(!student, !top_hifdip) %>%
-  plot_inflection_custom(`Happiness 1-100`, `Household income`, 
-                         knots = c(43.7, 49.2, 47.4, 53.2, 72.3)) +
-  scale_fill_manual(values = "#C79999") +
-  scale_color_manual(values = "#8F2727")
 ```
 ```{r figure_1, fig.height=10, fig.width=9, dpi=300}
-p3 / p4 / p1 / p2
+p1 / p2 / p3 / p4
 ```
@@
 The nonlinear relationship between happiness and income (Figure 1, 2nd row) was consistently and negatively inflected (happiness increased less with income after the change point).
-By contrast the nonlinear relationship between satisfaction and income shown in the 4th row was as likely to be negatively inflected as positively inflected: positively inflected in 2002 and 2006; negatively inflected 2014 and 2018; and no apparent inflection in 2010.
+By contrast the nonlinear relationship between satisfaction and income shown in the 4th row exhibited negative inflection (2001, 2015, 2019), positive inflection (2005), and no apparent inflection (2010).
 
-The posterior evidence from model selection revealed the nonlinear (piecewise) fit of _happiness_, but not _satisfaction_, was credibly superior to a linear fit in each year (Supplementary Materials, Figure S2).
-Overall the posterior evidence suggests that happiness and satisfaction have distinct relationships with household income; satisfaction tends to increase linearly with income, while a change point exists in the relationship between happiness and household income.
+The evidence from model selection revealed the nonlinear (piecewise) fit of _happiness_, but not _satisfaction_, was superior to a linear fit in each year (Supplementary Materials, Figure S1).
+The WAIC difference indicated the piecewise fit of happiness had reliably less deviance than the linear fit, indicating the piecewise model predicted happiness from income more accurately using this out-of-sample metric.
+Overall the model comparison suggested that happiness and satisfaction have distinct relationships with household income; satisfaction tends to increase linearly with income, while a change point exists in the relationship between happiness and household income.
 
@@
 
-#### The evolution of income on happiness 2002-2018
+#### The evolution of the association between income and happiness 2001-2019
 
-Figure 2 below presents the posterior distribution of each parameter from the piecewise model regressing happiness on income: the change point (&omega;), the intercept (&beta;~0~), the pre-change slope (&beta;~1~), and the post-change slope (&beta;~2~).
-Over 12,000 samples were drawn from each piecewise model to determine the posterior distribution of each parameter.
-Horizontal bars represent the 95% credible interval and so intervals which fall completely to the right of the vertical grey dotted line are credibly higher than our base year 2002.
+Figure 2 below presents the posterior distribution of each parameter from the piecewise model regressing happiness on income: the change point (&omega;), the intercept (&beta;~0~), the pre-change point slope (&beta;~1~), and the post-change point slope (&beta;~2~).
+Over 4,000 samples were drawn from each piecewise model to determine the posterior distribution of each parameter (smallest bulk effective sample size = 661, see Table S2 for the complete fit statistics in Supplementary materials).
+Horizontal bars represent the 95% credible interval of the posterior distribution and so intervals which fall completely to the right of the vertical grey dotted line are credibly higher than the expected value of our base year, 2001.
 
@@
 
-##### Figure 2. Posterior parameters of income on happiness  
+##### Figure 2. Posterior parameters of the dependency of happiness on income (real 2019 dollars)   
 ```{r get_posteriors}
-path_to_gh9_fits <- list.files(
-  path = "../results/incl_topcodes/with_pweights/",
-  pattern = "^fit_nl_hifdip_gh9_4cov_pwt_.*.RDS$",
-  full.names = TRUE
-)
+income_summary <- read_rds("../results/gh9_hhoecd_data.RDS")
+fits <- read_rds("../results/gh9_hhoecd_0cov_tight.RDS")
 
-posteriors_b0 <- list()
-posteriors_b1 <- list()
-posteriors_b2 <- list()
-posteriors_alpha <- list()
-
-for (pathtofile in path_to_gh9_fits) {
-
-  df <- read_rds(pathtofile) %>%
-    as.data.frame()
-
-  b0 <- select(df, b_b0_Intercept)
-  b1 <- select(df, b_b1_Intercept)
-  b2 <- select(df, b_b2_Intercept)
-  alpha <- select(df, b_alpha_Intercept)
-
-  column_name <- paste0("param_",
-                        substring(pathtofile, nchar(pathtofile)-7, nchar(pathtofile)-4))
-
-  colnames(b0) <- column_name
-  colnames(b1) <- column_name
-  colnames(b2) <- column_name
-  colnames(alpha) <- column_name
-
-  posteriors_b0 <- bind_cols(posteriors_b0, b0)
-  posteriors_b1 <- bind_cols(posteriors_b1, b1)
-  posteriors_b2 <- bind_cols(posteriors_b2, b2)
-  posteriors_alpha <- bind_cols(posteriors_alpha, alpha)
-
-}
-
-smooth_by_year <- function(posteriors) {
-
-  posteriors %>%
-    transmute(
-      param_2002 = (param_2001 + param_2002)/2,
-      param_2003 = (param_2002 + param_2004)/2,
-      param_2004 = (param_2003 + param_2004)/2,
-      param_2005 = (param_2004 + param_2005)/2,
-      param_2006 = (param_2005 + param_2006)/2,
-      param_2007 = (param_2006 + param_2007)/2,
-      param_2008 = (param_2007 + param_2008)/2,
-      param_2009 = (param_2008 + param_2009)/2,
-      param_2010 = (param_2009 + param_2010)/2,
-      param_2011 = (param_2010 + param_2011)/2,
-      param_2012 = (param_2011 + param_2012)/2,
-      param_2013 = (param_2012 + param_2013)/2,
-      param_2014 = (param_2013 + param_2014)/2,
-      param_2015 = (param_2014 + param_2015)/2,
-      param_2016 = (param_2016 + param_2016)/2,
-      param_2017 = (param_2017 + param_2017)/2,
-      param_2018 = (param_2018 + param_2018)/2
-    ) -> smoothed
-
-  return(smoothed)
-
-}
+posteriors <- map_dfr(fits, as.data.frame, .id = "fit") %>%
+  left_join(bind_rows(income_summary, .id = "fit")) %>%
+  mutate(omega_dollars = (b_omega_Intercept * sd + mean)/1000) %>%
+  as_tibble()
 ```
@@
 ```{r figure_2, fig.width = 6, fig.height = 7, dpi=300}
-posteriors_omega <- posteriors_alpha %>%
-  mutate_all(.funs = ~brms::inv_logit_scaled(., 0, 100)*10)
+param_key = c(
+  omega_dollars = "Change point ($000s)",
+  b_b0_Intercept = paste0("Intercept (\u03b2", "0)"),
+  b_b1_Intercept = paste0("Pre-slope (\u03b2", "1)"),
+  b_b2_Intercept = paste0("Post-slope (\u03b2", "2)")
+)
 
-smooth_by_year(posteriors_omega) %>% 
-  mutate(param = "omega ($000s)") %>% # (\u03c9)
-  bind_rows(., (mutate(smooth_by_year(posteriors_b0), 
-                       param = paste0("intercept (\u03b2", "0)")))) %>%
-  bind_rows(., (mutate(smooth_by_year(posteriors_b1), 
-                       param = paste0("preslope (\u03b2", "1)")))) %>%
-  bind_rows(., (mutate(smooth_by_year(posteriors_b2), 
-                       param = paste0("postslope (\u03b2", "2)")))) %>%
-  gather(key = "year", val = "sample", -param) %>% 
+posteriors %>%
+  select(fit, contains("b0"), contains("b1"), contains("b2"), omega_dollars) %>%
+  gather(key = "param", val = "sample", -fit) %>%
   mutate(
-    # year = str_remove(year, "param_"),
-    year = extract_numeric(year),
-    param = fct_relevel(param, "omega ($000s)")) %>% 
-  filter(year > 2001) %>%
+    year = extract_numeric(fit),
+    param = recode_factor(param, !!!param_key)
+    ) %>% 
   group_by(param, year) %>%
@@
     upp = quantile(sample, probs = 0.95, na.rm=T)
-  ) -> d0
+  ) -> parameter_summary
 
-d1 <- filter(d0, year == 2002)
+base_year <- filter(parameter_summary, year == 2001)
 
-ggplot(d0, aes(y = year, x = expected, color = param)) +
+ggplot(parameter_summary, aes(y = year, x = expected, color = param)) +
   geom_pointrange(aes(xmin = low, xmax = upp)) +
-  geom_vline(aes(xintercept = expected), d1, color = "gray", 
+  geom_vline(aes(xintercept = expected), base_year, color = "gray", 
              linetype = 3) +
-  scale_y_continuous(breaks = c(2002, 2006, 2010, 2014, 2018)) +
+  scale_y_continuous(breaks = c(2001, 2007, 2013, 2019)) +
   facet_wrap(~param, scales = "free_x") +
   ft_theme(base_size = 12) +
-  scale_color_manual(values = c("red", "grey65", "grey65", "grey65")) +
+  scale_color_manual(values = c("#A25050", "grey65", "grey65", "grey65")) +
   theme(legend.position = "none",
@@
 ```
-Figure 2 legend: Posterior distributions of the change point parameter omega representing the location in real household income (2018 dollars), as well as the intercept, preslope and postslope parameters in arbitrary units.
+Figure 2 legend: Posterior distributions of the change point parameter representing the location in real household income (real 2019 dollars), as well as the intercept, pre-slope and post-slope parameters in SD units (happiness).
 Horizontal bar represents the 95% credible region and the solid point indicates the expected value (median) of each distribution.
-Vertical dotted line indicates the 2002 parameter value (median) as a base year comparison.
+Vertical dotted line indicates the 2001 expected value (median) as a base year comparison.
 ```
@@
 
-The posterior estimates of the change point between happiness and household income indicates that the location (i.e., the real income value) of the change point shows a systematic increasing trend since 2011.
-Changes to the other parameters of the function between income and happiness also occured between 2002 and 2018 (i.e., the pre-slope, post-slope and intercept), however these did not show any sustained trend over the period nor result in a reliable change from the baseline year 2002.
+The posterior estimates of the change point in the association between happiness and household income indicates that the probable location (i.e., the real income value) of the change point shows a systematic increasing trend since 2001.
+Changes to the other parameters of the function between income and happiness also occurred between 2001 and 2019 (i.e., the pre-slope, post-slope and intercept), but did not show any sustained trend over the period.
 
-It is also worth noting that both the pre-slope and post-slope parameters (&beta;~1~ and &beta;~2~) were credibly larger than zero in most years, indicating there was a reliable dependency between happiness and income at income levels below and above the change point.
-However comparing the pre-slope with the post-slope values makes clear that the relationship between happiness and income among the majority of people in lower income households was an order of magnitude steeper than those in high income households.
+Both the pre-slope and post-slope parameters (&beta;~1~ and &beta;~2~) were credibly larger than zero in each year, indicating there was a reliable dependency between happiness and income at income levels both below and above the change point.
+However the pre-slope parameter was in general four to five times greater than the post-slope values, indicating that the relationship between happiness and income among people in lower income households was an order of magnitude stronger than those in high income households.
 
-The change in parameter values between 2002 to 2018 indicates the relationship between happiness and income evolved over time.
-We determined the impact of this evolution on the distribution of happiness over the range of household income in 2018 in a counterfactual analysis.
-The counterfactual analysis is a hypothetical demonstration of how happiness would change if people in 2018 were subject to the function that existed in 2002.
-That is, would happiness increase or decrease if the 2002 function was in place in 2018?
-It controls for changes in the sample which occur over time that are not related to happiness but could nevertheless contribute to changes in the distribution of happiness.
-For example, the increase in the range of (real) income levels in the economy between 2002 and 2018 could produce a greater divergence in happiness due to the increasing gap between the rich and poor - even with a stable relationship between income and happiness.
+Changes in the other demographic variables (e.g., age, education levels and chronic illness) did not materially alter the trends just described.
+The parameters of conditional piecewise models which included age, age^2^, education, and illness as covariates are compared in Supplementary materials (Figure S2).
+
+The change in parameter values between 2001 to 2019 indicates the relationship between happiness and income evolved over time.
+We determined the impact of this evolution on the distribution of happiness over the range of household income in 2019 in a counterfactual analysis.
+The counterfactual analysis is a hypothetical demonstration of how happiness would change if people in 2019 were subject to the function that existed in 2001, i.e., would happiness increase or decrease if the 2001 function was in place in 2019?
+This controls for changes in the sample which occur over time that are not related to happiness but could nevertheless contribute to changes in the distribution of happiness.
+For example, an increase in the range of (real) income levels in the economy between 2001 and 2019 could produce a greater divergence in happiness due to the increasing gap between the rich and poor - even with a stable relationship between income and happiness.
 Such changes in the sample characteristics may mask or confound the impact of the change point on the distribution of happiness without careful control.
-Because we were interested in the implications of the evolution of the function rather than changes in our sample characteristics _per se_, we estimated happiness levels for each _n_ = 14,280 person in 2018 using the 2002 function.
-These 2002 model-estimates were compared to (subtracted from) the 2018 model-estimates generated from the same sample (*n* = 14,280), to obtain the change (delta) in happiness for each person under the counterfactual.
-This delta must be entirely due to the evolution of the function between 2002 and 2018, since the analysis has held everything else constant (i.e., exactly the same people with the same characteristics were used to estimate happiness from two different functions).
-Figure 3 below presents the 14,280 deltas from such a comparison, along with a smoothed mean (blue overlay) to summarize how happiness evolved with the function across the entire income distribution.
+Because we were interested in the implications of the evolution of the function rather than changes in our sample characteristics _per se_, we estimated happiness levels for each person in 2019 (*n* = 14,459) using the 2001 function.
+These 2001 model-estimates were compared to (subtracted from) the 2019 model-estimates generated from the same sample (*n* = 14,459), to obtain the change (delta) in happiness for each person under the counterfactual.
+This delta is attributable to the evolution of the function between 2001 and 2019.
+Figure 3 below presents the 14,459 deltas from such a comparison, along with a smoothed mean (solid line) to summarize how the distribution of happiness evolved across the income range.
 
-##### Figure 3. Counterfactual difference in happiness between 2002 and 2018
-```{r counterfactual_estimates, eval=F}
-# Data from 2018
-data2018 <- happywealth %>%
-  filter(!student, year == 2018) %>%
+##### Figure 3. Counterfactual difference in happiness between 2019 and 2001
+```{r counterfactual_estimates}
+fits <- read_rds("../results/gh9_hhoecd_0cov_tight.RDS")
+
+# Data from 2019
+data2019 <- hilda_data %>%
+    filter(
+      year == 2019,
+      !student,
+      hh_disp_inc_eq >= 0,
+      !top_hifdip
+    ) %>%
     transmute(
-    y = as.vector(scale(gh9)), # losat or gh9
-    dollars = hifdip_adj/10000, 
-    pweights = as.vector(scale(weights)),
-    male = as.logical(male),
-    age = as.vector(scale(age)),
-    age_sq = age^2,
-    grad = edu %in% c("Grad", "Postgrad")
+      region,
+      y = as.vector(scale(gh9)),
+      income = hh_disp_inc_eq/1000,
+      shh_disp_inc_eq = scale(hh_disp_inc_eq),
+      dollars = as.vector(shh_disp_inc_eq),
     ) %>% 
@@
 
-# Model fits from 2002 and 2018
-fit2002 <- read_rds(path_to_fits[2])
-fit2018 <- read_rds(path_to_fits[18])
+data2019$Est.2019 <- predict(fits[["fit2019"]])[, 'Estimate']
+data2019$Est.2001 <- predict(fits[["fit2001"]], newdata = data2019)[, 'Estimate']
 
-# Model estimates from fits and data 
-fits_2018 <- fitted(fit2018, newdata = dat2018)
-fits_2002 <- fitted(fit2002, newdata = dat2018)
+# counterfacts <- map_dfr(.x = fits, .f = ~{
+#   predict(.x, newdata = data2019) %>%
+#     as_tibble()
+# }, .id = "fit")
+# 
+# counterfacts %>%
+#   group_by(fit) %>%
+#   mutate(
+#     Est.2019 = data2019$Est.2019,
+#     income = data2019$income) -> counterfacts
 
-write_rds(lst(fits_2002, fits_2018), path = "../results/counterfactual_estimates.rds")
+# write_rds(counterfacts, path = "../results/counterfactual_estimates.rds")
 ```
@@
 ```{r figure_3, fig.width=7, fig.height=7, dpi=300}
-estimates2002 <- read_rds("../results/counterfactual_estimates.rds")$fits_2002
-estimates2018 <- read_rds("../results/counterfactual_estimates.rds")$fits_2018
+data2019 %>%
+  mutate(
+    delta = Est.2019 - Est.2001
+  ) %>%
+  ggplot(aes(x = income)) +
+  geom_hline(yintercept = 0, color = "grey50", size = 0.4) +
+  geom_point(aes(y = delta), size = 0.1, alpha = 0.1) +
+  # geom_density(fill = "grey50", color = "grey50") +
+  stat_smooth(aes(y = delta), geom = "line", method = "loess", se = F, span = 5) +
+  xlim(0, 200) +
+  labs(
+    subtitle = "Happiness delta(s) from 2019 predictions less 2001 predictions (SD units)",
+    x = "Household income ($000s)", y = "",
+    caption = "Source: Predicted happiness scores from 2019 HILDA sample (n = 14,459)"
+  ) +
+  theme_test()
 
-# Data from 2018
-data2018 <- happywealth %>%
-  filter(!student, year == 2018) %>%
-  transmute(
-    y = as.vector(scale(gh9)), # losat or gh9
-    dollars = hifdip_adj/10000, 
-    pweights = as.vector(scale(weights)),
-    male = as.logical(male),
-    age = as.vector(scale(age)),
-    age_sq = age^2,
-    grad = edu %in% c("Grad", "Postgrad")
-  ) %>%
-  na.omit()
-
-# Difference between estimates (∆ = 2018 - 2002)
-data2018$delta = estimates2018[, 'Estimate'] - estimates2002[, 'Estimate']
-
-# Plot deltas
-ggplot(data = data2018, aes(x = dollars*10, y = delta)) +
-  geom_point(size = 0.1, alpha = 0.1) + 
-  geom_smooth() +
-  geom_vline(xintercept = 72, color = "grey50", size = 0.4) +
-  geom_hline(yintercept = 0, color = "grey50", size = 0.4) +
-  annotate("text", x = 96, y = -.25, label = "$72K (2018 change point)") +
-  xlim(0, 150) +
-  ylim(-.3, .3) +
-  labs(subtitle = "Happiness delta(s) from 2018 predictions less 2002 predictions (SD units)",
-       x = "Household income ($000s)", y = "",
-       caption = "Source: Predicted happiness scores from 2018 HILDA sample (n = 14,280)") +
-  theme_test() +
-  theme(
-    plot.caption = ggplot2::element_text(
-      hjust = 0,
-      size = ggplot2::rel(0.8),
-      margin = margin(t = 6)
-    )
-  )
+# counterfacts %>%
+#   mutate(
+#     year = str_remove(fit, "fit"),
+#     delta = Est.2019 - Estimate
+#   ) %>%
+#   filter(year %in% c(2001, 2010)) %>%
+#   ggplot(aes(x = income)) +
+#   geom_hline(yintercept = 0, color = "grey50", size = 0.4) +
+#   geom_density(fill = "grey50", color = "grey50") +
+#   stat_smooth(aes(y = delta), geom = "line", method = "loess", se = F, span = 50) +
+#   xlim(0, 200) +
+#   labs(
+#     subtitle = "Mean change in happiness (SD units)",
+#     x = "Household income ($000s)", y = "") +
+#   facet_wrap(~year) +
+#   theme_test()
 ```
@@
 ```
-Figure 3 legend: The difference (∆) in model-estimated happiness between 2018 and 2002 for the same n = 14,280 individuals from the 2018 survey.
-Values below zero on the y-axis indicate lower happiness estimates in 2018 relative to 2002.
-Points to the left of the vertial ($72K) are individuals below the 2018 change point.
-The smoothed overlay (blue) indicates how the average delta changes across the income distribution.
+Figure 3 legend: The difference (∆) in model-estimated happiness between 2019 and 2001 for the same n = 14,459 individuals from the 2019 survey.
+Values below zero on the y-axis indicate lower happiness estimates in 2019 relative to 2001.
+The smoothed overlay (solid line) indicates how the average happiness changes across the income distribution.
 ```
@@
 
-Figure 3 shows that, on average, happiness estimates improved for people with household incomes above \$72K when compared to 2002, as indicated by the average delta (smooth blue line) falling above zero on the right side of the plot.
-Conversely, people with household incomes below \$72K, on average, suffered a decrease in estimated happiness compared to 2002.
-And the greatest decrease in happiness occurred for people whose income levels were transliminal: that is, their income was above the changepoint in 2002 but the shift in the change point resulted in their income falling below the change point in 2018.
+Figure 3 shows that, on average, happiness estimates improved for people with household incomes above \$55K in 2019 when compared to 2001, as indicated by the average delta (solid line) falling above zero on the right side of the plot.
+Conversely, people with household incomes below \$55K, on average, suffered a decrease in estimated happiness in 2019 compared to 2001, particularly those whose income levels were transliminal: that is, their income was above the changepoint in 2001 but the shift in the change point resulted in their income falling below the change point in 2019 (i.e., ~$50K, see Figure S6 in Supplementary materials).
 Of course the obtained deltas are due to changes in all the parameters of the function, including the slope before and after the change point.
-However because this comparison was performed on the same individuals (i.e., from the 2018 survey), it held characteristics such as age, income, etc, constant that would otherwise be expected to change over time and possibly contribute to any difference in happiness distribution.
-In this way these results isolate the amount of change entirely due to the _evolution_ of the function between 2002 and 2018, and demonstrates how this has contributed to the unequal distribution of happiness between the rich and the poor over time.
+However because this comparison was performed on the same individuals from the 2019 survey, it held characteristics such as age, income, etc, constant that would otherwise be expected to change over time and possibly contribute to any difference in happiness distribution.
+In this way these results isolate the amount of change entirely due to the _evolution_ of the function between 2001 and 2019, and demonstrates how this has contributed to a more unequal distribution of happiness between the rich and the poor over time in Australia.
 
@@
 
-A less hypothetical implication of the shifting change point in the relationship between income and happiness is that over time fewer people have an income that affords them happiness, which is (relatively) independent of their financial security.
-Assuming median income is also not increasing at a faster rate, any increase in the change point will reduce the number of people who fall above it over time.
-Figure 5 presents median household income levels weighted for the Australian population (by age, sex, marital status, labour force participation and geographical region).
-It also shows the change point between income and happiness increased faster than rises in median household income between 2002 and 2018.
-The third panel shows that as a result, a smaller proportion of the Australian population in 2018 had a household income above the changepoint than in 2002.
+Assuming median income is also not increasing at a faster rate, any increase in the change point will reduce the number of people who fall above it over time; an implication being that over time more people's happiness is more dependent on their income level than previously.
+Figure 4 presents median household income levels weighted for the Australian population (by age, sex, marital status, labour force participation and geographical region).
+This shows the change point between income and happiness increased faster than rises in median household income between 2001 and 2019.
+The third panel shows that as a result, a smaller proportion of the Australian population in 2019 had a household income above the changepoint than in 2001.
 
-##### Figure 4. Rise in median income, change point, and population in Australia 2002-2018 
+##### Figure 4. Rise in median income, change point, and population in Australia 2001-2019 
 ```{r figure_4, fig.width=7, fig.height=6, dpi=300}
-trimmed_omega <- posteriors_omega %>%
+fits <- read_rds("../results/gh9_hhoecd_0cov_tight.RDS")
+income_params <- read_rds("../results/gh9_hhoecd_summary.RDS")
+
+posteriors <- map_dfr(fits, as.data.frame, .id = "fit") %>%
+  left_join(bind_rows(income_params, .id = "fit")) %>%
+  mutate(omega_dollars = b_omega_Intercept * sd + mean)
+
+posteriors %>%
+  select(fit, omega_dollars) %>%
+  group_by(fit) %>%
   mutate(
-        # param_2002 = (param_2001 + param_2002)/2,
-        param_2003 = (param_2002 + param_2004)/2,
-        # param_2004 = (param_2003 + param_2004)/2,
-        # param_2005 = (param_2004 + param_2005)/2,
-        # param_2006 = (param_2005 + param_2006)/2,
-        # param_2007 = (param_2006 + param_2007)/2,
-        # param_2008 = (param_2007 + param_2008)/2,
-        param_2009 = (param_2008 + param_2009)/2,
-        param_2010 = (param_2009 + param_2011)/2,
-        # param_2011 = (param_2010 + param_2011)/2,
-        # param_2012 = (param_2011 + param_2012)/2,
-        # param_2013 = (param_2012 + param_2013)/2,
-        # param_2014 = (param_2013 + param_2014)/2,
-        # param_2015 = (param_2014 + param_2015)/2,
-        # param_2016 = (param_2015 + param_2016)/2,
-        # param_2017 = (param_2016 + param_2017)/2,
-        # param_2018 = (param_2017 + param_2018)/2
-        )
+    sampleid = seq(1, 4000),
+    omega_dollars = omega_dollars/1000
+    ) %>%
+  ungroup() %>%
+  spread(fit, omega_dollars) %>%
+  select(-sampleid) -> df
 
 # Tabulate against median income
-happywealth %>%
+hilda_data %>%
   filter(!student) %>%
@@
   select(year, xwaveid, income = hh_disp_inc_eq, weights) %>%
-  filter(year %in% 2002:2018) %>%
+  filter(year %in% 2001:2019) %>%
   mutate(
-    year = paste0("param_", year),
+    year = paste0("fit", year),
     income = income/1000) %>%
@@
     gather(
-      summarise_all(trimmed_omega, list(median)), 
+      summarise_all(df, list(median)), 
       year, 
@@
     `Median household income\n ($000s)` = round(median(income), 2),
-    # `Median household income\n ($000s)` = round(matrixStats::weightedMedian(income, weights)),
-
     `Change point\n ($000s)` = round(mean(change_point), 2),
-
     `Population above change\n(%)` = round(
       sum(weighted_over) / sum(weights) *100)
-
   ) %>%
@@
 df.plot %>%
-  select(key) %>% distinct() %>%
-  mutate(string = c("post-GFC\n(2009 onwards)", "", "")) -> gfc.note
-
-df.plot %>%
-  filter(year %in% c(min(year), 2018)) %>%
+  filter(year %in% c(min(year), 2019)) %>%
   mutate(
@@
               paste0("$", round(value), "K"))),
-    vpos = if_else(val %in% c("$49K", "$72K", "42%"), -0.75, 1.65),
+    vpos = if_else(val %in% c("$51K", "$74K"), -0.75, 1.65),
     hpos = if_else(year == min(year), 0.3, 0.8)
@@
 
-p.econ <- p + 
-  labs(title = "Income & happiness in Australia between 2002-2018",
-       subtitle = " ") +
-  ggthemes::scale_color_economist() +
-  ggthemes::theme_economist(base_family="Verdana", base_size = 8) +
-  theme(legend.position = "none",
-        strip.text = element_text(size = 10)
-        )
-
 p + 
@@
 Figure 4 legend: Real household income has stagnated in Australia since 2009 (post GFC) while the change point between happiness and income has increased.
-Consequently fewer Australians had an income that affords them happiness, which is independent of their financial security in 2018.
+Consequently fewer Australians had an income that affords them happiness, which is independent of their financial security in 2019.
 ```
@@
 
-We confirmed previous findings [e.g., @howell2008relation; @kahneman2010high], that the relationship between both types of subjective wellbeing with household income was positive but quite different over a 17 year period: Satisfaction increased linearly with income, while happiness increased rapidly up to a point after which higher levels of income were associated with less improvement.
-For the first time we have shown that this change point in the relationship between household income and happiness increased faster than inflation or the median household income over time between 2002 and 2018.
+We confirmed previous findings [e.g., @howell2008relation; @kahneman2010high], that the relationship between both types of subjective wellbeing with household income was positive but quite different: Satisfaction increased linearly with income, while happiness increased rapidly up to a point after which higher levels of income were associated with less improvement.
+Our novel finding was that this change point in the relationship between household income and happiness increased faster than both inflation and median household income between 2001 and 2019.
+For the first time we have shown there has been a temporal shift in the change point between income and happiness over the 19 year period, such that happiness has become more dependent on income for more people - especially the poor and middle-class.
 
@@
 After this point, happiness is no longer as dependent on household income, and the economic security it represents.
-Presumably after this point further increases in happiness depend more on other life factors (e.g., leisure time, social connections).
+Presumably after this point further increases in happiness depend more on other life factors (e.g., leisure time, social connections) than financial security.
 Life satisfaction on the other hand appeared to show consistent increases with household income and we found no evidence of any change point.
@@
 
-An implication of the changing relationship between happiness and income since 2002 is that income inequality may be driving increasing inequities in wellbeing.
-This is demonstrated in Figure 3, where the difference in happiness between 2002 and 2018 increased for incomes above $72K/year and decreased for incomes below that level.
-The inequity was also highlighted in Figure 4 where the change point of happiness represented a 9% increase over median income in 2002, while in 2018 it represented a 42% increase over median income.
-This increase relative to median income also represented a reduction from 42% to 24% in the proportion of people whose income fell above the changepoint.
-Thus we can see that over the last sixteen years the difference in happiness between the rich and the poor has increased; while the proportion of people whose happiness no longer depends on their financial security has decreased.
+An implication of the changing relationship between happiness and income is that income inequality may be driving increasing inequities in wellbeing.
+This is demonstrated in Figure 3, where the difference in happiness between 2001 and 2019 increased for incomes above $50K/year and decreased for incomes below that level (see also Figure S6 in Supplementary materials).
+The inequity was also highlighted in Figure 4 where the change point of happiness represented a 9% increase over median income in 2001, while in 2019 it represented a 42% increase over median income.
+This increase relative to median income also represented a reduction from 43% to 26% in the proportion of people whose income fell above the changepoint.
+Thus we can see that over the last nineteen years the difference in happiness between the rich and the poor has increased; while the happiness of an increasing proportion of people, including the middle-class, is more dependent on their financial security.
 
-Australia has low levels of income disparity relative to many other OECD countries, and the Gini coefficient has not changed a great deal between 2002 and 2018 [@australian2018rising].
+Australia has low levels of income disparity relative to many other OECD countries, and the Gini coefficient has not changed a great deal between 2001 and 2019 [@australian2018rising].
 A stable Gini coefficient shows income inequality has remained steady over the time period, and our results do not conflict with this conclusion.
-Rather what we are revealing is that even a static income distribution may have dynamic effects on happiness over time.
-We think this highlights the issue that while traditional measures of wealth and income inequality may be relatively stable and exhibit little change, their impact on wellbeing and health can still vary.
-As such, we believe these results may well have relevance to other developed nations in North America and Europe which have enjoyed stable economic growth overall, but have stagnating incomes, and declining happiness levels [@sachs2019world].
+Rather we are revealing that even a static income distribution may have dynamic effects on happiness over time.
+We think this highlights the issue that while traditional univariate measures of wealth and income inequality may be relatively stable and exhibit little change, their impact on wellbeing and health can still vary.
+As such, these results may well have relevance to other developed nations in North America and Europe which have also enjoyed stable economic growth overall, but have stagnating incomes, and declining happiness levels [@sachs2019world].
 As focus shifts from traditional financial indicators towards health and wellbeing measures, findings such as this may become more prevalent.
 
-Some recent studies have challenged either the notion that the positive effect of  income plateaus after some level, or that the effect on happiness and satisfaction are distinct [@killingsworth2021experienced; @twenge2020expanding; @jebb2018happiness].
-Most recently the point has been made that experienced happiness continues to increase on a log scale with income and does not plateau after USD75,000 [@killingsworth2021experienced].
-It is worth noting that log-linear models are typically used to investigate income effects, and we would agree that our data do not support a strict view of a "satiety" point or flat gradient between income and happiness which can persist on a log scale [@jebb2018happiness; kahneman2010high].
-But log-linear models mask the fact that marginal dollars matter less the more one earns, which is what we observed for happiness here.
-Furthermore, log models [and more flexible additive models @jebb2018happiness] may produce a better fit, but will not provide a clear or a distinct change point allowing us to test a change in the marginal effect of income.
-And so one implication of our model choice is that the change in slope may be much smoother than implied by our model.
-We concede this point but would still contend the location of the change in slope will display a comparable shift over time, if it can be detected with certainty.
+The relationship between happiness and income is often described as log-linear [e.g., @kahneman2010high; @killingsworth2021experienced].
+Under this view, similar proportional changes in income or wealth produce similar evaluations of satisfaction or happiness, rather than the absolute change.
+Likewise, the piecewise model reported here has a slope that reduces with wealth, albeit with a single identifiable change-point rather than a smooth curve (*nb*., in the Bayesian setting, the slope is smooth over the posterior of the change-point).
+However our main thesis is not committed to a function with a single or smooth changepoint; only that the slope between income and happiness is unequal at different levels of income, and the function is shifting to the right over time (e.g., Figure 2).
+The choice between functions is somewhat arbitrary, since a log-linear function is as easily consistent with our thesis as the piecewise function we describe.
+Nevertheless, the model evidence favoured a piecewise- over a log-linear function in a supplementary analysis (Figure S3), which validates our otherwise arbitrary model choice here.
 
-Past researchers have also argued there is no distinct relationship between income and the different forms of subjective wellbeing (e.g., happiness, satisfaction), and that income has the same impact on both constructs.
-It is rare for such research to directly test differences in the form of the function, usually instead inferring no difference after observing null results.
-One benefit of our model selection was the quantification of the posterior evidence for a linear over a nonlinear function (e.g., Figure S1, Supplementary Materials).
-That showed the evidence for a linear effect was clearly different between our two wellbeing variables, and the evidence for a linear function was also decreasing over time.
-It may be that with more recent data, or in the near future, there is no distinct function between income and the different wellbeing constructs.
-However it is also worth noting that the only meta-analysis conducted in this area, over ten years ago, also suggested the linear relationship between income and life satisfaction is stronger than that with happiness [@howell2008relation].
-We would also note that a linear function between income and life satisfaction as we report here, is consistent with the majority of earlier research [@stevenson2013subjective].
-In general, the importance of differences between the distinct constructs of subjective wellbeing will need to be adjudicated when deciding how to measure income effects.
+Some recent studies have challenged the notion that the positive effect of income plateaus after some level ("satiety") [@killingsworth2021experienced; @twenge2020expanding; stevenson2013subjective].
+Our results also do not support a "satiety" point or flat gradient between income and happiness, as the average post-change slope between 2001-2019 was reliably above zero.
+The pre-change slope however was also four to five times greater than that, and we argue this difference in slopes indicates a previously unacknowledged inequity in the distribution of happiness between the rich and poor.
+While the amount of variance in happiness explained or predicted by household income may be small, the slope is reliably above zero and small changes in happiness over an entire economy of 25M people or more will aggregate.
+Moreover, happiness is measured on a subjective scale, where small changes at different points on the scale may represent a large change in some other functional/real-world outcome (e.g., risk of suicide).
+This is all to say that quantifying the real-world impact of small changes in a subjective measure such as happiness is difficult and speculative without further investigation.
+Our aim here is to merely determine whether the inequity is present at all.
 
@@
 
-Supplementary information is available [here](supplementary.html).
+Supplementary information is available [online](https://datarichard.github.io/The-increasing-cost-of-happiness/supplementary.html){target="_blank"}.
 

### 2021-03-12 src/draft0_word.Rmd
Major update v0.4

--- 2021-03-12 src/draft0.Rmd
@@
 ```{r get_posteriors}
-income_summary <- read_rds("../results/gh9_hhoecd_data.RDS")
-fits <- read_rds("../results/gh9_hhoecd_0cov_tight.RDS")
+income_summary <- read_rds("../results/fits/excl_topcodes/sd/gh9_hhoecd_data.RDS")
+fits <- read_rds("../results/fits/excl_topcodes/sd/gh9_hhoecd_0cov_tight.RDS")
 
@@
 ```{r counterfactual_estimates}
-fits <- read_rds("../results/gh9_hhoecd_0cov_tight.RDS")
+fits <- read_rds("../results/fits/excl_topcodes/sd/gh9_hhoecd_0cov_tight.RDS")
 
@@
 ```{r figure_4, fig.width=7, fig.height=6, dpi=300}
-fits <- read_rds("../results/gh9_hhoecd_0cov_tight.RDS")
-income_params <- read_rds("../results/gh9_hhoecd_summary.RDS")
+fits <- read_rds("../results/fits/excl_topcodes/sd/gh9_hhoecd_0cov_tight.RDS")
+income_params <- read_rds("../results/fits/excl_topcodes/sd/gh9_hhoecd_data.RDS")
 
@@
 
-Supplementary information is available [online](https://datarichard.github.io/The-increasing-cost-of-happiness/supplementary.html){target="_blank"}.
+Supplementary information is available [online](https://datarichard.github.io/The-increasing-cost-of-happiness/supplementary.html).
 

### 2021-03-12 src/supplementary0.Rmd
Major update v0.4

---
title: "Supplementary material: The increasing cost of happiness"
author: "R.W. Morris^1,2^, N. Kettlewell^2,3,4^  & N.Glozier^1,5^"
bibliography: ../src/references.bib
csl: ../src/elsevier-harvard.csl
output:  
  html_document:
    reference_docx: ../src/academic_style.docx
always_allow_html: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(brms)
library(flextable)
library(gtsummary)
library(patchwork)

library(bayesplot)

knitr::opts_chunk$set(eval = TRUE,
                      echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      fig.path = "../figures/",
                      ft.align="left")

source("ft_themes.r")
```
\  
\

1. Central Clinical School, Faculty of Medicine and Health, University of Sydney, NSW, Australia
2. ARC Centre of Excellence for Children and Families over the Life Course, University of Queensland, QLD, Australia
3. School of Economics, University of Technology, NSW, Australia
4. Institute of Labor Economics (IZA), Bonn, Germany
5. Brain and Mind Centre, University of Sydney, NSW, Australia

\

**Corresponding author:**  
Professor Nick Glozier  
Faculty of Medicine and Health,   
University of Sydney,  
NSW 2050,  
Australia  
email: nick.glozier@sydney.edu.au

\

Draft: `r format(Sys.time(), '%d %B, %Y')`  
Supplementary tables:      4  
Supplementary figures:     6

**keywords:** Subjective wellbeing, household income, HILDA

\

***

## Supplementary Material

<br>

#### Sample characteristics

The broad demographic characteristics of the sample at each sixth year wave are presented below in Table S1. The linear trend as a function of time was estimated for each variable (linear regression of time for continuous variables and Cochrane-Armitage test in the case of binary variables), and Bon-Ferroni adjusted p-values for multiple comparisons are presented (k = 11).

```{r table_s1}
prop_trend_test <- function(data, variable, by, ...) {

  data %>%
    select(x = {{variable}}, y = {{by}}) %>%
    dplyr::filter(complete.cases(.)) %>%
    dplyr::group_by(y) %>%
    dplyr::summarise(
      xsum = sum(x, na.rm = T),
      n = n()
    ) %>%
    {prop.trend.test(x = .$xsum, n = .$n)} %>%
    broom::tidy() %>%
    mutate(method = "Cochrane-Armitage test",
           p.value = p.value * 11,
           p.value = if_else(p.value > 1, 1, p.value))
}

linear_trend_test <- function(data, variable, by, ...) {

  data %>%
    select(y = {{variable}}, x = {{by}}) %>%
    dplyr::filter(complete.cases(.)) %>%
    {lm(formula = y ~ ordered(x), data = .)} %>%
    broom::tidy() %>%
    filter(str_detect(term, ".L")) %>%
    mutate(method = "linear trend analysis (parametric)",
           p.value = p.value * 11,
           p.value = if_else(p.value > 1, 1, p.value))

}

hilda_data <- read_rds("../data/hilda_data.rds")

hilda_data %>%
  filter(year %in% c(2001, 2007, 2013, 2019)) %>%
  filter(!student, !top_hifdip, (!is.na(losat)|!is.na(gh9))) %>%
  mutate(
    `University Graduates` = edu == "Grad",
    ) %>%
  select(
    year, `Household income ($)` = hh_disp_inc_eq,
    `Life satisfaction` = losat, `Happiness` = gh9, 
    Males = male, `Age (years)` = age, `University Graduates`, #Workforce,
    Unemployed = unemployed, Couples = coupled, #`Relationship status`, 
    `Chronic illness` = chronic, SEIFA, 
    `Household size` = hhsize,
  ) %>% 
  tbl_summary(
    by = year,
    missing = "no",
    statistic = 
      list(all_continuous() ~ "{mean} (±{sd})",
           all_categorical() ~ "{n} ({p}%)"),
    digits = 
      list(all_continuous() ~ 1,
           `Household income ($)` ~ 0),
    missing_text = "(Missing)" 
    ) %>%
  add_p(
    test = list(
      all_continuous() ~ "linear_trend_test",
      all_dichotomous() ~ "prop_trend_test"
    )
  ) %>%
  modify_header(stat_by = "{level}\nN = {style_number(n)}") %>%
  as_flex_table() %>%
  set_caption(caption = "Table S1. Demographic characteristics of the sample") %>%
  align(part = "body") %>%
  valign(valign = "top", part = "header") %>%
  fontsize(size = 9.5, part = "all") %>%
  width(j = 1, width = 1.5) %>%
  width(j = 2:6, width = 1)
```

<br>

Household income and average life satisfaction increased between 2001-2019, while average happiness score decreased slightly over the 19 years. The proportions of each sex were stable over time, as was the proportion of people living as a couple (Couples), average household size and SEIFA index. The percentage unemployed varied with economic conditions however there was no significant positive or negative linear trend. Age and chronic health conditions tended to increase over time. The slight increase in the age of the panel over time (2.3 years) was substantially less than that of a cohort study which would be 18 years. Education levels (i.e., percentage of university graduates) also tended to increase over time.

Generally, Table S1 supports the view that the HILDA sample was a relatively stable representation of Australian socioeconomic conditions between 2001 and 2019. The importance of the specific demographic variables with a significant linear trend (age, age^2^, education, illness) were tested in regression models of happiness as covariates (see below Figure S3).

<br><br>

#### Model Comparisons

**Linear vs non-linear models.** We compared the out-of-sample evidence for the non-linear (piecewise) models with the linear models of income on wellbeing in each year, using the difference in WAIC deviance scores (i.e., WAIC~linear~ — WAIC~piecewise~). Thus a WAIC difference greater than zero indicated evidence for a linear relationship; a WAIC difference less than zero indicated evidence for a nonlinear (piecewise) relationship.

<br>

##### Figure S1. Linear model evidence (WAIC~linear~ — WAIC~piecewise~)
```{r figure_s1, dpi=300}
here <- "~/Dropbox/HILDA/The-increasing-cost-of-happiness/"

lm_fits = read_rds(paste0(here, "results/lm_gh9_hhoecd_0cov.RDS"))

nl_fits = read_rds(paste0(here, "results/gh9_hhoecd_0cov_tight.RDS"))

# get waic results
map2_dfr(.x = nl_fits, 
         .y = lm_fits, 
         .f = ~{
           loo_compare(.x, .y, criterion = "waic") %>%
             as.data.frame() %>%
             rownames_to_column("model") %>%
             as_tibble()
         },
         .id = "year") %>%
  filter(elpd_diff != 0) %>%
  mutate(
    year = extract_numeric(year),
    model = if_else(model == ".x", "nl", "lm"),
    waic_diff = 2 * elpd_diff, # lm disadvantage
    se = 2 * se_diff,
    wellbeing = "Happiness"
  ) -> waic_happiness

nl_fits_losat = read_rds(paste0(here, "results/losat_hhoecd_0cov.RDS"))

lm_fits_losat = read_rds(paste0(here, "results/lm_losat_hhoecd_0cov.RDS"))

# plot waic results
map2_dfr(.x = nl_fits_losat, 
         .y = lm_fits_losat, 
         .f = ~{
           loo_compare(.x, .y, criterion = "waic") %>%
             as.data.frame() %>%
             rownames_to_column("model") %>%
             as_tibble()
         },
         .id = "year") %>%
  filter(elpd_diff != 0) %>%
  mutate(
    year = extract_numeric(year),
    model = if_else(model == ".x", "nl", "lm"),
    waic_diff = if_else(model == "lm", 
                        2 * elpd_diff,   # nl advantage
                        -2 * elpd_diff), # nl advantage
    se = 2 * se_diff,
    wellbeing = "Satisfaction"
  ) -> waic_satisfaction

bind_rows(waic_happiness, waic_satisfaction) %>% 
  ggplot(aes(x = waic_diff, y = year, color = wellbeing)) +
    geom_vline(aes(xintercept = 0), color = "grey90") +
    geom_pointrange(aes(xmin = waic_diff - se*1.96, xmax = waic_diff + se*1.96)) +
    scale_y_continuous(breaks = c(2001, 2007, 2013, 2019)) +
    ft_theme() +
    labs(x = "deviance", y = "") +
    facet_wrap(~wellbeing, ncol = 1) +
    scale_color_manual(values  = c("#8F2727", "#005b96")) +
    theme(legend.position = "none")
```

```
Figure S1 legend: Differences in WAIC scores (deviance units) for a linear fit over a piecewise fit for happiness (red) and satisfaction (blue). The filled circle indicates the mean of the difference and the horizontal bars represents the 95% interval. A difference below zero (grey vertical line) is support for the piecewise model (a difference above zero is support for the linear model).
```

<br>

The difference in WAIC scores for each model revealed the nonlinear (piecewise) fits of _happiness_, but not _satisfaction_, provided superior out-of-sample accuracy over a linear fit in each year. Figure S1 shows the 95% interval of the differences in WAIC scores clearly distinguished the advantage of the nonlinear fits for happiness with no overlap with zero in any year. By contrast, the difference in WAIC between the linear and nonlinear models of life satisfaction was not distinguished from zero for most years (with the exception of 2017 and 2008), indicating the (out-of-sample) accuracy of the two models was largely indistinguishable over the 19 years. Overall the evidence suggested that household income has a distinct nonlinear relationship with happiness; implying the existence of a change point or difference in slopes at different levels of household income. This is consistent with the only meta-analysis conducted in this area, over ten years ago, which concluded the linear relationship between income and life satisfaction is stronger than that with happiness [@howell2008relation]. It is also worth noting that a linear function between income and life satisfaction is consistent with the majority of  research since then [@stevenson2013subjective].

**Conditional vs unconditional models of happiness.** Table S1 (above) showed the specific variables that changed over the 19 years between 2001-2019. In particular, the number of university graduates, average age and the incidence of chronic illness increased slighty over the period. Including these variables as covariates in any model (i.e,. a conditional model such that *P*( $y$ | income, age, education illness)) will improve the in-sample fit (e.g., R^2^), and also adjust the parameter estimates of interest (e.g., change point) conditional upon their common association with happiness. Given the general aim of the main paper was to describe the temporal change in functional form between happiness and income in Australia, regardless of any other variable (i.e., an unconditional model of *P*($y$| income)) or sampling differences, neither the in-sample fit nor the conditional model estimates are directly relevant. Nevertheless, the importance of these variables in mediating the changes we observed in the unconditional model are relevant, and so we calculated the estimates from conditional piecewise models including age, age^2^, sex, education  (university degree), chronic illness for each year below to confirm the pattern of temporal change was similar. Figure S2 shows the addition of these covariates did not change the trend in change point parameter estimates (red) relative to the unconditional parameter estimates (grey) over the 19 years.

<br>

##### Figure S2. Conditional model parameters  
```{r figure_s2, fig.height=8, dpi=300}
here <- "~/Dropbox/HILDA/The-increasing-cost-of-happiness/"

nl_fits = read_rds(paste0(here, "results/gh9_hhoecd_0cov_tight.RDS"))

nl_fits_5cov = read_rds(paste0(here, "results/gh9_hhoecd_5cov_tight.RDS"))

income_params <- read_rds("../results/gh9_hhoecd_summary.RDS")

posteriors_5c <- map_dfr(nl_fits_5cov, as.data.frame, .id = "fit") %>%
  left_join(bind_rows(income_params, .id = "fit")) %>%
  mutate(omega_dollars = (b_omega_Intercept * sd + mean)/1000) %>%
  as_tibble()

param_key = c(
  omega_dollars = "Change point ($000s)",
  b_b0_Intercept = paste0("Intercept (\u03b2", "0)"),
  b_b1_Intercept = paste0("Pre-slope (\u03b2", "1)"),
  b_b2_Intercept = paste0("Post-slope (\u03b2", "2)")
)

posteriors_5c %>%
  select(fit, contains("b0"), contains("b1"), contains("b2"), omega_dollars) %>%
  gather(key = "param", val = "sample", -fit) %>%
  mutate(
    year = extract_numeric(fit),
    param = recode_factor(param, !!!param_key)
  ) %>% 
  group_by(param, year) %>%
  summarise(
    low = quantile(sample, probs = 0.05, na.rm=T),
    expected = quantile(sample, probs = 0.5, na.rm=T),
    upp = quantile(sample, probs = 0.95, na.rm=T)
  ) %>%
  mutate(model = "conditional") -> parameter_summary_5c

posteriors_0c <- map_dfr(nl_fits, as.data.frame, .id = "fit") %>%
  left_join(bind_rows(income_params, .id = "fit")) %>%
  mutate(omega_dollars = (b_omega_Intercept * sd + mean)/1000) %>%
  as_tibble()

posteriors_0c %>%
  select(fit, contains("b0"), contains("b1"), contains("b2"), omega_dollars) %>%
  gather(key = "param", val = "sample", -fit) %>%
  mutate(
    year = extract_numeric(fit),
    param = recode_factor(param, !!!param_key)
  ) %>% 
  group_by(param, year) %>%
  summarise(
    low = quantile(sample, probs = 0.05, na.rm=T),
    expected = quantile(sample, probs = 0.5, na.rm=T),
    upp = quantile(sample, probs = 0.95, na.rm=T)
  ) %>%
  mutate(model = "unconditional") -> parameter_summary_0c

bind_rows(parameter_summary_0c, parameter_summary_5c) %>%
  ggplot(aes(y = year, x = expected, color = model)) +
  geom_pointrange(aes(xmin = low, xmax = upp), position = position_dodge(width = 0.75)) +
  scale_y_continuous(breaks = c(2001, 2007, 2013, 2019)) +
  facet_wrap(~param, scales = "free_x") +
  ft_theme(base_size = 12) +
  scale_color_manual(values = c("#C79999", "grey65")) +
  theme(legend.position = "none",
        axis.title = element_blank())
```
```
Figure S2 legend: Comparison of the conditional parameter estimates (red) and unconditional parameter estimates (grey) in the change point model of income on happiness. The conditional models included covariates for age, age squared, sex, education, and illness. Horizontal bar represents the 95% credible region and the solid point indicates the expected value (median) of each distribution.
```

<br>

The parameter estimates of the conditional model (red) were adjusted for the broad demographic variables which tended to increase over the 19 years in our sample (age, education, illness); however the pattern of changes over time was comparable to the unconditional model (grey), indicating these temporal trends were not responsible for the evolution in the change point between income and happiness.

<br><br>

#### Additional non-linear fits

**Log-linear models.** The relationship between happiness and income is often described as log-linear [e.g., @kahneman2010high]. Under this view, similar proportional changes in income or wealth produce similar evaluations of satisfaction or happiness, rather than the absolute change. That is, a \$100 increase in wealth will produce a greater impact on someone earning a minimum wage than someone earning many orders of magnitude of that amount. Likewise, the piecewise model that we employed can also adopt an increasing monotonic function whose slope reduces with wealth, albeit with a single identifiable change-point rather than a smooth curve. However our main thesis is not committed to a function with a single discrete changepoint; only that the slope between income and happiness is unequal at different levels of income, and the function is shifting to the right over time. The choice between functions is somewhat arbitrary, since a log-linear function is as easily consistent with our thesis as the piecewise function we described in the main paper. To resolve this, we compared the evidence for a log-linear fit against the piecewise fits according to their WAIC scores. Thus a WAIC difference less than zero indicated evidence for the piecewise fit over the log-linear fit (i.e., greater out-of-sample accuracy).

<br>

##### Figure S3. Log-linear model evidence (WAIC~log-linear~ — WAIC~piecewise~)
```{r, figure_s3, fig.height=3.5, dpi=300}
here <- "~/Dropbox/HILDA/The-increasing-cost-of-happiness/"

nl_fits = read_rds(paste0(here, "results/gh9_hhoecd_0cov_tight_x0.RDS"))

log_fits = read_rds(paste0(here, "results/log_gh9_hhoecd_0cov.RDS"))

# get waic results
map2_dfr(.x = nl_fits, 
         .y = log_fits, 
         .f = ~{
           loo_compare(.x, .y, criterion = "waic") %>%
             as.data.frame() %>%
             rownames_to_column("model") %>%
             as_tibble()
         },
         .id = "year") %>%
  filter(elpd_diff != 0) %>%
  mutate(
    year = extract_numeric(year),
    model = if_else(model == ".x", "nl", "log-linear"),
    waic_diff = 2 * elpd_diff, # lm disadvantage
    se = 2 * se_diff
  ) %>%
  ggplot(aes(x = waic_diff, y = year)) +
  geom_vline(aes(xintercept = 0), color = "grey90") +
  geom_pointrange(aes(xmin = waic_diff - se*1.96, xmax = waic_diff + se*1.96),
                 color = "#ffcc80") +
  scale_y_continuous(breaks = c(2001, 2007, 2013, 2019)) +
  ft_theme() +
  labs(x = "deviance", y = "")
```
```
Figure S3 legend: The joint distribution between wellbeing and income for a single year, imposed by the priors.
```

<br>

The piecewise fit was superior to a log-linear fit of income on happiness in each year, although the difference was less distinguishable in 2003, 2006, and 2008 where the 95% interval overlapped zero. This indicates the evidence for a piecewise model was generally greater than for a log-linear model, however it does not necessarily imply the presence of a discrete change point exists. In a Bayesian change point model, the estimated location of the change point in each year is not a single point but is rather a probability distribution of income values that extends over the entire range of household income. Importantly the width of the 95% posterior probability distribution in Figure 2 indicates the change point location is likely over a range of values, on average ±$9.15K, which implies a more or less smooth function over that range.

<br><br>

#### Prior and residual checks

We used regularizing Guassian priors centred on zero, with Normal(0, 0.1) for each *&beta;*, and a Normal(0, 0.5) for *&omega;*. A plot of the joint distribution between income and wellbeing imposed by the priors alone confirmed there was no apparent slope or change point implied by their regularizing effect.

##### Figure S4. Prior predictive check  
```{r figure_s4, dpi=300}
load(paste0(here, "results/prior_predictions.Rdata"))

prior_predictions %>%
  ggplot(aes(x = hh_disp_inc_eq/1000, y = Estimate, group = 1)) +
  geom_point(size = 0.1, alpha = 0.1) +
  stat_smooth(color = "#8F2727", size = 0.5, se = F) +
  xlim(c(0, 300)) +
  labs(y = "Wellbeing estimate (SD units)", x = "Household income ($000s)") +
  theme_test()
```
```
Figure S4 legend: The joint distribution between wellbeing and income for a single year, imposed by the priors. The solid blue line indicates a smoothed average (loess).
```  
<br><br>

Examination of the residual plots from the piecewise models of happiness confirms the residuals were evenly distributed around zero and there was little indication of heteroscedascity. Overall, there was little evidence that the piecewise models of happiness were not appropriate for the data.

<br>

##### Figure S5. Residual plots for piecewise happiness models
```{r figure_s5, fig.height=9, fig.width=9}
# https://www.flutterbys.com.au/stats/tut/tut7.2b.html
# 
# 
# residual_fits <- map_dfr(nl_fits, ~{
#   
#   d0 <- resid(.x) %>%
#     as_tibble() %>%
#     select(resids = Estimate)
#   
#   d1 <- fitted(.x) %>%
#     as_tibble() %>%
#     select(fitteds = Estimate)
#   
#   bind_cols(d0, d1)
#   
#   }, .id = "fit") %>%
#   mutate(year = extract_numeric(fit))
# 
# save(residual_fits, file = paste0(here, "results/residual_fits.Rdata"))

load(paste0(here, "results/residual_fits.Rdata"))

residual_fits %>%
  group_by(year) %>%
  slice_sample(n = 1000) %>% # to reduce overplotting
  ggplot(aes(x = fitteds, y = resids)) +
    geom_point(alpha = 0.1, size = 0.1) +
    facet_wrap(~year, scales = "free") +
    theme_test()
```

```
Figure S5 legend: The joint distribution between residual and fitted values for each year.
```

<br><br>

#### Parameter and fit summary statistics

There were _N_ = 59,876 total observations. 4000 samples (post-warmup) were drawn using sampling(NUTS). For each parameter, Bulk_ESS and Tail_ESS are effective sample size (ESS) measures, and Rhat is the potential scale reduction factor on split chains. In each case the Rhat approached 1, indicating convergance. Each ESS _n_ > 300, which is a conservative threshold for estimation.

<br>

```{r table_s2, ft.align="left"}
param_key = c(
  b0_Intercept = "Intercept",
  b1_Intercept = "Pre-slope",
  b2_Intercept = "Post-slope",
  omega_Intercept = "Change-point"
)

map_dfr(nl_fits, 
        .f = ~{
          fit_summary <- summary(.x)

          sigma_summary <- fit_summary$spec_pars %>%
            as_tibble(rownames = "Parameter")

          fit_summary$fixed %>%
            as_tibble(rownames = "Parameter") %>%
            bind_rows(sigma_summary) %>%
            select(Parameter, everything())

          },
        .id = "Year") %>%
  mutate(
    Year = str_remove(Year, "fit"),
    Parameter = recode_factor(Parameter, !!!param_key)
    ) %>%
  flextable() %>%
  merge_v(j = 1) %>%
  valign(j = 1, valign = "top") %>%
  bg(i = ~Year %in% seq(2001, 2019, 2), 
     bg = "grey90", 
     part = "body") %>%
  colformat_num(j = 3:6, digits= 3) %>%
  colformat_num(j = 7, digits= 1) %>%
  set_caption("Table S2. Piecewise parameter estimates and fit statistics for happiness") %>%
  autofit()
```

<br>

```{r table_s3, ft.align="left"}
param_key = c(
  b0_Intercept = "Intercept",
  b1_Intercept = "Pre-slope",
  b2_Intercept = "Post-slope",
  omega_Intercept = "Change-point"
)

map_dfr(lm_fits_losat, 
        .f = ~{
          fit_summary <- summary(.x)

          sigma_summary <- fit_summary$spec_pars %>%
            as_tibble(rownames = "Parameter")

          fit_summary$fixed %>%
            as_tibble(rownames = "Parameter") %>%
            bind_rows(sigma_summary) %>%
            select(Parameter, everything())

          },
        .id = "Year") %>%
  mutate(
    Year = str_remove(Year, "fit"),
    Parameter = recode_factor(Parameter, !!!param_key)
    ) %>%
  flextable() %>%
  merge_v(j = 1) %>%
  valign(j = 1, valign = "top") %>%
  bg(i = ~Year %in% seq(2001, 2019, 2), 
     bg = "grey90", 
     part = "body") %>%
  colformat_num(j = 3:6, digits= 3) %>%
  colformat_num(j = 7, digits= 1) %>%
  set_caption("Table S3. Linear parameter estimates and fit statistics for life satisfaction") %>%
  autofit()
```

<br><br>

#### Model predictions (2001-2019)

The posterior estimates of the piecewise model indicated that the relationship between happiness and income changed over time between 2001 and 2019. One implication of such a change is the disparity in happiness between income groups has increased. This will occur as more people fall below the change point over time and so are subject to the steep region of the function where income and happiness are strongly related. To directly examine the implications of our model for such inequities in happiness, we used the model to estimate or _predict_ the happiness of two different income levels: one income level which was above the change point in 2001 but fell below it by 2019 (\$50K/yr); and another level which always remained above the change points over the same period (\$75K/yr). Figure S8 below presents the happiness levels (in SD units/year) for each income level as well as the difference in happiness between them (∆).

<br>

##### Figure S6. Happiness (SD units) at $50K/yr and $75K/yr from 2001-2019  
```{r figure_s6, fig.width=7, fig.height=5, dpi=300}
here <- "~/Dropbox/HILDA/The-increasing-cost-of-happiness/"

nl_fits = read_rds(paste0(here, "results/gh9_hhoecd_0cov_tight.RDS"))

# Optional
# nl_fits_5cov = read_rds(
#   paste0(here, "results/fits/excl_topcodes/sd/gh9_hhoecd_5cov_tight.RDS")
# )

income_params <- read_rds("../results/gh9_hhoecd_summary.RDS")

bind_rows(income_params, .id = "fit") %>%
  mutate(low = (40000 - mean)/sd,
         high = (75000 - mean)/sd) %>%
  column_to_rownames(var = "fit") %>%
  as.matrix() -> fixed_income_hash

colnames(fixed_income_hash) <- NULL

fixed_income_fit <- tibble()

for (name in names(nl_fits)) {

  newdat = data.frame(
    dollars = fixed_income_hash[name, 3:4]
  )

  df <- fitted(nl_fits[[name]], newdata = newdat) %>% 
    as_tibble() %>%
    mutate(Dollars = c("$50K", "$75K"),

           fit = name)

  fixed_income_fit <- bind_rows(fixed_income_fit, df)

}

fixed_income_fit %>%
  mutate(year = extract_numeric(fit)) %>%
  select(Estimate, Dollars, year) %>%
  filter(Dollars %in% c("$75K", "$50K")) %>%
  spread(Dollars, Estimate) %>%
  mutate(`$75K - $50K (∆)` = `$75K` - `$50K`) %>%
  gather(income, happiness, -year) %>%
  mutate(income = fct_relevel(income, "$50K", "$75K")) %>%
  ggplot(aes(x = year, y = happiness, color = income)) +
  geom_point(size = 4, alpha = 0.25) +
  geom_smooth(method = "loess", se = F, span = 3) +
  facet_wrap(~income) +
  labs(caption = "Source:     HILDA, University of Melbourne
                 Predicted happiness scores (SD unit) from piecewise regression on real household income", 
       x = "", y = "") -> p

p +
  ft_theme() +
  ggthemes::scale_colour_wsj() +
  theme(legend.position = "none",
        plot.title = element_text(size = 19),
        strip.text = element_text(size = 12))
```

```
Figure S6 legend: The difference (∆) in happiness between a household income of $50K per year and a household income of $75K per year has increased between 2001 and 2019.
```

<br>

Figure S6 shows the disparity in happiness between two fixed income levels, \$50K/year (traversing the change points) and \$75K/year (above the change points). In 2001, a household income of \$50K/year achieved an above average level of happiness relative to everyone else that year, however by 2019 happiness had declined to lower than average levels relative to the population. By contrast, a household income of \$75K/year enjoyed a higher than average level of happiness for the entire period. The difference (∆) in happiness between these two income levels doubled over the period. The increasing disparity in happiness between the rich and the poor implies that happiness has became more inequitable over time.

<br><br>

#### In-sample fit statistics

Our primary objective was to understand the relationships between the parameters of our model, and so in-sample fit statistics such as R^2^ are irrelevant for this objective. Moreover, our use of regularized priors will bias such in-sample estimates towards zero as regularization deliberately sacrifices in-sample variance for out-of-sample accuracy, which obviously hinders interpretation of the in-sample values. Nevertheless, some readers may find it useful to compare the in-sample fit statistics between the three different model types employed here: piecewise, linear and log-linear, as well as between the conditional and unconditional models. Because the usual fit statistics such as R^2^ present a problem for Bayesian fits, as the variance of the predicted values can be larger than the variance of the data results in R^2^ values greater than 1, we present the in-sample statistics from frequentist OLS model fits (without regularization or penalized likelihoods).

```{r table_s4}
library(segmented)

hilda_data %>%
  filter(
    !student,
    !top_hifdip
  ) %>%
  filter(hh_disp_inc_eq > 0) %>%
  group_by(year) %>%
  transmute(
    y = as.vector(scale(gh9)),
    dollars = hh_disp_inc_eq,
    male = male,
    age = as.vector(scale(age)),
    age_sq = age^2,
    grad = edu %in% c("Grad", "Postgrad"),
    illness = chronic,
    weights = as.vector(scale(weights))
  ) %>%
  na.omit() %>%
  nest() -> nested_data

nested_data %>%
  mutate(
    pw_fit = map(data, ~segmented(lm(
      formula = y ~ dollars, 
      data = .x),
      seg.Z = ~dollars)),
    pw_fit_c = map(data, ~segmented(lm(
      formula = y ~ dollars + male + age + age_sq + grad + illness + weights, 
      data = .x),
      seg.Z = ~dollars)),
    lin_fit = map(data, ~lm(
      formula = y ~ dollars, 
      data = .x)),
    lin_fit_c = map(data, ~lm(
      formula = y ~ dollars + male + age + age_sq + grad + illness + weights, 
      data = .x)),
    log_fit = map(data, ~lm(
      formula = y ~ log(dollars), 
      data = .x)),
    log_fit_c = map(data, ~lm(
      formula = y ~ log(dollars) + male + age + age_sq + grad + illness + weights, 
      data = .x))
    ) -> nested_models

bind_rows(
  map_dfr(nested_models$log_fit, broom::glance) %>%
    mutate(model = "log", year = 2001:2019),
  map_dfr(nested_models$lin_fit, broom::glance) %>%
    mutate(model = "lin", year = 2001:2019),
  map_dfr(nested_models$pw_fit, broom::glance) %>%
    mutate(model = "pw", year = 2001:2019),
  map_dfr(nested_models$log_fit_c, broom::glance) %>%
    mutate(model = "logc", year = 2001:2019),
  map_dfr(nested_models$lin_fit_c, broom::glance) %>%
    mutate(model = "linc", year = 2001:2019),
  map_dfr(nested_models$pw_fit_c, broom::glance) %>%
    mutate(model = "pwc", year = 2001:2019)
) -> model_summaries

model_summaries %>%
  select(model, year, r.squared) %>%
  spread(model, r.squared) %>%
  flextable() %>%
  colformat_num(j = 2:7, digits = 3) %>%
  set_caption("Table S4. In-sample fit statistics (R-squared)") %>%
  set_header_labels(
    lin = "Linear", 
    linc = "Linear\n(conditional)",
    log = "Log-linear",
    logc = "Log-linear\n(conditional)",
    pw = "Piecewise",
    pwc = "Piecewise\n(conditional)") %>%
  valign(i = 1, valign = "top", part = "header") %>%
  width(j = 2:7, width = 1)
```

<br>

A comparison of the relative differences in R^2^ values between models indicates the piecewise model produced a slightly higher in-sample fit over the other two models, while the linear model tended to have the lowest R^2^. The log-linear fit varied between both over the 19 years. As expected, adding covariates increased the R-squared in each case, but it did not change the relative difference between models.

Three things need to be observed when interpreting the R^2^ values: 1) While the R^2^ between income and happiness is low (~2% and ~14% in the unconditional and conditional models, respectively), the aggregate effect over the wider population (e.g., 25M in Australia) will be much larger; 2) Happiness (and to a lesser extent, income) is an imperfect and noisy measure so any effect of income on happiness is likely to be larger than that measured here; 3) Finally, happiness is measured on a subjective and arbitrary Likert scale so a small change in happiness at different points in the scale may represent a large change in some other functional/real-world outcome (e.g., risk of suicide). This is all to say that quantifying the real-world impact of small changes in variation reported here is difficult and speculative without further investigation.

<br><br>

## References

### 2021-03-12 src/supplementary0_word.Rmd
Major update v0.4

--- 2021-03-12 src/supplementary0.Rmd
@@
 output:  
-  html_document:
-    reference_docx: ../src/academic_style.docx
+  word_document:
+    reference_docx: ../src/Rmarkdown-for-Word-styles.docx
 always_allow_html: yes
@@
 
-lm_fits = read_rds(paste0(here, "results/lm_gh9_hhoecd_0cov.RDS"))
+lm_fits = read_rds(
+  paste0(here, "results/fits/excl_topcodes/sd/lm_gh9_hhoecd_0cov.RDS")
+)
 
-nl_fits = read_rds(paste0(here, "results/gh9_hhoecd_0cov_tight.RDS"))
+nl_fits = read_rds(
+  paste0(here, "results/fits/excl_topcodes/sd/gh9_hhoecd_0cov_tight.RDS")
+)
 
@@
 
-nl_fits_losat = read_rds(paste0(here, "results/losat_hhoecd_0cov.RDS"))
+nl_fits_losat = read_rds(
+  paste0(here, "results/fits/excl_topcodes/sd/losat_hhoecd_0cov.RDS")
+)
 
-lm_fits_losat = read_rds(paste0(here, "results/lm_losat_hhoecd_0cov.RDS"))
+lm_fits_losat = read_rds(
+  paste0(here, "results/fits/excl_topcodes/sd/lm_losat_hhoecd_0cov.RDS")
+)
 
@@
 
-nl_fits = read_rds(paste0(here, "results/gh9_hhoecd_0cov_tight.RDS"))
+nl_fits = read_rds(
+  paste0(here, "results/fits/excl_topcodes/sd/gh9_hhoecd_0cov_tight.RDS")
+)
 
-nl_fits_5cov = read_rds(paste0(here, "results/gh9_hhoecd_5cov_tight.RDS"))
+nl_fits_5cov = read_rds(
+  paste0(here, "results/fits/excl_topcodes/sd/gh9_hhoecd_5cov_tight.RDS")
+)
 
-income_params <- read_rds("../results/gh9_hhoecd_summary.RDS")
+income_summary <- read_rds("../results/fits/excl_topcodes/sd/gh9_hhoecd_data.RDS")
 
 posteriors_5c <- map_dfr(nl_fits_5cov, as.data.frame, .id = "fit") %>%
-  left_join(bind_rows(income_params, .id = "fit")) %>%
+  left_join(bind_rows(income_summary, .id = "fit")) %>%
   mutate(omega_dollars = (b_omega_Intercept * sd + mean)/1000) %>%
@@
 posteriors_0c <- map_dfr(nl_fits, as.data.frame, .id = "fit") %>%
-  left_join(bind_rows(income_params, .id = "fit")) %>%
+  left_join(bind_rows(income_summary, .id = "fit")) %>%
   mutate(omega_dollars = (b_omega_Intercept * sd + mean)/1000) %>%
@@
 
-nl_fits = read_rds(paste0(here, "results/gh9_hhoecd_0cov_tight_x0.RDS"))
+nl_fits = read_rds(
+  paste0(here, "results/fits/excl_topcodes/sd/exc_zeros/gh9_hhoecd_0cov_tight.RDS")
+)
 
-log_fits = read_rds(paste0(here, "results/log_gh9_hhoecd_0cov.RDS"))
+log_fits = read_rds(
+  paste0(here, "results/fits/excl_topcodes/sd/log_gh9_hhoecd_0cov.RDS")
+)
 
@@
 
-nl_fits = read_rds(paste0(here, "results/gh9_hhoecd_0cov_tight.RDS"))
+nl_fits = read_rds(
+  paste0(here, "results/fits/excl_topcodes/sd/exc_zeros/gh9_hhoecd_0cov_tight.RDS")
+)
 
@@
 
-income_params <- read_rds("../results/gh9_hhoecd_summary.RDS")
+income_summary <- read_rds("../results/fits/excl_topcodes/sd/gh9_hhoecd_data.RDS")
 
-bind_rows(income_params, .id = "fit") %>%
+bind_rows(income_summary, .id = "fit") %>%
   mutate(low = (40000 - mean)/sd,

### 2021-03-12 src/happyfun.R

@@ -1,258 +1,0 @@
-plot_linear <- function(.data, swb_col, dollars_col) {
- 
-
-  swb_colname <- rlang::enexpr(swb_col)
-  dollars_colname <- rlang::enexpr(dollars_col)
-  
-  .data %>%
-    select(xwaveid, 
-           year, 
-           # swb = enexpr(swb_col), dollars = enexpr(dollars_col)) %>%
-           swb = !!swb_colname, dollars = !!dollars_colname) %>%
-    filter(swb >= 0) %>%
-    filter(dollars >= 0) %>%
-    filter(year %in% c(2002, 2006, 2010, 2014, 2018)) -> df
-  
-  df %>%
-    group_by(year) %>%
-    summarise(n = n()) %>%
-    mutate(note = paste("n =", format(n, big.mark = ","))) -> df.sum
-  
-  df %>%
-    group_by(year) %>%
-    mutate(
-      dollar_bin = ntile(dollars, 10)
-    ) %>%
-    group_by(year, dollar_bin) %>%
-    summarize(
-      wellbeing = mean(swb, na.rm=T),
-      wealth = mean(dollars, na.rm=T)/1000
-      ) -> df.deciles
-  
-  df %>%
-    mutate(
-      wealth = dollars/1000,
-      wellbeing = swb
-    ) -> df.plot
-  
-  # Linear model
-  ggplot(df.plot, aes(x = wealth, y = wellbeing)) +
-    geom_smooth(aes(fill = "manual1", color = "manual2"),
-      method = "lm", 
-      size = 0.5
-      ) +
-    coord_cartesian(xlim = c(0, max(df.deciles$wealth)),
-                    ylim = c(min(df.deciles$wellbeing), max(df.deciles$wellbeing))) +
-    facet_wrap(~year, nrow = 1) +
-    labs(subtitle = paste(rlang::as_name(swb_colname), "(linear)"),
-         x = paste(rlang::as_name(dollars_colname), "($000s)"),
-         y = "") +
-    theme_test() +
-    theme(
-      legend.position = "none") -> p1
-  
-  p1 +
-    geom_point(data = df.deciles) + # size = 3, alpha = 0.25
-    geom_text(data = df.sum, aes(x = Inf, y = -Inf, label = note),
-              hjust = 1.1, vjust = -.75, colour = 1, size = 3)
-  
-}
-
-
[…]

### 2021-03-12 src/model fitting.R

@@ -1,243 +1,0 @@
-#### The Increasing Cost of Happiness ####
-# RW Morris, N Kettlewell, N Glozier
-# 
-# Model specification and fitting
-# Jan 8th 2021
-# 
-#### libraries ####
-library(dplyr)
-library(ggplot2)
-library(brms)
-options(mc.cores = 4)
-
-#### Import ####
-happywealth <- readRDS("../data/happywealth.rds")
-
-#### Preprocessing ####
-happywealth %>%
-  filter(!student, year == 2001) %>%
-  transmute(
-    y = as.vector(scale(gh9)), # gh9
-    dollars = hifdip_adj/10000, 
-    male = as.numeric(male),
-    age = as.vector(scale(age)),
-    age_sq = age^2,
-    grad = as.numeric(edu %in% c("Grad", "Postgrad"))
-  ) %>%
-  na.omit() -> df
-
-# Checking preprocessing...
-df %>%
-  mutate(decile = ntile(dollars, 10)) %>% 
-  group_by(decile) %>%
-  summarise(dollars = mean(dollars), 
-            y = mean(y)) %>%
-  ggplot(aes(x = dollars, y = y)) + geom_point()
-
-summary(df$dollars)
-#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
-#  0.000   2.697   4.176   4.758   6.065  40.311 2002 | gh9
-#  0.000   3.588   5.362   6.173   7.706  85.139 2018 | gh9
-
-summary(df$y)
-#  Min.     1st Qu.  Median  Mean    3rd Qu. Max. 
-#  -4.0191  -0.6054  0.1824  0.0000  0.7076  1.8893  2002 | gh9
-#  -4.90231 -0.54261 0.08021 0.00000 0.70302 1.32584 2002 | losat
-
-hist(df$dollars, breaks = 100)
-hist(df$y, breaks = 10)
-
-#### Fit the model ####
-#
-# This model was suggested at: 
-# https://discourse.mc-stan.org/t/piecewise-linear-mixed-models-with-a-random-change-point/5306/5
-# 
-# Model definition
-# b0 is the intercept denoting the expected value of wellbeing (y) at the changepoint
-# b1 is the linear slope before the changepoint
-# b2 is the linear slope after the changepoint
-# alpha is the coefficient of wealth (x) at the changepoint
-
-bfixed <- bf(
-  y ~ b0 + b1 * (dollars - omega) * step(omega - dollars) + 
[…]

### 2021-03-12 src/model selection.R

@@ -1,83 +1,0 @@
-#### Comparing models ####
-# https://bookdown.org/ajkurz/Statistical_Rethinking_recoded/overfitting-
-# regularization-and-information-criteria.html#the-problem-with-parameters
-# 
-# WAIC is a (penalized) estimate of out-of-sample deviance so smaller values (closer to 
-# zero) are better. Note also that p_waic is the penalty term of WAIC, and so it
-# will be close to the number of dimensions in the posterior of each model  
-
-lm_fit <- readRDS("../results/no_topcodes/fit_lm_hifdip_losat_cov_2002.RDS")
-pw_fit <- readRDS("../results/no_topcodes/fit_nl_hifdip_losat_cov_2002.RDS")
-
-# Comparing WAIC using loo_compare()
-# The best model will be reported at the top of the list with elpd_diff = 0
-ans <- loo_compare(lm_fit, pw_fit, criterion = "waic")
-
-print(ans, simplify = F)
-#         elpd_diff se_diff
-# pw_fit   0.0       0.0  
-# lm_fit -69.6      12.2  
-
-waic_results <- tibble()
-
-for (i in c(2001:2018)) {
-  
-  lm_fit <- readRDS(paste0("../results/no_topcodes/fit_lm_hifdip_losat_cov_", i, ".RDS"))
-  pw_fit <- readRDS(paste0("../results/no_topcodes/fit_nl_hifdip_losat_cov_", i, ".RDS"))
-  
-  ans <- loo_compare(lm_fit, pw_fit, criterion = "waic")
-  
-  as_tibble(ans, rownames = "model") %>%
-    mutate(year = i) %>%
-    mutate(across(elpd_diff:year, as.numeric)) -> tb
-  
-  # Add R-squared
-  lm_r2 <- bayes_R2(lm_fit) %>%
-    as_tibble() %>%
-    rename(R2 = Estimate, R2.Error = Est.Error, R2_Q2.5 = Q2.5, R2_Q97.5 = Q97.5) %>%
-    mutate(year = i, model = "lm_fit")
-  
-  pw_r2 <- bayes_R2(pw_fit) %>%
-    as_tibble() %>%
-    rename(R2 = Estimate, R2.Error = Est.Error, R2_Q2.5 = Q2.5, R2_Q97.5 = Q97.5) %>%
-    mutate(year = i, model = "pw_fit")
-  
-  tb %>%
[…]

### 2021-03-12 src/piecewise model simulation.R

@@ -1,480 +1,0 @@
-#### Author: RMorris
-#### Date: Jan 12th 2021
-
-library(tidyverse)
-library(brms)
-
-tibble(
-  year = rep(1:16, each = 20),
-  dollars = rep(1:10, 32)
-  ) %>% 
-  rowwise() %>%
-  mutate(
-    year = as.character(year),
-    dollars = dollars + rnorm(1),
-    b0 = 8,
-    b1 = 2,
-    b2 = 0.05,
-    omega = 5, #ceiling((year + 2)/2.5),
-    y = -log(1/(dollars + 1.5)) + rnorm(1),
-    age = sample(18:100, 1),
-    male = as.logical(sample(0:1, 1)),
-    grad = as.logical(rbinom(1, 1, 0.25))) %>%
-  ungroup() %>%
-  mutate(
-    y = as.vector(scale(y)),
-    age = as.vector(scale(age)),
-    agesq = age^2
-    ) -> df
-
-ggplot(df, aes(x = dollars, y = y)) +
-  geom_point() +
-  facet_wrap(~year)
-
-hist(df$dollars)
-hist(df$y)
-summary(df$dollars)
-#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
-# -2.094   3.038   5.559   5.478   7.930  12.232 
-
-#### Fit the model ####
-#
-# This model was suggested at: 
-# https://discourse.mc-stan.org/t/piecewise-linear-mixed-models-with-a-random-change-point/5306/5
-# 
-# Model definition
-# b0 is the intercept denoting the expected value of wellbeing (y) at the changepoint
-# b1 is the linear slope before the changepoint
-# b2 is the linear slope after the changepoint
-# alpha is the coefficient of wealth (x) at the changepoint
-
-
-bfixed <- bf(
-  y ~ b0 + b1 * (dollars - omega) * step(omega - dollars) + 
-    b2 * (dollars - omega) * step(dollars - omega) + b3*age + b4*agesq + b5*male + b6*grad,
-  # b0 + b1 + b2 + b3 + b4 + b5 + b6 + alpha ~ 1 + year, # fixed intercept/year (complete pooling)
-  # b0 + b1 + b2 + b3 + b4 + b5 + b6 + alpha ~ 1 + (1|year), # random intercept/year (no pooling)
-  # Sigmoid transform to keep omega within the range of 0 to x
-  nlf(omega ~ inv_logit(alpha) * 15),
-  nl = TRUE
-)
-
-# priors                                      
[…]

### 2021-03-12 src/preprocessing.R

@@ -148,29 +148,30 @@ happiness <- gh9_items %>%
 # prior to 2012. See Appendix in Wilkins (2014) 'Derived income variables in the
 # HILDA survey' for explanation
 # 
-# CPI values from https://www.pc.gov.au/research/completed/rising-inequality
+# CPI values from https://www.rba.gov.au/calculator/annualDecimal.html
 CPI = data.frame(
-  wave = letters[1:18],
-  deflator = c(1.52547554, # 2000/01
-               1.48315720, # 2001/02
-               1.43988458, # 2002/03
-               1.40607389, # 2003/04
-               1.37297463, # 2004/05
-               1.33027251, # 2005/06
-               1.29200230, # 2006/07
-               1.24993042, # 2007/08
-               1.21214575, # 2008/09
-               1.18464785, # 2009/10
-               1.14888718, # 2010/11
-               1.12303076, # 2011/12
-               1.09804401, # 2012/13
-               1.06903118, # 2013/14
-               1.05101802, # 2014/15
-               1.03670360, # 2015/16
-               1.01929187, # 2016/17
-               1)) %>% mutate(inflator = rev(deflator))
+  wave = letters[1:19],
+  deflator = c(1.542713570,
+               1.498048150,
+               1.458201390,
+               1.424814360,
+               1.387466100,
+               1.339831250,
+               1.309354560,
+               1.254768390,
+               1.232931730,
+               1.197970860,
+               1.159657520,
+               1.139569410,
+               1.112318840,
+               1.085316990,
+               1.069189690,
+               1.055708390,
+               1.035529570,
+               1.016107680,
+               1.000000000)) #%>% mutate(inflator = rev(deflator))
 
-thresholds <- read_csv("../../docs/Frequencies 180c/HILDA-thresholds-by-wave-180.csv") %>%
+thresholds <- read_csv("~/Dropbox/HILDA/data/HILDA-thresholds-by-wave-190.csv") %>%
   filter(variable %in% c("hifdip")) %>%
   gather("wave", "threshold", -variable, -label) %>%
   separate(wave, into = c("temp", "wave"), sep = "e") %>%
@@ -194,8 +195,8 @@ income <- gather_hilda(hilda, c(
[…]

### 2021-03-12 src/residual data.R

@@ -1,29 +1,0 @@
-#### create residual plot ####
-
-library(tidyverse)
-
-path_to_fits <- list.files(
-  path = 'results',
-  pattern = '^fit_cp_hifdip_gh9_cov_.*.RDS$',
-  full.names = TRUE
-)
-
-residual_df <- data.frame()
-
-for (path in path_to_fits) {
-  fit <- readRDS(path)
-  resids = resid(fit)[, "Estimate"]
-  fitteds = fitted(fit)[, "Estimate"]
-  df <- data.frame(resids, fitteds)
-  df$year = substring(path, 31, 34)
-  
-  residual_df <- bind_rows(residual_df, df)
-}
-
-write_rds(residual_df, "results/residual_data.RDS")
-
-resids = resid(pw_fit)[, "Estimate"]
-fitteds = fitted(pw_fit)[, "Estimate"]
-cor(fitteds, resids)
-hist(resids)
-ggplot() + geom_point(data = NULL, aes(y = resids, x = fitteds))

### 2021-03-12 src/draft0_word.Rmd
Submitted to PNAS

--- 2021-03-12 src/draft0.Rmd
@@
 title: "The increasing cost of happiness"
-author: "R.W. Morris^1,2^, N. Kettlewell^2,3,4^  & N.Glozier^1,5^"
+author: "R.W. Morris^1,2^, N. Kettlewell^2,3,4^  & N.Glozier^1,2^"
 bibliography: ../src/references.bib
@@
 1. Central Clinical School, Faculty of Medicine and Health, University of Sydney, NSW, Australia
-2. ARC Centre of Excellence for Children and Families over the Life Course, University of Sydney, NSW, Australia
+2. ARC Centre of Excellence for Children and Families over the Life Course
 3. School of Economics, University of Technology, NSW, Australia
 4. Institute of Labor Economics (IZA), Bonn, Germany
-5. Brain and Mind Centre, University of Sydney, NSW, Australia
 
@@
 | Draft      | `r format(Sys.time(), '%d %B, %Y')` |
-| Words      | 4399                                |
+| Words      | 4395                                |
 | Tables     | 0                                   |
@@
 A fundamental question for society is how much happiness does a dollar buy?
-The accepted view among economists and psychologists is that money and happiness increase together up to a point, after which there is little further gain from increasing wealth.
+The accepted view is that income and happiness increase together up to a point, after which there is little further gain from increasing wealth.
 While the location of this change point reportedly ranges between USD\$60K to $95K, there has been no investigation as to whether this has increased or decreased over time.
-We tested the temporal relationship between income and affective wellbeing (happiness), and income and cognitive wellbeing (life satisfaction), using household economic data from Australia between 2001-2019.
-We discovered the change point between happiness and income has increased over those 19 years faster than inflation (i.e., cost of living).
+We determined the change over time in the relationship between income and both affective wellbeing (happiness), and cognitive wellbeing (life satisfaction), using household economic data from Australia between 2001-2019.
+The change point in the slope between happiness and income has increased over those 19 years faster than the cost of living, but the association of life satisfaction with income has remained linear.
 This suggests that inequalities in income may be driving increasing inequities in happiness between the rich and the poor, with implications for health and recent government policy-goals to monitor and improve wellbeing.
@@
 Substantial research has shown that higher incomes have decreasing gains on wellbeing and happiness, across countries, cultures and ethnicities, with an inflection or change point at or near USD75,000.
-At incomes below this point happiness is dependent on financial security while at income levels above, happiness is relatively independent of income; revealing an inequality in the distribution of happiness between rich and poor.
+At incomes below this point happiness is more steeply dependent on income levels; revealing an inequality in the distribution of happiness between rich and poor.
 However we do not know whether or how this point has itself changed over time.
@@
 We examine 19 years of data from a single developed western nation (Australia) and determine that the income level at which happiness becomes less influenced by income has increased relative to rises in income and inflation.
-The happiness of an increasing proportion of the population is more dependent on financial security than before, whereas the association between income and life satisfaction has remained linear this millennium.
+The happiness of an increasing proportion of the population is more dependent on differences in financial security than before, whereas the association between income and life satisfaction has remained linear this millennium.
 
@@
 Inequities in the distribution of wellbeing are increasingly relevant to governments and policy-makers due to the growing recognition that increasing income does not necessarily lead to equal changes in wellbeing [@clark2018four; @frijters2020happy].
-Even prior to COVID-19, the World Gallup Poll has observed that happiness has been decreasing over the past decade in western Europe, north America, Australia and New Zealand, despite increases in income in the same countries [@sachs2019world].
+Even prior to COVID-19, the World Gallup Poll has observed that happiness has been decreasing over the past decade in western Europe, North America, Australia and New Zealand, despite increases in income in the same countries [@sachs2019world].
 However to date there has been no investigation of whether the relationship between income and happiness has changed over time which may contribute to these social trends.
-In particular has the change point between income and happiness, and therefore the distribution of happiness between rich and poor, become more or less equitable in the last few decades?
+In particular has the change point between income and happiness increased or decreased, and therefore the distribution of happiness between rich and poor, become more or less equitable in the last two decades?
 
@@
 
-The parameters of the piecewise model used Gaussian priors centred at zero, with a regularized parameterization for the slopes and intercept of Normal(0, 0.1) for each *&beta;*, and a slightly less skeptical regularization for the change point Normal(0, 0.5).
+The parameters of the piecewise model used Gaussian priors centred at zero, with a regularized parameterization for the slopes and intercept of Normal(0, 0.1) for each *&beta;* prior, and a slightly less restrictive regularization for the change point prior Normal(0, 0.5).
 These regularizing priors were selected to reduce overfitting while still allowing the model to learn the regular features of the sample, and so provide a more robust population estimate.
@@
 Cross-sectional population weights for Australia provided by the University of Melbourne for each year were included as a covariate to adjust for differences in the sample representativeness.
-Full-time students were removed, as well as individuals with an annual household disposable income that was indicated as _topcoded_ by the University of Melbourne.
+Full-time students were removed, as well as individuals with an annual household disposable income that was negative or indicated as _topcoded_ by the University of Melbourne.
 
@@
 For each wellbeing variable we show the results of a linear fit (rows 1 and 3) and a piecewise fit (rows 2 and 4).
-For visualization purposes only, due to the large numbers of individual data in each year, we display the mean levels of income and wellbeing for each (equal-sized) income decile rather than every individual data point, whereas the line-of-best-fit and 95% credible intervals (shaded) in each regression model are derived from _all individuals_.
+For visualization purposes only, we display the mean levels of income and wellbeing for each (equal-sized) income decile, whereas the line-of-best-fit and 95% credible intervals (shaded) in each regression model are derived from the underlying _individual_ data-points.
 
@@
 
-#### The evolution of the association between income and happiness 2001-2019
+#### Temporal trends in the association between income and happiness 2001-2019
 
@@
 
-##### Figure 2. Posterior parameters of the dependency of happiness on income (real 2019 dollars)   
+##### Figure 2. Posterior parameters of the happiness ~ income piecewise model (real 2019 dollars)   
 ```{r get_posteriors}
-income_summary <- read_rds("../results/gh9_hhoecd_data.RDS")
+income_params <- read_rds("../results/gh9_hhoecd_summary.RDS")
 fits <- read_rds("../results/gh9_hhoecd_0cov_tight.RDS")
@@
 posteriors <- map_dfr(fits, as.data.frame, .id = "fit") %>%
-  left_join(bind_rows(income_summary, .id = "fit")) %>%
+  left_join(bind_rows(income_params, .id = "fit")) %>%
   mutate(omega_dollars = (b_omega_Intercept * sd + mean)/1000) %>%
@@
 
-The change in parameter values between 2001 to 2019 indicates the relationship between happiness and income evolved over time.
+The change in parameter values between 2001 to 2019 indicates the relationship between happiness and income has evolved over time.
 We determined the impact of this evolution on the distribution of happiness over the range of household income in 2019 in a counterfactual analysis.
@@
 This controls for changes in the sample which occur over time that are not related to happiness but could nevertheless contribute to changes in the distribution of happiness.
-For example, an increase in the range of (real) income levels in the economy between 2001 and 2019 could produce a greater divergence in happiness due to the increasing gap between the rich and poor - even with a stable relationship between income and happiness.
+For example, an increase in the range of (real) income levels in the economy between 2001 and 2019 could produce an increasing gap in happiness between the rich and poor - even with a stable relationship between income and happiness.
 Such changes in the sample characteristics may mask or confound the impact of the change point on the distribution of happiness without careful control.
@@
 
-##### Figure 3. Counterfactual difference in happiness between 2019 and 2001
+##### Figure 3. Counterfactual difference in happiness between 2019 and 2001 for a given level of income
 ```{r counterfactual_estimates}
@@
       dollars = as.vector(shh_disp_inc_eq),
+      male = male,
+      age = as.vector(scale(age)),
+      age_sq = age^2,
+      grad = edu %in% c("Grad", "Postgrad"),
+      weights = as.vector(scale(weights))
     ) %>% 
@@
   geom_point(aes(y = delta), size = 0.1, alpha = 0.1) +
-  # geom_density(fill = "grey50", color = "grey50") +
-  stat_smooth(aes(y = delta), geom = "line", method = "loess", se = F, span = 5) +
+  stat_smooth(aes(y = delta), geom = "line", method = "loess", se = F, span = .5) +
   xlim(0, 200) +
   labs(
-    subtitle = "Happiness delta(s) from 2019 predictions less 2001 predictions (SD units)",
+    subtitle = "Happiness delta (SD units)",
     x = "Household income ($000s)", y = "",
@@
 ```
-Figure 3 legend: The difference (∆) in model-estimated happiness between 2019 and 2001 for the same n = 14,459 individuals from the 2019 survey.
-Values below zero on the y-axis indicate lower happiness estimates in 2019 relative to 2001.
-The smoothed overlay (solid line) indicates how the average happiness changes across the income distribution.
+Figure 3 legend: The difference (∆) in predicted happiness between 2019 and 2001 for the same n = 14,459 individuals from the 2019 survey.
+Values below zero on the y-axis indicate lower happiness predictions using the 2019 model compared to using the 2001 model.
+Solid line overlay represents the (loess) smoothed average delta over the income distribution.
+Variance in these posterior samples is from the sample weights, uncertainty in the model parameters as well as observation-level (residual) variance.
 ```
@@
 
-Figure 3 shows that, on average, happiness estimates improved for people with household incomes above \$55K in 2019 when compared to 2001, as indicated by the average delta (solid line) falling above zero on the right side of the plot.
-Conversely, people with household incomes below \$55K, on average, suffered a decrease in estimated happiness in 2019 compared to 2001, particularly those whose income levels were transliminal: that is, their income was above the changepoint in 2001 but the shift in the change point resulted in their income falling below the change point in 2019 (i.e., ~$50K, see Figure S6 in Supplementary materials).
+Figure 3 shows that, on average, predicted happiness is higher for people with household incomes above \$50K under the 2019 function, when compared to the function from 2001.
+This is indicated by the average delta (solid line) falling above zero on the right side of the plot.
+Conversely, people with household incomes below \$50K, on average, had a decrease in predicted happiness under the 2019 function compared to 2001 function.
 Of course the obtained deltas are due to changes in all the parameters of the function, including the slope before and after the change point.
@@
 
-Assuming median income is also not increasing at a faster rate, any increase in the change point will reduce the number of people who fall above it over time; an implication being that over time more people's happiness is more dependent on their income level than previously.
+Any increase in the change point is likely to reduce the number of people who fall above it over time if this is greater than any increase in median income; i.e., over time a larger proportion of the population's happiness is responsive to marginal changes in their income than previously.
 Figure 4 presents median household income levels weighted for the Australian population (by age, sex, marital status, labour force participation and geographical region).
@@
 Figure 4 legend: Real household income has stagnated in Australia since 2009 (post GFC) while the change point between happiness and income has increased.  
-Consequently fewer Australians had an income that affords them happiness, which is independent of their financial security in 2019.
 ```
@@
 
-We confirmed previous findings [e.g., @howell2008relation; @kahneman2010high], that the relationship between both types of subjective wellbeing with household income was positive but quite different: Satisfaction increased linearly with income, while happiness increased rapidly up to a point after which higher levels of income were associated with less improvement.
+This analysis of the Australian population since the millennium confirms previous findings [e.g., @howell2008relation; @kahneman2010high], that the relationship between the two types of subjective wellbeing and household income was positive but quite different: Satisfaction increased linearly with income, while happiness increased relatively steeply up to a point after which higher levels of income were associated with much more modest differences.
+We refer to the change point after which increases in income correspond to smaller increases in happiness as the _cost of happiness_.
+After this point, happiness is no longer as responsive to differences in household income, and the economic security it represents.
+Presumably after this point further increases in happiness depend more on other life factors (e.g., leisure time, social connections) than financial security.
 Our novel finding was that this change point in the relationship between household income and happiness increased faster than both inflation and median household income between 2001 and 2019.
-For the first time we have shown there has been a temporal shift in the change point between income and happiness over the 19 year period, such that happiness has become more dependent on income for more people - especially the poor and middle-class.
+For the first time we have shown there has been a temporal shift in the change point between income and happiness over the 19 year period, such that happiness of an increasing proportion of the population has become more responsive to differences in income - especially the poor and middle-income earners.
 
-We refer to the change point after which increases in income no longer produce similar increases in happiness as the _cost of happiness_.
-After this point, happiness is no longer as dependent on household income, and the economic security it represents.
-Presumably after this point further increases in happiness depend more on other life factors (e.g., leisure time, social connections) than financial security.
 Life satisfaction on the other hand appeared to show consistent increases with household income and we found no evidence of any change point.
@@
 An implication of the changing relationship between happiness and income is that income inequality may be driving increasing inequities in wellbeing.
-This is demonstrated in Figure 3, where the difference in happiness between 2001 and 2019 increased for incomes above $50K/year and decreased for incomes below that level (see also Figure S6 in Supplementary materials).
+This is demonstrated in Figure 3, where the difference in predicted happiness between 2001 and 2019 increased for incomes above $50K/year and decreased for incomes below that level (see also Figure S6 in Supplementary materials).
 The inequity was also highlighted in Figure 4 where the change point of happiness represented a 9% increase over median income in 2001, while in 2019 it represented a 42% increase over median income.
 This increase relative to median income also represented a reduction from 43% to 26% in the proportion of people whose income fell above the changepoint.
-Thus we can see that over the last nineteen years the difference in happiness between the rich and the poor has increased; while the happiness of an increasing proportion of people, including the middle-class, is more dependent on their financial security.
+Thus we can see that over the last nineteen years the difference in income-related happiness between the rich and the poor has increased; while the happiness of an increasing proportion of people, including the middle-class, is more dependent on their financial security.
 
-Australia has low levels of income disparity relative to many other OECD countries, and the Gini coefficient has not changed a great deal between 2001 and 2019 [@australian2018rising].
-A stable Gini coefficient shows income inequality has remained steady over the time period, and our results do not conflict with this conclusion.
+Australia has low levels of income disparity relative to many other OECD countries, and the Gini coefficient has changed little between 2001 and 2019 [@australian2018rising], suggesting income inequality has remained steady over this time period.
+Our results do not conflict with this conclusion.
 Rather we are revealing that even a static income distribution may have dynamic effects on happiness over time.
-We think this highlights the issue that while traditional univariate measures of wealth and income inequality may be relatively stable and exhibit little change, their impact on wellbeing and health can still vary.
-As such, these results may well have relevance to other developed nations in North America and Europe which have also enjoyed stable economic growth overall, but have stagnating incomes, and declining happiness levels [@sachs2019world].
+This highlights the issue that while traditional univariate measures of wealth and income inequality may be relatively stable and exhibit little change, their impact on wellbeing and health can still vary.
+As such, these results may well have relevance to other developed nations in North America and Europe which also enjoy low levels of income inequality, but have stagnating incomes, and declining happiness levels [@sachs2019world].
 As focus shifts from traditional financial indicators towards health and wellbeing measures, findings such as this may become more prevalent.
@@
 The relationship between happiness and income is often described as log-linear [e.g., @kahneman2010high; @killingsworth2021experienced].
-Under this view, similar proportional changes in income or wealth produce similar evaluations of satisfaction or happiness, rather than the absolute change.
-Likewise, the piecewise model reported here has a slope that reduces with wealth, albeit with a single identifiable change-point rather than a smooth curve (*nb*., in the Bayesian setting, the slope is smooth over the posterior of the change-point).
+Under this view, similar proportional changes in income produce similar evaluations of satisfaction or happiness, rather than the absolute change.
+Likewise, the piecewise model reported here has a slope that reduces with income, albeit with a single identifiable change-point rather than a smooth curve.
 However our main thesis is not committed to a function with a single or smooth changepoint; only that the slope between income and happiness is unequal at different levels of income, and the function is shifting to the right over time (e.g., Figure 2).
 The choice between functions is somewhat arbitrary, since a log-linear function is as easily consistent with our thesis as the piecewise function we describe.
-Nevertheless, the model evidence favoured a piecewise- over a log-linear function in a supplementary analysis (Figure S3), which validates our otherwise arbitrary model choice here.
+Nevertheless, the model evidence favoured a piecewise- over a log-linear function in a supplementary analysis (Figure S3), which supports our model choice here.
 
-Some recent studies have challenged the notion that the positive effect of income plateaus after some level ("satiety") [@killingsworth2021experienced; @twenge2020expanding; stevenson2013subjective].
+Some recent studies have challenged the notion that the positive effect of income plateaus after some level ("satiety") [@killingsworth2021experienced; @twenge2020expanding; @stevenson2013subjective].
 Our results also do not support a "satiety" point or flat gradient between income and happiness, as the average post-change slope between 2001-2019 was reliably above zero.
 The pre-change slope however was also four to five times greater than that, and we argue this difference in slopes indicates a previously unacknowledged inequity in the distribution of happiness between the rich and poor.
-While the amount of variance in happiness explained or predicted by household income may be small, the slope is reliably above zero and small changes in happiness over an entire economy of 25M people or more will aggregate.
-Moreover, happiness is measured on a subjective scale, where small changes at different points on the scale may represent a large change in some other functional/real-world outcome (e.g., risk of suicide).
-This is all to say that quantifying the real-world impact of small changes in a subjective measure such as happiness is difficult and speculative without further investigation.
-Our aim here is to merely determine whether the inequity is present at all.
+
+While the amount of variance in happiness explained by household income is small (Table S4, Supplementary materials), the slope is reliably above zero and small changes in happiness over an entire population will aggregate.
+Moreover, financial gains have one of the largest impacts on individual happiness in within-subject models (e.g., fixed-effects), second only to marriage by our estimate [@kettlewell2020differential], and there are many reasons why cross-sectional estimates of the covariance would under-represent this.
+Finally, happiness is measured on a subjective scale, where small changes at different points on the scale may represent a large change in some other functional/real-world outcome (e.g., risk of suicide).
+Ultimately, the real-world impact of small changes in a subjective measure such as happiness is inherently difficult to determine.
+Our aim here was to determine whether the relationship between income and happiness is changing at all, and whether this may be contributing to increasing inequity, and our results indicate this is the case.
 
@@
 According to @frijters2020happy, coming up with a consensus to translate income into wellbeing features high on the wider wellbeing research agenda.
-Establishing the links between wealth, household income, wellbeing and health, and how inequalities in one drives inequities in the other, will be a critical step in the success of that agenda.
+Establishing the links between wealth, household income, wellbeing and health, and how inequalities in one drives inequities in the other, will be a critical step in that agenda.
 
@@
 
-RWM analysed the data and wrote the paper. NK designed the analysis and wrote the paper. NG proposed the idea, planned the analysis, and wrote the paper.
+RWM designed the analysis, analysed the data and wrote the paper. NK designed the analysis and wrote the paper. NG proposed the idea, planned the analysis, and wrote the paper.
 
@@
 
-Supplementary information is available [online](https://datarichard.github.io/The-increasing-cost-of-happiness/supplementary.html){target="_blank"}.
+Supplementary information is available [online](https://datarichard.github.io/The-increasing-cost-of-happiness/supplementary.html.
 

### 2021-03-12 src/supplementary0.Rmd
Submitted to PNAS

@@
 title: "Supplementary material: The increasing cost of happiness"
-author: "R.W. Morris^1,2^, N. Kettlewell^2,3,4^  & N.Glozier^1,5^"
+author: "R.W. Morris^1,2^, N. Kettlewell^2,3,4^  & N.Glozier^1,2^"
 bibliography: ../src/references.bib
@@
 4. Institute of Labor Economics (IZA), Bonn, Germany
-5. Brain and Mind Centre, University of Sydney, NSW, Australia
 
@@
 ```{r figure_s1, dpi=300}
-here <- "~/Dropbox/HILDA/The-increasing-cost-of-happiness/"
+lm_fits = read_rds("../results/lm_gh9_hhoecd_0cov.RDS")
 
-lm_fits = read_rds(paste0(here, "results/lm_gh9_hhoecd_0cov.RDS"))
-
-nl_fits = read_rds(paste0(here, "results/gh9_hhoecd_0cov_tight.RDS"))
+nl_fits = read_rds("../results/gh9_hhoecd_0cov_tight.RDS")
 
@@
 
-nl_fits_losat = read_rds(paste0(here, "results/losat_hhoecd_0cov.RDS"))
+nl_fits_losat = read_rds("../results/losat_hhoecd_0cov.RDS")
 
-lm_fits_losat = read_rds(paste0(here, "results/lm_losat_hhoecd_0cov.RDS"))
+lm_fits_losat = read_rds("../results/lm_losat_hhoecd_0cov.RDS")
 
@@
 ##### Figure S2. Conditional model parameters  
-```{r figure_s2, fig.height=8, dpi=300}
-here <- "~/Dropbox/HILDA/The-increasing-cost-of-happiness/"
+```{r figure_s2, fig.width = 6, fig.height = 7.5, dpi=300}
+nl_fits = read_rds("../results/gh9_hhoecd_0cov_tight.RDS")
 
-nl_fits = read_rds(paste0(here, "results/gh9_hhoecd_0cov_tight.RDS"))
+nl_fits_5cov = read_rds("../results/gh9_hhoecd_5cov_tight.RDS")
 
-nl_fits_5cov = read_rds(paste0(here, "results/gh9_hhoecd_5cov_tight.RDS"))
-
-income_params <- read_rds("../results/gh9_hhoecd_summary.RDS")
+income_params <- read_rds("../results/gh9_hhoecd_summary.RDS") %>%
+  bind_rows(.id = "fit")
 
 posteriors_5c <- map_dfr(nl_fits_5cov, as.data.frame, .id = "fit") %>%
-  left_join(bind_rows(income_params, .id = "fit")) %>%
+  left_join(income_params) %>%
   mutate(omega_dollars = (b_omega_Intercept * sd + mean)/1000) %>%
@@
 posteriors_0c <- map_dfr(nl_fits, as.data.frame, .id = "fit") %>%
-  left_join(bind_rows(income_params, .id = "fit")) %>%
+  left_join(income_params) %>%
   mutate(omega_dollars = (b_omega_Intercept * sd + mean)/1000) %>%
@@
 ```{r, figure_s3, fig.height=3.5, dpi=300}
-here <- "~/Dropbox/HILDA/The-increasing-cost-of-happiness/"
+nl_fits = read_rds("../results/gh9_hhoecd_0cov_tight_x0.RDS")
 
-nl_fits = read_rds(paste0(here, "results/gh9_hhoecd_0cov_tight_x0.RDS"))
-
-log_fits = read_rds(paste0(here, "results/log_gh9_hhoecd_0cov.RDS"))
+log_fits = read_rds("../results/log_gh9_hhoecd_0cov.RDS")
 
@@
 ```
-Figure S3 legend: The joint distribution between wellbeing and income for a single year, imposed by the priors.
+Figure S3 legend: Differences in WAIC scores (deviance units) for a log-linear fit over a piecewise fit for happiness.
 ```
@@
 ```{r figure_s4, dpi=300}
-load(paste0(here, "results/prior_predictions.Rdata"))
+load("../results/prior_predictions.Rdata")
 
@@
 
-load(paste0(here, "results/residual_fits.Rdata"))
+load("../results/residual_fits.Rdata")
 
@@
 ```{r table_s2, ft.align="left"}
+nl_fits = read_rds("../results/gh9_hhoecd_0cov_tight.RDS")
+
+income_params %>%
+  mutate(fit = str_remove(fit, "fit")) %>%
+  column_to_rownames("fit") -> income
+
 param_key = c(
@@
   b2_Intercept = "Post-slope",
-  omega_Intercept = "Change-point"
+  omega_Intercept = "Change-point ($000s)"
 )
@@
         .f = ~{
-          fit_summary <- summary(.x)
+          fit_summary <- summary(.x, prob = 0.95)
 
@@
     Year = str_remove(Year, "fit"),
-    Parameter = recode_factor(Parameter, !!!param_key)
+    Parameter = recode_factor(Parameter, !!!param_key),
+    Estimate = if_else(str_detect(Parameter, "Change"),
+                       (Estimate * income[Year, "sd"] + income[Year, "mean"])/1000,
+                       Estimate),
+    Est.Error = if_else(str_detect(Parameter, "Change"),
+                        (Est.Error * income[Year, "sd"] + income[Year, "mean"])/1000,
+                        Est.Error),
+    `l-95% CI` = if_else(str_detect(Parameter, "Change"),
+                        (`l-95% CI` * income[Year, "sd"] + income[Year, "mean"])/1000,
+                        `l-95% CI`),
+    `u-95% CI` = if_else(str_detect(Parameter, "Change"),
+                         (`u-95% CI` * income[Year, "sd"] + income[Year, "mean"])/1000,
+                         `u-95% CI`)
     ) %>%
@@
      part = "body") %>%
-  colformat_num(j = 3:6, digits= 3) %>%
-  colformat_num(j = 7, digits= 1) %>%
+
+  colformat_double(j = 3:6, digits= 3) %>%
+  colformat_double(j = 7, digits= 1) %>%
   set_caption("Table S2. Piecewise parameter estimates and fit statistics for happiness") %>%
-  autofit()
+  fontsize(size = 9.5, part = "all") %>%
+  width(j = 1, width = .5) %>%
+  width(j = 2, width = 1) %>%
+  width(j = c(3:7), width = .7) %>%
+  width(j = c(8, 9), width = .8)
 ```
@@
      part = "body") %>%
-  colformat_num(j = 3:6, digits= 3) %>%
-  colformat_num(j = 7, digits= 1) %>%
+  colformat_double(j = 3:6, digits= 3) %>%
+  colformat_double(j = 7, digits= 1) %>%
   set_caption("Table S3. Linear parameter estimates and fit statistics for life satisfaction") %>%
+  fontsize(size = 9.5, part = "all") %>%
   autofit()
@@
 ```{r figure_s6, fig.width=7, fig.height=5, dpi=300}
-here <- "~/Dropbox/HILDA/The-increasing-cost-of-happiness/"
-
-nl_fits = read_rds(paste0(here, "results/gh9_hhoecd_0cov_tight.RDS"))
+nl_fits = read_rds("../results/gh9_hhoecd_0cov_tight.RDS")
 
@@
 
-income_params <- read_rds("../results/gh9_hhoecd_summary.RDS")
-
-bind_rows(income_params, .id = "fit") %>%
+fixed_income_hash <- read_rds("../results/gh9_hhoecd_summary.RDS") %>%
+  bind_rows(.id = "fit") %>%
   mutate(low = (40000 - mean)/sd,
@@
   column_to_rownames(var = "fit") %>%
-  as.matrix() -> fixed_income_hash
+  as.matrix()
 
@@
 model_summaries %>%
-  select(model, year, r.squared) %>%
+  select(model, Year = year, r.squared) %>%
   spread(model, r.squared) %>%
   flextable() %>%
-  colformat_num(j = 2:7, digits = 3) %>%
+  colformat_num(j = 1, big.mark = "") %>%
+  colformat_double(j = 2:7, digits = 3) %>%
   set_caption("Table S4. In-sample fit statistics (R-squared)") %>%
@@
   valign(i = 1, valign = "top", part = "header") %>%
-  width(j = 2:7, width = 1)
+  fontsize(size = 9.5, part = "all") %>%
+  width(j = 2:7, width = 1) %>%
+  width(j = 1, width = .5)
 ```
@@
 
-Three things need to be observed when interpreting the R^2^ values: 1) While the R^2^ between income and happiness is low (~2% and ~14% in the unconditional and conditional models, respectively), the aggregate effect over the wider population (e.g., 25M in Australia) will be much larger; 2) Happiness (and to a lesser extent, income) is an imperfect and noisy measure so any effect of income on happiness is likely to be larger than that measured here; 3) Finally, happiness is measured on a subjective and arbitrary Likert scale so a small change in happiness at different points in the scale may represent a large change in some other functional/real-world outcome (e.g., risk of suicide).
+Three things need to be observed when interpreting the R^2^ values: 1) While the R^2^ between income and happiness is low (~2% and ~14% in the unconditional and conditional models, respectively), the effect over the wider population will aggregate to be much larger; 2) Happiness (and to a lesser extent, income) is an imperfect and noisy measure so any effect of income on happiness is likely to be larger than that measured here; 3) Finally, happiness is measured on a subjective and arbitrary Likert scale so a small change in happiness at different points in the scale may represent a large change in some other functional/real-world outcome (e.g., risk of suicide).
 This is all to say that quantifying the real-world impact of small changes in variation reported here is difficult and speculative without further investigation.

### 2021-03-12 src/supplementary0_word.Rmd
Submitted to PNAS

--- 2021-03-12 src/supplementary0.Rmd
@@
 1. Central Clinical School, Faculty of Medicine and Health, University of Sydney, NSW, Australia
-2. ARC Centre of Excellence for Children and Families over the Life Course, University of Queensland, QLD, Australia
+2. ARC Centre of Excellence for Children and Families over the Life Course
 3. School of Economics, University of Technology, NSW, Australia
 4. Institute of Labor Economics (IZA), Bonn, Germany
-5. Brain and Mind Centre, University of Sydney, NSW, Australia
 

### 2021-05-27 src/draft-PNAS0.Rmd
SSM0 Population Health submission

--- 2021-03-12 src/draft0_word.Rmd
@@
 The accepted view is that income and happiness increase together up to a point, after which there is little further gain from increasing wealth.
-While the location of this change point reportedly ranges between USD\$60K to $95K, there has been no investigation as to whether this has increased or decreased over time.
+While the location of this change point reportedly ranges between USD\$60K to \$95K, there has been no investigation as to whether this has increased or decreased over time.
 We determined the change over time in the relationship between income and both affective wellbeing (happiness), and cognitive wellbeing (life satisfaction), using household economic data from Australia between 2001-2019.
 The change point in the slope between happiness and income has increased over those 19 years faster than the cost of living, but the association of life satisfaction with income has remained linear.
+As a result, the happiness of a larger proportion of the population is more dependent on differences in income, despite stable differences in income distribution (e.g., Gini coefficient) in that time.
 This suggests that inequalities in income may be driving increasing inequities in happiness between the rich and the poor, with implications for health and recent government policy-goals to monitor and improve wellbeing.
-(149 words)
 
@@
 ## Significance Statement
-Substantial research has shown that higher incomes have decreasing gains on wellbeing and happiness, across countries, cultures and ethnicities, with an inflection or change point at or near USD75,000.
-At incomes below this point happiness is more steeply dependent on income levels; revealing an inequality in the distribution of happiness between rich and poor.
-However we do not know whether or how this point has itself changed over time.
-Has it decreased and reduced inequality, or increased the difference between the rich and poor?
-We examine 19 years of data from a single developed western nation (Australia) and determine that the income level at which happiness becomes less influenced by income has increased relative to rises in income and inflation.
-The happiness of an increasing proportion of the population is more dependent on differences in financial security than before, whereas the association between income and life satisfaction has remained linear this millennium.
+Income has decreasing returns on happiness, across countries, cultures and ethnicities, with an inflection or change point near USD75,000.
+Below this point, happiness is relatively dependent on income levels; revealing an inequity in the distribution of happiness between rich and poor.
+We do not know whether or how this point changes over time.
+Has it decreased and reduced inequity, or increased the difference between the rich and poor?
+We examine 19 years of data from Australia and determine that the income level at which happiness becomes less influenced by income has increased, relative to the cost of living.
+The happiness of an increasing proportion of the population is more dependent on their financial security than ever this millennium.
+(120 max)
 

### 2021-05-27 src/draft-SSM0.Rmd
SSM0 Population Health submission

--- 2021-05-27 src/draft-PNAS0.Rmd
@@
 | Draft      | `r format(Sys.time(), '%d %B, %Y')` |
-| Words      | 4395                                |
+| Words      | 4084                                |
 | Tables     | 0                                   |
@@
 
-**keywords:** Subjective wellbeing, household income, HILDA
+Declarations of interest: none.
+
+Funding sources: This research did not receive any specific grant from funding agencies in the public, commercial, or not-for-profit sectors
 
@@
 A fundamental question for society is how much happiness does a dollar buy?
-The accepted view is that income and happiness increase together up to a point, after which there is little further gain from increasing wealth.
+The accepted view among economists and psychologists is that money and happiness increase together up to a point, after which there is little further gain from increasing wealth.
 While the location of this change point reportedly ranges between USD\$60K to \$95K, there has been no investigation as to whether this has increased or decreased over time.
-We determined the change over time in the relationship between income and both affective wellbeing (happiness), and cognitive wellbeing (life satisfaction), using household economic data from Australia between 2001-2019.
-The change point in the slope between happiness and income has increased over those 19 years faster than the cost of living, but the association of life satisfaction with income has remained linear.
-As a result, the happiness of a larger proportion of the population is more dependent on differences in income, despite stable differences in income distribution (e.g., Gini coefficient) in that time.
+We tested the temporal relationship between income and affective wellbeing (happiness), and income and cognitive wellbeing (life satisfaction), using household economic data from Australia between 2001-2019.
+We discovered the change point between happiness and income has increased over those 19 years faster than inflation (i.e., cost of living).
 This suggests that inequalities in income may be driving increasing inequities in happiness between the rich and the poor, with implications for health and recent government policy-goals to monitor and improve wellbeing.
+
+<br><br>
+
+**keywords:** Subjective wellbeing, household income, HILDA
 
@@
 
-<br><br>
+## Highlights Statement
 
-## Significance Statement
-Income has decreasing returns on happiness, across countries, cultures and ethnicities, with an inflection or change point near USD75,000.
-Below this point, happiness is relatively dependent on income levels; revealing an inequity in the distribution of happiness between rich and poor.
-We do not know whether or how this point changes over time.
-Has it decreased and reduced inequity, or increased the difference between the rich and poor?
-We examine 19 years of data from Australia and determine that the income level at which happiness becomes less influenced by income has increased, relative to the cost of living.
-The happiness of an increasing proportion of the population is more dependent on their financial security than ever this millennium.
-(120 max)
+- We investigate how the relationship between income and subjective wellbeing (affective and cognitive) has changed over time, using a dataset with highly detailed measurements of each
+- Nineteen years of data from a well-developed, economically stable G20 nation is examined (Australia), allowing a relatively undisturbed view of the changes without major economic shocks
+- We are able to show for the first time that the income level at which happiness is less determined by income has risen since 2001, resulting in greater happiness inequality between the rich and the poor
+- Conversely the linear relationship between income and life-satisfaction has remained constant over this period
+- The happiness of an increasing proportion of the population is more dependent on their financial security than ever this millennium
 
@@
 
-## Background
-A fundamental question for governments and people is just how much wellbeing does a dollar buy?
+## 1.0 Introduction
+A fundamental question for society is just how much wellbeing does a dollar buy?
 Increasing income is commonly associated with increasing happiness and subjective wellbeing, however a point at which subjective wellbeing no longer increases with income has also been widely observed [@clark2008relative; @dolan2008we; @easterlin1974growth].
-Given that a central goal of nations and governments is to improve income under the assumption that higher income always increases wellbeing, challenges to this notion have far reaching consequences [@frijters2020happy; @laws2014world].
 
-Subjective wellbeing is not a unitary entity [@diener2017findings]; studies typically distinguish between life satisfaction, the cognitive appraisal of one’s own accomplishments, and affective wellbeing, one's prevailing affective state, emotional mood, or everyday experience of happiness.
-Money can have different effects on each.
-For instance, we have recently reported that large increases in wealth, such as a major financial windfall, have a greater impact on an individual’s life satisfaction than their happiness [@kettlewell2020differential].
-The distinct effect of money on satisfaction and happiness was observed within individuals over time, however a distinct effect of income has also been observed between people with different income levels.
-For instance, Kahneman & Deaton [-@kahneman2010high] showed that self-reported levels of happiness increased with household income up to a point (USD75,000), but after that, increasing income had little further effect on happiness.
+Kahneman & Deaton [-@kahneman2010high] showed that self-reported happiness increased with household income up to a point (USD75,000), but after that, increasing income had little further effect.
 Conversely life satisfaction continued to increase with income beyond USD75,000.
+Subjective wellbeing is not a unitary entity [@diener2017findings]; studies typically distinguish between life satisfaction, the cognitive appraisal of one’s own accomplishments, and affective wellbeing, one’s prevailing affective state, emotional mood, or everyday experience of happiness [e.g., @kettlewell2020differential].
 Indeed, the difference between the two questions: “How satisfied are you with your life?” and “How happy are you these days?” has been identified as a crucial mediating factor in a meta-analysis of 111 studies on income and wellbeing [@howell2008relation; also @veenhoven2006rising].
-Larger and newer data sets now also suggest that higher incomes are associated with more happiness and life satisfaction both within and across nations [@sacks2012new; @jebb2018happiness; @killingsworth2021experienced].
-Results such as these have provided a more nuanced view of the relationship between income and wellbeing; namely that while both happiness and satisfaction continue to increase with income, the happiness returns are decreasing [@diener2010wealth; @haushofer2014psychology].
+Results since then have also shown that while both happiness and satisfaction increase with income, the happiness returns are decreasing [@killingsworth2021experienced; @sacks2012new].
 
 Fundamentally, the existence of a change point in the relationship between income and happiness reveals an unacknowledged source of inequality in the distribution of wellbeing (i.e., happiness) in the economy.
-For instance, the change point of USD75,000 reported by Kahneman in 2008 was substantially more than the US median income of USD52,000 in the same year, indicating that the happiness of the poorest majority of the US population was tied to marginal changes in income while the happiness of a richer minority was not.
+For instance, a change point of USD75,000 in 2008 was substantially more than the US median income of USD52,000 in the same year, indicating that the happiness of the poorest majority of the US population was tied to marginal changes in income while the happiness of a richer minority was not.
 Thus the change point represents the dollar value up to which income drives inequities in the distribution of happiness, such that a lower valued change point represents a more equitable distribution of happiness in the economy.
 Inequities in the distribution of wellbeing are increasingly relevant to governments and policy-makers due to the growing recognition that increasing income does not necessarily lead to equal changes in wellbeing [@clark2018four; @frijters2020happy].
-Even prior to COVID-19, the World Gallup Poll has observed that happiness has been decreasing over the past decade in western Europe, North America, Australia and New Zealand, despite increases in income in the same countries [@sachs2019world].
-However to date there has been no investigation of whether the relationship between income and happiness has changed over time which may contribute to these social trends.
-In particular has the change point between income and happiness increased or decreased, and therefore the distribution of happiness between rich and poor, become more or less equitable in the last two decades?
+Even prior to COVID-19, the World Gallup Poll has observed that happiness has decreased over the past decade in western Europe, North America, Australia and New Zealand, despite increases in income in the same countries [@sachs2019world].
+However to date there has been no investigation of whether the relationship between income and happiness has changed over time which may contribute to these trends.
+In particular has the change point between income and happiness, and therefore the distribution of happiness between rich and poor, become more or less equitable in the last few decades?
 
@@
 
-## Methods
+## 2.0 Methods
 
 We used household economic panel data from Australia to provide the first investigation of whether changes in income and wellbeing have shifted the change point between 2001 to 2019.
-Australia is a developed western nation that has experienced a relatively stable period of economic growth in the last 19 years, which makes it a good candidate to observe how the function between income and subjective wellbeing has evolved over time without major economic shocks.
-HILDA (the Household, Income, and Labour Dynamics in Australia survey) provides a representative sample of households in Australia with detailed measurements of income and subjective wellbeing each year, which makes it an excellent data source to investigate the present question.
+HILDA (the Household, Income, and Labour Dynamics in Australia survey) provides a representative sample of households in Australia with detailed measurements of income and subjective wellbeing each year.
 We distinguished between satisfaction and happiness as different components of subjective wellbeing, and evaluated how each varied with household income over time using piecewise linear regression to determine the change points in each year.
@@
 
-#### Income
-To allow comparison with other major studies of income and wellbeing we used household after-tax income as the indicator of income and economic security [e..g, @kahneman2010high].
-Household income better represents economic security than personal income, since members of the same household share expenses as well as risks; i.e., they can provide a direct and immediate support network when financial shocks occur, which a priori might affect wellbeing.
+#### 2.1 Income
+We used household after-tax income as the indicator of income and economic security [e..g, @kahneman2010high].
 The ‘real household annual disposable income’ was calculated from the self-reported combined income of all household members after receipt of government pensions and benefits and deduction of income taxes in the financial year ended 30th June of the year of the wave (e.g., 2001 in wave 1).
@@
 
-The _equivalised_ household income was obtained by adjusting for household size (the number of adult and child household members).
-We used the 'modified OECD' scale [@hagenaars1994poverty], which divides household income by 1 for the first household member plus 0.5 for each other household member aged 15 or over, plus 0.3 for each child under 15.
-A family comprising two adults and two children under 15 years of age would therefore have an equivalence scale of 2.1 (1 + 0.5 + 0.3 + 0.3), meaning that the family would need to have an income 2.1 times that of a single-person household in order to achieve the same standard of living.
-This scale recognises that larger households require more income, but it also recognises that there are economies of scale in consumption and that children require less than adults.
-The equivalised income calculated for a household is then assigned to each member of the household.
+The _equivalised_ household income was obtained by adjusting for household size (the number of adult and child household members) using the 'modified OECD' scale [@hagenaars1994poverty].
+Household income was divided by 1 for the first household member plus 0.5 for each other household member aged 15 or over, plus 0.3 for each child under 15.
+The equivalised income calculated for a household was then assigned to each member of the household.
 
@@
 
-#### Subjective Wellbeing
-There are a variety of variables related to subjective well-being collected annually in HILDA, but the two we used here matched the variables we used in our previous paper [@kettlewell2020differential], namely, life satisfaction as a measure of cognitive wellbeing, and item 9 from the SF-36 (9a-9i) as a measure of affective wellbeing or happiness.
+#### 2.2 Subjective Wellbeing
+Life satisfaction was assessed by a single item question asked each survey: "All things considered, how satisfied are you with your life (0 to 10)".
 
-Life satisfaction [(losat)](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=losat){target="_blank"} was assessed by a single item question asked each survey: "All things considered, how satisfied are you with your life (0 to 10)".
-
-Happiness was determined by 9 questions in the SF-36 [(9a to 9i)](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=gh9a){target="_blank"}.
+Happiness was determined by 9 questions in the SF-36 [(9a to 9i)](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=gh9a).
 The SF-36 is a widely used self-completion measure of various aspects of physical, emotional and mental health [@ware2000sf].
@@
 
-#### Modelling
+#### 2.3 Modelling
 
-We modelled the relationship between income and each wellbeing variable (happiness and satisfaction) using a simple linear model and a piecewise model with a single change-point.
+We modelled the relationship between income and each wellbeing variable (happiness and satisfaction) using a simple linear model and a piecewise model with a single change-point as a free parameter estimated from the data.
 The piecewise model was chosen as the simplest extension of a linear model which can identify a change point (inflection) in the relationship between wellbeing and income.
-The location of the change point was a free parameter which revealed where wellbeing no longer increased at a uniform rate with income.
-We then compared the linear model against the piecewise model to determine if a change point existed in any year between household income and each wellbeing variable (see Model Selection).
-Finally, where a change point existed, we determined the location of the change point for that year (see Parameter Estimation).
-
 For modelling, both measures of wellbeing and income were rescaled with a mean of zero and a SD of 1 (z-scores) for each year.
@@
 
-The piecewise model was a simple extension of this to include a free parameter to represent the changepoint in income ($\omega$) as well as the slope before the change point ($\beta_1$) and the slope after the change point ($\beta_2$):
+The piecewise model included a free parameter to represent the changepoint in income ($\omega$) as well as the slope before the change point ($\beta_1$) and the slope after the change point ($\beta_2$):
 
@@
 
-The parameters of the piecewise model used Gaussian priors centred at zero, with a regularized parameterization for the slopes and intercept of Normal(0, 0.1) for each *&beta;* prior, and a slightly less restrictive regularization for the change point prior Normal(0, 0.5).
+The parameters of the piecewise model used Gaussian priors centred at zero, with a regularized parameterization for the slopes and intercept of Normal(0, 0.1) for each *&beta;*, and a slightly less skeptical regularization for the change point Normal(0, 0.5).
 These regularizing priors were selected to reduce overfitting while still allowing the model to learn the regular features of the sample, and so provide a more robust population estimate.
@@
 _Model Selection_  
-To determine whether wellbeing was a linear or non-linear (e.g., piecewise) function of income, we compared the out-of-sample deviance of the linear and piecewise model fits using the [Widely Applicable Information Criterion (WAIC)](https://bookdown.org/ajkurz/Statistical_Rethinking_recoded/overfitting-regularization-and-information-criteria.html#the-problem-with-parameters){target="_blank"}.
-The WAIC is the log-posterior predictive density plus a penalty proportional to the variance in the posterior distribution, so it represents an approximation of the out-of-sample deviance that converges to the cross-validation accuracy in a large sample, with a penalty for the effective number of parameters (degrees of freedom).
+We compared the out-of-sample deviance of the linear and piecewise model fits using the [Widely Applicable Information Criterion (WAIC)](https://bookdown.org/ajkurz/Statistical_Rethinking_recoded/overfitting-regularization-and-information-criteria.html#the-problem-with-parameters){target="_blank"}.
+The WAIC represents an approximation of the out-of-sample deviance that converges to the cross-validation accuracy in a large sample, with a penalty for the effective number of parameters (degrees of freedom).
 Thus using the out-of-sample deviance for model selection in combination with regularizing priors in our model design is a dual strategy to reduce overfitting and penalize overfitting.
-This makes it a useful approach for comparing models of varying complexity, such as our linear and piecewise (nonlinear) models.
 As with other deviance metrics, smaller negative WAIC values are better (i.e., indicate more accuracy).
@@
 To determine the location of the change point (*&omega;*) between wellbeing and income, we modelled the relationship between income and wellbeing across individuals using the piecewise model described above, and sampled the posterior probability of *&omega;* over 4000 interations.
-The complete posterior distribution of *&omega;* for each year is presented along with the expected value (mean).
+The complete posterior distribution of *&omega;* for each year is presented along with the expected value (median).
 
@@
 
-## Results
+## 3.0 Results
 
@@
 
-#### Happiness and satisfaction have distinct relationships with income
+#### 3.1 Happiness and satisfaction have distinct relationships with income
 
@@
 For each wellbeing variable we show the results of a linear fit (rows 1 and 3) and a piecewise fit (rows 2 and 4).
-For visualization purposes only, we display the mean levels of income and wellbeing for each (equal-sized) income decile, whereas the line-of-best-fit and 95% credible intervals (shaded) in each regression model are derived from the underlying _individual_ data-points.
+For visualization purposes only, due to the large number of individuals in each year, we display the mean levels of income and wellbeing for each (equal-sized) income decile, whereas the line-of-best-fit and 95% credible intervals (shaded) in each regression model are derived from _all individuals_.
 
@@
 The evidence from model selection revealed the nonlinear (piecewise) fit of _happiness_, but not _satisfaction_, was superior to a linear fit in each year (Supplementary Materials, Figure S1).
-The WAIC difference indicated the piecewise fit of happiness had reliably less deviance than the linear fit, indicating the piecewise model predicted happiness from income more accurately using this out-of-sample metric.
 Overall the model comparison suggested that happiness and satisfaction have distinct relationships with household income; satisfaction tends to increase linearly with income, while a change point exists in the relationship between happiness and household income.
@@
 
-#### Temporal trends in the association between income and happiness 2001-2019
+#### 3.2 Temporal trends in the association between income and happiness 2001-2019
 
 Figure 2 below presents the posterior distribution of each parameter from the piecewise model regressing happiness on income: the change point (&omega;), the intercept (&beta;~0~), the pre-change point slope (&beta;~1~), and the post-change point slope (&beta;~2~).
-Over 4,000 samples were drawn from each piecewise model to determine the posterior distribution of each parameter (smallest bulk effective sample size = 661, see Table S2 for the complete fit statistics in Supplementary materials).
 Horizontal bars represent the 95% credible interval of the posterior distribution and so intervals which fall completely to the right of the vertical grey dotted line are credibly higher than the expected value of our base year, 2001.
@@
 
-The change in parameter values between 2001 to 2019 indicates the relationship between happiness and income has evolved over time.
+The change in parameter values between 2001 to 2019 indicates the relationship between happiness and income evolved over time.
 We determined the impact of this evolution on the distribution of happiness over the range of household income in 2019 in a counterfactual analysis.
@@
 ```
-Figure 3 legend: The difference (∆) in predicted happiness between 2019 and 2001 for the same n = 14,459 individuals from the 2019 survey.
+Figure 3 legend: The difference (∆) in predicted happiness between 2019 and 2001 for the same n = 14,459 individuals.
 Values below zero on the y-axis indicate lower happiness predictions using the 2019 model compared to using the 2001 model.
-Solid line overlay represents the (loess) smoothed average delta over the income distribution.
-Variance in these posterior samples is from the sample weights, uncertainty in the model parameters as well as observation-level (residual) variance.
+The smoothed overlay (solid line) indicates how the average happiness changes across the income distribution.
 ```
@@
 
-#### The increasing cost of happiness in Australia
+#### 3.4 The increasing cost of happiness in Australia
 
@@
 
-## Conclusions
+## 4.0 Discussion
 
-This analysis of the Australian population since the millennium confirms previous findings [e.g., @howell2008relation; @kahneman2010high], that the relationship between the two types of subjective wellbeing and household income was positive but quite different: Satisfaction increased linearly with income, while happiness increased relatively steeply up to a point after which higher levels of income were associated with much more modest differences.
-We refer to the change point after which increases in income correspond to smaller increases in happiness as the _cost of happiness_.
-After this point, happiness is no longer as responsive to differences in household income, and the economic security it represents.
+We confirmed previous findings [e.g., @howell2008relation; @kahneman2010high], that the relationship between both types of subjective wellbeing with household income was positive but quite different: Satisfaction increased linearly with income, while happiness increased rapidly up to a point after which higher levels of income were associated with less improvement.
+Our novel finding was that this change point in the relationship between household income and happiness increased faster than both inflation and median household income between 2001 and 2019.
+For the first time we have shown there has been a temporal shift in the change point between income and happiness over the 19 year period, such that happiness has become more dependent on income for more people - especially the poor and middle-class.
+
+We refer to the change point after which increases in income no longer produce similar increases in happiness as the _cost of happiness_.
+After this point, happiness is no longer as dependent on household income, and the economic security it represents.
 Presumably after this point further increases in happiness depend more on other life factors (e.g., leisure time, social connections) than financial security.
-Our novel finding was that this change point in the relationship between household income and happiness increased faster than both inflation and median household income between 2001 and 2019.
-For the first time we have shown there has been a temporal shift in the change point between income and happiness over the 19 year period, such that happiness of an increasing proportion of the population has become more responsive to differences in income - especially the poor and middle-income earners.
-
 Life satisfaction on the other hand appeared to show consistent increases with household income and we found no evidence of any change point.
@@
 As such, these results may well have relevance to other developed nations in North America and Europe which also enjoy low levels of income inequality, but have stagnating incomes, and declining happiness levels [@sachs2019world].
-As focus shifts from traditional financial indicators towards health and wellbeing measures, findings such as this may become more prevalent.
-
-The relationship between happiness and income is often described as log-linear [e.g., @kahneman2010high; @killingsworth2021experienced].
-Under this view, similar proportional changes in income produce similar evaluations of satisfaction or happiness, rather than the absolute change.
-Likewise, the piecewise model reported here has a slope that reduces with income, albeit with a single identifiable change-point rather than a smooth curve.
-However our main thesis is not committed to a function with a single or smooth changepoint; only that the slope between income and happiness is unequal at different levels of income, and the function is shifting to the right over time (e.g., Figure 2).
-The choice between functions is somewhat arbitrary, since a log-linear function is as easily consistent with our thesis as the piecewise function we describe.
-Nevertheless, the model evidence favoured a piecewise- over a log-linear function in a supplementary analysis (Figure S3), which supports our model choice here.
-
-Some recent studies have challenged the notion that the positive effect of income plateaus after some level ("satiety") [@killingsworth2021experienced; @twenge2020expanding; @stevenson2013subjective].
-Our results also do not support a "satiety" point or flat gradient between income and happiness, as the average post-change slope between 2001-2019 was reliably above zero.
-The pre-change slope however was also four to five times greater than that, and we argue this difference in slopes indicates a previously unacknowledged inequity in the distribution of happiness between the rich and poor.
-
-While the amount of variance in happiness explained by household income is small (Table S4, Supplementary materials), the slope is reliably above zero and small changes in happiness over an entire population will aggregate.
-Moreover, financial gains have one of the largest impacts on individual happiness in within-subject models (e.g., fixed-effects), second only to marriage by our estimate [@kettlewell2020differential], and there are many reasons why cross-sectional estimates of the covariance would under-represent this.
-Finally, happiness is measured on a subjective scale, where small changes at different points on the scale may represent a large change in some other functional/real-world outcome (e.g., risk of suicide).
-Ultimately, the real-world impact of small changes in a subjective measure such as happiness is inherently difficult to determine.
-Our aim here was to determine whether the relationship between income and happiness is changing at all, and whether this may be contributing to increasing inequity, and our results indicate this is the case.
-
-As governments and policy-makers begin to focus on wellbeing, it will be critical to understand how traditional economic indicators such as household income, wealth inequality, and consumption interact with wellbeing and health.
-According to @frijters2020happy, coming up with a consensus to translate income into wellbeing features high on the wider wellbeing research agenda.
-Establishing the links between wealth, household income, wellbeing and health, and how inequalities in one drives inequities in the other, will be a critical step in that agenda.
+Establishing the links between income, wellbeing and health, and how inequalities in one drives inequities in the other, will be a critical aim as government begin to shift focus away from traditional measures of economic prosperity.
 
@@
 
-## Acknowledgements
-
-This research was not supported by any funding.
-
-<br><br>
-
-## Author contributions
-
-RWM designed the analysis, analysed the data and wrote the paper. NK designed the analysis and wrote the paper. NG proposed the idea, planned the analysis, and wrote the paper.
-
-<br><br>
-
-## Competing interests
-
-The authors declare no competing interests.
-
-<br><br>
-
-## Additional information
-
-This paper uses unit record data from Household, Income and Labour Dynamics in Australia Survey [HILDA](https://melbourneinstitute.unimelb.edu.au/hilda) conducted by the Australian Government Department of Social Services (DSS).
-The findings and views reported in this paper, however, are those of the author[s] and should not be attributed to the Australian Government, DSS, or any of DSS’ contractors or partners.
-All code and scripts used in the analysis are available at https://github.com/datarichard/The-increasing-cost-of-happiness
-
-Supplementary information is available [online](https://datarichard.github.io/The-increasing-cost-of-happiness/supplementary.html.
+## Data Statement  
+This paper uses unit record data from Household, Income and Labour Dynamics in Australia Survey HILDA conducted by the Australian Government Department of Social Services (DSS).
+All code and scripts used in the analysis are available [online](https://github.com/datarichard/The-increasing-cost-of-happiness).
 

### 2021-05-27 src/draft-html0.Rmd
SSM0 Population Health submission

--- 2021-05-27 src/draft-PNAS0.Rmd
@@
 output:  
-  word_document:
-    reference_docx: ../src/Rmarkdown-for-Word-styles.docx
+  html_document:
+    toc: true
+    toc_depth: 4
+    toc_float:
+      collapsed: false
 always_allow_html: yes
@@
 ## Significance Statement
+
 Income has decreasing returns on happiness, across countries, cultures and ethnicities, with an inflection or change point near USD75,000.
@@
 
-Supplementary information is available [online](https://datarichard.github.io/The-increasing-cost-of-happiness/supplementary.html.
+Supplementary information is available [online](https://datarichard.github.io/The-increasing-cost-of-happiness/supplementary.html){target="_blank"}.
 

### 2021-05-27 src/supplementary-SSM0.Rmd
SSM0 Population Health submission

--- 2021-03-12 src/supplementary0_word.Rmd
@@
 title: "Supplementary material: The increasing cost of happiness"
-author: "R.W. Morris^1,2^, N. Kettlewell^2,3,4^  & N.Glozier^1,2^"
 bibliography: ../src/references.bib
@@
 output:  
-  word_document:
-    reference_docx: ../src/Rmarkdown-for-Word-styles.docx
+  html_document:
+    reference_docx: ../src/academic_style.docx
 always_allow_html: yes
@@
 \  
-\
-
-1. Central Clinical School, Faculty of Medicine and Health, University of Sydney, NSW, Australia
-2. ARC Centre of Excellence for Children and Families over the Life Course
-3. School of Economics, University of Technology, NSW, Australia
-4. Institute of Labor Economics (IZA), Bonn, Germany
-
-\
-
-**Corresponding author:**  
-Professor Nick Glozier  
-Faculty of Medicine and Health,   
-University of Sydney,  
-NSW 2050,  
-Australia  
-email: nick.glozier@sydney.edu.au
-
 \
@@
 
+Model fitting was performed in RStan (v2.21.2) using the brms (v2.14.4) package [@burkner2017; @rstan2019].
 There were _N_ = 59,876 total observations.

### 2021-05-27 src/plotfun.R

@@ -1,0 +1,375 @@
+plot_linear <- function(.data, swb_col, dollars_col) {
+ 
+
+  swb_colname <- rlang::enexpr(swb_col)
+  dollars_colname <- rlang::enexpr(dollars_col)
+  
+  .data %>%
+    select(xwaveid, 
+           year, 
+           swb = !!swb_colname, dollars = !!dollars_colname) %>%
+    filter(swb >= 0) -> df
+  
+  df %>%
+    group_by(year) %>%
+    summarise(n = n()) %>%
+    mutate(note = paste("n =", format(n, big.mark = ","))) -> df.sum
+  
+  df %>%
+    group_by(year) %>%
+    mutate(
+      dollar_bin = ntile(dollars, 10)
+    ) %>%
+    group_by(year, dollar_bin) %>%
+    summarize(
+      wellbeing = mean(swb, na.rm=T),
+      wealth = mean(dollars, na.rm=T)/1000
+      ) -> df.deciles
+  
+  df %>%
+    mutate(
+      wealth = dollars/1000,
+      wellbeing = swb
+    ) -> df.plot
+  
+  # Linear model
+  ggplot(df.plot, aes(x = wealth, y = wellbeing)) +
+    geom_smooth(aes(fill = "manual1", color = "manual2"),
+      method = "lm", 
+      size = 0.5
+      ) +
+    coord_cartesian(xlim = c(0, max(df.deciles$wealth)),
+                    ylim = c(min(df.deciles$wellbeing), max(df.deciles$wellbeing))) +
+    facet_wrap(~year, nrow = 1) +
+    labs(subtitle = paste(rlang::as_name(swb_colname), "(linear)"),
+         x = paste(rlang::as_name(dollars_colname), "($000s)"),
+         y = "") +
+    theme_test() +
+    theme(
+      legend.position = "none") -> p1
+  
+  p1 +
+    geom_point(data = df.deciles) + # size = 3, alpha = 0.25
+    geom_text(data = df.sum, aes(x = Inf, y = -Inf, label = note),
+              hjust = 1.1, vjust = -.75, colour = 1, size = 3)
+  
+}
+
+
+plot_inflection_piecewise <- function(.data, swb_col, dollars_col, knots = 1) {
+  
+  swb_colname <- rlang::enexpr(swb_col)
+  dollars_colname <- rlang::enexpr(dollars_col)
+  
+  .data %>%
+    select(xwaveid, year, 
[…]

### 2021-05-27 src/preprocessing.R

@@ -142,6 +142,62 @@ happiness <- gh9_items %>%
             wave, 
             gh9 = scales::rescale(gh9, to = c(1, 100))) 
 
+# Split gh9 into positive and negative components
+positive_items <- c("gh9a", "gh9d", "gh9e", "gh9h")
+negative_items <- c("gh9b", "gh9c", "gh9f", "gh9g", "gh9i")
+
+positive_affect <- gh9_items %>%
+  select(xwaveid, wave, all_of(positive_items)) %>%
+  # Reverse score
+  mutate_at(vars(any_of(reversed_items)), list(~ 7 - .)) %>%
+  gather(code, val, -xwaveid, -wave) %>%
+  # Impute average if less than half missing (see ghmh data dictionary)
+  group_by(xwaveid, wave) %>%
+  mutate(
+    sum_na = sum(is.na(val)),
+    mean_na = mean(val, na.rm = TRUE),
+    imputed = ifelse(is.na(val) & sum_na < 2, mean_na, val),
+    sum_imputed = sum(imputed),
+    code = "gh9_sum"
+  ) %>% 
+  ungroup() %>%
+  select(xwaveid, wave, gh9 = sum_imputed) %>%
+  distinct() %>%
+  left_join(select(gh9_items, xwaveid, wave), 
+            by = c("xwaveid", "wave")
+  ) %>%
+  transmute(xwaveid, 
+            wave, 
+            gh9_positive = scales::rescale(gh9, to = c(1, 100))) 
+
+negative_affect <- gh9_items %>%
+  select(xwaveid, wave, all_of(negative_items)) %>%
+  # Reverse score
+  mutate_at(vars(any_of(reversed_items)), list(~ 7 - .)) %>%
+  gather(code, val, -xwaveid, -wave) %>%
+  # Impute average if less than half missing (see ghmh data dictionary)
+  group_by(xwaveid, wave) %>%
+  mutate(
+    sum_na = sum(is.na(val)),
+    mean_na = mean(val, na.rm = TRUE),
+    imputed = ifelse(is.na(val) & sum_na < 3, mean_na, val),
+    sum_imputed = sum(imputed),
+    code = "gh9_sum"
+  ) %>% 
+  ungroup() %>%
+  select(xwaveid, wave, gh9 = sum_imputed) %>%
+  distinct() %>%
+  left_join(select(gh9_items, xwaveid, wave), 
+            by = c("xwaveid", "wave")
+  ) %>%
+  transmute(xwaveid, 
+            wave, 
[…]

### 2021-06-02 src/draft-SSM1.Rmd
Numbered equations (closes issue #23) and updates to Figure 1 (issue #10)

--- 2021-05-27 src/draft-SSM0.Rmd
@@
 
-We modelled the relationship between income and each wellbeing variable (happiness and satisfaction) using a simple linear model and a piecewise model with a single change-point as a free parameter estimated from the data.
-The piecewise model was chosen as the simplest extension of a linear model which can identify a change point (inflection) in the relationship between wellbeing and income.
-For modelling, both measures of wellbeing and income were rescaled with a mean of zero and a SD of 1 (z-scores) for each year.
+We modelled the relationship between income and each wellbeing variable (happiness and satisfaction) using a log-linear model and a piecewise-linear model with a single change-point as a free parameter estimated from the data.
+The piecewise-linear model was chosen as the simplest extension of a linear model which can identify a change point (inflection) in the relationship between wellbeing and income.
+For modelling, income values were log transformed and then both measures of wellbeing and income were rescaled with a mean of zero and a SD of 1 (z-scores) for each year.
 
 _Model Design_  
-We adopted a Bayesian approach for estimating the linear and piecewise model in the software Stan [@burkner2017; @rstan2019]. In each case the linear model was estimated as:
+We adopted a Bayesian approach for estimating the log-linear and piecewise-linear model in the software Stan [@burkner2017; @rstan2019]. In each case the log-linear model was estimated as:
 
 $$
-y_i \sim N(\mu_i, \sigma^2_y)
+  y_i \sim N(\mu_i, \sigma^2_y) \ \ \ \ \ \ \ \ \ \ (1)
 $$
@@
 $$
-\mu_i = \beta_0 + \beta_1 X_i 
+  \mu_i = \beta_0 + \beta_p X_i \ \ \ \ \ \ \ \ \ \ (2)
 $$
 
-Where $X_i$ was an individual's household income (\$) as well as other covariates (age, age^2^, sex, education, chronic illness), and $y_i$ was an individual's wellbeing.
+Where $X_i$ was an individual's log household income (log(\$)) as well as other covariates (i.e., age, age^2^, sex, education, chronic illness, HILDA population weights; see below), and $y_i$ was an individual's wellbeing.
+Note that $\beta_p$ represents a term 1...p for each parameter in $X_i$.
 
@@
 
-The piecewise model included a free parameter to represent the changepoint in income ($\omega$) as well as the slope before the change point ($\beta_1$) and the slope after the change point ($\beta_2$):
+The piecewise-linear model included a free parameter to represent the changepoint in income ($\omega$) as well as the slope before the change point ($\beta_1$) and the slope after the change point ($\beta_2$):
 
 $$
-\mu_i = \beta_0 + \beta_1 (x_i - \omega) (x_i ≤ \omega) + \beta_2 (x_i - \omega) (x_i > \omega) + \beta_3 X_i
+\mu_i = \beta_0 + \beta_1 (x_i - \omega) (x_i ≤ \omega) + \beta_2 (x_i - \omega) (x_i > \omega) + \beta_p X_i \ \ \ (3)
 $$
 
-Where $x_i$ was an individual's household income, and $X_i$ were the cross-sectional HILDA population weights to adjust for differences in the sample representativeness according to sex by broad age, marital status, region, and labour force status.
+Where $x_i$ was an individual's household income (\$), and $X_i$ were the covariates-of-no-interest (see below for complete description).
 
@@
 
-The parameters of the piecewise model used Gaussian priors centred at zero, with a regularized parameterization for the slopes and intercept of Normal(0, 0.1) for each *&beta;*, and a slightly less skeptical regularization for the change point Normal(0, 0.5).
+The parameters of the piecewise-linear model used Gaussian priors centred at zero, with a regularized parameterization for the slopes and intercept of Normal(0, 0.1) for each *&beta;*, and a slightly less skeptical regularization for the change point Normal(0, 0.5).
 These regularizing priors were selected to reduce overfitting while still allowing the model to learn the regular features of the sample, and so provide a more robust population estimate.
@@
 _Model Selection_  
-We compared the out-of-sample deviance of the linear and piecewise model fits using the [Widely Applicable Information Criterion (WAIC)](https://bookdown.org/ajkurz/Statistical_Rethinking_recoded/overfitting-regularization-and-information-criteria.html#the-problem-with-parameters){target="_blank"}.
+We compared the out-of-sample deviance of the log-linear and piecewise-linear model fits using the [Widely Applicable Information Criterion (WAIC)](https://bookdown.org/ajkurz/Statistical_Rethinking_recoded/overfitting-regularization-and-information-criteria.html#the-problem-with-parameters){target="_blank"}.
 The WAIC represents an approximation of the out-of-sample deviance that converges to the cross-validation accuracy in a large sample, with a penalty for the effective number of parameters (degrees of freedom).
@@
 _Covariates_  
-Cross-sectional population weights for Australia provided by the University of Melbourne for each year were included as a covariate to adjust for differences in the sample representativeness.
+Covariates for age, age^2^, sex, education, and chronic illness were included in each model above.
+Cross-sectional population weights for Australia provided by the University of Melbourne for each year were also included as a covariate to adjust for differences in the sample representativeness (by sex, broad age, marital status, region, and labour force status).
+Thus each model includes terms for the linear effect of each covariate, and age^2^ includes a term for the quadratic effect.
 Full-time students were removed, as well as individuals with an annual household disposable income that was negative or indicated as _topcoded_ by the University of Melbourne.
+The results of sensitivity analyses to determine the effect of the covariates and exclusions on the change point parameters are included in Supplementary Materials (**add section reference here**).
 
@@
     year %in% c(2001, 2005, 2010, 2015, 2019), # 2001, 2007, 2013, 2019
-    `Household income` >= 0
+    `Household income` > 0
   ) -> figure1_data
+
+log_breaks <- c(log(12000), log(25000), log(50000), log(100000))
 
@@
   rename(`Happiness 1-100` = gh9) %>% 
-  plot_linear(`Happiness 1-100`, `Household income`) +
+  plot_log2(`Happiness 1-100`, `Household income`) +
+  scale_x_continuous(breaks = log_breaks, labels = c(12, 25, 50, 100)) +
   scale_fill_manual(values = "#C79999") +
@@
   rename(`Satisfaction 0-10` = losat) %>% 
-  plot_linear(`Satisfaction 0-10`, `Household income`) +
+  plot_log2(`Satisfaction 0-10`, `Household income`) +
+  scale_x_continuous(breaks = log_breaks, labels = c(12, 25, 50, 100)) +
   scale_fill_manual(values = "#b3cde0") +
@@
 ```
+
 ```{r figure_1, fig.height=10, fig.width=9, dpi=300}

### 2021-06-02 src/draft-SSM1.Rmd
Replace wealth with income (closes issue #6)

@@
 A fundamental question for society is how much happiness does a dollar buy?
-The accepted view among economists and psychologists is that money and happiness increase together up to a point, after which there is little further gain from increasing wealth.
+The accepted view among economists and psychologists is that money and happiness increase together up to a point, after which there is little further gain from increasing income.
 While the location of this change point reportedly ranges between USD\$60K to \$95K, there has been no investigation as to whether this has increased or decreased over time.
@@
 Rather we are revealing that even a static income distribution may have dynamic effects on happiness over time.
-This highlights the issue that while traditional univariate measures of wealth and income inequality may be relatively stable and exhibit little change, their impact on wellbeing and health can still vary.
+This highlights the issue that while traditional univariate measures of income and income inequality may be relatively stable and exhibit little change, their impact on wellbeing and health can still vary.
 As such, these results may well have relevance to other developed nations in North America and Europe which also enjoy low levels of income inequality, but have stagnating incomes, and declining happiness levels [@sachs2019world].

### 2021-06-02 src/draft-SSM1.Rmd
Cognitive- and affective wellbeing are used throughout (closes issue #12)

@@
 We tested the temporal relationship between income and affective wellbeing (happiness), and income and cognitive wellbeing (life satisfaction), using household economic data from Australia between 2001-2019.
-We discovered the change point between happiness and income has increased over those 19 years faster than inflation (i.e., cost of living).
+We discovered the change point between affective wellbeing and income has increased over those 19 years faster than inflation (i.e., cost of living).
 This suggests that inequalities in income may be driving increasing inequities in happiness between the rich and the poor, with implications for health and recent government policy-goals to monitor and improve wellbeing.
@@
 - Nineteen years of data from a well-developed, economically stable G20 nation is examined (Australia), allowing a relatively undisturbed view of the changes without major economic shocks
-- We are able to show for the first time that the income level at which happiness is less determined by income has risen since 2001, resulting in greater happiness inequality between the rich and the poor
-- Conversely the linear relationship between income and life-satisfaction has remained constant over this period
+- We are able to show for the first time that the income level at which affective wellbeing is less determined by income has risen since 2001, resulting in greater inequality in affective wellbeing (happiness) between the rich and the poor
+- Conversely the linear relationship between income and cognitive wellbeing (life-satisfaction) has remained constant over this period
 - The happiness of an increasing proportion of the population is more dependent on their financial security than ever this millennium
@@
 
-Kahneman & Deaton [-@kahneman2010high] showed that self-reported happiness increased with household income up to a point (USD75,000), but after that, increasing income had little further effect.
-Conversely life satisfaction continued to increase with income beyond USD75,000.
 Subjective wellbeing is not a unitary entity [@diener2017findings]; studies typically distinguish between life satisfaction, the cognitive appraisal of one’s own accomplishments, and affective wellbeing, one’s prevailing affective state, emotional mood, or everyday experience of happiness [e.g., @kettlewell2020differential].
+Kahneman & Deaton [-@kahneman2010high] showed that affective wellbeing increased with household income up to a point (USD75,000), but after that, increasing income had little further effect.
+Conversely cognitive wellbeing continued to increase with income beyond USD75,000.
 Indeed, the difference between the two questions: “How satisfied are you with your life?” and “How happy are you these days?” has been identified as a crucial mediating factor in a meta-analysis of 111 studies on income and wellbeing [@howell2008relation; also @veenhoven2006rising].
-Results since then have also shown that while both happiness and satisfaction increase with income, the happiness returns are decreasing [@killingsworth2021experienced; @sacks2012new].
+Results since then have also shown that while both affective and cognitive wellbeing increase with income, the affective returns are decreasing [@killingsworth2021experienced; @sacks2012new].
 
-Fundamentally, the existence of a change point in the relationship between income and happiness reveals an unacknowledged source of inequality in the distribution of wellbeing (i.e., happiness) in the economy.
-For instance, a change point of USD75,000 in 2008 was substantially more than the US median income of USD52,000 in the same year, indicating that the happiness of the poorest majority of the US population was tied to marginal changes in income while the happiness of a richer minority was not.
-Thus the change point represents the dollar value up to which income drives inequities in the distribution of happiness, such that a lower valued change point represents a more equitable distribution of happiness in the economy.
+Fundamentally, the existence of a change point in the relationship between income and affective wellbeing reveals an unacknowledged source of inequality in the distribution of wellbeing (i.e., happiness) in the economy.
+For instance, a change point of USD75,000 in 2008 was substantially more than the US median income of USD52,000 in the same year, indicating that the affective wellbeing of the poorest majority of the US population was tied to marginal changes in income while the affective wellbeing of a richer minority was not.
+Thus the change point represents the dollar value up to which income drives inequities in the distribution of affective wellbeing, such that a lower valued change point represents a more equitable distribution of affective wellbeing in the economy.
 Inequities in the distribution of wellbeing are increasingly relevant to governments and policy-makers due to the growing recognition that increasing income does not necessarily lead to equal changes in wellbeing [@clark2018four; @frijters2020happy].
-Even prior to COVID-19, the World Gallup Poll has observed that happiness has decreased over the past decade in western Europe, North America, Australia and New Zealand, despite increases in income in the same countries [@sachs2019world].
-However to date there has been no investigation of whether the relationship between income and happiness has changed over time which may contribute to these trends.
-In particular has the change point between income and happiness, and therefore the distribution of happiness between rich and poor, become more or less equitable in the last few decades?
+Even prior to COVID-19, the World Gallup Poll has observed that affective wellbeing has decreased over the past decade in western Europe, North America, Australia and New Zealand, despite increases in income in the same countries [@sachs2019world].
+However to date there has been no investigation of whether the relationship between income and affective wellbeing has changed over time which may contribute to these trends.
+In particular has the change point between income and affective wellbeing, and therefore the distribution of happiness between rich and poor, become more or less equitable in the last few decades?
 
@@
 HILDA (the Household, Income, and Labour Dynamics in Australia survey) provides a representative sample of households in Australia with detailed measurements of income and subjective wellbeing each year.
-We distinguished between satisfaction and happiness as different components of subjective wellbeing, and evaluated how each varied with household income over time using piecewise linear regression to determine the change points in each year.
+We distinguished between cognitive- and affective wellbeing as different components of subjective wellbeing, and evaluated how each varied with household income over time using piecewise linear regression to determine the change points in each year.
 
@@
 #### 2.2 Subjective Wellbeing
-Life satisfaction was assessed by a single item question asked each survey: "All things considered, how satisfied are you with your life (0 to 10)".
+Cognitive wellbeing was assessed by a single item question asked each survey: "All things considered, how satisfied are you with your life (0 to 10)".
 
-Happiness was determined by 9 questions in the SF-36 [(9a to 9i)](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=gh9a).
+Affective wellbeing was determined by 9 questions in the SF-36 [(9a to 9i)](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=gh9a).
 The SF-36 is a widely used self-completion measure of various aspects of physical, emotional and mental health [@ware2000sf].
@@
 We reverse scored negatively phrased questions and calculated the sum of the nine questions so that higher scores represented better wellbeing.
-To aid interpretability, we rescaled the final sum to a score between 1-100, where 100 represents the maximum happiness achievable.
+To aid interpretability, we rescaled the final sum to a score between 1-100, where 100 represents the maximum affective wellbeing achievable.
 
@@
 
-We modelled the relationship between income and each wellbeing variable (happiness and satisfaction) using a log-linear model and a piecewise-linear model with a single change-point as a free parameter estimated from the data.
+We modelled the relationship between income and each wellbeing variable (cognitive and affective) using a log-linear model and a piecewise-linear model with a single change-point as a free parameter estimated from the data.
 The piecewise-linear model was chosen as the simplest extension of a linear model which can identify a change point (inflection) in the relationship between wellbeing and income.
@@
 $$
-  \mu_i = \beta_0 + \beta_p X_i \ \ \ \ \ \ \ \ \ \ (2)
+  \mu_i = \beta_0 + B X_i \ \ \ \ \ \ \ \ \ \ (2)
 $$
@@
 Where $X_i$ was an individual's log household income (log(\$)) as well as other covariates (i.e., age, age^2^, sex, education, chronic illness, HILDA population weights; see below), and $y_i$ was an individual's wellbeing.
-Note that $\beta_p$ represents a term 1...p for each parameter in $X_i$.
+Note that $B$ is a column matrix with a term 1...p for each parameter in $X_i$.
 
@@
 $$
-\mu_i = \beta_0 + \beta_1 (x_i - \omega) (x_i ≤ \omega) + \beta_2 (x_i - \omega) (x_i > \omega) + \beta_p X_i \ \ \ (3)
+\mu_i = \beta_0 + \beta_1 (x_i - \omega) (x_i ≤ \omega) + \beta_2 (x_i - \omega) (x_i > \omega) + B X_i \ \ \ (3)
 $$
@@
 The broad demographic characteristics of the sample are presented in Supplementary Materials Table S1.
-Household income and average life satisfaction levels increased between 2001-2019, while average happiness scores decreased slightly over the 19 years.
+Household income and average cognitive wellbeing levels increased between 2001-2019, while average affective wellbeing scores decreased slightly over the 19 years.
 The proportions of each sex and couples were stable over time, as were the average household size and SEIFA index.
@@
 
-#### 3.1 Happiness and satisfaction have distinct relationships with income
+#### 3.1 Cognitive- and Affective wellbeing have distinct relationships with income
 
-The relationship between household income and happiness (red) and satisfaction (blue) every four years is shown in Figure 1.
+The relationship between household income and affective wellbeing (red) and satisfaction (blue) every four years is shown in Figure 1.
 For each wellbeing variable we show the results of a linear fit (rows 1 and 3) and a piecewise fit (rows 2 and 4).
@@
 
-##### Figure 1. Household income and happiness (red) and satisfaction (blue)  
+##### Figure 1. Household income and affective- (red) and cognitive- (blue) wellbeing  
 ```{r hh_disp_inc_eq}
@@
 p1 <- figure1_data %>%
-  rename(`Happiness 1-100` = gh9) %>% 
-  plot_log2(`Happiness 1-100`, `Household income`) +
+  rename(`Affective 1-100` = gh9) %>% 
+  plot_log2(`Affective 1-100`, `Household income`) +
   scale_x_continuous(breaks = log_breaks, labels = c(12, 25, 50, 100)) +
@@
 p2 <- figure1_data %>%
-  rename(`Happiness 1-100` = gh9) %>% 
-  plot_inflection_custom(`Happiness 1-100`, `Household income`) +
+  rename(`Affective 1-100` = gh9) %>% 
+  plot_inflection_custom(`Affective 1-100`, `Household income`) +
   scale_fill_manual(values = "#C79999") +
@@
 p3 <- figure1_data %>%
-  rename(`Satisfaction 0-10` = losat) %>% 
-  plot_log2(`Satisfaction 0-10`, `Household income`) +
+  rename(`Cognitive 0-10` = losat) %>% 
+  plot_log2(`Cognitive 0-10`, `Household income`) +
   scale_x_continuous(breaks = log_breaks, labels = c(12, 25, 50, 100)) +
@@
 p4 <- figure1_data %>%
-  rename(`Satisfaction 0-10` = losat) %>% 
-  plot_inflection_custom(`Satisfaction 0-10`, `Household income`) +
+  rename(`Cognitive 0-10` = losat) %>% 
+  plot_inflection_custom(`Cognitive 0-10`, `Household income`) +
   scale_fill_manual(values = "#b3cde0") +
@@
 Figure 1 legend: The relationship between income and wellbeing across equal-sized income deciles, overlaid by regression lines from linear and piecewise models (±95%CI).
-Wellbeing was measured as happiness (red) or life satisfaction (blue).
+Wellbeing was measured as affective- (red) or cognitive- (blue) wellbeing.
 The total number of individuals contributing to each regression in each year are noted (n).
@@
 
-The nonlinear relationship between happiness and income (Figure 1, 2nd row) was consistently and negatively inflected (happiness increased less with income after the change point).
-By contrast the nonlinear relationship between satisfaction and income shown in the 4th row exhibited negative inflection (2001, 2015, 2019), positive inflection (2005), and no apparent inflection (2010).
+The nonlinear relationship between affective wellbeing and income (Figure 1, 2nd row) was consistently and negatively inflected (affective wellbeing increased less with income after the change point).
+By contrast the nonlinear relationship between cognitive wellbeing and income shown in the 4th row exhibited negative inflection (2001, 2015, 2019), positive inflection (2005), and no apparent inflection (2010).
 
-The evidence from model selection revealed the nonlinear (piecewise) fit of _happiness_, but not _satisfaction_, was superior to a linear fit in each year (Supplementary Materials, Figure S1).
-Overall the model comparison suggested that happiness and satisfaction have distinct relationships with household income; satisfaction tends to increase linearly with income, while a change point exists in the relationship between happiness and household income.
+The evidence from model selection revealed the nonlinear (piecewise) fit of _affective_-, but not _cognitive_ wellbeing, was superior to a linear fit in each year (Supplementary Materials, Figure S1).
+Overall the model comparison suggested that cognitive- and affective wellbeing have distinct relationships with household income; cognitive wellbeing tends to increase linearly with income, while a change point exists in the relationship between affective wellbeing and household income.
 
@@
 
-#### 3.2 Temporal trends in the association between income and happiness 2001-2019
+#### 3.2 Temporal trends in the association between income and affective wellbeing 2001-2019
 
-Figure 2 below presents the posterior distribution of each parameter from the piecewise model regressing happiness on income: the change point (&omega;), the intercept (&beta;~0~), the pre-change point slope (&beta;~1~), and the post-change point slope (&beta;~2~).
+Figure 2 below presents the posterior distribution of each parameter from the piecewise model regressing affective wellbeing on income: the change point (&omega;), the intercept (&beta;~0~), the pre-change point slope (&beta;~1~), and the post-change point slope (&beta;~2~).
 Horizontal bars represent the 95% credible interval of the posterior distribution and so intervals which fall completely to the right of the vertical grey dotted line are credibly higher than the expected value of our base year, 2001.
@@
 
-##### Figure 2. Posterior parameters of the happiness ~ income piecewise model (real 2019 dollars)   
+##### Figure 2. Posterior parameters of the affective-wellbeing ~ income piecewise model (real 2019 dollars)   
 ```{r get_posteriors}
@@
 ```
-Figure 2 legend: Posterior distributions of the change point parameter representing the location in real household income (real 2019 dollars), as well as the intercept, pre-slope and post-slope parameters in SD units (happiness).
+Figure 2 legend: Posterior distributions of the change point parameter representing the location in real household income (real 2019 dollars), as well as the intercept, pre-slope and post-slope parameters in SD units (affective wellbeing).
 Horizontal bar represents the 95% credible region and the solid point indicates the expected value (median) of each distribution.
@@
 
-The posterior estimates of the change point in the association between happiness and household income indicates that the probable location (i.e., the real income value) of the change point shows a systematic increasing trend since 2001.
-Changes to the other parameters of the function between income and happiness also occurred between 2001 and 2019 (i.e., the pre-slope, post-slope and intercept), but did not show any sustained trend over the period.
+The posterior estimates of the change point in the association between affective wellbeing and household income indicates that the probable location (i.e., the real income value) of the change point shows a systematic increasing trend since 2001.
+Changes to the other parameters of the function between income and affective wellbeing also occurred between 2001 and 2019 (i.e., the pre-slope, post-slope and intercept), but did not show any sustained trend over the period.
 
-Both the pre-slope and post-slope parameters (&beta;~1~ and &beta;~2~) were credibly larger than zero in each year, indicating there was a reliable dependency between happiness and income at income levels both below and above the change point.
-However the pre-slope parameter was in general four to five times greater than the post-slope values, indicating that the relationship between happiness and income among people in lower income households was an order of magnitude stronger than those in high income households.
+Both the pre-slope and post-slope parameters (&beta;~1~ and &beta;~2~) were credibly larger than zero in each year, indicating there was a reliable dependency between affective wellbeing and income at income levels both below and above the change point.
+However the pre-slope parameter was in general four to five times greater than the post-slope values, indicating that the relationship between affective wellbeing and income among people in lower income households was an order of magnitude stronger than those in high income households.
 
@@
 
-The change in parameter values between 2001 to 2019 indicates the relationship between happiness and income evolved over time.
-We determined the impact of this evolution on the distribution of happiness over the range of household income in 2019 in a counterfactual analysis.
-The counterfactual analysis is a hypothetical demonstration of how happiness would change if people in 2019 were subject to the function that existed in 2001, i.e., would happiness increase or decrease if the 2001 function was in place in 2019?
-This controls for changes in the sample which occur over time that are not related to happiness but could nevertheless contribute to changes in the distribution of happiness.
-For example, an increase in the range of (real) income levels in the economy between 2001 and 2019 could produce an increasing gap in happiness between the rich and poor - even with a stable relationship between income and happiness.
-Such changes in the sample characteristics may mask or confound the impact of the change point on the distribution of happiness without careful control.
-Because we were interested in the implications of the evolution of the function rather than changes in our sample characteristics _per se_, we estimated happiness levels for each person in 2019 (*n* = 14,459) using the 2001 function.
-These 2001 model-estimates were compared to (subtracted from) the 2019 model-estimates generated from the same sample (*n* = 14,459), to obtain the change (delta) in happiness for each person under the counterfactual.
+The change in parameter values between 2001 to 2019 indicates the relationship between affective wellbeing and income evolved over time.
+We determined the impact of this evolution on the distribution of affective wellbeing over the range of household income in 2019 in a counterfactual analysis.
+The counterfactual analysis is a hypothetical demonstration of how affective wellbeing would change if people in 2019 were subject to the function that existed in 2001, i.e., would affective wellbeing increase or decrease if the 2001 function was in place in 2019?
+This controls for changes in the sample which occur over time that are not related to affective wellbeing but could nevertheless contribute to changes in the distribution of affective wellbeing.
+For example, an increase in the range of (real) income levels in the economy between 2001 and 2019 could produce an increasing gap in affective wellbeing between the rich and poor - even with a stable relationship between income and affective wellbeing.
+Such changes in the sample characteristics may mask or confound the impact of the change point on the distribution of affective wellbeing without careful control.
+Because we were interested in the implications of the evolution of the function rather than changes in our sample characteristics _per se_, we estimated affective wellbeing levels for each person in 2019 (*n* = 14,459) using the 2001 function.
+These 2001 model-estimates were compared to (subtracted from) the 2019 model-estimates generated from the same sample (*n* = 14,459), to obtain the change (delta) in affective wellbeing for each person under the counterfactual.
 This delta is attributable to the evolution of the function between 2001 and 2019.
-Figure 3 below presents the 14,459 deltas from such a comparison, along with a smoothed mean (solid line) to summarize how the distribution of happiness evolved across the income range.
+Figure 3 below presents the 14,459 deltas from such a comparison, along with a smoothed mean (solid line) to summarize how the distribution of affective wellbeing evolved across the income range.
 
-##### Figure 3. Counterfactual difference in happiness between 2019 and 2001 for a given level of income
+##### Figure 3. Counterfactual difference in affective wellbeing between 2019 and 2001 for a given level of income
 ```{r counterfactual_estimates}
@@
   labs(
-    subtitle = "Happiness delta (SD units)",
+    subtitle = "Affective delta (SD units)",
     x = "Household income ($000s)", y = "",
-    caption = "Source: Predicted happiness scores from 2019 HILDA sample (n = 14,459)"
+    caption = "Source: Predicted affective wellbeing scores from 2019 HILDA sample (n = 14,459)"
   ) +
@@
 ```
-Figure 3 legend: The difference (∆) in predicted happiness between 2019 and 2001 for the same n = 14,459 individuals.
-Values below zero on the y-axis indicate lower happiness predictions using the 2019 model compared to using the 2001 model.
-The smoothed overlay (solid line) indicates how the average happiness changes across the income distribution.
+Figure 3 legend: The difference (∆) in predicted affective wellbeing between 2019 and 2001 for the same n = 14,459 individuals.
+Values below zero on the y-axis indicate lower affective wellbeing predictions using the 2019 model compared to using the 2001 model.
+The smoothed overlay (solid line) indicates how the average affective wellbeing changes across the income distribution.
 ```
@@
 
-Figure 3 shows that, on average, predicted happiness is higher for people with household incomes above \$50K under the 2019 function, when compared to the function from 2001.
+Figure 3 shows that, on average, predicted affective wellbeing is higher for people with household incomes above \$50K under the 2019 function, when compared to the function from 2001.
 This is indicated by the average delta (solid line) falling above zero on the right side of the plot.
-Conversely, people with household incomes below \$50K, on average, had a decrease in predicted happiness under the 2019 function compared to 2001 function.
+Conversely, people with household incomes below \$50K, on average, had a decrease in predicted affective wellbeing under the 2019 function compared to 2001 function.
 Of course the obtained deltas are due to changes in all the parameters of the function, including the slope before and after the change point.
-However because this comparison was performed on the same individuals from the 2019 survey, it held characteristics such as age, income, etc, constant that would otherwise be expected to change over time and possibly contribute to any difference in happiness distribution.
-In this way these results isolate the amount of change entirely due to the _evolution_ of the function between 2001 and 2019, and demonstrates how this has contributed to a more unequal distribution of happiness between the rich and the poor over time in Australia.
+However because this comparison was performed on the same individuals from the 2019 survey, it held characteristics such as age, income, etc, constant that would otherwise be expected to change over time and possibly contribute to any difference in affective wellbeing distribution.
+In this way these results isolate the amount of change entirely due to the _evolution_ of the function between 2001 and 2019, and demonstrates how this has contributed to a more unequal distribution of affective wellbeing between the rich and the poor over time in Australia.
 
@@
 
-Any increase in the change point is likely to reduce the number of people who fall above it over time if this is greater than any increase in median income; i.e., over time a larger proportion of the population's happiness is responsive to marginal changes in their income than previously.
+Any increase in the change point is likely to reduce the number of people who fall above it over time if this is greater than any increase in median income; i.e., over time a larger proportion of the population's affective wellbeing is responsive to marginal changes in their income than previously.
 Figure 4 presents median household income levels weighted for the Australian population (by age, sex, marital status, labour force participation and geographical region).
-This shows the change point between income and happiness increased faster than rises in median household income between 2001 and 2019.
+This shows the change point between income and affective wellbeing increased faster than rises in median household income between 2001 and 2019.
 The third panel shows that as a result, a smaller proportion of the Australian population in 2019 had a household income above the changepoint than in 2001.
@@
 ```
-Figure 4 legend: Real household income has stagnated in Australia since 2009 (post GFC) while the change point between happiness and income has increased.  
+Figure 4 legend: Real household income has stagnated in Australia since 2009 (post GFC) while the change point between affective wellbeing and income has increased.  
 ```
@@
 
-We confirmed previous findings [e.g., @howell2008relation; @kahneman2010high], that the relationship between both types of subjective wellbeing with household income was positive but quite different: Satisfaction increased linearly with income, while happiness increased rapidly up to a point after which higher levels of income were associated with less improvement.
-Our novel finding was that this change point in the relationship between household income and happiness increased faster than both inflation and median household income between 2001 and 2019.
-For the first time we have shown there has been a temporal shift in the change point between income and happiness over the 19 year period, such that happiness has become more dependent on income for more people - especially the poor and middle-class.
+We confirmed previous findings [e.g., @howell2008relation; @kahneman2010high], that the relationship between both types of subjective wellbeing with household income was positive but quite different: Cognitive wellbeing increased linearly with income, while affective wellbeing increased rapidly up to a point after which higher levels of income were associated with less improvement.
+Our novel finding was that this change point in the relationship between household income and affective wellbeing increased faster than both inflation and median household income between 2001 and 2019.
+For the first time we have shown there has been a temporal shift in the change point between income and affective wellbeing over the 19 year period, such that affective wellbeing has become more dependent on income for more people - especially the poor and middle-class.
 
-We refer to the change point after which increases in income no longer produce similar increases in happiness as the _cost of happiness_.
-After this point, happiness is no longer as dependent on household income, and the economic security it represents.
-Presumably after this point further increases in happiness depend more on other life factors (e.g., leisure time, social connections) than financial security.
-Life satisfaction on the other hand appeared to show consistent increases with household income and we found no evidence of any change point.
+We refer to the change point after which increases in income no longer produce similar increases in affective wellbeing as the _cost of happiness_.
+After this point, affective wellbeing is no longer as dependent on household income, and the economic security it represents.
+Presumably after this point further increases in affective wellbeing depend more on other life factors (e.g., leisure time, social connections) than financial security.
+Cognitive wellbeing on the other hand appeared to show consistent increases with household income and we found no evidence of any change point.
 The difference may reflect the importance of a numerical dollar value (e.g., bank balance, house value) when cognitively appraising one’s life achievements, versus the relevance of that number to our everyday experience of joy and our prevailing mood.
 
-An implication of the changing relationship between happiness and income is that income inequality may be driving increasing inequities in wellbeing.
-This is demonstrated in Figure 3, where the difference in predicted happiness between 2001 and 2019 increased for incomes above $50K/year and decreased for incomes below that level (see also Figure S6 in Supplementary materials).
-The inequity was also highlighted in Figure 4 where the change point of happiness represented a 9% increase over median income in 2001, while in 2019 it represented a 42% increase over median income.
+An implication of the changing relationship between affective wellbeing and income is that income inequality may be driving increasing inequities in wellbeing.
+This is demonstrated in Figure 3, where the difference in predicted affective wellbeing between 2001 and 2019 increased for incomes above $50K/year and decreased for incomes below that level (see also Figure S6 in Supplementary materials).
+The inequity was also highlighted in Figure 4 where the change point of affective wellbeing represented a 9% increase over median income in 2001, while in 2019 it represented a 42% increase over median income.
 This increase relative to median income also represented a reduction from 43% to 26% in the proportion of people whose income fell above the changepoint.
-Thus we can see that over the last nineteen years the difference in income-related happiness between the rich and the poor has increased; while the happiness of an increasing proportion of people, including the middle-class, is more dependent on their financial security.
+Thus we can see that over the last nineteen years the difference in income-related affective wellbeing between the rich and the poor has increased; while the affective wellbeing of an increasing proportion of people, including the middle-class, is more dependent on their financial security.
 

### 2021-06-02 src/draft-SSM1.Rmd
Added item 9 questions (closes issue #13)

@@
 The SF-36 is a widely used self-completion measure of various aspects of physical, emotional and mental health [@ware2000sf].
-A subset of 9 questions assess mental health and vitality, with five questions measuring positive and negative aspects of mental health (e.g., "Felt so down in the dumps nothing could cheer me up", "Been happy"), and four questions on positive and negative aspects of vitality (e.g., "feel full of life", "felt worn out").
-The response scale timeframe is the past four weeks and agreement was indicated on a six-point Likert scale.
-We reverse scored negatively phrased questions and calculated the sum of the nine questions so that higher scores represented better wellbeing.
+A subset of 9 questions assess mental health and vitality, with four questions measuring agreement with positive aspects of mental health and vitality (i.e., "Feel full of life", "Felt calm and peaceful", "Have a lot of energy", "Been happy"), and five questions measuring agreement with negative aspects ("Felt so down in the dumps nothing could cheer me up", "Felt worn out", "Been a nervous person", "Felt down", "Felt tired").
+The response scale timeframe was the past four weeks and agreement was indicated on a six-point Likert scale.
+Supplementary analysis on the subset of positive and negative items did not reveal any difference in the pattern of results between them (**insert Supplementary section name here**), and so we reverse scored negatively-phrased questions and calculated the sum of the nine questions so that higher scores represented better wellbeing.
 To aid interpretability, we rescaled the final sum to a score between 1-100, where 100 represents the maximum affective wellbeing achievable.

### 2021-06-06 src/draft-SSM1.Rmd
Added log-linear models. Closes issue #10

@@
 Indeed, the difference between the two questions: “How satisfied are you with your life?” and “How happy are you these days?” has been identified as a crucial mediating factor in a meta-analysis of 111 studies on income and wellbeing [@howell2008relation; also @veenhoven2006rising].
-Results since then have also shown that while both affective and cognitive wellbeing increase with income, the affective returns are decreasing [@killingsworth2021experienced; @sacks2012new].
+Studies since then have also shown that while both affective and cognitive wellbeing increase with income, the affective returns are a decreasing function of income [@killingsworth2021experienced; @sacks2012new].
 
-Fundamentally, the existence of a change point in the relationship between income and affective wellbeing reveals an unacknowledged source of inequality in the distribution of wellbeing (i.e., happiness) in the economy.
+While studies agree that cognitive wellbeing tends to increase linearly, or log-linearly, with income [e.g., @stevenson2013subjective; but see @jebb2018happiness], the exact form of the functional relationship between affective wellbeing and income is contested: some studies show the presence of a changepoint after which increasing income has little further effect [@kahneman2010high; @jebb2018happiness]; other studies have found that affective wellbeing is a constant (albiet decreasing) log-linear function of income [@killingsworth2021experienced; @sacks2012newl]; and indeed at least one recent study has indicated a constant linear relationship [@twenge2020expanding].
+Few studies have formally tested the functional form of the relationship, relying instead upon visual comparison after binning the data over income deciles or quintiles.
+In this regard, arbitrary decisions about bin width may have salient effects on the functional form, which in turn will influence the inferences drawn.
+For this reason, a formal comparison between linear, piecewise-linear and log-linear functions in the relationship between affective wellbeing and income (unbinned) is warranted.
+More fundementally, the existence of a change point in the relationship between income and affective wellbeing would reveal an unacknowledged source of inequality in the distribution of wellbeing (i.e., happiness) in the economy.
 For instance, a change point of USD75,000 in 2008 was substantially more than the US median income of USD52,000 in the same year, indicating that the affective wellbeing of the poorest majority of the US population was tied to marginal changes in income while the affective wellbeing of a richer minority was not.
-Thus the change point represents the dollar value up to which income drives inequities in the distribution of affective wellbeing, such that a lower valued change point represents a more equitable distribution of affective wellbeing in the economy.
+Thus a change point would represent the dollar value up to which income drives inequities in the distribution of affective wellbeing, such that a lower valued change point represents a more equitable distribution of affective wellbeing in the economy.
+
 Inequities in the distribution of wellbeing are increasingly relevant to governments and policy-makers due to the growing recognition that increasing income does not necessarily lead to equal changes in wellbeing [@clark2018four; @frijters2020happy].
 Even prior to COVID-19, the World Gallup Poll has observed that affective wellbeing has decreased over the past decade in western Europe, North America, Australia and New Zealand, despite increases in income in the same countries [@sachs2019world].
-However to date there has been no investigation of whether the relationship between income and affective wellbeing has changed over time which may contribute to these trends.
-In particular has the change point between income and affective wellbeing, and therefore the distribution of happiness between rich and poor, become more or less equitable in the last few decades?
+However to date there has been little investigation of whether the relationship between income and affective wellbeing has changed over time, which may contribute to these trends.
+In particular has the functional relationship between income and affective wellbeing, and therefore the distribution of happiness between rich and poor, become more or less equitable in the last few decades?
 
@@
 
-We used household economic panel data from Australia to provide the first investigation of whether changes in income and wellbeing have shifted the change point between 2001 to 2019.
+We used household economic panel data from Australia to provide the first investigation of whether changes in the functional form between income and wellbeing have shifted between 2001 to 2019.
 HILDA (the Household, Income, and Labour Dynamics in Australia survey) provides a representative sample of households in Australia with detailed measurements of income and subjective wellbeing each year.
-We distinguished between cognitive- and affective wellbeing as different components of subjective wellbeing, and evaluated how each varied with household income over time using piecewise linear regression to determine the change points in each year.
+We distinguished between cognitive- and affective wellbeing as different components of subjective wellbeing, and evaluated how each varied with household income over time using linear, log-linear and piecewise linear regression in each year.
 
@@
 
-We modelled the relationship between income and each wellbeing variable (cognitive and affective) using a log-linear model and a piecewise-linear model with a single change-point as a free parameter estimated from the data.
+We modelled the relationship between income and each wellbeing variable (cognitive and affective) using a linear model, log-linear model, and a piecewise-linear model with a single change-point as a free parameter estimated from the data.
 The piecewise-linear model was chosen as the simplest extension of a linear model which can identify a change point (inflection) in the relationship between wellbeing and income.
-For modelling, income values were log transformed and then both measures of wellbeing and income were rescaled with a mean of zero and a SD of 1 (z-scores) for each year.
+For modelling, both measures of wellbeing and income were rescaled with a mean of zero and a SD of 1 (z-scores) for each year.
+As each year was modelled separately this does not affect the results.
+Importantly, the coefficient for the change point in each year was returned to the original scale and transformed to real 2019 dollars when reported.
 
 _Model Design_  
-We adopted a Bayesian approach for estimating the log-linear and piecewise-linear model in the software Stan [@burkner2017; @rstan2019]. In each case the log-linear model was estimated as:
+We adopted a Bayesian approach for estimating the linear, log-linear and piecewise-linear model in the software Stan [@burkner2017; @rstan2019].
+The log-linear model was estimated after log transforming the income values, and the linear and log-linear model were then both estimated as:
 
@@
 The WAIC represents an approximation of the out-of-sample deviance that converges to the cross-validation accuracy in a large sample, with a penalty for the effective number of parameters (degrees of freedom).
-Thus using the out-of-sample deviance for model selection in combination with regularizing priors in our model design is a dual strategy to reduce overfitting and penalize overfitting.
-As with other deviance metrics, smaller negative WAIC values are better (i.e., indicate more accuracy).
+Thus using the out-of-sample deviance for model comparison in combination with regularizing priors in our model design is a dual strategy to reduce overfitting and penalize overfitting.
+As with other deviance metrics, smaller WAIC values are better (i.e., indicate more accuracy).
 
@@
 
-The nonlinear relationship between affective wellbeing and income (Figure 1, 2nd row) was consistently and negatively inflected (affective wellbeing increased less with income after the change point).
-By contrast the nonlinear relationship between cognitive wellbeing and income shown in the 4th row exhibited negative inflection (2001, 2015, 2019), positive inflection (2005), and no apparent inflection (2010).
+The piecewise relationship between affective wellbeing and income (Figure 1, 2nd row) was consistently and negatively inflected (affective wellbeing increased less with income after the change point).
+By contrast the piecewise relationship between cognitive wellbeing and income shown in the 4th row exhibited negative inflection (2001, 2015, 2019), positive inflection (2005), and no apparent inflection (2010).
 
-The evidence from model selection revealed the nonlinear (piecewise) fit of _affective_-, but not _cognitive_ wellbeing, was superior to a linear fit in each year (Supplementary Materials, Figure S1).
+The evidence from model selection revealed the piecewise-linear fit of _affective_-, but not _cognitive_ wellbeing, was superior to a log-linear or linear fit in each year (Supplementary Materials, Figure S1).
 Overall the model comparison suggested that cognitive- and affective wellbeing have distinct relationships with household income; cognitive wellbeing tends to increase linearly with income, while a change point exists in the relationship between affective wellbeing and household income.

### 2021-06-06 src/supplementary-SSM1.Rmd
Added log-linear models. Closes issue #10

--- 2021-05-27 src/supplementary-SSM0.Rmd
@@
 
+##### Table S1. Sample characteristics  
 ```{r table_s1}
@@
   as_flex_table() %>%
-  set_caption(caption = "Table S1. Demographic characteristics of the sample") %>%
   align(part = "body") %>%
@@
 
-**Linear vs non-linear models.** We compared the out-of-sample evidence for the non-linear (piecewise) models with the linear models of income on wellbeing in each year, using the difference in WAIC deviance scores (i.e., WAIC~linear~ — WAIC~piecewise~).
-Thus a WAIC difference greater than zero indicated evidence for a linear relationship; a WAIC difference less than zero indicated evidence for a nonlinear (piecewise) relationship.
+We compared the out-of-sample evidence for the piecewise-linear models with the log-linear and linear models of income on wellbeing in each year, using the difference in WAIC deviance scores (i.e., WAIC~piecewise~ — WAIC~log~, WAIC~piecewise~ — WAIC~linear~).
+Since smaller deviance scores indicate better fit, a difference in deviance scores less than zero (i.e., negative) indicated evidence for the piecewise-linear model; a difference greater than zero (i.e., positive) indicated evidence for the other model (log-linear or linear).
 
@@
 
-##### Figure S1. Linear model evidence (WAIC~linear~ — WAIC~piecewise~)
-```{r figure_s1, dpi=300}
+##### Figure S1. Model comparisons (WAIC~piecewise~ vs WAIC~linear~, WAIC~log~)
+```{r figure_s1, fig.width=6, fig.height = 4, dpi=300}
 lm_fits = read_rds("../results/lm_gh9_hhoecd_0cov.RDS")
-
 nl_fits = read_rds("../results/gh9_hhoecd_0cov_tight.RDS")
@@
     year = extract_numeric(year),
-    model = if_else(model == ".x", "nl", "lm"),
+    model = if_else(model == ".x", "nl", "piecewise vs. linear"),
     waic_diff = 2 * elpd_diff, # lm disadvantage
     se = 2 * se_diff,
-    wellbeing = "Happiness"
-  ) -> waic_happiness
+    wellbeing = "Affective wellbeing"
+  ) -> waic_happiness_lm
+
+nl_fits_x0 = read_rds("../results/gh9_hhoecd_0cov_tight_x0.RDS")
+log_fits = read_rds("../results/log_gh9_hhoecd_0cov.RDS")
+
+# get waic results
+map2_dfr(.x = nl_fits_x0, 
+         .y = log_fits, 
+         .f = ~{
+           loo_compare(.x, .y, criterion = "waic") %>%
+             as.data.frame() %>%
+             rownames_to_column("model") %>%
+             as_tibble()
+         },
+         .id = "year") %>%
+  filter(elpd_diff != 0) %>%
+  mutate(
+    year = extract_numeric(year),
+    model = if_else(model == ".x", "nl", "piecewise vs. log"),
+    waic_diff = 2 * elpd_diff, # lm disadvantage
+    se = 2 * se_diff,
+    wellbeing = "Affective wellbeing"
+  ) -> waic_happiness_log
 
 nl_fits_losat = read_rds("../results/losat_hhoecd_0cov.RDS")
-
 lm_fits_losat = read_rds("../results/lm_losat_hhoecd_0cov.RDS")
 
-# plot waic results
 map2_dfr(.x = nl_fits_losat, 
@@
     year = extract_numeric(year),
-    model = if_else(model == ".x", "nl", "lm"),
-    waic_diff = if_else(model == "lm", 
+    model = if_else(model == ".x", "nl", "linear"),
+    waic_diff = if_else(model == "linear", 
                         2 * elpd_diff,   # nl advantage
@@
     se = 2 * se_diff,
-    wellbeing = "Satisfaction"
-  ) -> waic_satisfaction
+    wellbeing = "Cognitive wellbeing"
+  ) %>%
+  mutate(model = "piecewise vs. linear") -> waic_satisfaction_lm
 
-bind_rows(waic_happiness, waic_satisfaction) %>% 
-  ggplot(aes(x = waic_diff, y = year, color = wellbeing)) +
+log_fits_losat = read_rds("../results/fits/excl_topcodes/sd/exc_zeros/log_losat_hhoecd_0cov_x0.RDS")
+nl_fits_losat_x0 = read_rds("../results/fits/excl_topcodes/sd/exc_zeros/losat_hhoecd_0cov_tight_x0.RDS")
+
+map2_dfr(.x = nl_fits_losat_x0, 
+         .y = log_fits_losat, 
+         .f = ~{
+           loo_compare(.x, .y, criterion = "waic") %>%
+             as.data.frame() %>%
+             rownames_to_column("model") %>%
+             as_tibble()
+         },
+         .id = "year") %>%
+  filter(elpd_diff != 0) %>%
+  mutate(
+    year = extract_numeric(year),
+    model = if_else(model == ".x", "nl", "piecewise vs. log"),
+    waic_diff = if_else(model == "piecewise vs. log", 
+                        2 * elpd_diff,   # nl advantage
+                        -2 * elpd_diff), # nl advantage
+    se = 2 * se_diff,
+    wellbeing = "Cognitive wellbeing"
+  ) -> waic_satisfaction_log
+
+bind_rows(waic_happiness_lm, 
+          waic_happiness_log,
+          waic_satisfaction_lm, 
+          waic_satisfaction_log) %>%
+  ggplot(aes(x = waic_diff, y = year, color = model)) +
     geom_vline(aes(xintercept = 0), color = "grey90") +
-    geom_pointrange(aes(xmin = waic_diff - se*1.96, xmax = waic_diff + se*1.96)) +
+    geom_pointrange(aes(xmin = waic_diff - se*1.96, xmax = waic_diff + se*1.96),
+                    position = position_dodge(width = .5)) +
+    facet_wrap(~wellbeing) +
     scale_y_continuous(breaks = c(2001, 2007, 2013, 2019)) +
-    ft_theme() +
-    labs(x = "deviance", y = "") +
-    facet_wrap(~wellbeing, ncol = 1) +
-    scale_color_manual(values  = c("#8F2727", "#005b96")) +
-    theme(legend.position = "none")
+    scale_color_manual(values = c("#ffcc80", "#8F2727")) + # "#005b96"
+    ft_theme(base_size = 12) +
+    theme(legend.title = element_blank(),
+          legend.position = "bottom") +
+    labs(x = "deviance", y = "")
 ```
@@
 ```
-Figure S1 legend: Differences in WAIC scores (deviance units) for a linear fit over a piecewise fit for happiness (red) and satisfaction (blue).
+Figure legend: Differences in WAIC scores (deviance units) for a piecewise-linear fit over a log-linear fit (red) or over a linear fit (yellow).
 The filled circle indicates the mean of the difference and the horizontal bars represents the 95% interval.
-A difference below zero (grey vertical line) is support for the piecewise model (a difference above zero is support for the linear model).
+A difference below zero (grey vertical line) represents support for the piecewise model, and a difference above zero is support for the other model.
 ```
@@
 
-The difference in WAIC scores for each model revealed the nonlinear (piecewise) fits of _happiness_, but not _satisfaction_, provided superior out-of-sample accuracy over a linear fit in each year.
-Figure S1 shows the 95% interval of the differences in WAIC scores clearly distinguished the advantage of the nonlinear fits for happiness with no overlap with zero in any year.
-By contrast, the difference in WAIC between the linear and nonlinear models of life satisfaction was not distinguished from zero for most years (with the exception of 2017 and 2008), indicating the (out-of-sample) accuracy of the two models was largely indistinguishable over the 19 years.
-Overall the evidence suggested that household income has a distinct nonlinear relationship with happiness; implying the existence of a change point or difference in slopes at different levels of household income.
-This is consistent with the only meta-analysis conducted in this area, over ten years ago, which concluded the linear relationship between income and life satisfaction is stronger than that with happiness [@howell2008relation].
-It is also worth noting that a linear function between income and life satisfaction is consistent with the majority of  research since then [@stevenson2013subjective].
+The difference in WAIC scores for each model revealed the piecewise-linear fits of affective wellbeing, but not cognitive wellbeing, provided superior out-of-sample accuracy over a log-linear or linear fit in each year.
+Figure S1 shows the differences in WAIC scores clearly distinguished the advantage of the piecewise-linear fits for affective wellbeing with most 95% intervals entirely below zero for each year (exceptions occurred for the piecewise vs.
+log fit in 2004, 2006 and 2008, where the piecewise fit was still superior on average).
+By contrast, the difference in WAIC scores for the cognitive wellbeing models was not as clearly distinguished from zero for most years, since the 95% intervals of the WAIC differences with either the log-linear fit or the linear fit overlapped zero in most years.
+Indeed in the years between 2002 to 2010 the linear fit was superior to the piecewise-linear fit on average, with the mean WAIC difference above zero (positive).
+Overall the evidence suggested that affective wellbeing has a piecewise relationship with household income; implying the existence of a change point or difference in slopes at different levels of household income.
+Evidence of the effect of income on cognitive wellbeing, on the other hand, was more equivocal, albeit with some indication of a linear relationship in the early part of the 21st C.
+Others have also noted a linear relationship between cognitive wellbeing (life satisfaction) and income [@howell2008relation; @stevenson2013subjective], and the results of our formal comparison provides some support for this notion.
 
@@
 ##### Figure S2. Conditional model parameters  
-```{r figure_s2, fig.width = 6, fig.height = 7.5, dpi=300}
+```{r figure_s2, fig.width = 6, fig.height = 8, dpi=300}
 nl_fits = read_rds("../results/gh9_hhoecd_0cov_tight.RDS")
@@
 ```
-Figure S2 legend: Comparison of the conditional parameter estimates (red) and unconditional parameter estimates (grey) in the change point model of income on happiness.
+Figure legend: Comparison of the conditional parameter estimates (red) and unconditional parameter estimates (grey) in the change point model of income on happiness.
 The conditional models included covariates for age, age squared, sex, education, and illness.
@@
 
-<br><br>
-
-#### Additional non-linear fits
-
-**Log-linear models.** The relationship between happiness and income is often described as log-linear [e.g., @kahneman2010high].
-Under this view, similar proportional changes in income or wealth produce similar evaluations of satisfaction or happiness, rather than the absolute change.
-That is, a \$100 increase in wealth will produce a greater impact on someone earning a minimum wage than someone earning many orders of magnitude of that amount.
-Likewise, the piecewise model that we employed can also adopt an increasing monotonic function whose slope reduces with wealth, albeit with a single identifiable change-point rather than a smooth curve.
-However our main thesis is not committed to a function with a single discrete changepoint; only that the slope between income and happiness is unequal at different levels of income, and the function is shifting to the right over time.
-The choice between functions is somewhat arbitrary, since a log-linear function is as easily consistent with our thesis as the piecewise function we described in the main paper.
-To resolve this, we compared the evidence for a log-linear fit against the piecewise fits according to their WAIC scores.
-Thus a WAIC difference less than zero indicated evidence for the piecewise fit over the log-linear fit (i.e., greater out-of-sample accuracy).
-
-<br>
-
-##### Figure S3. Log-linear model evidence (WAIC~log-linear~ — WAIC~piecewise~)
-```{r, figure_s3, fig.height=3.5, dpi=300}
-nl_fits = read_rds("../results/gh9_hhoecd_0cov_tight_x0.RDS")
-
-log_fits = read_rds("../results/log_gh9_hhoecd_0cov.RDS")
-
-# get waic results
-map2_dfr(.x = nl_fits, 
-         .y = log_fits, 
-         .f = ~{
-           loo_compare(.x, .y, criterion = "waic") %>%
-             as.data.frame() %>%
-             rownames_to_column("model") %>%
-             as_tibble()
-         },
-         .id = "year") %>%
-  filter(elpd_diff != 0) %>%
-  mutate(
-    year = extract_numeric(year),
-    model = if_else(model == ".x", "nl", "log-linear"),
-    waic_diff = 2 * elpd_diff, # lm disadvantage
-    se = 2 * se_diff
-  ) %>%
-  ggplot(aes(x = waic_diff, y = year)) +
-  geom_vline(aes(xintercept = 0), color = "grey90") +
-  geom_pointrange(aes(xmin = waic_diff - se*1.96, xmax = waic_diff + se*1.96),
-                 color = "#ffcc80") +
-  scale_y_continuous(breaks = c(2001, 2007, 2013, 2019)) +
-  ft_theme() +
-  labs(x = "deviance", y = "")
-```
-```
-Figure S3 legend: Differences in WAIC scores (deviance units) for a log-linear fit over a piecewise fit for happiness.
-```
-
-<br>
-
-The piecewise fit was superior to a log-linear fit of income on happiness in each year, although the difference was less distinguishable in 2003, 2006, and 2008 where the 95% interval overlapped zero.
-This indicates the evidence for a piecewise model was generally greater than for a log-linear model, however it does not necessarily imply the presence of a discrete change point exists.
-In a Bayesian change point model, the estimated location of the change point in each year is not a single point but is rather a probability distribution of income values that extends over the entire range of household income.
-Importantly the width of the 95% posterior probability distribution in Figure 2 indicates the change point location is likely over a range of values, on average ±$9.15K, which implies a more or less smooth function over that range.
-
-<br><br>
-
-#### Prior and residual checks
-
-We used regularizing Guassian priors centred on zero, with Normal(0, 0.1) for each *&beta;*, and a Normal(0, 0.5) for *&omega;*.
-A plot of the joint distribution between income and wellbeing imposed by the priors alone confirmed there was no apparent slope or change point implied by their regularizing effect.
-
-##### Figure S4. Prior predictive check  
-```{r figure_s4, dpi=300}
-load("../results/prior_predictions.Rdata")
-
-prior_predictions %>%
-  ggplot(aes(x = hh_disp_inc_eq/1000, y = Estimate, group = 1)) +
-  geom_point(size = 0.1, alpha = 0.1) +
-  stat_smooth(color = "#8F2727", size = 0.5, se = F) +
-  xlim(c(0, 300)) +
-  labs(y = "Wellbeing estimate (SD units)", x = "Household income ($000s)") +
-  theme_test()
-```
-```
-Figure S4 legend: The joint distribution between wellbeing and income for a single year, imposed by the priors. The solid blue line indicates a smoothed average (loess).
-```  
-<br><br>
-
-Examination of the residual plots from the piecewise models of happiness confirms the residuals were evenly distributed around zero and there was little indication of heteroscedascity.
-Overall, there was little evidence that the piecewise models of happiness were not appropriate for the data.
-
-<br>
-
-##### Figure S5. Residual plots for piecewise happiness models
-```{r figure_s5, fig.height=9, fig.width=9}
-# https://www.flutterbys.com.au/stats/tut/tut7.2b.html
-# 
-# 
-# residual_fits <- map_dfr(nl_fits, ~{
-#   
-#   d0 <- resid(.x) %>%
-#     as_tibble() %>%
-#     select(resids = Estimate)
-#   
-#   d1 <- fitted(.x) %>%
-#     as_tibble() %>%
-#     select(fitteds = Estimate)
-#   
-#   bind_cols(d0, d1)
-#   
-#   }, .id = "fit") %>%
-#   mutate(year = extract_numeric(fit))
-# 
-# save(residual_fits, file = paste0(here, "results/residual_fits.Rdata"))
-
-load("../results/residual_fits.Rdata")
-
-residual_fits %>%
-  group_by(year) %>%
-  slice_sample(n = 1000) %>% # to reduce overplotting
-  ggplot(aes(x = fitteds, y = resids)) +
-    geom_point(alpha = 0.1, size = 0.1) +
-    facet_wrap(~year, scales = "free") +
-    theme_test()
-```
-
-```
-Figure S5 legend: The joint distribution between residual and fitted values for each year.
-```
+The unconditional parameter values are also presented in Table S2 below (Parameter and fit summary statistics).
 
@@
 There were _N_ = 59,876 total observations.
-4000 samples (post-warmup) were drawn using sampling(NUTS).
-For each parameter, Bulk_ESS and Tail_ESS are effective sample size (ESS) measures, and Rhat is the potential scale reduction factor on split chains.
-In each case the Rhat approached 1, indicating convergance.
-Each ESS _n_ > 300, which is a conservative threshold for estimation.
+For each Bayesian model presented in this report, 4000 samples (post-warmup) were drawn using sampling(NUTS).
 
-<br>
+For all models estimated with RStan presented in this report we confirmed the absence of divergent transitions; the $\hat{R}$ approached 1 indicating convergance; and the effective sample size (ESS) for each parameter was _n_ > 300, which is a conservative threshold for estimation.
 
+Table S2 below shows the fit and summary statistics for each piecewise-linear model fit of affective wellbeing and income in each year.
+Estimates for the intercept, pre-slope and post-slope are in year-wise SD units while the change-point is presented in real 2019 dollars.
+
+##### Table S2. Parameter estimates and fit statistics for piecewise-linear fit of affective wellbeing and income
 ```{r table_s2, ft.align="left"}
@@
 param_key = c(
-  b0_Intercept = "Intercept",
-  b1_Intercept = "Pre-slope",
-  b2_Intercept = "Post-slope",
-  omega_Intercept = "Change-point ($000s)"
+  omega_Intercept = "Change point ($000s)",
+  b0_Intercept = paste0("Intercept (\u03b2", "0)"),
+  b1_Intercept = paste0("Pre-slope (\u03b2", "1)"),
+  b2_Intercept = paste0("Post-slope (\u03b2", "2)")
 )
+
+# param_key = c(
+#   b0_Intercept = "Intercept",
+#   b1_Intercept = "Pre-slope",
+#   b2_Intercept = "Post-slope",
+#   omega_Intercept = "Change-point ($000s)"
+# )
 
@@
      part = "body") %>%
-
   colformat_double(j = 3:6, digits= 3) %>%
   colformat_double(j = 7, digits= 1) %>%
-  set_caption("Table S2. Piecewise parameter estimates and fit statistics for happiness") %>%
-  fontsize(size = 9.5, part = "all") %>%
+  footnote(i = 1, j = 7,
+           value = as_paragraph(
+             c("Rhat is the potential scale reduction factor on split chains")),
+           ref_symbols = c("a"),
+           part = "header") %>%
+  footnote(i = 1, j = c(8:9),
+           value = as_paragraph(
+             c("Bulk_ESS and Tail_ESS are effective sample size (ESS) measures")),
+           ref_symbols = c("b"),
+           part = "header") %>%
+  fontsize(size = 9, part = "all") %>%
   width(j = 1, width = .5) %>%
@@
 
+<br><br>
+
+The coeficients (in SD units) for the piecewise-linear, log-linear and linear Bayesian models are shown in the table below.
+
+##### Table S3. Bayesian coefficients from linear, log-linear and piecewise-linear models  
+```{r table_s3}
+map_df(log_fits, .f = ~{
+  fixef(.x) %>%
+    as_tibble(rownames = "terms") %>%
+    transmute(terms, coef = paste0(round(Estimate, 2), " (±", round(Est.Error, 3), ")")) %>%
+    spread(terms, coef)},
+  .id = "Year"
+  ) %>%
+  rename("b0_log" = 2, "b1_log" = 3) -> coefs_gh9_log
+
+map_df(lm_fits, .f = ~{
+  fixef(.x) %>%
+    as_tibble(rownames = "terms") %>%
+    transmute(terms, coef = paste0(round(Estimate, 2), " (±", round(Est.Error, 3), ")")) %>%
+    spread(terms, coef)},
+  .id = "Year"
+  ) %>%
+  select(Year, "b0_lin" = 3, "b1_lin" = 2) -> coefs_gh9_lm
+
+map_df(nl_fits, .f = ~{
+  fixef(.x) %>%
+    as_tibble(rownames = "terms") %>%
+    transmute(terms, coef = paste0(round(Estimate, 2), " (±", round(Est.Error, 3), ")")) %>%
+    spread(terms, coef)},
+  .id = "Year"
+  ) %>%
+  rename("b0_pw" = 2, "b1_pw" = 3, 
+         "b2_pw" = 4, "Omega" = 5) -> coefs_gh9_pw
+
+coefs_affective <- left_join(coefs_gh9_pw, coefs_gh9_log) %>%
+  left_join(coefs_gh9_lm) %>%
+  mutate(Year = str_remove(Year, "fit"),
+         Wellbeing = "Affective")
+
+map_df(log_fits_losat, .f = ~{
+  fixef(.x) %>%
+    as_tibble(rownames = "terms") %>%
+    transmute(terms, coef = paste0(round(Estimate, 2), " (±", round(Est.Error, 3), ")")) %>%
+    spread(terms, coef)},
+  .id = "Year"
+) %>%
+  rename("b0_log" = 2, "b1_log" = 3) -> coefs_losat_log
+
+map_df(lm_fits_losat, .f = ~{
+  fixef(.x) %>%
+    as_tibble(rownames = "terms") %>%
+    transmute(terms, coef = paste0(round(Estimate, 2), " (±", round(Est.Error, 3), ")")) %>%
+    spread(terms, coef)},
+  .id = "Year"
+) %>%
+  select(Year, "b0_lin" = 3, "b1_lin" = 2) -> coefs_losat_lm
+
+map_df(nl_fits_losat, .f = ~{
+  fixef(.x) %>%
+    as_tibble(rownames = "terms") %>%
+    transmute(terms, coef = paste0(round(Estimate, 2), " (±", round(Est.Error, 3), ")")) %>%
+    spread(terms, coef)},
+  .id = "Year"
+) %>%
+  rename("b0_pw" = 2, "b1_pw" = 3, 
+         "b2_pw" = 4, "Omega" = 5) -> coefs_losat_pw
+
+coefs_cognitive <- left_join(coefs_losat_pw, coefs_losat_log) %>%
+  left_join(coefs_losat_lm) %>%
+  mutate(Year = str_remove(Year, "fit"),
+         Wellbeing = "Cognitive")
+
+bind_rows(coefs_affective, coefs_cognitive) %>%
+  select(Wellbeing, everything()) %>%
+  flextable() %>%
+  merge_v(j = 1) %>%
+  set_header_labels(values = c(b0_pw = paste0("\u03b2", "0"),
+                               b1_pw = paste0("\u03b2", "1"),
+                               b2_pw = paste0("\u03b2", "2"),
+                               Omega = "\u03C9",
+                               b0_log = paste0("\u03b2", "0"),
+                               b1_log = paste0("\u03b2", "1"),
+                               b0_lin = paste0("\u03b2", "0"),
+                               b1_lin = paste0("\u03b2", "1"))) %>%
+  add_header(top = TRUE,
+             values = c(b0_pw = "Piecewise-linear (±se)",
+                        b1_pw = "Piecewise-linear (±se)",
+                        b2_pw = "Piecewise-linear (±se)",
+                        Omega = "Piecewise-linear (±se)",
+                        b0_log = "Log-linear (±se)",
+                        b1_log = "Log-linear (±se)",
+                        b0_lin = "Linear (±se)",
+                        b1_lin = "Linear (±se)")) %>%
+  merge_h(part = "header") %>%
+  width(j = 1:2, width = c(0.75, 0.5)) %>%
+  width(j = 3:10, width = 0.65) %>%
+  fontsize(size = 9, part = "all") %>%
+  # flextable::theme_vanilla() %>% # theme_vanilla() 
+  valign(valign = "top") %>%
+  valign(valign = "bottom", part = "header") %>%
+  border_remove() %>%
+  hline_top(part="header", border = officer::fp_border()) %>%
+  hline_bottom(part="header", border = officer::fp_border()) %>%
+  hline(i = 19, border = officer::fp_border()) %>%
+  hline_bottom(part="body", border = officer::fp_border())
+```
+
+<br><br>
+
+#### Prior and residual checks
+
+We used regularizing Guassian priors centred on zero, with Normal(0, 0.1) for each *&beta;*, and a Normal(0, 0.5) for *&omega;*.
+A plot of the joint distribution between income and wellbeing imposed by the priors alone confirmed there was no apparent slope or change point implied by their regularizing effect.
+
+##### Figure S3. Prior predictive check  
+```{r figure_s3, fig.width=6, fig.height = 6, dpi=300}
+load("../results/prior_predictions.Rdata")
+
+prior_predictions %>%
+  ggplot(aes(x = hh_disp_inc_eq/1000, y = Estimate, group = 1)) +
+  geom_point(size = 0.1, alpha = 0.1) +
+  stat_smooth(color = "#8F2727", size = 0.5, se = F) +
+  xlim(c(0, 300)) +
+  labs(y = "Wellbeing estimate (SD units)", x = "Household income ($000s)") +
+  theme_test()
+```
+```
+Figure legend: The joint distribution between wellbeing and income for a single year, imposed by the priors. The solid blue line indicates a smoothed average (loess).
+```  
+<br><br>
+
+Examination of the residual plots from the piecewise models of happiness confirms the residuals were evenly distributed around zero and there was little indication of heteroscedascity.
+Overall, there was little evidence that the piecewise models of happiness were not appropriate for the data.
+
 <br>
 
-```{r table_s3, ft.align="left"}
-param_key = c(
-  b0_Intercept = "Intercept",
-  b1_Intercept = "Pre-slope",
-  b2_Intercept = "Post-slope",
-  omega_Intercept = "Change-point"
-)
+##### Figure S4. Residual plots for piecewise happiness models
+```{r figure_s4, fig.height=9, fig.width=9, dpi=600}
+# https://www.flutterbys.com.au/stats/tut/tut7.2b.html
+# 
+# 
+# residual_fits <- map_dfr(nl_fits, ~{
+#   
+#   d0 <- resid(.x) %>%
+#     as_tibble() %>%
+#     select(resids = Estimate)
+#   
+#   d1 <- fitted(.x) %>%
+#     as_tibble() %>%
+#     select(fitteds = Estimate)
+#   
+#   bind_cols(d0, d1)
+#   
+#   }, .id = "fit") %>%
+#   mutate(year = extract_numeric(fit))
+# 
+# save(residual_fits, file = paste0(here, "results/residual_fits.Rdata"))
 
-map_dfr(lm_fits_losat, 
-        .f = ~{
-          fit_summary <- summary(.x)
+load("../results/residual_fits.Rdata")
 
-          sigma_summary <- fit_summary$spec_pars %>%
-            as_tibble(rownames = "Parameter")
+residual_fits %>%
+  group_by(year) %>%
+  slice_sample(n = 1000) %>% # to reduce overplotting
+  ggplot(aes(x = fitteds, y = resids)) +
+    geom_point(alpha = 0.1, size = 0.1) +
+    facet_wrap(~year, scales = "free") +
+    theme_test()
+```
 
-          fit_summary$fixed %>%
-            as_tibble(rownames = "Parameter") %>%
-            bind_rows(sigma_summary) %>%
-            select(Parameter, everything())
-
-          },
-        .id = "Year") %>%
-  mutate(
-    Year = str_remove(Year, "fit"),
-    Parameter = recode_factor(Parameter, !!!param_key)
-    ) %>%
-  flextable() %>%
-  merge_v(j = 1) %>%
-  valign(j = 1, valign = "top") %>%
-  bg(i = ~Year %in% seq(2001, 2019, 2), 
-     bg = "grey90", 
-     part = "body") %>%
-  colformat_double(j = 3:6, digits= 3) %>%
-  colformat_double(j = 7, digits= 1) %>%
-  set_caption("Table S3. Linear parameter estimates and fit statistics for life satisfaction") %>%
-  fontsize(size = 9.5, part = "all") %>%
-  autofit()
+```
+Figure legend: The joint distribution between residual and fitted values for each year.
 ```
@@
 
-##### Figure S6. Happiness (SD units) at $50K/yr and $75K/yr from 2001-2019  
-```{r figure_s6, fig.width=7, fig.height=5, dpi=300}
+##### Figure S5. Happiness (SD units) at $50K/yr and $75K/yr from 2001-2019  
+```{r figure_s5, fig.width=7, fig.height=5, dpi=300}
 nl_fits = read_rds("../results/gh9_hhoecd_0cov_tight.RDS")
@@
 
-#### In-sample fit statistics
+## Appendix
+
+#### OLS Insample fit statistics
 
@@
 Nevertheless, some readers may find it useful to compare the in-sample fit statistics between the three different model types employed here: piecewise, linear and log-linear, as well as between the conditional and unconditional models.
-Because the usual fit statistics such as R^2^ present a problem for Bayesian fits, as the variance of the predicted values can be larger than the variance of the data results in R^2^ values greater than 1, we present the in-sample statistics from frequentist OLS model fits (without regularization or penalized likelihoods).
+Because the usual fit statistics such as R^2^ present a problem for Bayesian fits, as the variance of the predicted values can be larger than the variance of the data results in R^2^ values greater than 1, we present the in-sample statistics from OLS model fits (without regularization or penalized likelihoods).
 
-```{r table_s4}
+```{r model_summaries_yearly}
 library(segmented)
@@
     y = as.vector(scale(gh9)),
-    dollars = hh_disp_inc_eq,
+    dollars = hh_disp_inc_eq/10000,
     male = male,
@@
 ) -> model_summaries
+```
+```{r model_summaries_allyears}
+df <- unnest(nested_data, cols = c(data)) %>% ungroup()
 
-model_summaries %>%
-  select(model, Year = year, r.squared) %>%
-  spread(model, r.squared) %>%
+#### Testing log-linear vs piecewise all years ####
+lm_fit = lm(formula = y ~ dollars, 
+             data = df)
+
+lm_fit_c = lm(
+  formula = y ~ dollars + male + age + age_sq + grad + illness + weights,
+  data = df)
+
+log_fit = lm(formula = y ~ log(dollars), 
+             data = df)
+
+log_fit_c = lm(
+  formula = y ~ log(dollars) + male + age + age_sq + grad + illness + weights,
+  data = df)
+
+pw_fit = segmented(lm(formula = y ~ dollars,
+                      data = df),
+                   seg.Z = ~dollars)
+
+pw_fit_c = segmented(lm(
+  formula = y ~ dollars + male + age + age_sq + grad + illness + weights, 
+  data = df),
+  seg.Z = ~dollars)
+
+model_summaries_all <- rbind(
+  broom::glance(lm_fit) %>% mutate(model = "lin", year = "All"),
+  broom::glance(lm_fit_c) %>% mutate(model = "linc", year = "All"),
+  broom::glance(log_fit) %>% mutate(model = "log", year = "All"),
+  broom::glance(log_fit_c) %>% mutate(model = "logc", year = "All"),
+  broom::glance(pw_fit) %>% mutate(model = "pw", year = "All"),
+  broom::glance(pw_fit_c) %>% mutate(model = "pwc", year = "All"),
+  model_summaries
+)
+```
+
+##### Table S4. In-sample fit statistics (adj. R-squared)  
+```{r table_s4}
+model_summaries_all %>%
+  select(model, Year = year, adj.r.squared) %>%
+  spread(model, adj.r.squared) %>%
   flextable() %>%
-  colformat_num(j = 1, big.mark = "") %>%
   colformat_double(j = 2:7, digits = 3) %>%
-  set_caption("Table S4. In-sample fit statistics (R-squared)") %>%
   set_header_labels(
@@
 
+```{r table_s5}
+lm_coef <- map_dfr(nested_models$lin_fit, broom::tidy, .id = "Year") %>%
+  filter(term == "dollars") %>%
+  mutate(coefficient = paste0(round(estimate, 3), " (±", round(std.error, 3), ")")) %>%
+  select(Year, `linear (±se)` = coefficient)
+
+log_coef <- map_dfr(nested_models$log_fit, broom::tidy, .id = "Year") %>%
+  filter(term == "log(dollars)") %>%
+  mutate(coefficient = paste0(round(estimate, 3), " (±", round(std.error, 3), ")")) %>%
+  select(Year, `log-linear (±se)` = coefficient)
+
+pw_coefs <- map_dfr(nested_models$pw_fit, broom::tidy, .id = "Year") %>%
+  filter(term %in% c("dollars", "U1.dollars")) %>%
+  select(Year, term, estimate, std.error) %>%
+  pivot_wider(id_cols = Year, 
+              names_from = term, 
+              values_from = estimate:std.error) %>%
+  mutate(
+    estimate_post = estimate_U1.dollars + estimate_dollars,
+    `pre-slope (±se)` = paste0(
+      round(estimate_dollars, 3), " ±", round(std.error_dollars, 3)),
+    `post-slope (±se)` = paste0(
+      round(estimate_post, 3), " ±", round(std.error_U1.dollars, 3))
+    ) %>%
+  select(Year, `pre-slope (±se)`, `post-slope (±se)`)
+
+pw_coefs <- map_dfr(nested_models$pw_fit, ~as_tibble(.x$psi), .id = "Year") %>%
+  mutate(term = "psi.dollars") %>%
+  mutate(coefficient = paste0(round(`Est.`, 2)*10, " (±", round(`St.Err`, 2)*10, ")")) %>%
+  select(Year, `change point ($000s ±se)` = coefficient) %>%
+  left_join(pw_coefs)
+
+left_join(lm_coef, log_coef, by = "Year") %>%
+  left_join(pw_coefs, by = "Year") %>%
+  mutate(Year = as.character(2000 + as.numeric(Year))) %>%
+  flextable() %>%
+  set_caption("OLS coefficients from linear, log-linear and piecewise models") %>%
+  autofit()
+```
+
+<br><br>
+
 ## References

### 2021-06-07 src/draft-SSM1.Rmd
Added validation of wellbeing measures. Closes issue #11 and #24

@@
 Affective wellbeing was determined by 9 questions in the SF-36 [(9a to 9i)](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=gh9a).
-The SF-36 is a widely used self-completion measure of various aspects of physical, emotional and mental health [@ware2000sf].
+The SF-36 is a widely used self-completion measure of various aspects of physical, emotional and mental health [@ware2000sf], which has been validated in the Australian HILDA sample as a measure of health inequality [@butterworth2004validity].
 A subset of 9 questions assess mental health and vitality, with four questions measuring agreement with positive aspects of mental health and vitality (i.e., "Feel full of life", "Felt calm and peaceful", "Have a lot of energy", "Been happy"), and five questions measuring agreement with negative aspects ("Felt so down in the dumps nothing could cheer me up", "Felt worn out", "Been a nervous person", "Felt down", "Felt tired").
 The response scale timeframe was the past four weeks and agreement was indicated on a six-point Likert scale.
-Supplementary analysis on the subset of positive and negative items did not reveal any difference in the pattern of results between them (**insert Supplementary section name here**), and so we reverse scored negatively-phrased questions and calculated the sum of the nine questions so that higher scores represented better wellbeing.
+We have previously shown a summed score of these 9 questions distinguishes the impact of good and bad major life events in a bidirectional manner [@kettlewell2020differential].
+Supplementary analysis on the subset of positive and negative questions did not reveal any difference in the pattern of results between them (Supplementary materials, Figure S3), and so we reverse scored negatively-phrased questions and calculated the sum of the nine questions so that higher scores represented better wellbeing.
 To aid interpretability, we rescaled the final sum to a score between 1-100, where 100 represents the maximum affective wellbeing achievable.
@@
 Cognitive wellbeing on the other hand appeared to show consistent increases with household income and we found no evidence of any change point.
+We have previously observed that cognitive- and affective-wellbeing appear to track distinct responses to financial gains and losses [@kettlewell2020differential], such that a major financial windfall produced less benefit to affective-wellbeing than cognitive-wellbeing whilst a major financial loss (e.g., bankruptcy) produced equal effects on both.
 The difference may reflect the importance of a numerical dollar value (e.g., bank balance, house value) when cognitively appraising one’s life achievements, versus the relevance of that number to our everyday experience of joy and our prevailing mood.
@@
 Thus we can see that over the last nineteen years the difference in income-related affective wellbeing between the rich and the poor has increased; while the affective wellbeing of an increasing proportion of people, including the middle-class, is more dependent on their financial security.
+
+#### 4.1 Limitations of the present study. 
+The amount of variance in wellbeing explained by income was small.
+While the amount of variance was small, the pre-change slope was reliably above zero and small changes in happiness over an entire population will aggregate.
+Moreover, financial gains and losses have one of the largest impacts on subjective wellbeing in within-subject/fixed-effect models [e.g., @kettlewell2020differential], and there are many reasons why cross-sectional estimates of the covariance would under-represent this.
+Another limitation is that wellbeing is measured on a subjective scale, and we do not know whether small changes at different points on the scale represent larger changes in some other functional/real-world outcome (e.g., risk of suicide).
+At present, the real-world impact of changes in a subjective measure such as wellbeing is difficult to determine.
 

### 2021-06-07 src/supplementary-SSM1.Rmd
Added validation of wellbeing measures. Closes issue #11 and #24

@@
 ```
-Figure legend: Comparison of the conditional parameter estimates (red) and unconditional parameter estimates (grey) in the change point model of income on happiness.
+Figure legend: Comparison of the conditional parameter estimates (red) and unconditional parameter estimates (grey) in the piecewise-linear model of income on happiness.
 The conditional models included covariates for age, age squared, sex, education, and illness.
@@
 The unconditional parameter values are also presented in Table S2 below (Parameter and fit summary statistics).
+
+<br><br>
+
+**Positive and negative components of affective wellbeing**.
+Affective wellbeing includes both positive and negative components, and both these components are assessed by multiple quesstions in the SF-36 item 9.
+To determine whether the increasing change-point in income on affective wellbeing we observed was due to positive or negative affect, we fit a piecewise-linear model of income on positive and negative affect separately.
+The four questions dealing with positive affect (i.e., "Feel full of life", "Felt calm and peaceful", "Have a lot of energy", "Been happy") and the five negative affect questions ("Felt so down in the dumps nothing could cheer me up", "Felt worn out", "Been a nervous person", "Felt down", "Felt tired") were summed together separately to create distinct scores for positive and negative affect.
+The negative affect score was reversed so that higher scoes indicated less negative affect.
+Household income was then regressed against each score using the piecewise linear model (Equation 3), as described for the overall affective wellbeing score in the main text.
+Figure S3 below shows that the change-point in income increased over the period 2001-2019 for both positive (pink) and negative (blue) affect, indicating a similar trend in both components of affective wellbeing over the 19 years.
+
+##### Figure S3. Positive and negative affect parameters  
+```{r figure_s3, fig.width=6, fig.height=8, dpi=300}
+gh9neg.fits <- read_rds("../results/fits/excl_topcodes/sd/gh9neg_hhoecd_0cov_tight.RDS")
+gh9pos.fits <- read_rds("../results/fits/excl_topcodes/sd/gh9pos_hhoecd_0cov_tight.RDS")
+
+income_params <- read_rds("../results/gh9_hhoecd_summary.RDS") %>%
+  bind_rows(.id = "fit")
+
+posteriors_pos <- map_dfr(gh9pos.fits, as.data.frame, .id = "fit") %>%
+  left_join(income_params) %>%
+  mutate(omega_dollars = (b_omega_Intercept * sd + mean)/1000) %>%
+  as_tibble()
+
+param_key = c(
+  omega_dollars = "Change point ($000s)",
+  b_b0_Intercept = paste0("Intercept (\u03b2", "0)"),
+  b_b1_Intercept = paste0("Pre-slope (\u03b2", "1)"),
+  b_b2_Intercept = paste0("Post-slope (\u03b2", "2)")
+)
+
+posteriors_pos %>%
+  select(fit, contains("b0"), contains("b1"), contains("b2"), omega_dollars) %>%
+  gather(key = "param", val = "sample", -fit) %>%
+  mutate(
+    year = extract_numeric(fit),
+    param = recode_factor(param, !!!param_key)
+  ) %>% 
+  group_by(param, year) %>%
+  summarise(
+    low = quantile(sample, probs = 0.05, na.rm=T),
+    expected = quantile(sample, probs = 0.5, na.rm=T),
+    upp = quantile(sample, probs = 0.95, na.rm=T)
+  ) %>%
+  mutate(model = "positive") -> parameter_summary_pos
+
+posteriors_neg <- map_dfr(gh9neg.fits, as.data.frame, .id = "fit") %>%
+  left_join(income_params) %>%
+  mutate(omega_dollars = (b_omega_Intercept * sd + mean)/1000) %>%
+  as_tibble()
+
+posteriors_neg %>%
+  select(fit, contains("b0"), contains("b1"), contains("b2"), omega_dollars) %>%
+  gather(key = "param", val = "sample", -fit) %>%
+  mutate(
+    year = extract_numeric(fit),
+    param = recode_factor(param, !!!param_key)
+  ) %>% 
+  group_by(param, year) %>%
+  summarise(
+    low = quantile(sample, probs = 0.05, na.rm=T),
+    expected = quantile(sample, probs = 0.5, na.rm=T),
+    upp = quantile(sample, probs = 0.95, na.rm=T)
+  ) %>%
+  mutate(model = "negative") -> parameter_summary_neg
+
+bind_rows(parameter_summary_pos, parameter_summary_neg) %>%
+  ggplot(aes(y = year, x = expected, color = model)) +
+  geom_pointrange(aes(xmin = low, xmax = upp), position = position_dodge(width = 0.75)) +
+  scale_y_continuous(breaks = c(2001, 2007, 2013, 2019)) +
+  facet_wrap(~param, scales = "free_x") +
+  ft_theme(base_size = 12) +
+  scale_color_manual(values = c("steelblue", "salmon")) +
+  theme(legend.position = "none",
+        axis.title = element_blank())
+```
+```
+Figure legend: Comparison of the parameter estimates from positive items (pink) and negative items (blue) in the piecewise-linear model of income on affective wellbeing.
+The positive model regressed income against a sum of positive affect questions in the SF-36, while the negative model regressed income against a sum of negative affect questions (reversed scaled to match the positive affect score).
+Horizontal bar represents the 95% credible region and the solid point indicates the expected value (median) of each distribution.
+```
+
+<br>
+
+Regressing income against the positive and negative components of affective wellbeing separately did not reveal substantially different trends for each.
+Both showed an increase in change-point income value over the 19 years, which by 2019 was credibly different from the base values in 2001.
+Changes in the other slope and intercept parameters were either not substantial or not credibly different from the 2001 values according to the 95% credible intervals.
 
@@
 
-##### Figure S3. Prior predictive check  
-```{r figure_s3, fig.width=6, fig.height = 6, dpi=300}
+##### Figure S4. Prior predictive check  
+```{r figure_s4, fig.width=6, fig.height = 6, dpi=300}
 load("../results/prior_predictions.Rdata")
@@
 
-##### Figure S4. Residual plots for piecewise happiness models
-```{r figure_s4, fig.height=9, fig.width=9, dpi=600}
+##### Figure S5. Residual plots for piecewise happiness models
+```{r figure_s5, fig.height=9, fig.width=9, dpi=600}
 # https://www.flutterbys.com.au/stats/tut/tut7.2b.html
@@
 
-##### Figure S5. Happiness (SD units) at $50K/yr and $75K/yr from 2001-2019  
-```{r figure_s5, fig.width=7, fig.height=5, dpi=300}
+##### Figure S6. Happiness (SD units) at $50K/yr and $75K/yr from 2001-2019  
+```{r figure_s6, fig.width=7, fig.height=5, dpi=300}
 nl_fits = read_rds("../results/gh9_hhoecd_0cov_tight.RDS")
@@
 model_summaries_all <- rbind(
-  broom::glance(lm_fit) %>% mutate(model = "lin", year = "All"),
-  broom::glance(lm_fit_c) %>% mutate(model = "linc", year = "All"),
-  broom::glance(log_fit) %>% mutate(model = "log", year = "All"),
-  broom::glance(log_fit_c) %>% mutate(model = "logc", year = "All"),
-  broom::glance(pw_fit) %>% mutate(model = "pw", year = "All"),
-  broom::glance(pw_fit_c) %>% mutate(model = "pwc", year = "All"),
+  broom::glance(lm_fit) %>% mutate(model = "lin", year = "Overall"),
+  broom::glance(lm_fit_c) %>% mutate(model = "linc", year = "Overall"),
+  broom::glance(log_fit) %>% mutate(model = "log", year = "Overall"),
+  broom::glance(log_fit_c) %>% mutate(model = "logc", year = "Overall"),
+  broom::glance(pw_fit) %>% mutate(model = "pw", year = "Overall"),
+  broom::glance(pw_fit_c) %>% mutate(model = "pwc", year = "Overall"),
   model_summaries
@@
   flextable() %>%
+  fontsize(size = 9.5, part = "all") %>%
   set_caption("OLS coefficients from linear, log-linear and piecewise models") %>%

### 2021-06-07 src/draft-SSM1.Rmd
Results are compared to previous research in discussion (closes issue #19)

@@
 
-#### 4.1 Limitations of the present study. 
+Other researchers have argued income has a satiety point on affective well-being, after which increasing income has no further effect [@kahneman2010high; @jebb2018happiness].
+This is consistent with a change point after which the slope is zero.
+While we observed a change-point, the slope afterwards was small but credibly larger than zero so our results are not strictly consistent with a "satiety" point.
+Our results emphasize the slope over income is not constant.
+Other researchers have argued that affective wellbeing is a constant (albiet decreasing) log-linear function of income [@killingsworth2021experienced; @sacks2012newl] or a constant linear relationship [@twenge2020expanding], however these studies typically used a binned measure of income which may reduce the sensitivity to change-points in the final functional form.
+
+<br>
+
+#### 4.1 Limitations of the present study  
 The amount of variance in wellbeing explained by income was small.
@@
 
+<br>
+
+#### 4.2 Conclusions  
 Australia has low levels of income disparity relative to many other OECD countries, and the Gini coefficient has changed little between 2001 and 2019 [@australian2018rising], suggesting income inequality has remained steady over this time period.

### 2021-06-07 src/draft-SSM1.Rmd
Discussion of Twenge added (closes issue #16)

@@
 Our results emphasize the slope over income is not constant.
-Other researchers have argued that affective wellbeing is a constant (albiet decreasing) log-linear function of income [@killingsworth2021experienced; @sacks2012newl] or a constant linear relationship [@twenge2020expanding], however these studies typically used a binned measure of income which may reduce the sensitivity to change-points in the final functional form.
+Other researchers have argued that affective wellbeing is a constant (albiet decreasing) log-linear function of income [@killingsworth2021experienced; @sacks2012newl] or a constant linear relationship [@twenge2020expanding].
+We also show the change-point has increased over time.
+Indeed Twenge et al report the linear relationship between income and happiness has become steeper over time in the US General Social Survey (GSS).
+While we did not see any consistent trend over time in pre or post-slope changes in the piecewise-linear model of affective-wellbeing (e.g., Figure 2), there was a credibly increasing slope in the linear model of cognitive-wellbeing with time (Table S3).
+It may be the single-item on general happiness in the GSS has more in common with the single item on life-satisfaction in HILDA than the more comprehensive 9-item, dual-component questions on affective-wellbeing in the SF-36.
+Nevertheless, like Twenge et al, we report that the difference in happiness betweeen the rich and poor has been increasing over time (e.g., Figure S6), however our piecewise-model reveals the income level around which this difference occurs.
 

### 2021-06-10 src/draft-SSM1.Rmd
Updates to text and Table S4. Closes issue #14 and issue #20

@@
 Subjective wellbeing is not a unitary entity [@diener2017findings]; studies typically distinguish between life satisfaction, the cognitive appraisal of one’s own accomplishments, and affective wellbeing, one’s prevailing affective state, emotional mood, or everyday experience of happiness [e.g., @kettlewell2020differential].
-Kahneman & Deaton [-@kahneman2010high] showed that affective wellbeing increased with household income up to a point (USD75,000), but after that, increasing income had little further effect.
-Conversely cognitive wellbeing continued to increase with income beyond USD75,000.
-Indeed, the difference between the two questions: “How satisfied are you with your life?” and “How happy are you these days?” has been identified as a crucial mediating factor in a meta-analysis of 111 studies on income and wellbeing [@howell2008relation; also @veenhoven2006rising].
-Studies since then have also shown that while both affective and cognitive wellbeing increase with income, the affective returns are a decreasing function of income [@killingsworth2021experienced; @sacks2012new].
+While measures of these distinct concepts are correlated, the distinction seems to be critical in understanding the impact of income on wellbeing [@howell2008relation].
+For instance, we have recently observed that a major financial gain substantially improved cognitive wellbeing for individuals without much impact on their affective wellbeing [@kettlewell2020differential].
+In a now classic paper, Kahneman & Deaton [-@kahneman2010high] showed that affective wellbeing increased with log household income up to a point (USD75,000), but after that, increasing income had no further effect (i.e., "satiation").
+Conversely cognitive wellbeing continued to increase with (log) income beyond USD75,000.
+Studies since then have confirmed that while both affective and cognitive wellbeing increase with income, income is more strongly related to cognitive- than affective-wellbeing [@diener2013rising; @diener2017findings; @howell2008relation].
 
-While studies agree that cognitive wellbeing tends to increase linearly, or log-linearly, with income [e.g., @stevenson2013subjective; but see @jebb2018happiness], the exact form of the functional relationship between affective wellbeing and income is contested: some studies show the presence of a changepoint after which increasing income has little further effect [@kahneman2010high; @jebb2018happiness]; other studies have found that affective wellbeing is a constant (albiet decreasing) log-linear function of income [@killingsworth2021experienced; @sacks2012newl]; and indeed at least one recent study has indicated a constant linear relationship [@twenge2020expanding].
-Few studies have formally tested the functional form of the relationship, relying instead upon visual comparison after binning the data over income deciles or quintiles.
-In this regard, arbitrary decisions about bin width may have salient effects on the functional form, which in turn will influence the inferences drawn.
-For this reason, a formal comparison between linear, piecewise-linear and log-linear functions in the relationship between affective wellbeing and income (unbinned) is warranted.
-More fundementally, the existence of a change point in the relationship between income and affective wellbeing would reveal an unacknowledged source of inequality in the distribution of wellbeing (i.e., happiness) in the economy.
-For instance, a change point of USD75,000 in 2008 was substantially more than the US median income of USD52,000 in the same year, indicating that the affective wellbeing of the poorest majority of the US population was tied to marginal changes in income while the affective wellbeing of a richer minority was not.
-Thus a change point would represent the dollar value up to which income drives inequities in the distribution of affective wellbeing, such that a lower valued change point represents a more equitable distribution of affective wellbeing in the economy.
+Fundementally, a difference in slope between affective wellbeing and income for the rich and poor represents an unacknowledged source of inequality in the distribution of wellbeing (i.e., happiness) in the economy.
+For instance, the "satiety" point of USD75,000 in 2008 was substantially more than the US median income of USD52,000 in the same year, indicating that the affective wellbeing of the poorest majority of the US population was tied to marginal changes in income while the affective wellbeing of a richer minority was not.
+While much commentary has focused on the zero slope at high income levels reported by Kahneman & Diener ("satiety" point), any decrease in slope for the richer minority would also represent a relatively unequal distribution of happiness among the poor compared to the rich.
+Thus, even a non-zero changepoint indicates that income is tied to inequities in the distribution of affective wellbeing in the economy.
+Conversely, a more equitable distribution of wellbeing in the economy would correspond to a lower valued change point.
+
+While studies agree that affective wellbeing tends to increase with income, the exact form of the functional relationship between affective wellbeing and income is contested: some studies show the presence of a zero-slope changepoint after which increasing income has no further effect [@kahneman2010high; @jebb2018happiness]; other studies have found that affective wellbeing is a constant (albiet decreasing) log-linear function of income [@killingsworth2021experienced; @sacks2012new; @twenge2020expanding].
+Few studies have formally tested the functional form of the relationship, relying instead upon binning the data (i.e., over income deciles or quintiles) and comparing adjacent bins either visually or with null hypothesis tests which do not provide evidence for a changepoint.
+In this regard, arbitrary choices about bin width can have salient effects on the function form, which in turn will influence the inferences drawn.
+And no studies to our knowledge have ever tested the presence of a changepoint between wellbeing and income using standard modelling approaches.
+For this reason, a formal comparison between log-linear, linear and a changepoint function (piecewise-linear) in the relationship between affective wellbeing and income is warranted.
 
@@
 The evidence from model selection revealed the piecewise-linear fit of _affective_-, but not _cognitive_ wellbeing, was superior to a log-linear or linear fit in each year (Supplementary Materials, Figure S1).
-Overall the model comparison suggested that cognitive- and affective wellbeing have distinct relationships with household income; cognitive wellbeing tends to increase linearly with income, while a change point exists in the relationship between affective wellbeing and household income.
+Overall the model comparison suggested that affective- and cognitive wellbeing have distinct relationships with household income; affective wellbeing increases with income more rapidly at lower household income levels than higher income levels with a distinct changepoint in each year, while cognitive wellbeing tends to increase with income, most likely in a linear fashion.
+Others have also noted cognitive wellbeing (life satisfaction) has a stronger linear relationship with income, than affective wellbeing and income [@howell2008relation], and the results of our formal comparison provides some support for this notion.
 
@@
 
-We confirmed previous findings [e.g., @howell2008relation; @kahneman2010high], that the relationship between both types of subjective wellbeing with household income was positive but quite different: Cognitive wellbeing increased linearly with income, while affective wellbeing increased rapidly up to a point after which higher levels of income were associated with less improvement.
-Our novel finding was that this change point in the relationship between household income and affective wellbeing increased faster than both inflation and median household income between 2001 and 2019.
+We found the marginal effect of household income on affective- and cognitive-wellbeing was positive but quite different: Affective wellbeing increased rapidly up to a point after which higher levels of income were associated with less improvement, while cognitive wellbeing tended to increase linearly or constantly with income.
+Other studies have also reported that income has distinct effects on cognitive and affective wellbeing [@howell2008relation; @kahneman2010high], with a change point in affective wellbeing after which increasing income has little further effect.
+Our novel finding was revealing the change point increased faster than both inflation and median household income between 2001 and 2019.
 For the first time we have shown there has been a temporal shift in the change point between income and affective wellbeing over the 19 year period, such that affective wellbeing has become more dependent on income for more people - especially the poor and middle-class.

### 2021-06-10 src/supplementary-SSM1.Rmd
Updates to text and Table S4. Closes issue #14 and issue #20

@@
 Evidence of the effect of income on cognitive wellbeing, on the other hand, was more equivocal, albeit with some indication of a linear relationship in the early part of the 21st C.
-Others have also noted a linear relationship between cognitive wellbeing (life satisfaction) and income [@howell2008relation; @stevenson2013subjective], and the results of our formal comparison provides some support for this notion.
 
@@
 
-##### Table S4. In-sample fit statistics (adj. R-squared)  
+##### Table S4. In-sample fit statistics (adj. R-squared) for affective wellbeing and income 
 ```{r table_s4}
@@
 
-A comparison of the relative differences in R^2^ values between models indicates the piecewise model produced a slightly higher in-sample fit over the other two models, while the linear model tended to have the lowest R^2^.
+A comparison of the relative differences in adjusted R^2^ values between models indicates the piecewise model produced a slightly higher in-sample fit over the other two models, while the linear model tended to have the lowest adjusted R^2^.
 The log-linear fit varied between both over the 19 years.
-As expected, adding covariates increased the R-squared in each case, but it did not change the relative difference between models.
+As expected, adding covariates increased the adjusted R-squared in each case, but it did not change the relative difference between models.
 

### 2021-06-10 src/draft-SSM1.Rmd
Edits to discussion. Closes issue #22 #18

@@
 
-Both the pre-slope and post-slope parameters (&beta;~1~ and &beta;~2~) were credibly larger than zero in each year, indicating there was a reliable dependency between affective wellbeing and income at income levels both below and above the change point.
+Both the pre-slope and post-slope parameters (&beta;~1~ and &beta;~2~) were credibly larger than zero in each year, indicating there was a reliable dependency between affective wellbeing and income at income levels both below and above the changepoint.
 However the pre-slope parameter was in general four to five times greater than the post-slope values, indicating that the relationship between affective wellbeing and income among people in lower income households was an order of magnitude stronger than those in high income households.
@@
 We found the marginal effect of household income on affective- and cognitive-wellbeing was positive but quite different: Affective wellbeing increased rapidly up to a point after which higher levels of income were associated with less improvement, while cognitive wellbeing tended to increase linearly or constantly with income.
-Other studies have also reported that income has distinct effects on cognitive and affective wellbeing [@howell2008relation; @kahneman2010high], with a change point in affective wellbeing after which increasing income has little further effect.
+Other studies have also reported that income has distinct effects on cognitive and affective wellbeing [@howell2008relation; @kahneman2010high], with a zero-slope change point in affective wellbeing after which increasing income has little further effect.
+In our case the changepoint did not precede a zero-slope, as the post-slope for each year was credibly above zero (Figure 2), but it did represent a substantial decrease in dependency between income and affective wellbeing for those higher income levels.
 Our novel finding was revealing the change point increased faster than both inflation and median household income between 2001 and 2019.
@@
 This is demonstrated in Figure 3, where the difference in predicted affective wellbeing between 2001 and 2019 increased for incomes above $50K/year and decreased for incomes below that level (see also Figure S6 in Supplementary materials).
-The inequity was also highlighted in Figure 4 where the change point of affective wellbeing represented a 9% increase over median income in 2001, while in 2019 it represented a 42% increase over median income.
+Of course the counterfactual analysis represents the effects of temporal changes in all the model parameters not only the changepoint, but the changepoint was the only parameter which reliably changed over the 19 year period.
+For instance, Figure 4 shows the change point represented a 9% increase over median income in 2001, while in 2019 it represented a 42% increase over median income.
 This increase relative to median income also represented a reduction from 43% to 26% in the proportion of people whose income fell above the changepoint.

### 2021-06-10 src/draft-SSM1.Rmd
Closes issue #21

@@
 This increase relative to median income also represented a reduction from 43% to 26% in the proportion of people whose income fell above the changepoint.
+The increasing disparity in affective wellbeing between the rich and poor implied by the model is also shown in Figure S6 (Supplementary materials).
 Thus we can see that over the last nineteen years the difference in income-related affective wellbeing between the rich and the poor has increased; while the affective wellbeing of an increasing proportion of people, including the middle-class, is more dependent on their financial security.
@@
 Our results do not conflict with this conclusion.
-Rather we are revealing that even a static income distribution may have dynamic effects on happiness over time.
+Rather we are revealing that even a static income distribution may have dynamic effects on happiness over time via changes in their joint distribution.
 This highlights the issue that while traditional univariate measures of income and income inequality may be relatively stable and exhibit little change, their impact on wellbeing and health can still vary.

### 2021-06-10 src/supplementary-SSM1.Rmd
Closes issue #21
Update to discussion of Table S3. Closes issue #17

@@
 
+<br>
+
+The table of model parameters (untransformed) for each year show that the $\beta_1$ slopes for piecewise-linear, log-linear and linear model for affective wellbeing and income do not show a sustained trend ove the 19 year period.
+The $\beta_1$ coefficients increase in the initial 5-10 years and then decline.
+Despite this common pattern, the linear slope appears to show the most sustained increase while the log-linear slope shows a relatively less sustained increase.
+Overall, the pattern of changes among the models indicates the function between income and happiness has in fact become slightly steeper over the period.
+
 <br><br>
@@
 To directly examine the implications of our model for such inequities in happiness, we used the model to estimate or _predict_ the happiness of two different income levels: one income level which was above the change point in 2001 but fell below it by 2019 (\$50K/yr); and another level which always remained above the change points over the same period (\$75K/yr).
-Figure S8 below presents the happiness levels (in SD units/year) for each income level as well as the difference in happiness between them (∆).
+The figure below presents the happiness levels (in SD units/year) for each income level as well as the difference in happiness between them (∆).
 

### 2021-06-12 src/supplementary-SSM1.Rmd
Added sensitivity analyses. Closes issue #15

@@
 Draft: `r format(Sys.time(), '%d %B, %Y')`  
-Supplementary tables:      4  
-Supplementary figures:     6
+Supplementary tables:      5  
+Supplementary figures:     8
 
@@
 
-Household income and average life satisfaction increased between 2001-2019, while average happiness score decreased slightly over the 19 years.
+Household income and average life satisfaction increased between 2001-2019, while average affective wellbeing score decreased slightly over the 19 years.
 The proportions of each sex were stable over time, as was the proportion of people living as a couple (Couples), average household size and SEIFA index.
@@
 Generally, Table S1 supports the view that the HILDA sample was a relatively stable representation of Australian socioeconomic conditions between 2001 and 2019.
-The importance of the specific demographic variables with a significant linear trend (age, age^2^, education, illness) were tested in regression models of happiness as covariates (see below Figure S3).
+The importance of the specific demographic variables with a significant linear trend (age, age^2^, education, illness) were tested in regression models of affective wellbeing as covariates (see below Figure S2).
 
@@
 
-**Conditional vs unconditional models of happiness.** Table S1 (above) showed the specific variables that changed over the 19 years between 2001-2019.
+**Conditional vs unconditional models of affective wellbeing**.
+Table S1 (above) showed the specific variables that changed over the 19 years between 2001-2019.
 In particular, the number of university graduates, average age and the incidence of chronic illness increased slighty over the period.
 Including these variables as covariates in any model (i.e,.
-a conditional model such that *P*( $y$ | income, age, education illness)) will improve the in-sample fit (e.g., R^2^), and also adjust the parameter estimates of interest (e.g., change point) conditional upon their common association with happiness.
-Given the general aim of the main paper was to describe the temporal change in functional form between happiness and income in Australia, regardless of any other variable (i.e., an unconditional model of *P*($y$| income)) or sampling differences, neither the in-sample fit nor the conditional model estimates are directly relevant.
+a conditional model such that *P*( $y$ | income, age, education illness)) will improve the in-sample fit (e.g., R^2^), and also adjust the parameter estimates of interest (e.g., change point) conditional upon their common association with affective wellbeing.
+Given the general aim of the main paper was to describe the temporal change in functional form between affective wellbeing and income in Australia, regardless of any other variable (i.e., an unconditional model of *P*($y$| income)) or sampling differences, neither the in-sample fit nor the conditional model estimates are directly relevant.
 Nevertheless, the importance of these variables in mediating the changes we observed in the unconditional model are relevant, and so we calculated the estimates from conditional piecewise models including age, age^2^, sex, education  (university degree), chronic illness for each year below to confirm the pattern of temporal change was similar.
@@
 ```
-Figure legend: Comparison of the conditional parameter estimates (red) and unconditional parameter estimates (grey) in the piecewise-linear model of income on happiness.
+Figure legend: Comparison of the conditional parameter estimates (red) and unconditional parameter estimates (grey) in the piecewise-linear model of income on affective wellbeing.
 The conditional models included covariates for age, age squared, sex, education, and illness.
@@
 
-The parameter estimates of the conditional model (red) were adjusted for the broad demographic variables which tended to increase over the 19 years in our sample (age, education, illness); however the pattern of changes over time was comparable to the unconditional model (grey), indicating these temporal trends were not responsible for the evolution in the change point between income and happiness.
+The parameter estimates of the conditional model (red) were adjusted for the broad demographic variables which tended to increase over the 19 years in our sample (age, education, illness); however the pattern of changes over time was comparable to the unconditional model (grey), indicating these temporal trends were not responsible for the evolution in the change point between income and affective wellbeing.
 
 The unconditional parameter values are also presented in Table S2 below (Parameter and fit summary statistics).
+
+<br><br>
+
+**Including income topcodes in models of affective wellbeing**.
+
+Household income values for high income individuals are substituted with the average income value of all those above a certain threshold (topcoded).
+This is performed by the University of Melbourne to preserve anonymity, and we excluded people marked as topcoded in the main analysis (highest _n_ in any year was _n_ = 103 in 2019).
+Here we present the parameter estimates of the affective wellbeing models with topcodes included, and compare them with parameters from models which exclude them.
+
+##### Figure S3. Topcoded model parameters  
+```{r figure_s3, fig.width = 6, fig.height = 8, dpi=300}
+nl_fits_topcodes = read_rds("../results/fits/inc_topcodes/gh9_hhoecd_0cov_tight.RDS")
+
+topcode_income_params <- read_rds("../results/fits/inc_topcodes/gh9_hhoecd_data.RDS") %>%
+  bind_rows(.id = "fit")
+
+posteriors_topcodes <- map_dfr(nl_fits_topcodes, as.data.frame, .id = "fit") %>%
+  left_join(topcode_income_params) %>%
+  mutate(omega_dollars = (b_omega_Intercept * sd + mean)/1000) %>%
+  as_tibble()
+
+posteriors_topcodes %>%
+  select(fit, contains("b0"), contains("b1"), contains("b2"), omega_dollars) %>%
+  gather(key = "param", val = "sample", -fit) %>%
+  mutate(
+    year = extract_numeric(fit),
+    param = recode_factor(param, !!!param_key)
+  ) %>% 
+  group_by(param, year) %>%
+  summarise(
+    low = quantile(sample, probs = 0.05, na.rm=T),
+    expected = quantile(sample, probs = 0.5, na.rm=T),
+    upp = quantile(sample, probs = 0.95, na.rm=T)
+  ) %>%
+  mutate(model = "topcodes") -> parameter_summary_topcodes
+
+parameter_summary_0c %>%
+  mutate(model = "no topcodes") -> parameter_summary_excl
+
+bind_rows(parameter_summary_excl, parameter_summary_topcodes) %>%
+  ggplot(aes(y = year, x = expected, color = model)) +
+  geom_pointrange(aes(xmin = low, xmax = upp), position = position_dodge(width = 0.75)) +
+  scale_y_continuous(breaks = c(2001, 2007, 2013, 2019)) +
+  facet_wrap(~param, scales = "free_x") +
+  ft_theme(base_size = 12) +
+  scale_color_manual(values = c("grey65", "#C79999")) +
+  theme(legend.position = "none",
+        axis.title = element_blank())
+```
+```
+Figure legend: Comparison of the topcoded parameter estimates (red) and parameter estimates excluding topcodes (grey) in the piecewise-linear model of income on affective wellbeing.
+The topcoded models included n = 26 to 130 extra (high income) individuals in each year.
+Horizontal bar represents the 95% credible region and the solid point indicates the expected value (median) of each distribution.
+```
+
+<br>
+
+Included topcoded individuals adds some variation in the parameter estimates (e.g., 2008), but did not change the overall pattern observed over time (e.g., the changepoint continued to increase over time).
+
+<br><br>
+
+**Unadjusted household income models of affective wellbeing**.
+
+Household income values were adjusted for household size in the main analysis.
+The _maximum_ household sizes varied from 10 (2017, 2019) to as high as 17 (2016), and so adjustment accounts for larger households needing more resources to achieve the same standard of living as a smaller household, and for some ‘economies of scale’ when sharing living costs.
+For example, two people living alone in two separate households typically incur higher living costs in total than two people living in one household.
+Here we present the parameter estimates of the affective wellbeing models without adjustment for household size and with adjustment.
+
+##### Figure S4. Model parameters unadjusted for household size  
+```{r figure_s4, fig.width = 6, fig.height = 8, dpi=300}
+nl_fits_unadjsize = read_rds("../results/fits/excl_hhsize/gh9_hifdip_0cov_tight.RDS")
+
+posteriors_unadjsize <- map_dfr(nl_fits_unadjsize, as.data.frame, .id = "fit") %>%
+  left_join(income_params) %>%
+  mutate(omega_dollars = (b_omega_Intercept * sd + mean)/1000) %>%
+  as_tibble()
+
+posteriors_unadjsize %>%
+  select(fit, contains("b0"), contains("b1"), contains("b2"), omega_dollars) %>%
+  gather(key = "param", val = "sample", -fit) %>%
+  mutate(
+    year = extract_numeric(fit),
+    param = recode_factor(param, !!!param_key)
+  ) %>% 
+  group_by(param, year) %>%
+  summarise(
+    low = quantile(sample, probs = 0.05, na.rm=T),
+    expected = quantile(sample, probs = 0.5, na.rm=T),
+    upp = quantile(sample, probs = 0.95, na.rm=T)
+  ) %>%
+  mutate(model = "unadjusted") -> parameter_summary_unadjsize
+
+parameter_summary_0c %>%
+  mutate(model = "adjusted") -> parameter_summary_adj
+
+bind_rows(parameter_summary_unadjsize, parameter_summary_adj) %>%
+  ggplot(aes(y = year, x = expected, color = model)) +
+  geom_pointrange(aes(xmin = low, xmax = upp), position = position_dodge(width = 0.75)) +
+  scale_y_continuous(breaks = c(2001, 2007, 2013, 2019)) +
+  facet_wrap(~param, scales = "free_x") +
+  ft_theme(base_size = 12) +
+  scale_color_manual(values = c("grey65", "#C79999")) +
+  theme(legend.position = "none",
+        axis.title = element_blank())
+```
+```
+Figure legend: Comparison of the parameter estimates unadjusted for household size (red) and parameter estimates adjusted for size (grey) in the piecewise-linear model of income on affective wellbeing.
+Horizontal bar represents the 95% credible region and the solid point indicates the expected value (median) of each distribution.
+```
+
+<br>
+
+Omitting the adjustment for household size on household income appears to reduce the $\beta_1$ parameters in each year, but there is little systematic effect on the other parameters and the changepoint continues to increase over time.
 
@@
 Household income was then regressed against each score using the piecewise linear model (Equation 3), as described for the overall affective wellbeing score in the main text.
-Figure S3 below shows that the change-point in income increased over the period 2001-2019 for both positive (pink) and negative (blue) affect, indicating a similar trend in both components of affective wellbeing over the 19 years.
-
-##### Figure S3. Positive and negative affect parameters  
-```{r figure_s3, fig.width=6, fig.height=8, dpi=300}
+Figure S5 below shows that the change-point in income increased over the period 2001-2019 for both positive (pink) and negative (blue) affect, indicating a similar trend in both components of affective wellbeing over the 19 years.
+
+##### Figure S5. Positive and negative affect parameters  
+```{r figure_s5, fig.width=6, fig.height=8, dpi=300}
 gh9neg.fits <- read_rds("../results/fits/excl_topcodes/sd/gh9neg_hhoecd_0cov_tight.RDS")
@@
 Despite this common pattern, the linear slope appears to show the most sustained increase while the log-linear slope shows a relatively less sustained increase.
-Overall, the pattern of changes among the models indicates the function between income and happiness has in fact become slightly steeper over the period.
+Overall, the pattern of changes among the models indicates the function between income and affective wellbeing has in fact become slightly steeper over the period.
 
@@
 
-##### Figure S4. Prior predictive check  
-```{r figure_s4, fig.width=6, fig.height = 6, dpi=300}
+##### Figure S6. Prior predictive check  
+```{r figure_s6, fig.width=6, fig.height = 6, dpi=300}
 load("../results/prior_predictions.Rdata")
@@
 
-Examination of the residual plots from the piecewise models of happiness confirms the residuals were evenly distributed around zero and there was little indication of heteroscedascity.
-Overall, there was little evidence that the piecewise models of happiness were not appropriate for the data.
-
-<br>
-
-##### Figure S5. Residual plots for piecewise happiness models
-```{r figure_s5, fig.height=9, fig.width=9, dpi=600}
+Examination of the residual plots from the piecewise models of affective wellbeing confirms the residuals were evenly distributed around zero and there was little indication of heteroscedascity.
+Overall, there was little evidence that the piecewise models of affective wellbeing were not appropriate for the data.
+
+<br>
+
+##### Figure S7. Residual plots for piecewise affective wellbeing models
+```{r figure_s7, fig.height=9, fig.width=9, dpi=600}
 # https://www.flutterbys.com.au/stats/tut/tut7.2b.html
@@
 
-The posterior estimates of the piecewise model indicated that the relationship between happiness and income changed over time between 2001 and 2019.
-One implication of such a change is the disparity in happiness between income groups has increased.
-This will occur as more people fall below the change point over time and so are subject to the steep region of the function where income and happiness are strongly related.
-To directly examine the implications of our model for such inequities in happiness, we used the model to estimate or _predict_ the happiness of two different income levels: one income level which was above the change point in 2001 but fell below it by 2019 (\$50K/yr); and another level which always remained above the change points over the same period (\$75K/yr).
-The figure below presents the happiness levels (in SD units/year) for each income level as well as the difference in happiness between them (∆).
-
-<br>
-
-##### Figure S6. Happiness (SD units) at $50K/yr and $75K/yr from 2001-2019  
-```{r figure_s6, fig.width=7, fig.height=5, dpi=300}
+The posterior estimates of the piecewise model indicated that the relationship between affective wellbeing and income changed over time between 2001 and 2019.
+One implication of such a change is the disparity in affective wellbeing between income groups has increased.
+This will occur as more people fall below the change point over time and so are subject to the steep region of the function where income and affective wellbeing are strongly related.
+To directly examine the implications of our model for such inequities in affective wellbeing, we used the model to estimate or _predict_ the affective wellbeing of two different income levels: one income level which was above the change point in 2001 but fell below it by 2019 (\$50K/yr); and another level which always remained above the change points over the same period (\$75K/yr).
+The figure below presents the affective wellbeing levels (in SD units/year) for each income level as well as the difference in affective wellbeing between them (∆).
+
+<br>
+
+##### Figure S8. affective wellbeing (SD units) at $50K/yr and $75K/yr from 2001-2019  
+```{r figure_s8, fig.width=7, fig.height=5, dpi=300}
 nl_fits = read_rds("../results/gh9_hhoecd_0cov_tight.RDS")
@@
   labs(caption = "Source:     HILDA, University of Melbourne
-                 Predicted happiness scores (SD unit) from piecewise regression on real household income", 
+                 Predicted affective wellbeing scores (SD unit) from piecewise regression on real household income", 
        x = "", y = "") -> p
@@
 ```
-Figure S6 legend: The difference (∆) in happiness between a household income of $50K per year and a household income of $75K per year has increased between 2001 and 2019.
-```
-
-<br>
-
-Figure S6 shows the disparity in happiness between two fixed income levels, \$50K/year (traversing the change points) and \$75K/year (above the change points).
-In 2001, a household income of \$50K/year achieved an above average level of happiness relative to everyone else that year, however by 2019 happiness had declined to lower than average levels relative to the population.
-By contrast, a household income of \$75K/year enjoyed a higher than average level of happiness for the entire period.
-The difference (∆) in happiness between these two income levels doubled over the period.
-The increasing disparity in happiness between the rich and the poor implies that happiness has became more inequitable over time.
+Figure legend: The difference (∆) in affective wellbeing between a household income of $50K per year and a household income of $75K per year has increased between 2001 and 2019.
+```
+
+<br>
+
+The figure above shows the disparity in affective wellbeing between two fixed income levels, \$50K/year (traversing the change points) and \$75K/year (above the change points).
+In 2001, a household income of \$50K/year achieved an above average level of affective wellbeing relative to everyone else that year, however by 2019 affective wellbeing had declined to lower than average levels relative to the population.
+By contrast, a household income of \$75K/year enjoyed a higher than average level of affective wellbeing for the entire period.
+The difference (∆) in affective wellbeing between these two income levels doubled over the period.
+The increasing disparity in affective wellbeing between the rich and the poor implies that affective wellbeing has became more inequitable over time.
 
@@
 
-Three things need to be observed when interpreting the R^2^ values: 1) While the R^2^ between income and happiness is low (~2% and ~14% in the unconditional and conditional models, respectively), the effect over the wider population will aggregate to be much larger; 2) Happiness (and to a lesser extent, income) is an imperfect and noisy measure so any effect of income on happiness is likely to be larger than that measured here; 3) Finally, happiness is measured on a subjective and arbitrary Likert scale so a small change in happiness at different points in the scale may represent a large change in some other functional/real-world outcome (e.g., risk of suicide).
+Three things need to be observed when interpreting the R^2^ values: 1) While the R^2^ between income and affective wellbeing is low (~2% and ~14% in the unconditional and conditional models, respectively), the effect over the wider population will aggregate to be much larger; 2) affective wellbeing (and to a lesser extent, income) is an imperfect and noisy measure so any effect of income on happiness is likely to be larger than that measured here; 3) Finally, affective wellbeing is measured on a subjective and arbitrary Likert scale so a small change in affective wellbeing at different points in the scale may represent a large change in some other functional/real-world outcome (e.g., risk of suicide).
 This is all to say that quantifying the real-world impact of small changes in variation reported here is difficult and speculative without further investigation.
@@
 
+##### Table S5. OLS coefficients from linear, log-linear and piecewise models  
 ```{r table_s5}
@@
   fontsize(size = 9.5, part = "all") %>%
-  set_caption("OLS coefficients from linear, log-linear and piecewise models") %>%
   autofit()

### 2021-06-12 src/draft-SSM1.Rmd
Minor updates to intro and discusssion

@@
 - Nineteen years of data from a well-developed, economically stable G20 nation is examined (Australia), allowing a relatively undisturbed view of the changes without major economic shocks
-- We are able to show for the first time that the income level at which affective wellbeing is less determined by income has risen since 2001, resulting in greater inequality in affective wellbeing (happiness) between the rich and the poor
-- Conversely the linear relationship between income and cognitive wellbeing (life-satisfaction) has remained constant over this period
+- We are able to show for the first time that the income level at which affective wellbeing (happiness) is less determined by income has risen since 2001, resulting in greater inequality in happiness between the rich and the poor
+- Conversely the linear relationship between income and cognitive wellbeing (life-satisfaction) has remained relatively constant over this period
 - The happiness of an increasing proportion of the population is more dependent on their financial security than ever this millennium
@@
 For instance, the "satiety" point of USD75,000 in 2008 was substantially more than the US median income of USD52,000 in the same year, indicating that the affective wellbeing of the poorest majority of the US population was tied to marginal changes in income while the affective wellbeing of a richer minority was not.
-While much commentary has focused on the zero slope at high income levels reported by Kahneman & Diener ("satiety" point), any decrease in slope for the richer minority would also represent a relatively unequal distribution of happiness among the poor compared to the rich.
+While much commentary has focused on the zero slope at high income levels reported by Kahneman & Deaton ("satiety" point), any decrease in slope for the richer minority would also represent a relatively unequal distribution of happiness among the poor compared to the rich.
 Thus, even a non-zero changepoint indicates that income is tied to inequities in the distribution of affective wellbeing in the economy.
@@
 And no studies to our knowledge have ever tested the presence of a changepoint between wellbeing and income using standard modelling approaches.
-For this reason, a formal comparison between log-linear, linear and a changepoint function (piecewise-linear) in the relationship between affective wellbeing and income is warranted.
+For this reason, a formal comparison between log-linear, linear and a changepoint function (i.e., piecewise) in the relationship between affective wellbeing and income is warranted.
 
@@
 We found the marginal effect of household income on affective- and cognitive-wellbeing was positive but quite different: Affective wellbeing increased rapidly up to a point after which higher levels of income were associated with less improvement, while cognitive wellbeing tended to increase linearly or constantly with income.
-Other studies have also reported that income has distinct effects on cognitive and affective wellbeing [@howell2008relation; @kahneman2010high], with a zero-slope change point in affective wellbeing after which increasing income has little further effect.
-In our case the changepoint did not precede a zero-slope, as the post-slope for each year was credibly above zero (Figure 2), but it did represent a substantial decrease in dependency between income and affective wellbeing for those higher income levels.
-Our novel finding was revealing the change point increased faster than both inflation and median household income between 2001 and 2019.
-For the first time we have shown there has been a temporal shift in the change point between income and affective wellbeing over the 19 year period, such that affective wellbeing has become more dependent on income for more people - especially the poor and middle-class.
+Other studies have also reported that income has distinct effects on cognitive and affective wellbeing [@howell2008relation], with a zero-slope change point in affective wellbeing after which increasing income has little further effect [@kahneman2010high].
+In our case we did not observe a zero-slope changepoint ("satiety" point), as the post-slope for each year was credibly above zero (Figure 2); but the changepoint did represent a substantial decrease in dependency between income and affective wellbeing for those higher income levels.
+Our novel finding was revealing the changepoint increased faster than both inflation and median household income between 2001 and 2019.
+For the first time we have shown there has been a temporal shift in the changepoint between income and affective wellbeing over the 19 year period, such that affective wellbeing has become more dependent on income for more people - especially the poor and middle-class.
 
@@
 Our results emphasize the slope over income is not constant.
-Other researchers have argued that affective wellbeing is a constant (albiet decreasing) log-linear function of income [@killingsworth2021experienced; @sacks2012newl] or a constant linear relationship [@twenge2020expanding].
+Other researchers have argued that affective wellbeing is a constant (albiet decreasing) log-linear function of income [@killingsworth2021experienced; @sacks2012newl; @twenge2020expanding].
 We also show the change-point has increased over time.
-Indeed Twenge et al report the linear relationship between income and happiness has become steeper over time in the US General Social Survey (GSS).
+Indeed Twenge & Cooper report the linear relationship between income and happiness has become steeper over time in the US General Social Survey (GSS).
 While we did not see any consistent trend over time in pre or post-slope changes in the piecewise-linear model of affective-wellbeing (e.g., Figure 2), there was a credibly increasing slope in the linear model of cognitive-wellbeing with time (Table S3).
-It may be the single-item on general happiness in the GSS has more in common with the single item on life-satisfaction in HILDA than the more comprehensive 9-item, dual-component questions on affective-wellbeing in the SF-36.
-Nevertheless, like Twenge et al, we report that the difference in happiness betweeen the rich and poor has been increasing over time (e.g., Figure S6), however our piecewise-model reveals the income level around which this difference occurs.
+It may be the single-item on general happiness in the GSS has more in common with the single item on life-satisfaction in HILDA than the more comprehensive 9-item, dual-valence questions on affective-wellbeing in the SF-36.
+Nevertheless, like Twenge et al, we report that the difference in happiness betweeen the rich and poor has been increasing over time (e.g., Figure S6), moreover our piecewise-model reveals the income level around which this difference appears.
 

### 2021-06-21 src/draft-SSM1.Rmd
Restored a consistent linear scale to Figure 1 x-axis

@@
 
-log_breaks <- c(log(12000), log(25000), log(50000), log(100000))
-
 p1 <- figure1_data %>%
   rename(`Affective 1-100` = gh9) %>% 
-  plot_log2(`Affective 1-100`, `Household income`) +
-  scale_x_continuous(breaks = log_breaks, labels = c(12, 25, 50, 100)) +
+  plot_log(`Affective 1-100`, `Household income`) +
   scale_fill_manual(values = "#C79999") +
@@
   rename(`Cognitive 0-10` = losat) %>% 
-  plot_log2(`Cognitive 0-10`, `Household income`) +
-  scale_x_continuous(breaks = log_breaks, labels = c(12, 25, 50, 100)) +
+  plot_log(`Cognitive 0-10`, `Household income`) +
   scale_fill_manual(values = "#b3cde0") +

### 2021-06-21 src/draft-SSM1.Rmd
Nick's revisions

@@
 
-Funding sources: This research did not receive any specific grant from funding agencies in the public, commercial, or not-for-profit sectors
+Funding sources: This research was supported by the Australian Government through the Australian Research Council's Centre of Excellence for Children and Families over the Life Course (Project ID CE200100025)
 
@@
 HILDA (the Household, Income, and Labour Dynamics in Australia survey) provides a representative sample of households in Australia with detailed measurements of income and subjective wellbeing each year.
-We distinguished between cognitive- and affective wellbeing as different components of subjective wellbeing, and evaluated how each varied with household income over time using linear, log-linear and piecewise linear regression in each year.
+We distinguished between cognitive- and affective wellbeing as different components of subjective wellbeing, and evaluated how each varied with household income over time using linear, log-linear and piecewise-linear regression in each year.
 
@@
 _Parameter Estimation_  
-To determine the location of the change point (*&omega;*) between wellbeing and income, we modelled the relationship between income and wellbeing across individuals using the piecewise model described above, and sampled the posterior probability of *&omega;* over 4000 interations.
+To determine the location of the change point (*&omega;*) between wellbeing and income, we modelled the relationship between income and wellbeing across individuals using the piecewise-linear model described above, and sampled the posterior probability of *&omega;* over 4000 interations.
 The complete posterior distribution of *&omega;* for each year is presented along with the expected value (median).
@@
 The relationship between household income and affective wellbeing (red) and satisfaction (blue) every four years is shown in Figure 1.
-For each wellbeing variable we show the results of a linear fit (rows 1 and 3) and a piecewise fit (rows 2 and 4).
+For each wellbeing variable we show the results of a linear fit (rows 1 and 3) and a piecewise-linear fit (rows 2 and 4).
 For visualization purposes only, due to the large number of individuals in each year, we display the mean levels of income and wellbeing for each (equal-sized) income decile, whereas the line-of-best-fit and 95% credible intervals (shaded) in each regression model are derived from _all individuals_.
@@
 
-##### Figure 1. Household income and affective- (red) and cognitive- (blue) wellbeing  
+##### Figure 1. Association Between Household Income and Affective- (red) and Cognitive- (blue) Wellbeing from 5-years 2001-2019  
 ```{r hh_disp_inc_eq}
@@
 ```
-Figure 1 legend: The relationship between income and wellbeing across equal-sized income deciles, overlaid by regression lines from linear and piecewise models (±95%CI).
+Figure 1 legend: The relationship between income and wellbeing across equal-sized income deciles, overlaid by regression lines from log-linear and piecewise-linear models (±95%CI).
 Wellbeing was measured as affective- (red) or cognitive- (blue) wellbeing.
@@
 
-The piecewise relationship between affective wellbeing and income (Figure 1, 2nd row) was consistently and negatively inflected (affective wellbeing increased less with income after the change point).
+The piecewise-linear relationship between affective wellbeing and income (Figure 1, 2nd row) was consistently and negatively inflected (affective wellbeing increased less with income after the change point).
 By contrast the piecewise relationship between cognitive wellbeing and income shown in the 4th row exhibited negative inflection (2001, 2015, 2019), positive inflection (2005), and no apparent inflection (2010).
 
-The evidence from model selection revealed the piecewise-linear fit of _affective_-, but not _cognitive_ wellbeing, was superior to a log-linear or linear fit in each year (Supplementary Materials, Figure S1).
+The evidence from model selection revealed the piecewise-linear fit of _affective_- was superior to a log-linear or linear fit in each year with a penalised WAIC score credibly (95% interval) smaller than the log-linear fits for all but three of the nineteen years (and for all 19 years for the linear fits, see Supplementary Materials, Figure S1).
+There was little evidence that any of the three models were reliably or credibly superior in representing the association between income and cognitive-wellbeing over the same period, albeit the linear model was better on average between 2002-2004 and 2007-2010 while the piecewise-linear model was better the remaining years.
 Overall the model comparison suggested that affective- and cognitive wellbeing have distinct relationships with household income; affective wellbeing increases with income more rapidly at lower household income levels than higher income levels with a distinct changepoint in each year, while cognitive wellbeing tends to increase with income, most likely in a linear fashion.
-Others have also noted cognitive wellbeing (life satisfaction) has a stronger linear relationship with income, than affective wellbeing and income [@howell2008relation], and the results of our formal comparison provides some support for this notion.
+Others have also noted cognitive wellbeing (life satisfaction) has a stronger linear relationship with income than affective wellbeing and income [@howell2008relation], and so the results of our formal comparison provides some support for this.
 
@@
 
-Figure 2 below presents the posterior distribution of each parameter from the piecewise model regressing affective wellbeing on income: the change point (&omega;), the intercept (&beta;~0~), the pre-change point slope (&beta;~1~), and the post-change point slope (&beta;~2~).
+Figure 2 below presents the posterior distribution of each parameter from the piecewise-linear model regressing affective wellbeing on income: the change point (&omega;), the intercept (&beta;~0~), the pre-change point slope (&beta;~1~), and the post-change point slope (&beta;~2~).
 Horizontal bars represent the 95% credible interval of the posterior distribution and so intervals which fall completely to the right of the vertical grey dotted line are credibly higher than the expected value of our base year, 2001.
@@
 
-##### Figure 2. Posterior parameters of the affective-wellbeing ~ income piecewise model (real 2019 dollars)   
+##### Figure 2. Posterior parameters of the affective-wellbeing ~ income piecewise-linear model (real 2019 dollars)   
 ```{r get_posteriors}
@@
 Changes in the other demographic variables (e.g., age, education levels and chronic illness) did not materially alter the trends just described.
-The parameters of conditional piecewise models which included age, age^2^, education, and illness as covariates are compared in Supplementary materials (Figure S2).
+The parameters of conditional piecewise-linear models which included age, age^2^, education, and illness as covariates are compared in Supplementary materials (Figure S2).
 
@@
 It may be the single-item on general happiness in the GSS has more in common with the single item on life-satisfaction in HILDA than the more comprehensive 9-item, dual-valence questions on affective-wellbeing in the SF-36.
-Nevertheless, like Twenge et al, we report that the difference in happiness betweeen the rich and poor has been increasing over time (e.g., Figure S6), moreover our piecewise-model reveals the income level around which this difference appears.
+Nevertheless, like Twenge et al, we report that the difference in happiness between the rich and poor has been increasing over time (e.g., Figure S6).
+Moreover our piecewise-linear model reveals the income level around which this difference appears.
 
@@
 Moreover, financial gains and losses have one of the largest impacts on subjective wellbeing in within-subject/fixed-effect models [e.g., @kettlewell2020differential], and there are many reasons why cross-sectional estimates of the covariance would under-represent this.
-Another limitation is that wellbeing is measured on a subjective scale, and we do not know whether small changes at different points on the scale represent larger changes in some other functional/real-world outcome (e.g., risk of suicide).
+Another limitation is that both types of wellbeing are measured on subjective scales, with different numbers of items for each.
+This would lead to differential measurement error, although _a priori_ this would not cause the association between income and wellbeing to differ but rather reduce the precision of parameter estimates (which may explain our failure to distinguish the models of cognitive wellbeing).
+We also do not know whether small changes at different points on the scale represent larger changes in some other functional/real-world outcome (e.g., risk of suicide).
 At present, the real-world impact of changes in a subjective measure such as wellbeing is difficult to determine.
@@
 #### 4.2 Conclusions  
+We justify a novel method, that challenges standard approaches, to show that the marginal effect of income on affective wellbeing is weaker for higher income groups (although not zero, i.e., "satiety"), and this point has been increasing over the past two decades.
 Australia has low levels of income disparity relative to many other OECD countries, and the Gini coefficient has changed little between 2001 and 2019 [@australian2018rising], suggesting income inequality has remained steady over this time period.
 Our results do not conflict with this conclusion.
-Rather we are revealing that even a static income distribution may have dynamic effects on happiness over time via changes in their joint distribution.
+Rather we suggest that even a static income distribution may have dynamic effects on happiness over time via changes in their joint distribution.
 This highlights the issue that while traditional univariate measures of income and income inequality may be relatively stable and exhibit little change, their impact on wellbeing and health can still vary.

### 2021-06-24 src/draft-SSM1.Rmd
Final resubmission revisions. Closes #20, #17, #4

@@
 title: "The increasing cost of happiness"
-author: "R.W. Morris^1,2^, N. Kettlewell^2,3,4^  & N.Glozier^1,2^"
+author: "R.W.Morris^1,2^, N.Kettlewell^2,3,4^  & N.Glozier^1,2^"
 bibliography: ../src/references.bib
@@
 | Draft      | `r format(Sys.time(), '%d %B, %Y')` |
-| Words      | 4084                                |
+| Words      | 4704                                |
 | Tables     | 0                                   |
@@
 The accepted view among economists and psychologists is that money and happiness increase together up to a point, after which there is little further gain from increasing income.
-While the location of this change point reportedly ranges between USD\$60K to \$95K, there has been no investigation as to whether this has increased or decreased over time.
-We tested the temporal relationship between income and affective wellbeing (happiness), and income and cognitive wellbeing (life satisfaction), using household economic data from Australia between 2001-2019.
-We discovered the change point between affective wellbeing and income has increased over those 19 years faster than inflation (i.e., cost of living).
-This suggests that inequalities in income may be driving increasing inequities in happiness between the rich and the poor, with implications for health and recent government policy-goals to monitor and improve wellbeing.
+The location of this change point was reportedly USD\$75K in 2008, which was substantially more than the median income in the same year (USD\$52K), indicating the happiness or wellbeing of the poorest majority of the population was tied to marginal changes in income while the wellbeing of a richer minority was not.
+Thus the location of the change point has implications for the distribution of wellbeing between the rich and poor.
+Inequalities in wellbeing are increasingly relevant due to the growing recognition that increasing income does not neccesarily lead to equal changes in wellbeing.
+In this paper we estimate the relationship between income and subjective wellbeing over a 19-year period focusing on whether this change point has shifted over time, and explore its implications for inequalities in wellbeing.
+To date there has been no investigation as to whether the change point has increased or decreased over time so we formally tested for the presence of a change point as well temporal changes in the relationship between income and affective wellbeing (happiness), and income and cognitive wellbeing (life satisfaction), using household economic data from Australia between 2001-2019.
+The results revealed the change point between affective wellbeing and income has increased over those 19 years faster than inflation (i.e., cost of living).
+This suggests that inequalities in income may be driving increasing _inequities_ in happiness between the rich and the poor, with implications for health and recent government policy-goals to monitor and improve wellbeing.
 
@@
 - Nineteen years of data from a well-developed, economically stable G20 nation is examined (Australia), allowing a relatively undisturbed view of the changes without major economic shocks
-- We are able to show for the first time that the income level at which affective wellbeing (happiness) is less determined by income has risen since 2001, resulting in greater inequality in happiness between the rich and the poor
+- We are able to show for the first time that the income level at which affective wellbeing (happiness) is less responsive to income has risen since 2001, resulting in a greater inequality in the distribution of affective wellbeing between the rich and the poor
 - Conversely the linear relationship between income and cognitive wellbeing (life-satisfaction) has remained relatively constant over this period
-- The happiness of an increasing proportion of the population is more dependent on their financial security than ever this millennium
+- The association between happiness and income is becoming more prevalent in the population this millennium, and likely to drive inequities in wellbeing across the economy
 
@@
 
-Fundementally, a difference in slope between affective wellbeing and income for the rich and poor represents an unacknowledged source of inequality in the distribution of wellbeing (i.e., happiness) in the economy.
+Fundamentally, a difference in slope between affective wellbeing and income for the rich and poor represents an unacknowledged source of inequity (i.e., unfairness) in the distribution of wellbeing (i.e., happiness) in the economy.
 For instance, the "satiety" point of USD75,000 in 2008 was substantially more than the US median income of USD52,000 in the same year, indicating that the affective wellbeing of the poorest majority of the US population was tied to marginal changes in income while the affective wellbeing of a richer minority was not.
-While much commentary has focused on the zero slope at high income levels reported by Kahneman & Deaton ("satiety" point), any decrease in slope for the richer minority would also represent a relatively unequal distribution of happiness among the poor compared to the rich.
-Thus, even a non-zero changepoint indicates that income is tied to inequities in the distribution of affective wellbeing in the economy.
-Conversely, a more equitable distribution of wellbeing in the economy would correspond to a lower valued change point.
+While much commentary has focused on the zero slope at high income levels reported by Kahneman & Deaton, the positive slope at lower income levels means wellbeing is distributed unequally over those income ranges.
+Moreover, the location of the change point in slope has implications for the distribution of wellbeing over the entire income range.
+For instance, an increase in the change point (i.e., a right shift on the income distribution) indicates more inequality in wellbeing between the rich and poor.
+Conversely, a decrease (left shift) represents a more equal distribution of wellbeing across income levels.
+To our knowledge, no studies to date have examined whether the location of the change point has shifted over time, but such shifts may help us understand how the distribution of wellbeing has been changing for different parts of the economy.
 
-While studies agree that affective wellbeing tends to increase with income, the exact form of the functional relationship between affective wellbeing and income is contested: some studies show the presence of a zero-slope changepoint after which increasing income has no further effect [@kahneman2010high; @jebb2018happiness]; other studies have found that affective wellbeing is a constant (albiet decreasing) log-linear function of income [@killingsworth2021experienced; @sacks2012new; @twenge2020expanding].
-Few studies have formally tested the functional form of the relationship, relying instead upon binning the data (i.e., over income deciles or quintiles) and comparing adjacent bins either visually or with null hypothesis tests which do not provide evidence for a changepoint.
-In this regard, arbitrary choices about bin width can have salient effects on the function form, which in turn will influence the inferences drawn.
-And no studies to our knowledge have ever tested the presence of a changepoint between wellbeing and income using standard modelling approaches.
-For this reason, a formal comparison between log-linear, linear and a changepoint function (i.e., piecewise) in the relationship between affective wellbeing and income is warranted.
+While studies agree that affective wellbeing tends to increase with income, the exact form of the functional relationship between affective wellbeing and income is contested: some studies show the presence of a zero-slope change point after which increasing income has no further effect [@kahneman2010high; @jebb2018happiness]; other studies have found that affective wellbeing is approximately log-linear [@killingsworth2021experienced; @sacks2012new; @twenge2020expanding].
+Few studies have formally tested the functional form of the relationship, relying instead upon p-value comparisons after dividing the data into arbitrary bins (i.e., deciles, quintiles, quartiles), which can have salient effects on the functional form observed.
+When p-values have been determined, they have not evaluated the probability of a change point given the data, but rather the probability of the data assuming no change - which is not a direct test.
+binning the data (i.e., over income deciles or quintiles) and comparing adjacent bins either visually or with null hypothesis tests which do not provide evidence for a change point.
+In this regard, no studies to our knowledge have ever directly tested for the presence of a change point between wellbeing and income using standard modelling approaches.
+For this reason, a formal comparison between log-linear, linear and a change point function (i.e., piecewise) in the relationship between affective wellbeing and income is warranted.
 
-Inequities in the distribution of wellbeing are increasingly relevant to governments and policy-makers due to the growing recognition that increasing income does not necessarily lead to equal changes in wellbeing [@clark2018four; @frijters2020happy].
+Inequalities in the distribution of wellbeing are increasingly relevant to governments and policy-makers due to the growing recognition that increasing income does not necessarily lead to equal changes in wellbeing [@clark2018four; @frijters2020happy].
 Even prior to COVID-19, the World Gallup Poll has observed that affective wellbeing has decreased over the past decade in western Europe, North America, Australia and New Zealand, despite increases in income in the same countries [@sachs2019world].
 However to date there has been little investigation of whether the relationship between income and affective wellbeing has changed over time, which may contribute to these trends.
-In particular has the functional relationship between income and affective wellbeing, and therefore the distribution of happiness between rich and poor, become more or less equitable in the last few decades?
+In particular, has the functional relationship between income and affective wellbeing, and therefore the distribution of happiness between rich and poor, become more or less equitable in the last few decades?
 
@@
 We have previously shown a summed score of these 9 questions distinguishes the impact of good and bad major life events in a bidirectional manner [@kettlewell2020differential].
-Supplementary analysis on the subset of positive and negative questions did not reveal any difference in the pattern of results between them (Supplementary materials, Figure S3), and so we reverse scored negatively-phrased questions and calculated the sum of the nine questions so that higher scores represented better wellbeing.
+Supplementary analysis on the subset of positive and negative questions did not reveal any difference in the pattern of results between them (Supplementary materials, Figure S5), and so we reverse scored negatively-phrased questions and calculated the sum of the nine questions so that higher scores represented better wellbeing.
 To aid interpretability, we rescaled the final sum to a score between 1-100, where 100 represents the maximum affective wellbeing achievable.
@@
 
-We modelled the relationship between income and each wellbeing variable (cognitive and affective) using a linear model, log-linear model, and a piecewise-linear model with a single change-point as a free parameter estimated from the data.
+We modelled the relationship between income and each wellbeing variable (cognitive and affective) using a linear model, log-linear model, and a piecewise-linear model with a single change point as a free parameter estimated from the data.
 The piecewise-linear model was chosen as the simplest extension of a linear model which can identify a change point (inflection) in the relationship between wellbeing and income.
@@
 $$
-  \mu_i = \beta_0 + B X_i \ \ \ \ \ \ \ \ \ \ (2)
+  \mu_i = \beta_0 +  X_iB \ \ \ \ \ \ \ \ \ \ (2)
 $$
@@
 Where $X_i$ was an individual's log household income (log(\$)) as well as other covariates (i.e., age, age^2^, sex, education, chronic illness, HILDA population weights; see below), and $y_i$ was an individual's wellbeing.
-Note that $B$ is a column matrix with a term 1...p for each parameter in $X_i$.
+Note that $B$ is a column vector with a term 1...p for each parameter in $X_i$.
 
@@
 
-The piecewise-linear model included a free parameter to represent the changepoint in income ($\omega$) as well as the slope before the change point ($\beta_1$) and the slope after the change point ($\beta_2$):
+The piecewise-linear model included a free parameter to represent the change point in income ($\omega$) as well as the slope before the change point ($\beta_1$) and the slope after the change point ($\beta_2$):
 
@@
 These regularizing priors were selected to reduce overfitting while still allowing the model to learn the regular features of the sample, and so provide a more robust population estimate.
-They are skeptical priors and a prior predictive check confirmed they assume no relationship between income and wellbeing, with no difference in gradient before or after the change point (Supplementary materials, Figure S4).
+They are skeptical priors and a prior predictive check confirmed they assume no relationship between income and wellbeing, with no difference in gradient before or after the change point (Supplementary materials, Figure S6).
 
@@
 To determine the location of the change point (*&omega;*) between wellbeing and income, we modelled the relationship between income and wellbeing across individuals using the piecewise-linear model described above, and sampled the posterior probability of *&omega;* over 4000 interations.
-The complete posterior distribution of *&omega;* for each year is presented along with the expected value (median).
+The complete posterior distribution of *&omega;* for each year is presented along with the median.
 
@@
 Full-time students were removed, as well as individuals with an annual household disposable income that was negative or indicated as _topcoded_ by the University of Melbourne.
-The results of sensitivity analyses to determine the effect of the covariates and exclusions on the change point parameters are included in Supplementary Materials (**add section reference here**).
+The results of sensitivity analyses to determine the effect of the covariates and exclusions on the change point parameters are included in Supplementary Materials (Figure S2, S3, S4).
 
@@
 The relationship between household income and affective wellbeing (red) and satisfaction (blue) every four years is shown in Figure 1.
-For each wellbeing variable we show the results of a linear fit (rows 1 and 3) and a piecewise-linear fit (rows 2 and 4).
+For each wellbeing variable we show the results of a log-linear fit (rows 1 and 3) and a piecewise-linear fit (rows 2 and 4).
 For visualization purposes only, due to the large number of individuals in each year, we display the mean levels of income and wellbeing for each (equal-sized) income decile, whereas the line-of-best-fit and 95% credible intervals (shaded) in each regression model are derived from _all individuals_.
@@
 There was little evidence that any of the three models were reliably or credibly superior in representing the association between income and cognitive-wellbeing over the same period, albeit the linear model was better on average between 2002-2004 and 2007-2010 while the piecewise-linear model was better the remaining years.
-Overall the model comparison suggested that affective- and cognitive wellbeing have distinct relationships with household income; affective wellbeing increases with income more rapidly at lower household income levels than higher income levels with a distinct changepoint in each year, while cognitive wellbeing tends to increase with income, most likely in a linear fashion.
+Overall the model comparison suggested that affective- and cognitive wellbeing have distinct relationships with household income; affective wellbeing increases with income more rapidly at lower household income levels than higher income levels with a distinct change point in each year, while cognitive wellbeing tends to increase with income, most likely in a linear fashion.
 Others have also noted cognitive wellbeing (life satisfaction) has a stronger linear relationship with income than affective wellbeing and income [@howell2008relation], and so the results of our formal comparison provides some support for this.
@@
 
-Both the pre-slope and post-slope parameters (&beta;~1~ and &beta;~2~) were credibly larger than zero in each year, indicating there was a reliable dependency between affective wellbeing and income at income levels both below and above the changepoint.
+Both the pre-slope and post-slope parameters (&beta;~1~ and &beta;~2~) were credibly larger than zero in each year, indicating there was a reliable dependency between affective wellbeing and income at income levels both below and above the change point.
 However the pre-slope parameter was in general four to five times greater than the post-slope values, indicating that the relationship between affective wellbeing and income among people in lower income households was an order of magnitude stronger than those in high income households.
@@
 Changes in the other demographic variables (e.g., age, education levels and chronic illness) did not materially alter the trends just described.
-The parameters of conditional piecewise-linear models which included age, age^2^, education, and illness as covariates are compared in Supplementary materials (Figure S2).
+The effect of adding the covariates (age, age^2^, education, illness) on the piecewise-linear model parameters are compared in Supplementary materials (Figure S2).
 
@@
 These 2001 model-estimates were compared to (subtracted from) the 2019 model-estimates generated from the same sample (*n* = 14,459), to obtain the change (delta) in affective wellbeing for each person under the counterfactual.
-This delta is attributable to the evolution of the function between 2001 and 2019.
-Figure 3 below presents the 14,459 deltas from such a comparison, along with a smoothed mean (solid line) to summarize how the distribution of affective wellbeing evolved across the income range.
+Thus, the delta is attributable to the evolution of the function over the 19 years.
+Figure 3 below presents the 14,459 deltas from such a comparison, along with a smoothed mean (solid line) to summarize how the predicted difference in affective wellbeing is unequally distributed across the income range under the two models.
 
@@
 Conversely, people with household incomes below \$50K, on average, had a decrease in predicted affective wellbeing under the 2019 function compared to 2001 function.
+These results are consistent with the picture in Figure 1, which shows for example that affective wellbeing changed by -5.79% for people with income in the bottom decile between 2001 and 2019 compared to -2.92% for people in the top decile.
+
 Of course the obtained deltas are due to changes in all the parameters of the function, including the slope before and after the change point.
@@
 This shows the change point between income and affective wellbeing increased faster than rises in median household income between 2001 and 2019.
-The third panel shows that as a result, a smaller proportion of the Australian population in 2019 had a household income above the changepoint than in 2001.
+The third panel shows that as a result, a smaller proportion of the Australian population in 2019 had a household income above the change point than in 2001.
 
@@
 Other studies have also reported that income has distinct effects on cognitive and affective wellbeing [@howell2008relation], with a zero-slope change point in affective wellbeing after which increasing income has little further effect [@kahneman2010high].
-In our case we did not observe a zero-slope changepoint ("satiety" point), as the post-slope for each year was credibly above zero (Figure 2); but the changepoint did represent a substantial decrease in dependency between income and affective wellbeing for those higher income levels.
-Our novel finding was revealing the changepoint increased faster than both inflation and median household income between 2001 and 2019.
-For the first time we have shown there has been a temporal shift in the changepoint between income and affective wellbeing over the 19 year period, such that affective wellbeing has become more dependent on income for more people - especially the poor and middle-class.
+In our case we did not observe a zero-slope change point ("satiety" point), as the post-slope for each year was credibly above zero (Figure 2); but the change point did represent a substantial decrease in dependency between income and affective wellbeing for those on higher income levels.
+Our novel finding was revealing the change point increased faster than both inflation and median household income between 2001 and 2019.
+For the first time we have shown there has been a temporal shift in the change point between income and affective wellbeing over the 19 year period, such that affective wellbeing has become more dependent on income for more people - especially the poor and middle-class.
 
@@
 
-An implication of the changing relationship between affective wellbeing and income is that income inequality may be driving increasing inequities in wellbeing.
-This is demonstrated in Figure 3, where the difference in predicted affective wellbeing between 2001 and 2019 increased for incomes above $50K/year and decreased for incomes below that level (see also Figure S6 in Supplementary materials).
-Of course the counterfactual analysis represents the effects of temporal changes in all the model parameters not only the changepoint, but the changepoint was the only parameter which reliably changed over the 19 year period.
+An implication of the changing relationship between affective wellbeing and income is that income inequality may be driving increasing inequality in wellbeing.
+This is depicted in Figure 3, where the difference in predicted affective wellbeing between 2001 and 2019 increased for incomes above $50K/year and decreased for incomes below that level (see also Figure S8 in Supplementary Materials).
+Of course the counterfactual analysis represents the effects of temporal changes in all the model parameters not only the change point, but the change point was the only parameter which reliably changed over the 19 year period.
 For instance, Figure 4 shows the change point represented a 9% increase over median income in 2001, while in 2019 it represented a 42% increase over median income.
-This increase relative to median income also represented a reduction from 43% to 26% in the proportion of people whose income fell above the changepoint.
-The increasing disparity in affective wellbeing between the rich and poor implied by the model is also shown in Figure S6 (Supplementary materials).
+This increase relative to median income also represented a reduction from 43% to 26% in the proportion of people whose income fell above the change point.
 Thus we can see that over the last nineteen years the difference in income-related affective wellbeing between the rich and the poor has increased; while the affective wellbeing of an increasing proportion of people, including the middle-class, is more dependent on their financial security.
@@
 This is consistent with a change point after which the slope is zero.
-While we observed a change-point, the slope afterwards was small but credibly larger than zero so our results are not strictly consistent with a "satiety" point.
+While we observed a change point, the slope afterwards was small but credibly larger than zero so our results are not strictly consistent with a "satiety" point.
 Our results emphasize the slope over income is not constant.
-Other researchers have argued that affective wellbeing is a constant (albiet decreasing) log-linear function of income [@killingsworth2021experienced; @sacks2012newl; @twenge2020expanding].
-We also show the change-point has increased over time.
+Other researchers have argued that affective wellbeing is a constant (albiet decreasing) log-linear function of income [@killingsworth2021experienced; @sacks2012new; @twenge2020expanding].
+We also show the change point has increased over time.
 Indeed Twenge & Cooper report the linear relationship between income and happiness has become steeper over time in the US General Social Survey (GSS).
@@
 It may be the single-item on general happiness in the GSS has more in common with the single item on life-satisfaction in HILDA than the more comprehensive 9-item, dual-valence questions on affective-wellbeing in the SF-36.
-Nevertheless, like Twenge et al, we report that the difference in happiness between the rich and poor has been increasing over time (e.g., Figure S6).
+Nevertheless, consistent with Twenge et al, our results imply that the difference in happiness between the rich and poor has been increasing over time (e.g., Figure 6 and S8).
 Moreover our piecewise-linear model reveals the income level around which this difference appears.
@@
 While the amount of variance was small, the pre-change slope was reliably above zero and small changes in happiness over an entire population will aggregate.
-Moreover, financial gains and losses have one of the largest impacts on subjective wellbeing in within-subject/fixed-effect models [e.g., @kettlewell2020differential], and there are many reasons why cross-sectional estimates of the covariance would under-represent this.
+Moreover, financial gains and losses have one of the largest impacts on subjective wellbeing in within-subject/fixed-effect models [e.g., @kettlewell2020differential], and there are many reasons why cross-sectional estimates of the covariance would under-represent this (e.g., measurement error).
 Another limitation is that both types of wellbeing are measured on subjective scales, with different numbers of items for each.
@@
 Our results do not conflict with this conclusion.
-Rather we suggest that even a static income distribution may have dynamic effects on happiness over time via changes in their joint distribution.
-This highlights the issue that while traditional univariate measures of income and income inequality may be relatively stable and exhibit little change, their impact on wellbeing and health can still vary.
-As such, these results may well have relevance to other developed nations in North America and Europe which also enjoy low levels of income inequality, but have stagnating incomes, and declining happiness levels [@sachs2019world].
+Rather we suggest that even a static income distribution may have dynamic effects on happiness over time via changes in their joint relationship.
+This highlights the issue that while traditional measures of income and income inequality may be relatively stable and exhibit little change, their impact on wellbeing and health can still vary and change over time.
+As such, these results may well have relevance to other developed nations in North America and Europe which also enjoy stable levels of income inequality, but have stagnating incomes and declining happiness levels [@sachs2019world].
 Establishing the links between income, wellbeing and health, and how inequalities in one drives inequities in the other, will be a critical aim as government begin to shift focus away from traditional measures of economic prosperity.

### 2021-06-24 src/supplementary-SSM1.Rmd
Final resubmission revisions. Closes #20, #17, #4

@@
     statistic = 
-      list(all_continuous() ~ "{mean} (±{sd})",
+      list(all_continuous() ~ "{mean} ({sd})",
            all_categorical() ~ "{n} ({p}%)"),
@@
   width(j = 1, width = 1.5) %>%
-  width(j = 2:6, width = 1)
+  width(j = 2:5, width = 1.1) %>%
+  width(j = 6, width = .6)
 ```
@@
     Est.Error = if_else(str_detect(Parameter, "Change"),
-                        (Est.Error * income[Year, "sd"] + income[Year, "mean"])/1000,
+                        (Est.Error * income[Year, "sd"])/1000,
                         Est.Error),
@@
     as_tibble(rownames = "terms") %>%
-    transmute(terms, coef = paste0(round(Estimate, 2), " (±", round(Est.Error, 3), ")")) %>%
+    transmute(terms, coef = paste0(round(Estimate, 2), " (", round(Est.Error, 3), ")")) %>%
     spread(terms, coef)},
@@
   add_header(top = TRUE,
-             values = c(b0_pw = "Piecewise-linear (±se)",
-                        b1_pw = "Piecewise-linear (±se)",
-                        b2_pw = "Piecewise-linear (±se)",
-                        Omega = "Piecewise-linear (±se)",
-                        b0_log = "Log-linear (±se)",
-                        b1_log = "Log-linear (±se)",
-                        b0_lin = "Linear (±se)",
-                        b1_lin = "Linear (±se)")) %>%
+             values = c(b0_pw = "Piecewise-linear (se)",
+                        b1_pw = "Piecewise-linear (se)",
+                        b2_pw = "Piecewise-linear (se)",
+                        Omega = "Piecewise-linear (se)",
+                        b0_log = "Log-linear (se)",
+                        b1_log = "Log-linear (se)",
+                        b0_lin = "Linear (se)",
+                        b1_lin = "Linear (se)")) %>%
   merge_h(part = "header") %>%
@@
   filter(term == "dollars") %>%
-  mutate(coefficient = paste0(round(estimate, 3), " (±", round(std.error, 3), ")")) %>%
-  select(Year, `linear (±se)` = coefficient)
+  mutate(coefficient = paste0(round(estimate, 3), " (", round(std.error, 3), ")")) %>%
+  select(Year, `linear (se)` = coefficient)
 
@@
   filter(term == "log(dollars)") %>%
-  mutate(coefficient = paste0(round(estimate, 3), " (±", round(std.error, 3), ")")) %>%
-  select(Year, `log-linear (±se)` = coefficient)
+  mutate(coefficient = paste0(round(estimate, 3), " (", round(std.error, 3), ")")) %>%
+  select(Year, `log-linear (se)` = coefficient)
 
@@
     estimate_post = estimate_U1.dollars + estimate_dollars,
-    `pre-slope (±se)` = paste0(
+    `pre-slope (se)` = paste0(
       round(estimate_dollars, 3), " ±", round(std.error_dollars, 3)),
-    `post-slope (±se)` = paste0(
+    `post-slope (se)` = paste0(
       round(estimate_post, 3), " ±", round(std.error_U1.dollars, 3))
     ) %>%
-  select(Year, `pre-slope (±se)`, `post-slope (±se)`)
+  select(Year, `pre-slope (se)`, `post-slope (se)`)
 
@@
   mutate(term = "psi.dollars") %>%
-  mutate(coefficient = paste0(round(`Est.`, 2)*10, " (±", round(`St.Err`, 2)*10, ")")) %>%
+  mutate(coefficient = paste0(round(`Est.`, 2)*10, " (", round(`St.Err`, 2)*10, ")")) %>%
   select(Year, `change point ($000s ±se)` = coefficient) %>%

### 2021-10-27 src/draft-SSM2.Rmd
Final response to Reviewer 1. Closes issue #27

--- 2021-06-24 src/draft-SSM1.Rmd
@@
 2. ARC Centre of Excellence for Children and Families over the Life Course
-3. School of Economics, University of Technology, NSW, Australia
+3. Economic Discipline Group, University of Technology Sydney, NSW, Australia
 4. Institute of Labor Economics (IZA), Bonn, Germany
@@
 | Draft      | `r format(Sys.time(), '%d %B, %Y')` |
-| Words      | 4704                                |
+| Words      | 4688                                |
 | Tables     | 0                                   |
@@
 A fundamental question for society is how much happiness does a dollar buy?
-The accepted view among economists and psychologists is that money and happiness increase together up to a point, after which there is little further gain from increasing income.
-The location of this change point was reportedly USD\$75K in 2008, which was substantially more than the median income in the same year (USD\$52K), indicating the happiness or wellbeing of the poorest majority of the population was tied to marginal changes in income while the wellbeing of a richer minority was not.
-Thus the location of the change point has implications for the distribution of wellbeing between the rich and poor.
-Inequalities in wellbeing are increasingly relevant due to the growing recognition that increasing income does not neccesarily lead to equal changes in wellbeing.
-In this paper we estimate the relationship between income and subjective wellbeing over a 19-year period focusing on whether this change point has shifted over time, and explore its implications for inequalities in wellbeing.
-To date there has been no investigation as to whether the change point has increased or decreased over time so we formally tested for the presence of a change point as well temporal changes in the relationship between income and affective wellbeing (happiness), and income and cognitive wellbeing (life satisfaction), using household economic data from Australia between 2001-2019.
-The results revealed the change point between affective wellbeing and income has increased over those 19 years faster than inflation (i.e., cost of living).
+The accepted view among economists and psychologists is that income has diminishing marginal returns on happiness: money and happiness increase together up to a point after which there is relatively little further gain.
+In this paper we estimate the relationship between income and subjective wellbeing over a 19-year period focusing on where the greatest change in the marginal return on income occurs and whether this change point has shifted over time.
+We formally test for the presence of a change point as well temporal changes in the relationship between income and affective wellbeing (happiness), and income and cognitive wellbeing (life satisfaction), using household economic data from Australia between 2001-2019.
+The results indicate that the change point between affective wellbeing and income has increased over those 19 years faster than inflation (i.e., cost of living).
 This suggests that inequalities in income may be driving increasing _inequities_ in happiness between the rich and the poor, with implications for health and recent government policy-goals to monitor and improve wellbeing.
@@
 **keywords:** Subjective wellbeing, household income, HILDA
+
+<br><br>
+
+<br><br>
+
+<br><br>
+
+## Highlights Statement
+
+\
+
+- We investigate how the relationship between income and subjective wellbeing (happiness and life satisfaction) has changed over time
+- Detailed wellbeing data from Australia over an economically stable 19-year period provided an undisturbed view of the changes
+- Happiness was less responsive to income after a certain income level, and this level had almost doubled since 2001
+- Conversely the relationship between income and life-satisfaction had remained relatively constant over the same period
+- A steep association between happiness and income has become more prevalent this millennium, and likely to drive inequities in wellbeing between the rich and poor
 
@@
 
-## Highlights Statement
-
-- We investigate how the relationship between income and subjective wellbeing (affective and cognitive) has changed over time, using a dataset with highly detailed measurements of each
-- Nineteen years of data from a well-developed, economically stable G20 nation is examined (Australia), allowing a relatively undisturbed view of the changes without major economic shocks
-- We are able to show for the first time that the income level at which affective wellbeing (happiness) is less responsive to income has risen since 2001, resulting in a greater inequality in the distribution of affective wellbeing between the rich and the poor
-- Conversely the linear relationship between income and cognitive wellbeing (life-satisfaction) has remained relatively constant over this period
-- The association between happiness and income is becoming more prevalent in the population this millennium, and likely to drive inequities in wellbeing across the economy
+<br><br>
 
@@
 
-Subjective wellbeing is not a unitary entity [@diener2017findings]; studies typically distinguish between life satisfaction, the cognitive appraisal of one’s own accomplishments, and affective wellbeing, one’s prevailing affective state, emotional mood, or everyday experience of happiness [e.g., @kettlewell2020differential].
+Subjective wellbeing is not a unitary entity [@diener2017findings].
+Psychology studies typically distinguish between life satisfaction: the cognitive appraisal of one’s own accomplishments; and affective wellbeing: one’s prevailing affective state, emotional mood, or everyday experience of happiness [e.g., @kettlewell2020differential].
 While measures of these distinct concepts are correlated, the distinction seems to be critical in understanding the impact of income on wellbeing [@howell2008relation].
 For instance, we have recently observed that a major financial gain substantially improved cognitive wellbeing for individuals without much impact on their affective wellbeing [@kettlewell2020differential].
-In a now classic paper, Kahneman & Deaton [-@kahneman2010high] showed that affective wellbeing increased with log household income up to a point (USD75,000), but after that, increasing income had no further effect (i.e., "satiation").
-Conversely cognitive wellbeing continued to increase with (log) income beyond USD75,000.
-Studies since then have confirmed that while both affective and cognitive wellbeing increase with income, income is more strongly related to cognitive- than affective-wellbeing [@diener2013rising; @diener2017findings; @howell2008relation].
+In a now classic paper, Kahneman & Deaton [-@kahneman2010high] showed that affective wellbeing increased with log household income up to a point (USD75,000), but after that the slope between income and wellbeing was effectively zero: increasing income had no further effect (i.e., "income satiation").
+Conversely cognitive wellbeing continued to increase with log income beyond USD75,000.
+Studies since then have confirmed that while both affective and cognitive wellbeing increase with income, (log) income is more strongly related to cognitive than affective wellbeing [@diener2013rising; @diener2017findings; @howell2008relation].
 
-Fundamentally, a difference in slope between affective wellbeing and income for the rich and poor represents an unacknowledged source of inequity (i.e., unfairness) in the distribution of wellbeing (i.e., happiness) in the economy.
-For instance, the "satiety" point of USD75,000 in 2008 was substantially more than the US median income of USD52,000 in the same year, indicating that the affective wellbeing of the poorest majority of the US population was tied to marginal changes in income while the affective wellbeing of a richer minority was not.
-While much commentary has focused on the zero slope at high income levels reported by Kahneman & Deaton, the positive slope at lower income levels means wellbeing is distributed unequally over those income ranges.
-Moreover, the location of the change point in slope has implications for the distribution of wellbeing over the entire income range.
-For instance, an increase in the change point (i.e., a right shift on the income distribution) indicates more inequality in wellbeing between the rich and poor.
-Conversely, a decrease (left shift) represents a more equal distribution of wellbeing across income levels.
-To our knowledge, no studies to date have examined whether the location of the change point has shifted over time, but such shifts may help us understand how the distribution of wellbeing has been changing for different parts of the economy.
+Fundamentally, any difference in slope between affective wellbeing (i.e., happiness) and income for the rich and poor represents an unacknowledged source of inequity (i.e., unfairness) in the distribution of wellbeing in the economy.
+While much commentary has focused on the zero slope at high income levels reported by Kahneman & Deaton, any steeper slope at lower income levels (relative to high income levels) means wellbeing is distributed more unequally among people with low incomes.
+For instance, the "satiety" point at USD75,000 in 2008 reported by Kahneman & Deaton was substantially more than the US median income of USD52,000 in the same year, indicating that the affective wellbeing of the poorest majority of the US population was tied to marginal changes in income while that of the richer minority was not.
 
-While studies agree that affective wellbeing tends to increase with income, the exact form of the functional relationship between affective wellbeing and income is contested: some studies show the presence of a zero-slope change point after which increasing income has no further effect [@kahneman2010high; @jebb2018happiness]; other studies have found that affective wellbeing is approximately log-linear [@killingsworth2021experienced; @sacks2012new; @twenge2020expanding].
-Few studies have formally tested the functional form of the relationship, relying instead upon p-value comparisons after dividing the data into arbitrary bins (i.e., deciles, quintiles, quartiles), which can have salient effects on the functional form observed.
-When p-values have been determined, they have not evaluated the probability of a change point given the data, but rather the probability of the data assuming no change - which is not a direct test.
-binning the data (i.e., over income deciles or quintiles) and comparing adjacent bins either visually or with null hypothesis tests which do not provide evidence for a change point.
-In this regard, no studies to our knowledge have ever directly tested for the presence of a change point between wellbeing and income using standard modelling approaches.
-For this reason, a formal comparison between log-linear, linear and a change point function (i.e., piecewise) in the relationship between affective wellbeing and income is warranted.
+While studies agree the marginal returns of income on affective wellbeing tend to diminish, the exact form of the functional relationship between affective wellbeing and income is contested: some studies show the presence of a zero-slope change point after which increasing income has no further effect [@kahneman2010high; @jebb2018happiness]; other studies have found the relationship is approximately log-linear with no discontinuity in slope [@killingsworth2021experienced; @sacks2012new; @twenge2020expanding].
+Rather than assume a log-linear relationship to characterise income and wellbeing, we formally test and compare different functional forms (linear, log-linear and piecewise-linear) to identify the point of greatest change in the marginal effect of income (including but not limited to a zero-slope change); and then determine whether that point is shifting over time.
+For instance, a right shift of the change point on the income distribution would represent an increase in the income level, and thus indicate more inequality in wellbeing among the poor than the rich.
+Conversely, a decrease (left shift) would represent a more equal distribution of wellbeing across income levels.
 
 Inequalities in the distribution of wellbeing are increasingly relevant to governments and policy-makers due to the growing recognition that increasing income does not necessarily lead to equal changes in wellbeing [@clark2018four; @frijters2020happy].
-Even prior to COVID-19, the World Gallup Poll has observed that affective wellbeing has decreased over the past decade in western Europe, North America, Australia and New Zealand, despite increases in income in the same countries [@sachs2019world].
+Even prior to COVID-19, the World Gallup Poll has observed that affective wellbeing has decreased over the past decade in western Europe, North America, Australia and New Zealand, despite increases in average income in the same countries [@sachs2019world].
 However to date there has been little investigation of whether the relationship between income and affective wellbeing has changed over time, which may contribute to these trends.
@@
 #### 2.1 Income
-We used household after-tax income as the indicator of income and economic security [e..g, @kahneman2010high].
+We used household after-tax income as the indicator of income and economic security [e.g., @kahneman2010high].
 The ‘real household annual disposable income’ was calculated from the self-reported combined income of all household members after receipt of government pensions and benefits and deduction of income taxes in the financial year ended 30th June of the year of the wave (e.g., 2001 in wave 1).
@@
 We modelled the relationship between income and each wellbeing variable (cognitive and affective) using a linear model, log-linear model, and a piecewise-linear model with a single change point as a free parameter estimated from the data.
-The piecewise-linear model was chosen as the simplest extension of a linear model which can identify a change point (inflection) in the relationship between wellbeing and income.
+The piecewise-linear model was chosen as the simplest extension of a linear model which can identify a change point (discontinuity in slope) in the relationship between wellbeing and income.
 For modelling, both measures of wellbeing and income were rescaled with a mean of zero and a SD of 1 (z-scores) for each year.
-As each year was modelled separately this does not affect the results.
-Importantly, the coefficient for the change point in each year was returned to the original scale and transformed to real 2019 dollars when reported.
+As each year was modelled separately this does not affect the results, and the model parameter estimates were returned to their raw score units (e.g., real 2019 dollars) in figures and tables.
 
@@
 Full-time students were removed, as well as individuals with an annual household disposable income that was negative or indicated as _topcoded_ by the University of Melbourne.
-The results of sensitivity analyses to determine the effect of the covariates and exclusions on the change point parameters are included in Supplementary Materials (Figure S2, S3, S4).
+The results of sensitivity analyses to determine the effect of the covariates and exclusions on the change point parameters are included in Supplementary Materials (Figures S2, S3, and S4).
 
@@
 
-The evidence from model selection revealed the piecewise-linear fit of _affective_- was superior to a log-linear or linear fit in each year with a penalised WAIC score credibly (95% interval) smaller than the log-linear fits for all but three of the nineteen years (and for all 19 years for the linear fits, see Supplementary Materials, Figure S1).
+The evidence from model selection revealed the piecewise-linear fit of _affective_ wellbeing was superior to a log-linear or linear fit in each year with a penalised WAIC score credibly (95% interval) smaller than the log-linear fits for all but three of the nineteen years (and for all 19 years for the linear fits, see Supplementary Materials, Figure S1).
 There was little evidence that any of the three models were reliably or credibly superior in representing the association between income and cognitive-wellbeing over the same period, albeit the linear model was better on average between 2002-2004 and 2007-2010 while the piecewise-linear model was better the remaining years.
-Overall the model comparison suggested that affective- and cognitive wellbeing have distinct relationships with household income; affective wellbeing increases with income more rapidly at lower household income levels than higher income levels with a distinct change point in each year, while cognitive wellbeing tends to increase with income, most likely in a linear fashion.
+Overall the model comparison suggested that affective and cognitive wellbeing have distinct relationships with household income; affective wellbeing increases with income more rapidly at lower household income levels than higher income levels with a distinct change point in each year, while cognitive wellbeing tends to increase with income, most likely in a linear fashion.
 Others have also noted cognitive wellbeing (life satisfaction) has a stronger linear relationship with income than affective wellbeing and income [@howell2008relation], and so the results of our formal comparison provides some support for this.
@@
 ```{r get_posteriors}
-income_params <- read_rds("../results/gh9_hhoecd_summary.RDS")
+# income_params <- read_rds("../results/gh9_hhoecd_summary.RDS")
 fits <- read_rds("../results/gh9_hhoecd_0cov_tight.RDS")
 
+# Recover the parameters from standardization to return raw score units
+std_params <- list()
+
+for (i in c(2001:2019)) {
+
+  hilda_data %>%
+    filter(
+      year == i,
+      !student,
+      hh_disp_inc_eq >= 0,
+      !top_hifdip
+    ) %>%
+    transmute(
+      sgh9 = scale(gh9),
+      shh_disp_inc_eq = scale(hh_disp_inc_eq),
+    ) %>% 
+    na.omit() -> df
+
+  income_sd_year <- attr(df$shh_disp_inc_eq, "scaled:scale")
+  income_m_year <- attr(df$shh_disp_inc_eq, "scaled:center")
+  gh9_sd_year <- attr(df$sgh9, "scaled:scale")
+  gh9_m_year <- attr(df$sgh9, "scaled:center")
+
+  std_params[[paste0("fit", i)]] <- c(`mean_income` = income_m_year, 
+                       `sd_income` = income_sd_year,
+                       `mean_gh9` = gh9_m_year,
+                       `sd_gh9` = gh9_sd_year)
+
+}
+
 posteriors <- map_dfr(fits, as.data.frame, .id = "fit") %>%
-  left_join(bind_rows(income_params, .id = "fit")) %>%
-  mutate(omega_dollars = (b_omega_Intercept * sd + mean)/1000) %>%
+  left_join(bind_rows(std_params, .id = "fit")) %>%
+  transmute(
+    fit,
+    omega_dollars = (b_omega_Intercept * sd_income + mean_income)/1000,
+    b0 = b_b0_Intercept * sd_gh9 + mean_gh9,
+    b1 = b_b1_Intercept * sd_gh9,
+    b2 = b_b2_Intercept * sd_gh9
+    ) %>%
   as_tibble()
@@
   omega_dollars = "Change point ($000s)",
-  b_b0_Intercept = paste0("Intercept (\u03b2", "0)"),
-  b_b1_Intercept = paste0("Pre-slope (\u03b2", "1)"),
-  b_b2_Intercept = paste0("Post-slope (\u03b2", "2)")
+  b0 = paste0("Intercept (\u03b2", "0)"),
+  b1 = paste0("Pre-slope (\u03b2", "1)"),
+  b2 = paste0("Post-slope (\u03b2", "2)")
 )
@@
 posteriors %>%
-  select(fit, contains("b0"), contains("b1"), contains("b2"), omega_dollars) %>%
   gather(key = "param", val = "sample", -fit) %>%
@@
 ```
-Figure 2 legend: Posterior distributions of the change point parameter representing the location in real household income (real 2019 dollars), as well as the intercept, pre-slope and post-slope parameters in SD units (affective wellbeing).
+Figure 2 legend: Posterior distributions of the change point parameter representing the location in real household income (real 2019 dollars), as well as the intercept, pre-slope and post-slope parameters in happiness units (0-100).
 Horizontal bar represents the 95% credible region and the solid point indicates the expected value (median) of each distribution.
@@
 
-We found the marginal effect of household income on affective- and cognitive-wellbeing was positive but quite different: Affective wellbeing increased rapidly up to a point after which higher levels of income were associated with less improvement, while cognitive wellbeing tended to increase linearly or constantly with income.
-Other studies have also reported that income has distinct effects on cognitive and affective wellbeing [@howell2008relation], with a zero-slope change point in affective wellbeing after which increasing income has little further effect [@kahneman2010high].
+We found the marginal effect of household income on affective- and cognitive-wellbeing was positive but quite different: Affective wellbeing increased rapidly up to a point after which higher levels of income were associated with relatively less improvement, while cognitive wellbeing tended to increase linearly with income.
+Other studies have also reported that income has distinct effects on cognitive and affective wellbeing [@howell2008relation], and our results are consistent with that distinction.
+We also found a change point in the marginal returns of income on affective wellbeing in each year.
+Some influential studies have found a zero-slope change point in affective wellbeing after which increasing income had no further effect [@kahneman2010high].
 In our case we did not observe a zero-slope change point ("satiety" point), as the post-slope for each year was credibly above zero (Figure 2); but the change point did represent a substantial decrease in dependency between income and affective wellbeing for those on higher income levels.
-Our novel finding was revealing the change point increased faster than both inflation and median household income between 2001 and 2019.
+Our novel finding was the change point increased faster than both inflation and median household income between 2001 and 2019.
 For the first time we have shown there has been a temporal shift in the change point between income and affective wellbeing over the 19 year period, such that affective wellbeing has become more dependent on income for more people - especially the poor and middle-class.
@@
 Our results emphasize the slope over income is not constant.
-Other researchers have argued that affective wellbeing is a constant (albiet decreasing) log-linear function of income [@killingsworth2021experienced; @sacks2012new; @twenge2020expanding].
+Other researchers have also argued that affective wellbeing is a continuous log-linear function of income [@killingsworth2021experienced; @sacks2012new; @twenge2020expanding], which our results do not support and instead suggest there is a point (at least one) of discontinuity in the change of slope over income.
 We also show the change point has increased over time.
@@
 Nevertheless, consistent with Twenge et al, our results imply that the difference in happiness between the rich and poor has been increasing over time (e.g., Figure 6 and S8).
-Moreover our piecewise-linear model reveals the income level around which this difference appears.
+Moreover our model reveals the income level around which the greatest difference appears.
 

### 2021-10-27 src/supplementary-SSM2.Rmd
Final response to Reviewer 1. Closes issue #27

--- 2021-06-24 src/supplementary-SSM1.Rmd
@@
 
-<br>
+<br><br>
+
+#### Sensitivity tests of model parameters
+
+```{r std_params}
+std_params <- list()
+
+for (i in c(2001:2019)) {
+
+  hilda_data %>%
+    filter(
+      year == i,
+      !student,
+      hh_disp_inc_eq >= 0,
+      !top_hifdip
+    ) %>%
+    transmute(
+      sgh9 = scale(gh9),
+      shh_disp_inc_eq = scale(hh_disp_inc_eq),
+    ) %>% 
+    na.omit() -> df
+
+  income_sd_year <- attr(df$shh_disp_inc_eq, "scaled:scale")
+  income_m_year <- attr(df$shh_disp_inc_eq, "scaled:center")
+  gh9_sd_year <- attr(df$sgh9, "scaled:scale")
+  gh9_m_year <- attr(df$sgh9, "scaled:center")
+
+  std_params[[paste0("fit", i)]] <- c(`mean_income` = income_m_year, 
+                       `sd_income` = income_sd_year,
+                       `mean_gh9` = gh9_m_year,
+                       `sd_gh9` = gh9_sd_year)
+
+}
+
+std_params <- bind_rows(std_params, .id = "fit")
+```
 
@@
 nl_fits = read_rds("../results/gh9_hhoecd_0cov_tight.RDS")
-
 nl_fits_5cov = read_rds("../results/gh9_hhoecd_5cov_tight.RDS")
 
-income_params <- read_rds("../results/gh9_hhoecd_summary.RDS") %>%
-  bind_rows(.id = "fit")
-
-posteriors_5c <- map_dfr(nl_fits_5cov, as.data.frame, .id = "fit") %>%
-  left_join(income_params) %>%
-  mutate(omega_dollars = (b_omega_Intercept * sd + mean)/1000) %>%
+summarise_posteriors <- function(.fits, .params = std_params, .model = NULL) {
+
+  posteriors <- map_dfr(.fits, as.data.frame, .id = "fit") %>%
+    left_join(.params, by = "fit") %>%
+    transmute(
+      fit,
+      omega_dollars = (b_omega_Intercept * sd_income + mean_income)/1000,
+      b0 = b_b0_Intercept * sd_gh9 + mean_gh9,
+      b1 = (b_b1_Intercept * sd_gh9)/10, # change per $1000
+      b2 = (b_b2_Intercept * sd_gh9)/10  # change per $1000
+    ) %>%
     as_tibble()
@@
     omega_dollars = "Change point ($000s)",
-  b_b0_Intercept = paste0("Intercept (\u03b2", "0)"),
-  b_b1_Intercept = paste0("Pre-slope (\u03b2", "1)"),
-  b_b2_Intercept = paste0("Post-slope (\u03b2", "2)")
+    b0 = paste0("Intercept (\u03b2", "0)"),
+    b1 = paste0("Pre-slope (\u03b2", "1)"),
+    b2 = paste0("Post-slope (\u03b2", "2)")
   )
 
-posteriors_5c %>%
-  select(fit, contains("b0"), contains("b1"), contains("b2"), omega_dollars) %>%
+  posteriors %>%
     gather(key = "param", val = "sample", -fit) %>%
@@
     ) %>%
-  mutate(model = "conditional") -> parameter_summary_5c
-
-posteriors_0c <- map_dfr(nl_fits, as.data.frame, .id = "fit") %>%
-  left_join(income_params) %>%
-  mutate(omega_dollars = (b_omega_Intercept * sd + mean)/1000) %>%
-  as_tibble()
-
-posteriors_0c %>%
-  select(fit, contains("b0"), contains("b1"), contains("b2"), omega_dollars) %>%
-  gather(key = "param", val = "sample", -fit) %>%
-  mutate(
-    year = extract_numeric(fit),
-    param = recode_factor(param, !!!param_key)
-  ) %>% 
-  group_by(param, year) %>%
-  summarise(
-    low = quantile(sample, probs = 0.05, na.rm=T),
-    expected = quantile(sample, probs = 0.5, na.rm=T),
-    upp = quantile(sample, probs = 0.95, na.rm=T)
-  ) %>%
-  mutate(model = "unconditional") -> parameter_summary_0c
+    mutate(model = .model) -> parameter_summary
+
+  return(parameter_summary)
+
+}
+
+parameter_summary_5c <- summarise_posteriors(.fits = nl_fits_5cov, 
+                                             .model = "conditional")
+
+parameter_summary_0c <- summarise_posteriors(.fits = nl_fits,
+                                             .model = "unconditional")
 
@@
   ft_theme(base_size = 12) +
-  scale_color_manual(values = c("#C79999", "grey65")) +
+  scale_color_manual(values = c("#A25050", "grey65")) +
   theme(legend.position = "none",
@@
 This is performed by the University of Melbourne to preserve anonymity, and we excluded people marked as topcoded in the main analysis (highest _n_ in any year was _n_ = 103 in 2019).
-Here we present the parameter estimates of the affective wellbeing models with topcodes included, and compare them with parameters from models which exclude them.
-
-##### Figure S3. Topcoded model parameters  
+Here we present the parameter estimates of the affective wellbeing models with topcodes included, and compare them with parameters from models which exclude them (as in the main text).
+
+##### Figure S3. Model parameters including topcoded individuals 
 ```{r figure_s3, fig.width = 6, fig.height = 8, dpi=300}
@@
 
-topcode_income_params <- read_rds("../results/fits/inc_topcodes/gh9_hhoecd_data.RDS") %>%
-  bind_rows(.id = "fit")
-
-posteriors_topcodes <- map_dfr(nl_fits_topcodes, as.data.frame, .id = "fit") %>%
-  left_join(topcode_income_params) %>%
-  mutate(omega_dollars = (b_omega_Intercept * sd + mean)/1000) %>%
-  as_tibble()
-
-posteriors_topcodes %>%
-  select(fit, contains("b0"), contains("b1"), contains("b2"), omega_dollars) %>%
-  gather(key = "param", val = "sample", -fit) %>%
-  mutate(
-    year = extract_numeric(fit),
-    param = recode_factor(param, !!!param_key)
-  ) %>% 
-  group_by(param, year) %>%
-  summarise(
-    low = quantile(sample, probs = 0.05, na.rm=T),
-    expected = quantile(sample, probs = 0.5, na.rm=T),
-    upp = quantile(sample, probs = 0.95, na.rm=T)
-  ) %>%
-  mutate(model = "topcodes") -> parameter_summary_topcodes
-
-parameter_summary_0c %>%
-  mutate(model = "no topcodes") -> parameter_summary_excl
+# topcode_income_params <- read_rds("../results/fits/inc_topcodes/gh9_hhoecd_data.RDS") %>%
+#   bind_rows(.id = "fit")
+
+topcode_params <- list()
+
+for (i in c(2001:2019)) {
+
+  hilda_data %>%
+    filter(
+      year == i,
+      !student,
+      hh_disp_inc_eq >= 0
+    ) %>%
+    transmute(
+      sgh9 = scale(gh9),
+      shh_disp_inc_eq = scale(hh_disp_inc_eq),
+    ) %>% 
+    na.omit() -> df
+
+  income_sd_year <- attr(df$shh_disp_inc_eq, "scaled:scale")
+  income_m_year <- attr(df$shh_disp_inc_eq, "scaled:center")
+  gh9_sd_year <- attr(df$sgh9, "scaled:scale")
+  gh9_m_year <- attr(df$sgh9, "scaled:center")
+
+  topcode_params[[paste0("fit", i)]] <- c(`mean_income` = income_m_year, 
+                       `sd_income` = income_sd_year,
+                       `mean_gh9` = gh9_m_year,
+                       `sd_gh9` = gh9_sd_year)
+
+}
+
+topcode_params <- bind_rows(topcode_params, .id = "fit")
+
+parameter_summary_topcodes <- summarise_posteriors(nl_fits_topcodes,
+                                                   .params = topcode_params,
+                                                   .model = "topcodes")
+
+parameter_summary_excl <- parameter_summary_0c %>% mutate(model = "no topcodes")
 
@@
   ft_theme(base_size = 12) +
-  scale_color_manual(values = c("grey65", "#C79999")) +
+  scale_color_manual(values = c("grey65", "#A25050")) +
   theme(legend.position = "none",
@@
 For example, two people living alone in two separate households typically incur higher living costs in total than two people living in one household.
-Here we present the parameter estimates of the affective wellbeing models without adjustment for household size and with adjustment.
+Here we present the parameter estimates of the affective wellbeing models without adjustment for household size, and with adjustment (as in the main text).
 
@@
 
-posteriors_unadjsize <- map_dfr(nl_fits_unadjsize, as.data.frame, .id = "fit") %>%
-  left_join(income_params) %>%
-  mutate(omega_dollars = (b_omega_Intercept * sd + mean)/1000) %>%
-  as_tibble()
-
-posteriors_unadjsize %>%
-  select(fit, contains("b0"), contains("b1"), contains("b2"), omega_dollars) %>%
-  gather(key = "param", val = "sample", -fit) %>%
-  mutate(
-    year = extract_numeric(fit),
-    param = recode_factor(param, !!!param_key)
-  ) %>% 
-  group_by(param, year) %>%
-  summarise(
-    low = quantile(sample, probs = 0.05, na.rm=T),
-    expected = quantile(sample, probs = 0.5, na.rm=T),
-    upp = quantile(sample, probs = 0.95, na.rm=T)
-  ) %>%
-  mutate(model = "unadjusted") -> parameter_summary_unadjsize
+parameter_summary_unadjsize <- summarise_posteriors(nl_fits_unadjsize, 
+                                                    .model = "unadjusted")
 
@@
 
-income_params <- read_rds("../results/gh9_hhoecd_summary.RDS") %>%
-  bind_rows(.id = "fit")
-
-posteriors_pos <- map_dfr(gh9pos.fits, as.data.frame, .id = "fit") %>%
-  left_join(income_params) %>%
-  mutate(omega_dollars = (b_omega_Intercept * sd + mean)/1000) %>%
-  as_tibble()
-
-param_key = c(
-  omega_dollars = "Change point ($000s)",
-  b_b0_Intercept = paste0("Intercept (\u03b2", "0)"),
-  b_b1_Intercept = paste0("Pre-slope (\u03b2", "1)"),
-  b_b2_Intercept = paste0("Post-slope (\u03b2", "2)")
-)
-
-posteriors_pos %>%
-  select(fit, contains("b0"), contains("b1"), contains("b2"), omega_dollars) %>%
-  gather(key = "param", val = "sample", -fit) %>%
-  mutate(
-    year = extract_numeric(fit),
-    param = recode_factor(param, !!!param_key)
-  ) %>% 
-  group_by(param, year) %>%
-  summarise(
-    low = quantile(sample, probs = 0.05, na.rm=T),
-    expected = quantile(sample, probs = 0.5, na.rm=T),
-    upp = quantile(sample, probs = 0.95, na.rm=T)
-  ) %>%
-  mutate(model = "positive") -> parameter_summary_pos
-
-posteriors_neg <- map_dfr(gh9neg.fits, as.data.frame, .id = "fit") %>%
-  left_join(income_params) %>%
-  mutate(omega_dollars = (b_omega_Intercept * sd + mean)/1000) %>%
-  as_tibble()
-
-posteriors_neg %>%
-  select(fit, contains("b0"), contains("b1"), contains("b2"), omega_dollars) %>%
-  gather(key = "param", val = "sample", -fit) %>%
-  mutate(
-    year = extract_numeric(fit),
-    param = recode_factor(param, !!!param_key)
-  ) %>% 
-  group_by(param, year) %>%
-  summarise(
-    low = quantile(sample, probs = 0.05, na.rm=T),
-    expected = quantile(sample, probs = 0.5, na.rm=T),
-    upp = quantile(sample, probs = 0.95, na.rm=T)
-  ) %>%
-  mutate(model = "negative") -> parameter_summary_neg
+parameter_summary_pos <- summarise_posteriors(gh9pos.fits, .model = "positive")
+parameter_summary_neg <- summarise_posteriors(gh9neg.fits, .model = "negative")
 
@@
 
-#### Parameter and fit summary statistics
+#### Sampling and fit statistics
+
+Below we provide sampling statistics (e.g., $\hat{R}$), as well as prior predictions and residual plots.
 
@@
 
-For all models estimated with RStan presented in this report we confirmed the absence of divergent transitions; the $\hat{R}$ approached 1 indicating convergance; and the effective sample size (ESS) for each parameter was _n_ > 300, which is a conservative threshold for estimation.
-
-Table S2 below shows the fit and summary statistics for each piecewise-linear model fit of affective wellbeing and income in each year.
-Estimates for the intercept, pre-slope and post-slope are in year-wise SD units while the change-point is presented in real 2019 dollars.
-
-##### Table S2. Parameter estimates and fit statistics for piecewise-linear fit of affective wellbeing and income
+For all models estimated with RStan presented in this report we confirmed the absence of divergent transitions; the $\hat{R}$ for each parameter approached 1 indicating convergance; and the effective sample size (ESS) for each parameter was _n_ > 300, which is a conservative minimum threshold for estimation.
+
+Table S2 below shows the sampling statistics for each piecewise-linear model fit of affective wellbeing and income in each year.
+
+##### Table S2. Sampling statistics for piecewise-linear fit of affective wellbeing and income
 ```{r table_s2, ft.align="left"}
@@
 
-income_params %>%
+params_df <- std_params %>% 
   mutate(fit = str_remove(fit, "fit")) %>%
-  column_to_rownames("fit") -> income
-
-param_key = c(
-  omega_Intercept = "Change point ($000s)",
-  b0_Intercept = paste0("Intercept (\u03b2", "0)"),
-  b1_Intercept = paste0("Pre-slope (\u03b2", "1)"),
-  b2_Intercept = paste0("Post-slope (\u03b2", "2)")
-)
-
-# param_key = c(
-#   b0_Intercept = "Intercept",
-#   b1_Intercept = "Pre-slope",
-#   b2_Intercept = "Post-slope",
-#   omega_Intercept = "Change-point ($000s)"
-# )
+  column_to_rownames("fit")
 
@@
         .f = ~{
-          fit_summary <- summary(.x, prob = 0.95)
+          fit_summary <- summary(.x)
 
@@
             bind_rows(sigma_summary) %>%
-            select(Parameter, everything())
+            select(Parameter, Bulk_ESS, Tail_ESS)
 
@@
     Year = str_remove(Year, "fit"),
-    Parameter = recode_factor(Parameter, !!!param_key),
-    Estimate = if_else(str_detect(Parameter, "Change"),
-                       (Estimate * income[Year, "sd"] + income[Year, "mean"])/1000,
-                       Estimate),
-    Est.Error = if_else(str_detect(Parameter, "Change"),
-                        (Est.Error * income[Year, "sd"])/1000,
-                        Est.Error),
-    `l-95% CI` = if_else(str_detect(Parameter, "Change"),
-                        (`l-95% CI` * income[Year, "sd"] + income[Year, "mean"])/1000,
-                        `l-95% CI`),
-    `u-95% CI` = if_else(str_detect(Parameter, "Change"),
-                         (`u-95% CI` * income[Year, "sd"] + income[Year, "mean"])/1000,
-                         `u-95% CI`)
-    ) %>%
+  ) %>%
+  pivot_wider(names_from = Parameter, 
+              values_from = c(Bulk_ESS, Tail_ESS)
+  ) %>% 
+  select(-one_of(c("Bulk_ESS_sigma", "Tail_ESS_sigma"))) %>%
   flextable() %>%
-  merge_v(j = 1) %>%
-  valign(j = 1, valign = "top") %>%
-  bg(i = ~Year %in% seq(2001, 2019, 2), 
-     bg = "grey90", 
-     part = "body") %>%
-  colformat_double(j = 3:6, digits= 3) %>%
-  colformat_double(j = 7, digits= 1) %>%
-  footnote(i = 1, j = 7,
-           value = as_paragraph(
-             c("Rhat is the potential scale reduction factor on split chains")),
-           ref_symbols = c("a"),
-           part = "header") %>%
-  footnote(i = 1, j = c(8:9),
-           value = as_paragraph(
-             c("Bulk_ESS and Tail_ESS are effective sample size (ESS) measures")),
-           ref_symbols = c("b"),
-           part = "header") %>%
+  set_header_labels(
+    values = c(Bulk_ESS_b0_Intercept = paste0("\u03b2", "0"),
+               Bulk_ESS_b1_Intercept = paste0("\u03b2", "1"),
+               Bulk_ESS_b2_Intercept = paste0("\u03b2", "2"),
+               Bulk_ESS_omega_Intercept = "\u03C9",
+               Tail_ESS_b0_Intercept = paste0("\u03b2", "0"),
+               Tail_ESS_b1_Intercept = paste0("\u03b2", "1"),
+               Tail_ESS_b2_Intercept = paste0("\u03b2", "2"),
+               Tail_ESS_omega_Intercept = "\u03C9")
+    ) %>%
+  add_header(
+    top = TRUE,
+    values = c(Bulk_ESS_b2_Intercept = "Bulk ESS",
+               Tail_ESS_b2_Intercept = "Tail ESS")) %>%
+  merge_h(part = "header") %>%
+  width(j = 1, width = 0.5) %>%
+  width(j = 2:9, width = 0.625) %>%
   fontsize(size = 9, part = "all") %>%
-  width(j = 1, width = .5) %>%
-  width(j = 2, width = 1) %>%
-  width(j = c(3:7), width = .7) %>%
-  width(j = c(8, 9), width = .8)
-```
-
-<br><br>
-
-The coeficients (in SD units) for the piecewise-linear, log-linear and linear Bayesian models are shown in the table below.
-
-##### Table S3. Bayesian coefficients from linear, log-linear and piecewise-linear models  
-```{r table_s3}
-map_df(log_fits, .f = ~{
-  fixef(.x) %>%
-    as_tibble(rownames = "terms") %>%
-    transmute(terms, coef = paste0(round(Estimate, 2), " (", round(Est.Error, 3), ")")) %>%
-    spread(terms, coef)},
-  .id = "Year"
-  ) %>%
-  rename("b0_log" = 2, "b1_log" = 3) -> coefs_gh9_log
-
-map_df(lm_fits, .f = ~{
-  fixef(.x) %>%
-    as_tibble(rownames = "terms") %>%
-    transmute(terms, coef = paste0(round(Estimate, 2), " (", round(Est.Error, 3), ")")) %>%
-    spread(terms, coef)},
-  .id = "Year"
-  ) %>%
-  select(Year, "b0_lin" = 3, "b1_lin" = 2) -> coefs_gh9_lm
-
-map_df(nl_fits, .f = ~{
-  fixef(.x) %>%
-    as_tibble(rownames = "terms") %>%
-    transmute(terms, coef = paste0(round(Estimate, 2), " (", round(Est.Error, 3), ")")) %>%
-    spread(terms, coef)},
-  .id = "Year"
-  ) %>%
-  rename("b0_pw" = 2, "b1_pw" = 3, 
-         "b2_pw" = 4, "Omega" = 5) -> coefs_gh9_pw
-
-coefs_affective <- left_join(coefs_gh9_pw, coefs_gh9_log) %>%
-  left_join(coefs_gh9_lm) %>%
-  mutate(Year = str_remove(Year, "fit"),
-         Wellbeing = "Affective")
-
-map_df(log_fits_losat, .f = ~{
-  fixef(.x) %>%
-    as_tibble(rownames = "terms") %>%
-    transmute(terms, coef = paste0(round(Estimate, 2), " (", round(Est.Error, 3), ")")) %>%
-    spread(terms, coef)},
-  .id = "Year"
-) %>%
-  rename("b0_log" = 2, "b1_log" = 3) -> coefs_losat_log
-
-map_df(lm_fits_losat, .f = ~{
-  fixef(.x) %>%
-    as_tibble(rownames = "terms") %>%
-    transmute(terms, coef = paste0(round(Estimate, 2), " (", round(Est.Error, 3), ")")) %>%
-    spread(terms, coef)},
-  .id = "Year"
-) %>%
-  select(Year, "b0_lin" = 3, "b1_lin" = 2) -> coefs_losat_lm
-
-map_df(nl_fits_losat, .f = ~{
-  fixef(.x) %>%
-    as_tibble(rownames = "terms") %>%
-    transmute(terms, coef = paste0(round(Estimate, 2), " (", round(Est.Error, 3), ")")) %>%
-    spread(terms, coef)},
-  .id = "Year"
-) %>%
-  rename("b0_pw" = 2, "b1_pw" = 3, 
-         "b2_pw" = 4, "Omega" = 5) -> coefs_losat_pw
-
-coefs_cognitive <- left_join(coefs_losat_pw, coefs_losat_log) %>%
-  left_join(coefs_losat_lm) %>%
-  mutate(Year = str_remove(Year, "fit"),
-         Wellbeing = "Cognitive")
-
-bind_rows(coefs_affective, coefs_cognitive) %>%
-  select(Wellbeing, everything()) %>%
-  flextable() %>%
-  merge_v(j = 1) %>%
-  set_header_labels(values = c(b0_pw = paste0("\u03b2", "0"),
-                               b1_pw = paste0("\u03b2", "1"),
-                               b2_pw = paste0("\u03b2", "2"),
-                               Omega = "\u03C9",
-                               b0_log = paste0("\u03b2", "0"),
-                               b1_log = paste0("\u03b2", "1"),
-                               b0_lin = paste0("\u03b2", "0"),
-                               b1_lin = paste0("\u03b2", "1"))) %>%
-  add_header(top = TRUE,
-             values = c(b0_pw = "Piecewise-linear (se)",
-                        b1_pw = "Piecewise-linear (se)",
-                        b2_pw = "Piecewise-linear (se)",
-                        Omega = "Piecewise-linear (se)",
-                        b0_log = "Log-linear (se)",
-                        b1_log = "Log-linear (se)",
-                        b0_lin = "Linear (se)",
-                        b1_lin = "Linear (se)")) %>%
-  merge_h(part = "header") %>%
-  width(j = 1:2, width = c(0.75, 0.5)) %>%
-  width(j = 3:10, width = 0.65) %>%
-  fontsize(size = 9, part = "all") %>%
-  # flextable::theme_vanilla() %>% # theme_vanilla() 
   valign(valign = "top") %>%
@@
   hline_bottom(part="header", border = officer::fp_border()) %>%
-  hline(i = 19, border = officer::fp_border()) %>%
   hline_bottom(part="body", border = officer::fp_border())
@@
 
-<br>
-
-The table of model parameters (untransformed) for each year show that the $\beta_1$ slopes for piecewise-linear, log-linear and linear model for affective wellbeing and income do not show a sustained trend ove the 19 year period.
-The $\beta_1$ coefficients increase in the initial 5-10 years and then decline.
-Despite this common pattern, the linear slope appears to show the most sustained increase while the log-linear slope shows a relatively less sustained increase.
-Overall, the pattern of changes among the models indicates the function between income and affective wellbeing has in fact become slightly steeper over the period.
-
-<br><br>
-
-#### Prior and residual checks
+<br><br>
 
@@
 To directly examine the implications of our model for such inequities in affective wellbeing, we used the model to estimate or _predict_ the affective wellbeing of two different income levels: one income level which was above the change point in 2001 but fell below it by 2019 (\$50K/yr); and another level which always remained above the change points over the same period (\$75K/yr).
-The figure below presents the affective wellbeing levels (in SD units/year) for each income level as well as the difference in affective wellbeing between them (∆).
-
-<br>
-
-##### Figure S8. affective wellbeing (SD units) at $50K/yr and $75K/yr from 2001-2019  
+The figure below presents the affective wellbeing levels for each income level as well as the difference in affective wellbeing between them (∆).
+
+<br>
+
+##### Figure S8. Affective wellbeing at $50K/yr and $75K/yr from 2001-2019  
 ```{r figure_s8, fig.width=7, fig.height=5, dpi=300}
 nl_fits = read_rds("../results/gh9_hhoecd_0cov_tight.RDS")
-
-# Optional
-# nl_fits_5cov = read_rds(
-#   paste0(here, "results/fits/excl_topcodes/sd/gh9_hhoecd_5cov_tight.RDS")
-# )
 
@@
   mutate(low = (40000 - mean)/sd,
-         high = (75000 - mean)/sd) %>%
+         high = (80000 - mean)/sd) %>%
   column_to_rownames(var = "fit") %>%
@@
 fixed_income_fit %>%
-  mutate(year = extract_numeric(fit)) %>%
-  select(Estimate, Dollars, year) %>%
-  filter(Dollars %in% c("$75K", "$50K")) %>%
-  spread(Dollars, Estimate) %>%
-  mutate(`$75K - $50K (∆)` = `$75K` - `$50K`) %>%
+  mutate(year = str_remove(fit, "fit"),
+         Happiness = Estimate * params_df[year, "sd_gh9"] + params_df[year, "mean_gh9"]) %>%
+  select(Happiness, Dollars, year) %>%
+  spread(Dollars, Happiness) %>%
+  mutate(`$75K - $50K (∆)` = (`$75K` - `$50K`)) %>%
   gather(income, happiness, -year) %>%
-  mutate(income = fct_relevel(income, "$50K", "$75K")) %>%
+  mutate(
+    income = fct_relevel(income, "$50K", "$75K"),
+    year = as.numeric(year)
+    ) -> d1
+
+d1 %>%
+  filter(income %in% c("$75K", "$50K")) %>%
   ggplot(aes(x = year, y = happiness, color = income)) +
@@
     facet_wrap(~income) +
-  labs(caption = "Source:     HILDA, University of Melbourne
-                 Predicted affective wellbeing scores (SD unit) from piecewise regression on real household income", 
-       x = "", y = "") -> p
-
-p +
+    labs(x = "", y = "") +
     ft_theme() +
@@
           plot.title = element_text(size = 19),
-        strip.text = element_text(size = 12))
+          strip.text = element_text(size = 12)) -> p1
+
+d1 %>%
+  filter(!income %in% c("$75K", "$50K")) %>%
+  ggplot(aes(x = year, y = happiness, color = income)) +
+  geom_point(size = 4, alpha = 0.25) +
+  geom_smooth(method = "loess", se = F, span = 3) +
+  facet_wrap(~income) +
+  labs(x = "", y = "") +
+  ft_theme() +
+  scale_colour_manual(values = "#be9c2e") +
+  theme(legend.position = "none",
+        plot.title = element_text(size = 19),
+        strip.text = element_text(size = 12)) -> p2
+
+library(patchwork)
+p1 + p2 + plot_layout(widths = c(2, 1))
 ```
@@
 In 2001, a household income of \$50K/year achieved an above average level of affective wellbeing relative to everyone else that year, however by 2019 affective wellbeing had declined to lower than average levels relative to the population.
-By contrast, a household income of \$75K/year enjoyed a higher than average level of affective wellbeing for the entire period.
-The difference (∆) in affective wellbeing between these two income levels doubled over the period.
+By comparison, a household income of \$75K/year enjoyed a higher level of affective wellbeing which decreased more slowly over the same period.
+The difference (∆) in affective wellbeing between these two income levels doubled over this period.
 The increasing disparity in affective wellbeing between the rich and the poor implies that affective wellbeing has became more inequitable over time.
@@
 
-#### OLS Insample fit statistics
-
-Our primary objective was to understand the relationships between the parameters of our model, and so in-sample fit statistics such as R^2^ are irrelevant for this objective.
-Moreover, our use of regularized priors will bias such in-sample estimates towards zero as regularization deliberately sacrifices in-sample variance for out-of-sample accuracy, which obviously hinders interpretation of the in-sample values.
-Nevertheless, some readers may find it useful to compare the in-sample fit statistics between the three different model types employed here: piecewise, linear and log-linear, as well as between the conditional and unconditional models.
-Because the usual fit statistics such as R^2^ present a problem for Bayesian fits, as the variance of the predicted values can be larger than the variance of the data results in R^2^ values greater than 1, we present the in-sample statistics from OLS model fits (without regularization or penalized likelihoods).
-
+<br>
+
+#### Bayesian linear, log-linear and piecewise-linear models
+
+The mean and standard deviation (±se) values of the posterior distributions for each coeficient from the piecewise-linear, log-linear and linear Bayesian models are shown in the table below.
+
+##### Table S3. Bayesian coefficients  
+```{r table_s3}
+map_df(log_fits, .f = ~{
+  fixef(.x) %>%
+    as_tibble(rownames = "terms") %>%
+    select(terms, Estimate, Est.Error)},
+  .id = "Year"
+) %>%
+  mutate(Year = str_remove(Year, "fit"),
+         Estimate = if_else(
+           str_detect(terms, "Intercept"),
+           (Estimate * sqrt(params_df[Year, "sd_gh9"]) + params_df[Year, "mean_gh9"]),
+           Estimate * params_df[Year, "sd_gh9"]/10),
+         Est.Error = if_else(
+           str_detect(terms, "Intercept"), 
+           Est.Error * sqrt(params_df[Year, "sd_gh9"]),
+           Est.Error * params_df[Year, "sd_gh9"]/10)
+  ) %>%
+  transmute(
+    Year,
+    terms, 
+    coef = paste0(round(Estimate, 2), "\n(", round(Est.Error, 3), ")")
+  ) %>%
+  spread(terms, coef) %>%
+  select(Year, "b0_log" = 2, "b1_log" = 3) -> coefs_gh9_log
+
+map_df(lm_fits, .f = ~{
+  fixef(.x) %>%
+    as_tibble(rownames = "terms") %>%
+    select(terms, Estimate, Est.Error)},
+  .id = "Year"
+  ) %>%
+  mutate(Year = str_remove(Year, "fit"),
+         Estimate = if_else(
+           str_detect(terms, "Intercept"),
+           (Estimate * params_df[Year, "sd_gh9"] + params_df[Year, "mean_gh9"]),
+           Estimate * params_df[Year, "sd_gh9"]/10),
+         Est.Error = if_else(
+           str_detect(terms, "Intercept"), 
+           Est.Error * params_df[Year, "sd_gh9"],
+           Est.Error * params_df[Year, "sd_gh9"]/10)
+  ) %>%
+  transmute(
+    Year,
+    terms, 
+    coef = paste0(round(Estimate, 2), "\n(", round(Est.Error, 3), ")")
+  ) %>%
+  spread(terms, coef) %>%
+  select(Year, "b0_lin" = 3, "b1_lin" = 2) -> coefs_gh9_lm
+
+coefs_gh9_pw <- map_df(nl_fits, .f = ~{
+  fixef(.x) %>%
+    as_tibble(rownames = "terms") 
+    }, 
+  .id = "Year") %>%
+  select(Year, terms, Estimate, Est.Error) %>%
+  mutate(
+    Year = str_remove(Year,"fit"),
+    terms = str_remove(terms, "_Intercept")) %>%
+  pivot_wider(names_from = terms, 
+              values_from = c(Estimate, Est.Error),
+              names_glue = "{terms}_{.value}") %>%
+  mutate(
+    b0_Estimate = b0_Estimate * params_df[Year, "sd_gh9"] + 
+      params_df[Year, "mean_gh9"],
+    omega_Estimate = (omega_Estimate * params_df[Year, "sd_income"] + 
+                        params_df[Year, "mean_income"])/1000,
+    omega_Est.Error = omega_Est.Error * params_df[Year, "sd_income"]/1000,
+    across(c(b1_Estimate, b2_Estimate, b0_Est.Error, b1_Est.Error, b2_Est.Error),
+           ~ {.x * params_df[Year, "sd_gh9"]/10})
+    ) %>%
+  transmute(
+    Year,
+    b0_pw = paste0(round(b0_Estimate, 2), "\n(", round(b0_Est.Error, 3), ")"),
+    b1_pw = paste0(round(b1_Estimate, 2), "\n(", round(b1_Est.Error, 3), ")"),
+    b2_pw = paste0(round(b2_Estimate, 2), "\n(", round(b2_Est.Error, 3), ")"),
+    Omega = paste0(round(omega_Estimate, 2), "\n(", round(omega_Est.Error, 3), ")"))
+
+coefs_affective <- left_join(coefs_gh9_pw, coefs_gh9_log) %>%
+  left_join(coefs_gh9_lm)
+
+coefs_affective %>%
+  flextable(
+    col_keys = c("Year", "b0_pw", "b1_pw", "b2_pw", "Omega", "break1", "b0_log", 
+                 "b1_log", "break2", "b0_lin", "b1_lin") # 11 columns
+  ) %>%
+  merge_v(j = 1) %>%
+  set_header_labels(values = c(b0_pw = paste0("\u03b2", "0"),
+                               b1_pw = paste0("\u03b2", "1"),
+                               b2_pw = paste0("\u03b2", "2"),
+                               Omega = "\u03C9",
+                               b0_log = paste0("\u03b2", "0"),
+                               b1_log = paste0("\u03b2", "1"),
+                               b0_lin = paste0("\u03b2", "0"),
+                               b1_lin = paste0("\u03b2", "1"))) %>%
+  add_header(top = TRUE,
+             values = c(b0_pw = "Piecewise-linear (±se)",
+                        b1_pw = "Piecewise-linear (±se)",
+                        b2_pw = "Piecewise-linear (±se)",
+                        Omega = "Piecewise-linear (±se)",
+                        b0_log = "Log-linear (±se)",
+                        b1_log = "Log-linear (±se)",
+                        b0_lin = "Linear (±se)",
+                        b1_lin = "Linear (±se)")) %>%
+  merge_h(part = "header") %>%
+  width(j = 1, width = 0.5) %>%
+  width(j = 2:11, width = 0.625) %>%
+  width(j = c(6, 9), width = 0.1) %>%
+  fontsize(size = 9, part = "all") %>%
+  # flextable::theme_vanilla() %>% # theme_vanilla()
+  valign(valign = "top") %>%
+  valign(valign = "bottom", part = "header") %>%
+  border_remove() %>%
+  hline_top(part="header", border = officer::fp_border()) %>%
+  hline_bottom(part="header", border = officer::fp_border()) %>%
+  hline_bottom(part="body", border = officer::fp_border())
+```
+
+<br>
+
+The table of model parameters for each year show that the $\beta_1$ slopes for piecewise-linear, log-linear and linear model for affective wellbeing and income do not show a sustained trend over the 19 year period.
+The $\beta_1$ coefficients increase in the initial 5-10 years and then decline.
+Despite this common pattern, the linear slope appears to show the most sustained increase while the log-linear slope shows a relatively less sustained increase.
+Overall, the pattern of changes among the models indicates the function between income and affective wellbeing has in fact become slightly steeper over the period.
+
+<br><br>
+
+#### OLS linear, log-linear and piecewise-linear models
+
+The temporal trends in coefficients from OLS models were largely consistent with those observed in the Bayesian models above.
+The change point (in real income) increased between 2001 and 2019, however there were few other trends over time.
 ```{r model_summaries_yearly}
@@
   transmute(
-    y = as.vector(scale(gh9)),
+    # y = as.vector(scale(gh9)),
+    y = gh9,
     dollars = hh_disp_inc_eq/10000,
@@
 ```
+
+##### Table S4. MLE coefficients from piecewise-linear, linear, log-linear OLS models   
+```{r table_s4}
+lm_coef <- map_dfr(nested_models$lin_fit, broom::tidy, .id = "Year") %>%
+  mutate(
+    coefficient = paste0(round(estimate, 2), "\n(", round(std.error, 2), ")")
+    ) %>%
+  select(Year, term, coefficient) %>%
+  spread(term, coefficient) %>%
+  rename("b0_lin" = 2, "b1_lin" = 3)
+
+log_coef <- map_dfr(nested_models$log_fit, broom::tidy, .id = "Year") %>%
+  mutate(
+    estimate = ifelse(str_detect(term, "log"), estimate/10, estimate),
+    std.error = ifelse(str_detect(term, "log"), std.error/10, std.error),
+    coefficient = paste0(round(estimate, 2), "\n(", round(std.error, 2), ")")
+    ) %>%
+  select(Year, term, coefficient) %>%
+  spread(term, coefficient) %>%
+  rename("b0_log" = 2, "b1_log" = 3)
+
+pw_coefs <- map_dfr(nested_models$pw_fit, broom::tidy, .id = "Year") %>%
+  filter(term %in% c("(Intercept)", "dollars", "U1.dollars")) %>%
+  select(Year, term, estimate, std.error) %>%
+  pivot_wider(id_cols = Year, 
+              names_from = term, 
+              values_from = estimate:std.error) %>%
+  mutate(
+    b0_pw = paste0(
+      round(`estimate_(Intercept)`, 2), "\n(", round(`std.error_(Intercept)`, 2), ")"),
+    estimate_post = estimate_U1.dollars + estimate_dollars,
+    b1_pw = paste0(
+      round(estimate_dollars, 2), "\n(", round(std.error_dollars, 2), ")"),
+    b2_pw = paste0(
+      round(estimate_post, 2), "\n(", round(std.error_U1.dollars, 2), ")")
+    ) %>%
+  select(Year, b0_pw, b1_pw, b2_pw)
+
+pw_coefs <- map_dfr(nested_models$pw_fit, ~as_tibble(.x$psi), .id = "Year") %>%
+  mutate(term = "psi.dollars") %>%
+  mutate(Omega = paste0(round(`Est.`, 3)*10, "\n(", round(`St.Err`, 3)*10, ")")) %>%
+  select(Year, Omega) %>%
+  left_join(pw_coefs)
+
+left_join(lm_coef, log_coef, by = "Year") %>%
+  left_join(pw_coefs, by = "Year") %>%
+  mutate(Year = as.character(2000 + as.numeric(Year))) %>%
+  arrange(Year) %>%
+  flextable(
+    col_keys = c("Year", "b0_pw", "b1_pw", "b2_pw", "Omega", "break1", "b0_log", 
+                 "b1_log", "break2", "b0_lin", "b1_lin") # 11 columns
+  ) %>%
+  merge_v(j = 1) %>%
+  set_header_labels(values = c(b0_pw = paste0("\u03b2", "0"),
+                               b1_pw = paste0("\u03b2", "1"),
+                               b2_pw = paste0("\u03b2", "2"),
+                               Omega = "\u03C9",
+                               b0_log = paste0("\u03b2", "0"),
+                               b1_log = paste0("\u03b2", "1"),
+                               b0_lin = paste0("\u03b2", "0"),
+                               b1_lin = paste0("\u03b2", "1"))) %>%
+  add_header(top = TRUE,
+             values = c(b0_pw = "Piecewise-linear (±se)",
+                        b1_pw = "Piecewise-linear (±se)",
+                        b2_pw = "Piecewise-linear (±se)",
+                        Omega = "Piecewise-linear (±se)",
+                        b0_log = "Log-linear (±se)",
+                        b1_log = "Log-linear (±se)",
+                        b0_lin = "Linear (±se)",
+                        b1_lin = "Linear (±se)")) %>%
+  merge_h(part = "header") %>%
+  width(j = 1, width = 0.5) %>%
+  width(j = 2:11, width = 0.625) %>%
+  width(j = c(6, 9), width = 0.1) %>%
+  fontsize(size = 9, part = "all") %>%
+  # flextable::theme_vanilla() %>% # theme_vanilla()
+  valign(valign = "top") %>%
+  valign(valign = "bottom", part = "header") %>%
+  border_remove() %>%
+  hline_top(part="header", border = officer::fp_border()) %>%
+  hline_bottom(part="header", border = officer::fp_border()) %>%
+  hline_bottom(part="body", border = officer::fp_border())
+```
+
+<br><br>
+
+**In-sample fit statistics**.
+Our primary objective was to understand the relationships between the parameters of our model, and so in-sample fit statistics such as R^2^ are irrelevant for this objective.
+Moreover, our use of regularized priors will bias such in-sample estimates towards zero as regularization deliberately sacrifices in-sample variance for out-of-sample accuracy, which obviously hinders interpretation of the in-sample values.
+Nevertheless, some readers may find it useful to compare the in-sample fit statistics between the three different model types employed here: piecewise, linear and log-linear, as well as between the conditional and unconditional models.
+Because the usual fit statistics such as R^2^ present a problem for Bayesian fits, as the variance of the predicted values can be larger than the variance of the data results in R^2^ values greater than 1, we present the in-sample statistics from OLS model fits (without regularization or penalized likelihoods).
+
+##### Table S5. Adjusted R-squared for affective wellbeing and income
 ```{r model_summaries_allyears}
@@
 ```
-
-##### Table S4. In-sample fit statistics (adj. R-squared) for affective wellbeing and income 
-```{r table_s4}
+```{r table_s5}
 model_summaries_all %>%
@@
 
-##### Table S5. OLS coefficients from linear, log-linear and piecewise models  
-```{r table_s5}
-lm_coef <- map_dfr(nested_models$lin_fit, broom::tidy, .id = "Year") %>%
-  filter(term == "dollars") %>%
-  mutate(coefficient = paste0(round(estimate, 3), " (", round(std.error, 3), ")")) %>%
-  select(Year, `linear (se)` = coefficient)
-
-log_coef <- map_dfr(nested_models$log_fit, broom::tidy, .id = "Year") %>%
-  filter(term == "log(dollars)") %>%
-  mutate(coefficient = paste0(round(estimate, 3), " (", round(std.error, 3), ")")) %>%
-  select(Year, `log-linear (se)` = coefficient)
-
-pw_coefs <- map_dfr(nested_models$pw_fit, broom::tidy, .id = "Year") %>%
-  filter(term %in% c("dollars", "U1.dollars")) %>%
-  select(Year, term, estimate, std.error) %>%
-  pivot_wider(id_cols = Year, 
-              names_from = term, 
-              values_from = estimate:std.error) %>%
-  mutate(
-    estimate_post = estimate_U1.dollars + estimate_dollars,
-    `pre-slope (se)` = paste0(
-      round(estimate_dollars, 3), " ±", round(std.error_dollars, 3)),
-    `post-slope (se)` = paste0(
-      round(estimate_post, 3), " ±", round(std.error_U1.dollars, 3))
-    ) %>%
-  select(Year, `pre-slope (se)`, `post-slope (se)`)
-
-pw_coefs <- map_dfr(nested_models$pw_fit, ~as_tibble(.x$psi), .id = "Year") %>%
-  mutate(term = "psi.dollars") %>%
-  mutate(coefficient = paste0(round(`Est.`, 2)*10, " (", round(`St.Err`, 2)*10, ")")) %>%
-  select(Year, `change point ($000s ±se)` = coefficient) %>%
-  left_join(pw_coefs)
-
-left_join(lm_coef, log_coef, by = "Year") %>%
-  left_join(pw_coefs, by = "Year") %>%
-  mutate(Year = as.character(2000 + as.numeric(Year))) %>%
-  flextable() %>%
-  fontsize(size = 9.5, part = "all") %>%
-  autofit()
-```
-
-<br><br>
-
 ## References
